### 2018-01-18 project_1/article/article.tex
Add start of project article

\documentclass[12pt, a4paper]{article}

% It's nice to be able to write your own name
\usepackage[T1]{fontenc}
% Nice fonts for nice people
\usepackage[sc, osf]{mathpazo}
\linespread{1.025} % Make more room for Palatino font
\usepackage[euler-digits, small]{eulervm}
% Automatic clickable links
\usepackage{hyperref}
% SI-units
\usepackage{siunitx}
% Enhanced math formatting
\usepackage{amsmath}
% Extended math symbols
\usepackage{amssymb}
% Include proof environment
\usepackage{amsthm}
% Import physics package to include bra-ket
\usepackage{physics}
\usepackage{enumerate}
% Import the tensor package for tensors
\usepackage{tensor}
% Include dod and dpd fracs
\usepackage{commath}
% Include font for the identity operator
\usepackage{dsfont}
% Include package for drawing Feynman diagrams
\usepackage{tikz-feynman}
% Include package for typesetting contractions
\usepackage{simpler-wick}
% Include the Feynman-slash
\usepackage{slashed}
% Include Tikz
\usepackage{tikz}
% Make Latex a bit wider
\usepackage[a4paper, total={6in, 8in}]{geometry}

% Macro for latin-letter vectors
\newcommand{\vf}{\mathbf}
% Macro for greek-letter vectors
\newcommand{\vfg}{\boldsymbol}

% Fast macro for real-numbers R
\newcommand{\R}{\mathbb{R}}
% Fast macro for complex-numbers C
\newcommand{\C}{\mathbb{C}}
% Fast macro for polynomial room P
\renewcommand{\P}{\mathbb{P}}
% New command for the identity operator
\newcommand{\1}{\mathds{1}}
% New command for the Lagrangian density
\newcommand{\cL}{\mathcal{L}}
% New command for the Hamiltonian density
\newcommand{\cH}{\mathcal{H}}
% Fast macro for partial differential tensors
\newcommand{\tpl}[1]{\tensor{\partial}{_#1}} % Lower
\newcommand{\tpu}[1]{\tensor{\partial}{^#1}} % Upper
% Fast macro for tensors
\newcommand{\te}[1]{\tensor{#1}}
% Fast macro for commutation and anti-commutation relations
\newcommand{\com}[2]{\left[#1, #2\right]}
\newcommand{\acom}[2]{\left\{#1, #2\right\}}

% Macros for writing auto-sized paranthesis, brackets and braces
\newcommand{\para}[1]{\left(#1\right)}
\newcommand{\brak}[1]{\left[#1\right]}
\newcommand{\brac}[1]{\left\{#1\right\}}

% Macro for creating orbital bra-ket's, i.e., bra-ket with paranthesis edges
\newcommand{\obra}[1]{( #1 \lvert}
\newcommand{\oket}[1]{\rvert #1 )}
\newcommand{\obraket}[2]{( #1 \lvert #2 )}

% Macro for expectation value
\newcommand{\expv}[1]{\langle #1 \rangle}

\title{Variational Monte Carlo on bosonic systems\\FYS4411}
\author{
    Winther-Larsen, Sebastian Gregorius
    \and
    Schøyen, Øyvind Sigmundson}
\date{\today}

\begin{document}

\maketitle

Start

\end{document}

### 2018-01-22 project_1/article/article.tex
Add introduction to the project
Use Revtex for article

@@
-\documentclass[12pt, a4paper]{article}
+\documentclass[
+    a4paper, aps, twocolumn, floatfix, superscriptaddress]{revtex4-1}
 
@@
 \usepackage[T1]{fontenc}
-% Nice fonts for nice people
-\usepackage[sc, osf]{mathpazo}
-\linespread{1.025} % Make more room for Palatino font
-\usepackage[euler-digits, small]{eulervm}
 % Automatic clickable links
@@
 \usepackage{tikz}
-% Make Latex a bit wider
-\usepackage[a4paper, total={6in, 8in}]{geometry}
 
@@
 
-\title{Variational Monte Carlo on bosonic systems\\FYS4411}
-\author{
-    Winther-Larsen, Sebastian Gregorius
-    \and
-    Schøyen, Øyvind Sigmundson}
+\begin{document}
+
+\title{Variational Monte Carlo on bosonic systems}
+\author{Winther-Larsen, Sebastian Gregorius}
+\affiliation{University of Oslo}
+\author{Schøyen, Øyvind Sigmundson}
+\affiliation{University of Oslo}
 \date{\today}
-
-\begin{document}
 
@@
 
-Start
+\section{Introduction}
+    We will in this project study the Variational Monte Carlo (VMC) method, and
+    use it to evaluate the ground state energy of a trapped, hard sphere Bose
+    gas.
 

### 2018-01-22 project_1/article/article.tex
Add Hamiltonian of the system

@@
 
+\newcommand{\half}{\frac{1}{2}}
+
 \begin{document}
@@
 
+\section{Theory}
+    To model the trapped bosonic gas particles we use the potential
+    \begin{align}
+        V_{\text{ext}}(\vf{r})
+        &=
+        \begin{cases}
+            \half m\omega^2r^2 & (\text{S}), \\
+            \half m [
+                \omega^2(x^2 + y^2) + \omega_z^2z^2
+            ] & (\text{E}),
+        \end{cases}
+    \end{align}
+    where we can choose between a spherical (S) or an elliptical (E) harmonic
+    trap. The two-body Hamiltonian of the system is given by
+    \begin{align}
+        H = \sum_{i}^{N}h(\vf{r}_i) + \sum_{i < j}^{N}w(\vf{r}_i, \vf{r}_j),
+    \end{align}
+    where the single particle one body operator, $h$, is given by
+    \begin{align}
+        h(\vf{r}_i) = -\frac{\hbar^2}{2m}\nabla_i^2
+        + V_{\text{ext}}(\vf{r}_i),
+    \end{align}
+    and the two-body interaction operator, $w$, is
+    \begin{align}
+        w(\vf{r}_i, \vf{r}_j)
+        = \begin{cases}
+            \infty & |\vf{r}_i - \vf{r}_j| \leq a, \\
+            0 & |\vf{r}_i - \vf{r}_j| > a,
+        \end{cases}
+    \end{align}
+    where $a$ is the hard sphere of the particle.
+
 \end{document}

### 2018-01-22 project_1/article/article.tex
Add trial wavefunction
Add motivation to finding the local energy

@@
             \half m\omega^2r^2 & (\text{S}), \\
-            \half m [
+            \half m \bigl[
                 \omega^2(x^2 + y^2) + \omega_z^2z^2
-            ] & (\text{E}),
+            \bigr] & (\text{E}),
         \end{cases}
@@
     \end{align}
-    where $a$ is the hard sphere of the particle.
+    where $a$ is the hard sphere of the particle. The trial wavefunction we will
+    be looking at is given by
+    \begin{align}
+        \Psi_T(\vf{r})
+        &= \bra{\vf{r}}\ket{\Psi_T}
+        = \prod_i^N g(\alpha, \beta, \vf{r}_i)
+        \prod_{i < j}^N f(a, \vf{r}_i, \vf{r}_j),
+    \end{align}
+    where $\alpha$ and $\beta$ are variational parameters and
+    \begin{align}
+        \vf{r} = (\vf{r}_1, \vf{r}_2, \dots, \vf{r}_N, \alpha, \beta).
+    \end{align}
+    Here $g$ are the single-particle wavefunctions given by
+    \begin{align}
+        g(\alpha, \beta, \vf{r}_i)
+        = \exp\bigl[
+            -\alpha(x_i^2 + y_i^2 + \beta z_i^2)
+        \bigr] \equiv \phi(\vf{r}_i),
+    \end{align}
+    and the correlation wavefunction, $f$, is
+    \begin{align}
+        f(a, \vf{r}_i, \vf{r}_j)
+        &=
+        \begin{cases}
+            0 & |\vf{r}_i - \vf{r}_j| \leq a, \\
+            \Bigl(
+                1 - \frac{a}{|\vf{r}_i - \vf{r}_j|}
+            \Bigr) & |\vf{r}_i - \vf{r}_j| > a.
+        \end{cases}
+    \end{align}
+
+    \subsection{Local energy}
+        As the many-body wavefunction creates a very large configuration space,
+        where much of the wavefunction is small, we use the Metropolis algorithm
+        in order to move towards regions in configuration space with
+        ``sensible'' values. We define the \textit{local energy}, $E_L{\vf{r}}$,
+        by
+        \begin{align}
+            E_L(\vf{r})
+            &= \frac{\bra{\vf{r}}H\ket{\Psi_T}}{\bra{\vf{r}}\ket{\Psi_T}}.
+        \end{align}
+        If $\ket{\Psi_T}$ is an exact eigenfunction of the Hamiltonian, $E_L$
+        will be constant. The closer $\ket{\Psi_T}$ is to the exact wave
+        function, the less variation in $E_L$ as a function of $\vf{r}$ we get.
+        One of the most computationally intensive part of the VMC algorithm will
+        be to compute $E_L$. We therefore find an analytical expression for
+        $E_L$ in terms of the trial wavefunction.
 

### 2018-01-22 project_1/article/article.tex
Add description of the simple case
	This is for the local energy when we have a non-interacting
	harmonic oscillator.

@@
 
+        \subsubsection{Non-interacting harmonic oscillator}
+            We start by finding an analytical expression for $a = 0$ and $\beta
+            = 1$, i.e., a non-interaction harmonic oscillator given by the
+            wavefunction
+            \begin{align}
+                \bra{\vf{r}}\ket{\Psi_T}
+                = \prod_i^N \exp\bigl[
+                    -\alpha |\vf{r}_i|^2
+                \bigr],
+            \end{align}
+            where $|\vf{r}_i| = r_i$. As $a = 0$ the interaction term,
+            $w(\vf{r}_i, \vf{r}_j)$, vanishes and the Hamiltonian is given by
+            (in the spherical case)
+            \begin{align}
+                H &= \sum_i^N h(\vf{r}_i)
+                = \sum_i^N \Bigl(
+                    -\frac{\hbar^2}{2m}\nabla_i^2
+                    + \half m \omega^2 r_i^2.
+                \Bigr)
+            \end{align}
+
 \end{document}

### 2018-01-22 project_1/article/article.tex
Add the gradient and Laplacian operators

@@
     \end{align}
+    (we assume equal mass)
     and the two-body interaction operator, $w$, is
@@
                     -\frac{\hbar^2}{2m}\nabla_i^2
-                    + \half m \omega^2 r_i^2.
-                \Bigr)
+                    + \half m \omega^2 r_i^2
+                \Bigr).
+            \end{align}
+            Working in spherical coordinates the gradient and the Laplace
+            operator (applied to a function) becomes
+            \begin{align}
+                \nabla_if
+                &= \vf{e}_r\tpl{r}f
+                + \vf{e}_{\theta}\frac{1}{r}\tpl{\theta}f
+                + \vf{e}_{\phi}\frac{1}{r\sin(\theta)}\tpl{\phi}f, \\
+                \nabla_i^2f
+                &= \frac{1}{r^2}\tpl{r}
+                \bigl[
+                    r^2\tpl{r}f
+                \bigr]
+                + \frac{1}{r^2\sin(\theta)}\tpl{\theta}
+                \bigl[
+                    \sin(\theta)\tpl{\theta}f
+                \bigr] \nonumber\\
+                &\qquad
+                + \frac{1}{r^2\sin^2(\theta)}\tpl{\phi}^2f,
+            \end{align}
+            where $\vf{e}_r$ signifies the unit vector in the radial direction
+            and we use the shorthand notation
+            \begin{align}
+                \tpl{r}f \equiv \dpd[]{f}{r}.
             \end{align}

### 2018-01-22 project_1/article/article.tex
Add theory on the non-interacting HO system

--- 2018-01-22 project_1/article/article.tex
@@
 
+        \subsubsection{Non-interacting harmonic oscillator}
+            We start by finding an analytical expression for $a = 0$ and $\beta
+            = 1$, i.e., a non-interaction harmonic oscillator given by the
+            wavefunction
+            \begin{align}
+                \bra{\vf{r}}\ket{\Psi_T}
+                = \prod_i^N \exp\bigl[
+                    -\alpha |\vf{r}_i|^2
+                \bigr],
+            \end{align}
+            where $|\vf{r}_i| = r_i$. As $a = 0$ the interaction term,
+            $w(\vf{r}_i, \vf{r}_j)$, vanishes and the Hamiltonian is given by
+            (in the spherical case)
+            \begin{align}
+                H &= \sum_i^N h(\vf{r}_i)
+                = \sum_i^N \Bigl(
+                    -\frac{\hbar^2}{2m}\nabla_i^2
+                    + \half m \omega^2 r_i^2
+                \Bigr).
+            \end{align}
+            Working in spherical coordinates the gradient and the Laplace
+            operator (applied to a function) becomes
+            \begin{align}
+                \nabla f
+                &= \vf{e}_r\tpl{r}f
+                + \vf{e}_{\theta}\frac{1}{r}\tpl{\theta}f
+                + \vf{e}_{\phi}\frac{1}{r\sin(\theta)}\tpl{\phi}f, \\
+                \nabla^2 f
+                &= \frac{1}{r^2}\tpl{r}
+                \bigl[
+                    r^2\tpl{r}f
+                \bigr]
+                + \frac{1}{r^2\sin(\theta)}\tpl{\theta}
+                \bigl[
+                    \sin(\theta)\tpl{\theta}f
+                \bigr] \nonumber\\
+                &\qquad
+                + \frac{1}{r^2\sin^2(\theta)}\tpl{\phi}^2f,
+            \end{align}
+            where $\vf{e}_r$ signifies the unit vector in the radial direction
+            and we use the shorthand notation
+            \begin{align}
+                \tpl{r}f \equiv \dpd[]{f}{r}.
+            \end{align}
+            As the trial wavefunction for $\beta = 1$ is purely radial we get
+            the same expression in one, two and three dimensions by setting
+            \begin{align}
+                r^2 =
+                \begin{cases}
+                    x & \text{one dimension}, \\
+                    x^2 + y^2 & \text{two dimensions}, \\
+                    x^2 + y^2 + z^2 & \text{three dimensions}.
+                \end{cases}
+            \end{align}
+            The gradient then becomes
+            \begin{align}
+                \bra{\vf{r}}\nabla_i\ket{\Psi_T}
+                &= \vf{e}_{{r_i}}\tpl{{r_i}}\prod_j^N\exp\bigl[
+                    -\alpha r_j^2
+                \bigr] \\
+                &=
+                -2\alpha r_i\prod_j^N\exp\bigl[
+                    -\alpha r_j^2
+                \bigr]\vf{e}_{{r_i}} \\
+                &=
+                -2\alpha r_i\bra{\vf{r}}\ket{\Psi_T}\vf{e}_{{r_i}}.
+            \end{align}
+            The Laplacian is given by
+            \begin{align}
+                \bra{\vf{r}}\nabla_i^2\ket{\Psi_T}
+                &= \frac{1}{r_i^2}\tpl{{r_i}}
+                \biggl\{
+                    r_i^2\tpl{{r_i}}
+                    \prod_j^N\exp\bigl[
+                        -\alpha r_j^2
+                    \bigr]
+                \biggr\} \\
+                &=
+                \frac{1}{r_i^2}\tpl{{r_i}}
+                \biggl\{
+                    -2 \alpha r_i^3
+                    \prod_j^N\exp\bigl[
+                        -\alpha r_j^2
+                    \bigr]
+                \biggr\} \\
+                &=
+                \frac{1}{r_i^2}
+                \Bigl\{
+                    -6 \alpha r_i^2
+                    + 4\alpha^2 r_i^4
+                \Bigr\}
+                \prod_j^N\exp\bigl[
+                    -\alpha r_j^2
+                \bigr]
+                \\
+                &=
+                \Bigl\{
+                    -6 \alpha
+                    + 4\alpha^2 r_i^2
+                \Bigr\}
+                \bra{\vf{r}}\ket{\Psi_T}.
+            \end{align}
+            This gives the analytical expression for the local energy to be
+            \begin{align}
+                E_L(\vf{r})
+                &= \frac{\bra{\vf{r}}H\ket{\Psi_T}}{\bra{\vf{r}}\ket{\Psi_T}} \\
+                &= \sum_i^N \Bigl(
+                    4\alpha^2 r_i^2
+                    + \half m \omega^2 r_i^2
+                \Bigr) - 6\alpha.
+            \end{align}
+            The \textit{drift force} of the system is given by
+            \begin{align}
+                \vf{F}_i
+                &=
+                \frac{2\bra{\vf{r}}\nabla_i\ket{\Psi_T}}
+                {\bra{\vf{r}}\ket{\Psi_T}} \\
+                &= -4\alpha r_i\vf{e}_{{r_i}}.
+            \end{align}
+
 \end{document}

### 2018-01-22 project_1/article/article.tex
Add description of the full problem
Add solution to the first derivative

@@
 
+        \subsubsection{The full system}
+            Moving to the full system allowing $\beta$ to vary and setting $a
+            \neq 0$ we write the trial wavefunction as
+            \begin{align}
+                \bra{\vf{r}}\ket{\Psi_T}
+                &= \prod_i^N \phi(\vf{r}_i)
+                \exp\Biggl(
+                    \sum_{i < j}^N u(r_{ij})
+                \Biggr),
+            \end{align}
+            where $r_{ij} = |\vf{r}_i - \vf{r}_j|$ and
+            \begin{align}
+                u(r_{ij}) = \ln\bigl[f(a, \vf{r}_i, \vf{r}_j)\bigr],
+            \end{align}
+            we wish to compute the local energy. The gradient is thus given by
+            \begin{align}
+                \bra{\vf{r}}\nabla_k\ket{\Psi_T}
+                &= \nabla_k\brac{
+                    \prod_i^N\phi(\vf{r}_i)
+                    \exp\Biggl(
+                        \sum_{i < j}^N u(r_{ij})
+                    \Biggr)
+                } \\
+                &=
+                \Biggl\{
+                    \nabla_k
+                    \prod_i^N\phi(\vf{r}_i)
+                \Biggr\}
+                \exp\Biggl(
+                    \sum_{i < j}^N u(r_{ij})
+                \Biggr) \\
+                &\qquad
+                +
+                \prod_i^N\phi(\vf{r}_i)
+                \nabla_k
+                \exp\Biggl(
+                    \sum_{i < j}^N u(r_{ij})
+                \Biggr) \\
+                &=
+                \nabla_k\phi(\vf{r}_k)
+                \prod_{i \neq k}^N \phi(\vf{r}_i)
+                \exp\Biggl(
+                    \sum_{i < j}^N u(r_{ij})
+                \Biggr) \\
+                &\qquad
+                +
+                \prod_i^N\phi(\vf{r}_i)
+                \exp\Biggl(
+                    \sum_{i < j}^N u(r_{ij})
+                \Biggr) \\
+                &\qquad
+                \times
+                \nabla_k
+                \Biggl(
+                    \sum_{i < l}^N u(r_{il})
+                \Biggr).
+            \end{align}
+            The last term is extracted from the chain rule and can be evaluated
+            as
+            \begin{align}
+                \nabla_k
+                \Biggl(
+                    \sum_{i < l}^N u(r_{il})
+                \Biggr)
+                &= \begin{cases}
+                    \nabla_k u(r_{ik}) & i < k, \\
+                    \sum_{k < l}^N \nabla_k u(r_{kl}) & i = k.
+                \end{cases}
+            \end{align}
+            As $r_{ij} = r_{ji}$ we can rewrite the first case to
+            \begin{align}
+                \nabla_k u(r_{ik}) = \nabla_k u(r_{ki}).
+            \end{align}
+            Here $i$ varies with the product thus we get
+            \begin{align}
+                \nabla_k
+                \Biggl(
+                    \sum_{i < l}^N u(r_{il})
+                \Biggr)
+                &=
+                \sum_{l \neq k}^N\nabla_k u(r_{kl}).
+            \end{align}
+            We are thus left with
+            \begin{align}
+                \bra{\vf{r}}\nabla_k\ket{\Psi_T}
+                &=
+                \nabla_k\phi(\vf{r}_k)
+                \prod_{i \neq k}^N \phi(\vf{r}_i)
+                \exp\Biggl(
+                    \sum_{i < j}^N u(r_{ij})
+                \Biggr) \\
+                &\qquad
+                +
+                \prod_i^N\phi(\vf{r}_i)
+                \exp\Biggl(
+                    \sum_{i < j}^N u(r_{ij})
+                \Biggr) \\
+                &\qquad
+                \times
+                \sum_{l \neq k}^N\nabla_k u(r_{kl}).
+            \end{align}
+
 \end{document}

### 2018-01-23 project_1/article/article.tex
Fix notation and correct the gradient calculation

--- 2018-01-22 project_1/article/article.tex
@@
         &= \bra{\vf{r}}\ket{\Psi_T}
-        = \prod_i^N g(\alpha, \beta, \vf{r}_i)
-        \prod_{i < j}^N f(a, \vf{r}_i, \vf{r}_j),
+        = \bra{\vf{r}}\ket{\Phi_T}
+        \prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k) \\
+        &= \Biggl(
+            \prod_i^N g(\alpha, \beta, \vf{r}_i)
+        \Biggr)
+        \prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k),
     \end{align}
@@
     \end{align}
-    Here $g$ are the single-particle wavefunctions given by
+    Here $g$ are the single particle wavefunctions given by
     \begin{align}
@@
     \end{align}
-    and the correlation wavefunction, $f$, is
+    and $\ket{\Phi_T}$ the \textit{Slater permanent} consisting of the $N$ first
+    single particle wavefunctions, and the correlation wavefunction, $f$, given
+    by
     \begin{align}
-        f(a, \vf{r}_i, \vf{r}_j)
+        f(a, \vf{r}_j, \vf{r}_k)
         &=
         \begin{cases}
-            0 & |\vf{r}_i - \vf{r}_j| \leq a, \\
+            0 & |\vf{r}_j - \vf{r}_k| \leq a, \\
             \Bigl(
-                1 - \frac{a}{|\vf{r}_i - \vf{r}_j|}
-            \Bigr) & |\vf{r}_i - \vf{r}_j| > a.
+                1 - \frac{a}{|\vf{r}_j - \vf{r}_k|}
+            \Bigr) & |\vf{r}_j - \vf{r}_k| > a.
         \end{cases}
@@
         \subsubsection{Non-interacting harmonic oscillator}
-            We start by finding an analytical expression for $a = 0$ and $\beta
-            = 1$, i.e., a non-interaction harmonic oscillator given by the
-            wavefunction
+            We start by finding an analytical expression for the local energy
+            with a system where we set $a = 0$ and $\beta = 1$, i.e., a system
+            of non-interacting harmonic oscillators.
             \begin{align}
                 \bra{\vf{r}}\ket{\Psi_T}
+                = \bra{\vf{r}}\ket{\Phi_T}
                 = \prod_i^N \exp\bigl[
@@
                 H &= \sum_i^N h(\vf{r}_i)
-                = \sum_i^N \Bigl(
+                = \sum_i^N \biggl(
                     -\frac{\hbar^2}{2m}\nabla_i^2
                     + \half m \omega^2 r_i^2
-                \Bigr).
+                \biggr).
             \end{align}
             Working in spherical coordinates the gradient and the Laplace
-            operator (applied to a function) becomes
+            operator (applied to a scalar function) becomes
             \begin{align}
@@
                 \begin{cases}
-                    x & \text{one dimension}, \\
+                    x^2 & \text{one dimension}, \\
                     x^2 + y^2 & \text{two dimensions}, \\
@@
 
+        \subsubsection{The full system}
+            Moving to the full system allowing $\beta$ to vary and setting $a
+            \neq 0$ we write the trial wavefunction as
+            \begin{align}
+                \bra{\vf{r}}\ket{\Psi_T}
+                &=
+                \bra{\vf{r}}\ket{\Phi_T}
+                \exp\Biggl(
+                    \sum_{j < k}^N u(r_{jk})
+                \Biggr) \\
+                &=
+                \Biggl(
+                    \prod_i^N \phi(\vf{r}_i)
+                \Biggr)
+                \exp\Biggl(
+                    \sum_{j < k}^N u(r_{jk})
+                \Biggr),
+            \end{align}
+            where $r_{jk} = |\vf{r}_j - \vf{r}_k|$ and
+            \begin{align}
+                u(r_{jk}) = \ln\bigl[f(a, \vf{r}_j, \vf{r}_k)\bigr],
+            \end{align}
+            we wish to compute the local energy. The gradient is thus given by
+            \begin{widetext}
+                \begin{align}
+                    \bra{\vf{r}}\nabla_k\ket{\Psi_T}
+                    &= \nabla_k\brac{
+                        \bra{\vf{r}}\ket{\Phi_T}
+                        \exp\Biggl(
+                            \sum_{j < l}^N u(r_{jl})
+                        \Biggr)
+                    } \\
+                    &=
+                    \Biggl\{
+                        \nabla_k
+                        \prod_i^N\phi(\vf{r}_i)
+                    \Biggr\}
+                    \exp\Biggl(
+                        \sum_{j < l}^N u(r_{jl})
+                    \Biggr)
+                    +
+                    \bra{\vf{r}}\ket{\Phi_T}
+                    \nabla_k
+                    \exp\Biggl(
+                        \sum_{j < l}^N u(r_{jl})
+                    \Biggr) \\
+                    &=
+                    \Biggl\{
+                        \nabla_k\phi(\vf{r}_k)
+                        \prod_{i \neq k}^N \phi(\vf{r}_i)
+                    \Biggr\}
+                    \exp\Biggl(
+                        \sum_{j < l}^N u(r_{jl})
+                    \Biggr)
+                    +
+                    \bra{\vf{r}}\ket{\Psi_T}
+                    \nabla_k
+                    \Biggl(
+                        \sum_{m < n}^N u(r_{mn})
+                    \Biggr).
+                \end{align}
+            The last term is extracted from the chain rule and can be evaluated
+            as
+            \begin{align}
+                \nabla_k
+                    \sum_{m < n}^N u(r_{mn})
+                &=
+                \sum_{m = 1}^{k - 1}
+                \nabla_k u(r_{mk})
+                +
+                \sum_{n = k + 1}^N
+                \nabla_k u(r_{kn}) \\
+                &=
+                \sum_{m \neq k}^N\nabla_k u(r_{km}),
+            \end{align}
+            as $r_{ij} = r_{ji}$.  We are thus left with
+            \begin{align}
+                \bra{\vf{r}}\nabla_k\ket{\Psi_T}
+                &=
+                \Biggl\{
+                    \nabla_k\phi(\vf{r}_k)
+                    \prod_{i \neq k}^N \phi(\vf{r}_i)
+                \Biggr\}
+                \exp\Biggl(
+                    \sum_{j < k}^N u(r_{jk})
+                \Biggr)
+                +
+                \bra{\vf{r}}\ket{\Psi_T}
+                \sum_{m \neq k}^N\nabla_k u(r_{km}).
+            \end{align}
+            \end{widetext}
+
 \end{document}

### 2018-01-23 project_1/article/article.tex
Indent text and equations in widetext
Add more systematic derivation of the gradient
	This makes it easier to expand the expression for the Laplace
	operator.

@@
         \prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k),
+        \label{eq:initial_trial_wavefunction}
     \end{align}
@@
                 \bra{\vf{r}}\ket{\Phi_T}
+                J(\vf{r}),
+            \end{align}
+            where $\ket{\Phi_T}$ is the same Slater permament as in
+            \autoref{eq:initial_trial_wavefunction} and $J(\vf{r})$ is the
+            \textit{Jastrow factor} given by
+            \begin{align}
+                J(\vf{r})
+                &=
                 \exp\Biggl(
-                    \sum_{j < k}^N u(r_{jk})
-                \Biggr) \\
-                &=
-                \Biggl(
-                    \prod_i^N \phi(\vf{r}_i)
-                \Biggr)
-                \exp\Biggl(
-                    \sum_{j < k}^N u(r_{jk})
+                    \sum_{j < l}^N u(r_{jl})
                 \Biggr),
@@
             \begin{align}
-                u(r_{jk}) = \ln\bigl[f(a, \vf{r}_j, \vf{r}_k)\bigr],
+                u(r_{jk}) = \ln\bigl[f(a, \vf{r}_j, \vf{r}_k)\bigr].
             \end{align}
-            we wish to compute the local energy. The gradient is thus given by
-            \begin{widetext}
+            We wish to find an analytical expression for the local energy.
+            Beginning with the gradient we get
             \begin{align}
                 \bra{\vf{r}}\nabla_k\ket{\Psi_T}
-                    &= \nabla_k\brac{
+                &= \nabla_k\Bigl[
                     \bra{\vf{r}}\ket{\Phi_T}
-                        \exp\Biggl(
-                            \sum_{j < l}^N u(r_{jl})
-                        \Biggr)
-                    } \\
+                    J(\vf{r})
+                \Bigr] \\
                 &=
-                    \Biggl\{
+                \Bigl[
                     \nabla_k
-                        \prod_i^N\phi(\vf{r}_i)
-                    \Biggr\}
-                    \exp\Biggl(
-                        \sum_{j < l}^N u(r_{jl})
-                    \Biggr)
+                    \bra{\vf{r}}\ket{\Phi_T}
+                \Bigr]
+                J(\vf{r})
+                + \bra{\vf{r}}\ket{\Phi_T}
+                \nabla_k J(\vf{r}).
+            \end{align}
+            The gradient of the Slater permament gives
+            \begin{align}
+                \nabla_k
+                \bra{\vf{r}}\ket{\Phi_T}
+                &=
+                \nabla_k\phi(\vf{r}_k)
+                \prod_{i \neq k}^N\phi(\vf{r}_i)
+                = \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                \bra{\vf{r}}\ket{\Phi_T}.
+            \end{align}
+            The gradient of the Jastrow factor is given by
+            \begin{align}
+                \nabla_k J(\vf{r})
+                &=
+                J(\vf{r})
+                \nabla_k\sum_{m < n}^N u(r_{mn}) \\
+                &=
+                J(\vf{r})
+                \Biggl(
+                    \sum_{m = 1}^{k - 1}
+                    \nabla_k u(r_{mk}) \nonumber \\
+                    &\qquad
                     +
-                    \bra{\vf{r}}\ket{\Phi_T}
-                    \nabla_k
-                    \exp\Biggl(
-                        \sum_{j < l}^N u(r_{jl})
+                    \sum_{n = k + 1}^N
+                    \nabla_k u(r_{kn})
                 \Biggr) \\
                 &=
-                    \Biggl\{
-                        \nabla_k\phi(\vf{r}_k)
-                        \prod_{i \neq k}^N \phi(\vf{r}_i)
-                    \Biggr\}
-                    \exp\Biggl(
-                        \sum_{j < l}^N u(r_{jl})
-                    \Biggr)
-                    +
-                    \bra{\vf{r}}\ket{\Psi_T}
-                    \nabla_k
-                    \Biggl(
-                        \sum_{m < n}^N u(r_{mn})
-                    \Biggr).
-                \end{align}
-            The last term is extracted from the chain rule and can be evaluated
-            as
-            \begin{align}
-                \nabla_k
-                    \sum_{m < n}^N u(r_{mn})
-                &=
-                \sum_{m = 1}^{k - 1}
-                \nabla_k u(r_{mk})
-                +
-                \sum_{n = k + 1}^N
-                \nabla_k u(r_{kn}) \\
-                &=
+                J(\vf{r})
                 \sum_{m \neq k}^N\nabla_k u(r_{km}),
             \end{align}
+%            \begin{align}
+%                &=
+%                \Biggl\{
+%                    \nabla_k
+%                    \prod_i^N\phi(\vf{r}_i)
+%                \Biggr\}
+%                \exp\Biggl(
+%                    \sum_{j < l}^N u(r_{jl})
+%                \Biggr)
+%                +
+%                \bra{\vf{r}}\ket{\Phi_T}
+%                \nabla_k
+%                \exp\Biggl(
+%                    \sum_{j < l}^N u(r_{jl})
+%                \Biggr) \\
+%                &=
+%                \Biggl\{
+%                    \nabla_k\phi(\vf{r}_k)
+%                    \prod_{i \neq k}^N \phi(\vf{r}_i)
+%                \Biggr\}
+%                \exp\Biggl(
+%                    \sum_{j < l}^N u(r_{jl})
+%                \Biggr)
+%                +
+%                \bra{\vf{r}}\ket{\Psi_T}
+%                \nabla_k
+%                \Biggl(
+%                    \sum_{m < n}^N u(r_{mn})
+%                \Biggr).
+%            \end{align}
+%            The last term is extracted from the chain rule and can be
+%            evaluated as
+%            \begin{align}
+%                \nabla_k
+%                    \sum_{m < n}^N u(r_{mn})
+%                &=
+%                \sum_{m = 1}^{k - 1}
+%                \nabla_k u(r_{mk})
+%                +
+%                \sum_{n = k + 1}^N
+%                \nabla_k u(r_{kn}) \\
+%                &=
+%                \sum_{m \neq k}^N\nabla_k u(r_{km}),
+%            \end{align}
             as $r_{ij} = r_{ji}$.  We are thus left with
             \begin{align}
-                \bra{\vf{r}}\nabla_k\ket{\Psi_T}
+                \nabla_k\bra{\vf{r}}\ket{\Psi_T}
                 &=
-                \Biggl\{
-                    \nabla_k\phi(\vf{r}_k)
-                    \prod_{i \neq k}^N \phi(\vf{r}_i)
-                \Biggr\}
-                \exp\Biggl(
-                    \sum_{j < k}^N u(r_{jk})
-                \Biggr)
-                +
-                \bra{\vf{r}}\ket{\Psi_T}
-                \sum_{m \neq k}^N\nabla_k u(r_{km}).
+                \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                \bra{\vf{r}}\ket{\Psi_T} \nonumber \\
+                &\qquad
+                + \bra{\vf{r}}\ket{\Psi_T}
+                \sum_{m \neq k}^N\nabla_k u(r_{km})
             \end{align}
-            \end{widetext}
+%            \begin{align}
+%                \bra{\vf{r}}\nabla_k\ket{\Psi_T}
+%                &=
+%                \Biggl\{
+%                    \nabla_k\phi(\vf{r}_k)
+%                    \prod_{i \neq k}^N \phi(\vf{r}_i)
+%                \Biggr\}
+%                \exp\Biggl(
+%                    \sum_{j < k}^N u(r_{jk})
+%                \Biggr)
+%                +
+%                \bra{\vf{r}}\ket{\Psi_T}
+%                \sum_{m \neq k}^N\nabla_k u(r_{km}).
+%            \end{align}
 

### 2018-01-25 project_1/article/article.tex
Move nabla-operator outside bra-ket
Break line with gradient of Slater permanent
Add general expression for the Laplacian

@@
             \begin{align}
-                \bra{\vf{r}}\nabla_i\ket{\Psi_T}
+                \nabla_i\bra{\vf{r}}\ket{\Psi_T}
                 &= \vf{e}_{{r_i}}\tpl{{r_i}}\prod_j^N\exp\bigl[
@@
             \begin{align}
-                \bra{\vf{r}}\nabla_i^2\ket{\Psi_T}
+                \nabla_i^2\bra{\vf{r}}\ket{\Psi_T}
                 &= \frac{1}{r_i^2}\tpl{{r_i}}
@@
                 &=
-                \frac{2\bra{\vf{r}}\nabla_i\ket{\Psi_T}}
+                \frac{2\nabla_i\bra{\vf{r}}\ket{\Psi_T}}
                 {\bra{\vf{r}}\ket{\Psi_T}} \\
@@
             \end{align}
-            where $\ket{\Phi_T}$ is the same Slater permament as in
+            where $\ket{\Phi_T}$ is the same Slater permanent as in
             \autoref{eq:initial_trial_wavefunction} and $J(\vf{r})$ is the
@@
             \begin{align}
-                \bra{\vf{r}}\nabla_k\ket{\Psi_T}
+                \nabla_k\bra{\vf{r}}\ket{\Psi_T}
                 &= \nabla_k\Bigl[
@@
             \end{align}
-            The gradient of the Slater permament gives
+            The gradient of the Slater permanent gives
             \begin{align}
@@
                 \nabla_k\phi(\vf{r}_k)
-                \prod_{i \neq k}^N\phi(\vf{r}_i)
-                = \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                \prod_{i \neq k}^N\phi(\vf{r}_i) \\
+                &= \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
                 \bra{\vf{r}}\ket{\Phi_T}.
             \end{align}
+            \begin{widetext}
                 The gradient of the Jastrow factor is given by
@@
                         \sum_{m = 1}^{k - 1}
-                    \nabla_k u(r_{mk}) \nonumber \\
-                    &\qquad
+                        \nabla_k u(r_{mk})
                         +
@@
                 \end{align}
-%            \begin{align}
-%                &=
-%                \Biggl\{
-%                    \nabla_k
-%                    \prod_i^N\phi(\vf{r}_i)
-%                \Biggr\}
-%                \exp\Biggl(
-%                    \sum_{j < l}^N u(r_{jl})
-%                \Biggr)
-%                +
-%                \bra{\vf{r}}\ket{\Phi_T}
-%                \nabla_k
-%                \exp\Biggl(
-%                    \sum_{j < l}^N u(r_{jl})
-%                \Biggr) \\
-%                &=
-%                \Biggl\{
-%                    \nabla_k\phi(\vf{r}_k)
-%                    \prod_{i \neq k}^N \phi(\vf{r}_i)
-%                \Biggr\}
-%                \exp\Biggl(
-%                    \sum_{j < l}^N u(r_{jl})
-%                \Biggr)
-%                +
-%                \bra{\vf{r}}\ket{\Psi_T}
-%                \nabla_k
-%                \Biggl(
-%                    \sum_{m < n}^N u(r_{mn})
-%                \Biggr).
-%            \end{align}
-%            The last term is extracted from the chain rule and can be
-%            evaluated as
-%            \begin{align}
-%                \nabla_k
-%                    \sum_{m < n}^N u(r_{mn})
-%                &=
-%                \sum_{m = 1}^{k - 1}
-%                \nabla_k u(r_{mk})
-%                +
-%                \sum_{n = k + 1}^N
-%                \nabla_k u(r_{kn}) \\
-%                &=
-%                \sum_{m \neq k}^N\nabla_k u(r_{km}),
-%            \end{align}
                 as $r_{ij} = r_{ji}$.  We are thus left with
@@
                     \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                \bra{\vf{r}}\ket{\Psi_T} \nonumber \\
+                    \bra{\vf{r}}\ket{\Psi_T}
+                    + \bra{\vf{r}}\ket{\Psi_T}
+                    \sum_{m \neq k}^N\nabla_k u(r_{km}) \\
+                    &=
+                    \Biggl(
+                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                        +
+                        \sum_{m \neq k}^N\nabla_k u(r_{km})
+                    \Biggr)
+                    \bra{\vf{r}}\ket{\Psi_T}.
+                \end{align}
+                We can now find the Laplacian of the trial wavefunction. This
+                gives
+                \begin{align}
+                    \nabla_k^2\bra{\vf{r}}\ket{\Psi_T}
+                    &= \nabla_k\nabla_k\bra{\vf{r}}\ket{\Psi_T}
+                    = \nabla_k
+                    \Biggl(
+                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                        +
+                        \sum_{m \neq k}^N\nabla_k u(r_{km})
+                    \Biggr)
+                    \bra{\vf{r}}\ket{\Psi_T} \\
+                    &=
+                    \Biggl(
+                        \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                        + \left[
+                            \nabla\frac{1}{\phi(\vf{r}_k)}
+                        \right]
+                        \nabla_k\phi(\vf{r}_k)
+                        +
+                        \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
+                    \Biggr)
+                    \bra{\vf{r}}\ket{\Psi_T}
+                    \\
                     &\qquad
-                + \bra{\vf{r}}\ket{\Psi_T}
+                    +
+                    \Biggl(
+                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                        +
                         \sum_{m \neq k}^N\nabla_k u(r_{km})
+                    \Biggr)
+                    \nabla_k\bra{\vf{r}}\ket{\Psi_T} \\
+                    &=
+                    \Biggl(
+                        \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                        - \left[
+                            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                        \right]^2
+                        +
+                        \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
+                    \Biggr)
+                    \bra{\vf{r}}\ket{\Psi_T}
+                    \\
+                    &\qquad
+                    +
+                    \Biggl(
+                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                        +
+                        \sum_{m \neq k}^N\nabla_k u(r_{km})
+                    \Biggr)^2
+                    \bra{\vf{r}}\ket{\Psi_T}.
                 \end{align}
-%            \begin{align}
-%                \bra{\vf{r}}\nabla_k\ket{\Psi_T}
-%                &=
-%                \Biggl\{
-%                    \nabla_k\phi(\vf{r}_k)
-%                    \prod_{i \neq k}^N \phi(\vf{r}_i)
-%                \Biggr\}
-%                \exp\Biggl(
-%                    \sum_{j < k}^N u(r_{jk})
-%                \Biggr)
-%                +
-%                \bra{\vf{r}}\ket{\Psi_T}
-%                \sum_{m \neq k}^N\nabla_k u(r_{km}).
-%            \end{align}
+                We now divide by the trial wavefunction. This simplifies the
+                calculations and is more similar to the expression for the local
+                energy.
+                \begin{align}
+                    \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
+                    {\bra{\vf{r}}\ket{\Psi_T}}
+                    &=
+                    \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                    - \left[
+                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                    \right]^2
+                    +
+                    \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
+                    +
+                    \Biggl(
+                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                        +
+                        \sum_{m \neq k}^N\nabla_k u(r_{km})
+                    \Biggr)^2 \\
+                    &=
+                    \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                    +
+                    \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
+                    +
+                    \frac{2\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+                    \sum_{m \neq k}^N\nabla_k u(r_{km})
+                    +
+                    \Biggl(
+                        \sum_{m \neq k}^N\nabla_k u(r_{km})
+                    \Biggr)^2
+                \end{align}
+            \end{widetext}
 

### 2018-01-25 project_1/article/article.tex
Add the gradient and Laplacian of the SPF

@@
             \end{widetext}
+            To go from here we have to find the gradient and the Laplacian of
+            the single particle functions, $\phi(\vf{r}_k)$, and the interaction
+            functions $u(r_{km})$. For the single particle functions we use
+            Cartesian coordinates when finding the derivatives whereas we for
+            the interaction functions will use spherical coordinates and do a
+            variable substitution. Beginning with the gradient of the single
+            particle functions we get
+            \begin{align}
+                \nabla_k\phi(\vf{r}_k)
+                &=
+                \nabla_k\exp\bigl[
+                    -\alpha(x_k^2 + y_k^2 + \beta z_k^2)
+                \bigr] \\
+                &=
+                -2\alpha
+                (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
+                \phi(\vf{r}_k),
+            \end{align}
+            note that the subscripts on the unit vectors $\vf{e}_i$ are
+            \textit{not} the same as the subscripts used for its components. The
+            Laplacian yields
+            \begin{align}
+                \nabla_k^2\phi(\vf{r}_k)
+                &=
+                -2\alpha
+                \big(
+                    2 + \beta
+                \bigr)
+                \phi(\vf{r}_k)
+                \nonumber \\
+                &\qquad
+                + 4\alpha^2
+                \bigl(
+                    x_k^2 + y_k^2 + \beta^2z_k^2
+                \bigr)
+                \phi(\vf{r}_k).
+            \end{align}
 

### 2018-01-25 project_1/article/article.tex
Add start of variable substitution
Add calculation of interaction functions

@@
             \end{align}
+            In order to derive the interaction functions we have to do a
+            variable substitution. We replace the derivative of the radial
+            component for particle $k$ by
+            \begin{align}
+                \tpl{{r_k}}
+                = \dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}.
+            \end{align}
+            The derivative of the distance $r_{km}$ is given by
+            \begin{gather}
+                r_{km} = |r_k - r_m|
+                \implies
+                \dpd[]{r_{km}}{r_k}
+                = \frac{r_k - r_m}{|r_k - r_m|}.
+            \end{gather}
+            Working in spherical coordinates we see that $u(r_{km})$ will only
+            be a function of radii of particles $k$ and $m$. The gradient thus
+            simplifies to the radial contribution giving
+            \begin{align}
+                \nabla_k u(r_{km})
+                &=
+                \vf{e}_{r}\tpl{{r_k}}u(r_{km})
+                \\
+                &=
+                \vf{e}_r\dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}u(r_{km})
+                \\
+                &= \vf{e}_r\frac{r_k - r_m}{r_{km}}
+                \tpl{{r_{km}}}u(r_{km}).
+            \end{align}
+            Using this expression we find the Laplacian to be
+            \begin{align}
+                \nabla_k^2 u(r_{km})
+                &=
+                \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
+                \nonumber \\
+                &\qquad
+                + \frac{(r_k - r_m)^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
+                \nonumber \\
+                &\qquad
+                +
+                \frac{r_k - r_m}{r_{km}}\tpl{{r_{km}}}^2 u(r_{km})
+            \end{align}
 

### 2018-01-27 project_1/article/article.tex
Add explicit expressions for derivates
	This is for the interaction functions.

@@
             \begin{gather}
-                r_{km} = |r_k - r_m|
+                r_{km} = |\vf{r}_k - \vf{r}_m|
                 \implies
                 \dpd[]{r_{km}}{r_k}
-                = \frac{r_k - r_m}{|r_k - r_m|}.
+                = \frac{r_k - r_m}{|\vf{r}_k - \vf{r}_m|},
             \end{gather}
-            Working in spherical coordinates we see that $u(r_{km})$ will only
-            be a function of radii of particles $k$ and $m$. The gradient thus
-            simplifies to the radial contribution giving
+            where in spherical coordinates $\vf{r}_k = r_k\vf{e}_r$.  The
+            gradient of $u(r_{km})$ thus simplifies to only include the radial
+            contribution.
             \begin{align}
@@
             \end{align}
+            \begin{widetext}
                 Using this expression we find the Laplacian to be
@@
                     \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
+                    + \frac{(r_k - r_m)^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
+                    +
+                    \frac{r_k - r_m}{r_{km}}
+                    \dpd[]{r_{km}}{r_k}
+                    \partial_{{r_{km}}}^2 u(r_{km}) \\
+                    &=
+                    \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
+                    + \frac{r_{km}^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
+                    +
+                    \frac{(r_k - r_m)^2}{r_{km}^2}
+                    \partial_{{r_{km}}}^2 u(r_{km}) \\
+                    &=
+                    \frac{2}{r_{km}}\tpl{{r_{km}}}u(r_{km})
+                    + \partial_{{r_{km}}}^2u(r_{km}).
+                \end{align}
+            \end{widetext}
+            We now take the derivate with respect to $r_{km}$ of the interaction
+            functions $u(r_{km})$ to find the closed form expressions. We get
+            \begin{align}
+                \partial_{r_{km}}u(r_{km})
+                &=
+                \partial_{r_{km}}
+                \ln\Biggl(
+                    1 - \frac{a}{r_{km}}
+                \Biggr) \\
+                &=
+                \partial_{r_{km}}
+                \ln\Biggl(
+                    \frac{1}{r_{km}}
+                    (r_{km} - a)
+                \Biggr) \\
+                &=
+                \partial_{r_{km}}
+                \Bigl[
+                    \ln(r_{km} - a)
+                    - \ln(r_{km})
+                \Bigr] \\
+                &=
+                \frac{1}{r_{km} - a} - \frac{1}{r_{km}} \\
+                &=
+                \frac{a}{r_{km}(r_{km} - a)} \\
+                \partial_{r_{km}}^2u(r_{km})
+                &= a\Bigl(
+                    \partial_{r_{km}}\frac{1}{r_{km}}
+                \Bigr)
+                \partial_{r_{km}}\frac{1}{r_{km} - a}
                 \nonumber \\
                 &\qquad
-                + \frac{(r_k - r_m)^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
-                \nonumber \\
-                &\qquad
-                +
-                \frac{r_k - r_m}{r_{km}}\tpl{{r_{km}}}^2 u(r_{km})
+                + a\frac{1}{r_{km}}
+                \Bigl(
+                    \partial_{r_{km}}\frac{1}{r_{km} - a}
+                \Bigr) \\
+                &=
+                -\frac{a}{r_{km}^2(r_{km} - a)}
+                -\frac{a}{r_{km}(r_{km} - a)^2} \\
+                &=
+                - \frac{2a(r_{km} - 1)}{r_{km}^2(r_{km} - a)^2}.
             \end{align}

### 2018-01-28 project_1/article/article.tex
Add explicit expression for the Laplacian

@@
             \end{align}
+            Using this we are now able to write the closed form expression of
+            the Laplacian of the trial wavefunction.
+            \begin{widetext}
+                \begin{align}
+                    \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
+                    {\bra{\vf{r}}\ket{\Psi_T}}
+                    &=
+                    -2\alpha
+                    \big(
+                        2 + \beta
+                    \bigr)
+                    + 4\alpha^2
+                    \bigl(
+                        x_k^2 + y_k^2 + \beta^2z_k^2
+                    \bigr)
+                    %\nonumber \\
+                    %&\qquad
+                    +
+                    2a
+                    \sum_{m \neq k}^N
+                    \Biggl[
+                        \frac{1}{r_{km}^2(r_{km} - a)}
+                        - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
+                    \Biggr]
+                    \nonumber \\
+                    &\qquad
+                    -4\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
+                    \cdot
+                    a\sum_{m \neq k}^N
+                    \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
+                    + \Biggl[
+                        a\sum_{m \neq k}^N
+                        \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
+                    \Biggr]^2.
+                \end{align}
+            \end{widetext}
+            To ease the load on the CPU we try to restrict the number of times
+            we evaluate the sums. For brevity we introduce the following
+            functions
+            \begin{align}
+                \xi_k(\vf{r}, a)
+                &= \sum_{m \neq k}^N
+                \Biggl[
+                    \frac{1}{r_{km}^2(r_{km} - a)}
+                    - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
+                \Biggr] \\
+                &=
+                \sum_{m \neq k}^N
+                \frac{1 - a}{r_{km}^2(r_{km} - a)^2}, \\
+                \vfg{\zeta}_k(\vf{r}, a)
+                &=
+                \sum_{m \neq k}^N
+                \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}.
+            \end{align}
+            We can then rewrite the expression for the Laplacian to
+            \begin{widetext}
+                \begin{align}
+                    \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
+                    {\bra{\vf{r}}\ket{\Psi_T}}
+                    &=
+                    -2\alpha
+                    \big(
+                        2 + \beta
+                    \bigr)
+                    + 4\alpha^2
+                    \bigl(
+                        x_k^2 + y_k^2 + \beta^2z_k^2
+                    \bigr)
+                    %\nonumber \\
+                    %&\qquad
+                    +
+                    2a\xi_k(\vf{r}, a)
+                    \nonumber \\
+                    &\qquad
+                    -4a\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
+                    \cdot
+                    \vfg{\zeta}_k(\vf{r}, a)
+                    + a^2\vfg{\zeta}_k(\vf{r}, a)^2.
+                \end{align}
+                Both $\xi_k$ and $\vfg{\zeta}_k$ can be evaluated in a common
+                loop. Using this expression for the Laplacian we can find the
+                local energy from
+                \begin{align}
+                    E_L(\vf{r})
+                    &=
+                    \frac{\bra{\vf{r}}H\ket{\Psi_T}}
+                    {\bra{\vf{r}}\ket{\Psi_T}}
+                    =
+                    \sum_{k}^N
+                    \Biggl(
+                        -\frac{\hbar^2}{2m}
+                        \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
+                        {\bra{\vf{r}}\ket{\Psi_T}}
+                        + V_{\text{ext}}(\vf{r})
+                    \Biggr)
+                    +
+                    \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j).
+                \end{align}
+            \end{widetext}
 

### 2018-01-28 project_1/vmc_hybrid/setup.py
Add skeleton of setup-script

@@ -1,0 +1,24 @@
+from distutils.core import setup
+from distutils.extension import Extension
+from Cython.Build import cythonize
+
+import numpy as np
+
+source_files = []
+
+
+include_dirs = [
+        "include",
+        np.get_include()
+]
+
+
+libraries = []
+
+extensions = []
+
+setup(
+    name="Bosonic vmc",
+    version="0.0.1",
+    ext_modules=cythonize(extensions)
+)

### 2018-02-06 project_1/article/article.tex
Add abstract section
Fix non-interacting case

@@
 
+\begin{abstract}
+    Something very abstract and clever should go here.
+\end{abstract}
+
 \maketitle
@@
     \end{align}
-    where $a$ is the hard sphere of the particle. The trial wavefunction we will
+    where $a$ is the hard sphere of the particle. The trial wavefunction,
+    $\ket{\Psi_T}$, we will
     be looking at is given by
@@
         \Psi_T(\vf{r})
-        &= \bra{\vf{r}}\ket{\Psi_T}
-        = \bra{\vf{r}}\ket{\Phi_T}
+        &= \Phi_T(\vf{r})
         \prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k) \\
@@
             E_L(\vf{r})
-            &= \frac{\bra{\vf{r}}H\ket{\Psi_T}}{\bra{\vf{r}}\ket{\Psi_T}}.
+            &= \frac{H\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
         \end{align}
@@
         function, the less variation in $E_L$ as a function of $\vf{r}$ we get.
-        One of the most computationally intensive part of the VMC algorithm will
-        be to compute $E_L$. We therefore find an analytical expression for
+        One of the most computationally intensive parts of the VMC algorithm
+        will be to compute $E_L$. We therefore find an analytical expression for
         $E_L$ in terms of the trial wavefunction.
@@
         \subsubsection{Non-interacting harmonic oscillator}
-            We start by finding an analytical expression for the local energy
-            with a system where we set $a = 0$ and $\beta = 1$, i.e., a system
-            of non-interacting harmonic oscillators.
+            We start by finding an analytical expression for the local energy in
+            a system where we set $a = 0$ and $\beta = 1$, i.e., a system of
+            non-interacting harmonic oscillators.
             \begin{align}
-                \bra{\vf{r}}\ket{\Psi_T}
-                = \bra{\vf{r}}\ket{\Phi_T}
+                \Psi_T(\vf{r})
+                = \Phi_T(\vf{r})
                 = \prod_i^N \exp\bigl[
@@
                 H &= \sum_i^N h(\vf{r}_i)
-                = \sum_i^N \biggl(
+                = \sum_i^N \Biggl(
                     -\frac{\hbar^2}{2m}\nabla_i^2
                     + \half m \omega^2 r_i^2
-                \biggr).
+                \Biggr).
             \end{align}
-            Working in spherical coordinates the gradient and the Laplace
-            operator (applied to a scalar function) becomes
+            The hardest part of finding the local energy is computing the
+            kinetic term. Starting with the gradient of the wavefunction in the
+            non-interacting case, we get
             \begin{align}
-                \nabla f
-                &= \vf{e}_r\tpl{r}f
-                + \vf{e}_{\theta}\frac{1}{r}\tpl{\theta}f
-                + \vf{e}_{\phi}\frac{1}{r\sin(\theta)}\tpl{\phi}f, \\
-                \nabla^2 f
-                &= \frac{1}{r^2}\tpl{r}
-                \bigl[
-                    r^2\tpl{r}f
-                \bigr]
-                + \frac{1}{r^2\sin(\theta)}\tpl{\theta}
-                \bigl[
-                    \sin(\theta)\tpl{\theta}f
-                \bigr] \nonumber\\
-                &\qquad
-                + \frac{1}{r^2\sin^2(\theta)}\tpl{\phi}^2f,
+                \nabla_k\Psi_T(\vf{r})
+                &= -2\alpha \vf{r}_k\Psi_T(\vf{r}).
             \end{align}
-            where $\vf{e}_r$ signifies the unit vector in the radial direction
-            and we use the shorthand notation
+            The Laplacian yields
             \begin{align}
-                \tpl{r}f \equiv \dpd[]{f}{r}.
+                \nabla^2_k\Psi_T(\vf{r})
+                &= \big(-2d\alpha + 4\alpha^2 r_k^2\bigr)\Psi_T(\vf{r}),
             \end{align}
-            As the trial wavefunction for $\beta = 1$ is purely radial we get
-            the same expression in one, two and three dimensions by setting
-            \begin{align}
-                r^2 =
-                \begin{cases}
-                    x^2 & \text{one dimension}, \\
-                    x^2 + y^2 & \text{two dimensions}, \\
-                    x^2 + y^2 + z^2 & \text{three dimensions}.
-                \end{cases}
-            \end{align}
-            The gradient then becomes
-            \begin{align}
-                \nabla_i\bra{\vf{r}}\ket{\Psi_T}
-                &= \vf{e}_{{r_i}}\tpl{{r_i}}\prod_j^N\exp\bigl[
-                    -\alpha r_j^2
-                \bigr] \\
-                &=
-                -2\alpha r_i\prod_j^N\exp\bigl[
-                    -\alpha r_j^2
-                \bigr]\vf{e}_{{r_i}} \\
-                &=
-                -2\alpha r_i\bra{\vf{r}}\ket{\Psi_T}\vf{e}_{{r_i}}.
-            \end{align}
-            The Laplacian is given by
-            \begin{align}
-                \nabla_i^2\bra{\vf{r}}\ket{\Psi_T}
-                &= \frac{1}{r_i^2}\tpl{{r_i}}
-                \biggl\{
-                    r_i^2\tpl{{r_i}}
-                    \prod_j^N\exp\bigl[
-                        -\alpha r_j^2
-                    \bigr]
-                \biggr\} \\
-                &=
-                \frac{1}{r_i^2}\tpl{{r_i}}
-                \biggl\{
-                    -2 \alpha r_i^3
-                    \prod_j^N\exp\bigl[
-                        -\alpha r_j^2
-                    \bigr]
-                \biggr\} \\
-                &=
-                \frac{1}{r_i^2}
-                \Bigl\{
-                    -6 \alpha r_i^2
-                    + 4\alpha^2 r_i^4
-                \Bigr\}
-                \prod_j^N\exp\bigl[
-                    -\alpha r_j^2
-                \bigr]
-                \\
-                &=
-                \Bigl\{
-                    -6 \alpha
-                    + 4\alpha^2 r_i^2
-                \Bigr\}
-                \bra{\vf{r}}\ket{\Psi_T}.
-            \end{align}
-            This gives the analytical expression for the local energy to be
+            where $d$ is the dimensionality of the problem determined by
+            $\vf{r}_k \in \mathbb{R}^d$. This yields the analytical expression
+            for the local energy in the non-interacting case to be
             \begin{align}
                 E_L(\vf{r})
-                &= \frac{\bra{\vf{r}}H\ket{\Psi_T}}{\bra{\vf{r}}\ket{\Psi_T}} \\
-                &= \sum_i^N \Bigl(
-                    4\alpha^2 r_i^2
-                    + \half m \omega^2 r_i^2
-                \Bigr) - 6\alpha.
+                &=
+                \sum_i^N
+                \Biggl(
+                    -\frac{\hbar^2}{2m}
+                    \bigl[
+                        -2d\alpha + 4\alpha^2 r_i^2
+                    \bigr]
+                    + \half m\omega^2 r_i^2
+                \Biggr).
             \end{align}
-            The \textit{drift force} of the system is given by
+            In natural units, with $\hbar = c = m = 1$, this reduces to
             \begin{align}
-                \vf{F}_i
+                E_L(\vf{r})
                 &=
-                \frac{2\nabla_i\bra{\vf{r}}\ket{\Psi_T}}
-                {\bra{\vf{r}}\ket{\Psi_T}} \\
-                &= -4\alpha r_i\vf{e}_{{r_i}}.
+                \alpha dN
+                + \biggl(
+                    \half\omega^2
+                    - 2\alpha^2
+                \biggr)
+                \sum_i^N r_i^2.
             \end{align}
+            It is worth noting that for $\omega = 1$ we will find a stable value
+            for the local energy when $\alpha = \half$. This happens as the
+            entire sum over all the random walkers disappears.
+
+%            Working in spherical coordinates the gradient and the Laplace
+%            operator (applied to a scalar function) becomes
+%            \begin{align}
+%                \nabla f
+%                &= \vf{e}_r\tpl{r}f
+%                + \vf{e}_{\theta}\frac{1}{r}\tpl{\theta}f
+%                + \vf{e}_{\phi}\frac{1}{r\sin(\theta)}\tpl{\phi}f, \\
+%                \nabla^2 f
+%                &= \frac{1}{r^2}\tpl{r}
+%                \bigl[
+%                    r^2\tpl{r}f
+%                \bigr]
+%                + \frac{1}{r^2\sin(\theta)}\tpl{\theta}
+%                \bigl[
+%                    \sin(\theta)\tpl{\theta}f
+%                \bigr] \nonumber\\
+%                &\qquad
+%                + \frac{1}{r^2\sin^2(\theta)}\tpl{\phi}^2f,
+%            \end{align}
+%            where $\vf{e}_r$ signifies the unit vector in the radial direction
+%            and we use the shorthand notation
+%            \begin{align}
+%                \tpl{r}f \equiv \dpd[]{f}{r}.
+%            \end{align}
+%            As the trial wavefunction for $\beta = 1$ is purely radial we get
+%            the same expression in one, two and three dimensions by setting
+%            \begin{align}
+%                r^2 =
+%                \begin{cases}
+%                    x^2 & \text{one dimension}, \\
+%                    x^2 + y^2 & \text{two dimensions}, \\
+%                    x^2 + y^2 + z^2 & \text{three dimensions}.
+%                \end{cases}
+%            \end{align}
+%            The gradient then becomes
+%            \begin{align}
+%                \nabla_i\bra{\vf{r}}\ket{\Psi_T}
+%                &= \vf{e}_{{r_i}}\tpl{{r_i}}\prod_j^N\exp\bigl[
+%                    -\alpha r_j^2
+%                \bigr] \\
+%                &=
+%                -2\alpha r_i\prod_j^N\exp\bigl[
+%                    -\alpha r_j^2
+%                \bigr]\vf{e}_{{r_i}} \\
+%                &=
+%                -2\alpha r_i\bra{\vf{r}}\ket{\Psi_T}\vf{e}_{{r_i}}.
+%            \end{align}
+%            The Laplacian is given by
+%            \begin{align}
+%                \nabla_i^2\bra{\vf{r}}\ket{\Psi_T}
+%                &= \frac{1}{r_i^2}\tpl{{r_i}}
+%                \biggl\{
+%                    r_i^2\tpl{{r_i}}
+%                    \prod_j^N\exp\bigl[
+%                        -\alpha r_j^2
+%                    \bigr]
+%                \biggr\} \\
+%                &=
+%                \frac{1}{r_i^2}\tpl{{r_i}}
+%                \biggl\{
+%                    -2 \alpha r_i^3
+%                    \prod_j^N\exp\bigl[
+%                        -\alpha r_j^2
+%                    \bigr]
+%                \biggr\} \\
+%                &=
+%                \frac{1}{r_i^2}
+%                \Bigl\{
+%                    -6 \alpha r_i^2
+%                    + 4\alpha^2 r_i^4
+%                \Bigr\}
+%                \prod_j^N\exp\bigl[
+%                    -\alpha r_j^2
+%                \bigr]
+%                \\
+%                &=
+%                \Bigl\{
+%                    -6 \alpha
+%                    + 4\alpha^2 r_i^2
+%                \Bigr\}
+%                \bra{\vf{r}}\ket{\Psi_T}.
+%            \end{align}
+%            This gives the analytical expression for the local energy to be
+%            \begin{align}
+%                E_L(\vf{r})
+%                &= \frac{\bra{\vf{r}}H\ket{\Psi_T}}{\bra{\vf{r}}\ket{\Psi_T}} \\
+%                &= \sum_i^N \Bigl(
+%                    4\alpha^2 r_i^2
+%                    + \half m \omega^2 r_i^2
+%                \Bigr) - 6\alpha.
+%            \end{align}
+%            The \textit{drift force} of the system is given by
+%            \begin{align}
+%                \vf{F}_i
+%                &=
+%                \frac{2\nabla_i\bra{\vf{r}}\ket{\Psi_T}}
+%                {\bra{\vf{r}}\ket{\Psi_T}} \\
+%                &= -4\alpha r_i\vf{e}_{{r_i}}.
+%            \end{align}
 

### 2018-02-06 project_1/article/article.tex
Update derivation of the full problem

@@
     \end{align}
+    We will for brevity use the notation $\phi(\vf{r}_i) = \phi_i$ and $r_{jk} =
+    |\vf{r}_j - \vf{r}_k|$.
 
@@
             \begin{align}
-                \bra{\vf{r}}\ket{\Psi_T}
+                \Psi_T(\vf{r})
                 &=
-                \bra{\vf{r}}\ket{\Phi_T}
+                \Phi_T(\vf{r})
                 J(\vf{r}),
@@
             \end{align}
+            To further shorten the notation we will use $u_{jk} = u(r_{jk})$.
             We wish to find an analytical expression for the local energy.
@@
             \begin{align}
-                \nabla_k\bra{\vf{r}}\ket{\Psi_T}
-                &= \nabla_k\Bigl[
-                    \bra{\vf{r}}\ket{\Phi_T}
-                    J(\vf{r})
-                \Bigr] \\
+                \nabla_k\Psi_T(\vf{r})
                 &=
@@
                     \nabla_k
-                    \bra{\vf{r}}\ket{\Phi_T}
+                    \Phi_T(\vf{r})
                 \Bigr]
                 J(\vf{r})
-                + \bra{\vf{r}}\ket{\Phi_T}
+                + \Phi_T(\vf{r})
                 \nabla_k J(\vf{r}).
             \end{align}
-            The gradient of the Slater permanent gives
+            The gradient of the Slater permament for particle $k$ is given by
             \begin{align}
                 \nabla_k
-                \bra{\vf{r}}\ket{\Phi_T}
+                \Phi_T(\vf{r})
                 &=
-                \nabla_k\phi(\vf{r}_k)
-                \prod_{i \neq k}^N\phi(\vf{r}_i) \\
-                &= \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                \bra{\vf{r}}\ket{\Phi_T}.
+                \nabla_k\phi_k
+                \prod_{i \neq k}^N\phi_i
+                = \frac{\nabla_k\phi_k}{\phi_k}
+                \Phi_T(\vf{r}).
             \end{align}
-            \begin{widetext}
             The gradient of the Jastrow factor is given by
@@
                 J(\vf{r})
-                    \nabla_k\sum_{m < n}^N u(r_{mn}) \\
+                \nabla_k\sum_{m < n}^N u_{mn} \\
+                &= J(\vf{r})
+                \Biggl(
+                    \sum_{m = 1}^{k - 1}\nabla_k u_{mk}
+                    \sum_{n = k + 1}^N\nabla_k u_{kn}
+                \Biggr)
+                \\
                 &=
                 J(\vf{r})
-                    \Biggl(
-                        \sum_{m = 1}^{k - 1}
-                        \nabla_k u(r_{mk})
-                        +
-                        \sum_{n = k + 1}^N
-                        \nabla_k u(r_{kn})
-                    \Biggr) \\
-                    &=
-                    J(\vf{r})
-                    \sum_{m \neq k}^N\nabla_k u(r_{km}),
+                \sum_{m \neq k}^N\nabla_k u_{km},
             \end{align}
-                as $r_{ij} = r_{ji}$.  We are thus left with
+            where the gradient of the interaction term splits the antisymmetric
+            sum into two parts. As $r_{ij} = r_{ji}$ we can combine these sums
+            into a single sum. This in total yields the gradient
             \begin{align}
-                    \nabla_k\bra{\vf{r}}\ket{\Psi_T}
-                    &=
-                    \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                    \bra{\vf{r}}\ket{\Psi_T}
-                    + \bra{\vf{r}}\ket{\Psi_T}
-                    \sum_{m \neq k}^N\nabla_k u(r_{km}) \\
+                \nabla_k\Psi_T(\vf{r})
                 &=
                 \Biggl(
-                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                        +
-                        \sum_{m \neq k}^N\nabla_k u(r_{km})
+                    \frac{\nabla_k\phi_k}{\phi_k}
+                    + \sum_{m \neq k}^N
+                    \nabla_k u_{km}
                 \Biggr)
-                    \bra{\vf{r}}\ket{\Psi_T}.
+                \Psi_T(\vf{r}).
+                \label{eq:gradient_full_wavefunction}
             \end{align}
-                We can now find the Laplacian of the trial wavefunction. This
-                gives
+            We now find the Laplacian of the trial wavefunction by taking the
+            gradient of \autoref{eq:gradient_full_wavefunction}.
             \begin{align}
-                    \nabla_k^2\bra{\vf{r}}\ket{\Psi_T}
-                    &= \nabla_k\nabla_k\bra{\vf{r}}\ket{\Psi_T}
-                    = \nabla_k
-                    \Biggl(
-                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                        +
-                        \sum_{m \neq k}^N\nabla_k u(r_{km})
-                    \Biggr)
-                    \bra{\vf{r}}\ket{\Psi_T} \\
+                \nabla_k^2\Psi_T(\vf{r})
                 &=
                 \Biggl(
-                        \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                        + \left[
-                            \nabla\frac{1}{\phi(\vf{r}_k)}
-                        \right]
-                        \nabla_k\phi(\vf{r}_k)
+                    \nabla_k\Biggl[
+                        \frac{\nabla_k\phi_k}{\phi_k}
+                    \Biggr]
                     +
-                        \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
-                    \Biggr)
-                    \bra{\vf{r}}\ket{\Psi_T}
+                    \sum_{m \neq k}^N \nabla_k^2 u_{km}
+                \Biggr)\Psi_T(\vf{r})
                 \\
@@
                 \Biggl(
-                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                        +
-                        \sum_{m \neq k}^N\nabla_k u(r_{km})
-                    \Biggr)
-                    \nabla_k\bra{\vf{r}}\ket{\Psi_T} \\
+                    \frac{\nabla_k\phi_k}{\phi_k}
+                    + \sum_{m \neq k}^N
+                    \nabla_k u_{km}
+                \Biggr)^2
+                \Psi_T(\vf{r}),
+            \end{align}
+            where the squared term came from taking the gradient of the trial
+            wavefunction.  To further simplify we divide by the trial
+            wavefunction. This yields
+            \begin{widetext}
+                \begin{align}
+                    \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
                     &=
                     \Biggl(
-                        \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                        - \left[
-                            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                        \right]^2
+                        \nabla_k\Biggl[
+                            \frac{\nabla_k\phi_k}{\phi_k}
+                        \Biggr]
                         +
-                        \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
+                        \sum_{m \neq k}^N \nabla_k^2 u_{km}
                     \Biggr)
-                    \bra{\vf{r}}\ket{\Psi_T}
-                    \\
-                    &\qquad
                     +
                     \Biggl(
-                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                        +
-                        \sum_{m \neq k}^N\nabla_k u(r_{km})
-                    \Biggr)^2
-                    \bra{\vf{r}}\ket{\Psi_T}.
-                \end{align}
-                We now divide by the trial wavefunction. This simplifies the
-                calculations and is more similar to the expression for the local
-                energy.
-                \begin{align}
-                    \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
-                    {\bra{\vf{r}}\ket{\Psi_T}}
-                    &=
-                    \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                    - \left[
-                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                    \right]^2
-                    +
-                    \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
-                    +
-                    \Biggl(
-                        \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                        +
-                        \sum_{m \neq k}^N\nabla_k u(r_{km})
+                        \frac{\nabla_k\phi_k}{\phi_k}
+                        + \sum_{m \neq k}^N
+                        \nabla_k u_{km}
                     \Biggr)^2 \\
                     &=
-                    \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                    +
-                    \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
-                    +
-                    \frac{2\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-                    \sum_{m \neq k}^N\nabla_k u(r_{km})
-                    +
-                    \Biggl(
-                        \sum_{m \neq k}^N\nabla_k u(r_{km})
-                    \Biggr)^2
+                    \frac{\nabla_k^2\phi_k}{\phi_k}
+                    + 2\frac{\nabla_k\phi_k}{\phi_k}
+                    \sum_{m \neq k}\nabla_k u_{km}
+                    + \sum_{m\neq k}^N\nabla_k^2 u_{km}
+                    + \Biggl(
+                        \sum_{m \neq k}^N\nabla_k u_{km}
+                    \Biggr)^2.
                 \end{align}
             \end{widetext}
+            %\begin{widetext}
+            %    We can now find the Laplacian of the trial wavefunction. This
+            %    gives
+            %    \begin{align}
+            %        \nabla_k^2\bra{\vf{r}}\ket{\Psi_T}
+            %        &= \nabla_k\nabla_k\bra{\vf{r}}\ket{\Psi_T}
+            %        = \nabla_k
+            %        \Biggl(
+            %            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %            +
+            %            \sum_{m \neq k}^N\nabla_k u(r_{km})
+            %        \Biggr)
+            %        \bra{\vf{r}}\ket{\Psi_T} \\
+            %        &=
+            %        \Biggl(
+            %            \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %            + \left[
+            %                \nabla\frac{1}{\phi(\vf{r}_k)}
+            %            \right]
+            %            \nabla_k\phi(\vf{r}_k)
+            %            +
+            %            \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
+            %        \Biggr)
+            %        \bra{\vf{r}}\ket{\Psi_T}
+            %        \\
+            %        &\qquad
+            %        +
+            %        \Biggl(
+            %            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %            +
+            %            \sum_{m \neq k}^N\nabla_k u(r_{km})
+            %        \Biggr)
+            %        \nabla_k\bra{\vf{r}}\ket{\Psi_T} \\
+            %        &=
+            %        \Biggl(
+            %            \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %            - \left[
+            %                \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %            \right]^2
+            %            +
+            %            \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
+            %        \Biggr)
+            %        \bra{\vf{r}}\ket{\Psi_T}
+            %        \\
+            %        &\qquad
+            %        +
+            %        \Biggl(
+            %            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %            +
+            %            \sum_{m \neq k}^N\nabla_k u(r_{km})
+            %        \Biggr)^2
+            %        \bra{\vf{r}}\ket{\Psi_T}.
+            %    \end{align}
+            %    We now divide by the trial wavefunction. This simplifies the
+            %    calculations and is more similar to the expression for the local
+            %    energy.
+            %    \begin{align}
+            %        \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
+            %        {\bra{\vf{r}}\ket{\Psi_T}}
+            %        &=
+            %        \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %        - \left[
+            %            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %        \right]^2
+            %        +
+            %        \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
+            %        +
+            %        \Biggl(
+            %            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %            +
+            %            \sum_{m \neq k}^N\nabla_k u(r_{km})
+            %        \Biggr)^2 \\
+            %        &=
+            %        \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %        +
+            %        \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
+            %        +
+            %        \frac{2\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
+            %        \sum_{m \neq k}^N\nabla_k u(r_{km})
+            %        +
+            %        \Biggl(
+            %            \sum_{m \neq k}^N\nabla_k u(r_{km})
+            %        \Biggr)^2
+            %    \end{align}
+            %\end{widetext}
             To go from here we have to find the gradient and the Laplacian of
-            the single particle functions, $\phi(\vf{r}_k)$, and the interaction
-            functions $u(r_{km})$. For the single particle functions we use
+            the single particle functions, $\phi_k$, and the interaction
+            functions $u_{km}$. For the single particle functions we use
             Cartesian coordinates when finding the derivatives whereas we for
@@
             \begin{align}
-                \nabla_k\phi(\vf{r}_k)
+                \nabla_k\phi_k
                 &=
@@
                 (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
-                \phi(\vf{r}_k),
+                \phi_k.
             \end{align}
-            note that the subscripts on the unit vectors $\vf{e}_i$ are
+            Note that the subscripts on the unit vectors $\vf{e}_i$ are
             \textit{not} the same as the subscripts used for its components. The
@@
             \begin{align}
-                \nabla_k^2\phi(\vf{r}_k)
+                \nabla_k^2\phi_k
                 &=
+                \Bigl[
                     -2\alpha
@@
                     \bigr)
-                \phi(\vf{r}_k)
-                \nonumber \\
-                &\qquad
                     + 4\alpha^2
@@
                     \bigr)
-                \phi(\vf{r}_k).
+                \Bigr]
+                \phi_k.
             \end{align}

### 2018-02-07 project_1/article/article.tex
Add proper substitution for the nabla operator
Add explicit expression for the Laplacian

@@
             \end{align}
-            We now find the Laplacian of the trial wavefunction by taking the
-            gradient of \autoref{eq:gradient_full_wavefunction}.
+            We now find the Laplacian of the trial wavefunction by finding the
+            divergence of \autoref{eq:gradient_full_wavefunction}.
             \begin{align}
@@
             In order to derive the interaction functions we have to do a
-            variable substitution. We replace the derivative of the radial
+            variable substitution using $r_{km} = |\vf{r}_k - \vf{r}_m|$. We can
+            then rewrite the $\nabla_k$-operator as
+            \begin{align}
+                \nabla_k
+                &=
+                \nabla_k
+                \dpd[]{r_{km}}{r_{km}}
+                =
+                \nabla_k r_{km} \dpd[]{}{r_{km}}
+                \\
+                &=
+                \frac{\vf{r}_k - \vf{r}_m}{r_{km}}\dpd[]{}{r_{km}}.
+            \end{align}
+            Applying this version of the $\nabla_k$-operator to $u_{km}$ yields
+            \begin{align}
+                \nabla_k u_{km}
+                &=
+                \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+                \dpd[]{u_{km}}{r_{km}}.
+            \end{align}
+            For the Laplacian we switch a little back and forth between the two
+            ways of representing the $\nabla_k$-operator. We thus get
+            \begin{widetext}
+                \begin{align}
+                    \nabla_k^2 u_{km}
+                    &=
+                    \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+                    + \Biggl[
+                        \nabla_k \frac{1}{r_{km}}
+                    \Biggr]
+                    (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
+                    + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+                    \nabla_k \dpd[]{u_{km}}{r_{km}} \\
+                    &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+                    - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
+                    \dpd[]{u_{km}}{r_{km}}
+                    + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
+                    \dpd[2]{u_{km}}{r_{km}} \\
+                    &=
+                    \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+                    + \dpd[2]{u_{km}}{r_{km}},
+                \end{align}
+                where $d$ is again the dimensionality of the problem. In total
+                we can state an intermediate version of the local energy as
+                \begin{align}
+                    \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
+                    &=
+                    \frac{\nabla_k^2\phi_k}{\phi_k}
+                    + 2\frac{\nabla_k\phi_k}{\phi_k}
+                    \sum_{m \neq k}^N
+                    \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+                    \dpd[]{u_{km}}{r_{km}}
+                    %\nonumber \\
+                    %&\qquad
+                    + \sum_{m\neq k}^N
+                    \Biggl(
+                        \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+                        + \dpd[2]{u_{km}}{r_{km}}
+                    \Biggr)
+                    \nonumber \\
+                    &\qquad
+                    +
+                    \sum_{m, n \neq k}^N
+                    \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+                    \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
+                    \dpd[]{u_{km}}{r_{km}}
+                    \dpd[]{u_{kn}}{r_{kn}}.
+                \end{align}
+            \end{widetext}
+
             component for particle $k$ by

### 2018-02-07 project_1/article/article.tex
Add derivation of the interaction terms
	Could you try to reproduce equations 40 and 41 @gregwinther?

@@
                 where $d$ is again the dimensionality of the problem. In total
-                we can state an intermediate version of the local energy as
+                we can state an intermediate version of the Laplacian occuring
+                in the local energy as
                 \begin{align}
@@
             \end{widetext}
+            Moving on to the derivatives of the interaction terms, $u_{km}$, to
+            get an explicit expression for the Laplacian.
+            \begin{align}
+                \dpd[]{u_{km}}{r_{km}}
+                &=
+                \frac{a}{r_{km}(r_{km} - a)},
+                \\
+                \dpd[2]{u_{km}}{r_{km}}
+                &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.
+            \end{align}
 
-            component for particle $k$ by
-            \begin{align}
-                \tpl{{r_k}}
-                = \dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}.
-            \end{align}
-            The derivative of the distance $r_{km}$ is given by
-            \begin{gather}
-                r_{km} = |\vf{r}_k - \vf{r}_m|
-                \implies
-                \dpd[]{r_{km}}{r_k}
-                = \frac{r_k - r_m}{|\vf{r}_k - \vf{r}_m|},
-            \end{gather}
-            where in spherical coordinates $\vf{r}_k = r_k\vf{e}_r$.  The
-            gradient of $u(r_{km})$ thus simplifies to only include the radial
-            contribution.
-            \begin{align}
-                \nabla_k u(r_{km})
-                &=
-                \vf{e}_{r}\tpl{{r_k}}u(r_{km})
-                \\
-                &=
-                \vf{e}_r\dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}u(r_{km})
-                \\
-                &= \vf{e}_r\frac{r_k - r_m}{r_{km}}
-                \tpl{{r_{km}}}u(r_{km}).
-            \end{align}
-            \begin{widetext}
-                Using this expression we find the Laplacian to be
-                \begin{align}
-                    \nabla_k^2 u(r_{km})
-                    &=
-                    \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
-                    + \frac{(r_k - r_m)^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
-                    +
-                    \frac{r_k - r_m}{r_{km}}
-                    \dpd[]{r_{km}}{r_k}
-                    \partial_{{r_{km}}}^2 u(r_{km}) \\
-                    &=
-                    \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
-                    + \frac{r_{km}^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
-                    +
-                    \frac{(r_k - r_m)^2}{r_{km}^2}
-                    \partial_{{r_{km}}}^2 u(r_{km}) \\
-                    &=
-                    \frac{2}{r_{km}}\tpl{{r_{km}}}u(r_{km})
-                    + \partial_{{r_{km}}}^2u(r_{km}).
-                \end{align}
-            \end{widetext}
-            We now take the derivate with respect to $r_{km}$ of the interaction
-            functions $u(r_{km})$ to find the closed form expressions. We get
-            \begin{align}
-                \partial_{r_{km}}u(r_{km})
-                &=
-                \partial_{r_{km}}
-                \ln\Biggl(
-                    1 - \frac{a}{r_{km}}
-                \Biggr) \\
-                &=
-                \partial_{r_{km}}
-                \ln\Biggl(
-                    \frac{1}{r_{km}}
-                    (r_{km} - a)
-                \Biggr) \\
-                &=
-                \partial_{r_{km}}
-                \Bigl[
-                    \ln(r_{km} - a)
-                    - \ln(r_{km})
-                \Bigr] \\
-                &=
-                \frac{1}{r_{km} - a} - \frac{1}{r_{km}} \\
-                &=
-                \frac{a}{r_{km}(r_{km} - a)} \\
-                \partial_{r_{km}}^2u(r_{km})
-                &= a\Bigl(
-                    \partial_{r_{km}}\frac{1}{r_{km}}
-                \Bigr)
-                \partial_{r_{km}}\frac{1}{r_{km} - a}
-                \nonumber \\
-                &\qquad
-                + a\frac{1}{r_{km}}
-                \Bigl(
-                    \partial_{r_{km}}\frac{1}{r_{km} - a}
-                \Bigr) \\
-                &=
-                -\frac{a}{r_{km}^2(r_{km} - a)}
-                -\frac{a}{r_{km}(r_{km} - a)^2} \\
-                &=
-                - \frac{2a(r_{km} - 1)}{r_{km}^2(r_{km} - a)^2}.
-            \end{align}
-            Using this we are now able to write the closed form expression of
-            the Laplacian of the trial wavefunction.
-            \begin{widetext}
-                \begin{align}
-                    \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
-                    {\bra{\vf{r}}\ket{\Psi_T}}
-                    &=
-                    -2\alpha
-                    \big(
-                        2 + \beta
-                    \bigr)
-                    + 4\alpha^2
-                    \bigl(
-                        x_k^2 + y_k^2 + \beta^2z_k^2
-                    \bigr)
-                    %\nonumber \\
-                    %&\qquad
-                    +
-                    2a
-                    \sum_{m \neq k}^N
-                    \Biggl[
-                        \frac{1}{r_{km}^2(r_{km} - a)}
-                        - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
-                    \Biggr]
-                    \nonumber \\
-                    &\qquad
-                    -4\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
-                    \cdot
-                    a\sum_{m \neq k}^N
-                    \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
-                    + \Biggl[
-                        a\sum_{m \neq k}^N
-                        \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
-                    \Biggr]^2.
-                \end{align}
-            \end{widetext}
-            To ease the load on the CPU we try to restrict the number of times
-            we evaluate the sums. For brevity we introduce the following
-            functions
-            \begin{align}
-                \xi_k(\vf{r}, a)
-                &= \sum_{m \neq k}^N
-                \Biggl[
-                    \frac{1}{r_{km}^2(r_{km} - a)}
-                    - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
-                \Biggr] \\
-                &=
-                \sum_{m \neq k}^N
-                \frac{1 - a}{r_{km}^2(r_{km} - a)^2}, \\
-                \vfg{\zeta}_k(\vf{r}, a)
-                &=
-                \sum_{m \neq k}^N
-                \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}.
-            \end{align}
-            We can then rewrite the expression for the Laplacian to
-            \begin{widetext}
-                \begin{align}
-                    \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
-                    {\bra{\vf{r}}\ket{\Psi_T}}
-                    &=
-                    -2\alpha
-                    \big(
-                        2 + \beta
-                    \bigr)
-                    + 4\alpha^2
-                    \bigl(
-                        x_k^2 + y_k^2 + \beta^2z_k^2
-                    \bigr)
-                    %\nonumber \\
-                    %&\qquad
-                    +
-                    2a\xi_k(\vf{r}, a)
-                    \nonumber \\
-                    &\qquad
-                    -4a\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
-                    \cdot
-                    \vfg{\zeta}_k(\vf{r}, a)
-                    + a^2\vfg{\zeta}_k(\vf{r}, a)^2.
-                \end{align}
-                Both $\xi_k$ and $\vfg{\zeta}_k$ can be evaluated in a common
-                loop. Using this expression for the Laplacian we can find the
-                local energy from
-                \begin{align}
-                    E_L(\vf{r})
-                    &=
-                    \frac{\bra{\vf{r}}H\ket{\Psi_T}}
-                    {\bra{\vf{r}}\ket{\Psi_T}}
-                    =
-                    \sum_{k}^N
-                    \Biggl(
-                        -\frac{\hbar^2}{2m}
-                        \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
-                        {\bra{\vf{r}}\ket{\Psi_T}}
-                        + V_{\text{ext}}(\vf{r})
-                    \Biggr)
-                    +
-                    \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j).
-                \end{align}
-            \end{widetext}
+            %component for particle $k$ by
+            %\begin{align}
+            %    \tpl{{r_k}}
+            %    = \dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}.
+            %\end{align}
+            %The derivative of the distance $r_{km}$ is given by
+            %\begin{gather}
+            %    r_{km} = |\vf{r}_k - \vf{r}_m|
+            %    \implies
+            %    \dpd[]{r_{km}}{r_k}
+            %    = \frac{r_k - r_m}{|\vf{r}_k - \vf{r}_m|},
+            %\end{gather}
+            %where in spherical coordinates $\vf{r}_k = r_k\vf{e}_r$.  The
+            %gradient of $u(r_{km})$ thus simplifies to only include the radial
+            %contribution.
+            %\begin{align}
+            %    \nabla_k u(r_{km})
+            %    &=
+            %    \vf{e}_{r}\tpl{{r_k}}u(r_{km})
+            %    \\
+            %    &=
+            %    \vf{e}_r\dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}u(r_{km})
+            %    \\
+            %    &= \vf{e}_r\frac{r_k - r_m}{r_{km}}
+            %    \tpl{{r_{km}}}u(r_{km}).
+            %\end{align}
+            %\begin{widetext}
+            %    Using this expression we find the Laplacian to be
+            %    \begin{align}
+            %        \nabla_k^2 u(r_{km})
+            %        &=
+            %        \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
+            %        + \frac{(r_k - r_m)^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
+            %        +
+            %        \frac{r_k - r_m}{r_{km}}
+            %        \dpd[]{r_{km}}{r_k}
+            %        \partial_{{r_{km}}}^2 u(r_{km}) \\
+            %        &=
+            %        \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
+            %        + \frac{r_{km}^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
+            %        +
+            %        \frac{(r_k - r_m)^2}{r_{km}^2}
+            %        \partial_{{r_{km}}}^2 u(r_{km}) \\
+            %        &=
+            %        \frac{2}{r_{km}}\tpl{{r_{km}}}u(r_{km})
+            %        + \partial_{{r_{km}}}^2u(r_{km}).
+            %    \end{align}
+            %\end{widetext}
+            %We now take the derivate with respect to $r_{km}$ of the interaction
+            %functions $u(r_{km})$ to find the closed form expressions. We get
+            %\begin{align}
+            %    \partial_{r_{km}}u(r_{km})
+            %    &=
+            %    \partial_{r_{km}}
+            %    \ln\Biggl(
+            %        1 - \frac{a}{r_{km}}
+            %    \Biggr) \\
+            %    &=
+            %    \partial_{r_{km}}
+            %    \ln\Biggl(
+            %        \frac{1}{r_{km}}
+            %        (r_{km} - a)
+            %    \Biggr) \\
+            %    &=
+            %    \partial_{r_{km}}
+            %    \Bigl[
+            %        \ln(r_{km} - a)
+            %        - \ln(r_{km})
+            %    \Bigr] \\
+            %    &=
+            %    \frac{1}{r_{km} - a} - \frac{1}{r_{km}} \\
+            %    &=
+            %    \frac{a}{r_{km}(r_{km} - a)} \\
+            %    \partial_{r_{km}}^2u(r_{km})
+            %    &= a\Bigl(
+            %        \partial_{r_{km}}\frac{1}{r_{km}}
+            %    \Bigr)
+            %    \partial_{r_{km}}\frac{1}{r_{km} - a}
+            %    \nonumber \\
+            %    &\qquad
+            %    + a\frac{1}{r_{km}}
+            %    \Bigl(
+            %        \partial_{r_{km}}\frac{1}{r_{km} - a}
+            %    \Bigr) \\
+            %    &=
+            %    -\frac{a}{r_{km}^2(r_{km} - a)}
+            %    -\frac{a}{r_{km}(r_{km} - a)^2} \\
+            %    &=
+            %    - \frac{2a(r_{km} - 1)}{r_{km}^2(r_{km} - a)^2}.
+            %\end{align}
+            %Using this we are now able to write the closed form expression of
+            %the Laplacian of the trial wavefunction.
+            %\begin{widetext}
+            %    \begin{align}
+            %        \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
+            %        {\bra{\vf{r}}\ket{\Psi_T}}
+            %        &=
+            %        -2\alpha
+            %        \big(
+            %            2 + \beta
+            %        \bigr)
+            %        + 4\alpha^2
+            %        \bigl(
+            %            x_k^2 + y_k^2 + \beta^2z_k^2
+            %        \bigr)
+            %        %\nonumber \\
+            %        %&\qquad
+            %        +
+            %        2a
+            %        \sum_{m \neq k}^N
+            %        \Biggl[
+            %            \frac{1}{r_{km}^2(r_{km} - a)}
+            %            - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
+            %        \Biggr]
+            %        \nonumber \\
+            %        &\qquad
+            %        -4\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
+            %        \cdot
+            %        a\sum_{m \neq k}^N
+            %        \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
+            %        + \Biggl[
+            %            a\sum_{m \neq k}^N
+            %            \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
+            %        \Biggr]^2.
+            %    \end{align}
+            %\end{widetext}
+            %To ease the load on the CPU we try to restrict the number of times
+            %we evaluate the sums. For brevity we introduce the following
+            %functions
+            %\begin{align}
+            %    \xi_k(\vf{r}, a)
+            %    &= \sum_{m \neq k}^N
+            %    \Biggl[
+            %        \frac{1}{r_{km}^2(r_{km} - a)}
+            %        - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
+            %    \Biggr] \\
+            %    &=
+            %    \sum_{m \neq k}^N
+            %    \frac{1 - a}{r_{km}^2(r_{km} - a)^2}, \\
+            %    \vfg{\zeta}_k(\vf{r}, a)
+            %    &=
+            %    \sum_{m \neq k}^N
+            %    \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}.
+            %\end{align}
+            %We can then rewrite the expression for the Laplacian to
+            %\begin{widetext}
+            %    \begin{align}
+            %        \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
+            %        {\bra{\vf{r}}\ket{\Psi_T}}
+            %        &=
+            %        -2\alpha
+            %        \big(
+            %            2 + \beta
+            %        \bigr)
+            %        + 4\alpha^2
+            %        \bigl(
+            %            x_k^2 + y_k^2 + \beta^2z_k^2
+            %        \bigr)
+            %        %\nonumber \\
+            %        %&\qquad
+            %        +
+            %        2a\xi_k(\vf{r}, a)
+            %        \nonumber \\
+            %        &\qquad
+            %        -4a\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
+            %        \cdot
+            %        \vfg{\zeta}_k(\vf{r}, a)
+            %        + a^2\vfg{\zeta}_k(\vf{r}, a)^2.
+            %    \end{align}
+            %    Both $\xi_k$ and $\vfg{\zeta}_k$ can be evaluated in a common
+            %    loop. Using this expression for the Laplacian we can find the
+            %    local energy from
+            %    \begin{align}
+            %        E_L(\vf{r})
+            %        &=
+            %        \frac{\bra{\vf{r}}H\ket{\Psi_T}}
+            %        {\bra{\vf{r}}\ket{\Psi_T}}
+            %        =
+            %        \sum_{k}^N
+            %        \Biggl(
+            %            -\frac{\hbar^2}{2m}
+            %            \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
+            %            {\bra{\vf{r}}\ket{\Psi_T}}
+            %            + V_{\text{ext}}(\vf{r})
+            %        \Biggr)
+            %        +
+            %        \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j).
+            %    \end{align}
+            %\end{widetext}
 

### 2018-02-20 project_1/vmc_hybrid/vmc/src/solvers/steepest_descent_metropolis.cpp
Troubleshooting. No major changes

@@ -41,12 +41,14 @@ double SteepestDescentMetropolis::steepest_descent(
 
     unsigned int iteration = 0;
 
+    bool direction = true;
+
     while (iteration <= MAX_ITER) 
     {
         iteration++;
 
         // The main attraction
-        alpha = alpha - gamma * gradient;
+        alpha = alpha + ((direction) ? (- gamma * gradient) : (gamma * gradient));
 
         // Storing old energy and running Metropolis-Hastings
         energy_prev = energy;
@@ -58,7 +60,10 @@ double SteepestDescentMetropolis::steepest_descent(
 
         // The crude gamma adaption
         if (energy >= energy_prev) {
-            gamma = 0.5 * gamma;
+            if (gamma > 0.001) {
+                gamma = 0.5 * gamma;
+            }
+            direction = !direction;
         }
 
         std::cout << "---------------" << std::endl;

### 2018-02-20 project_1/article/article.tex

@@
 \documentclass[
-    a4paper, aps, twocolumn, floatfix, superscriptaddress]{revtex4-1}
+    a4paper, aps, twocolumn, floatfix, superscriptaddress, nofootinbib]{revtex4-1}
 
@@
     Something very abstract and clever should go here.
+    \begin{figure}[h!]
+	\centering
+	\includegraphics[width=0.7\textwidth]{loopholes.jpg}
+\end{figure}
 \end{abstract}
@@
 \maketitle
+
+%\begin{widetext}
+
+%\end{widetext}
 
@@
             \end{align}
-            It is worth noting that for $\omega = 1$ we will find a stable value
-            for the local energy when $\alpha = \half$. This happens as the
-            entire sum over all the random walkers disappears.
-
-%            Working in spherical coordinates the gradient and the Laplace
-%            operator (applied to a scalar function) becomes
-%            \begin{align}
-%                \nabla f
-%                &= \vf{e}_r\tpl{r}f
-%                + \vf{e}_{\theta}\frac{1}{r}\tpl{\theta}f
-%                + \vf{e}_{\phi}\frac{1}{r\sin(\theta)}\tpl{\phi}f, \\
-%                \nabla^2 f
-%                &= \frac{1}{r^2}\tpl{r}
-%                \bigl[
-%                    r^2\tpl{r}f
-%                \bigr]
-%                + \frac{1}{r^2\sin(\theta)}\tpl{\theta}
-%                \bigl[
-%                    \sin(\theta)\tpl{\theta}f
-%                \bigr] \nonumber\\
-%                &\qquad
-%                + \frac{1}{r^2\sin^2(\theta)}\tpl{\phi}^2f,
-%            \end{align}
-%            where $\vf{e}_r$ signifies the unit vector in the radial direction
-%            and we use the shorthand notation
-%            \begin{align}
-%                \tpl{r}f \equiv \dpd[]{f}{r}.
-%            \end{align}
-%            As the trial wavefunction for $\beta = 1$ is purely radial we get
-%            the same expression in one, two and three dimensions by setting
-%            \begin{align}
-%                r^2 =
-%                \begin{cases}
-%                    x^2 & \text{one dimension}, \\
-%                    x^2 + y^2 & \text{two dimensions}, \\
-%                    x^2 + y^2 + z^2 & \text{three dimensions}.
-%                \end{cases}
-%            \end{align}
-%            The gradient then becomes
-%            \begin{align}
-%                \nabla_i\bra{\vf{r}}\ket{\Psi_T}
-%                &= \vf{e}_{{r_i}}\tpl{{r_i}}\prod_j^N\exp\bigl[
-%                    -\alpha r_j^2
-%                \bigr] \\
-%                &=
-%                -2\alpha r_i\prod_j^N\exp\bigl[
-%                    -\alpha r_j^2
-%                \bigr]\vf{e}_{{r_i}} \\
-%                &=
-%                -2\alpha r_i\bra{\vf{r}}\ket{\Psi_T}\vf{e}_{{r_i}}.
-%            \end{align}
-%            The Laplacian is given by
-%            \begin{align}
-%                \nabla_i^2\bra{\vf{r}}\ket{\Psi_T}
-%                &= \frac{1}{r_i^2}\tpl{{r_i}}
-%                \biggl\{
-%                    r_i^2\tpl{{r_i}}
-%                    \prod_j^N\exp\bigl[
-%                        -\alpha r_j^2
-%                    \bigr]
-%                \biggr\} \\
-%                &=
-%                \frac{1}{r_i^2}\tpl{{r_i}}
-%                \biggl\{
-%                    -2 \alpha r_i^3
-%                    \prod_j^N\exp\bigl[
-%                        -\alpha r_j^2
-%                    \bigr]
-%                \biggr\} \\
-%                &=
-%                \frac{1}{r_i^2}
-%                \Bigl\{
-%                    -6 \alpha r_i^2
-%                    + 4\alpha^2 r_i^4
-%                \Bigr\}
-%                \prod_j^N\exp\bigl[
-%                    -\alpha r_j^2
-%                \bigr]
-%                \\
-%                &=
-%                \Bigl\{
-%                    -6 \alpha
-%                    + 4\alpha^2 r_i^2
-%                \Bigr\}
-%                \bra{\vf{r}}\ket{\Psi_T}.
-%            \end{align}
-%            This gives the analytical expression for the local energy to be
-%            \begin{align}
-%                E_L(\vf{r})
-%                &= \frac{\bra{\vf{r}}H\ket{\Psi_T}}{\bra{\vf{r}}\ket{\Psi_T}} \\
-%                &= \sum_i^N \Bigl(
-%                    4\alpha^2 r_i^2
-%                    + \half m \omega^2 r_i^2
-%                \Bigr) - 6\alpha.
-%            \end{align}
-%            The \textit{drift force} of the system is given by
-%            \begin{align}
-%                \vf{F}_i
-%                &=
-%                \frac{2\nabla_i\bra{\vf{r}}\ket{\Psi_T}}
-%                {\bra{\vf{r}}\ket{\Psi_T}} \\
-%                &= -4\alpha r_i\vf{e}_{{r_i}}.
-%            \end{align}
+            It is worth noting that for $\alpha = \half\omega$ we will find a
+            stable value. This happens as the entire sum over all the random
+            walkers disappears.
 
@@
             \end{widetext}
-            %\begin{widetext}
-            %    We can now find the Laplacian of the trial wavefunction. This
-            %    gives
-            %    \begin{align}
-            %        \nabla_k^2\bra{\vf{r}}\ket{\Psi_T}
-            %        &= \nabla_k\nabla_k\bra{\vf{r}}\ket{\Psi_T}
-            %        = \nabla_k
-            %        \Biggl(
-            %            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %            +
-            %            \sum_{m \neq k}^N\nabla_k u(r_{km})
-            %        \Biggr)
-            %        \bra{\vf{r}}\ket{\Psi_T} \\
-            %        &=
-            %        \Biggl(
-            %            \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %            + \left[
-            %                \nabla\frac{1}{\phi(\vf{r}_k)}
-            %            \right]
-            %            \nabla_k\phi(\vf{r}_k)
-            %            +
-            %            \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
-            %        \Biggr)
-            %        \bra{\vf{r}}\ket{\Psi_T}
-            %        \\
-            %        &\qquad
-            %        +
-            %        \Biggl(
-            %            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %            +
-            %            \sum_{m \neq k}^N\nabla_k u(r_{km})
-            %        \Biggr)
-            %        \nabla_k\bra{\vf{r}}\ket{\Psi_T} \\
-            %        &=
-            %        \Biggl(
-            %            \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %            - \left[
-            %                \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %            \right]^2
-            %            +
-            %            \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
-            %        \Biggr)
-            %        \bra{\vf{r}}\ket{\Psi_T}
-            %        \\
-            %        &\qquad
-            %        +
-            %        \Biggl(
-            %            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %            +
-            %            \sum_{m \neq k}^N\nabla_k u(r_{km})
-            %        \Biggr)^2
-            %        \bra{\vf{r}}\ket{\Psi_T}.
-            %    \end{align}
-            %    We now divide by the trial wavefunction. This simplifies the
-            %    calculations and is more similar to the expression for the local
-            %    energy.
-            %    \begin{align}
-            %        \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
-            %        {\bra{\vf{r}}\ket{\Psi_T}}
-            %        &=
-            %        \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %        - \left[
-            %            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %        \right]^2
-            %        +
-            %        \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
-            %        +
-            %        \Biggl(
-            %            \frac{\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %            +
-            %            \sum_{m \neq k}^N\nabla_k u(r_{km})
-            %        \Biggr)^2 \\
-            %        &=
-            %        \frac{\nabla_k^2\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %        +
-            %        \sum_{m \neq k}^N\nabla_k^2 u(r_{km})
-            %        +
-            %        \frac{2\nabla_k\phi(\vf{r}_k)}{\phi(\vf{r}_k)}
-            %        \sum_{m \neq k}^N\nabla_k u(r_{km})
-            %        +
-            %        \Biggl(
-            %            \sum_{m \neq k}^N\nabla_k u(r_{km})
-            %        \Biggr)^2
-            %    \end{align}
-            %\end{widetext}
             To go from here we have to find the gradient and the Laplacian of
@@
                     \dpd[]{u_{km}}{r_{km}}
-                    %\nonumber \\
-                    %&\qquad
                     + \sum_{m\neq k}^N
@@
 
-            %component for particle $k$ by
-            %\begin{align}
-            %    \tpl{{r_k}}
-            %    = \dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}.
-            %\end{align}
-            %The derivative of the distance $r_{km}$ is given by
-            %\begin{gather}
-            %    r_{km} = |\vf{r}_k - \vf{r}_m|
-            %    \implies
-            %    \dpd[]{r_{km}}{r_k}
-            %    = \frac{r_k - r_m}{|\vf{r}_k - \vf{r}_m|},
-            %\end{gather}
-            %where in spherical coordinates $\vf{r}_k = r_k\vf{e}_r$.  The
-            %gradient of $u(r_{km})$ thus simplifies to only include the radial
-            %contribution.
-            %\begin{align}
-            %    \nabla_k u(r_{km})
-            %    &=
-            %    \vf{e}_{r}\tpl{{r_k}}u(r_{km})
-            %    \\
-            %    &=
-            %    \vf{e}_r\dpd[]{r_{km}}{r_k}\tpl{{r_{km}}}u(r_{km})
-            %    \\
-            %    &= \vf{e}_r\frac{r_k - r_m}{r_{km}}
-            %    \tpl{{r_{km}}}u(r_{km}).
-            %\end{align}
-            %\begin{widetext}
-            %    Using this expression we find the Laplacian to be
-            %    \begin{align}
-            %        \nabla_k^2 u(r_{km})
-            %        &=
-            %        \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
-            %        + \frac{(r_k - r_m)^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
-            %        +
-            %        \frac{r_k - r_m}{r_{km}}
-            %        \dpd[]{r_{km}}{r_k}
-            %        \partial_{{r_{km}}}^2 u(r_{km}) \\
-            %        &=
-            %        \frac{1}{r_{km}} \tpl{{r_{km}}} u(r_{km})
-            %        + \frac{r_{km}^2}{r_{km}^3} \tpl{{r_{km}}} u(r_{km})
-            %        +
-            %        \frac{(r_k - r_m)^2}{r_{km}^2}
-            %        \partial_{{r_{km}}}^2 u(r_{km}) \\
-            %        &=
-            %        \frac{2}{r_{km}}\tpl{{r_{km}}}u(r_{km})
-            %        + \partial_{{r_{km}}}^2u(r_{km}).
-            %    \end{align}
-            %\end{widetext}
-            %We now take the derivate with respect to $r_{km}$ of the interaction
-            %functions $u(r_{km})$ to find the closed form expressions. We get
-            %\begin{align}
-            %    \partial_{r_{km}}u(r_{km})
-            %    &=
-            %    \partial_{r_{km}}
-            %    \ln\Biggl(
-            %        1 - \frac{a}{r_{km}}
-            %    \Biggr) \\
-            %    &=
-            %    \partial_{r_{km}}
-            %    \ln\Biggl(
-            %        \frac{1}{r_{km}}
-            %        (r_{km} - a)
-            %    \Biggr) \\
-            %    &=
-            %    \partial_{r_{km}}
-            %    \Bigl[
-            %        \ln(r_{km} - a)
-            %        - \ln(r_{km})
-            %    \Bigr] \\
-            %    &=
-            %    \frac{1}{r_{km} - a} - \frac{1}{r_{km}} \\
-            %    &=
-            %    \frac{a}{r_{km}(r_{km} - a)} \\
-            %    \partial_{r_{km}}^2u(r_{km})
-            %    &= a\Bigl(
-            %        \partial_{r_{km}}\frac{1}{r_{km}}
-            %    \Bigr)
-            %    \partial_{r_{km}}\frac{1}{r_{km} - a}
-            %    \nonumber \\
-            %    &\qquad
-            %    + a\frac{1}{r_{km}}
-            %    \Bigl(
-            %        \partial_{r_{km}}\frac{1}{r_{km} - a}
-            %    \Bigr) \\
-            %    &=
-            %    -\frac{a}{r_{km}^2(r_{km} - a)}
-            %    -\frac{a}{r_{km}(r_{km} - a)^2} \\
-            %    &=
-            %    - \frac{2a(r_{km} - 1)}{r_{km}^2(r_{km} - a)^2}.
-            %\end{align}
-            %Using this we are now able to write the closed form expression of
-            %the Laplacian of the trial wavefunction.
-            %\begin{widetext}
-            %    \begin{align}
-            %        \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
-            %        {\bra{\vf{r}}\ket{\Psi_T}}
-            %        &=
-            %        -2\alpha
-            %        \big(
-            %            2 + \beta
-            %        \bigr)
-            %        + 4\alpha^2
-            %        \bigl(
-            %            x_k^2 + y_k^2 + \beta^2z_k^2
-            %        \bigr)
-            %        %\nonumber \\
-            %        %&\qquad
-            %        +
-            %        2a
-            %        \sum_{m \neq k}^N
-            %        \Biggl[
-            %            \frac{1}{r_{km}^2(r_{km} - a)}
-            %            - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
-            %        \Biggr]
-            %        \nonumber \\
-            %        &\qquad
-            %        -4\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
-            %        \cdot
-            %        a\sum_{m \neq k}^N
-            %        \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
-            %        + \Biggl[
-            %            a\sum_{m \neq k}^N
-            %            \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}
-            %        \Biggr]^2.
-            %    \end{align}
-            %\end{widetext}
-            %To ease the load on the CPU we try to restrict the number of times
-            %we evaluate the sums. For brevity we introduce the following
-            %functions
-            %\begin{align}
-            %    \xi_k(\vf{r}, a)
-            %    &= \sum_{m \neq k}^N
-            %    \Biggl[
-            %        \frac{1}{r_{km}^2(r_{km} - a)}
-            %        - \frac{r_{km} - 1}{r_{km}^2(r_{km} - a)^2}
-            %    \Biggr] \\
-            %    &=
-            %    \sum_{m \neq k}^N
-            %    \frac{1 - a}{r_{km}^2(r_{km} - a)^2}, \\
-            %    \vfg{\zeta}_k(\vf{r}, a)
-            %    &=
-            %    \sum_{m \neq k}^N
-            %    \frac{\vf{r}_k - \vf{r}_m}{r_{km}^2(r_{km} - a)}.
-            %\end{align}
-            %We can then rewrite the expression for the Laplacian to
-            %\begin{widetext}
-            %    \begin{align}
-            %        \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
-            %        {\bra{\vf{r}}\ket{\Psi_T}}
-            %        &=
-            %        -2\alpha
-            %        \big(
-            %            2 + \beta
-            %        \bigr)
-            %        + 4\alpha^2
-            %        \bigl(
-            %            x_k^2 + y_k^2 + \beta^2z_k^2
-            %        \bigr)
-            %        %\nonumber \\
-            %        %&\qquad
-            %        +
-            %        2a\xi_k(\vf{r}, a)
-            %        \nonumber \\
-            %        &\qquad
-            %        -4a\alpha(x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
-            %        \cdot
-            %        \vfg{\zeta}_k(\vf{r}, a)
-            %        + a^2\vfg{\zeta}_k(\vf{r}, a)^2.
-            %    \end{align}
-            %    Both $\xi_k$ and $\vfg{\zeta}_k$ can be evaluated in a common
-            %    loop. Using this expression for the Laplacian we can find the
-            %    local energy from
-            %    \begin{align}
-            %        E_L(\vf{r})
-            %        &=
-            %        \frac{\bra{\vf{r}}H\ket{\Psi_T}}
-            %        {\bra{\vf{r}}\ket{\Psi_T}}
-            %        =
-            %        \sum_{k}^N
-            %        \Biggl(
-            %            -\frac{\hbar^2}{2m}
-            %            \frac{\nabla_k^2\bra{\vf{r}}\ket{\Psi_T}}
-            %            {\bra{\vf{r}}\ket{\Psi_T}}
-            %            + V_{\text{ext}}(\vf{r})
-            %        \Biggr)
-            %        +
-            %        \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j).
-            %    \end{align}
-            %\end{widetext}
+\section{Algorithms}
+
+In the project we rely on a Monte Carlo approach of random sampling to obtain numerical results. 
+Ironic, is it not? To rely randomness to solve problems that must be naturally deterministic? 
+We simulate random walks over a volume in order to find optimal parameters in our trial wavefunctions. 
+The most common of such methods, which we make use of herein, is the Metropolis-Hastings algorithm.
+
+\subsection{Metropolis-Hastings Algorithm}
+The Metropolis-Hastings algorithm can in our particular situation be condensed down to the following steps,
+\begin{enumerate}
+	\item The system is initialised by a certain number $N$ of randomly generated positions, or particles. This allows us to evaluate the
+	 wavefunction at these points and compute the local energy $E_L$.
+	\item The initial configuration is changed by setting a new position for one of these particles. The particle is picked at random.
+	\item A ratio between new wavefunction density and the previous (initial) density is computed and compared to a random number. 
+	This acceptance probability decides if the particle move is rejected or accepted. 
+	The particle is only allowed to move a predetermined step length.
+	\item If the particle movement is accepted and the local energy $E_L$ is computed for the new system. 
+	\item Repeat steps until convergence and an optimum is reached.
+\end{enumerate}
+
+The algorithm described above can be applied in an "exhaustive" search of the parameter space in order to find the optimal parameters. 
+Whether a proposed move is accepted or not is determined by a transition probability and the acceptance probability. 
+The strength of the algorithm is that the transition algorithm need not be known.
+
+\subsubsection{Importance Sampling}
+A problem with the naïve Metropolis-Hastings sampling approach is that the sampling of the position space is done with 
+no regard for where we are likely to find a particle. This problem can be remedied through a \emph{importance sampling}. 
+It is reasonable to assume that the particles we erratically scatter in space are prone to movement towards the peaks 
+of the probability density as dictated by the wave function. Consider therefore the Fokker-Planck equation,
+\begin{equation}
+	\frac{\partial \psi}{\partial t} = D \frac{\partial }{\partial x}\left(\frac{\partial }{\partial x} - F \right) \psi,	
+	\label{eq:fokker_planck}
+\end{equation}
+which described the evolution in time of a probability density function, here exchanged for the wavefunction $\psi$. 
+Originaly an equation that models diffusion, we have a diffusion term $D$ and a drift force, 
+\begin{equation}
+	F = \frac{2}{\psi} \frac{\partial \psi}{\partial x}.
+\end{equation}
+
+The one-particle model corresponding to the Fokker-Planck equation is the Langevin equation,
+\begin{equation}
+	\frac{\partial x}{\partial t} = D F(x) + \eta,
+	\label{eq:langevin}
+\end{equation}
+where $\eta$ is a uniformly distributed stochastic variable. Solving Langevin's equation by Eurler's method gives a recursive relation for
+the subsequent new positions of a particle,
+\begin{equation}
+	x_{t+\Delta t} = x_t + D F(x) \Delta t +\xi \sqrt{\Delta t},
+\end{equation}
+given a time step $\Delta t$\footnote{Bear in mind that \autoref{eq:langevin} is only valid as $\Delta t \to 0$, a property stemming from the use of Euler's method.} and a normally distributed stochastic variable $\xi$.
+
+Now we need to change the acceptance probability of the metropolis algorithm to something that takes 
+the new sampling method into account,
+\begin{equation}
+	q(x_{\text{new}}, x) = \frac{G(x, x_{\text{new}}, \Delta t)\abs{\psi_T(x_{\text{new}})}}{G(x_{\text{new}}, x, \Delta t)\abs{\psi_T(x)}}
+\end{equation}
+where $G$ is the Green's function to the Fokker-Planck equation,
+\begin{align}
+	G&(x_{\text{new}}, x, \Delta x) \nonumber \\
+	&= \frac{1}{(4\pi D \Delta t)^{3N/2}}\exp\left(-\frac{(x_{\text{new}} - x - D\Delta t F(x))^2}{4D\Delta t} \right).
+\end{align}
 

### 2018-02-27 project_1/article/article.tex
Add subsection for the drift force

@@
         $E_L$ in terms of the trial wavefunction.
+
+    \subsection{The drift force}
+        A disadvantage in the use of the brute-force Metropolis algorithm is
+        that we might be spending much computational resources in an
+        uninteresting part of configuration space. To make smarter moves we will
+        use the Metropolis-Hastings algorithm (which will be discussed in due
+        time). This algorithm is dependent on the drift force of the system.
+        \begin{align}
+            \vf{F}(\vf{r})
+            &=
+            \frac{2\nabla\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
+        \end{align}
+        Using this expression we are able to move towards parts of
+        configuration space where the gradient increases or decreases yielding a
+        better choice of movements.
 

### 2018-02-27 project_1/vmc_hybrid/scripts/solver_run.py
Imporance metropolis algorithm now give (skewed) results.

@@ -8,13 +8,13 @@ import numpy as np
 num_particles = 1
 num_dimensions = 1
 num_parameters = 1
-spread = 5.0
+spread = 2.0
 step_length = 0.05
-num_samples = 10
+num_samples = 1000000
 mass = 1
 omega = 1
 
-alphas = np.linspace(0.2, 1.1, 22)
+alphas = np.linspace(0.1, 1.1, 21)
 
 wavefunction = PySimpleGaussian(num_particles, num_dimensions, num_parameters, mass, omega)
 wavefunction.set_parameters(np.array([0.5]))
@@ -25,7 +25,7 @@ solver = PyImportanceMetropolis(num_particles)
 solver.run(wavefunction, hamiltonian, step_length, num_samples)
 
 energies = [] 
-'''
+
 for alpha in alphas:
 
     wavefunction.redistribute()
@@ -39,6 +39,3 @@ for alpha in alphas:
 
 plt.plot(alphas, energies)
 plt.show()
-
-'''
-

### 2018-02-27 project_1/vmc_hybrid/vmc/src/solvers/importance_metropolis.cpp

@@ -8,33 +8,38 @@
 
 
 // Function to compute the Green's functions fraction
-double ImportanceMetropolis::greensFraction(Wavefunction *wavefunction, double new_pos, double old_pos, double time_step, double D) {
-
+double ImportanceMetropolis::greensFraction(Wavefunction *wavefunction, double* new_pos, double* old_pos, double time_step, double D) 
+{
     double result;
     double drift_force_old, drift_force_new;
-    double G_1 = 0;
-    double G_2 = 0;
+    double G_numerator = 0;
+    double G_denominator = 0;
     unsigned int i, num_dimensions;
+
+    double first_vector = 0;
+    double second_vector = 0;
 
     num_dimensions = wavefunction->get_num_dimensions();
 
     for (i = 0; i < num_dimensions; i++) {
         
-        drift_force_old = wavefunction->compute_drift_force_component(i, old_pos);
-        drift_force_new = wavefunction->compute_drift_force_component(i, new_pos);
+        drift_force_old = wavefunction->compute_drift_force_component(old_pos[i]);
+        drift_force_new = wavefunction->compute_drift_force_component(new_pos[i]);
 
-        double x_old, x_new;
-        x_old = m_particles[i, old_pos*num_dimensions]
-        x_new = m_particles[i, new_pos*num_dimensions]
+        //double x_old, x_new;
+        //x_old = m_particles[i + old_pos*num_dimensions];
+        //x_new = m_particles[i + new_pos*num_dimensions];
 
-        // Part of this computation must probably be down outside the loop
-        G_1 += exp(- SQUARE(x_old - x_new - D*time_step*driftforce_new) / (4*D*time_step));
-        G_2 += exp(- SQUARE(x_new - x_old - D*time_step*driftforce_old) / (4*D*time_step));
+        first_vector  += SQUARE(new_pos[i] - old_pos[i] - D*time_step*drift_force_new);
+        second_vector += SQUARE(new_pos[i] - old_pos[i] - D*time_step*drift_force_old);
     }
 
-    result = G1 / G2;
+    G_numerator = exp(-  first_vector / (4*D*time_step));
+    G_denominator = exp(- second_vector / (4*D*time_step));
+    
+    result = G_numerator / G_denominator;
 
-   return result; 
+    return result; 
 }
 
 bool ImportanceMetropolis::step(Wavefunction *wavefunction, double step_length)
@@ -51,14 +56,14 @@ bool ImportanceMetropolis::step(Wavefunction *wavefunction, double step_length)
     */
 
[…]

### 2018-02-27 project_1/vmc_hybrid/vmc/src/solvers/importance_metropolis.h

@@ -9,11 +9,11 @@ class ImportanceMetropolis : public MonteCarloMethod
     public:
         using MonteCarloMethod::MonteCarloMethod;
 
+        double greensFraction(Wavefunction *wavefunction, double *new_pos, double *old_pos, double time_step, double D);
+
         bool step(Wavefunction *wavefunction, double step_length);
 
         double run(
                 Wavefunction *wavefunction, Hamiltonian *hamiltonian,
                 double step_length, unsigned int num_samples);
-        
-        double green(double pos1, double pos2, double time_step, double D);
 };

### 2018-02-27 project_1/vmc_hybrid/vmc/src/wavefunctions/simple_gaussian.h

@@ -19,5 +19,6 @@ class SimpleGaussian : public Wavefunction
 
         double evaluate();
         double compute_laplacian();
-        double compute_drift_force_component(unsigned int dimension, unsigned int particle);
+        //double compute_drift_force_component(unsigned int dimension, unsigned int particle);
+        double compute_drift_force_component(double coordinate);
 };

### 2018-02-27 project_1/vmc_hybrid/vmc/src/wavefunctions/wavefunction.h

@@ -71,5 +71,6 @@ class Wavefunction
 
         virtual double evaluate() = 0;
         virtual double compute_laplacian() = 0;
-        virtual double compute_gradient() = 0;
+        //virtual double compute_drift_force_component(unsigned int dimension, unsigned int particle) = 0;
+        virtual double compute_drift_force_component(double coordinate) = 0;
 };

### 2018-02-27 project_1/article/article.tex
Add drift force to the non-interacting system

@@
     \begin{align}
-        H = \sum_{i}^{N}h(\vf{r}_i) + \sum_{i < j}^{N}w(\vf{r}_i, \vf{r}_j),
+        H = \sum_{i = 1}^{N}h(\vf{r}_i) + \sum_{i < j}^{N}w(\vf{r}_i, \vf{r}_j),
     \end{align}
@@
         &= \Biggl(
-            \prod_i^N g(\alpha, \beta, \vf{r}_i)
+            \prod_{i = 1}^N g(\alpha, \beta, \vf{r}_i)
         \Biggr)
@@
             &=
-            \frac{2\nabla\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
+            \sum_{k = 1}^N
+            \vf{F}_k(\vf{r})
+            =
+            \sum_{k = 1}^N
+            \frac{2\nabla_k\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
         \end{align}
@@
         configuration space where the gradient increases or decreases yielding a
-        better choice of movements.
+        better choice of movements. We will mainly be interested in the drift
+        force of a single particle $k$.
 
-        \subsubsection{Non-interacting harmonic oscillator}
-            We start by finding an analytical expression for the local energy in
-            a system where we set $a = 0$ and $\beta = 1$, i.e., a system of
-            non-interacting harmonic oscillators.
+\section{Non-interacting harmonic oscillators}
+    We start by looking at a simple system of non-interacting harmonic
+    oscillators. That is, where $a = 0$ and $\beta = 1$. We thus get the trial
+    wavefunction
     \begin{align}
@@
         = \Phi_T(\vf{r})
-                = \prod_i^N \exp\bigl[
+        = \prod_{i = 1}^N \exp\bigl[
             -\alpha |\vf{r}_i|^2
@@
     \begin{align}
-                H &= \sum_i^N h(\vf{r}_i)
-                = \sum_i^N \Biggl(
+        H &= \sum_{i = 1}^N h(\vf{r}_i)
+        = \sum_{i = 1}^N \Biggl(
             -\frac{\hbar^2}{2m}\nabla_i^2
@@
     \end{align}
-            The hardest part of finding the local energy is computing the
-            kinetic term. Starting with the gradient of the wavefunction in the
-            non-interacting case, we get
+    To find the drift force and the local energy we have to compute the gradient
+    and the Laplacian of the trial wavefunction. The gradient is given by
     \begin{align}
         \nabla_k\Psi_T(\vf{r})
-                &= -2\alpha \vf{r}_k\Psi_T(\vf{r}).
+        &= -2\alpha \vf{r}_k\Psi_T(\vf{r}),
     \end{align}
-            The Laplacian yields
+    whereas the Laplacian yields
     \begin{align}
@@
     where $d$ is the dimensionality of the problem determined by
-            $\vf{r}_k \in \mathbb{R}^d$. This yields the analytical expression
-            for the local energy in the non-interacting case to be
+    $\vf{r}_k \in \mathbb{R}^d$. We can thus use the gradient to find an
+    expression for the drift force for particle $k$.
+    \begin{align}
+        \vf{F}_k(\vf{r})
+        &= -2\alpha\vf{r}_k.
+    \end{align}
+    Using the Laplcian we can compute the kinteic term in the expression for the
+    local energy. We get
     \begin{align}
@@
         &=
-                \sum_i^N
+        \sum_{i = 1}^N
         \Biggl(
@@
         \biggr)
-                \sum_i^N r_i^2.
+        \sum_{i = 1}^N r_i^2.
     \end{align}
-            It is worth noting that for $\alpha = \half\omega$ we will find a
-            stable value. This happens as the entire sum over all the random
-            walkers disappears.
+    It is worth noting that for $\alpha = \pm\half\omega$ we will find a stable
+    value which turns out to be the exact energy minimum. This happens as the
+    entire sum over all the random walkers disappears.
 
-        \subsubsection{The full system}
+\section{Interacting hard sphere bosons}
     Moving to the full system allowing $\beta$ to vary and setting $a
-            \neq 0$ we write the trial wavefunction as
+    \neq 0$ we can write the trial wavefunction as
     \begin{align}

### 2018-02-27 project_1/article/article.tex
Fix layout of the full problem

@@
     \end{align}
-            To further shorten the notation we will use $u_{jk} = u(r_{jk})$.
-            We wish to find an analytical expression for the local energy.
-            Beginning with the gradient we get
+    To further shorten the notation we will use $u_{jk} = u(r_{jk})$. Computing
+    the gradient of the wavefunction we get
     \begin{align}
@@
     \end{align}
-            We now find the Laplacian of the trial wavefunction by finding the
-            divergence of \autoref{eq:gradient_full_wavefunction}.
+
+    The Laplcian of the trial wavefunction is found by finding the divergence of
+    \autoref{eq:gradient_full_wavefunction}.
     \begin{align}
@@
     wavefunction. This yields
-            \begin{widetext}
     \begin{align}
         \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
-                    &=
-                    \Biggl(
-                        \nabla_k\Biggl[
-                            \frac{\nabla_k\phi_k}{\phi_k}
-                        \Biggr]
-                        +
-                        \sum_{m \neq k}^N \nabla_k^2 u_{km}
-                    \Biggr)
-                    +
-                    \Biggl(
-                        \frac{\nabla_k\phi_k}{\phi_k}
-                        + \sum_{m \neq k}^N
-                        \nabla_k u_{km}
-                    \Biggr)^2 \\
         &=
@@
         \sum_{m \neq k}\nabla_k u_{km}
+        \nonumber \\
+        &\qquad
         + \sum_{m\neq k}^N\nabla_k^2 u_{km}
@@
     \end{align}
-            \end{widetext}
     To go from here we have to find the gradient and the Laplacian of
@@
             \big(
-                        2 + \beta
+                d - 1 + \beta
             \bigr)
+            \nonumber \\
+            &\qquad
             + 4\alpha^2
@@
         \Bigr]
-                \phi_k.
+        \phi_k,
     \end{align}
+    with $d$ as the dimensionality of the problem.
     In order to derive the interaction functions we have to do a
@@
     \end{align}
-            For the Laplacian we switch a little back and forth between the two
-            ways of representing the $\nabla_k$-operator. We thus get
+
     \begin{widetext}
+        For the Laplacian we switch a little back and forth between the
+        two ways of representing the $\nabla_k$-operator. We thus get
         \begin{align}

### 2018-02-27 project_1/vmc_hybrid/scripts/solver_run.py
Importance sampling gives correct results.

@@ -5,8 +5,8 @@ from matplotlib import pyplot as plt
 from mpl_toolkits.mplot3d import Axes3D
 import numpy as np
 
-num_particles = 1
-num_dimensions = 1
+num_particles = 100
+num_dimensions = 3
 num_parameters = 1
 spread = 2.0
 step_length = 0.05

### 2018-02-27 project_1/vmc_hybrid/vmc/src/solvers/importance_metropolis.cpp

@@ -30,7 +30,7 @@ double ImportanceMetropolis::greensFraction(Wavefunction *wavefunction, double* 
         //x_old = m_particles[i + old_pos*num_dimensions];
         //x_new = m_particles[i + new_pos*num_dimensions];
 
-        first_vector  += SQUARE(new_pos[i] - old_pos[i] - D*time_step*drift_force_new);
+        first_vector  += SQUARE(old_pos[i] - new_pos[i] - D*time_step*drift_force_new);
         second_vector += SQUARE(new_pos[i] - old_pos[i] - D*time_step*drift_force_old);
     }
 
@@ -113,7 +113,7 @@ bool ImportanceMetropolis::step(Wavefunction *wavefunction, double step_length)
     double greens_fraction = greensFraction(wavefunction, new_pos, old_pos, time_step, D);
 
     // Acceptance weight
-    weight = greens_fraction * (abs(current_wavefunction) / abs(previous_wavefunction));
+    weight = greens_fraction * (SQUARE(current_wavefunction) / SQUARE(previous_wavefunction));
 
     // The same kind of test as before
     if (weight >= m_random_step(m_engine)) {

### 2018-02-27 project_1/article/article.tex
Add explicit drift force for full system
Important typo in eq. 46, Green's function fraction

--- 2018-02-27 project_1/article/article.tex
@@
         $E_L$ in terms of the trial wavefunction.
-
-    \subsection{The drift force}
-        A disadvantage in the use of the brute-force Metropolis algorithm is
-        that we might be spending much computational resources in an
-        uninteresting part of configuration space. To make smarter moves we will
-        use the Metropolis-Hastings algorithm (which will be discussed in due
-        time). This algorithm is dependent on the drift force of the system.
-        \begin{align}
-            \vf{F}(\vf{r})
-            &=
-            \frac{2\nabla\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
-        \end{align}
-        Using this expression we are able to move towards parts of
-        configuration space where the gradient increases or decreases yielding a
-        better choice of movements.
 
@@
 \begin{equation}
-    q(x_{\text{new}}, x) = \frac{G(x, x_{\text{new}}, \Delta t)\abs{\psi_T(x_{\text{new}})}}{G(x_{\text{new}}, x, \Delta t)\abs{\psi_T(x)}}
+	q(x_{\text{new}}, x) = \frac{G(x, x_{\text{new}}, \Delta t)\abs{\psi_T(x_{\text{new}})}^2}{G(x_{\text{new}}, x, \Delta t)\abs{\psi_T(x)}^2}
 \end{equation}

### 2018-02-27 project_1/vmc_hybrid/vmc/src/config/constants.h
Remove constant

@@ -1,4 +1,3 @@
 #pragma once
 
 #define HBAR 1
-#define DELTA_H ((double) 1e-7)

### 2018-02-27 project_1/article/article.tex
Replace tab with spaces
Remove explicit expressions
Add ratio test to algorithm

--- 2018-02-27 project_1/article/article.tex
@@
     \end{align}
+    The local energy and the drift force can now be found by combining these
+    expressions. For brevity, we will not write out the explicit expressions as
+    these will be called by separated functions in our programs.
 
 \section{Algorithms}
+    In the project we rely on a Monte Carlo approach of random sampling to
+    obtain numerical results.  We simulate random walks over a volume in order
+    to find optimal parameters in our trial wavefunctions.  The most common of
+    such methods, which we make use of herein, is the Metropolis-Hastings
+    algorithm.
 
-In the project we rely on a Monte Carlo approach of random sampling to obtain numerical results.
-Ironic, is it not? To rely randomness to solve problems that must be naturally deterministic?
-We simulate random walks over a volume in order to find optimal parameters in our trial wavefunctions.
-The most common of such methods, which we make use of herein, is the Metropolis-Hastings algorithm.
+    \subsection{The Metropolis-Hastings Algorithm}
+        The Metropolis-Hastings algorithm can in our particular situation be
+        condensed down to the following steps:
 
-\subsection{Metropolis-Hastings Algorithm}
-The Metropolis-Hastings algorithm can in our particular situation be condensed down to the following steps,
         \begin{enumerate}
-    \item The system is initialised by a certain number $N$ of randomly generated positions, or particles. This allows us to evaluate the
-     wavefunction at these points and compute the local energy $E_L$.
-    \item The initial configuration is changed by setting a new position for one of these particles. The particle is picked at random.
-    \item A ratio between new wavefunction density and the previous (initial) density is computed and compared to a random number.
-    This acceptance probability decides if the particle move is rejected or accepted.
-    The particle is only allowed to move a predetermined step length.
-    \item If the particle movement is accepted and the local energy $E_L$ is computed for the new system.
+            \item The system is initialised by a certain number $N$ of randomly
+                generated positions, or particles. This allows us to evaluate
+                the wavefunction at these points and compute the local energy
+                $E_L$.
+
+            \item The initial configuration is changed by setting a new position
+                for one of these particles. The particle is picked at random.
+
+            \item A ratio between new wavefunction density and the previous
+                (initial) density is computed and compared to a random number.
+                This acceptance probability decides if the particle move is
+                rejected or accepted.  The particle is only allowed to move a
+                predetermined step length.
+
+            \item If the particle movement is accepted and the local energy
+                $E_L$ is computed for the new system.
+
             \item Repeat steps until convergence and an optimum is reached.
+
         \end{enumerate}
 
-The algorithm described above can be applied in an "exhaustive" search of the parameter space in order to find the optimal parameters.
-Whether a proposed move is accepted or not is determined by a transition probability and the acceptance probability.
-The strength of the algorithm is that the transition algorithm need not be known.
+        The algorithm described above can be applied in an "exhaustive" search
+        of the parameter space in order to find the optimal parameters.  Whether
+        a proposed move is accepted or not is determined by a transition
+        probability and the acceptance probability.  The strength of the
+        algorithm is that the transition algorithm need not be known. For
+        example, the simplest case is to accept the new state, i.e., the new
+        position for the random walker, if the ratio
+        \begin{align}
+            w &= \frac{\Psi_T(\vf{r}_{i + 1})}{\Psi_T(\vf{r}_{i})},
+        \end{align}
+        where $\vf{r}_{i + 1}$ are all the positions at step $i + 1$, is greater
+        than a uniform probability $p \in [0, 1)$.
 
@@
 \begin{equation}
-    q(x_{\text{new}}, x) = \frac{G(x, x_{\text{new}}, \Delta t)\abs{\psi_T(x_{\text{new}})}}{G(x_{\text{new}}, x, \Delta t)\abs{\psi_T(x)}}
+    q(x_{\text{new}}, x) = \frac{G(x, x_{\text{new}}, \Delta t)\abs{\psi_T(x_{\text{new}})}^2}{G(x_{\text{new}}, x, \Delta t)\abs{\psi_T(x)}^2}
 \end{equation}

### 2018-02-27 project_1/vmc_hybrid/scripts/importance_compare.py
Added script that easily compares metropolis with and without importance sampling.

@@ -1,0 +1,54 @@
+import config
+
+from vmc.interface import PySimpleGaussian, PyHarmonicOscillator, PyMetropolisAlgorithm, PySteepestDescentMetropolis, PyImportanceMetropolis
+from matplotlib import pyplot as plt 
+from mpl_toolkits.mplot3d import Axes3D
+import numpy as np
+
+num_particles = 100
+num_dimensions = 3
+num_parameters = 1
+spread = 5.0
+step_length = 0.05
+num_samples = 1000000
+mass = 1
+omega = 1
+
+no_of_alphas = 11
+alphas = np.linspace(0.3, 0.7, no_of_alphas)
+
+wavefunction_1 = PySimpleGaussian(num_particles, num_dimensions, num_parameters, mass, omega)
+wavefunction_2 = PySimpleGaussian(num_particles, num_dimensions, num_parameters, mass, omega)
+hamiltonian_1 = PyHarmonicOscillator(mass, omega)
+hamiltonian_2 = PyHarmonicOscillator(mass, omega)
+regular_metropolis_solver = PyMetropolisAlgorithm(num_particles)
+importance_metropolis_solver = PyImportanceMetropolis(num_particles)
+
+energies = np.zeros((2, no_of_alphas)) 
+
+for i in range(no_of_alphas):
+
+    wavefunction_1.redistribute()
+    wavefunction_2.redistribute()
+    wavefunction_1.set_parameters(np.array([alphas[i]]))
+    wavefunction_2.set_parameters(np.array([alphas[i]]))
+
+    regular_energy    = regular_metropolis_solver.run(wavefunction_1, hamiltonian_1, step_length, num_samples)
+    importance_energy = importance_metropolis_solver.run(wavefunction_2, hamiltonian_2, step_length, num_samples) 
+
+    regular_energy = regular_energy / num_samples
+    importance_energy = importance_energy / num_samples
+
+    energies[0, i] = regular_energy 
+    energies[1, i] = importance_energy 
+
+    print("alpha: {:.3f}, Reg. metr. energy: {:.3f}, Imp. samp. energy. {:.3f}.".format(alphas[i], regular_energy, importance_energy))
+
+
+plt.plot(alphas, energies[0, :], '--r', label="Regular metropolis")
+plt.plot(alphas, energies[1, :], '--b', label="With importance sampling")
+plt.title(str(num_dimensions) + "D, N = " + str(num_particles))
+plt.ylabel("Energy")
+plt.xlabel("alpha")
+plt.legend()
+plt.show()

### 2018-02-27 project_1/vmc_hybrid/scripts/solver_run.py

@@ -1,41 +1,0 @@
-import config
-
-from vmc.interface import PySimpleGaussian, PyHarmonicOscillator, PyMetropolisAlgorithm, PySteepestDescentMetropolis, PyImportanceMetropolis
-from matplotlib import pyplot as plt 
-from mpl_toolkits.mplot3d import Axes3D
-import numpy as np
-
-num_particles = 100
-num_dimensions = 3
-num_parameters = 1
-spread = 2.0
-step_length = 0.05
-num_samples = 1000000
-mass = 1
-omega = 1
-
-alphas = np.linspace(0.1, 1.1, 21)
-
-wavefunction = PySimpleGaussian(num_particles, num_dimensions, num_parameters, mass, omega)
-wavefunction.set_parameters(np.array([0.5]))
-hamiltonian = PyHarmonicOscillator(mass, omega)
-#solver = PyMetropolisAlgorithm(num_particles)
-solver = PyImportanceMetropolis(num_particles)
-
-solver.run(wavefunction, hamiltonian, step_length, num_samples)
-
-energies = [] 
-
-for alpha in alphas:
-
-    wavefunction.redistribute()
-    wavefunction.set_parameters(np.array([alpha]))
-
-    energy = solver.run(wavefunction, hamiltonian, step_length, num_samples)
-
-    energies.append(energy/num_samples)
-
-    print("alpha: {:.3f}, Energy: {:.3f}".format(alpha, energy/num_samples))
-
-plt.plot(alphas, energies)
-plt.show()

### 2018-02-27 project_1/article/article.tex
Add multi-dimensional importance sampling

@@
             \frac{2\nabla_k\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
+            \label{eq:drift_force}
         \end{align}
@@
         \begin{align}
-            w &= \frac{\Psi_T(\vf{r}_{i + 1})}{\Psi_T(\vf{r}_{i})},
+            q(\vf{r}_{i + 1}, \vf{r}_i)
+            &=
+            \frac{\left|\Psi_T(\vf{r}_{i + 1})\right|^2}
+            {\left|\Psi_T(\vf{r}_{i})\right|^2},
         \end{align}
@@
         \subsubsection{Importance Sampling}
-A problem with the naïve Metropolis-Hastings sampling approach is that the sampling of the position space is done with
-no regard for where we are likely to find a particle. This problem can be remedied through a \emph{importance sampling}.
-It is reasonable to assume that the particles we erratically scatter in space are prone to movement towards the peaks
-of the probability density as dictated by the wave function. Consider therefore the Fokker-Planck equation,
-\begin{equation}
-    \frac{\partial \psi}{\partial t} = D \frac{\partial }{\partial x}\left(\frac{\partial }{\partial x} - F \right) \psi,
+            A problem with the naïve Metropolis-Hastings sampling approach is
+            that the sampling of position space is done with no regard for where
+            we are likely to find a particle. This problem can be remedied
+            through \emph{importance sampling}.  It is reasonable to assume that
+            the particles we erratically scatter in space are prone to movement
+            towards the peaks of the probability density as dictated by the wave
+            function. Consider therefore the Fokker-Planck equation,
+            \begin{align}
+                \dpd[]{\Psi_T}{t}
+                &=
+                D \nabla\cdot
+                \left(
+                    \nabla
+                    - \vf{F}
+                \right) \Psi_T,
                 \label{eq:fokker_planck}
-\end{equation}
-which described the evolution in time of a probability density function, here exchanged for the wavefunction $\psi$.
-Originaly an equation that models diffusion, we have a diffusion term $D$ and a drift force,
-\begin{equation}
-    F = \frac{2}{\psi} \frac{\partial \psi}{\partial x}.
-\end{equation}
+            \end{align}
+            which describes the evolution in time of a probability density
+            function. In our case this is the trial wavefunction $\Psi_T$.
+            Originaly an equation that models diffusion, we have a diffusion
+            term $D$ and the drift force \autoref{eq:drift_force}. In our case
+            the diffusion term $D$ is simply $1/2$ from the kinetic energy (in
+            natural units).
 
-The one-particle model corresponding to the Fokker-Planck equation is the Langevin equation,
-\begin{equation}
-    \frac{\partial x}{\partial t} = D F(x) + \eta,
+            We use the Langevin equation to find the new position of the
+            particle.
+            \begin{align}
+                \dpd[]{\vf{r}}{t}
+                &=
+                D\vf{F}(\vf{r}) + \vfg{\eta},
                 \label{eq:langevin}
-\end{equation}
-where $\eta$ is a uniformly distributed stochastic variable. Solving Langevin's equation by Eurler's method gives a recursive relation for
-the subsequent new positions of a particle,
-\begin{equation}
-    x_{t+\Delta t} = x_t + D F(x) \Delta t +\xi \sqrt{\Delta t},
-\end{equation}
-given a time step $\Delta t$\footnote{Bear in mind that \autoref{eq:langevin} is only valid as $\Delta t \to 0$, a property stemming from the use of Euler's method.} and a normally distributed stochastic variable $\xi$.
+            \end{align}
+            where $\vfg{\eta}$ is a uniformly distributed stochastic variable
+            for each dimenion.  Solving Langevin's equation by Euler's method
+            gives a recursive relation for the subsequent new positions of a
+            particle.
+            \begin{align}
+                \vf{r}_{i + 1}
+                &=
+                \vf{r}_i + D\vf{F}(\vf{r}_i)\Delta t
+                + \vfg{\xi}\sqrt{\Delta t},
+            \end{align}
+            for a given time step $\Delta t$
+            \footnote{
+                Bear in mind that \autoref{eq:langevin} is only valid as $\Delta
+                t \to 0$, a property stemming from the use of Euler's method.
+            }
+            and a normally distributed stochastic variable $\vfg{\xi}$.
 
-Now we need to change the acceptance probability of the metropolis algorithm to something that takes
-the new sampling method into account,
-\begin{equation}
-    q(x_{\text{new}}, x) = \frac{G(x, x_{\text{new}}, \Delta t)\abs{\psi_T(x_{\text{new}})}^2}{G(x_{\text{new}}, x, \Delta t)\abs{\psi_T(x)}^2}
-\end{equation}
-where $G$ is the Green's function to the Fokker-Planck equation,
+            Now we need to change the acceptance probability of the metropolis
+            algorithm to something that takes the new sampling method into
+            account.
             \begin{align}
-    G&(x_{\text{new}}, x, \Delta x) \nonumber \\
-    &= \frac{1}{(4\pi D \Delta t)^{3N/2}}\exp\left(-\frac{(x_{\text{new}} - x - D\Delta t F(x))^2}{4D\Delta t} \right).
+                q(\vf{r}_{i + 1}, \vf{r}_i)
+                &=
+                \frac{
+                    G(\vf{r}_{i + 1}, \vf{r}_{i}, \Delta t)
+                    \left|\Psi_T(\vf{r}_{i + 1})\right|^2
+                }
+                {
+                    G(\vf{r}_{i}, \vf{r}_{i + 1}, \Delta t)
+                    \left|\Psi_T(\vf{r}_{i})\right|^2
+                },
             \end{align}
+            where $G$ is the Green's function of the Fokker-Planck equation
+            given by
+            \begin{align}
+                G(\vf{r}_{i + 1}, \vf{r}_i, \Delta t)
+                &=
+                \exp\Biggl(
+                    -\frac{\bigl[
+                        \vf{r}_{i + 1}
+                        - \vf{r}_{i}
+                        - D\vf{F}(\vf{r}_i)\Delta t
+                    \bigr]^2}{4D\Delta t}
+                \Biggr)
+                \nonumber \\
+                &\qquad
+                \times
+                \frac{1}{(4\pi D\Delta t)^{dN/2}},
+            \end{align}
+            where $d$ is the dimensionality.
 

### 2018-03-01 project_1/article/article.tex
Begun statistical analysis.

@@
 
+	\subsection{Statistical Analysis}
+		If the results of the metropolis sampling were completely uncorrelated,
+		it would be enough to compute the standard deviation in a familiar way,
+		\begin{equation}
+			\sigma = \sqrt{\frac{1}{N}(\ev{E^2_L} - \ev{E_L}^2},
+			\label{eq:standard_deviation}
+		\end{equation}
+		where $N$ is the number of samples, or Monte-Carlo cycles, in the experiment.
+		However, it is reasonable to assume that the data we are dealing with in this
+		study is liable to suffer from \emph{autocorrelation} and \autoref{eq:standard_deviation}
+		does not hold. The prevailing
+		definition of autocorrelation in a data stream or signal is correlation 
+		between a delay of the signal and the original signal. One would be interested to find the
+		delay, or lag, in the signal at which the "self-correlation" is highest. We refer to this
+		spacing as $d$, and define the following correlation function,
+		\begin{equation}
+			f_d = \frac{1}{n -d} \sum_{k=1}^{n-d} (x_k - \bar{x}_n)(x_{k+d} - \bar{x}_n).
+			\label{eq:correlation_f_d}
+		\end{equation}
+		The keen reader would have noticed that the function $f_d$ in \autoref{eq:correlation_f_d}
+		would be equal to the sample variance for $d=0$. We can now define the \emph{autocorrelation function}
+		\begin{equation}
+			\kappa_d = \frac{f_d}{\text{Var}(x)},
+			\label{eq:autocorrelation_function}
+		\end{equation}
+		which is equal to $1$ if the data exhibits no autocorrealtions, equating to $d=0$. From the autocorrelation
+		function (\ref{eq:autocorrelation_function}) we in turn define the \emph{autocorrelation time},
+		\begin{equation}
+			\tau = 1 + 2 \sum_{d=1}^{n-1},
+		\end{equation}
+		notice that the autocorrelation time is $1$ for a correlation free experiment.
 \end{document}

### 2018-03-01 project_1/vmc_hybrid/vmc/interface.pyx
Include thermalization in the sampling

@@ -120,7 +120,12 @@ cdef class PySampler:
     def get_ratio_of_accepted_steps(self):
         return self.sampler.get_ratio_of_accepted_steps()
 
-    def sample(self, unsigned int num_samples, double step_length):
+    def sample(self, unsigned int num_samples, double step_length,
+            unsigned int num_thermalization_steps=0):
+
+        if num_thermalization_steps > 0:
+            self.sampler.sample(num_thermalization_steps, step_length)
+
         self.sampler.sample(num_samples, step_length)
 
 cdef class PyMetropolisAlgorithm(PyMonteCarloMethod):

### 2018-03-01 project_1/article/article.tex
Blocking

@@
 		notice that the autocorrelation time is $1$ for a correlation free experiment.
+
+		We are now able to make a correction to the expression for the standard deviation 
+		improving on \autoref{eq:standard_deviation} by taking correlation into account,
+		\begin{equation}
+			\sigma =\sqrt{\frac{1 + 2\tau / \Delta t}{N}(\ev{E^2_L} - \ev{E_L}^2)},
+		\end{equation}
+		where $\Delta t$ is the time between each sample. The main problem at this point is to 
+		find $\tau$, which we do not know for any given system and it is generally very expensive to
+		compute. In order to find a good estimate of $\tau$ we use a procedure called \emph{blocking}.
+
+		\subsubsection{Blocking}
+		In the method of blocking we group the samples into blocks of increasing size. If one where to compute
+		the standard deviation for each block, one should see the variance increase with the block size.
+		The standard deviation would only increase up to a certain point, from whence it would stay
+		almost constant. What is happening is that we have reached a point where a particular
+		sample from one block is no longer correlated with a corresponding sample from an adjacent block.
+		The block size for this point of convergence now functions as an estimate for the autocorrelation 
+		time $\tau$.
+
 \end{document}

### 2018-03-02 project_1/vmc_hybrid/scripts/steepest_descent.py
First OK results for steepest descent acchieved

@@ -1,0 +1,82 @@
+'''
+    STEEPEST DESCENT (aka Gradient Descent)
+    The idea is that a function F would decrease fastest in
+    the direction of - Del F(alpha),
+    (alpha is a variational parameter).
+    We thereby have the iterative scheme:
+
+    alpha[k + 1] = alpha[k] - gamma * Del F(alpha[k])
+
+    if gamma is small enough we have $F(alpha[k]) >= F(alpha[k+1])
+    for all k, and hopefulle we have convergense to minimum.
+'''
+
+import config
+
+from vmc.interface import PySimpleGaussian, PyHarmonicOscillator, PyImportanceMetropolis, PySampler
+from matplotlib import pyplot as plt
+import numpy as np
+
+num_particles = 10
+num_dimensions = 1
+num_parameters = 1
+spread = 2.0
+step_length = 0.05
+num_samples = 1000000
+
+mass = 1
+omega = 1
+
+num_local_energies = 100
+stride_local_energies = num_samples // num_local_energies
+
+
+hamiltonian = PyHarmonicOscillator(mass, omega)
+wavefunction = PySimpleGaussian(num_particles, num_dimensions, num_parameters, mass, omega)
+solver = PyImportanceMetropolis(num_particles)
+sampler = PySampler(wavefunction, hamiltonian, solver, num_local_energies, stride_local_energies)
+
+alphas = np.linspace(0.1, 1.1, 11)
+
+'''
+for alpha in alphas:
+    print("Alpha = ", alpha)
+    wavefunction.set_parameters(np.array([alpha]))
+    print("Start: ", sampler.get_energy_gradient())
+    sampler.sample(num_samples, step_length)
+    print("End: ", sampler.get_energy_gradient())
+'''
+
+alpha = 0.7 # start
+
+MAX_ITER = 100
+iterations = 0
+gamma = 0.01
+wavefunction.set_parameters(np.array([alpha]))
+sampler.sample(num_samples, step_length)
+gradient = sampler.get_energy_gradient()
+energy = sampler.get_energy()
+
+alphas = np.zeros(MAX_ITER)
+
+while (iterations < MAX_ITER):
+
+    print("Alpha = ", alpha)
+    print("Gradient = ", gradient)
+    alpha = alpha - gamma * gradient
+    wavefunction.set_parameters(np.array([alpha]))
+    energy_prev = energy
+    energy = sampler.get_energy()
+    
[…]

### 2018-03-02 project_1/vmc_hybrid/vmc/interface.pyx

@@ -122,6 +122,9 @@ cdef class PySampler:
 
     def sample(self, unsigned int num_samples, double step_length):
         self.sampler.sample(num_samples, step_length)
+    
+    def get_energy_gradient(self):
+        return self.sampler.get_energy_gradient()
 
 cdef class PyMetropolisAlgorithm(PyMonteCarloMethod):
 

### 2018-03-02 project_1/vmc_hybrid/vmc/src/hamiltonians/hamiltonian.h

@@ -7,4 +7,5 @@ class Hamiltonian
     public:
         virtual double compute_local_energy(Wavefunction *wavefunction) = 0;
         virtual double compute_potential_energy(Wavefunction *wavefunction) = 0;
+        virtual double compute_local_energy_gradient(Wavefunction *wavefunction) = 0;
 };

### 2018-03-02 project_1/vmc_hybrid/vmc/src/sampler/sampler.cpp

@@ -65,3 +65,11 @@ void Sampler::sample(unsigned int num_samples, double step_length)
     /* Compute the variance */
     m_variance = m_energy_squared - SQUARE(m_energy);
 }
+
+double Sampler::get_energy_gradient()
+{
+    // Not done. Must sum up all particles.
+    // Never mind. Already done in Wavefunction:.compute_position_squared_sum()
+    // Consider moving to header file only
+    return m_hamiltonian->compute_local_energy_gradient(m_wavefunction);
+}

### 2018-03-02 project_1/vmc_hybrid/vmc/src/sampler/sampler.h

@@ -72,4 +72,6 @@ class Sampler
         {
             return ((double) m_num_accepted_steps)/((double) m_num_steps);
         }
+
+        double get_energy_gradient();
 };

### 2018-03-02 project_1/vmc_hybrid/vmc/src/wavefunctions/simple_gaussian.cpp

@@ -53,9 +53,9 @@ double SimpleGaussian::compute_laplacian_alpha_derivative()
     position_squared_sum = compute_position_squared_sum();
 
     /* Compute the alpha derivative of the laplacian */
-    laplacian_alpha_derivative = (double) -2 * m_num_dimensions
+    laplacian_alpha_derivative = (double) -2 * m_num_dimensions * m_num_particles
         + 8 * alpha * position_squared_sum;
-    
+
     return laplacian_alpha_derivative;
 }
 

### 2018-03-02 project_1/article/article.tex

@@
     \end{align}
-    It is worth noting that for $\alpha = \pm\half\omega$ we will find a stable
-    value which turns out to be the exact energy minimum. This happens as the
-    entire sum over all the random walkers disappears.
+    It is worth noting that for $\alpha = \half\omega$ ($\alpha$ is required to
+    be positive) we will find a stable value which turns out to be the exact
+    energy minimum. This happens as the entire sum over all the random walkers
+    disappears.
+
+    \subsection{Exact variational energy}
+        As the system is non-interacting and consisting of Gaussians we can
+        find an expression for the exact energy as a function of the
+        variational parameter $\alpha$, i.e.,
+        \begin{align}
+            E(\alpha)
+            &=
+            \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}.
+        \end{align}
+
+        % TODO: Add the computation of this expression
+        The final result for the energy is
+        \begin{align}
+            E(\alpha)
+            &=
+            \left(
+                \frac{\hbar^2 \alpha}{2m}
+                + \frac{m\omega^2}{8\alpha}
+            \right)dN.
+        \end{align}
+        By minimizing this expression, i.e., finding the derivative of the
+        energy with respect to $\alpha$ and equating this to zero, yields the
+        expected minimum of variational energy to be
+        \begin{align}
+            \dod[]{E(\alpha)}{\alpha} = 0
+            \implies
+            \alpha_0 = \frac{m\omega}{2\hbar},
+        \end{align}
+        which in natural units reduces to $\alpha_0 = \half\omega$. The energy at
+        this value of $\alpha$ (in natural units) is then
+        \begin{align}
+            E(\alpha_0)
+            &=
+            \frac{\omega dN}{2}.
+        \end{align}
 

### 2018-03-03 project_1/vmc_hybrid/scripts/steepest_descent.py
No significant changes

@@ -17,11 +17,11 @@ from vmc.interface import PySimpleGaussian, PyHarmonicOscillator, PyImportanceMe
 from matplotlib import pyplot as plt
 import numpy as np
 
-num_particles = 10
+num_particles = 20
 num_dimensions = 1
 num_parameters = 1
 spread = 1.0
-step_length = 0.05
+step_length = 0.01
 num_samples = int(1e6)
 num_thermalization_steps = int(1e6)
 

### 2018-03-03 project_1/vmc_hybrid/scripts/steepest_descent.py
Remove explicit float formatting

@@ -45,9 +45,9 @@ for i in range(1, experiments + 1):
     sampler.sample(num_samples, step_length, num_thermalization_steps)
     gradient += sampler.get_energy_gradient()
     wavefunction.redistribute()
-    print ("Intermediate gradient: {0:f}".format(gradient/i))
+    print ("Intermediate gradient: {0}".format(gradient/i))
 
-print ("Final gradient: {0:f}".format(gradient/experiments))
+print ("Final gradient: {0}".format(gradient/experiments))
 __import__('sys').exit()
 
 alphas = np.linspace(0.1, 1.1, 11)

### 2018-03-03 project_1/vmc_hybrid/vmc/interface.pyx
Use std::valarray for gradient

@@ -129,8 +129,17 @@ cdef class PySampler:
         self.sampler.sample(num_samples, step_length)
 
     def get_energy_gradient(self):
-        # TODO: This should return a numpy array
-        return self.sampler.get_energy_gradient()
+        cdef valarray[double] _energy_gradient
+        cdef np.ndarray[double, ndim=1, mode="c"] energy_gradient
+        cdef unsigned int i
+
+        _energy_gradient = self.sampler.get_energy_gradient()
+        energy_gradient = np.zeros(_energy_gradient.size())
+
+        for i in range(energy_gradient.size):
+            energy_gradient[i] = _energy_gradient[i]
+
+        return energy_gradient
 
 cdef class PyMetropolisAlgorithm(PyMonteCarloMethod):
 

### 2018-03-03 project_1/vmc_hybrid/vmc/src/hamiltonians/hamiltonian.h

@@ -1,5 +1,7 @@
 #pragma once
 
+#include <valarray>
+
 #include "wavefunction.h"
 
 class Hamiltonian
@@ -7,5 +9,6 @@ class Hamiltonian
     public:
         virtual double compute_local_energy(Wavefunction *wavefunction) = 0;
         virtual double compute_potential_energy(Wavefunction *wavefunction) = 0;
-        virtual double compute_local_energy_gradient(Wavefunction *wavefunction) = 0;
+        virtual std::valarray<double> compute_local_energy_gradient(
+                Wavefunction *wavefunction) = 0;
 };

### 2018-03-03 project_1/vmc_hybrid/vmc/src/hamiltonians/harmonic_oscillator.cpp

@@ -1,4 +1,5 @@
 #include <cassert>
+#include <valarray>
 
 #include "hamiltonian.h"
 #include "harmonic_oscillator.h"
@@ -35,13 +36,14 @@ double HarmonicOscillator::compute_potential_energy(Wavefunction *wavefunction)
     return potential_energy;
 }
 
-double HarmonicOscillator::compute_local_energy_gradient(Wavefunction *wavefunction)
+std::valarray<double> HarmonicOscillator::compute_local_energy_gradient(
+        Wavefunction *wavefunction)
 {
-    double local_energy_gradient;
+    std::valarray<double> local_energy_gradient;
 
     local_energy_gradient =
-        -SQUARE(HBAR)*wavefunction->compute_laplacian_alpha_derivative();
-    local_energy_gradient /= 2*m_mass;
+        wavefunction->compute_laplacian_variational_gradient();
+    local_energy_gradient *= -SQUARE(HBAR)/(2*m_mass);
 
     return local_energy_gradient;
 }

### 2018-03-03 project_1/vmc_hybrid/vmc/src/hamiltonians/harmonic_oscillator.h

@@ -1,5 +1,7 @@
 #pragma once
 
+#include <valarray>
+
 #include "hamiltonian.h"
 #include "wavefunction.h"
 
@@ -9,7 +11,8 @@ class HarmonicOscillator : public Hamiltonian
         HarmonicOscillator(double mass, double omega);
         double compute_local_energy(Wavefunction *wavefunction);
         double compute_potential_energy(Wavefunction *wavefunction);
-        double compute_local_energy_gradient(Wavefunction *wavefunction);
+        std::valarray<double> compute_local_energy_gradient(
+                Wavefunction *wavefunction);
 
     private:
         double m_omega;

### 2018-03-03 project_1/vmc_hybrid/vmc/src/sampler/sampler.cpp

@@ -1,3 +1,5 @@
+#include <valarray>
+
 #include "sampler.h"
 #include "monte_carlo_method.h"
 #include "wavefunction.h"
@@ -15,6 +17,8 @@ Sampler::Sampler(
     m_wavefunction = wavefunction;
     m_hamiltonian = hamiltonian;
     m_solver = solver;
+    m_energy_gradient =
+        std::valarray<double>(wavefunction->get_num_parameters());
     m_num_local_energies = num_local_energies;
     m_stride_local_energies = stride_local_energies;
     m_local_energies = local_energies;
@@ -52,7 +56,6 @@ void Sampler::sample(unsigned int num_samples, double step_length)
         /* Add local energy squared */
         m_energy_squared += SQUARE(current_local_energy);
         /* Add local energy gradient */
-        /* TODO: Should be a vector */
         m_energy_gradient +=
             m_hamiltonian->compute_local_energy_gradient(m_wavefunction);
 

### 2018-03-03 project_1/vmc_hybrid/vmc/src/sampler/sampler.h

@@ -1,5 +1,7 @@
 #pragma once
 
+#include <valarray>
+
 #include "monte_carlo_method.h"
 #include "wavefunction.h"
 #include "hamiltonian.h"
@@ -14,8 +16,7 @@ class Sampler
         double m_energy;
         double m_energy_squared;
         double m_variance;
-        /* TODO: This should be a vector for each variational parameter */
-        double m_energy_gradient;
+        std::valarray<double> m_energy_gradient;
 
         unsigned int m_num_local_energies;
         unsigned int m_stride_local_energies;
@@ -40,7 +41,6 @@ class Sampler
             m_energy = 0;
             m_energy_squared = 0;
             m_variance = 0;
-            /* TODO: Should be a vector */
             m_energy_gradient = 0;
 
             if (m_num_local_energies > 0) {
@@ -54,7 +54,6 @@ class Sampler
         {
             m_energy /= m_num_steps;
             m_energy_squared /= m_num_steps;
-            /* TODO: Vector */
             m_energy_gradient /= m_num_steps;
         }
 
@@ -80,9 +79,8 @@ class Sampler
             return ((double) m_num_accepted_steps)/((double) m_num_steps);
         }
 
-        double get_energy_gradient()
+        std::valarray<double> get_energy_gradient()
         {
-            /* TODO: Vector */
             return m_energy_gradient;
         }
 };

### 2018-03-03 project_1/vmc_hybrid/vmc/src/wavefunctions/simple_gaussian.cpp

@@ -1,10 +1,10 @@
 #include <math.h>
+#include <valarray>
 
 #include "wavefunction.h"
 #include "simple_gaussian.h"
 #include "math_macros.h"
 #include "constants.h"
-#include <iostream>
 
 SimpleGaussian::SimpleGaussian(
         unsigned int num_particles,
@@ -42,9 +42,10 @@ double SimpleGaussian::compute_laplacian()
     return laplacian;
 }
 
-double SimpleGaussian::compute_laplacian_alpha_derivative() 
+std::valarray<double> SimpleGaussian::compute_laplacian_variational_gradient()
 {
-    double alpha, position_squared_sum, laplacian_alpha_derivative;
+    double alpha, position_squared_sum;
+    std::valarray<double> laplacian_alpha_derivative(m_num_parameters);
 
     /* Fetch variational parameter alpha */
     alpha = m_parameters[0];
@@ -53,7 +54,8 @@ double SimpleGaussian::compute_laplacian_alpha_derivative()
     position_squared_sum = compute_position_squared_sum();
 
     /* Compute the alpha derivative of the laplacian */
-    laplacian_alpha_derivative = (double) -2 * m_num_dimensions * m_num_particles
+    laplacian_alpha_derivative[0] =
+        (double) -2 * m_num_dimensions * m_num_particles
         + 8 * alpha * position_squared_sum;
 
     return laplacian_alpha_derivative;
@@ -61,7 +63,7 @@ double SimpleGaussian::compute_laplacian_alpha_derivative()
 
 double SimpleGaussian::compute_drift_force_component(double coordinate)
 {
-    double alpha; 
+    double alpha;
 
     /* Fetch variational parameter alpha */
     alpha = m_parameters[0];

### 2018-03-03 project_1/vmc_hybrid/vmc/src/wavefunctions/simple_gaussian.h

@@ -1,5 +1,7 @@
 #pragma once
 
+#include <valarray>
+
 #include "wavefunction.h"
 
 class SimpleGaussian : public Wavefunction
@@ -20,5 +22,5 @@ class SimpleGaussian : public Wavefunction
         double evaluate();
         double compute_laplacian();
         double compute_drift_force_component(double coordinate);
-        double compute_laplacian_alpha_derivative();
+        std::valarray<double> compute_laplacian_variational_gradient();
 };

### 2018-03-03 project_1/vmc_hybrid/vmc/src/wavefunctions/wavefunction.h

@@ -1,5 +1,7 @@
 #pragma once
 
+#include <valarray>
+
 class Wavefunction
 {
     protected:
@@ -19,10 +21,10 @@ class Wavefunction
          *
          *      m_particles[j + i*m_num_dimensions] = m_particles[i][j],
          *
-         * where i is the particle index and j is the position index. 
-         * 
+         * where i is the particle index and j is the position index.
+         *
          *      Nice to know.
-         *          -Seb. 
+         *          -Seb.
          */
         double *m_particles;
 
@@ -40,6 +42,16 @@ class Wavefunction
         unsigned int get_num_dimensions()
         {
             return m_num_dimensions;
+        }
+
+        double *get_parameters()
+        {
+            return m_parameters;
+        }
+
+        unsigned int get_num_parameters()
+        {
+            return m_num_parameters;
         }
 
         void reset_particle_position(
@@ -66,11 +78,9 @@ class Wavefunction
                 double step, unsigned int particle_index,
                 unsigned int coordinate);
 
-        double *get_parameters() {return m_parameters;};
-        unsigned int get_num_parameters() {return m_num_parameters;};
-
         virtual double evaluate() = 0;
         virtual double compute_laplacian() = 0;
         virtual double compute_drift_force_component(double coordinate) = 0;
-        virtual double compute_laplacian_alpha_derivative() = 0;
+        virtual std::valarray<double> compute_laplacian_variational_gradient()
+            = 0;
 };

### 2018-03-06 project_1/article/article.tex
Add missing boy
Add references
	The placement of the citations is completely arbitrary.
Add theory on scaling the system

@@
     use it to evaluate the ground state energy of a trapped, hard sphere Bose
-    gas.
+    gas. \cite{dubois2001bose} \cite{nilsen2005vortices}
 
@@
 
+    \subsection{Scaling the system}
+        We now introduce a scaled distance $\vf{r}' = \vf{r}/a_{\text{ho}}$
+        \cite{dubois2001bose}, where
+        \begin{align}
+            a_{\text{ho}}
+            &=
+            \sqrt{\frac{\hbar}{m\omega}}.
+        \end{align}
+        We can then rewrite the Hamiltonian for the elliptic potential in terms
+        of this new distance. By doing a variable substitution for each
+        direction in the Laplace operator we get
+        \begin{align}
+            \nabla_k^2
+            &=
+            \frac{1}{a_{\text{ho}}^2}{\nabla_k'}^2.
+        \end{align}
+
+        Looking at the one-body part of the Hamiltonian we get
+        \begin{align}
+            H
+            &=
+            \sum_{k = 1}^N
+            \brac{
+                -\frac{\hbar^2}{2m}\nabla_k^2
+                + \half m\brak{
+                    \omega^2(x_k^2 + y_k^2)
+                    + \omega_z^2 z_k^2
+                }
+            }
+            \\
+            &=
+            \frac{\hbar\omega}{2}
+            \sum_{k = 1}^N
+            \brac{
+                -{\nabla_k'}^2
+                +
+                \brak{
+                    {x_k'}^2 + {y_k'}^2 + \lambda^2{z_k'}^2
+                }
+            },
+        \end{align}
+        where we have introduced the dimensionless frequency $\lambda =
+        \omega_z/\omega$. The single particle functions also gets a scaling
+        factor
+        \begin{align}
+            \phi_k
+            &=
+            \exp\brak{
+                -\frac{\alpha\hbar}{m\omega}
+                ({x_k'}^2 + {y_k'}^2 + \beta^2{z_k'}^2)
+            }
+            \equiv
+            \phi(\vf{r}_k').
+        \end{align}
+        The correlation wavefunction and the interaction potential remain the
+        same, but using the scaled hard core sphere radius $a/a_{\text{ho}} =
+        0.0043$. For brevity we will remove the ticks on the variables.
+
 \section{Algorithms}
@@
         \begin{equation}
-            \tau = 1 + 2 \sum_{d=1}^{n-1},
+            \tau = 1 + 2 \sum_{d=1}^{n-1}\kappa_d,
         \end{equation}
@@
 
+\bibliography{references}
 \end{document}

### 2018-03-08 project_1/vmc_hybrid/vmc/__init__.py
Add direct import of the runner module

@@ -1,0 +1,1 @@
+from vmc.runner import run_experiment

### 2018-03-08 project_1/vmc_hybrid/setup.py
Install vmc-package

@@ -51,5 +51,6 @@ extensions = [
 setup(
     name="Bosonic vmc",
     version="0.0.1",
+    packages=["vmc"],
     ext_modules=cythonize(extensions)
 )

### 2018-03-08 project_1/article/article.tex

@@
 
+        INSERT PSEUDO-CODE DESCRIPTION OF
+
+\section{Results}
+
+	\subsection{Non-Interacting Systems}
+	Initially we are interested in seeing how our implementation performs and to what extent it functions properly.
+	the logical thing to do would be to look at a non-interacting system with a simple Gaussian wavefunction and
+	harmonic oscillator potential. Such a system has a simple closed-form analytic solution\footnote{A fact that every 
+	undergraduate physics student should be well aware of.}.
+
 \bibliography{references}

### 2018-03-08 project_1/article/article.tex
Continued writing

@@
         \sum_{i = 1}^N r_i^2.
+        \label{eq:closed_form_natural_units_local_energy}
     \end{align}
@@
 	harmonic oscillator potential. Such a system has a simple closed-form analytic solution\footnote{A fact that every 
-	undergraduate physics student should be well aware of.}.
+	undergraduate physics student should be well aware of.}, which is further simplified by the "god-given" analytic
+	units (\autoref{eq:closed_form_natural_units_local_energy}). By insertion of the analytically optimal
+
+	We will start without importance sampling, that is by employing the simpler and more naïve "brute force" Monte Carlo
+	sampling. Moreover, we will
 

### 2018-03-09 project_1/article/article.tex
More results on problem b.

@@
 \usepackage{tikz}
+% Several figures in the same figure
+\usepackage{subfig}
 
@@
 
-        INSERT PSEUDO-CODE DESCRIPTION OF
+        INSERT PSEUDO-CODE DESCRIPTION OF ALGORITHM?
 
@@
 	undergraduate physics student should be well aware of.}, which is further simplified by the "god-given" analytic
-	units (\autoref{eq:closed_form_natural_units_local_energy}). By insertion of the analytically optimal
+	units (\autoref{eq:closed_form_natural_units_local_energy}). This expression is minimised for 
+	$\alpha = \frac{1}{2}\omega$. We also set $\omega = 1$ and get an expression for the 
+	expected minimum local energy,
+	\begin{equation}
+		E_L(\vb{r}) = \frac{dN}{2},
+	\end{equation}
+	which will be used as the benchmark we are ultimately aiming for.
 
 	We will start without importance sampling, that is by employing the simpler and more naïve "brute force" Monte Carlo
-	sampling. Moreover, we will
+	sampling. Moreover, we will compute the value for the local energy analytically as well as with a numerical approach. 
+	In the numerical approach.
+
+	\begin{table}
+		\caption{Analytic result. $N=1$, 1D, $2\times10^6$ cycles.}
+		\centering
+		\begin{tabular}{ccc} \hline
+			\alpha & E($\alpha$) & $\sigma^2$(E) \\ \hline \hline
+				0.30 & 0.5717 & 0.1504 \\
+				0.34 & 0.5397 & 0.0815 \\
+				0.38 & 0.5197 & 0.0390 \\
+				0.42 & 0.5084 & 0.0154 \\
+				0.46 & 0.5016 & 0.0034 \\
+				0.50 & 0.5000 & 0.0000 \\
+				0.54 & 0.5016 & 0.0030 \\
+				0.58 & 0.5047 & 0.0117 \\
+				0.62 & 0.5114 & 0.0239 \\
+				0.66 & 0.5203 & 0.0381 \\
+				0.70 & 0.5275 & 0.0590 \\ \hline \hline			
+		\end{tabular}	
+	\end{table}
+
+	\begin{table}
+		\caption{Numerical result. $N=1$, 1D, $2\times10^6$ cycles.}
+		\centering
+		\begin{tabular}{ccc} \hline
+			\alpha & E($\alpha$) & $\sigma^2$(E) \\ \hline \hline
+				0.30 & 0.5807 & 0.2535 \\
+				0.34 & 0.5379 & 0.1280 \\
+				0.38 & 0.5365 & 0.1120 \\
+				0.42 & 0.5220 & 0.0671 \\
+				0.46 & 0.5210 & 0.05250 \\
+				0.50 & 0.5265 & 0.0411\\
+				0.54 & 0.5264 & 0.0263 \\
+				0.58 & 0.5353 & 0.0259 \\
+				0.62 & 0.5481 & 0.0221 \\
+				0.66 & 0.5481 & 0.0194 \\
+				0.70 & 0.5596 & 0.0204 \\ \hline \hline
+		\end{tabular}
+	\end{table}
+
+	\begin{table*}[ht]
+		\centering
+		\caption{Brute force Metropolis-Hasting at optimum value $\alpha = \frac{1}{2}$ in three dimensions. Step length = 0.1.}
+		\begin{tabular}{c|ccc|ccc} 
+			  & \multicolumn{3}{c|}{Analytic} & \multicolumn{3}{c}{Numeric} \\ \hline
+			No. of particles & E($\alpha$) & $\sigma^2$(E) & CPU time [s] & E($\alpha$) & $\sigma^2$(E) & CPU time [s] \\ \hline
+			1					 	&   1.5000		  & 0.0000			&  0.42562		& 1.29538		 & 0.10489		   & 0.69979  \\
+			10						& 15.0000		  & 0.0000			&  0.78426 		& 7.70527		 & 3.91873		   & 7.06599	 \\
+			100					&  150.0000	  & 0.0000			&  4.34369		& 74.22956 	 & 29.77369		   & 481.69100\\
+			500					&	..				  & 0.0000			&  .. 					& ..	    			 & ..					   & \\ \hline
+		\end{tabular}
+	\end{table*}
 

### 2018-03-09 project_1/vmc_hybrid/scripts/problem_b.py

@@ -33,12 +33,13 @@ spread      = 0.1 # Unused !?!?!
 step_length = 0.1 
 
 # Parameter space, variational parameter alpha (include alpha = 0.5)
-num_alphas = 11
-alpha = np.linspace(0.3, 0.7, num_alphas).reshape(num_alphas, 1)
+num_alphas = 1
+#alpha = np.linspace(0.3, 0.7, num_alphas).reshape(num_alphas, 1)
+alpha = np.array([0.5]).reshape(1,1)
 
 # Particle-, dimension- and parameter configuration
-num_particles  = [1, 10, 100]
-num_dimensions = [1, 2, 3]
+num_particles  = [1, 10, 100, 500]
+num_dimensions = [3]
 num_parameters = 1
 
 # Storage
@@ -101,6 +102,6 @@ for i in range(len(num_particles)):
             analytic_wfn.redistribute()
             numerical_wfn.redistribute()
 
-            print("Alpha = {:5.3f}, Analytic E = {:9.5f}, CPU Time: {:8.5f}, Numerical E = {:9.5f}, CPU Time: {:8.5f}"
-                    .format(alpha.ravel()[k], energies_analytic[i, j, k], cpu_time_analytic[i, j, k], energies_numerical[i, j, k], cpu_time_numerical[i, j, k]))
+            print("Alpha = {:5.3f}, Ana E = {:9.5f}, var = {:9.5f}, CPU Time: {:8.5f}, Num E = {:9.5f}, var = {:9.5f}, CPU Time: {:8.5f}"
+                    .format(alpha.ravel()[k], energies_analytic[i, j, k], variance_analytic[i, j, k], cpu_time_analytic[i, j, k], energies_numerical[i, j, k], variance_numerical[i, j, k], cpu_time_numerical[i, j, k]))
 

### 2018-03-11 project_1/article/article.tex

@@
 			10						&  15.0000	  & 0.0000			&  0.78426 		& 7.70527		 & 3.91873		   & 7.06599	 \\
-			100					&  150.0000	  & 0.0000			&  4.34369		& 74.22956 	 & 29.77369		   & 481.69100\\
-			500					&	..				  & 0.0000			&  .. 					& ..	    			 & ..					   & \\ \hline
+			100					&  150.0000	  & 0.0000			&  4.34369		& 74.22956 	 & 29.77369		   & 481.69100 \\
+			500					&	750.0000	  & 0.0000			&  21.32269		& 362.01744	 & 51.52385         & 28409.41744\\ \hline
 		\end{tabular}

### 2018-03-15 project_1/vmc_hybrid/scripts/problem_b.py
Finally added the appendix figure for problem b

@@ -27,19 +27,18 @@ omega = 1.0
 hbar  = 1.0
 
 # Initial distribution spread
-spread      = 0.1 # Unused !?!?!
 
 # Pretermined step particles (walkers) can move for each iteration
-step_length = 0.1 
+step_length = 1.0 
 
 # Parameter space, variational parameter alpha (include alpha = 0.5)
-num_alphas = 1
+num_alphas = 11
 #alpha = np.linspace(0.3, 0.7, num_alphas).reshape(num_alphas, 1)
-alpha = np.array([0.5]).reshape(1,1)
+alpha = np.linspace(0.3, 0.7, num_alphas).reshape(num_alphas, 1)
 
 # Particle-, dimension- and parameter configuration
-num_particles  = [1, 10, 100, 500]
-num_dimensions = [3]
+num_particles  = [1, 10, 100]
+num_dimensions = [1, 2, 3]
 num_parameters = 1
 
 # Storage

### 2018-03-20 project_1/article/article.tex
Finally added the appendix figure for problem b
Add plots and tables for problem b

@@
 \usepackage{dsfont}
-% Include package for drawing Feynman diagrams
-\usepackage{tikz-feynman}
-% Include package for typesetting contractions
-\usepackage{simpler-wick}
-% Include the Feynman-slash
-\usepackage{slashed}
 % Include Tikz
@@
 \usepackage{subfig}
+% Appendix
+\usepackage[toc, page]{appendix}
 
@@
     \begin{table}
-		\caption{Analytic result. $N=1$, 1D, $2\times10^6$ cycles.}
         \centering
-		\begin{tabular}{ccc} \hline
-			\alpha & E($\alpha$) & $\sigma^2$(E) \\ \hline \hline
-				0.30 & 0.5717 & 0.1504 \\
-				0.34 & 0.5397 & 0.0815 \\
-				0.38 & 0.5197 & 0.0390 \\
-				0.42 & 0.5084 & 0.0154 \\
-				0.46 & 0.5016 & 0.0034 \\
-				0.50 & 0.5000 & 0.0000 \\
-				0.54 & 0.5016 & 0.0030 \\
-				0.58 & 0.5047 & 0.0117 \\
-				0.62 & 0.5114 & 0.0239 \\
-				0.66 & 0.5203 & 0.0381 \\
-				0.70 & 0.5275 & 0.0590 \\ \hline \hline			
+        \caption{One particle in one dimension for the analytic expression with
+        $2^{21}$ Monte Carlo cycles and step length of $0.5$.}
+        \begin{ruledtabular}
+            \begin{tabular}{cccccc}
+                $\alpha$ & $\langle  E_L\rangle$ & $\sigma$ & $\sigma_b$ & $A$
+                & $t_{\text{CPU}} [\si{\second}]$ \\
+                \hline
+                0.30&0.56661&0.00026&0.00133&0.89121&0.39765\\
+                0.34&0.53858&0.00019&0.00096&0.88419&0.34550\\
+                0.38&0.51822&0.00013&0.00062&0.87823&0.34849\\
+                0.42&0.50770&0.00009&0.00037&0.87136&0.36115\\
+                0.46&0.50197&0.00004&0.00018&0.86589&0.34705\\
+                0.50&0.50000&0.00000&0.00000&0.86092&0.35124\\
+                0.54&0.50146&0.00004&0.00016&0.85497&0.34875\\
+                0.58&0.50592&0.00007&0.00029&0.84999&0.34705\\
+                0.62&0.51180&0.00011&0.00041&0.84455&0.34861\\
+                0.66&0.51983&0.00014&0.00051&0.84014&0.36164\\
+                0.70&0.52742&0.00017&0.00063&0.83516&0.35525\\
             \end{tabular}
+        \end{ruledtabular}
+        \label{tab:1D1N_analytic}
     \end{table}
@@
     \begin{table}
-		\caption{Numerical result. $N=1$, 1D, $2\times10^6$ cycles.}
         \centering
-		\begin{tabular}{ccc} \hline
-			\alpha & E($\alpha$) & $\sigma^2$(E) \\ \hline \hline
-				0.30 & 0.5807 & 0.2535 \\
-				0.34 & 0.5379 & 0.1280 \\
-				0.38 & 0.5365 & 0.1120 \\
-				0.42 & 0.5220 & 0.0671 \\
-				0.46 & 0.5210 & 0.05250 \\
-				0.50 & 0.5265 & 0.0411\\
-				0.54 & 0.5264 & 0.0263 \\
-				0.58 & 0.5353 & 0.0259 \\
-				0.62 & 0.5481 & 0.0221 \\
-				0.66 & 0.5481 & 0.0194 \\
-				0.70 & 0.5596 & 0.0204 \\ \hline \hline
+        \caption{One particle in one dimension for the numeric expression with
+        $2^{21}$ Monte Carlo cycles and a step length of $0.5$.}
+        \begin{ruledtabular}
+            \begin{tabular}{cccccc}
+                $\alpha$ & $\langle  E_L\rangle$ & $\sigma$ & $\sigma_b$
+                & $A$&$t_{\text{CPU}} [\si{\second}]$ \\
+                \hline
+                0.30&0.56494&0.00026&0.00131&0.89160&0.35031\\
+                0.34&0.53933&0.00020&0.00097&0.88427&0.37179\\
+                0.38&0.51841&0.00014&0.00063&0.87840&0.34806\\
+                0.42&0.50737&0.00008&0.00037&0.87182&0.34957\\
+                0.46&0.50211&0.00004&0.00018&0.86549&0.34976\\
+                0.50&0.50000&0.00000&0.00000&0.86041&0.34791\\
+                0.54&0.50136&0.00004&0.00015&0.85532&0.34858\\
+                0.58&0.50525&0.00007&0.00028&0.84945&0.34703\\
+                0.62&0.51181&0.00011&0.00040&0.84535&0.34517\\
+                0.66&0.52002&0.00014&0.00051&0.84053&0.34858\\
+                0.70&0.52778&0.00017&0.00061&0.83560&0.34818\\
             \end{tabular}
+        \end{ruledtabular}
+        \label{tab:1D1N_numeric}
+    \end{table}
+
+    \begin{table}
+        \centering
+        \caption{BLEH}
+        \begin{ruledtabular}
+            \begin{tabular}{c|ccc|ccc}
+                & \multicolumn{3}{c|}{Analytic}
+                & \multicolumn{3}{c}{Numeric} \\
+                \hline
+                $N$
+                & $\langle E_L\rangle$ & $\sigma_b$
+                & $t_{\text{CPU}} [\si{\second}]$
+                & $\langle E_L\rangle$ & $\sigma_b$
+                & $t_{\text{CPU}} [\si{\second}]$ \\
+                \hline
+                1  \\
+                10 \\
+                100 \\
+                500
+            \end{tabular}
+        \end{ruledtabular}
+        \label{tab:initial_optimum_comparison}
     \end{table}
@@
 
+\appendix
+\section{Brute Force Metropolis-Hastings}
+
+    \begin{figure*}
+        \makebox[\textwidth][c]{
+            \includegraphics[width=1.2\textwidth]{../data/figures/problem_b.pdf}
+        }
+        \caption{Comparison of analytic and numeric results with the exact
+        expression for the energy as a function of $\alpha$. The plots show the
+        standard deviation from the blocking method as error ticks. In the
+        figure $N$ is the number of particles and $d$ the number of dimensions.}
+        \label{fig:initial_problem_b}
+    \end{figure*}
+
 \bibliography{references}
+
 \end{document}

### 2018-03-21 project_1/article/article.tex
Add correct energies
Add run time for large system

@@
                 \hline
-                0.30&0.56661&0.00026&0.00133&0.89121&0.39765\\
-                0.34&0.53858&0.00019&0.00096&0.88419&0.34550\\
-                0.38&0.51822&0.00013&0.00062&0.87823&0.34849\\
-                0.42&0.50770&0.00009&0.00037&0.87136&0.36115\\
-                0.46&0.50197&0.00004&0.00018&0.86589&0.34705\\
-                0.50&0.50000&0.00000&0.00000&0.86092&0.35124\\
-                0.54&0.50146&0.00004&0.00016&0.85497&0.34875\\
-                0.58&0.50592&0.00007&0.00029&0.84999&0.34705\\
-                0.62&0.51180&0.00011&0.00041&0.84455&0.34861\\
-                0.66&0.51983&0.00014&0.00051&0.84014&0.36164\\
-                0.70&0.52742&0.00017&0.00063&0.83516&0.35525\\
+                0.30&0.56244&0.00026&0.00129&0.89185&0.43238\\
+                0.34&0.53652&0.00019&0.00094&0.88465&0.38340\\
+                0.38&0.51863&0.00013&0.00063&0.87844&0.37200\\
+                0.42&0.50805&0.00009&0.00038&0.87161&0.37978\\
+                0.46&0.50201&0.00004&0.00017&0.86594&0.37214\\
+                0.50&0.50000&0.00000&0.00000&0.86047&0.37162\\
+                0.54&0.50183&0.00004&0.00015&0.85543&0.37959\\
+                0.58&0.50566&0.00007&0.00029&0.84994&0.37778\\
+                0.62&0.51175&0.00011&0.00041&0.84479&0.37368\\
+                0.66&0.51906&0.00014&0.00052&0.84000&0.37257\\
+                0.70&0.52883&0.00017&0.00062&0.83570&0.37342\\
             \end{tabular}
@@
                 \hline
-                0.30&0.56494&0.00026&0.00131&0.89160&0.35031\\
-                0.34&0.53933&0.00020&0.00097&0.88427&0.37179\\
-                0.38&0.51841&0.00014&0.00063&0.87840&0.34806\\
-                0.42&0.50737&0.00008&0.00037&0.87182&0.34957\\
-                0.46&0.50211&0.00004&0.00018&0.86549&0.34976\\
-                0.50&0.50000&0.00000&0.00000&0.86041&0.34791\\
-                0.54&0.50136&0.00004&0.00015&0.85532&0.34858\\
-                0.58&0.50525&0.00007&0.00028&0.84945&0.34703\\
-                0.62&0.51181&0.00011&0.00040&0.84535&0.34517\\
-                0.66&0.52002&0.00014&0.00051&0.84053&0.34858\\
-                0.70&0.52778&0.00017&0.00061&0.83560&0.34818\\
+                0.30000&0.56675&0.00026&0.00135&0.89098&0.52634\\
+                0.34000&0.53714&0.00019&0.00096&0.88479&0.52186\\
+                0.38000&0.51836&0.00013&0.00062&0.87820&0.52351\\
+                0.42000&0.50717&0.00009&0.00038&0.87180&0.52892\\
+                0.46000&0.50184&0.00004&0.00017&0.86581&0.51759\\
+                0.50000&0.50000&0.00001&0.00001&0.86024&0.52427\\
+                0.54000&0.50133&0.00004&0.00015&0.85520&0.52052\\
+                0.58000&0.50520&0.00007&0.00028&0.84971&0.52156\\
+                0.62000&0.51120&0.00011&0.00039&0.84481&0.51830\\
+                0.66000&0.51896&0.00014&0.00052&0.83990&0.51865\\
+                0.70000&0.52794&0.00017&0.00061&0.83506&0.51255\\
             \end{tabular}
@@
     \begin{table}
+        \caption{Comparison of energy, standard deviation and the CPU time for
+        $\alpha = 0.5$ (the correct minimum) for the analytic and numeric
+        schemes in three dimensions and $N$ particles. The step length is
+        $0.5$.}
         \centering
-        \caption{BLEH}
         \begin{ruledtabular}
-            \begin{tabular}{c|ccc|ccc}
+            \begin{tabular}{r|rrr|rrr}
                 & \multicolumn{3}{c|}{Analytic}
@@
                 \hline
-                1  \\
-                10 \\
-                100 \\
-                500
+                1 & 1.5 & 0.0 & 0.50
+                & 1.49997 & 0.00008 & 0.73 \\
+                10 & 15.0 & 0.0 & 0.73
+                & 14.98278 & 0.00867 & 6.40 \\
+                100 & 150.0 & 0.0 & 4.13
+                & 149.81453 & 0.81501 & 387.60 \\
+                500 & 750.0 & 0.0 & 19.16
+                & 773.85435 & 20.69789 & 9595.02
             \end{tabular}
@@
     \end{table}
-
-	\begin{table*}[ht]
-		\centering
-		\caption{Brute force Metropolis-Hasting at optimum value $\alpha = \frac{1}{2}$ in three dimensions. Step length = 0.1.}
-		\begin{tabular}{c|ccc|ccc}
-			  & \multicolumn{3}{c|}{Analytic} & \multicolumn{3}{c}{Numeric} \\ \hline
-			No. of particles & E($\alpha$) & $\sigma^2$(E) & CPU time [s] & E($\alpha$) & $\sigma^2$(E) & CPU time [s] \\ \hline
-			1					 	&   1.5000		  & 0.0000			&  0.42562		& 1.29538		 & 0.10489		   & 0.69979  \\
-			10						&  15.0000	  & 0.0000			&  0.78426 		& 7.70527		 & 3.91873		   & 7.06599	 \\
-			100					&  150.0000	  & 0.0000			&  4.34369		& 74.22956 	 & 29.77369		   & 481.69100 \\
-			500					&	750.0000	  & 0.0000			&  21.32269		& 362.01744	 & 51.52385         & 28409.41744\\ \hline
-		\end{tabular}
-	\end{table*}
 

### 2018-03-21 project_1/vmc_hybrid/vmc/src/hamiltonians/hamiltonian.cpp
Remove repulsive interaction in Hamiltonian
        This will never be run as we will never accept a step where we
        have overlap.

@@ -14,34 +14,3 @@ double Hamiltonian::compute_local_energy(Wavefunction *wavefunction)
 
     return kinetic_energy + potential_energy;
 }
-
-double Hamiltonian::compute_interaction(Wavefunction *wavefunction)
-{
-    double a, distance;
-    const double TOL = 1e-10;
-    unsigned int i, j;
-
-    a = wavefunction->get_hard_sphere_radius();
-
-    /* Check if there is a hard sphere */
-    if (fabs(a - 0.0) < TOL) {
-        return 0.0;
-    }
-
-    /* Check all particle combinations */
-    for (i = 0; i < wavefunction->get_num_particles(); i++) {
-        for (j = i + 1; j < wavefunction->get_num_particles(); j++) {
-            distance = wavefunction->get_distance_between_particles(i, j);
-
-            /* Check if the particle hard spheres are overlapping */
-            if (fabs(a - distance) <= TOL) {
-                /* Abort the loop by returning infinity if any particles are
-                 * overlapping */
-                return std::numeric_limits<double>::infinity();
-            }
-        }
-    }
-
-    /* No particles are overlapping, return 0 */
-    return 0.0;
-}

### 2018-03-21 project_1/vmc_hybrid/vmc/src/hamiltonians/hamiltonian.h

@@ -11,5 +11,4 @@ class Hamiltonian
         virtual double compute_potential_energy(Wavefunction *wavefunction) = 0;
 
         double compute_local_energy(Wavefunction *wavefunction);
-        double compute_interaction(Wavefunction *wavefunction);
 };

### 2018-03-21 project_1/article/article.tex

@@
 \usepackage{tikz}
+% Tikz add-on
+\usetikzlibrary{shapes,arrows}
+\usetikzlibrary{positioning}
 % Several figures in the same figure
@@
 
-        INSERT PSEUDO-CODE DESCRIPTION OF ALGORITHM?
+        % Flow chart for description of automatic blocking algorithm
+		\begin{figure}
+			\centering
+
+        		% Defining blocks of flowchart
+        		\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
+   			 	text width=5em, text badly centered, node distance=3cm, inner sep=0pt]
+			\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
+    				text width=10em, text centered, rounded corners, minimum height=4em]
+			\tikzstyle{line} = [draw, -latex']
+
+    			% Actual figure
+    			\begin{tikzpicture}[node distance = 2.2cm, auto]
+
+    				% Nodes
+    				\node [block] (init) {Start with $\mathbf{X}_i$, $i=0$};
+    				\node [block, below of=init] (variance) {Compute variance and autocovariance of $\mathbf{X}_i$, $s_i$ and $\gamma_i$};
+    				\node [block, below of=variance] (transform) 
+    					{Transform data $(\mathbf{X}_{i+1})_k = \frac{1}{2}\left((\mathbf{X})_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right)$};
+    				\node [block, right=1.5cm of transform] (i+1) {Next iteration: $i = i + 1$.} ;
+    				\node [decision, below of = transform] (x_size) {Length of $\mathbf{X}_{i + 1} \geq 2$ ? }; 
+    				\node [block, below of=x_size] (m_test) {Compute test statistic $M_j$};
+    				\node [block, below of=m_test] (magic_numbers) {Find first $k$ so that $M_k \leq q_{d-k}(1-\alpha)$};
+    				\node [block, below of=magic_numbers] (return) {Return $s_k^2/n_k$};
+
+    				% Edges
+    				\path [line] (init) -- (variance);
+    				\path [line] (variance) -- (transform);
+    				\path [line] (transform) -- (x_size);
+    					\path [line] (x_size) -| node [near start] {yes} (i+1);
+    					\path [line] (i+1) |- (variance);
+    				\path [line] (x_size) -- node [near start] {no} (m_test);
+    				\path [line] (m_test) -- (magic_numbers);
+    				\path [line] (magic_numbers) -- (return);
+    			\end{tikzpicture}
+    			\caption{Flow chart, blocking algorithm.}
+		\end{figure}
 

### 2018-03-21 project_1/article/article.tex
Add table of contents
Add tables and figure
        The article looks insane. This needs to be fixed after adding
        all results.

@@
 \usepackage{subfig}
-% Appendix
-\usepackage[toc, page]{appendix}
+%% Appendix
+%\usepackage[toc, page]{appendix}
 
@@
 \maketitle
-
-%\begin{widetext}
-
-%\end{widetext}
+\tableofcontents
 
@@
     \subsection{Non-Interacting Systems}
-    Initially we are interested in seeing how our implementation performs and to what extent it functions properly.
-    the logical thing to do would be to look at a non-interacting system with a simple Gaussian wavefunction and
-    harmonic oscillator potential. Such a system has a simple closed-form analytic solution\footnote{A fact that every
-    undergraduate physics student should be well aware of.}, which is further simplified by the "god-given" analytic
-    units (\autoref{eq:closed_form_natural_units_local_energy}). This expression is minimised for
-    $\alpha = \frac{1}{2}\omega$. We also set $\omega = 1$ and get an expression for the
-    expected minimum local energy,
+    Initially we are interested in seeing how our implementation performs and to
+    what extent it functions properly.  the logical thing to do would be to look
+    at a non-interacting system with a simple Gaussian wavefunction and harmonic
+    oscillator potential. Such a system has a simple closed-form analytic
+    solution\footnote{A fact that every undergraduate physics student should be
+    well aware of.}, which is further simplified by the "god-given" analytic
+    units (\autoref{eq:closed_form_natural_units_local_energy}). This expression
+    is minimised for $\alpha = \frac{1}{2}\omega$. We also set $\omega = 1$ and
+    get an expression for the expected minimum local energy,
     \begin{equation}
@@
 
-    We will start without importance sampling, that is by employing the simpler and more naïve "brute force" Monte Carlo
-    sampling. Moreover, we will compute the value for the local energy analytically as well as with a numerical approach.
-    In the numerical approach.
+    We will start without importance sampling, that is by employing the simpler
+    and more naïve "brute force" Monte Carlo sampling. Moreover, we will compute
+    the value for the local energy analytically as well as with a numerical
+    approach.  In the numerical approach.
 
@@
                 \hline
-                0.30000&0.56675&0.00026&0.00135&0.89098&0.52634\\
-                0.34000&0.53714&0.00019&0.00096&0.88479&0.52186\\
-                0.38000&0.51836&0.00013&0.00062&0.87820&0.52351\\
-                0.42000&0.50717&0.00009&0.00038&0.87180&0.52892\\
-                0.46000&0.50184&0.00004&0.00017&0.86581&0.51759\\
-                0.50000&0.50000&0.00001&0.00001&0.86024&0.52427\\
-                0.54000&0.50133&0.00004&0.00015&0.85520&0.52052\\
-                0.58000&0.50520&0.00007&0.00028&0.84971&0.52156\\
-                0.62000&0.51120&0.00011&0.00039&0.84481&0.51830\\
-                0.66000&0.51896&0.00014&0.00052&0.83990&0.51865\\
-                0.70000&0.52794&0.00017&0.00061&0.83506&0.51255\\
+                0.30&0.56675&0.00026&0.00135&0.89098&0.52634\\
+                0.34&0.53714&0.00019&0.00096&0.88479&0.52186\\
+                0.38&0.51836&0.00013&0.00062&0.87820&0.52351\\
+                0.42&0.50717&0.00009&0.00038&0.87180&0.52892\\
+                0.46&0.50184&0.00004&0.00017&0.86581&0.51759\\
+                0.50&0.50000&0.00001&0.00001&0.86024&0.52427\\
+                0.54&0.50133&0.00004&0.00015&0.85520&0.52052\\
+                0.58&0.50520&0.00007&0.00028&0.84971&0.52156\\
+                0.62&0.51120&0.00011&0.00039&0.84481&0.51830\\
+                0.66&0.51896&0.00014&0.00052&0.83990&0.51865\\
+                0.70&0.52794&0.00017&0.00061&0.83506&0.51255\\
             \end{tabular}
@@
 
+    \subsection{Interacting elliptical harmonic oscillator}
+        \begin{table}
+            \caption{In this table we show the expected energy, the standard
+            deviation, the acceptance ratio and the time spent for each $alpha$
+            for the elliptical potential with the interacting wavefunction. We
+            have used importance sampling with $\delta t = 0.1$ and $2^{21}$
+            Monte Carlo cycles and an additional $10\%$ of that as
+            thermalization steps. The hard sphere radius is set to $0.0043$ and
+            a $\beta = \lambda = \sqrt{8}$.}
+            \centering
+            \begin{ruledtabular}
+                \begin{tabular}{ccccc}
+                    $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
+                    &$A$&$t_C$ \\
+                    \hline
+                    0.2&35.17548&0.05399&0.98995&24.48868\\
+                    0.3&27.62004&0.02311&0.98187&13.65132\\
+                    0.4&24.97850&0.00827&0.97248&13.54666\\
+                    0.5&24.39877&0.00030&0.96140&13.38388\\
+                    0.6&24.83863&0.00604&0.94998&13.31570\\
+                    0.7&25.82855&0.01059&0.93766&13.20579\\
+                    0.8&27.22895&0.01436&0.92375&13.11933\\
+                \end{tabular}
+            \end{ruledtabular}
+            \label{tab:10_interacting}
+        \end{table}
+
+        \begin{table}
+            \caption{BLEH}
+            \centering
+            \begin{ruledtabular}
+                \begin{tabular}{ccccc}
+                    $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
+                    &$A$&$t_C$ \\
+                    \hline
+                    0.2&181.69371&0.25533&0.98756&199.95366\\
+                    0.3&142.61591&0.10782&0.97849&195.80928\\
+                    0.4&129.88615&0.04051&0.96862&190.44553\\
+                    0.5&127.29926&0.00595&0.95749&195.10368\\
+                    0.6&129.97630&0.03300&0.94595&193.22755\\
+                    0.7&135.67382&0.05785&0.93330&191.96391\\
+                    0.8&143.23238&0.07286&0.92061&192.72399\\
+                \end{tabular}
+            \end{ruledtabular}
+            \label{tab:50_interacting}
+        \end{table}
+
+        \begin{table}
+            \caption{BLEH}
+            \centering
+            \begin{ruledtabular}
+                \begin{tabular}{ccccc}
+                    $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
+                    &$A$&$t_C$ \\
+                    \hline
+                    0.2&375.04401&0.51475&0.98527&745.21438\\
+                    0.3&296.76879&0.20214&0.97534&741.04425\\
+                    0.4&270.60830&0.08113&0.94522&723.06823\\
+                    0.5&266.37263&0.02020&0.95314&702.71000\\
+                    0.6&272.51171&0.07366&0.94112&689.12862\\
+                    0.7&285.10285&0.11381&0.92895&707.48248\\
+                    0.8&301.40609&0.15458&0.91648&707.86118\\
+                \end{tabular}
+            \end{ruledtabular}
+            \label{tab:100_interacting}
+        \end{table}
+
+        \begin{figure}
+            \includegraphics[width=250px]{../data/figures/problem_e.pdf}
+            \caption{BLEH}
+            \label{fig:interacting}
+        \end{figure}
+
 \appendix

### 2018-03-22 project_1/article/article.tex
Wrote about blocking.

--- 2018-03-21 project_1/article/article.tex
@@
 
+     	Instead of going by this "manual" method of looking at charts, we will construct a test statistic and
+     	let the computer do the necessary considerations in a more "automatic way.
+
+        The easiest way to do this is by use of the \emph{automatic blocking scheme}, nicely illustrated 
+        with a flow chart in \autoref{fig:blocking}. The input parameter for the method is an array of the 
+        local energies, but it can be any array of time series data, which is why it is herein referred to as
+        $\mathbf{X}$. Any array of sequential data, where any particular data point depends on the previous
+        data point inhibits all attributes of a time series. Our array of local energies is therefor most definitely 
+        a time series.
+
+		From $\mathbf{X}$ we compute the variance and sample covariance between the series itself and a 
+		one-point lag of the series. This covariance is usually referred to as the first order autocovariance.
+		\begin{equation}
+			\gamma_i = \frac{1}{n}
+				\left(\sum_{k=1}^n\left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)\right)
+				\left(\sum_{k=0}^{n-1}\left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)\right).
+		\end{equation}
+
+		Now we "block"! The next step is to perform a transformation of the our time series $\mathbf{x}_i$, where
+		we end up with an array that is half the length of the original time series,
+		\begin{equation}
+			(\mathbf{X}_i)_k = \frac{1}{2}\left((\mathbf{X}_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right).
+		\end{equation}
+		This process is were the term "blocking" stems from. There are a number of ways to perform the blocking -
+		we could have picked the data points used in the new ($i+1$) array randomly, or more orderly. Here we we have
+		picked them sequentially. We have obtained $\frac{n}{2}$ new "primed" stochastic variables. Assuming that the
+		reader is playing close attention, it is now needless to say that it is much easier to perform these sequential
+		blocking steps if we have an array of length $n= 2^d$, where $d$ is integer. Otherwise, problems would arise
+		because one would have arrays of different lengths when the original array is "split in two". The ailment is
+		easy to remedy by excluding an observation from the longest resulting array after splitting.
+
+		For each new time series $\mathbf{X}_{i+1}$ a new sample variance and sample covariance is computed.
+		The blocking transformation is allowed to continue until we are we run out of data, i.e. the length
+		of $\mathbf{X}_i < 2$. In other words, there is only a single number left. We now compute a test statistic,
+		\begin{equation}
+			M_j = \sum_{k=j}^{d-1} n_k \frac{\left[(n_k - 1)\frac{s^2_k}{n_k^2} y \gamma_k(1)\right] ^2}{s_k^4}.
+		\end{equation}
+		As luck would have it, this number $M$ is   $\chi^2$-distributed. The following step would be to look up a 
+		statistical table and find the first $k$, such that $M_k \leq q_{d-k}(1-\alpha)$, where $q$ is the statistical
+		limit and $\alpha$ is  the significance level that your hear desires. Finally, the correct sample standard deviation 
+		is $\frac{s_k^2}{n_k}$. Again, see \autoref{fig:blocking} for a summary of the algorithm.
+
         % Flow chart for description of automatic blocking algorithm
@@
 			\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
-    				text width=10em, text centered, rounded corners, minimum height=4em]
+    				text width=11em, text centered, rounded corners, minimum height=4em]
 			\tikzstyle{line} = [draw, -latex']
@@
     				\node [decision, below of = transform] (x_size) {Length of $\mathbf{X}_{i + 1} \geq 2$ ? }; 
-    				\node [block, below of=x_size] (m_test) {Compute test statistic $M_j$};
+    				\node [block, below of=x_size] (m_test) {Compute test statistic $M_j$ from $s_j$ and $\gamma_j$ for all $j$.};
     				\node [block, below of=m_test] (magic_numbers) {Find first $k$ so that $M_k \leq q_{d-k}(1-\alpha)$};
@@
     			\end{tikzpicture}
-    			\caption{Flow chart, blocking algorithm.}
+    			\caption{Flow chart depicting the automatic blocking method for finding a better estimate for sample variance.}
+    			\label{fig:blocking}
 		\end{figure}

### 2018-03-23 project_1/article/article.tex
Added results that were accidentally removed in a previous commit.

@@
 \maketitle
+\tableofcontents
 
@@
     sampling. Moreover, we will compute the value for the local energy analytically as well as with a numerical approach.
-    In the numerical approach.
+    In the numerical approach we employ the well-know approximation for the second derivative,
+    \begin{equation}
+    		f''(x= \approx \frac{f(x-h) - 2f(x) + f(x+h)}{h^2}.
+    \end{equation}
 
@@
 
+\subsection{Interacting elliptical harmonic oscillator}
+         \begin{table}
+             \caption{The expected energy, the standard
+             deviation, the acceptance ratio and CPU time for each $alpha$
+             for the elliptical potential with the interacting wavefunction. We
+             have used importance sampling with $\delta t = 0.1$ and $2^{21}$
+             Monte Carlo cycles and an additional $10\%$ of that as
+             thermalization steps. The hard sphere radius is set to $0.0043$ and
+             a $\beta = \lambda = \sqrt{8}$.}
+             \centering
+             \begin{ruledtabular}
+                 \begin{tabular}{ccccc}
+                     $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
+                     &$A$&$t_C$ \\
+                     \hline
+                     0.2&35.17548&0.05399&0.98995&24.48868\\
+                     0.3&27.62004&0.02311&0.98187&13.65132\\
+                     0.4&24.97850&0.00827&0.97248&13.54666\\
+                     0.5&24.39877&0.00030&0.96140&13.38388\\
+                     0.6&24.83863&0.00604&0.94998&13.31570\\
+                     0.7&25.82855&0.01059&0.93766&13.20579\\
+                     0.8&27.22895&0.01436&0.92375&13.11933\\
+                 \end{tabular}
+             \end{ruledtabular}
+             \label{tab:10_interacting}
+         \end{table}
+
+         \begin{table}
+             \caption{BLEH}
+             \centering
+             \begin{ruledtabular}
+                 \begin{tabular}{ccccc}
+                     $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
+                     &$A$&$t_C$ \\
+                     \hline
+                     0.2&181.69371&0.25533&0.98756&199.95366\\
+                     0.3&142.61591&0.10782&0.97849&195.80928\\
+                     0.4&129.88615&0.04051&0.96862&190.44553\\
+                     0.5&127.29926&0.00595&0.95749&195.10368\\
+                     0.6&129.97630&0.03300&0.94595&193.22755\\
+                     0.7&135.67382&0.05785&0.93330&191.96391\\
+                     0.8&143.23238&0.07286&0.92061&192.72399\\
+                 \end{tabular}
+             \end{ruledtabular}
+             \label{tab:50_interacting}
+         \end{table}
+
+         \begin{table}
+             \caption{BLEH}
+             \centering
+             \begin{ruledtabular}
+                 \begin{tabular}{ccccc}
+                     $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
+                     &$A$&$t_C$ \\
+                     \hline
+                     0.2&375.04401&0.51475&0.98527&745.21438\\
+                     0.3&296.76879&0.20214&0.97534&741.04425\\
+                     0.4&270.60830&0.08113&0.94522&723.06823\\
+                     0.5&266.37263&0.02020&0.95314&702.71000\\
+                     0.6&272.51171&0.07366&0.94112&689.12862\\
+                     0.7&285.10285&0.11381&0.92895&707.48248\\
+                     0.8&301.40609&0.15458&0.91648&707.86118\\
+                 \end{tabular}
+             \end{ruledtabular}
+             \label{tab:100_interacting}
+         \end{table}
+
+         \begin{figure}
+             \includegraphics[width=250px]{../data/figures/problem_e.pdf}
+             \caption{BLEH}
+             \label{fig:interacting}
+         \end{figure}
+
 \appendix

### 2018-03-23 project_1/article/article.tex
Initial comments on initial results.

@@
 
+		\subsection{Gradient Descent}
+
+		The method of
+
 \section{Results}
@@
     harmonic oscillator potential. Such a system has a simple closed-form analytic solution\footnote{A fact that every
-    undergraduate physics student should be well aware of.}, which is further simplified by the "god-given" analytic
+    undergraduate physics student should be well aware of after finishing their first course in quantum physics}, 
+    which is further simplified by the "god-given" analytic
     units (\autoref{eq:closed_form_natural_units_local_energy}). This expression is minimised for
@@
     \end{table}
+
+	For the first
+
+	\autoref{tab:1D1N_analytic} shows the expected local energy $\expval{E_L}$, standard deviations 
+	$\sigma$, $\sigma_b$, the share of accepted Monte Carlo reorientations $A$ and cpu time for 
+	the analytic computation at different values for the variational parameter $\alpha$.
+	\autoref{tab:1D1N_numeric} shows the same numbers but for the numeric scheme. The first column of
+	standard deviations $\sigma$ are the "naïve" standard deviations, while $\sigma_b$ marks the
+	standard deviations computed with the blocking method. One can clearly see, in both the analytic and numeric
+	case, that the former standard deviations are lower than the latter, meaning that autocorrelated data
+	can lead to a lower perceived uncertainty than what is reality. The blocking method takes this fact into account 
+	and for this reason, all other standard deviations will be computed with the blocking method from here on.
 

### 2018-03-23 project_1/article/article.tex
Wrote gradient descent theory

@@
 
-		The method of
+		Scanning the entire space of variational parameters can be a somewhat despairing task. To aid us in the search we
+		therefore implement the method of gradient descent\footnote{Also known as steepest descent, easily confused
+		with the steepest descent method for integation.}. In any optimum we generally have that the derivative should equal
+		zero. In our case, we want to find the minimum of the expected local energy with respect to some variational
+		parameter $\alpha$, so that
+		\begin{equation}
+			\nabla_\alpha \expval{E_L(\mathbf{R}; \alpha)} = 0.
+		\end{equation}	
+		The foundation for  the method of gradient descent lies in the fact that some analytic
+		\footnote{We use the term analytic quite recklessly.} function $F(\alpha)$ decreases fastest in the direction
+		of $-\nabla_\alpha F(\alpha)$. This means that we could eventually get to an ensuing minimum by trailing along the
+		pathway of the following recursive relation,
+		\begin{equation}
+			\alpha_{k+1} = \alpha_k - \gamma \nabla_{alpha k} F(\alpha_k).
+		\end{equation}
+		If $\gamma$ is small enough we should have $f(\alpha_k) \geq F(\alpha_{k+1})$ for all $k$ and hopefully we will
+		experience convergence to the desired minimum.
+
+		For our special case : INSERT DERIVATION OF LOCAL ENERGY??
 
@@
     \begin{equation}
-    		f''(x= \approx \frac{f(x-h) - 2f(x) + f(x+h)}{h^2}.
+    		f''(x)= \approx \frac{f(x-h) - 2f(x) + f(x+h)}{h^2}.
     \end{equation}

### 2018-03-23 project_1/article/article.tex
Filled in initial non-interacting system text

@@
     \begin{align}
-        h(\vf{r}_i) = -\frac{\hbar^2}{2m}\nabla_i^2
+        h(\vf{r}_i) = -\frac{\hbar^2}{2m}_i^2
         + V_{\text{ext}}(\vf{r}_i),
@@
 		\begin{equation}
-			\alpha_{k+1} = \alpha_k - \gamma \nabla_{alpha k} F(\alpha_k).
+			\alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha k} F(\alpha_k).
 		\end{equation}
@@
 
-		For our special case : INSERT DERIVATION OF LOCAL ENERGY??
+		For our special case : INSERT DERIVATION OF LOCAL ENERGY derivative wrt alpha??
 
@@
     \begin{equation}
-    		f''(x)= \approx \frac{f(x-h) - 2f(x) + f(x+h)}{h^2}.
+    		f''(x) \approx \frac{f(x-h) - 2f(x) + f(x+h)}{h^2}.
     \end{equation}
@@
 
-	For the first
-
 	\autoref{tab:1D1N_analytic} shows the expected local energy $\expval{E_L}$, standard deviations 
@@
 	and for this reason, all other standard deviations will be computed with the blocking method from here on.
+
+	The first initial observation of note is that we have found an optimum for the system in terms of a minimum
+	energy at $\alpha=\frac{1}{2}$. Moreover, notice how the standard deviation disappears completely.
+	This should already be apparent from \autoref{eq:closed_form_natural_units_local_energy}, as the term
+	involving particle positions disappears for the optimal $\alpha$.
 
@@
 
+	Systems of optimum variational parameter warrants further investigation, as summed up in 
+	\autoref{tab:initial_optimum_comparison}. The analytical results are quite predictable and 
+	monotonous; the expected local energy $\expval{E_l}$ is perfectly proportional by a factor
+	$\frac{3}{2}$ to the number of particles and the standard deviation $\sigma_b$ is zero for
+	systems of any number of particles. The numerical results tells a more tortuous tale. As the 
+	number of particles increase, the uncertainty also increases rapidly. For the largest system,
+	with $N = 500$ particles, the energy misses its target critically - the true energy is more than
+	one standard deviation away. Lastly, the cpu time is notably higher for the numeric scheme.
+
+	\subsubsection{Introducing Importance Sampling}
+
+	Next we will introduce importance sampling..
+
 \subsection{Interacting elliptical harmonic oscillator}
@@
          \begin{table}
-             \caption{BLEH}
+             \caption{$50$ interacting particles.}
              \centering
@@
          \begin{table}
-             \caption{BLEH}
+             \caption{$100$ Interacting particles.}
              \centering
@@
              \includegraphics[width=250px]{../data/figures/problem_e.pdf}
-             \caption{BLEH}
+             \caption{Local energies as a function of variational parameter $\alpha$
+             				 for different number of particles.}
              \label{fig:interacting}

### 2018-03-23 project_1/article/article.tex
Even more on initial results.

@@
 
+	In the appendices,  \autoref{fig:initial_problem_b} shows how both the analytic and numeric approach
+	matches the exact expression for the energy as a function of $\alpha$ for systems of increasing size. 
+	We see also here that the uncertainty increases in the numeric scheme as the system gets larger.
+
 	\subsubsection{Introducing Importance Sampling}
 
-	Next we will introduce importance sampling..
+	Next we will introduce importance sampling.
 

### 2018-03-24 project_1/article/article.tex
Importance sampling convergence results.

@@
 
-	Next we will introduce importance sampling.
+	Next we will introduce importance sampling, the first prediction for importance sampling compared to 
+	brute, naïve method is that we should approach an equilibrium much faster because we are now
+	making moves in a smarter way. This prediction is confirmed in \autoref{fig:problem_c1} and 
+	\autoref{fig:problem_c2}.
+
+	\begin{figure}
+		\centering
+			\includegraphics[width=250px]{../data/figures/problem_c1.pdf}
+			\caption{As the number of Monte Carlo cycles increases the standard deviation of the local energies 
+			decrease. This effect is stronger when the importance sampling scheme is employed.}
+			\label{fig:problem_c1}
+	\end{figure}
+
+	\begin{figure}
+		\centering
+			\includegraphics[width=250px]{../data/figures/problem_c2.pdf}
+			\caption{The absolute difference between the expected energy and the exact energy for both brute force
+			and importance sampling. }
+			\label{fig:problem_c2}
+	\end{figure}
+
+	\autoref{fig:problem_c1} shows the standard deviation of the local energies for increasing number of Monte
+	Carlo cycles for both the regular sampling and the importance sampling methods. We see that the standard 
+	deviation decreases for both methods for a larger number of Monte Carlo cycles, but the standard deviation
+	is always lower for the importance sampling method. This is in accordance with our expectations.
+
+	We have also compared the expected energy with the exact energy for different number of Monte Carlo
+	cycles. The result of this analysis is shown in \autoref{fig:problem_c2}. We see that both methods will  evnetually 
+	result in expected energies that are equal to the exact energies as $\abs{E - \expval{E}} \to 0$. However, for the 
+	method of importance sampling the error is consistently lower.
 

### 2018-03-24 project_1/article/article.tex
Wrote about importance sampling results.

@@
 
+	\begin{table}
+		\caption{Importance sampled simulations of $n=10$ particles in $d=3$ dimensions. Average acceptance
+			ratios and standard deviations for eleven alpha values between $0.3$ and $0.7$.}
+		\centering
+			\begin{ruledtabular}
+				\begin{tabular}{cccc}
+				Time step & $\bar{\sigma}_b$ & $\bar{A}$ & $\expval{E}\left(\alpha = \frac{1}{2}\right)$ \\ \hline
+				$2^{+3}$ & 0.09363 & 0.00531 & 15.000 \\
+				$2^{+2}$ & 0.01549  & 0.04123 & 15.000 \\
+				$2^{+1}$ & 0.00706 & 0.21604 & 15.000 \\ 
+				$2^{\ 0}$ & 0.00675 & 0.51030 & 15.000 \\
+				$2^{-1}$ & 0.00932 & 0.73779 & 15.000 \\
+				$2^{-2}$ & 0.01578 & 0.86633 & 15.000 \\
+				$2^{-3}$ & 0.02723 & 0.93305 & 15.000 \\
+				$2^{-4}$ & 0.05016 & 0.96654 & 15.000 \\
+				$2^{-5}$ & 0.08474 & 0.98316 & 15.000 \\
+				$2^{-6}$ & 0.11555 & 0.99174 & 15.000 \\
+				$2^{-7}$ & 0.17070 & 0.99612 & 15.000 \\
+				\end{tabular}
+			\end{ruledtabular}
+			\label{tab:imp_time_step}
+	\end{table}
+
+	\autoref{tab:imp_time_step} shows the standard deviations and acceptance ratios for different time steps. The
+	simulations in this tabel are from the same span $\alpha \in [0.3, 07]$ as in \autoref{tab:1D1N_analytic} and
+	\autoref{tab:1D1N_numeric}. The optimum energies for $\alpha = \frac{1}{2}$ gives the same result for importance
+	sampling as for the naïve way of sampling. Notice that by lowering the time step, the uncertainty starts to decrease
+	and then starts to increase again. Moreover, the acceptance ratios increases as the time step decreases.
+
 \subsection{Interacting elliptical harmonic oscillator}

### 2018-03-25 project_1/vmc/vmc/interface.pyx
Add a magnitude limit
	This is used to make sure that no parameters makes a giant leap.
	Especially in our case, low alphas (~0.1) will get a large
	gradient and stabilize at the wrong minimum. Using the magnitude
	limit we reduce the size of this jump to avoid overshooting.

@@ -258,6 +258,7 @@ cdef class PySampler:
     def find_minimum(self,
             np.ndarray[double, ndim=1, mode="c"] start_parameters,
             double gamma, double gamma_scale, unsigned int max_iterations,
+            *sample_args, quiet=False, double magnitude_limit=5,
             double tol=1e-9, **sample_kwargs):
 
         cdef np.ndarray[double, ndim=1, mode="c"] parameters, \
@@ -287,6 +288,12 @@ Old parameters: {2}""".format(i, gradient, parameters))
 
             parameters_prev = parameters.copy()
             parameters += -gamma*gradient
+
+            mask = np.abs(parameters) > magnitude_limit*np.abs(parameters_prev)
+            while np.any(mask):
+                parameters[mask] /= magnitude_limit
+                mask = np.abs(parameters) \
+                        > magnitude_limit*np.abs(parameters_prev)
 
             # We only use positive variational parameters
             mask = parameters < 0

### 2018-03-25 project_1/article/article.tex
insignificant changes
Started on interacting particles result text

@@
             },
+            \label{eq:one_body_hamiltonian}
         \end{align}
@@
 
-		For our special case : INSERT DERIVATION OF LOCAL ENERGY derivative wrt alpha??
+		For our special case we only have one variational parameter, $\alpha$.  The expression for the derivative is, 
+		\begin{equation}
+			\frac{d E_L}{d\alpha} = 
+				2\left( \expval{\frac{1}{\psi_T}\frac{d\psi_T}{d\alpha} E_L} - \expval{\frac{1}{\psi_T} \frac{d\psi_T}{d\alpha}}\expval{E_L} \right).
+		\end{equation}
+
+		INSERT DERIVATION OF LOCAL ENERGY??
 
@@
 
+	\subsubsection{Test of Gradient descent}
+
+	Before moving on to the interacting, more interesting systems, it would be beneficial to test the gradient descent implementation
+	now that we are certain that it moves towards the optimum variational parameter $\alpha = \frac{1}{2}$.
+
 \subsection{Interacting elliptical harmonic oscillator}
+
+		We now make two changes to the simple systems in the previous section in order to make everything
+		a bit more interesting. First, the potential in the Hamiltonian is perturbed in the $z$-direction according to
+		\autoref{eq:one_body_hamiltonian} by setting $\lambda = \sqrt{8} \approx 2.82843$. This corresponds to enclosing particles
+		of the simulation in a spherical trap instead of a circular trap. Second, a particle-to-particle interaction is 
+		introduces in form of a hard shell with radius $\frac{a}{a_\text{ho}} = 0.0043$. Similarly to the non-interacting
+		case, the simulation results for different values of $\alpha$ are displayed in \autoref{tab:10_interacting}.
+
+		In \autoref{tab:10_interacting} the system is set to have $10$ interacting particles. Notice first that the energy is
+		higher than for the interacting system than for a comparable non-interacting system (\autoref{tab:initial_optimum_comparison}).
+		Moreover, the standard deviation gets lover as the system is close to an optimal variational parameter.  The importnance 
+		sampling method is used in this simulations, with a relatively low time step $\delta t = 0.1$, leading to a correspondingly
+		high acceptance ratio.
+
          \begin{table}
              \caption{The expected energy, the standard
-             deviation, the acceptance ratio and CPU time for each $alpha$
+             deviation, the acceptance ratio and CPU time for each $\alpha$
              for the elliptical potential with the interacting wavefunction. We
@@
 
+		The result of simulations with a higher number of particles, $n=50$ and $n=100$, are  shown in \autoref{tab:50_interacting}
+		and \autoref{tab:100_interacting} respectively. We observe that the energy for systems of interacting particles increases
+		rapidly as the size of the system increases.
+
          \begin{table}
@@
 
+\section{Derivation of Local Energy alpha derivative}
+
+TO BE INSERTED HERE??? Or whaaaat boiii...
+
 \bibliography{references}

### 2018-03-25 project_1/article/article.tex
Add typo in drift force
Add labels
Use small delta in time step
Replace tabs with spaces
	![](https://media1.giphy.com/media/l0IylSajlbPRFxH8Y/giphy.gif)

--- 2018-03-24 project_1/article/article.tex
@@
         \vf{F}_k(\vf{r})
-        &= -2\alpha\vf{r}_k.
+        &= -4\alpha\vf{r}_k.
     \end{align}
@@
             \right)dN.
+            \label{eq:exact_energy}
         \end{align}
@@
             },
+            \label{eq:scaled_hamiltonian}
         \end{align}
@@
             \phi(\vf{r}_k').
+            \label{eq:scaled_spf}
         \end{align}
@@
                 &=
-                \vf{r}_i + D\vf{F}(\vf{r}_i)\Delta t
-                + \vfg{\xi}\sqrt{\Delta t},
-            \end{align}
-            for a given time step $\Delta t$
+                \vf{r}_i + D\vf{F}(\vf{r}_i)\delta t
+                + \vfg{\xi}\sqrt{\delta t},
+            \end{align}
+            for a given time step $\delta t$
             \footnote{
@@
                 \frac{
-                    G(\vf{r}_{i + 1}, \vf{r}_{i}, \Delta t)
+                    G(\vf{r}_{i + 1}, \vf{r}_{i}, \delta t)
                     \left|\Psi_T(\vf{r}_{i + 1})\right|^2
@@
                 {
-                    G(\vf{r}_{i}, \vf{r}_{i + 1}, \Delta t)
+                    G(\vf{r}_{i}, \vf{r}_{i + 1}, \delta t)
                     \left|\Psi_T(\vf{r}_{i})\right|^2
@@
             \begin{align}
-                G(\vf{r}_{i + 1}, \vf{r}_i, \Delta t)
+                G(\vf{r}_{i + 1}, \vf{r}_i, \delta t)
                 &=
@@
                         - \vf{r}_{i}
-                        - D\vf{F}(\vf{r}_i)\Delta t
-                    \bigr]^2}{4D\Delta t}
+                        - D\vf{F}(\vf{r}_i)\delta t
+                    \bigr]^2}{4D\delta t}
                 \Biggr)
@@
                 \times
-                \frac{1}{(4\pi D\Delta t)^{dN/2}},
+                \frac{1}{(4\pi D\delta t)^{dN/2}},
             \end{align}
@@
 
-		Scanning the entire space of variational parameters can be a somewhat despairing task. To aid us in the search we
-		therefore implement the method of gradient descent\footnote{Also known as steepest descent, easily confused
-		with the steepest descent method for integation.}. In any optimum we generally have that the derivative should equal
-		zero. In our case, we want to find the minimum of the expected local energy with respect to some variational
-		parameter $\alpha$, so that
+        Scanning the entire space of variational parameters can be a somewhat
+        despairing task. To aid us in the search we therefore implement the
+        method of gradient descent\footnote{Also known as steepest descent,
+        easily confused with the steepest descent method for integation.}. In
+        any optimum we generally have that the derivative should equal zero. In
+        our case, we want to find the minimum of the expected local energy with
+        respect to some variational parameter $\alpha$, so that
         \begin{equation}
@@
         \end{equation}
-		The foundation for  the method of gradient descent lies in the fact that some analytic
-		\footnote{We use the term analytic quite recklessly.} function $F(\alpha)$ decreases fastest in the direction
-		of $-\nabla_\alpha F(\alpha)$. This means that we could eventually get to an ensuing minimum by trailing along the
-		pathway of the following recursive relation,
+        The foundation for  the method of gradient descent lies in the fact that
+        some analytic \footnote{We use the term analytic quite recklessly.}
+        function $F(\alpha)$ decreases fastest in the direction of
+        $-\nabla_\alpha F(\alpha)$. This means that we could eventually get to
+        an ensuing minimum by trailing along the pathway of the following
+        recursive relation,
         \begin{equation}
-			\alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha k} F(\alpha_k).
+            \alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha} F(\alpha_k).
         \end{equation}
-		If $\gamma$ is small enough we should have $f(\alpha_k) \geq F(\alpha_{k+1})$ for all $k$ and hopefully we will
-		experience convergence to the desired minimum.
+        If $\gamma$ is small enough we should have $f(\alpha_k) \geq
+        F(\alpha_{k+1})$ for all $k$ and hopefully we will experience
+        convergence to the desired minimum.
 
@@
 
+	\subsubsection{Test of Gradient descent}
+
+	Before moving on to the interacting, more interesting systems, it would be beneficial to test the gradient descent implementation
+	now that we are certain that it moves towards the optimum variational parameter $\alpha = \frac{1}{2}$.
+
 \subsection{Interacting elliptical harmonic oscillator}

### 2018-03-25 project_1/article/article.tex
Include results on interacting case

@@
                 \begin{tabular}{cccc}
-				Time step & $\bar{\sigma}_b$ & $\bar{A}$ & $\expval{E}\left(\alpha = \frac{1}{2}\right)$ \\ \hline
+                    $\delta t$ & $\bar{\sigma}_b$ & $\bar{A}$
+                    & $\expval{E}\left(\alpha = \frac{1}{2}\right)$ \\
+                    \hline
                     $2^{+3}$ & 0.09363 & 0.00531 & 15.000 \\
@@
 
-	\autoref{tab:imp_time_step} shows the standard deviations and acceptance ratios for different time steps. The
-	simulations in this tabel are from the same span $\alpha \in [0.3, 07]$ as in \autoref{tab:1D1N_analytic} and
-	\autoref{tab:1D1N_numeric}. The optimum energies for $\alpha = \frac{1}{2}$ gives the same result for importance
-	sampling as for the naïve way of sampling. Notice that by lowering the time step, the uncertainty starts to decrease
-	and then starts to increase again. Moreover, the acceptance ratios increases as the time step decreases.
+    \autoref{tab:imp_time_step} shows the standard deviations and acceptance
+    ratios for different time steps. The simulations in this table are from the
+    same span $\alpha \in [0.3, 07]$ as in \autoref{tab:1D1N_analytic} and
+    \autoref{tab:1D1N_numeric}. The optimum energies for $\alpha = \frac{1}{2}$
+    gives the same result for importance sampling as for the naïve way of
+    sampling. Notice that by lowering the time step, the uncertainty starts to
+    decrease and then starts to increase again. Moreover, the acceptance ratios
+    increases as the time step decreases.
 
     \subsubsection{Test of Gradient descent}
-
-	Before moving on to the interacting, more interesting systems, it would be beneficial to test the gradient descent implementation
-	now that we are certain that it moves towards the optimum variational parameter $\alpha = \frac{1}{2}$.
+        Before moving on to the interacting, more interesting systems, it would
+        be beneficial to test the gradient descent implementation now that we
+        are certain that it moves towards the optimum variational parameter
+        $\alpha = \frac{1}{2}$.
 
     \subsection{Interacting elliptical harmonic oscillator}
+        We now look at the interacting, elliptical, harmonic oscillator with the
+        scaled system shown in \autoref{eq:scaled_hamiltonian} and scaled single
+        particle functions from \autoref{eq:scaled_spf} with the Jastrow factor.
+        Due to similarities with the spherical harmonic oscillator system we
+        make a guess that the true minimum of this new system should be situated
+        close the minimum of the previous system. Choosing seven values for
+        $\alpha \in [0.2, 0.7]$ we run importance sampling on this system for $N
+        = \{10, 50, 100\}$ particles in $d = 3$ dimensions using $2^{21}$ Monte
+        Carlo cycles\footnote{This turned out to be a little bit overkill due to
+        the fast convergence of importance sampling.} and an additional $10\%$
+        of the Monte Carlo cycles for thermalization of the system prior to
+        doing any sampling. We have used $\delta t = 0.1$ with $a_{\text{ho}} =
+        0.0043$ and $\lambda = \beta = \sqrt{8}$.
+
+        We have sampled the expectation value for the local energy, the standard
+        deviation using the blocking method, the acceptance ratio of the method
+        and the CPU time spent for each value $\alpha$. The results for $N = 10$
+        particles are shown in \autoref{tab:10_interacting}, $N = 50$ in
+        \autoref{tab:50_interacting} and finally for $N = 100$ in
+        \autoref{tab:100_interacting}.
+
         \begin{table}
-             \caption{The expected energy, the standard
-             deviation, the acceptance ratio and CPU time for each $alpha$
-             for the elliptical potential with the interacting wavefunction. We
-             have used importance sampling with $\delta t = 0.1$ and $2^{21}$
-             Monte Carlo cycles and an additional $10\%$ of that as
-             thermalization steps. The hard sphere radius is set to $0.0043$ and
-             a $\beta = \lambda = \sqrt{8}$.}
+            \caption{Simulation results for $N = 10$ bosonic interacting,
+            elliptical, harmonic oscillators.}
             \centering
@@
                     $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
-                     &$A$&$t_C$ \\
+                    &$A$ & $t_{\text{CPU}} [\si{\second}]$ \\
                     \hline
-                     0.2&35.17548&0.05399&0.98995&24.48868\\
-                     0.3&27.62004&0.02311&0.98187&13.65132\\
-                     0.4&24.97850&0.00827&0.97248&13.54666\\
-                     0.5&24.39877&0.00030&0.96140&13.38388\\
-                     0.6&24.83863&0.00604&0.94998&13.31570\\
-                     0.7&25.82855&0.01059&0.93766&13.20579\\
-                     0.8&27.22895&0.01436&0.92375&13.11933\\
+                    0.2&35.17548&0.05399&0.990%98995
+                    &24.5\\%48868\\
+                    0.3&27.62004&0.02311&0.981%98187
+                    &13.6\\%65132\\
+                    0.4&24.97850&0.00827&0.972%97248
+                    &13.5\\%54666\\
+                    0.5&24.39877&0.00030&0.961%96140
+                    &13.4\\%38388\\
+                    0.6&24.83863&0.00604&0.950%94998
+                    &13.3\\%31570\\
+                    0.7&25.82855&0.01059&0.938%93766
+                    &13.2\\%20579\\
+                    0.8&27.22895&0.01436&0.924%92375
+                    &13.1\\%11933\\
                 \end{tabular}
@@
         \begin{table}
-             \caption{$50$ interacting particles.}
+            \caption{Simulation results for $N = 50$ bosonic interacting,
+            elliptical, harmonic oscillators.}
             \centering
@@
                     $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
-                     &$A$&$t_C$ \\
+                    & $A$ & $t_{\text{CPU}} [\si{\second}]$ \\
                     \hline
-                     0.2&181.69371&0.25533&0.98756&199.95366\\
-                     0.3&142.61591&0.10782&0.97849&195.80928\\
-                     0.4&129.88615&0.04051&0.96862&190.44553\\
-                     0.5&127.29926&0.00595&0.95749&195.10368\\
-                     0.6&129.97630&0.03300&0.94595&193.22755\\
-                     0.7&135.67382&0.05785&0.93330&191.96391\\
-                     0.8&143.23238&0.07286&0.92061&192.72399\\
+                    0.2&181.69371&0.25533&0.987%98756
+                    &199.9\\%95366\\
+                    0.3&142.61591&0.10782&0.978%97849
+                    &195.8\\%80928\\
+                    0.4&129.88615&0.04051&0.969%96862
+                    &190.4\\%44553\\
+                    0.5&127.29926&0.00595&0.957%95749
+                    &195.1\\%10368\\
+                    0.6&129.97630&0.03300&0.946%94595
+                    &193.2\\%22755\\
+                    0.7&135.67382&0.05785&0.933%93330
+                    &192.0\\%191.96391\\
+                    0.8&143.23238&0.07286&0.921%92061
+                    &192.7\\%72399\\
                 \end{tabular}
@@
         \begin{table}
-             \caption{$100$ Interacting particles.}
+            \caption{Simulation results for $N = 100$ bosonic interacting,
+            elliptical, harmonic oscillators.}
             \centering
@@
                     $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
-                     &$A$&$t_C$ \\
+                    &$A$ & $t_{\text{CPU}} [\si{\second}]$ \\
                     \hline
-                     0.2&375.04401&0.51475&0.98527&745.21438\\
-                     0.3&296.76879&0.20214&0.97534&741.04425\\
-                     0.4&270.60830&0.08113&0.94522&723.06823\\
-                     0.5&266.37263&0.02020&0.95314&702.71000\\
-                     0.6&272.51171&0.07366&0.94112&689.12862\\
-                     0.7&285.10285&0.11381&0.92895&707.48248\\
-                     0.8&301.40609&0.15458&0.91648&707.86118\\
+                    0.2&375.04401&0.51475&0.985%98527
+                    &745.2\\%21438\\
+                    0.3&296.76879&0.20214&0.975%97534
+                    &741.0\\%04425\\
+                    0.4&270.60830&0.08113&0.945%94522
+                    &723.1\\%06823\\
+                    0.5&266.37263&0.02020&0.953%95314
+                    &702.8\\%71000\\
+                    0.6&272.51171&0.07366&0.941%94112
+                    &689.1\\%12862\\
+                    0.7&285.10285&0.11381&0.929%92895
+                    &707.5\\%48248\\
+                    0.8&301.40609&0.15458&0.916%91648
+                    &707.9\\%86118\\
                 \end{tabular}
@@
         \end{table}
+
+        \subsubsection{Finding the minimum}
+            Using the method of gradient descent we choose a set of six starting
+            $\alpha_0 \in [0.2, 0.8]$ and try to locate a value for $\alpha$
+            where the gradient goes to zero. We looked at $N = 10$ with $d = 3$,
+            $\omega = 1$ and $\beta = \lambda = \sqrt{8}$. This yields the value
+            for the minimum $\alpha$ to be
+            \begin{align}
+                \alpha = 0.49744 \pm 0.00002.
+            \end{align}
+            In \autoref{fig:gradient_descent_interacting} we can see how the
+            conjugate gradient method ``moves'' to find the optimal value of
+            $\alpha$.
+
+            \begin{figure}
+                \includegraphics[width=250px]{../data/figures/problem_f.pdf}
+                \caption{In this figure we can see the convergence of different
+                starting values for $\alpha$ towards the true value of $\alpha$
+                that minimizes the expected energy in the interacting case.}
+                \label{fig:gradient_descent_interacting}
+            \end{figure}
+
+\section{Discussion}
+    \subsection{Energy per particle}
+        For the non-interacting harmonic oscillator system there is no
+        difference between a new ``particle'' and new dimensions
+        \footnote{This is not entirely true in the case of our implementation of
+        the Metropolis algorithms as we draw a single random particle and move
+        all dimensions instead of iterating over all particles and all
+        dimensions}. This means that the inclusion of a new particle or
+        dimension yields a linear scaling in the energy. Thus if we compute the
+        energy per particle for systems of varying size we expect to get the
+        same energy. This is most easily seen in \autoref{fig:initial_problem_b}
+        and in the expression for the exact energy \autoref{eq:exact_energy}.
+        This, however, is \textit{not} the case in the interacting system due to
+        the Jastrow factor. Looking at \autoref{tab:10_interacting} and
+        \autoref{tab:100_interacting} we see that multiplying the expected
+        energy of theformer with $10$ yields a lower number than the latter. We
+        can also see this behavior in \autoref{fig:interacting}, where the
+        energy per particle for $N = \{10, 50, 100\}$ have been plotted in the
+        same figure.
 

### 2018-03-25 project_1/article/article.tex

@@
 
-        Scanning the entire space of variational parameters can be a somewhat
-        despairing task. To aid us in the search we therefore implement the
-        method of gradient descent\footnote{Also known as steepest descent,
-        easily confused with the steepest descent method for integation.}. In
-        any optimum we generally have that the derivative should equal zero. In
-        our case, we want to find the minimum of the expected local energy with
-        respect to some variational parameter $\alpha$, so that
+        Scanning the entire space of variational parameters can be a somewhat despairing task. To aid us in the search we
+        therefore implement the method of gradient descent\footnote{Also known as steepest descent, easily confused
+        with the steepest descent method for integation.}. In any optimum we generally have that the derivative should equal
+        zero. In our case, we want to find the minimum of the expected local energy with respect to some variational
+        parameter $\alpha$, so that
         \begin{equation}
@@
         \end{equation}
-        The foundation for  the method of gradient descent lies in the fact that
-        some analytic \footnote{We use the term analytic quite recklessly.}
-        function $F(\alpha)$ decreases fastest in the direction of
-        $-\nabla_\alpha F(\alpha)$. This means that we could eventually get to
-        an ensuing minimum by trailing along the pathway of the following
-        recursive relation,
+        The foundation for  the method of gradient descent lies in the fact that some analytic
+        \footnote{We use the term analytic quite recklessly.} function $F(\alpha)$ decreases fastest in the direction
+        of $-\nabla_\alpha F(\alpha)$. This means that we could eventually get to an ensuing minimum by trailing along the
+        pathway of the following recursive relation,
         \begin{equation}
-            \alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha} F(\alpha_k).
+            \alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha k} F(\alpha_k).
         \end{equation}
-        If $\gamma$ is small enough we should have $f(\alpha_k) \geq
-        F(\alpha_{k+1})$ for all $k$ and hopefully we will experience
-        convergence to the desired minimum.
-
-        For our special case : INSERT DERIVATION OF LOCAL ENERGY derivative wrt alpha??
+        If $\gamma$ is small enough we should have $f(\alpha_k) \geq F(\alpha_{k+1})$ for all $k$ and hopefully we will
+        experience convergence to the desired minimum.
+
+        For our special case we only have one variational parameter, $\alpha$.  The expression for the derivative is,
+        \begin{equation}
+            \frac{d E_L}{d\alpha} =
+                2\left( \expval{\frac{1}{\psi_T}\frac{d\psi_T}{d\alpha} E_L} - \expval{\frac{1}{\psi_T} \frac{d\psi_T}{d\alpha}}\expval{E_L} \right).
+        \end{equation}
+
+        INSERT DERIVATION OF LOCAL ENERGY??
 
@@
     \subsection{Non-Interacting Systems}
-    Initially we are interested in seeing how our implementation performs and to what extent it functions properly.
-    the logical thing to do would be to look at a non-interacting system with a simple Gaussian wavefunction and
-    harmonic oscillator potential. Such a system has a simple closed-form analytic solution\footnote{A fact that every
-    undergraduate physics student should be well aware of after finishing their first course in quantum physics},
-    which is further simplified by the "god-given" analytic
-    units (\autoref{eq:closed_form_natural_units_local_energy}). This expression is minimised for
-    $\alpha = \frac{1}{2}\omega$. We also set $\omega = 1$ and get an expression for the
-    expected minimum local energy,
+        Initially we are interested in seeing how our implementation performs
+        and to what extent it functions properly.  the logical thing to do would
+        be to look at a non-interacting system with a simple Gaussian
+        wavefunction and harmonic oscillator potential. Such a system has a
+        simple closed-form analytic solution\footnote{A fact that every
+        undergraduate physics student should be well aware of after finishing
+        their first course in quantum physics}, which is further simplified by
+        the "god-given" analytic units
+        (\autoref{eq:closed_form_natural_units_local_energy}). This expression
+        is minimised for $\alpha = \frac{1}{2}\omega$. We also set $\omega = 1$
+        and get an expression for the expected minimum local energy,
         \begin{equation}
@@
 
-    We will start without importance sampling, that is by employing the simpler and more naïve "brute force" Monte Carlo
-    sampling. Moreover, we will compute the value for the local energy analytically as well as with a numerical approach.
-    In the numerical approach we employ the well-know approximation for the second derivative,
+        We will start without importance sampling, that is by employing the
+        simpler and more naïve "brute force" Monte Carlo sampling. Moreover, we
+        will compute the value for the local energy analytically as well as with
+        a numerical approach.  In the numerical approach we employ the well-know
+        approximation for the second derivative,
         \begin{equation}
@@
             \centering
-        \caption{One particle in one dimension for the analytic expression with
-        $2^{21}$ Monte Carlo cycles and step length of $0.5$.}
+            \caption{One particle in one dimension for the analytic expression
+            with $2^{21}$ Monte Carlo cycles and step length of $0.5$.}
             \begin{ruledtabular}
@@
             \centering
-        \caption{One particle in one dimension for the numeric expression with
-        $2^{21}$ Monte Carlo cycles and a step length of $0.5$.}
+            \caption{One particle in one dimension for the numeric expression
+            with $2^{21}$ Monte Carlo cycles and a step length of $0.5$.}
             \begin{ruledtabular}
@@
 
-    \autoref{tab:1D1N_analytic} shows the expected local energy $\expval{E_L}$, standard deviations
-    $\sigma$, $\sigma_b$, the share of accepted Monte Carlo reorientations $A$ and cpu time for
-    the analytic computation at different values for the variational parameter $\alpha$.
-    \autoref{tab:1D1N_numeric} shows the same numbers but for the numeric scheme. The first column of
-    standard deviations $\sigma$ are the "naïve" standard deviations, while $\sigma_b$ marks the
-    standard deviations computed with the blocking method. One can clearly see, in both the analytic and numeric
-    case, that the former standard deviations are lower than the latter, meaning that autocorrelated data
-    can lead to a lower perceived uncertainty than what is reality. The blocking method takes this fact into account
-    and for this reason, all other standard deviations will be computed with the blocking method from here on.
-
-    The first initial observation of note is that we have found an optimum for the system in terms of a minimum
-    energy at $\alpha=\frac{1}{2}$. Moreover, notice how the standard deviation disappears completely.
-    This should already be apparent from \autoref{eq:closed_form_natural_units_local_energy}, as the term
+        \autoref{tab:1D1N_analytic} shows the expected local energy
+        $\expval{E_L}$, standard deviations $\sigma$, $\sigma_b$, the share of
+        accepted Monte Carlo reorientations $A$ and cpu time for the analytic
+        computation at different values for the variational parameter $\alpha$.
+        \autoref{tab:1D1N_numeric} shows the same numbers but for the numeric
+        scheme. The first column of standard deviations $\sigma$ are the "naïve"
+        standard deviations, while $\sigma_b$ marks the standard deviations
+        computed with the blocking method. One can clearly see, in both the
+        analytic and numeric case, that the former standard deviations are lower
+        than the latter, meaning that autocorrelated data can lead to a lower
+        perceived uncertainty than what is reality. The blocking method takes
+        this fact into account and for this reason, all other standard
+        deviations will be computed with the blocking method from here on.
+
+        The first initial observation of note is that we have found an optimum
+        for the system in terms of a minimum energy at $\alpha=\frac{1}{2}$.
+        Moreover, notice how the standard deviation disappears completely.  This
+        should already be apparent from
+        \autoref{eq:closed_form_natural_units_local_energy}, as the term
         involving particle positions disappears for the optimal $\alpha$.
@@
         \begin{table}
-        \caption{Comparison of energy, standard deviation and the CPU time for
-        $\alpha = 0.5$ (the correct minimum) for the analytic and numeric
-        schemes in three dimensions and $N$ particles. The step length is
-        $0.5$.}
+            \caption{Comparison of energy, standard deviation and the CPU time
+            for $\alpha = 0.5$ (the correct minimum) for the analytic and
+            numeric schemes in three dimensions and $N$ particles. The step
+            length is $0.5$.}
             \centering
@@
 
-    Systems of optimum variational parameter warrants further investigation, as summed up in
-    \autoref{tab:initial_optimum_comparison}. The analytical results are quite predictable and
-    monotonous; the expected local energy $\expval{E_l}$ is perfectly proportional by a factor
-    $\frac{3}{2}$ to the number of particles and the standard deviation $\sigma_b$ is zero for
-    systems of any number of particles. The numerical results tells a more tortuous tale. As the
-    number of particles increase, the uncertainty also increases rapidly. For the largest system,
-    with $N = 500$ particles, the energy misses its target critically - the true energy is more than
-    one standard deviation away. Lastly, the cpu time is notably higher for the numeric scheme.
-
-    In the appendices,  \autoref{fig:initial_problem_b} shows how both the analytic and numeric approach
-    matches the exact expression for the energy as a function of $\alpha$ for systems of increasing size.
-    We see also here that the uncertainty increases in the numeric scheme as the system gets larger.
+        Systems of optimum variational parameter warrants further investigation,
+        as summed up in \autoref{tab:initial_optimum_comparison}. The analytical
+        results are quite predictable and monotonous; the expected local energy
+        $\expval{E_l}$ is perfectly proportional by a factor $\frac{3}{2}$ to
+        the number of particles and the standard deviation $\sigma_b$ is zero
+        for systems of any number of particles. The numerical results tells a
+        more tortuous tale. As the number of particles increase, the uncertainty
+        also increases rapidly. For the largest system, with $N = 500$
+        particles, the energy misses its target critically - the true energy is
+        more than one standard deviation away. Lastly, the cpu time is notably
+        higher for the numeric scheme.
+
+        In the appendices,  \autoref{fig:initial_problem_b} shows how both the
+        analytic and numeric approach matches the exact expression for the
+        energy as a function of $\alpha$ for systems of increasing size.  We see
+        also here that the uncertainty increases in the numeric scheme as the
+        system gets larger.
 
@@
 
-    Next we will introduce importance sampling, the first prediction for importance sampling compared to
-    brute, naïve method is that we should approach an equilibrium much faster because we are now
-    making moves in a smarter way. This prediction is confirmed in \autoref{fig:problem_c1} and
-    \autoref{fig:problem_c2}.
+            Next we will introduce importance sampling, the first prediction for
+            importance sampling compared to brute, naïve method is that we
+            should approach an equilibrium much faster because we are now making
+            moves in a smarter way. This prediction is confirmed in
+            \autoref{fig:problem_c1} and \autoref{fig:problem_c2}.
 
@@
                 \centering
-            \includegraphics[width=250px]{../data/figures/problem_c1.pdf}
-            \caption{As the number of Monte Carlo cycles increases the standard deviation of the local energies
-            decrease. This effect is stronger when the importance sampling scheme is employed.}
+                    \includegraphics[width=250px]
+                    {../data/figures/problem_c1.pdf}
+                    \caption{As the number of Monte Carlo cycles increases the
+                    standard deviation of the local energies decrease. This
+                    effect is stronger when the importance sampling scheme is
+                    employed.}
                     \label{fig:problem_c1}
@@
                 \centering
-            \includegraphics[width=250px]{../data/figures/problem_c2.pdf}
-            \caption{The absolute difference between the expected energy and the exact energy for both brute force
-            and importance sampling. }
+                    \includegraphics[width=250px]
+                    {../data/figures/problem_c2.pdf}
+                    \caption{The absolute difference between the expected energy
+                    and the exact energy for both brute force and importance
+                    sampling. }
                     \label{fig:problem_c2}
@@
 
-    \autoref{fig:problem_c1} shows the standard deviation of the local energies for increasing number of Monte
-    Carlo cycles for both the regular sampling and the importance sampling methods. We see that the standard
-    deviation decreases for both methods for a larger number of Monte Carlo cycles, but the standard deviation
-    is always lower for the importance sampling method. This is in accordance with our expectations.
-
-    We have also compared the expected energy with the exact energy for different number of Monte Carlo
-    cycles. The result of this analysis is shown in \autoref{fig:problem_c2}. We see that both methods will  evnetually
-    result in expected energies that are equal to the exact energies as $\abs{E - \expval{E}} \to 0$. However, for the
-    method of importance sampling the error is consistently lower.
+            \autoref{fig:problem_c1} shows the standard deviation of the local
+            energies for increasing number of Monte Carlo cycles for both the
+            regular sampling and the importance sampling methods. We see that
+            the standard deviation decreases for both methods for a larger
+            number of Monte Carlo cycles, but the standard deviation is always
+            lower for the importance sampling method. This is in accordance with
+            our expectations.
+
+            We have also compared the expected energy with the exact energy for
+            different number of Monte Carlo cycles. The result of this analysis
+            is shown in \autoref{fig:problem_c2}. We see that both methods will
+            evnetually result in expected energies that are equal to the exact
+            energies as $\abs{E - \expval{E}} \to 0$. However, for the method of
+            importance sampling the error is consistently lower.
 
             \begin{table}
-        \caption{Importance sampled simulations of $n=10$ particles in $d=3$ dimensions. Average acceptance
-            ratios and standard deviations for eleven alpha values between $0.3$ and $0.7$.}
+                \caption{Importance sampled simulations of $n=10$ particles in
+                $d=3$ dimensions. Average acceptance ratios and standard
+                deviations for eleven alpha values between $0.3$ and $0.7$.}
                 \centering
@@
 
-    \autoref{tab:imp_time_step} shows the standard deviations and acceptance
-    ratios for different time steps. The simulations in this table are from the
-    same span $\alpha \in [0.3, 07]$ as in \autoref{tab:1D1N_analytic} and
-    \autoref{tab:1D1N_numeric}. The optimum energies for $\alpha = \frac{1}{2}$
-    gives the same result for importance sampling as for the naïve way of
-    sampling. Notice that by lowering the time step, the uncertainty starts to
-    decrease and then starts to increase again. Moreover, the acceptance ratios
+            \autoref{tab:imp_time_step} shows the standard deviations and
+            acceptance ratios for different time steps. The simulations in this
+            tabel are from the same span $\alpha \in [0.3, 07]$ as in
+            \autoref{tab:1D1N_analytic} and \autoref{tab:1D1N_numeric}. The
+            optimum energies for $\alpha = \frac{1}{2}$ gives the same result
+            for importance sampling as for the naïve way of sampling. Notice
+            that by lowering the time step, the uncertainty starts to decrease
+            and then starts to increase again. Moreover, the acceptance ratios
             increases as the time step decreases.
@@
         \subsubsection{Test of Gradient descent}
-        Before moving on to the interacting, more interesting systems, it would
-        be beneficial to test the gradient descent implementation now that we
-        are certain that it moves towards the optimum variational parameter
-        $\alpha = \frac{1}{2}$.
+
+            Before moving on to the interacting, more interesting systems, it
+            would be beneficial to test the gradient descent implementation now
+            that we are certain that it moves towards the optimum variational
+            parameter $\alpha = \frac{1}{2}$.
 
     \subsection{Interacting elliptical harmonic oscillator}
-        We now look at the interacting, elliptical, harmonic oscillator with the
-        scaled system shown in \autoref{eq:scaled_hamiltonian} and scaled single
-        particle functions from \autoref{eq:scaled_spf} with the Jastrow factor.
-        Due to similarities with the spherical harmonic oscillator system we
-        make a guess that the true minimum of this new system should be situated
-        close the minimum of the previous system. Choosing seven values for
-        $\alpha \in [0.2, 0.7]$ we run importance sampling on this system for $N
-        = \{10, 50, 100\}$ particles in $d = 3$ dimensions using $2^{21}$ Monte
-        Carlo cycles\footnote{This turned out to be a little bit overkill due to
-        the fast convergence of importance sampling.} and an additional $10\%$
-        of the Monte Carlo cycles for thermalization of the system prior to
-        doing any sampling. We have used $\delta t = 0.1$ with $a_{\text{ho}} =
-        0.0043$ and $\lambda = \beta = \sqrt{8}$.
-
-        We have sampled the expectation value for the local energy, the standard
-        deviation using the blocking method, the acceptance ratio of the method
-        and the CPU time spent for each value $\alpha$. The results for $N = 10$
-        particles are shown in \autoref{tab:10_interacting}, $N = 50$ in
-        \autoref{tab:50_interacting} and finally for $N = 100$ in
-        \autoref{tab:100_interacting}.
+        We now make a change to the simple systems in the previous section in
+        order to make everything a bit more interesting. The potential in the
+        one-body Hamiltonian is perturbed in the $z$-direction according to
+        \autoref{eq:scaled_hamiltonian} by setting $\lambda = \sqrt{8} \approx
+        2.82843$. This corresponds to enclosing particles of the simulation in
+        an elliptical trap instead of a spherical trap.
+
+        We also choose a new trial wavefunction with the elliptical gaussian
+        single particle functions in \autoref{eq:scaled_spf} and the Jastrow
+        factor. The hard sphere radius is set to $a/a_{\text{ho}} = 0.0043$.
+
+        Due to similarities with the spherical harmonic oscillator
+        system we make a guess that the true minimum of this new system should
+        be situated close to the minimum of the previous system. Choosing seven
+        values for $\alpha \in [0.2, 0.7]$ we run importance sampling on this
+        system for $N = \{10, 50, 100\}$ particles in $d = 3$ dimensions using
+        $2^{21}$ Monte Carlo cycles\footnote{This turned out to be a little bit
+        overkill due to the fast convergence of importance sampling.} and an
+        additional $10\%$ of the Monte Carlo cycles for thermalization of the
+        system prior to doing any sampling. We have used $\delta t = 0.1$ yields
+        a high acceptance ratio, but slower convergence.
+
+        In \autoref{tab:10_interacting} the system is set to have $N = 10$
+        interacting particles. Notice first that the energy is higher in
+        the interacting system than for a comparable non-interacting system
+        (\autoref{tab:initial_optimum_comparison}).  Moreover, the standard
+        deviation gets lower as the system is close to an optimal variational
+        parameter.
 
@@
 
+        The result of simulations with a higher number of particles, $N=50$ and
+        $N=100$, are  shown in \autoref{tab:50_interacting} and
+        \autoref{tab:100_interacting} respectively. We observe that the energy
+        for systems of interacting particles increases rapidly as the size of
+        the system increases.
+
         \begin{table}

### 2018-03-25 project_1/article/article.tex
Add expression for the expectation value of energy

@@
     where $a$ is the hard sphere of the particle. The trial wavefunction,
-    $\ket{\Psi_T}$, we will
-    be looking at is given by
+    $\Psi_T(\vf{r})$, we will be looking at is given by
     \begin{align}
@@
     \end{align}
-    and $\ket{\Phi_T}$ the \textit{Slater permanent} consisting of the $N$ first
-    single particle wavefunctions, and the correlation wavefunction, $f$, given
-    by
+    and $\Phi_T(\vf{r})$ the \textit{Slater permanent} consisting of the $N$
+    first single particle wavefunctions, and the correlation wavefunction, $f$,
+    given by
     \begin{align}
@@
         \end{align}
-        If $\ket{\Psi_T}$ is an exact eigenfunction of the Hamiltonian, $E_L$
-        will be constant. The closer $\ket{\Psi_T}$ is to the exact wave
+        If $\Psi_T(\vf{r})$ is an exact eigenfunction of the Hamiltonian, $E_L$
+        will be constant. The closer $\Psi_T(\vf{r})$ is to the exact wave
         function, the less variation in $E_L$ as a function of $\vf{r}$ we get.
@@
         $E_L$ in terms of the trial wavefunction.
+
+        When we are performing the Monte Carlo sampling we are interested in the
+        expected value of the local energy as this will serve as our estimate of
+        the true energy. Mathematically this is expressed as
+        \begin{align}
+            \expval{E_L}
+            &=
+            \frac{\int\dd \vf{r}\Psi_T^{*}(\vf{r})H\Psi_T{\vf{r}}}
+            {\int\dd \vf{r}|\Psi_T(\vf{r})|^2}
+            =
+            \frac{
+                \int\dd \vf{r}|\Psi_T(\vf{r})|^2 E_L
+            }{
+                \int\dd \vf{r}|\Psi_T(\vf{r})|^2
+            },
+        \end{align}
+        where we have multiplied and divided by the trial wavefunction to get
+        the latter expression.
 

### 2018-03-25 project_1/article/article.tex
Remove overfull hbox in equation and picture
Use general variational parameter gradient

@@
         one-point lag of the series. This covariance is usually referred to as the first order autocovariance.
-        \begin{equation}
-            \gamma_i = \frac{1}{n}
-                \left(\sum_{k=1}^n\left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)\right)
-                \left(\sum_{k=0}^{n-1}\left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)\right).
-        \end{equation}
+        \begin{align}
+            \gamma_i
+            &= \frac{1}{n}
+            \left(
+                \sum_{k=1}^n
+                \left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)
+            \right)
+            \nonumber \\
+            &\qquad
+            \times
+            \left(
+                \sum_{k=0}^{n-1}
+                \left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)
+            \right).
+        \end{align}
 
@@
             \centering
-
             % Defining blocks of flowchart
@@
                     {Transform data $(\mathbf{X}_{i+1})_k = \frac{1}{2}\left((\mathbf{X})_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right)$};
-                    \node [block, right=1.5cm of transform] (i+1) {Next iteration: $i = i + 1$.} ;
+                \node [block, right=1.0cm of transform] (i+1) {Next iteration: $i = i + 1$.} ;
                 \node [decision, below of = transform] (x_size) {Length of $\mathbf{X}_{i + 1} \geq 2$ ? };
@@
 
-        Scanning the entire space of variational parameters can be a somewhat despairing task. To aid us in the search we
-        therefore implement the method of gradient descent\footnote{Also known as steepest descent, easily confused
-        with the steepest descent method for integation.}. In any optimum we generally have that the derivative should equal
-        zero. In our case, we want to find the minimum of the expected local energy with respect to some variational
-        parameter $\alpha$, so that
+            Scanning the entire space of variational parameters can be a
+            somewhat despairing task. To aid us in the search we therefore
+            implement the method of gradient descent\footnote{Also known as
+            steepest descent, easily confused with the steepest descent method
+            for integation.}. In any optimum we generally have that the
+            derivative should equal zero. In our case, we want to find the
+            minimum of the expected local energy with respect to some
+            variational parameter $\alpha$, so that
             \begin{equation}
@@
             \end{equation}
-        The foundation for  the method of gradient descent lies in the fact that some analytic
-        \footnote{We use the term analytic quite recklessly.} function $F(\alpha)$ decreases fastest in the direction
-        of $-\nabla_\alpha F(\alpha)$. This means that we could eventually get to an ensuing minimum by trailing along the
-        pathway of the following recursive relation,
+            The foundation for  the method of gradient descent lies in the fact
+            that some analytic \footnote{We use the term analytic quite
+            recklessly.} function $F(\alpha)$ decreases fastest in the direction
+            of $-\nabla_\alpha F(\alpha)$. This means that we could eventually
+            get to an ensuing minimum by trailing along the pathway of the
+            following recursive relation,
             \begin{equation}
-            \alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha k} F(\alpha_k).
+                \alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha} F(\alpha_k).
             \end{equation}
-        If $\gamma$ is small enough we should have $f(\alpha_k) \geq F(\alpha_{k+1})$ for all $k$ and hopefully we will
-        experience convergence to the desired minimum.
-
-        For our special case we only have one variational parameter, $\alpha$.  The expression for the derivative is,
+            If $\gamma$ is small enough we should have $f(\alpha_k) \geq
+            F(\alpha_{k+1})$ for all $k$ and hopefully we will experience
+            convergence to the desired minimum.
+
+            For our special case we only have one variational parameter,
+            $\alpha$.  The expression for the derivative is,
             \begin{equation}
-            \frac{d E_L}{d\alpha} =
-                2\left( \expval{\frac{1}{\psi_T}\frac{d\psi_T}{d\alpha} E_L} - \expval{\frac{1}{\psi_T} \frac{d\psi_T}{d\alpha}}\expval{E_L} \right).
+                \nabla_{\alpha}\expval{E_L}
+                =
+                2\left(
+                    \expval{E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
+                    - \expval{\frac{1}{\Psi_T} \nabla_{\alpha}\Psi_T}\expval{E_L}
+                \right).
+                \label{eq:variational_parameter_gradient}
             \end{equation}
 
-        INSERT DERIVATION OF LOCAL ENERGY??
+            %TODO: INSERT DERIVATION OF LOCAL ENERGY??
 

### 2018-03-25 project_1/article/article.tex
Scale images to avoid overfull hbox

@@
                 \centering
-                    \includegraphics[width=250px]
+                    \includegraphics[width=244px]
                     {../data/figures/problem_c1.pdf}
@@
                 \centering
-                    \includegraphics[width=250px]
+                    \includegraphics[width=244px]
                     {../data/figures/problem_c2.pdf}
@@
             \begin{figure}
-                \includegraphics[width=250px]{../data/figures/problem_f.pdf}
+                \includegraphics[width=244px]{../data/figures/problem_f.pdf}
                 \caption{In this figure we can see the convergence of different
@@
         \begin{figure}
-            \includegraphics[width=250px]{../data/figures/problem_e.pdf}
-            \caption{Local energies as a function of variational parameter $\alpha$
-                             for different number of particles.}
+            \includegraphics[width=244px]{../data/figures/problem_e.pdf}
+            \caption{Local energies as a function of variational parameter
+            $\alpha$ for different number of particles.}
             \label{fig:interacting}

### 2018-03-25 project_1/article/article.tex
Quick maths!

@@
 
-\section{Derivation of Local Energy alpha derivative}
-
-TO BE INSERTED HERE??? Or whaaaat boiii...
+\section{Variational parameter gradient of the expectation energy}
+    Here we show how to arrive at the expression shown in
+    \autoref{eq:variational_parameter_gradient}. This requires us to restrict
+    our view to real trial wavefunctions, i.e., $\Psi_T^{*} = \Psi_T$, which is
+    the case for the wavefunctions we are exploring.
+    \begin{align}
+        \nabla_{\alpha}\expval{E_L}
+        &=
+        \nabla_{\alpha}
+        \para{
+            \frac{
+                \int\dd\vf{r}\Psi_T^{*}H\Psi_T
+            }{
+                \int\dd\vf{r}|\Psi_T|^2
+            }
+        }
+        \\
+        &=
+        -
+        \frac{
+            2
+            \para{
+                \int\dd\vf{r}\Psi_T H\Psi_T
+            }
+        }{\para{\int\dd\vf{r}\Psi_T}^2}
+        \int\dd\vf{r}\Psi_T\nabla_{\alpha}\Psi_T
+        \\
+        &\qquad
+        +
+        \frac{2}{\int\dd\vf{r}\Psi_T^2}
+        \int\dd\vf{r}\Psi_TH\brak{
+            \nabla_{\alpha}\Psi_T
+        }
+        \\
+        &=
+        - 2\expval{E_L}
+        \frac{1}{
+            \int\dd\vf{r}\Psi_T^2
+        }
+        \int\dd\vf{r}\Psi_T^2\para{
+            \frac{\nabla_{\alpha}\Psi_T}{\Psi_T}
+        }
+        \\
+        &\qquad
+        +
+        \frac{2}{\int\dd\vf{r}\Psi_T^2}
+        \int\dd\vf{r}\Psi_T^2 E_L\para{
+            \frac{\nabla_{\alpha}\Psi_T}{\Psi_T}
+        }.
+    \end{align}
+    We now use the definition of the expectation value to group terms together.
+    We are thus left with
+    \begin{align}
+        \nabla_{\alpha}\expval{E_L}
+        =
+        2
+        \para{
+            \expval{
+                E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T
+            }
+            - \expval{E_L}
+            \expval{\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
+        },
+    \end{align}
+    which is what we wanted to show.
 

### 2018-03-25 project_1/article/article.tex
Include link to source code
Reduce number of significant digits
Fix typo

@@
 \author{Winther-Larsen, Sebastian Gregorius}
+\homepage[Project code: ]{https://github.com/Schoyen/FYS4411}
 \affiliation{University of Oslo}
 \author{Schøyen, Øyvind Sigmundson}
+\homepage[Project code: ]{https://github.com/Schoyen/FYS4411}
 \affiliation{University of Oslo}
@@
                 \begin{tabular}{cccccc}
-                    $\alpha$ & $\langle  E_L\rangle$ & $\sigma$ & $\sigma_b$ & $A$
-                    & $t_{\text{CPU}} [\si{\second}]$ \\
+                    $\alpha$ & $\langle  E_L\rangle$ & $\sigma$ & $\sigma_b$
+                    & $A$ & $t_{\text{CPU}} [\si{\second}]$ \\
                     \hline
-                    0.30&0.56244&0.00026&0.00129&0.89185&0.43238\\
-                    0.34&0.53652&0.00019&0.00094&0.88465&0.38340\\
-                    0.38&0.51863&0.00013&0.00063&0.87844&0.37200\\
-                    0.42&0.50805&0.00009&0.00038&0.87161&0.37978\\
-                    0.46&0.50201&0.00004&0.00017&0.86594&0.37214\\
-                    0.50&0.50000&0.00000&0.00000&0.86047&0.37162\\
-                    0.54&0.50183&0.00004&0.00015&0.85543&0.37959\\
-                    0.58&0.50566&0.00007&0.00029&0.84994&0.37778\\
-                    0.62&0.51175&0.00011&0.00041&0.84479&0.37368\\
-                    0.66&0.51906&0.00014&0.00052&0.84000&0.37257\\
-                    0.70&0.52883&0.00017&0.00062&0.83570&0.37342\\
+                    0.30&0.56244&0.00026&0.00129&0.892&0.43\\
+                    0.34&0.53652&0.00019&0.00094&0.885&0.38\\
+                    0.38&0.51863&0.00013&0.00063&0.878&0.37\\
+                    0.42&0.50805&0.00009&0.00038&0.872&0.38\\
+                    0.46&0.50201&0.00004&0.00017&0.866&0.37\\
+                    0.50&0.50000&0.00000&0.00000&0.860&0.37\\
+                    0.54&0.50183&0.00004&0.00015&0.855&0.38\\
+                    0.58&0.50566&0.00007&0.00029&0.850&0.38\\
+                    0.62&0.51175&0.00011&0.00041&0.845&0.37\\
+                    0.66&0.51906&0.00014&0.00052&0.840&0.37\\
+                    0.70&0.52883&0.00017&0.00062&0.836&0.37\\
                 \end{tabular}
@@
                     \hline
-                    0.30&0.56675&0.00026&0.00135&0.89098&0.52634\\
-                    0.34&0.53714&0.00019&0.00096&0.88479&0.52186\\
-                    0.38&0.51836&0.00013&0.00062&0.87820&0.52351\\
-                    0.42&0.50717&0.00009&0.00038&0.87180&0.52892\\
-                    0.46&0.50184&0.00004&0.00017&0.86581&0.51759\\
-                    0.50&0.50000&0.00001&0.00001&0.86024&0.52427\\
-                    0.54&0.50133&0.00004&0.00015&0.85520&0.52052\\
-                    0.58&0.50520&0.00007&0.00028&0.84971&0.52156\\
-                    0.62&0.51120&0.00011&0.00039&0.84481&0.51830\\
-                    0.66&0.51896&0.00014&0.00052&0.83990&0.51865\\
-                    0.70&0.52794&0.00017&0.00061&0.83506&0.51255\\
+                    0.30&0.56675&0.00026&0.00135&0.891&0.53\\
+                    0.34&0.53714&0.00019&0.00096&0.885&0.52\\
+                    0.38&0.51836&0.00013&0.00062&0.878&0.52\\
+                    0.42&0.50717&0.00009&0.00038&0.872&0.53\\
+                    0.46&0.50184&0.00004&0.00017&0.866&0.52\\
+                    0.50&0.50000&0.00001&0.00001&0.860&0.52\\
+                    0.54&0.50133&0.00004&0.00015&0.855&0.52\\
+                    0.58&0.50520&0.00007&0.00028&0.850&0.52\\
+                    0.62&0.51120&0.00011&0.00039&0.845&0.52\\
+                    0.66&0.51896&0.00014&0.00052&0.840&0.52\\
+                    0.70&0.52794&0.00017&0.00061&0.835&0.51\\
                 \end{tabular}
@@
         results are quite predictable and monotonous; the expected local energy
-        $\expval{E_l}$ is perfectly proportional by a factor $\frac{3}{2}$ to
+        $\expval{E_L}$ is perfectly proportional by a factor $\frac{3}{2}$ to
         the number of particles and the standard deviation $\sigma_b$ is zero
@@
                             $\delta t$ & $\bar{\sigma}_b$ & $\bar{A}$
-                            & $\expval{E}\left(\alpha = \frac{1}{2}\right)$ \\
+                            & $\expval{E}\left(\alpha = 1/2\right)$ \\
                             \hline
-                            $2^{+3}$ & 0.09363 & 0.00531 & 15.000 \\
-                            $2^{+2}$ & 0.01549  & 0.04123 & 15.000 \\
-                            $2^{+1}$ & 0.00706 & 0.21604 & 15.000 \\
-                            $2^{\ 0}$ & 0.00675 & 0.51030 & 15.000 \\
-                            $2^{-1}$ & 0.00932 & 0.73779 & 15.000 \\
-                            $2^{-2}$ & 0.01578 & 0.86633 & 15.000 \\
-                            $2^{-3}$ & 0.02723 & 0.93305 & 15.000 \\
-                            $2^{-4}$ & 0.05016 & 0.96654 & 15.000 \\
-                            $2^{-5}$ & 0.08474 & 0.98316 & 15.000 \\
-                            $2^{-6}$ & 0.11555 & 0.99174 & 15.000 \\
-                            $2^{-7}$ & 0.17070 & 0.99612 & 15.000 \\
+                            $2^{+3}$ & 0.09363 & 0.005 & 15.000 \\
+                            $2^{+2}$ & 0.01549  & 0.041 & 15.000 \\
+                            $2^{+1}$ & 0.00706 & 0.216 & 15.000 \\
+                            $2^{\ 0}$ & 0.00675 & 0.510 & 15.000 \\
+                            $2^{-1}$ & 0.00932 & 0.738 & 15.000 \\
+                            $2^{-2}$ & 0.01578 & 0.866 & 15.000 \\
+                            $2^{-3}$ & 0.02723 & 0.933 & 15.000 \\
+                            $2^{-4}$ & 0.05016 & 0.966 & 15.000 \\
+                            $2^{-5}$ & 0.08474 & 0.983 & 15.000 \\
+                            $2^{-6}$ & 0.11555 & 0.992 & 15.000 \\
+                            $2^{-7}$ & 0.17070 & 0.996 & 15.000 \\
                         \end{tabular}

### 2018-03-25 project_1/article/article.tex
Add punctuation to footnote

@@
         all dimensions instead of iterating over all particles and all
-        dimensions}. This means that the inclusion of a new particle or
+        dimensions.}. This means that the inclusion of a new particle or
         dimension yields a linear scaling in the energy. Thus if we compute the

### 2018-04-01 project_1/article/article.tex
Introduction

@@
     gas. \cite{dubois2001bose} \cite{nilsen2005vortices}
+
+    By building up an increasingly intricate system we eloquently illustrate the
+    strength of the VMC methods. Any Monte Carlo method draws on the fact that
+    the computer will not complain when being tasked to perform a very similar
+    task up to several million times in rapid succession. In general Monte Carlo
+    methods is in essence a series of guesses where the current guess is only a minor 
+    alteration of the one prior to it. The trick lies in only accepting the subsequent 
+    guesses that improves upon our configuration up to a certain likelihood.
+
+    The method we employed is called a \emph{Variational} Monte Carlo method, 
+    because we make an educated guess at a wavefunction which we allow to contain 
+    certain \emph{variational parameters}. We shall see that the system converges to a proper
+    \footnote{Global, as far as we can tell.} minimum for only a certain set of parameters.
+
+    First, we start by explaining our theoretical quantum system of a bosonic gas trapped
+    in a potential well. Here we also outline the principle under which particles in the gas are
+    allowed to interact. Moreover, we introduce the concept of \emph{local energy} and
+    then quantum \emph{drift force}. Second, we outline in more detailed algebra the
+    nature of a non-interaction system. For this system we are able to derive an exact 
+    variational energy to which we can compare the results from our simulations. Third,
+    we give a description of the particle-particle interaction implemented as if each 
+    particle is a hard sphere. Fourth, we provide a thorough description of algorithms and
+    statistical methods employed in this study. Here we introduce the Metropolis-Hastings
+    algorithm. This algorithm can by a deft trick be expanded with the method of 
+    \emph{importance sampling} which make the way in which we pick the next step in
+   	the Monte-Carlo cycling much more intelligent. The main point of note with regards to 
+   	the statistical analysis is that the data generated by the Monte Carlo cycling is
+   	autocorrelated and computation of variance must be done with the \emph{blocking method}.
+   	As an automated way of finding the optimal variational parameter, we sketch out the
+   	method of \emph{gradient descent}. Fifth, we put the machinere described in the 
+   	previous methods to work. We test against the actual values in a non-interacting system
+   	before analysing the more interesting perturbed-trap system with interacting particles. 
+   	Lastly, we highlight the more interesting aspects of this study in a discussion of the results
+   	before ending on some summary remarks.
 

### 2018-04-02 project_1/article/article.tex
Removed unecessary subsection
Figure fix, a bit of discussion writing..

@@
 
-        \subsubsection{Test of Gradient descent}
-
-            Before moving on to the interacting, more interesting systems, it
-            would be beneficial to test the gradient descent implementation now
-            that we are certain that it moves towards the optimum variational
-            parameter $\alpha = \frac{1}{2}$.
-
     \subsection{Interacting elliptical harmonic oscillator}
@@
 \section{Discussion}
+
+	\subsection{Validity of Results}
+		The non-interacting tests compared against the analytic "solution" showed
+		to a very high degree the validity of the method. One can see this quite clearly 
+		from \autoref{fig:initial_problem_b}. Because we get results from simulations
+		that correspond to the analytic results we can be fairly certain that the method
+		can be expanded upon to include attributes outside the realm problems that are 
+		analytically solvable. The systems with elliptic traps and interacting particles are
+		such systems.
+
     \subsection{Energy per particle}
@@
 
+\section{Summary Remarks}
+
 \appendix

### 2018-04-03 project_1/article/article.tex
Add section on stability

@@
             that some analytic \footnote{We use the term analytic quite
-            recklessly.} function $F(\alpha)$ decreases fastest in the direction
-            of $-\nabla_\alpha F(\alpha)$. This means that we could eventually
+            recklessly.} function $\vf{F}(\alpha)$ decreases fastest in the direction
+            of $-\nabla_\alpha\cdot\vf{F}(\alpha)$. This means that we could eventually
             get to an ensuing minimum by trailing along the pathway of the
@@
             \begin{equation}
-                \alpha_{k+1} = \alpha_k - \gamma \nabla_{\alpha} F(\alpha_k).
+                \vfg{\alpha}_{k+1}
+                = \vfg{\alpha}_k - \gamma \nabla_{\alpha}\cdot \vf{F}(\alpha_k).
+                \label{eq:parameter_difference}
             \end{equation}
-            If $\gamma$ is small enough we should have $f(\alpha_k) \geq
-            F(\alpha_{k+1})$ for all $k$ and hopefully we will experience
+            If $\gamma$ is small enough we should have $\vf{F}(\alpha_k) \geq
+            \vf{F}(\alpha_{k+1})$ for all $k$ and hopefully we will experience
             convergence to the desired minimum.
@@
             $\alpha$. The expression for the derivative is,
-            \begin{equation}
+            \begin{align}
                 \nabla_{\alpha}\expval{E_L}
@@
                     \expval{E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
-                    - \expval{\frac{1}{\Psi_T} \nabla_{\alpha}\Psi_T}\expval{E_L}
+                    - \expval{\frac{1}{\Psi_T} \nabla_{\alpha}\Psi_T}
+                    \expval{E_L}
                 \right).
                 \label{eq:variational_parameter_gradient}
-            \end{equation}
-
-            %TODO: INSERT DERIVATION OF LOCAL ENERGY??
+            \end{align}
+            To evaluate this expression we sample the local energy, the
+            parameter gradient of the wavefunction, and the combined expression
+            of these at each Monte Carlo cycle.
+
+            \subsubsection{Stability of the algorithm}
+                When employing the method of gradient descent to locate a
+                minimum value for the variational parameters we expect some
+                issues due to stability of the Monte Carlo methods and the
+                steepnes of the expected energy curves. If our guess of the
+                location of the minimum is far off we can expect the method to
+                diverge as the gradient will become huge. This will yield a
+                new value for $\vfg{\alpha}$ in
+                \autoref{eq:parameter_difference} which will be further away and
+                might result in a local minimum ``trapping'' the method. To
+                prevent this from happening we have added a limit to the allowed
+                change in $\vfg{\alpha}$. If the change in $\vfg{\alpha}$ is
+                larger than a set order of magnitude we keep dividing
+                $\vfg{\alpha}$ with this number until this criteria is
+                fulfilled.
 

### 2018-04-03 project_1/article/article.tex
Insignificant changes
Theory revision.

--- 2018-04-02 project_1/article/article.tex
@@
     allowed to interact. Moreover, we introduce the concept of \emph{local energy} and
-    then quantum \emph{drift force}. Second, we outline in more detailed algebra the
+    then quantum \emph{drift force}. Here we also outline in more detailed algebra the
     nature of a non-interaction system. For this system we are able to derive an exact 
-    variational energy to which we can compare the results from our simulations. Third,
-    we give a description of the particle-particle interaction implemented as if each 
-    particle is a hard sphere. Fourth, we provide a thorough description of algorithms and
+    variational energy to which we can compare the results from our simulations. Naturally,
+    we furthermore give a description of the particle-particle interaction implemented as if each 
+    particle is a hard sphere. Second, we provide a thorough description of algorithms and
     statistical methods employed in this study. Here we introduce the Metropolis-Hastings
@@
    	As an automated way of finding the optimal variational parameter, we sketch out the
-   	method of \emph{gradient descent}. Fifth, we put the machinere described in the 
-   	previous methods to work. We test against the actual values in a non-interacting system
+   	method of \emph{gradient descent}. Third, we put the machinery described in the 
+   	previous sections to work. We test against the actual values in a non-interacting system
    	before analysing the more interesting perturbed-trap system with interacting particles. 
@@
     where we can choose between a spherical (S) or an elliptical (E) harmonic
-    trap. The two-body Hamiltonian of the system is given by
+    trap, where the oscillatory frequency is different in $z$-direction so that
+    $\omega_z \neq \omega$. 
+    The two-body Hamiltonian of the system is given by
     \begin{align}
@@
     \subsection{The drift force}
-        A disadvantage in the use of the brute-force Metropolis algorithm is
+        A disadvantage in the use of the brute-force Metropolis-Hastings algorithm is
         that we might be spending much computational resources in an
         uninteresting part of configuration space. To make smarter moves we will
-        use the Metropolis-Hastings algorithm (which will be discussed in due
-        time). This algorithm is dependent on the drift force of the system.
+        use the Metropolis-Hastings algorithm with the additions of 
+        \emph{Importance Sampling}  (which will be discussed in due
+        time). This alteration of the algorithm is dependent on the 
+        \emph{drift force} of the system.
         \begin{align}
@@
 
-\section{Non-interacting harmonic oscillators}
+\subsection{Non-interacting harmonic oscillators}
     We start by looking at a simple system of non-interacting harmonic
-    oscillators. That is, where $a = 0$ and $\beta = 1$. We thus get the trial
-    wavefunction
+    oscillators. That is, where $a = 0$ and $\beta = 1$, i.e. there is no hard-shell
+    interaction between particles and the particles are confined in a spherical trap.
+    We thus get the trial  wavefunction
     \begin{align}
@@
     \end{align}
-    Using the Laplcian we can compute the kinteic term in the expression for the
+    Using the Laplacian we can compute the kinetic term in the expression for the
     local energy. We get
@@
     be positive) we will find a stable value which turns out to be the exact
-    energy minimum. This happens as the entire sum over all the random walkers
+    energy minimum. This happens as the sum over all particles
     disappears.
 
-    \subsection{Exact variational energy}
+    \subsubsection{Exact variational energy}
         As the system is non-interacting and consisting of Gaussians we can
@@
 
-\section{Interacting hard sphere bosons}
+\subsection{Interacting hard sphere bosons}
     Moving to the full system allowing $\beta$ to vary and setting $a
@@
     \end{align}
-    To go from here we have to find the gradient and the Laplacian of
+    From here, the next step is to find the gradient and the Laplacian of
     the single particle functions, $\phi_k$, and the interaction
     functions $u_{km}$. For the single particle functions we use
-    Cartesian coordinates when finding the derivatives whereas we for
-    the interaction functions will use spherical coordinates and do a
+    Cartesian coordinates when finding the derivatives while for
+    the interaction functions we use spherical coordinates and do a
     variable substitution. Beginning with the gradient of the single
@@
     The local energy and the drift force can now be found by combining these
-    expressions. For brevity, we will not write out the explicit expressions as
+    expressions. We will not write out the explicit expressions as
     these will be called by separated functions in our programs.
 
-    \subsection{Scaling the system}
+    \subsubsection{Scaling the system}
         We now introduce a scaled distance $\vf{r}' = \vf{r}/a_{\text{ho}}$
@@
 \section{Algorithms}
-    In the project we rely on a Monte Carlo approach of random sampling to
+    In this project we rely on a Monte Carlo approach of random sampling to
     obtain numerical results.  We simulate random walks over a volume in order
@@
 
+		\subsubsection{One-Body Densities}
+
 \section{Discussion}

### 2018-04-03 project_1/article/article.tex
Added Oyvind's stuff back.

@@
             that some analytic \footnote{We use the term analytic quite
-            recklessly.} function $F(\alpha)$ decreases fastest in the direction
-            of $-\nabla_\alpha F(\alpha)$. This means that we could eventually
+            recklessly.} function $\vf{F}(\alpha)$ decreases fastest in the direction
+            of $-\nabla_\alpha \cdot \vf{F}(\alpha)$. This means that we could eventually
             get to an ensuing minimum by trailing along the pathway of the
@@
             $\alpha$.  The expression for the derivative is,
-            \begin{equation}
+            \begin{align}
                 \nabla_{\alpha}\expval{E_L}
@@
                     \expval{E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
-                    - \expval{\frac{1}{\Psi_T} \nabla_{\alpha}\Psi_T}\expval{E_L}
+                    - \expval{\frac{1}{\Psi_T} \nabla_{\alpha}\Psi_T}
+                    \expval{E_L}
                 \right).
                 \label{eq:variational_parameter_gradient}
-            \end{equation}
-
-            %TODO: INSERT DERIVATION OF LOCAL ENERGY??
+            \end{align}
+
+            To evaluate this expression we sample the local energy, the
+            parameter gradient of the wavefunction, and the combined expression
+             of these at each Monte Carlo cycle.
+
+            \subsubsection{Stability of the algorithm}
+                When employing the method of gradient descent to locate a
+                minimum value for the variational parameters we expect some
+                issues due to stability of the Monte Carlo methods and the
+                steepness of the expected energy curves. If our guess of the
+                location of the minimum is far off we can expect the method to
+                diverge as the gradient will become huge. This will yield a
+                new value for $\vfg{\alpha}$ in
+                \autoref{eq:parameter_difference} which will be further away and
+                might result in a local minimum ``trapping'' the method. To
+                prevent this from happening we have added a limit to the allowed
+                change in $\vfg{\alpha}$. If the change in $\vfg{\alpha}$ is
+                larger than a set order of magnitude we keep dividing
+                $\vfg{\alpha}$ with this number until this criteria is
+                fulfilled.
 

### 2018-04-03 project_1/vmc/vmc/interface.pyx
Add grid distribution of particles

@@ -109,30 +109,50 @@ cdef class PyInteractingEllipticalGaussian(PyWavefunction):
         self.radius = radius
         super().__init__(num_particles, num_dimensions, num_parameters, spread)
 
-        self._remove_overlap()
+        self.redistribute()
 
         self.wavefunction = new InteractingEllipticalGaussian(
                 num_particles, num_dimensions, mass, omega, beta, radius,
                 &self.parameters[0], &self.particles[0, 0])
 
-    def _remove_overlap(self):
-        cdef unsigned int p_i, i, p_j, j
+    def redistribute(self, double spread=-1):
+        if spread < 0:
+            spread = self.spread
+
+        assert spread/self.num_particles > (2*self.radius), \
+                "Distance between each particle must be larger than the " \
+                + "diameter"
+
+        cdef double start, end, step_length, center
+        cdef np.ndarray[double, ndim=1, mode="c"] steps
+        cdef list indices
+        cdef unsigned int p_i, i, carry
+
+        start = -spread
+        end = spread
+
+        step_length = (end - start)/(self.num_particles - 1)
+        center = step_length/2.0
+
+        steps = np.linspace(
+                -spread + center, spread - center,
+                int(np.ceil(self.num_particles**(1.0/self.num_dimensions))))
+
+        indices = [0]*self.num_dimensions
 
         for p_i in range(self.num_particles):
-            for p_j in range(self.num_particles):
-                if p_i == p_j:
-                    continue
+            for i in range(self.num_dimensions):
+                self.particles[p_i, i] = steps[indices[i]]
 
-                for i in range(self.num_dimensions):
+            carry = 1
+            for i in range(self.num_dimensions):
+                indices[i] += carry
+                if indices[i] >= steps.size:
+                    indices[i] = 0
+                    carry = 1
+                else:
+                    carry = 0
 
-                    while abs(self.particles[p_i, i] - self.particles[p_j, i]) \
-                            <= self.radius:
-                        self.particles[p_j, i] = \
-                                self.spread*(2.0*np.random.random() - 1.0)
-
-    def redistribute(self, double spread=-1):
-        super().redistribute(spread=spread)
-        self._remove_overlap()
 
 cdef class PyHamiltonian:
     cdef Hamiltonian *hamiltonian

### 2018-04-03 project_1/article/article.tex
Add table and plot of one-body densities

@@
                 \alpha = 0.49744 \pm 0.00002.
+                \label{eq:optimal_interacting_alpha}
             \end{align}
@@
         \subsubsection{One-Body Densities}
+            We now look at one-body densities for the value of $\alpha$ found in
+            \autoref{eq:optimal_interacting_alpha}.
+
+            \begin{table}
+                \caption{Comparison of the elliptical system with an elliptical
+                gaussian trial wavefunction with and without the Jastrow
+                factor.}
+                \centering
+                \begin{ruledtabular}
+                    \begin{tabular}{r|rr|rr}
+                        & \multicolumn{2}{c|}{Interacting}
+                        & \multicolumn{2}{c}{Non-interacting} \\
+                        \hline
+                        $N$ & $\expval{E}$ & $\sigma_b$ & $\expval{E}$
+                        & $\sigma_b$ \\
+                        \hline
+                        10 & 24.3986 & 0.0003 & 24.1418 & 0.0002 \\
+                        50 & 127.2852 & 0.0077 & 120.7129 & 0.0012 \\
+                        100 & 266.2904 & 0.0297 & 241.4248 & 0.0025
+                    \end{tabular}
+                \end{ruledtabular}
+                \label{tab:10_optimal}
+            \end{table}
+
+            \begin{figure}
+                \includegraphics[width=244px]
+                {../data/figures/problem_g_100N.pdf}
+                \caption{In this figure the interacting and non-interacting
+                densities correspond to the elliptical potential with and
+                without the Jastrow factor respectively. The ``ideal'' system is
+                the spherical potential with the non-interacting Gaussian trial
+                wavefunction.}
+                \label{fig:one_body_density_100}
+            \end{figure}
 
@@
     \subsection{Validity of Results}
-		The non-interacting tests compared against the analytic "solution" showed
-		to a very high degree the validity of the method. One can see this quite clearly 
-		from \autoref{fig:initial_problem_b}. Because we get results from simulations
-		that correspond to the analytic results we can be fairly certain that the method
-		can be expanded upon to include attributes outside the realm problems that are 
-		analytically solvable. The systems with elliptic traps and interacting particles are
-		such systems.
+        The non-interacting tests compared against the analytic "solution"
+        showed to a very high degree the validity of the method. One can see
+        this quite clearly from \autoref{fig:initial_problem_b}. Because we get
+        results from simulations that correspond to the analytic results we can
+        be fairly certain that the method can be expanded upon to include
+        attributes outside the realm problems that are analytically solvable.
+        The systems with elliptic traps and interacting particles are such
+        systems.
 

### 2018-04-03 project_1/article/article.tex
Add review of introduction
        This includes the first part of the theory section.

@@
 \documentclass[
-    a4paper, aps, twocolumn, floatfix, superscriptaddress, nofootinbib]{revtex4-1}
+    a4paper, aps, twocolumn, floatfix, superscriptaddress,
+    nofootinbib]{revtex4-1}
 
@@
 \begin{abstract}
-    Something very abstract and clever should go here.
+    In this project we use Quantum Variational Monte Carlo on bosonic systems.
     \begin{figure}[h!]
@@
 \section{Introduction}
-    We will in this project study the Variational Monte Carlo (VMC) method, and
-    use it to evaluate the ground state energy of a trapped, hard sphere Bose
-    gas. \cite{dubois2001bose} \cite{nilsen2005vortices}
+    We will in this project study the Variational Monte Carlo method, and use it
+    to evaluate the ground state energy of a trapped, hard sphere Bose gas.
+    \cite{dubois2001bose} \cite{nilsen2005vortices}
 
     By building up an increasingly intricate system we eloquently illustrate the
-    strength of the VMC methods. Any Monte Carlo method draws on the fact that
-    the computer will not complain when being tasked to perform a very similar
-    task up to several million times in rapid succession. In general Monte Carlo
-    methods is in essence a series of guesses where the current guess is only a minor
-    alteration of the one prior to it. The trick lies in only accepting the subsequent
-    guesses that improves upon our configuration up to a certain likelihood.
-
-    The method we employed is called a \emph{Variational} Monte Carlo method,
-    because we make an educated guess at a wavefunction which we allow to contain
-    certain \emph{variational parameters}. We shall see that the system converges to a proper
-    \footnote{Global, as far as we can tell.} minimum for only a certain set of parameters.
-
-    First, we start by explaining our theoretical quantum system of a bosonic gas trapped
-    in a potential well. Here we also outline the principle under which particles in the gas are
-    allowed to interact. Moreover, we introduce the concept of \emph{local energy} and
-    then quantum \emph{drift force}. Here we also outline in more detailed algebra the
-    nature of a non-interaction system. For this system we are able to derive an exact
-    variational energy to which we can compare the results from our simulations. Naturally,
-    we furthermore give a description of the particle-particle interaction implemented as if each
-    particle is a hard sphere. Second, we provide a thorough description of algorithms and
-    statistical methods employed in this study. Here we introduce the Metropolis-Hastings
-    algorithm. This algorithm can by a deft trick be expanded with the method of
-    \emph{importance sampling} which make the way in which we pick the next step in
-       the Monte-Carlo cycling much more intelligent. The main point of note with regards to
-       the statistical analysis is that the data generated by the Monte Carlo cycling is
-       autocorrelated and computation of variance must be done with the \emph{blocking method}.
-       As an automated way of finding the optimal variational parameter, we sketch out the
-       method of \emph{gradient descent}. Third, we put the machinery described in the
-       previous sections to work. We test against the actual values in a non-interacting system
-       before analysing the more interesting perturbed-trap system with interacting particles.
-       Lastly, we highlight the more interesting aspects of this study in a discussion of the results
-       before ending on some summary remarks.
+    strength of the Variational Monte Carlo methods. Any Monte Carlo method
+    draws on the fact that the computer will not complain when being tasked to
+    perform a very similar task up to several million times in rapid succession.
+    Monte Carlo methods is in essence a series of guesses where the current
+    guess is only a minor alteration of the one prior to it. The trick lies in
+    only accepting the subsequent guesses that improves upon our configuration
+    up to a certain likelihood.
+
+    The method we have employed is called a \emph{Variational} Monte Carlo
+    method, because we make an educated guess at a trial wavefunction which we
+    allow certain degrees of freedom called \emph{variational parameters}. We
+    shall see that the system converges to a proper \footnote{Global, as far as
+    we can tell.} minimum for only a certain set of parameters.
+
+    First, we start by explaining our theoretical quantum system of a bosonic
+    gas trapped in a potential well. Here we also outline the principle under
+    which particles in the gas are allowed to interact. Moreover, we introduce
+    the \emph{local energy} and the quantum \emph{drift force}. Here we also
+    outline in more detailed algebra the nature of a non-interacting system. For
+    this system we are able to derive an exact variational energy to which we
+    can compare the results from our simulations. We furthermore give a
+    description of the particle-particle interaction implemented as if each
+    particle is a hard sphere. Second, we provide a thorough description of
+    algorithms and statistical methods employed in this study. Here we introduce
+    the Metropolis-Hastings algorithm. This algorithm can by a deft trick be
+    expanded with the method of \emph{importance sampling} yielding a more
+    intelligent choice for the next step in the Monte Carlo steps.
+
+    The main point of note with regards to the statistical analysis is that the
+    data generated by the Monte Carlo cycling is autocorrelated and computation
+    of variance must be done with the \emph{blocking method}.  As an automated
+    way of finding the optimal variational parameter, we sketch out the method
+    of \emph{gradient descent}.  Third, we put the machinery described in the
+    previous sections to work. We test against the actual values in a
+    non-interacting system before analysing the more interesting perturbed-trap
+    system with interacting particles.  Lastly, we highlight the more
+    interesting aspects of this study in a discussion of the results before
+    ending on some summary remarks.
 
@@
     \begin{align}
-        V_{\text{ext}}(\vf{r})
+        v(\vf{r})
         &=
@@
     where we can choose between a spherical (S) or an elliptical (E) harmonic
-    trap, where the oscillatory frequency is different in $z$-direction so that
-    $\omega_z \neq \omega$.
-    The two-body Hamiltonian of the system is given by
+    oscillator trap.  In the latter potential the oscillatory frequency is
+    different in the $z$-direction, i.e., $\omega_z \neq \omega$.  The full
+    Hamiltonian of the system is given by
     \begin{align}
@@
     \end{align}
-    where the single particle one body operator, $h$, is given by
-    \begin{align}
-        h(\vf{r}_i) = -\frac{\hbar^2}{2m}_i^2
-        + V_{\text{ext}}(\vf{r}_i),
-    \end{align}
-    (we assume equal mass)
-    and the two-body interaction operator, $w$, is
+    where the single particle one-body operator, $h(\vf{r}_i)$, is given by
+    \begin{align}
+        h(\vf{r}_i) = -\frac{\hbar^2}{2m}\nabla_i^2
+        + v(\vf{r}_i),
+    \end{align}
+    for equal masses, and the two-body interaction operator, $w(\vf{r}_i,
+    \vf{r}_j)$, is
     \begin{align}
@@
         \end{cases}
+        \label{eq:two-body_interaction}
     \end{align}
@@
     \end{align}
-    where $\alpha$ and $\beta$ are variational parameters and
+    where $\alpha$ is a variational parameter and
     \begin{align}
@@
     \end{align}
-    and $\Phi_T(\vf{r})$ the \textit{Slater permanent} consisting of the $N$
+    and $\Phi_T(\vf{r})$ is the \textit{Slater permanent} consisting of the $N$
     first single particle wavefunctions, and the correlation wavefunction, $f$,
@@
         \end{cases}
+        \label{eq:correlation_wavefunction}
     \end{align}
     We will for brevity use the notation $\phi(\vf{r}_i) = \phi_i$ and $r_{jk} =
-    |\vf{r}_j - \vf{r}_k|$.
+    |\vf{r}_j - \vf{r}_k|$. In practice \autoref{eq:two-body_interaction} is
+    unnecessary as the situtation where $r_{jk} \leq a$ automatically yields a
+    probability density of zero from \autoref{eq:correlation_wavefunction} thus
+    rejecting the state completely.
 

### 2018-04-03 project_1/article/article.tex
Review local energy stub

@@
     \end{align}
-    and $\Phi_T(\vf{r})$ is the \textit{Slater permanent} consisting of the $N$
+    and $\Phi_T(\vf{r})$ is the \emph{Slater permanent} consisting of the $N$
     first single particle wavefunctions, and the correlation wavefunction, $f$,
@@
         in order to move towards regions in configuration space with
-        ``sensible'' values. We define the \textit{local energy}, $E_L{\vf{r}}$,
+        ``sensible'' values. We define the \emph{local energy}, $E_L(\vf{r})$,
         by
@@
         function, the less variation in $E_L$ as a function of $\vf{r}$ we get.
-        One of the most computationally intensive parts of the VMC algorithm
-        will be to compute $E_L$. We therefore find an analytical expression for
-        $E_L$ in terms of the trial wavefunction.
+        This leads us to searching for values of $\vf{r}$ yielding the lowest
+        variance in $E_L$.
 
@@
 
+        One of the most computationally intensive parts of the Variational Monte
+        Carlo algorithm will be to compute $E_L$. We will therefore find an
+        analytical expression for $E_L$ in terms of the trial wavefunctions and
+        the system we are exploring.
+
     \subsection{The drift force}
-        A disadvantage in the use of the brute-force Metropolis-Hastings algorithm is
-        that we might be spending much computational resources in an
-        uninteresting part of configuration space. To make smarter moves we will
-        use the Metropolis-Hastings algorithm with the additions of
-        \emph{Importance Sampling}  (which will be discussed in due
-        time). This alteration of the algorithm is dependent on the
-        \emph{drift force} of the system.
+        A disadvantage in the use of the brute force Metropolis-Hastings
+        algorithm is that we might be spending much computational resources in
+        an uninteresting part of configuration space. To make smarter moves we
+        will use the Metropolis-Hastings algorithm with the additions of
+        \emph{Importance Sampling}  (which will be discussed in due time). This
+        alteration of the algorithm is dependent on the \emph{drift force} of
+        the system.
         \begin{align}
@@
     \autoref{eq:initial_trial_wavefunction} and $J(\vf{r})$ is the
-    \textit{Jastrow factor} given by
+    \emph{Jastrow factor} given by
     \begin{align}
@@
     Note that the subscripts on the unit vectors $\vf{e}_i$ are
-    \textit{not} the same as the subscripts used for its components. The
+    \emph{not} the same as the subscripts used for its components. The
     Laplacian yields
@@
         We will start without importance sampling, that is by employing the
-        simpler and more naïve "brute force" Monte Carlo sampling. Moreover, we
+        simpler and more naïve ``brute force'' Monte Carlo sampling. Moreover, we
         will compute the value for the local energy analytically as well as with
@@
         and in the expression for the exact energy \autoref{eq:exact_energy}.
-        This, however, is \textit{not} the case in the interacting system due to
+        This, however, is \emph{not} the case in the interacting system due to
         the Jastrow factor. Looking at \autoref{tab:10_interacting} and

### 2018-04-03 project_1/article/article.tex
Review drift force stub

@@
         alteration of the algorithm is dependent on the \emph{drift force} of
-        the system.
+        the system\footnote{Also known as the \emph{quantum force}}.
         \begin{align}
@@
         better choice of movements. We will mainly be interested in the drift
-        force of a single particle $k$.
+        force of a single particle $k$ as the Monte Carlo sampling involves
+        choosing a single particle at a time.
 

### 2018-04-03 project_1/article/article.tex
Add review of the simple system

@@
     We start by looking at a simple system of non-interacting harmonic
-    oscillators. That is, where $a = 0$ and $\beta = 1$, i.e. there is no hard-shell
-    interaction between particles and the particles are confined in a spherical trap.
-    We thus get the trial  wavefunction
+    oscillators, where $a = 0$ and $\beta = 1$. This means that there is no
+    hard-shell interaction between the particles and that we use a spherical
+    potential trap.  We thus get the trial  wavefunction
     \begin{align}
@@
         = \Phi_T(\vf{r})
-        = \prod_{i = 1}^N \exp\bigl[
+        = \prod_{i = 1}^N \exp\para{
             -\alpha |\vf{r}_i|^2
-        \bigr],
+        },
     \end{align}
@@
     \end{align}
-    In natural units, with $\hbar = c = m = 1$, this reduces to
+    In natural units, $\hbar = c = 1$, and with the mass set to unity, $m = 1$,
+    this reduces to
     \begin{align}
@@
     \end{align}
-    It is worth noting that for $\alpha = \half\omega$ ($\alpha$ is required to
-    be positive) we will find a stable value which turns out to be the exact
-    energy minimum. This happens as the sum over all particles
-    disappears.
+    It is worth noting that for $\alpha = \omega/2$\footnote{Note that $\alpha$
+    is required to be positive.} we get a stable value for the local energy.
+    \begin{align}
+        E_L(\vf{r})
+        &\equiv
+        E_L
+        =
+        \frac{\omega d N}{2},
+        \label{eq:minimum_energy_simple}
+    \end{align}
+    which is the stable minimum for $\alpha = \omega/2$.
 
@@
             &=
-            \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}.
-        \end{align}
-
-        % TODO: Add the computation of this expression
-        The final result for the energy is
-        \begin{align}
-            E(\alpha)
+            \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
+            \\
             &=
@@
         \end{align}
-        By minimizing this expression, i.e., finding the derivative of the
-        energy with respect to $\alpha$ and equating this to zero, yields the
-        expected minimum of variational energy to be
+        By minimizing this expression with respect to $\alpha$ yields
         \begin{align}
@@
         \end{align}
-        which in natural units reduces to $\alpha_0 = \half\omega$. The energy at
+        which in natural units reduces to $\alpha_0 = \omega/2$. The energy at
         this value of $\alpha$ (in natural units) is then
@@
             &=
-            \frac{\omega dN}{2}.
-        \end{align}
+            \frac{\omega dN}{2},
+        \end{align}
+        in perfect accordance with \autoref{eq:minimum_energy_simple} which
+        yields high hopes for the latter equation and the expected minimum.
 

### 2018-04-03 project_1/article/article.tex
Review interacting subsection

@@
         \end{align}
-    To find the drift force and the local energy we have to compute the gradient
-    and the Laplacian of the trial wavefunction. The gradient is given by
+        To find the drift force and the local energy we have to compute the
+        gradient and the Laplacian of the trial wavefunction. The gradient is
+        given by
         \begin{align}
@@
         \end{align}
-    where $d$ is the dimensionality of the problem determined by
-    $\vf{r}_k \in \mathbb{R}^d$. We can thus use the gradient to find an
-    expression for the drift force for particle $k$.
+        where $d$ is the dimensionality of the problem determined by $\vf{r}_k
+        \in \mathbb{R}^d$. We can thus use the gradient to find an expression
+        for the drift force for particle $k$.
         \begin{align}
@@
         \end{align}
-    Using the Laplacian we can compute the kinetic term in the expression for the
-    local energy. We get
+        Using the Laplacian we can compute the kinetic term in the expression
+        for the local energy. We get
         \begin{align}
@@
         \end{align}
-    In natural units, $\hbar = c = 1$, and with the mass set to unity, $m = 1$,
-    this reduces to
+        In natural units, $\hbar = c = 1$, and with the mass set to unity, $m =
+        1$, this reduces to
         \begin{align}
@@
         \end{align}
-    It is worth noting that for $\alpha = \omega/2$\footnote{Note that $\alpha$
-    is required to be positive.} we get a stable value for the local energy.
+        It is worth noting that for $\alpha = \omega/2$\footnote{Note that
+        $\alpha$ is required to be positive.} we get a stable value for the
+        local energy.
         \begin{align}
@@
             \end{align}
-        which in natural units reduces to $\alpha_0 = \omega/2$. The energy at
-        this value of $\alpha$ (in natural units) is then
+            which in natural units reduces to $\alpha_0 = \omega/2$. The energy
+            at this value of $\alpha$ (in natural units) is then
             \begin{align}
@@
     \subsection{Interacting hard sphere bosons}
-    Moving to the full system allowing $\beta$ to vary and setting $a
+        Moving to the full elliptical system for $\beta \neq 1$ and setting $a
         \neq 0$ we can write the trial wavefunction as
@@
         \end{align}
-    where $\ket{\Phi_T}$ is the same Slater permanent as in
+        where $\Phi_T(\vf{r})$ is the same Slater permanent as in
         \autoref{eq:initial_trial_wavefunction} and $J(\vf{r})$ is the
@@
             &=
-        \exp\Biggl(
+            \exp\para{
                 \sum_{j < l}^N u(r_{jl})
-        \Biggr),
-    \end{align}
-    where $r_{jk} = |\vf{r}_j - \vf{r}_k|$ and
-    \begin{align}
-        u(r_{jk}) = \ln\bigl[f(a, \vf{r}_j, \vf{r}_k)\bigr].
-    \end{align}
-    To further shorten the notation we will use $u_{jk} = u(r_{jk})$. Computing
-    the gradient of the wavefunction we get
-    \begin{align}
-        \nabla_k\Psi_T(\vf{r})
-        &=
-        \Bigl[
-            \nabla_k
-            \Phi_T(\vf{r})
-        \Bigr]
-        J(\vf{r})
-        + \Phi_T(\vf{r})
-        \nabla_k J(\vf{r}).
-    \end{align}
-    The gradient of the Slater permament for particle $k$ is given by
-    \begin{align}
-        \nabla_k
-        \Phi_T(\vf{r})
-        &=
-        \nabla_k\phi_k
-        \prod_{i \neq k}^N\phi_i
-        = \frac{\nabla_k\phi_k}{\phi_k}
-        \Phi_T(\vf{r}).
-    \end{align}
-    The gradient of the Jastrow factor is given by
-    \begin{align}
-        \nabla_k J(\vf{r})
-        &=
-        J(\vf{r})
-        \nabla_k\sum_{m < n}^N u_{mn} \\
-        &= J(\vf{r})
-        \Biggl(
-            \sum_{m = 1}^{k - 1}\nabla_k u_{mk}
-            \sum_{n = k + 1}^N\nabla_k u_{kn}
-        \Biggr)
-        \\
-        &=
-        J(\vf{r})
-        \sum_{m \neq k}^N\nabla_k u_{km},
-    \end{align}
-    where the gradient of the interaction term splits the antisymmetric
-    sum into two parts. As $r_{ij} = r_{ji}$ we can combine these sums
-    into a single sum. This in total yields the gradient
-    \begin{align}
-        \nabla_k\Psi_T(\vf{r})
-        &=
-        \Biggl(
+            },
+        \end{align}
+        where
+        \begin{align}
+            u(r_{jk})
+            = \ln\bigl[f(a, \vf{r}_j, \vf{r}_k)\bigr]
+            \equiv u_{jk}
+        \end{align}
+        The drift force for particle $k$ for this system becomes
+        \begin{align}
+            \vf{F}_k(\vf{r})
+            &=
+            2\Biggl(
                 \frac{\nabla_k\phi_k}{\phi_k}
@@
                 \nabla_k u_{km}
-        \Biggr)
-        \Psi_T(\vf{r}).
-        \label{eq:gradient_full_wavefunction}
-    \end{align}
-
-    The Laplcian of the trial wavefunction is found by finding the divergence of
-    \autoref{eq:gradient_full_wavefunction}.
-    \begin{align}
-        \nabla_k^2\Psi_T(\vf{r})
-        &=
-        \Biggl(
-            \nabla_k\Biggl[
-                \frac{\nabla_k\phi_k}{\phi_k}
-            \Biggr]
-            +
-            \sum_{m \neq k}^N \nabla_k^2 u_{km}
-        \Biggr)\Psi_T(\vf{r})
-        \\
-        &\qquad
-        +
-        \Biggl(
-            \frac{\nabla_k\phi_k}{\phi_k}
-            + \sum_{m \neq k}^N
-            \nabla_k u_{km}
-        \Biggr)^2
-        \Psi_T(\vf{r}),
-    \end{align}
-    where the squared term came from taking the gradient of the trial
-    wavefunction.  To further simplify we divide by the trial
-    wavefunction. This yields
+            \Biggr).
+        \end{align}
+
+        The Laplacian of particle $k$ divided by the trial wavefunction is
+        \begin{widetext}
             \begin{align}
@@
                 + 2\frac{\nabla_k\phi_k}{\phi_k}
-        \sum_{m \neq k}\nabla_k u_{km}
-        \nonumber \\
-        &\qquad
-        + \sum_{m\neq k}^N\nabla_k^2 u_{km}
-        + \Biggl(
-            \sum_{m \neq k}^N\nabla_k u_{km}
-        \Biggr)^2.
-    \end{align}
-    From here, the next step is to find the gradient and the Laplacian of
-    the single particle functions, $\phi_k$, and the interaction
-    functions $u_{km}$. For the single particle functions we use
-    Cartesian coordinates when finding the derivatives while for
-    the interaction functions we use spherical coordinates and do a
-    variable substitution. Beginning with the gradient of the single
-    particle functions we get
-    \begin{align}
-        \nabla_k\phi_k
-        &=
-        \nabla_k\exp\bigl[
-            -\alpha(x_k^2 + y_k^2 + \beta z_k^2)
-        \bigr] \\
-        &=
+                \sum_{m \neq k}^N
+                \frac{\Delta\vf{r}_{km}}{r_{km}}
+                \dpd[]{u_{km}}{r_{km}}
+                + \sum_{m\neq k}^N
+                \para{
+                    \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+                    + \dpd[2]{u_{km}}{r_{km}}
+                }
+                +
+                \para{
+                    \sum_{m \neq k}^N
+                    \frac{\Delta\vf{r}_{km}}{r_{km}}
+                    \dpd[]{u_{km}}{r_{km}}
+                }^2,
+            \end{align}
+        \end{widetext}
+        where $\Delta \vf{r}_{km} = \vf{r}_k - \vf{r}_m$ and $d$ is the
+        dimensionality. Evaluating the gradient and Laplacian of the single
+        particle functions yields
+        \begin{gather}
+            \frac{\nabla_k\phi_k}{\phi_k}
+            =
             -2\alpha
-        (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
-        \phi_k.
-    \end{align}
-    Note that the subscripts on the unit vectors $\vf{e}_i$ are
-    \emph{not} the same as the subscripts used for its components. The
-    Laplacian yields
-    \begin{align}
-        \nabla_k^2\phi_k
-        &=
-        \Bigl[
+            (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k),
+            \\
+            \frac{\nabla_k^2\phi_k}{\phi_k}
+            =
             -2\alpha
@@
             \bigr)
-            \nonumber \\
-            &\qquad
             + 4\alpha^2
@@
                 x_k^2 + y_k^2 + \beta^2z_k^2
-            \bigr)
-        \Bigr]
-        \phi_k,
-    \end{align}
-    with $d$ as the dimensionality of the problem.
-    In order to derive the interaction functions we have to do a
-    variable substitution using $r_{km} = |\vf{r}_k - \vf{r}_m|$. We can
-    then rewrite the $\nabla_k$-operator as
-    \begin{align}
-        \nabla_k
-        &=
-        \nabla_k
-        \dpd[]{r_{km}}{r_{km}}
-        =
-        \nabla_k r_{km} \dpd[]{}{r_{km}}
-        \\
-        &=
-        \frac{\vf{r}_k - \vf{r}_m}{r_{km}}\dpd[]{}{r_{km}}.
-    \end{align}
-    Applying this version of the $\nabla_k$-operator to $u_{km}$ yields
-    \begin{align}
-        \nabla_k u_{km}
-        &=
-        \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
-        \dpd[]{u_{km}}{r_{km}}.
-    \end{align}
-
-    \begin{widetext}
-        For the Laplacian we switch a little back and forth between the
-        two ways of representing the $\nabla_k$-operator. We thus get
-        \begin{align}
-            \nabla_k^2 u_{km}
-            &=
-            \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
-            + \Biggl[
-                \nabla_k \frac{1}{r_{km}}
-            \Biggr]
-            (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
-            + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
-            \nabla_k \dpd[]{u_{km}}{r_{km}} \\
-            &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
-            - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
-            \dpd[]{u_{km}}{r_{km}}
-            + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
-            \dpd[2]{u_{km}}{r_{km}} \\
-            &=
-            \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
-            + \dpd[2]{u_{km}}{r_{km}},
-        \end{align}
-        where $d$ is again the dimensionality of the problem. In total
-        we can state an intermediate version of the Laplacian occuring
-        in the local energy as
-        \begin{align}
-            \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
-            &=
-            \frac{\nabla_k^2\phi_k}{\phi_k}
-            + 2\frac{\nabla_k\phi_k}{\phi_k}
-            \sum_{m \neq k}^N
-            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
-            \dpd[]{u_{km}}{r_{km}}
-            + \sum_{m\neq k}^N
-            \Biggl(
-                \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
-                + \dpd[2]{u_{km}}{r_{km}}
-            \Biggr)
-            \nonumber \\
-            &\qquad
-            +
-            \sum_{m, n \neq k}^N
-            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
-            \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
-            \dpd[]{u_{km}}{r_{km}}
-            \dpd[]{u_{kn}}{r_{kn}}.
-        \end{align}
-    \end{widetext}
-    Moving on to the derivatives of the interaction terms, $u_{km}$, to
-    get an explicit expression for the Laplacian.
+            \bigr).
+        \end{gather}
+        Repeating this for the correlation functions yields
         \begin{align}
@@
         \end{align}
-    The local energy and the drift force can now be found by combining these
-    expressions. We will not write out the explicit expressions as
-    these will be called by separated functions in our programs.
+        We will avoid writing out the full expression for the local energy for
+        this system due to the size of the equation.
 
@@
             same, but using the scaled hard core sphere radius $a/a_{\text{ho}} =
-        0.0043$. For brevity we will remove the ticks on the variables.
+            0.0043$.
 
@@
 \appendix
+\section{Computing the drift force and local energy of the full system}
+    Computing the gradient of the wavefunction we get
+    \begin{align}
+        \nabla_k\Psi_T(\vf{r})
+        &=
+        \Bigl[
+            \nabla_k
+            \Phi_T(\vf{r})
+        \Bigr]
+        J(\vf{r})
+        + \Phi_T(\vf{r})
+        \nabla_k J(\vf{r}).
+    \end{align}
+    The gradient of the Slater permament for particle $k$ is given by
+    \begin{align}
+        \nabla_k
+        \Phi_T(\vf{r})
+        &=
+        \nabla_k\phi_k
+        \prod_{i \neq k}^N\phi_i
+        = \frac{\nabla_k\phi_k}{\phi_k}
+        \Phi_T(\vf{r}).
+    \end{align}
+    The gradient of the Jastrow factor is given by
+    \begin{align}
+        \nabla_k J(\vf{r})
+        &=
+        J(\vf{r})
+        \nabla_k\sum_{m < n}^N u_{mn} \\
+        &= J(\vf{r})
+        \Biggl(
+            \sum_{m = 1}^{k - 1}\nabla_k u_{mk}
+            \sum_{n = k + 1}^N\nabla_k u_{kn}
+        \Biggr)
+        \\
+        &=
+        J(\vf{r})
+        \sum_{m \neq k}^N\nabla_k u_{km},
+    \end{align}
+    where the gradient of the interaction term splits the antisymmetric
+    sum into two parts. As $r_{ij} = r_{ji}$ we can combine these sums
+    into a single sum. This in total yields the gradient
+    \begin{align}
+        \nabla_k\Psi_T(\vf{r})
+        &=
+        \Biggl(
+            \frac{\nabla_k\phi_k}{\phi_k}
+            + \sum_{m \neq k}^N
+            \nabla_k u_{km}
+        \Biggr)
+        \Psi_T(\vf{r}).
+        \label{eq:gradient_full_wavefunction}
+    \end{align}
+
+    The Laplacian of the trial wavefunction is found by finding the
+    divergence of \autoref{eq:gradient_full_wavefunction}.
+    \begin{align}
+        \nabla_k^2\Psi_T(\vf{r})
+        &=
+        \Biggl(
+            \nabla_k\Biggl[
+                \frac{\nabla_k\phi_k}{\phi_k}
+            \Biggr]
+            +
+            \sum_{m \neq k}^N \nabla_k^2 u_{km}
+        \Biggr)\Psi_T(\vf{r})
+        \\
+        &\qquad
+        +
+        \Biggl(
+            \frac{\nabla_k\phi_k}{\phi_k}
+            + \sum_{m \neq k}^N
+            \nabla_k u_{km}
+        \Biggr)^2
+        \Psi_T(\vf{r}),
+    \end{align}
+    where the squared term came from taking the gradient of the trial
+    wavefunction.  To further simplify we divide by the trial
+    wavefunction. This yields
+    \begin{align}
+        \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
+        &=
+        \frac{\nabla_k^2\phi_k}{\phi_k}
+        + 2\frac{\nabla_k\phi_k}{\phi_k}
+        \sum_{m \neq k}\nabla_k u_{km}
+        \nonumber \\
+        &\qquad
+        + \sum_{m\neq k}^N\nabla_k^2 u_{km}
+        + \Biggl(
+            \sum_{m \neq k}^N\nabla_k u_{km}
+        \Biggr)^2.
+    \end{align}
+    From here, the next step is to find the gradient and the Laplacian of
+    the single particle functions, $\phi_k$, and the interaction
+    functions $u_{km}$. For the single particle functions we use
+    Cartesian coordinates when finding the derivatives while for
+    the interaction functions we use spherical coordinates and do a
+    variable substitution. Beginning with the gradient of the single
+    particle functions we get
+    \begin{align}
+        \nabla_k\phi_k
+        &=
+        \nabla_k\exp\bigl[
+            -\alpha(x_k^2 + y_k^2 + \beta z_k^2)
+        \bigr] \\
+        &=
+        -2\alpha
+        (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
+        \phi_k.
+    \end{align}
+    Note that the subscripts on the unit vectors $\vf{e}_i$ are
+    \emph{not} the same as the subscripts used for its components. The
+    Laplacian yields
+    \begin{align}
+        \nabla_k^2\phi_k
+        &=
+        \Bigl[
+            -2\alpha
+            \big(
+                d - 1 + \beta
+            \bigr)
+            \nonumber \\
+            &\qquad
+            + 4\alpha^2
+            \bigl(
+                x_k^2 + y_k^2 + \beta^2z_k^2
+            \bigr)
+        \Bigr]
+        \phi_k,
+    \end{align}
+    with $d$ as the dimensionality of the problem.
+    In order to derive the interaction functions we have to do a
+    variable substitution using $r_{km} = |\vf{r}_k - \vf{r}_m|$. We can
+    then rewrite the $\nabla_k$-operator as
+    \begin{align}
+        \nabla_k
+        &=
+        \nabla_k
+        \dpd[]{r_{km}}{r_{km}}
+        =
+        \nabla_k r_{km} \dpd[]{}{r_{km}}
+        \\
+        &=
+        \frac{\vf{r}_k - \vf{r}_m}{r_{km}}\dpd[]{}{r_{km}}.
+    \end{align}
+    Applying this version of the $\nabla_k$-operator to $u_{km}$ yields
+    \begin{align}
+        \nabla_k u_{km}
+        &=
+        \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+        \dpd[]{u_{km}}{r_{km}}.
+    \end{align}
+
+    \begin{widetext}
+        For the Laplacian we switch a little back and forth between the
+        two ways of representing the $\nabla_k$-operator. We thus get
+        \begin{align}
+            \nabla_k^2 u_{km}
+            &=
+            \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+            + \Biggl[
+                \nabla_k \frac{1}{r_{km}}
+            \Biggr]
+            (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
+            + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+            \nabla_k \dpd[]{u_{km}}{r_{km}} \\
+            &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+            - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
+            \dpd[]{u_{km}}{r_{km}}
+            + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
+            \dpd[2]{u_{km}}{r_{km}} \\
+            &=
+            \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+            + \dpd[2]{u_{km}}{r_{km}},
+        \end{align}
+        where $d$ is again the dimensionality of the problem. In total
+        we can state an intermediate version of the Laplacian occuring
+        in the local energy as
+        \begin{align}
+            \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
+            &=
+            \frac{\nabla_k^2\phi_k}{\phi_k}
+            + 2\frac{\nabla_k\phi_k}{\phi_k}
+            \sum_{m \neq k}^N
+            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+            \dpd[]{u_{km}}{r_{km}}
+            + \sum_{m\neq k}^N
+            \Biggl(
+                \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+                + \dpd[2]{u_{km}}{r_{km}}
+            \Biggr)
+            \nonumber \\
+            &\qquad
+            +
+            \sum_{m, n \neq k}^N
+            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+            \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
+            \dpd[]{u_{km}}{r_{km}}
+            \dpd[]{u_{kn}}{r_{kn}}.
+        \end{align}
+    \end{widetext}
+    Moving on to the derivatives of the interaction terms, $u_{km}$, to
+    get an explicit expression for the Laplacian.
+    \begin{align}
+        \dpd[]{u_{km}}{r_{km}}
+        &=
+        \frac{a}{r_{km}(r_{km} - a)},
+        \\
+        \dpd[2]{u_{km}}{r_{km}}
+        &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.
+    \end{align}
+    The local energy and the drift force can now be found by combining these
+    expressions. We will not write out the explicit expressions as
+    these will be called by separated functions in our programs.
+
 \section{Brute Force Metropolis-Hastings}

### 2018-04-03 project_1/article/article.tex
Start on theory of VMC

@@
 
-\section{Theory}
+\section{Variational Monte Carlo}
+    The variational principle states that for a given \emph{trial wavefunction},
+    $\ket{\Psi_T}$, the expectation value of the Hamiltonian, $H$, will be an
+    upper bound to the ground state energy, $E_0$ of the system.
+    \begin{align}
+        E_0 \leq E
+        = \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}},
+    \end{align}
+    The complicated part of the variational principle is finding the minimum, or
+    values close to the minimum. For a select few systems we can calculate this
+    analytically, but for the interesting systems we quickly hit an intractable
+    wall. To get around this we try to calculate the energy for a given
+    Hamiltonian and a trial wavefunction for different variational parameters.
+    The variance of the variational energy is given by
+    \begin{align}
+        \sigma^2
+        &=
+        \bra{\Psi_T}H^2\ket{\Psi_T}
+        - \bra{\Psi_T}H\ket{\Psi_T}^2.
+    \end{align}
+    If we have found the ``true'' trial wavefunction $\ket{\Psi_T}$, i.e., the
+    eigenfunction to the Hamiltonian $H$, we get $\sigma^2 = 0$. We are thus
+    interested in trying to find the variational parameters in the trial
+    wavefunction that minimizes the variance $\sigma^2$.
+
+    \subsection{Monte Carlo integration}
+        Unfortunately the evaluation of the expectation value of the energy for
+        a given trial wavefunction involves a large integral.
+        \begin{align}
+            E
+            &= \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
+            =
+            \frac{\int\dd\vf{r}\Psi_T^{*}(\vf{r})H\Psi_T(\vf{r})}
+            {\int\dd\vf{r}|\Psi_T|^2},
+        \end{align}
+        where $\vf{r} = (\vf{r}_1, \vf{r}_2, \dots \vf{r}_N)$ are the
+        positions\footnote{This can include both spatial and spin parts, but we
+        will in the remainder of the project ignore spin.} of the particles
+        contained in the system.
+
+    \subsection{Local energy}
+        As the many-body wavefunction creates a very large configuration space,
+        where much of the wavefunction is small, we use the Metropolis algorithm
+        in order to move towards regions in configuration space with
+        ``sensible'' values. We define the \emph{local energy}, $E_L(\vf{r})$,
+        by
+        \begin{align}
+            E_L(\vf{r})
+            &= \frac{H\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
+        \end{align}
+        If $\Psi_T(\vf{r})$ is an exact eigenfunction of the Hamiltonian, $E_L$
+        will be constant. The closer $\Psi_T(\vf{r})$ is to the exact wave
+        function, the less variation in $E_L$ as a function of $\vf{r}$ we get.
+        This leads us to searching for values of $\vf{r}$ yielding the lowest
+        variance in $E_L$.
+
+        When we are performing the Monte Carlo sampling we are interested in the
+        expected value of the local energy as this will serve as our estimate of
+        the true energy. Mathematically this is expressed as
+        \begin{align}
+            \expval{E_L}
+            &=
+            \frac{\int\dd \vf{r}\Psi_T^{*}(\vf{r})H\Psi_T{\vf{r}}}
+            {\int\dd \vf{r}|\Psi_T(\vf{r})|^2}
+            =
+            \frac{
+                \int\dd \vf{r}|\Psi_T(\vf{r})|^2 E_L
+            }{
+                \int\dd \vf{r}|\Psi_T(\vf{r})|^2
+            },
+        \end{align}
+        where we have multiplied and divided by the trial wavefunction to get
+        the latter expression.
+
+        One of the most computationally intensive parts of the Variational Monte
+        Carlo algorithm will be to compute $E_L$. We will therefore find an
+        analytical expression for $E_L$ in terms of the trial wavefunctions and
+        the system we are exploring.
+
+    \subsection{The drift force}
+        A disadvantage in the use of the brute force Metropolis-Hastings
+        algorithm is that we might be spending much computational resources in
+        an uninteresting part of configuration space. To make smarter moves we
+        will use the Metropolis-Hastings algorithm with the additions of
+        \emph{Importance Sampling}  (which will be discussed in due time). This
+        alteration of the algorithm is dependent on the \emph{drift force} of
+        the system\footnote{Also known as the \emph{quantum force}.}.
+        \begin{align}
+            \vf{F}(\vf{r})
+            &=
+            \sum_{k = 1}^N
+            \vf{F}_k(\vf{r})
+            =
+            \sum_{k = 1}^N
+            \frac{2\nabla_k\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
+            \label{eq:drift_force}
+        \end{align}
+        Using this expression we are able to move towards parts of
+        configuration space where the gradient increases or decreases yielding a
+        better choice of movements. We will mainly be interested in the drift
+        force of a single particle $k$ as the Monte Carlo sampling involves
+        choosing a single particle at a time.
+
+\section{Systems}
     To model the trapped bosonic gas particles we use the potential
@@
 
-    \subsection{Local energy}
-        As the many-body wavefunction creates a very large configuration space,
-        where much of the wavefunction is small, we use the Metropolis algorithm
-        in order to move towards regions in configuration space with
-        ``sensible'' values. We define the \emph{local energy}, $E_L(\vf{r})$,
-        by
-        \begin{align}
-            E_L(\vf{r})
-            &= \frac{H\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
-        \end{align}
-        If $\Psi_T(\vf{r})$ is an exact eigenfunction of the Hamiltonian, $E_L$
-        will be constant. The closer $\Psi_T(\vf{r})$ is to the exact wave
-        function, the less variation in $E_L$ as a function of $\vf{r}$ we get.
-        This leads us to searching for values of $\vf{r}$ yielding the lowest
-        variance in $E_L$.
-
-        When we are performing the Monte Carlo sampling we are interested in the
-        expected value of the local energy as this will serve as our estimate of
-        the true energy. Mathematically this is expressed as
-        \begin{align}
-            \expval{E_L}
-            &=
-            \frac{\int\dd \vf{r}\Psi_T^{*}(\vf{r})H\Psi_T{\vf{r}}}
-            {\int\dd \vf{r}|\Psi_T(\vf{r})|^2}
-            =
-            \frac{
-                \int\dd \vf{r}|\Psi_T(\vf{r})|^2 E_L
-            }{
-                \int\dd \vf{r}|\Psi_T(\vf{r})|^2
-            },
-        \end{align}
-        where we have multiplied and divided by the trial wavefunction to get
-        the latter expression.
-
-        One of the most computationally intensive parts of the Variational Monte
-        Carlo algorithm will be to compute $E_L$. We will therefore find an
-        analytical expression for $E_L$ in terms of the trial wavefunctions and
-        the system we are exploring.
-
-    \subsection{The drift force}
-        A disadvantage in the use of the brute force Metropolis-Hastings
-        algorithm is that we might be spending much computational resources in
-        an uninteresting part of configuration space. To make smarter moves we
-        will use the Metropolis-Hastings algorithm with the additions of
-        \emph{Importance Sampling}  (which will be discussed in due time). This
-        alteration of the algorithm is dependent on the \emph{drift force} of
-        the system\footnote{Also known as the \emph{quantum force}}.
-        \begin{align}
-            \vf{F}(\vf{r})
-            &=
-            \sum_{k = 1}^N
-            \vf{F}_k(\vf{r})
-            =
-            \sum_{k = 1}^N
-            \frac{2\nabla_k\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
-            \label{eq:drift_force}
-        \end{align}
-        Using this expression we are able to move towards parts of
-        configuration space where the gradient increases or decreases yielding a
-        better choice of movements. We will mainly be interested in the drift
-        force of a single particle $k$ as the Monte Carlo sampling involves
-        choosing a single particle at a time.
-
     \subsection{Non-interacting harmonic oscillators}
@@
         \end{align}
-        The drift force for particle $k$ for this system becomes
+        The drift force for particle $k$ in this system becomes
         \begin{align}
@@
                 }^2,
+                \label{eq:laplacian_full_system}
             \end{align}
@@
         \end{align}
-        We will avoid writing out the full expression for the local energy for
-        this system due to the size of the equation.
+        The local energy is thus
+        \begin{align}
+            E_L(\vf{r})
+            &=
+            \sum_{k = 1}^N
+            \brac{
+                -\frac{\hbar^2}{2m}
+                \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
+                + v(\vf{r}_k)
+            },
+        \end{align}
+        where the Laplcian is given by \autoref{eq:laplacian_full_system}.  We
+        will avoid writing out the full expression for the local energy for this
+        system due to the size of the equation.
 
@@
             where we have introduced the dimensionless frequency $\lambda =
-            \omega_z/\omega$. The single particle functions also gets a scaling
+            \omega_z/\omega$. The single particle functions also get a scaling
             factor
@@
             The correlation wavefunction and the interaction potential remain the
-            same, but using the scaled hard core sphere radius $a/a_{\text{ho}} =
+            same, but using the scaled hard core radius $a/a_{\text{ho}} =
             0.0043$.

### 2018-04-03 project_1/article/article.tex
Start on theory of MC integration
One-body results

--- 2018-04-03 project_1/article/article.tex
@@
         alteration of the algorithm is dependent on the \emph{drift force} of
-        the system\footnote{Also known as the \emph{quantum force}}.
+        the system.
         \begin{align}
@@
         better choice of movements. We will mainly be interested in the drift
-        force of a single particle $k$ as the Monte Carlo sampling involves
-        choosing a single particle at a time.
+        force of a single particle $k$.
 
@@
     We start by looking at a simple system of non-interacting harmonic
-    oscillators, where $a = 0$ and $\beta = 1$. This means that there is no
-    hard-shell interaction between the particles and that we use a spherical
-    potential trap.  We thus get the trial  wavefunction
+    oscillators. That is, where $a = 0$ and $\beta = 1$, i.e. there is no hard-shell
+    interaction between particles and the particles are confined in a spherical trap.
+    We thus get the trial  wavefunction
     \begin{align}
@@
         = \Phi_T(\vf{r})
-        = \prod_{i = 1}^N \exp\para{
+        = \prod_{i = 1}^N \exp\bigl[
             -\alpha |\vf{r}_i|^2
-        },
+        \bigr],
     \end{align}
@@
     \end{align}
-    In natural units, $\hbar = c = 1$, and with the mass set to unity, $m = 1$,
-    this reduces to
+    In natural units, with $\hbar = c = m = 1$, this reduces to
     \begin{align}
@@
     \end{align}
-    It is worth noting that for $\alpha = \omega/2$\footnote{Note that $\alpha$
-    is required to be positive.} we get a stable value for the local energy.
-    \begin{align}
-        E_L(\vf{r})
-        &\equiv
-        E_L
-        =
-        \frac{\omega d N}{2},
-        \label{eq:minimum_energy_simple}
-    \end{align}
-    which is the stable minimum for $\alpha = \omega/2$.
+    It is worth noting that for $\alpha = \half\omega$ ($\alpha$ is required to
+    be positive) we will find a stable value which turns out to be the exact
+    energy minimum. This happens as the sum over all particles
+    disappears.
 
@@
             &=
-            \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
-            \\
+            \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}.
+        \end{align}
+
+        % TODO: Add the computation of this expression
+        The final result for the energy is
+        \begin{align}
+            E(\alpha)
             &=
@@
         \end{align}
-        By minimizing this expression with respect to $\alpha$ yields
+        By minimizing this expression, i.e., finding the derivative of the
+        energy with respect to $\alpha$ and equating this to zero, yields the
+        expected minimum of variational energy to be
         \begin{align}
@@
         \end{align}
-        which in natural units reduces to $\alpha_0 = \omega/2$. The energy at
+        which in natural units reduces to $\alpha_0 = \half\omega$. The energy at
         this value of $\alpha$ (in natural units) is then
@@
             &=
-            \frac{\omega dN}{2},
-        \end{align}
-        in perfect accordance with \autoref{eq:minimum_energy_simple} which
-        yields high hopes for the latter equation and the expected minimum.
+            \frac{\omega dN}{2}.
+        \end{align}
 
@@
 
+            For this optimal, interacting $\alpha$ we have computed the expected
+            energy and standard deviation for systems of $n = \{10, 50, 100\}$
+            particles. These results are shown in \autoref{tab:10_optimal}, the 
+            non-interacting case, excluding Jastrow factor is included as a 
+            comparison.
+
             \begin{figure}
@@
             We now look at one-body densities for the value of $\alpha$ found in
-            \autoref{eq:optimal_interacting_alpha}.
+            \autoref{eq:optimal_interacting_alpha}. The results for a system consisting
+            of $n=100$ particles are displayed in \autoref{fig:one_body_density_100}.
+            In this figure we see the densities for for a system confined to a 
+            perturbed elliptic trap with and without interaction. We see that for the
+            interacting case, the one-body density is skewed towards a larger 
+            $\abs{\vf{r}}$, as is expected. The `ìdeal'' density, a system with spherical
+            trap, is also included in the figure.
 

### 2018-04-03 project_1/article/article.tex
Added Oyvind's stuff back. Again.

--- 2018-04-03 project_1/article/article.tex
@@
 
+        To get around\footnote{Not exactly \emph{get around} per se, but make
+        the problem solvable at all.} the trouble of evaluating the integral we
+        employ the method of Monte Carlo integration.
+
     \subsection{Local energy}
@@
 
-    \subsection{Interacting hard sphere bosons}
-        Moving to the full elliptical system for $\beta \neq 1$ and setting $a
-        \neq 0$ we can write the trial wavefunction as
+\subsection{Non-interacting harmonic oscillators}
+    We start by looking at a simple system of non-interacting harmonic
+    oscillators. That is, where $a = 0$ and $\beta = 1$, i.e. there is no hard-shell
+    interaction between particles and the particles are confined in a spherical trap.
+    We thus get the trial  wavefunction
     \begin{align}
         \Psi_T(\vf{r})
-            &=
-            \Phi_T(\vf{r})
-            J(\vf{r}),
-        \end{align}
-        where $\Phi_T(\vf{r})$ is the same Slater permanent as in
-        \autoref{eq:initial_trial_wavefunction} and $J(\vf{r})$ is the
-        \emph{Jastrow factor} given by
-        \begin{align}
-            J(\vf{r})
-            &=
-            \exp\para{
-                \sum_{j < l}^N u(r_{jl})
-            },
-        \end{align}
-        where
-        \begin{align}
-            u(r_{jk})
-            = \ln\bigl[f(a, \vf{r}_j, \vf{r}_k)\bigr]
-            \equiv u_{jk}
-        \end{align}
+        = \Phi_T(\vf{r})
+        = \prod_{i = 1}^N \exp\bigl[
+            -\alpha |\vf{r}_i|^2
+        \bigr],
+    \end{align}
+    where $|\vf{r}_i| = r_i$. As $a = 0$ the interaction term,
+    $w(\vf{r}_i, \vf{r}_j)$, vanishes and the Hamiltonian is given by
+    (in the spherical case)
+    \begin{align}
+        H &= \sum_{i = 1}^N h(\vf{r}_i)
+        = \sum_{i = 1}^N \Biggl(
+            -\frac{\hbar^2}{2m}\nabla_i^2
+            + \half m \omega^2 r_i^2
+        \Biggr).
+    \end{align}
+    To find the drift force and the local energy we have to compute the gradient
+    and the Laplacian of the trial wavefunction. The gradient is given by
+    \begin{align}
+        \nabla_k\Psi_T(\vf{r})
+        &= -2\alpha \vf{r}_k\Psi_T(\vf{r}),
+    \end{align}
+    whereas the Laplacian yields
+    \begin{align}
+        \nabla^2_k\Psi_T(\vf{r})
+        &= \big(-2d\alpha + 4\alpha^2 r_k^2\bigr)\Psi_T(\vf{r}),
+    \end{align}
+    where $d$ is the dimensionality of the problem determined by
+    $\vf{r}_k \in \mathbb{R}^d$. We can thus use the gradient to find an
+    expression for the drift force for particle $k$.
+    \begin{align}
+        \vf{F}_k(\vf{r})
+        &= -4\alpha\vf{r}_k.
+    \end{align}
+    Using the Laplacian we can compute the kinetic term in the expression for the
+    local energy. We get
+    \begin{align}
+        E_L(\vf{r})
+        &=
+        \sum_{i = 1}^N
+        \Biggl(
+            -\frac{\hbar^2}{2m}
+            \bigl[
+                -2d\alpha + 4\alpha^2 r_i^2
+            \bigr]
+            + \half m\omega^2 r_i^2
+        \Biggr).
+    \end{align}
+    In natural units, with $\hbar = c = m = 1$, this reduces to
+    \begin{align}
+        E_L(\vf{r})
+        &=
+        \alpha dN
+        + \biggl(
+            \half\omega^2
+            - 2\alpha^2
+        \biggr)
+        \sum_{i = 1}^N r_i^2.
+        \label{eq:closed_form_natural_units_local_energy}
+    \end{align}
+    It is worth noting that for $\alpha = \half\omega$ ($\alpha$ is required to
+    be positive) we will find a stable value which turns out to be the exact
+    energy minimum. This happens as the sum over all particles
+    disappears.
+
+    \subsubsection{Exact variational energy}
+        As the system is non-interacting and consisting of Gaussians we can
+        find an expression for the exact energy as a function of the
+        variational parameter $\alpha$, i.e.,
+        \begin{align}
+            E(\alpha)
+            &=
+            \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}.
+        \end{align}
+
+        % TODO: Add the computation of this expression
+        The final result for the energy is
+        \begin{align}
+            E(\alpha)
+            &=
+            \left(
+                \frac{\hbar^2 \alpha}{2m}
+                + \frac{m\omega^2}{8\alpha}
+            \right)dN.
+            \label{eq:exact_energy}
+        \end{align}
+        By minimizing this expression, i.e., finding the derivative of the
+        energy with respect to $\alpha$ and equating this to zero, yields the
+        expected minimum of variational energy to be
+        \begin{align}
+            \dod[]{E(\alpha)}{\alpha} = 0
+            \implies
+            \alpha_0 = \frac{m\omega}{2\hbar},
+        \end{align}
+
         The drift force for particle $k$ in this system becomes
-        \begin{align}
-            \vf{F}_k(\vf{r})
-            &=
-            2\Biggl(
-                \frac{\nabla_k\phi_k}{\phi_k}
-                + \sum_{m \neq k}^N
-                \nabla_k u_{km}
-            \Biggr).
+
+        \begin{align}
+            E(\alpha_0)
+            &=
+            \frac{\omega dN}{2}.
         \end{align}
@@
         \begin{gather}
+
             \frac{\nabla_k\phi_k}{\phi_k}
-            =
+            + \sum_{m \neq k}^N
+            \nabla_k u_{km}
+        \Biggr)
+        \Psi_T(\vf{r}).
+        \label{eq:gradient_full_wavefunction}
+    \end{align}
+
+    The Laplcian of the trial wavefunction is found by finding the divergence of
+    \autoref{eq:gradient_full_wavefunction}.
+    \begin{align}
+        \nabla_k^2\Psi_T(\vf{r})
+        &=
+        \Biggl(
+            \nabla_k\Biggl[
+                \frac{\nabla_k\phi_k}{\phi_k}
+            \Biggr]
+            +
+            \sum_{m \neq k}^N \nabla_k^2 u_{km}
+        \Biggr)\Psi_T(\vf{r})
+        \\
+        &\qquad
+        +
+        \Biggl(
+            \frac{\nabla_k\phi_k}{\phi_k}
+            + \sum_{m \neq k}^N
+            \nabla_k u_{km}
+        \Biggr)^2
+        \Psi_T(\vf{r}),
+    \end{align}
+    where the squared term came from taking the gradient of the trial
+    wavefunction.  To further simplify we divide by the trial
+    wavefunction. This yields
+    \begin{align}
+        \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
+        &=
+        \frac{\nabla_k^2\phi_k}{\phi_k}
+        + 2\frac{\nabla_k\phi_k}{\phi_k}
+        \sum_{m \neq k}\nabla_k u_{km}
+        \nonumber \\
+        &\qquad
+        + \sum_{m\neq k}^N\nabla_k^2 u_{km}
+        + \Biggl(
+            \sum_{m \neq k}^N\nabla_k u_{km}
+        \Biggr)^2.
+    \end{align}
+    From here, the next step is to find the gradient and the Laplacian of
+    the single particle functions, $\phi_k$, and the interaction
+    functions $u_{km}$. For the single particle functions we use
+    Cartesian coordinates when finding the derivatives while for
+    the interaction functions we use spherical coordinates and do a
+    variable substitution. Beginning with the gradient of the single
+    particle functions we get
+    \begin{align}
+        \nabla_k\phi_k
+        &=
+        \nabla_k\exp\bigl[
+            -\alpha(x_k^2 + y_k^2 + \beta z_k^2)
+        \bigr] \\
+        &=
         -2\alpha
-            (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k),
-            \\
-            \frac{\nabla_k^2\phi_k}{\phi_k}
-            =
+        (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
+        \phi_k.
+    \end{align}
+    Note that the subscripts on the unit vectors $\vf{e}_i$ are
+    \emph{not} the same as the subscripts used for its components. The
+    Laplacian yields
+    \begin{align}
+        \nabla_k^2\phi_k
+        &=
+        \Bigl[
             -2\alpha
@@
             \bigr)
+            \nonumber \\
+            &\qquad
             + 4\alpha^2
@@
                 x_k^2 + y_k^2 + \beta^2z_k^2
-            \bigr).
-        \end{gather}
-        Repeating this for the correlation functions yields
+            \bigr)
+        \Bigr]
+        \phi_k,
+    \end{align}
+    with $d$ as the dimensionality of the problem.
+    In order to derive the interaction functions we have to do a
+    variable substitution using $r_{km} = |\vf{r}_k - \vf{r}_m|$. We can
+    then rewrite the $\nabla_k$-operator as
+    \begin{align}
+        \nabla_k
+        &=
+        \nabla_k
+        \dpd[]{r_{km}}{r_{km}}
+        =
+        \nabla_k r_{km} \dpd[]{}{r_{km}}
+        \\
+        &=
+        \frac{\vf{r}_k - \vf{r}_m}{r_{km}}\dpd[]{}{r_{km}}.
+    \end{align}
+    Applying this version of the $\nabla_k$-operator to $u_{km}$ yields
+    \begin{align}
+        \nabla_k u_{km}
+        &=
+        \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+        \dpd[]{u_{km}}{r_{km}}.
+    \end{align}
+
+    \begin{widetext}
+        For the Laplacian we switch a little back and forth between the
+        two ways of representing the $\nabla_k$-operator. We thus get
+        \begin{align}
+            \nabla_k^2 u_{km}
+            &=
+            \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+            + \Biggl[
+                \nabla_k \frac{1}{r_{km}}
+            \Biggr]
+            (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
+            + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+            \nabla_k \dpd[]{u_{km}}{r_{km}} \\
+            &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+            - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
+            \dpd[]{u_{km}}{r_{km}}
+            + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
+            \dpd[2]{u_{km}}{r_{km}} \\
+            &=
+            \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+            + \dpd[2]{u_{km}}{r_{km}},
+        \end{align}
+        where $d$ is again the dimensionality of the problem. In total
+        we can state an intermediate version of the Laplacian occuring
+        in the local energy as
+        \begin{align}
+            \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
+            &=
+            \frac{\nabla_k^2\phi_k}{\phi_k}
+            + 2\frac{\nabla_k\phi_k}{\phi_k}
+            \sum_{m \neq k}^N
+            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+            \dpd[]{u_{km}}{r_{km}}
+            + \sum_{m\neq k}^N
+            \Biggl(
+                \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
+                + \dpd[2]{u_{km}}{r_{km}}
+            \Biggr)
+            \nonumber \\
+            &\qquad
+            +
+            \sum_{m, n \neq k}^N
+            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
+            \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
+            \dpd[]{u_{km}}{r_{km}}
+            \dpd[]{u_{kn}}{r_{kn}}.
+        \end{align}
+    \end{widetext}
+    Moving on to the derivatives of the interaction terms, $u_{km}$, to
+    get an explicit expression for the Laplacian.
     \begin{align}
@@
     \end{align}
+    The local energy and the drift force can now be found by combining these
+    expressions. We will not write out the explicit expressions as
+    these will be called by separated functions in our programs.
+
+    \subsubsection{Scaling the system}
+        We now introduce a scaled distance $\vf{r}' = \vf{r}/a_{\text{ho}}$
+        \cite{dubois2001bose}, where
+        \begin{align}
+            a_{\text{ho}}
+            &=
+            \sqrt{\frac{\hbar}{m\omega}}.
+        \end{align}
+        We can then rewrite the Hamiltonian for the elliptic potential in terms
+        of this new distance. By doing a variable substitution for each
+        direction in the Laplace operator we get
+        \begin{align}
+            \nabla_k^2
+            &=
+            \frac{1}{a_{\text{ho}}^2}{\nabla_k'}^2.
+        \end{align}
+
         The local energy is thus
@@
                     + \omega_z^2 z_k^2
-                    }
                 }

### 2018-04-03 project_1/article/article.tex
Is it fixed??
Summary remarks, abstract.

@@
 
-\title{Variational Monte Carlo on bosonic systems}
+\title{Variational Monte Carlo on Bosonic Systems}
 \author{Winther-Larsen, Sebastian Gregorius}
@@
 \begin{abstract}
-    In this project we use Quantum Variational Monte Carlo on bosonic systems.
-    \begin{figure}[h!]
-        \centering
-        \includegraphics[width=0.7\textwidth]{loopholes.jpg}
-    \end{figure}
+    In this project we use Quantum Variational Monte Carlo on a toy model of a  bosonic gas  in an elliptic trap. The framework
+    we have built is first tested against analytic solutions of simple non-interacting systems. From there the system
+    is extended to include perturbed harmonic oscillator traps and particle interaction. We compute one-body
+    densities for the interacting case and discuss why this tool is important and how our toy model has similar 
+    attributes to real-world systems.
 \end{abstract}
@@
         \begin{gather}
-
             \frac{\nabla_k\phi_k}{\phi_k}
@@
         		\label{eq:gradient_full_wavefunction}
-    \end{align}
+    		\end{gather}
 
@@
                     + \omega_z^2 z_k^2
-                }
+                }}
                 \\
@@
             interacting case, the one-body density is skewed towards a larger 
-            $\abs{\vf{r}}$, as is expected. The `ìdeal'' density, a system with spherical
+            $\abs{\vf{r}}$, as is expected. The ``ideal'' density, a system with spherical
             trap, is also included in the figure.
@@
 
+        That said, the way to get the best confirmation of the results presented 
+        above is in comparison to experiment. This would be quite difficult
+        as the systems presented are very theoretical in nature and most likely
+        far from anything we would see in nature. Without being able to compare 
+        our result with anything happening in the real world this entire study would
+        have no value except for the instrumental value, being no more than
+        a theoretical exercise.  However, the one-body
+        density is a measure of the role of many-body correlations which in
+        some cases can be related to experiment. For a system of charged 
+        particles it can be used to determine a charge distribution, which in
+        turn can be linked to experiment. It would be very fruitful to simulate a
+        system of such particles that is more natural, i.e. real, but we realise the
+        purpose of this study is first and foremost to introduce variational 
+        Monte Carlo as a tool in computational physics. We assume, however,
+        that this study is easy enough to expand into a study of alkali atoms like
+        $^{87}Rb$, $^{23}Na$ or $^7Li$.
+
 	\subsection{Energy per particle}
@@
         \autoref{tab:100_interacting} we see that multiplying the expected
-        energy of theformer with $10$ yields a lower number than the latter. We
+        energy of the former with $10$ yields a lower number than the latter. We
         can also see this behavior in \autoref{fig:interacting}, where the
@@
 
+        The reason for the increase in energy per particle can be ascribed to the
+        repulsive interaction added by the Jastrow factor. The repulsive force can
+        be explain thusly; The Jastrow factor  increases the value of the 
+        wavefunctions for any given particle if that particle is some distance away
+         from any other particle. An increased value in the wavefucntion is equivalent
+         to an increase in probability amplitude, again correpsonding to an increase in
+         probability density, giving rise the the fact that particles are more likely to 
+         appear further from one another.
+
+	\subsection{Low density system}
+		It is safe to say that the effect of adding a Jastrow factor is significant, 
+		i.e. the increase in
+        	energy for the system as the interaction is switched on is much higher
+        	than the uncertainty of the energy. A simple look at \autoref{fig:interacting}
+        	will be proof enough as the error bars of the figure are not discernible. 
+        On the other hand, the energy added to the system by including interaction
+        is not very high compared to the energy of the system as a whole. For the 
+        larger of the interacting systems we have looked at, with $n=100$ particles,
+        The interaction add only a mere $9\%$ extra energy (see \autoref{tab:10_optimal}).
+        Studying \autoref{fig:one_body_density_100} will reveal the same.
+
+        The reason for this stems from the assumption which justifies that
+        adding only two-body interaction is sufficient to model all interaction to a
+        reasonable extent. This assumption is justified by the fact that alkali and
+        atomic hydrogen systems\footnote{Systems similar to our toy model}
+        are dilute. For gases of alkali atoms confined in magnetic or optical traps
+        the average distance between atoms is much larger than the range of
+        inter-atomic interaction. The probability is high that atoms very seldom
+        interact and if so, just with one other atom.
+
 \section{Summary Remarks}
+
+	In this study we have built a functional variational Monte Carlo solver framework for 
+	a gas following Bose-Einstein statistics confined in an elliptic trap. Monte Carlo 
+	methods is a broad class of computational algorithms, whereas the variational
+	Monte Carlo is a subset. Other applications of such methods within the physical
+	sciences includes, but is not limited to, quantum chromodynamics, the modeling of
+	aearodynamic forms, particle physics detector design and galaxy evolution. 
+	Needless to say, Monte Carlo method is a valuable tool for a computational physicist.
+
+	To prove the 
+	validity of the framework we tested it against an analytically solved case, an ideal
+	non-interacting system. Thereafter we have shown how our implementation can
+	handle large systems of interacting particles and perturbed traps. Even though
+	this study is somewhat theoretic in nature, we have discussed the possibility of
+	expanding our machinery to a case bearing a larger resemblance to a system
+	found in nature. To this end we have computed one-body densities for the 
+	interacting system, which can be compared to real physical properties as
+	measured in an experiment.
 

### 2018-04-04 project_1/article/article.tex
Restore old crap

@@
 
-\subsection{Non-interacting harmonic oscillators}
-    We start by looking at a simple system of non-interacting harmonic
-    oscillators. That is, where $a = 0$ and $\beta = 1$, i.e. there is no hard-shell
-    interaction between particles and the particles are confined in a spherical trap.
-    We thus get the trial  wavefunction
+    \subsection{Interacting hard sphere bosons}
+        Moving to the full elliptical system for $\beta \neq 1$ and setting $a
+        \neq 0$ we can write the trial wavefunction as
         \begin{align}
             \Psi_T(\vf{r})
-        = \Phi_T(\vf{r})
-        = \prod_{i = 1}^N \exp\bigl[
-            -\alpha |\vf{r}_i|^2
-        \bigr],
-    \end{align}
-    where $|\vf{r}_i| = r_i$. As $a = 0$ the interaction term,
-    $w(\vf{r}_i, \vf{r}_j)$, vanishes and the Hamiltonian is given by
-    (in the spherical case)
-    \begin{align}
-        H &= \sum_{i = 1}^N h(\vf{r}_i)
-        = \sum_{i = 1}^N \Biggl(
-            -\frac{\hbar^2}{2m}\nabla_i^2
-            + \half m \omega^2 r_i^2
+            &=
+            \Phi_T(\vf{r})
+            J(\vf{r}),
+        \end{align}
+        where $\Phi_T(\vf{r})$ is the same Slater permanent as in
+        \autoref{eq:initial_trial_wavefunction} and $J(\vf{r})$ is the
+        \emph{Jastrow factor} given by
+        \begin{align}
+            J(\vf{r})
+            &=
+            \exp\para{
+                \sum_{j < l}^N u(r_{jl})
+            },
+        \end{align}
+        where
+        \begin{align}
+            u(r_{jk})
+            = \ln\bigl[f(a, \vf{r}_j, \vf{r}_k)\bigr]
+            \equiv u_{jk}
+        \end{align}
+        The drift force for particle $k$ in this system becomes
+        \begin{align}
+            \vf{F}_k(\vf{r})
+            &=
+            2\Biggl(
+                \frac{\nabla_k\phi_k}{\phi_k}
+                + \sum_{m \neq k}^N
+                \nabla_k u_{km}
             \Biggr).
-    \end{align}
-    To find the drift force and the local energy we have to compute the gradient
-    and the Laplacian of the trial wavefunction. The gradient is given by
-    \begin{align}
-        \nabla_k\Psi_T(\vf{r})
-        &= -2\alpha \vf{r}_k\Psi_T(\vf{r}),
-    \end{align}
-    whereas the Laplacian yields
-    \begin{align}
-        \nabla^2_k\Psi_T(\vf{r})
-        &= \big(-2d\alpha + 4\alpha^2 r_k^2\bigr)\Psi_T(\vf{r}),
-    \end{align}
-    where $d$ is the dimensionality of the problem determined by
-    $\vf{r}_k \in \mathbb{R}^d$. We can thus use the gradient to find an
-    expression for the drift force for particle $k$.
-    \begin{align}
-        \vf{F}_k(\vf{r})
-        &= -4\alpha\vf{r}_k.
-    \end{align}
-    Using the Laplacian we can compute the kinetic term in the expression for the
-    local energy. We get
-    \begin{align}
-        E_L(\vf{r})
-        &=
-        \sum_{i = 1}^N
-        \Biggl(
-            -\frac{\hbar^2}{2m}
-            \bigl[
-                -2d\alpha + 4\alpha^2 r_i^2
-            \bigr]
-            + \half m\omega^2 r_i^2
-        \Biggr).
-    \end{align}
-    In natural units, with $\hbar = c = m = 1$, this reduces to
-    \begin{align}
-        E_L(\vf{r})
-        &=
-        \alpha dN
-        + \biggl(
-            \half\omega^2
-            - 2\alpha^2
-        \biggr)
-        \sum_{i = 1}^N r_i^2.
-        \label{eq:closed_form_natural_units_local_energy}
-    \end{align}
-    It is worth noting that for $\alpha = \half\omega$ ($\alpha$ is required to
-    be positive) we will find a stable value which turns out to be the exact
-    energy minimum. This happens as the sum over all particles
-    disappears.
-
-    \subsubsection{Exact variational energy}
-        As the system is non-interacting and consisting of Gaussians we can
-        find an expression for the exact energy as a function of the
-        variational parameter $\alpha$, i.e.,
-        \begin{align}
-            E(\alpha)
-            &=
-            \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}.
-        \end{align}
-
-        % TODO: Add the computation of this expression
-        The final result for the energy is
-        \begin{align}
-            E(\alpha)
-            &=
-            \left(
-                \frac{\hbar^2 \alpha}{2m}
-                + \frac{m\omega^2}{8\alpha}
-            \right)dN.
-            \label{eq:exact_energy}
-        \end{align}
-        By minimizing this expression, i.e., finding the derivative of the
-        energy with respect to $\alpha$ and equating this to zero, yields the
-        expected minimum of variational energy to be
-        \begin{align}
-            \dod[]{E(\alpha)}{\alpha} = 0
-            \implies
-            \alpha_0 = \frac{m\omega}{2\hbar},
-        \end{align}
-
-        The drift force for particle $k$ in this system becomes
-
-        \begin{align}
-            E(\alpha_0)
-            &=
-            \frac{\omega dN}{2}.
         \end{align}
@@
             \frac{\nabla_k\phi_k}{\phi_k}
-            + \sum_{m \neq k}^N
-            \nabla_k u_{km}
-        		\Biggr)
-       		 \Psi_T(\vf{r}).
-        		\label{eq:gradient_full_wavefunction}
-    		\end{gather}
-
-    The Laplcian of the trial wavefunction is found by finding the divergence of
-    \autoref{eq:gradient_full_wavefunction}.
-    \begin{align}
-        \nabla_k^2\Psi_T(\vf{r})
-        &=
-        \Biggl(
-            \nabla_k\Biggl[
-                \frac{\nabla_k\phi_k}{\phi_k}
-            \Biggr]
-            +
-            \sum_{m \neq k}^N \nabla_k^2 u_{km}
-        \Biggr)\Psi_T(\vf{r})
+            =
+            -2\alpha
+            (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k),
             \\
-        &\qquad
-        +
-        \Biggl(
-            \frac{\nabla_k\phi_k}{\phi_k}
-            + \sum_{m \neq k}^N
-            \nabla_k u_{km}
-        \Biggr)^2
-        \Psi_T(\vf{r}),
-    \end{align}
-    where the squared term came from taking the gradient of the trial
-    wavefunction.  To further simplify we divide by the trial
-    wavefunction. This yields
-    \begin{align}
-        \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
-        &=
             \frac{\nabla_k^2\phi_k}{\phi_k}
-        + 2\frac{\nabla_k\phi_k}{\phi_k}
-        \sum_{m \neq k}\nabla_k u_{km}
-        \nonumber \\
-        &\qquad
-        + \sum_{m\neq k}^N\nabla_k^2 u_{km}
-        + \Biggl(
-            \sum_{m \neq k}^N\nabla_k u_{km}
-        \Biggr)^2.
-    \end{align}
-    From here, the next step is to find the gradient and the Laplacian of
-    the single particle functions, $\phi_k$, and the interaction
-    functions $u_{km}$. For the single particle functions we use
-    Cartesian coordinates when finding the derivatives while for
-    the interaction functions we use spherical coordinates and do a
-    variable substitution. Beginning with the gradient of the single
-    particle functions we get
-    \begin{align}
-        \nabla_k\phi_k
-        &=
-        \nabla_k\exp\bigl[
-            -\alpha(x_k^2 + y_k^2 + \beta z_k^2)
-        \bigr] \\
-        &=
-        -2\alpha
-        (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
-        \phi_k.
-    \end{align}
-    Note that the subscripts on the unit vectors $\vf{e}_i$ are
-    \emph{not} the same as the subscripts used for its components. The
-    Laplacian yields
-    \begin{align}
-        \nabla_k^2\phi_k
-        &=
-        \Bigl[
+            =
             -2\alpha
@@
         \end{align}
-    The local energy and the drift force can now be found by combining these
-    expressions. We will not write out the explicit expressions as
-    these will be called by separated functions in our programs.
-
-    \subsubsection{Scaling the system}
-        We now introduce a scaled distance $\vf{r}' = \vf{r}/a_{\text{ho}}$
-        \cite{dubois2001bose}, where
-        \begin{align}
-            a_{\text{ho}}
-            &=
-            \sqrt{\frac{\hbar}{m\omega}}.
-        \end{align}
-        We can then rewrite the Hamiltonian for the elliptic potential in terms
-        of this new distance. By doing a variable substitution for each
-        direction in the Laplace operator we get
-        \begin{align}
-            \nabla_k^2
-            &=
-            \frac{1}{a_{\text{ho}}^2}{\nabla_k'}^2.
-        \end{align}
-
         The local energy is thus
@@
                         + \omega_z^2 z_k^2
-                }}
+                    }
+                }
                 \\

### 2018-04-04 project_1/article/article.tex
Move references before appendix
Align abstract
Start on MC integration

@@
 \begin{abstract}
-    In this project we use Quantum Variational Monte Carlo on a toy model of a  bosonic gas  in an elliptic trap. The framework
-    we have built is first tested against analytic solutions of simple non-interacting systems. From there the system
-    is extended to include perturbed harmonic oscillator traps and particle interaction. We compute one-body
-    densities for the interacting case and discuss why this tool is important and how our toy model has similar 
-    attributes to real-world systems.
+    In this project we use Quantum Variational Monte Carlo on a toy model of a
+    bosonic gas  in an elliptic trap. The framework we have built is first
+    tested against analytic solutions of simple non-interacting systems. From
+    there the system is extended to include perturbed harmonic oscillator traps
+    and particle interaction. We compute one-body densities for the interacting
+    case and discuss why this tool is important and how our toy model has
+    similar attributes to real-world systems.
 \end{abstract}
@@
 \tableofcontents
-
-%\begin{widetext}
-
-%\end{widetext}
 
@@
         To get around\footnote{Not exactly \emph{get around} per se, but make
-        the problem solvable at all.} the trouble of evaluating the integral we
-        employ the method of Monte Carlo integration.
+        the problem solvable at all.} the trouble of evaluating the
+        multidimensional-integral we employ the method of Monte Carlo
+        integration.
 
@@
 
+\bibliography{references}
+\clearpage
+
 \appendix
@@
 
-\bibliography{references}
-
 \end{document}

### 2018-04-04 project_1/article/article.tex
Add first draft of MC integration

@@
         multidimensional-integral we employ the method of Monte Carlo
-        integration.
+        integration. This consists of approximating an integral of the
+        expectation value of a certain quantitity by a sum.
+        \begin{align}
+            \expval{x}
+            &=
+            \int_{\mathbb{R}}xp(x)\dd x
+            \approx
+            \frac{1}{N}\sum_{i = 1}^N x_ip(x_i),
+        \end{align}
+        where $x$ is a random variable with a probability density function
+        $p(x)$ and $\{x_i\}$ for $i \in \{1, \dots, N\}$ is the set of drawn
+        random variables. Here $N$ is the number of Monte Carlo samples chosen
+        to approximate the integral. By the law of large numbers we expect that
+        \begin{align}
+            \lim_{N \to \infty}\frac{1}{N}\sum_{i = 1}^N x_ip(x_i)
+            = \expval{x}.
+        \end{align}
+
+        For us to be able to take Monte Carlo integration to the quantum realm
+        we have to construct a probability density function from the
+        wavefunction. Luckily the wavefunction squared is just such a quantity.
+        Given a normalized wavefunction $\ket{\psi}$ we can compute the
+        expectation value of a quantum observable $\mathcal{O}$ using Monte
+        Carlo integration.
+        \begin{align}
+            \expval{\mathcal{O}}
+            &= \bra{\psi}\mathcal{O}\ket{\psi}
+            =
+            \int\dd\vf{r}
+            \psi^{*}(\vf{r})\mathcal{O}\psi(\vf{r})
+            \\
+            &=
+            \int\dd\vf{r}
+            |\psi(\vf{r})|^2
+            \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})},
+        \end{align}
+        where we have multiplied and divided by the wavefunction. We now define
+        the probability density function, $p(\vf{r})$, as
+        \begin{align}
+            p(\vf{r}) = |\psi(\vf{r})|^2,
+        \end{align}
+        and the local quantum observable
+        \begin{align}
+            \mathcal{O}_L(\vf{r})
+            &= \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})}.
+        \end{align}
+        We see that if the chosen wavefunction is an eigenstate of the quantum
+        operator, the local quantum observable becomes a constant.  We can thus
+        approximate the expecatation value of the quantum observable using Monte
+        Carlo integration.
+        \begin{align}
+            \expval{O}
+            &=
+            \int\dd\vf{r}
+            p(\vf{r})
+            \mathcal{O}_L(\vf{r})
+            \approx
+            \frac{1}{N}\sum_{i = 1}^N
+            p(\vf{r}_i)\mathcal{O}_L(\vf{r}_i).
+        \end{align}
+        In the case of a non-normalized wavefunction we need to divide by a
+        normalization factor in the probability density function.
 
@@
             &=
-            \frac{\int\dd \vf{r}\Psi_T^{*}(\vf{r})H\Psi_T{\vf{r}}}
+            \frac{\int\dd \vf{r}\Psi_T^{*}(\vf{r})H\Psi_T(\vf{r})}
             {\int\dd \vf{r}|\Psi_T(\vf{r})|^2}
@@
             \frac{
-                \int\dd \vf{r}|\Psi_T(\vf{r})|^2 E_L
+                \int\dd \vf{r}|\Psi_T(\vf{r})|^2 E_L(\vf{r})
             }{

### 2018-04-05 project_1/article/article.tex
Add TODO
Rewrote introduction.

@@
 
-    First, we start by explaining our theoretical quantum system of a bosonic
-    gas trapped in a potential well. Here we also outline the principle under
-    which particles in the gas are allowed to interact. Moreover, we introduce
-    the \emph{local energy} and the quantum \emph{drift force}. Here we also
-    outline in more detailed algebra the nature of a non-interacting system. For
-    this system we are able to derive an exact variational energy to which we
+	First, we provide the theoretical foundation for the Variational Monte Carlo
+	method. This includes explaining how Monte Carlo integration works, and
+	introducing the important quantities \emph{local energy}, \emph{drift force}
+	and \emph{one-body densities}.
+
+    Second, we outline in more detailed algebra the nature of a systems we will
+    be studying. For
+    the non-interacting system we are able to derive an exact variational energy 
+    to which we
     can compare the results from our simulations. We furthermore give a
-    description of the particle-particle interaction implemented as if each
-    particle is a hard sphere. Second, we provide a thorough description of
+    description of an interacting system implemented as if each
+    particle is a hard sphere.
+
+    Third, we provide a thorough description of
     algorithms and statistical methods employed in this study. Here we introduce
@@
     expanded with the method of \emph{importance sampling} yielding a more
-    intelligent choice for the next step in the Monte Carlo steps.
-
+    intelligent choice for the next step in the Monte Carlo cycling.
     The main point of note with regards to the statistical analysis is that the
     data generated by the Monte Carlo cycling is autocorrelated and computation
-    of variance must be done with the \emph{blocking method}.  As an automated
+    of variance must be done with the \emph{blocking method} for time series data. 
+    As an automated
     way of finding the optimal variational parameter, we sketch out the method
-    of \emph{gradient descent}.  Third, we put the machinery described in the
+    of \emph{gradient descent}.
+
+    Fourth, we put the machinery described in the
     previous sections to work. We test against the actual values in a
     non-interacting system before analysing the more interesting perturbed-trap
-    system with interacting particles.  Lastly, we highlight the more
+    system with interacting particles.
+
+    Lastly, we highlight
     interesting aspects of this study in a discussion of the results before
@@
         choosing a single particle at a time.
+
+    \subsection{One-Body Densities}
+        % TODO: Fill in this
 

### 2018-04-05 project_1/article/article.tex
Did TODO: one-body densities

@@
     \subsection{One-Body Densities}
-        % TODO: Fill in this
+        % TODO: Fill in this part.
+        % SEB: Sure thing, buddy
+        In order to make some conclusion towards where the particles in our systems
+        are most prone to exist, we need a good way to visualise our wavefunctions.        
+        It is impossible for a human being to visualise a probability density in three dimensions, 
+        because this measure would have to be represented in four dimensions. The fact that
+        we will be studying systems of up to $n=500$ particles only adds to the headache.
+        We therefore introduce the single-particle density function, or \emph{one-body density}
+        $\rho(x, t)$. The one-body density is defined by the integral over all coordinates
+        except one,
+        \begin{align}
+        		\rho(x_1) \equiv \int \abs{\Psi(x_1, \dots, x_n)}^2 dx_2 \dots dx_n.
+        \end{align}
+        The one-body density can be interpreted as one-dimensional probability density for
+        finding a particle in $x_1$-direction. We also have the following relation,
+        \begin{align}
+        		n = \int \rho(x_1)dx_1,
+        \end{align}
+        in other words, the integral over the entire one-dimensional space in $x_1$-direction
+        gives us the total number of particles in our system. This can be a bit odd at first sight
+        as we are used to getting the number $1$ when we integrate over the entire density.
+
+        It is best to think of the one-body density as the multivariate probability density
+        for the entire system collapsed down to one dimension in a representation than looks very 
+        familiar to every physicist with a basic knowledge of quantum mechanics.
 

### 2018-04-05 project_1/article/article.tex
Fought with last figure.
Add spellchecked introduction

@@
 
-    Second, we outline in more detailed algebra the nature of a systems we will
-    be studying. For
-    the non-interacting system we are able to derive an exact variational energy 
-    to which we
-    can compare the results from our simulations. We furthermore give a
-    description of an interacting system implemented as if each
-    particle is a hard sphere.
-
-    Third, we provide a thorough description of
-    algorithms and statistical methods employed in this study. Here we introduce
-    the Metropolis-Hastings algorithm. This algorithm can by a deft trick be
-    expanded with the method of \emph{importance sampling} yielding a more
-    intelligent choice for the next step in the Monte Carlo cycling.
-    The main point of note with regards to the statistical analysis is that the
-    data generated by the Monte Carlo cycling is autocorrelated and computation
-    of variance must be done with the \emph{blocking method} for time series data. 
-    As an automated
-    way of finding the optimal variational parameter, we sketch out the method
-    of \emph{gradient descent}.
-
-    Fourth, we put the machinery described in the
-    previous sections to work. We test against the actual values in a
-    non-interacting system before analysing the more interesting perturbed-trap
-    system with interacting particles.
-
-    Lastly, we highlight
-    interesting aspects of this study in a discussion of the results before
-    ending on some summary remarks.
+    Second, we outline in more detailed algebra the nature of the systems we
+    will be studying. For the non-interacting system we are able to derive an
+    exact variational energy to which we can compare the results from our
+    simulations. We furthermore give a description of an interacting system
+    implemented as if each particle is a hard sphere.
+
+    Third, we provide a thorough description of algorithms and statistical
+    methods employed in this study. Here we introduce the Metropolis-Hastings
+    algorithm. This algorithm can by a deft trick be expanded with the method of
+    \emph{importance sampling} yielding a more intelligent choice for the next
+    step in the Monte Carlo cycling.  The main point of note with regards to the
+    statistical analysis is that the data generated by the Monte Carlo cycling
+    is autocorrelated and computation of variance must be done with the
+    \emph{blocking method} for time series data.  As an automated way of finding
+    the optimal variational parameter, we sketch out the method of
+    \emph{gradient descent}.
+
+    Fourth, we put the machinery described in the previous sections to work. We
+    test against the actual values in a non-interacting system before analyzing
+    the more interesting perturbed-trap system with interacting particles.
+
+    Lastly, we highlight interesting aspects of this study in a discussion of
+    the results before ending on some summary remarks.
 
@@
 
-\section{Brute Force Metropolis-Hastings}
-
-    \begin{figure*}
-        \makebox[\textwidth][c]{
-            \includegraphics[width=1.2\textwidth]{../data/figures/problem_b.pdf}
-        }
-        \caption{Comparison of analytic and numeric results with the exact
-        expression for the energy as a function of $\alpha$. The plots show the
-        standard deviation from the blocking method as error ticks. In the
-        figure $N$ is the number of particles and $d$ the number of dimensions.}
-        \label{fig:initial_problem_b}
-    \end{figure*}
-
 \section{Variational parameter gradient of the expectation energy}
@@
 
+\clearpage
+\onecolumngrid
+\section{Brute Force Metropolis-Hastings}
+    \begin{figure}[b]
+        \makebox[\textwidth][c]{
+            \includegraphics[width=1.2\textwidth]{../data/figures/problem_b.pdf}      
+        }
+        \caption{Comparison of analytic and numeric results with the exact
+        expression for the energy as a function of $\alpha$. The plots show the
+        standard deviation from the blocking method as error ticks. In the
+        figure $N$ is the number of particles and $d$ the number of dimensions.}
+        \label{fig:initial_problem_b}       
+		\vspace{2cm}
+    \end{figure}
+
 \end{document}

### 2018-04-05 project_1/article/article.tex
Add spellcheck to VMC intro

@@
     \end{align}
-    The complicated part of the variational principle is finding the minimum, or
-    values close to the minimum. For a select few systems we can calculate this
+    The complicated part of the variational principle is finding the trial
+    wavefunction that minimizes the energy. The strength in this method lies in
+    introducing variational parameters in the trial wavefunction. Using these as
+    degrees of freedom we can minimize the energy with respect to these in order
+    to find an infimum to the ground state energy for the given trial
+    wavefunction and Hamiltonian. For a select few systems we can calculate this
     analytically, but for the interesting systems we quickly hit an intractable
-    wall. To get around this we try to calculate the energy for a given
-    Hamiltonian and a trial wavefunction for different variational parameters.
-    The variance of the variational energy is given by
+    wall as the systems become analytically unsolvable. To get around this we
+    try to calculate the energy for a given Hamiltonian and a trial wavefunction
+    for different variational parameters numerically.
+
+    To determine if we have located a minimum we can use the variance of the
+    variational energy.
     \begin{align}
@@
         &=
+        \frac{
             \bra{\Psi_T}H^2\ket{\Psi_T}
-        - \bra{\Psi_T}H\ket{\Psi_T}^2.
-    \end{align}
-    If we have found the ``true'' trial wavefunction $\ket{\Psi_T}$, i.e., the
-    eigenfunction to the Hamiltonian $H$, we get $\sigma^2 = 0$. We are thus
-    interested in trying to find the variational parameters in the trial
-    wavefunction that minimizes the variance $\sigma^2$.
+        }{
+            \bra{\Psi_T}\ket{\Psi_T}
+        }
+        - \para{
+            \frac{
+                \bra{\Psi_T}H\ket{\Psi_T}
+            }{
+                \bra{\Psi_T}\ket{\Psi_T}
+            }
+        }^2.
+    \end{align}
+    If we have found the ``true'' trial wavefunction $\ket{\Psi_T}$ with the
+    correct variational parameters, i.e., the eigenfunction to the Hamiltonian
+    we get $\sigma^2 = 0$. We are thus interested in trying to find the
+    variational parameters in the trial wavefunction that minimizes the
+    variance.
 

### 2018-04-05 project_1/article/article.tex
Spell check section on MC integration

@@
         positions\footnote{This can include both spatial and spin parts, but we
-        will in the remainder of the project ignore spin.} of the particles
+        will in the remainder of this project ignore spin.} of the particles
         contained in the system.
 
-        To get around\footnote{Not exactly \emph{get around} per se, but make
-        the problem solvable at all.} the trouble of evaluating the
-        multidimensional-integral we employ the method of Monte Carlo
-        integration. This consists of approximating an integral of the
-        expectation value of a certain quantitity by a sum.
+        To get around\footnote{Not exactly \emph{get around} per se. We still
+        have to evaluate the integral, but we make the problem solvable at all.}
+        the trouble of evaluating the multidimensional-integral we employ the
+        method of Monte Carlo integration. This consists of approximating an
+        integral of the expectation value of a certain quantitity by a sum.
         \begin{align}
@@
         Given a normalized wavefunction $\ket{\psi}$ we can compute the
-        expectation value of a quantum observable $\mathcal{O}$ using Monte
+        expectation value of a quantum operator $\mathcal{O}$ using Monte
         Carlo integration.
@@
         \begin{align}
-            p(\vf{r}) = |\psi(\vf{r})|^2,
-        \end{align}
-        and the local quantum observable
+            p(\vf{r}) \equiv |\psi(\vf{r})|^2,
+        \end{align}
+        and the local quantum operator
         \begin{align}
             \mathcal{O}_L(\vf{r})
-            &= \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})}.
+            &\equiv \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})}.
         \end{align}
         We see that if the chosen wavefunction is an eigenstate of the quantum
-        operator, the local quantum observable becomes a constant.  We can thus
-        approximate the expecatation value of the quantum observable using Monte
+        operator, the local quantum operator becomes a constant.  We can thus
+        approximate the expecatation value of the quantum operator using Monte
         Carlo integration.

### 2018-04-05 project_1/article/article.tex
Spell check section on local energy

@@
     \subsection{Local energy}
-        As the many-body wavefunction creates a very large configuration space,
-        where much of the wavefunction is small, we use the Metropolis algorithm
-        in order to move towards regions in configuration space with
-        ``sensible'' values. We define the \emph{local energy}, $E_L(\vf{r})$,
-        by
+        As is mostly the case in physics, the quantity we are looking for is the
+        energy. We therefore define the \emph{local energy} by
         \begin{align}
@@
         \end{align}
-        If $\Psi_T(\vf{r})$ is an exact eigenfunction of the Hamiltonian, $E_L$
-        will be constant. The closer $\Psi_T(\vf{r})$ is to the exact wave
-        function, the less variation in $E_L$ as a function of $\vf{r}$ we get.
-        This leads us to searching for values of $\vf{r}$ yielding the lowest
-        variance in $E_L$.
-
-        When we are performing the Monte Carlo sampling we are interested in the
-        expected value of the local energy as this will serve as our estimate of
-        the true energy. Mathematically this is expressed as
+        The energy of the system is thus found by computing the expectation
+        value of the local energy.
         \begin{align}
             \expval{E_L}
-            &=
-            \frac{\int\dd \vf{r}\Psi_T^{*}(\vf{r})H\Psi_T(\vf{r})}
-            {\int\dd \vf{r}|\Psi_T(\vf{r})|^2}
             =
@@
                 \int\dd \vf{r}|\Psi_T(\vf{r})|^2
-            },
-        \end{align}
-        where we have multiplied and divided by the trial wavefunction to get
-        the latter expression.
-
+            }
+            \approx
+            \frac{1}{N}\sum_{i = 1}^N
+            p(\vf{r}_i)E_L(\vf{r}_i).
+        \end{align}
+        To determine if we have found a minimum we compute the variance of the
+        expectation value of the local energy.
+        \begin{align}
+            \sigma^2 = \langle E_L^2\rangle - \expval{E_L}^2.
+        \end{align}
         One of the most computationally intensive parts of the Variational Monte

### 2018-04-05 project_1/article/article.tex
Spell check section on drift force

@@
         an uninteresting part of configuration space. To make smarter moves we
-        will use the Metropolis-Hastings algorithm with the additions of
-        \emph{Importance Sampling}  (which will be discussed in due time). This
-        alteration of the algorithm is dependent on the \emph{drift force} of
-        the system\footnote{Also known as the \emph{quantum force}.}.
+        will add importance sampling (which will be discussed in due time) to
+        the algorithm. This alteration of the algorithm needs a new quantity
+        called the \emph{drift force} of the system\footnote{Also known as the
+        \emph{quantum force}.}.
         \begin{align}
@@
         \end{align}
-        Using this expression we are able to move towards parts of
-        configuration space where the gradient increases or decreases yielding a
-        better choice of movements. We will mainly be interested in the drift
-        force of a single particle $k$ as the Monte Carlo sampling involves
-        choosing a single particle at a time.
+        Using the drift force we are able to move towards more interesting parts
+        of configuration space as we are following the gradient of the
+        wavefunction. We will mainly be interested in the drift force of a
+        single particle $k$ as the Monte Carlo sampling involves choosing a
+        single particle at a time.
 

### 2018-04-05 project_1/article/article.tex
Spell check one-body densities
        Also included is a reference to Jørgen.

@@
     \subsection{One-Body Densities}
-        % TODO: Fill in this part.
-        % SEB: Sure thing, buddy
-        In order to make some conclusion towards where the particles in our systems
-        are most prone to exist, we need a good way to visualise our wavefunctions.
-        It is impossible for a human being to visualise a probability density in three dimensions,
-        because this measure would have to be represented in four dimensions. The fact that
-        we will be studying systems of up to $n=500$ particles only adds to the headache.
-        We therefore introduce the single-particle density function, or \emph{one-body density}
-        $\rho(x, t)$. The one-body density is defined by the integral over all coordinates
-        except one,
-        \begin{align}
-                \rho(x_1) \equiv \int \abs{\Psi(x_1, \dots, x_n)}^2 dx_2 \dots dx_n.
-        \end{align}
-        The one-body density can be interpreted as one-dimensional probability density for
-        finding a particle in $x_1$-direction. We also have the following relation,
-        \begin{align}
-                n = \int \rho(x_1)dx_1,
-        \end{align}
-        in other words, the integral over the entire one-dimensional space in $x_1$-direction
-        gives us the total number of particles in our system. This can be a bit odd at first sight
-        as we are used to getting the number $1$ when we integrate over the entire density.
-
-        It is best to think of the one-body density as the multivariate probability density
-        for the entire system collapsed down to one dimension in a representation than looks very
-        familiar to every physicist with a basic knowledge of quantum mechanics.
+        In order to make some conclusion towards where the particles in our
+        systems are most prone to exist, we need a good way to visualise our
+        wavefunctions. If we are to visualize the combined wavefunction of two
+        three-dimensional particles we need four dimensions per particle and can
+        subtract one dimension as we look at a single probability density
+        yielding a grand total of seven dimensions. The fact that we will be
+        studying systems of up to $N = 500$ particles only adds to the headache.
+        We therefore introduce the single-particle density function, or
+        \emph{one-body density} $\rho(\vf{r})$. The one-body density is defined
+        by the integral over all coordinates except one,
+        \begin{align}
+            \rho(\vf{r}_1)
+            \equiv
+            \int \dd\vf{r}_2 \dots \dd\vf{r}_N
+            \abs{\Psi(\vf{r}_1, \dots, \vf{r}_N)}^2.
+        \end{align}
+        The one-body density represents the distribution of all the particles in
+        the system. Furthermore, $\rho(\vf{r}_1)\dd\vf{r}_1$ will yield the
+        probability of finding any particle within the volume element
+        $\dd\vf{r}_1$. By convention the one-body density should be normalized
+        such that the following relation holds,
+        \begin{align}
+            N = \int \rho(\vf{r}_1)\dd\vf{r}_1.
+        \end{align}
+        In other words, the integral over the entire volume in
+        $\vf{r}_1$-direction gives us the total number of particles in our
+        system. This can be a bit odd at first sight as we are used to getting
+        unity when we integrate over the entire probability
+        density.\cite{hogberget2013quantum}
+
+        It is best to think of the one-body density as the multivariate
+        probability density for the entire system collapsed down to one
+        coordinate in a representation than looks very familiar to every
+        physicist with a basic knowledge of quantum mechanics.
 

### 2018-04-05 project_1/article/article.tex
Spell check systems section

@@
     where we can choose between a spherical (S) or an elliptical (E) harmonic
-    oscillator trap.  In the latter potential the oscillatory frequency is
-    different in the $z$-direction, i.e., $\omega_z \neq \omega$.  The full
-    Hamiltonian of the system is given by
+    oscillator trap. The full Hamiltonian of the system is given by
     \begin{align}
@@
     \end{align}
-    for equal masses, and the two-body interaction operator, $w(\vf{r}_i,
+    for equal masses. The two-body interaction operator, $w(\vf{r}_i,
     \vf{r}_j)$, is
@@
     \end{align}
-    where $a$ is the hard sphere of the particle.  The trial wavefunction,
-    $\Psi_T(\vf{r})$, we will be looking at is given by
+    where $a$ is the hard sphere of the particle.  The trial wavefunction we
+    will be looking at is given by
     \begin{align}
@@
         &= \Phi_T(\vf{r})
-        \prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k) \\
-        &= \Biggl(
-            \prod_{i = 1}^N g(\alpha, \beta, \vf{r}_i)
-        \Biggr)
+        J(\vf{r}),
+        %\\
+        %&= \Biggl(
+        %    \prod_{i = 1}^N g(\alpha, \beta, \vf{r}_i)
+        %\Biggr)
+        %\prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k),
+        \label{eq:initial_trial_wavefunction}
+    \end{align}
+    where $\vf{r} = (\vf{r}_1, \dots, \vf{r}_N)$. The Slater permanent,
+    $\Phi_T(\vf{r})$, is given by the $N$ first single-particle functions
+    \begin{align}
+        \Phi_T(\vf{r})
+        &=
+        \prod_{i = 1}^N \phi(\vf{r}_i)
+        = \prod_{i = 1}^N\exp\bigl[
+            -\alpha(x_i^2 + y_i^2 + \beta z_i^2)
+        \bigr].
+    \end{align}
+    Here $\alpha$ is the variational parameter of the trial
+    wavefunction.\footnote{We can also treat $\beta$ as a variational parameter,
+    but we will in this project limit ourselves to a single variational
+    parameter, i.e., $\alpha$.} The Jastrow factor, $J(\vf{r})$ contributes with
+    correlations between the particles and is represented as
+    \begin{align}
+        J(\vf{r})
+        &=
         \prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k),
-        \label{eq:initial_trial_wavefunction}
-    \end{align}
-    where $\alpha$ is a variational parameter and
-    \begin{align}
-        \vf{r} = (\vf{r}_1, \vf{r}_2, \dots, \vf{r}_N, \alpha, \beta).
-    \end{align}
-    Here $g$ are the single particle wavefunctions given by
-    \begin{align}
-        g(\alpha, \beta, \vf{r}_i)
-        = \exp\bigl[
-            -\alpha(x_i^2 + y_i^2 + \beta z_i^2)
-        \bigr] \equiv \phi(\vf{r}_i),
-    \end{align}
-    and $\Phi_T(\vf{r})$ is the \emph{Slater permanent} consisting of the $N$
-    first single particle wavefunctions, and the correlation wavefunction, $f$,
-    given by
+    \end{align}
+    where the correlation wavefunctions are given by
     \begin{align}
@@
     \end{align}
-    We will for brevity use the notation $\phi(\vf{r}_i) = \phi_i$ and $r_{jk} =
-    |\vf{r}_j - \vf{r}_k|$. In practice \autoref{eq:two-body_interaction} is
+    We will for brevity define the notation
+    \begin{gather}
+        \phi_i \equiv \phi(\vf{r}_i),
+        \\
+        r_{jk} \equiv \abs{\vf{r}_j - \vf{r}_k}.
+    \end{gather}
+    In practice \autoref{eq:two-body_interaction} is
     unnecessary as the situtation where $r_{jk} \leq a$ automatically yields a

### 2018-04-05 project_1/article/article.tex
Spell check non-interacting section

@@
                 -\alpha |\vf{r}_i|^2
-            },
-        \end{align}
-        where $|\vf{r}_i| = r_i$. As $a = 0$ the interaction term,
-        $w(\vf{r}_i, \vf{r}_j)$, vanishes and the Hamiltonian is given by
-        (in the spherical case)
+            }.
+        \end{align}
+        As $a = 0$ the interaction term, $w(\vf{r}_i, \vf{r}_j)$, vanishes and
+        the Hamiltonian is given by (in the spherical case)
         \begin{align}
@@
             \nabla^2_k\Psi_T(\vf{r})
-            &= \big(-2d\alpha + 4\alpha^2 r_k^2\bigr)\Psi_T(\vf{r}),
-        \end{align}
-        where $d$ is the dimensionality of the problem determined by $\vf{r}_k
+            &= \big(-2d\alpha + 4\alpha^2 r_k^2\bigr)\Psi_T(\vf{r}).
+        \end{align}
+        Here $d$ is the dimensionality of the problem determined by $\vf{r}_k
         \in \mathbb{R}^d$. We can thus use the gradient to find an expression
@@
         \end{align}
-        In natural units, $\hbar = c = 1$, and with the mass set to unity, $m =
-        1$, this reduces to
+        In natural units, $\hbar = c = 1$, and with the mass set to unity this
+        reduces to
         \begin{align}
@@
         \begin{align}
-            E_L(\vf{r})
+            E_L(\vf{r}; \alpha = \omega/2)
             &\equiv
@@
         \end{align}
-        which is the stable minimum for $\alpha = \omega/2$.
+        which turns out to be the exact minimum for the ground state energy of
+        this system.
 
@@
                 \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
-                \\
-                &=
+                =
                 \left(

### 2018-04-05 project_1/article/article.tex
Add derivation of the exact energy

@@
             }.
+            \label{eq:trial_simple_gaussian}
         \end{align}
@@
             &= \big(-2d\alpha + 4\alpha^2 r_k^2\bigr)\Psi_T(\vf{r}).
+            \label{eq:laplacian_simple_gaussian}
         \end{align}
@@
 \appendix
+\section{Computing the exact energy of the non-interacting system}
+    Here we show how to arrive at \autoref{eq:exact_energy}. We need to solve
+    two integrals consisting of $N$, $d$-dimensional, Gaussians from
+    \autoref{eq:trial_simple_gaussian}. We will label the position in the $j$-th
+    coordinate for the $i$-th particle by $x_{ij}$. Starting with the
+    normalization factor we get
+    \begin{align}
+        \bra{\Psi_T}\ket{\Psi_T}
+        &=
+        \int\dd\vf{r}\abs{\Psi_T(\vf{r})}^2
+        \\
+        &=
+        \prod_{i = 1}^N
+        \int\dd\vf{r}_i\exp(-2\alpha\vf{r}_i^2)
+        \\
+        &=
+        \prod_{i = 1}^N
+        \prod_{j = 1}^d
+        \int\dd x_{ij}\exp(-2\alpha x_{ij}^2)
+        \\
+        &=
+        \prod_{i = 1}^N
+        \prod_{j = 1}^d
+        \sqrt{\frac{\pi}{2\alpha}}
+        = \para{
+            \sqrt{\frac{\pi}{2\alpha}}
+        }^{dN}.
+    \end{align}
+    From the Hamiltonian we get two integrals containing $\vf{r}_i^2$ for each
+    particle $i$.
+    \begin{align}
+        \bra{\Psi_T}\vf{r}_i^2\ket{\Psi_T}
+        &=
+        \int\dd\vf{r}\abs{\Psi_T(\vf{r})}^2\vf{r}_i^2
+        \\
+        &=
+        \int\dd\vf{r}_i\exp(-2\alpha\vf{r}_i^2)\vf{r}_i^2
+        \nonumber \\
+        &\qquad
+        \times
+        \prod_{j \neq i}^N
+        \int\dd\vf{r}_j\exp(-2\alpha\vf{r}_j^2)
+        \\
+        &=
+        \sum_{l = 1}^d
+        \int
+        \para{
+            \prod_{k = 1}^d
+            \dd x_{ik}\exp(-2\alpha x_{ik}^2)
+        } x_{il}^2
+        \nonumber\\
+        &\qquad
+        \times
+        \para{
+            \sqrt{\frac{\pi}{2\alpha}}
+        }^{d(N - 1)}
+        \\
+        &=
+        \sum_{l = 1}^d
+        \frac{1}{4\alpha}\sqrt{\frac{\pi}{2\alpha}}
+        \para{
+            \sqrt{\frac{\pi}{2\alpha}}
+        }^{d - 1}
+        \nonumber\\
+        &\qquad
+        \times
+        \para{
+            \sqrt{\frac{\pi}{2\alpha}}
+        }^{d(N - 1)}
+        \\
+        &= \frac{d}{4\alpha}\para{
+            \sqrt{\frac{\pi}{2\alpha}}
+        }^{dN}.
+    \end{align}
+    We thus have that
+    \begin{align}
+        \frac{\bra{\Psi_T}\vf{r}_i^2\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
+        &=
+        \frac{d}{4\alpha}.
+    \end{align}
+    Using \autoref{eq:laplacian_simple_gaussian} we can find the exact energy
+    expression we are looking for.
+    \begin{align}
+        E(\alpha)
+        &=
+        \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
+        \\
+        &=
+        \sum_{i = 1}^N
+        \brac{
+            -\frac{\hbar^2}{2m}
+            \brak{
+                -2d\alpha
+                + 4\alpha^2\frac{d}{4\alpha}
+            }
+            + \frac{1}{2}m\omega^2\frac{d}{4\alpha}
+        }
+        \\
+        &=
+        \para{
+            \frac{\alpha\hbar^2}{2m}
+            + \frac{m\omega^2}{8\alpha}
+        }dN,
+    \end{align}
+    which is what we wanted to show.
+
 \section{Computing the drift force and local energy of the full system}

### 2018-04-05 project_1/article/article.tex
Spell check exact variational energy

@@
         which turns out to be the exact minimum for the ground state energy of
-        this system.
+        this system as we shall see.
 
@@
             \end{align}
-            By minimizing this expression with respect to $\alpha$ yields
+            By minimizing this expression with respect to $\alpha$ we get
             \begin{align}

### 2018-04-05 project_1/article/article.tex
Spell check section on interacting system

@@
         Moving to the full elliptical system for $\beta \neq 1$ and setting $a
-        \neq 0$ we can write the trial wavefunction as
-        \begin{align}
-            \Psi_T(\vf{r})
-            &=
-            \Phi_T(\vf{r})
-            J(\vf{r}),
-        \end{align}
-        where $\Phi_T(\vf{r})$ is the same Slater permanent as in
-        \autoref{eq:initial_trial_wavefunction} and $J(\vf{r})$ is the
-        \emph{Jastrow factor} given by
+        \neq 0$ we use the trial wavefunction from
+        \autoref{eq:initial_trial_wavefunction}. We rewrite the Jastrow factor
+        to
         \begin{align}
@@
         \end{align}
-
-        The Laplacian of particle $k$ divided by the trial wavefunction is
+        The Laplacian of particle $k$ divided by the trial wavefunction is then
         \begin{widetext}
@@
         where $\Delta \vf{r}_{km} = \vf{r}_k - \vf{r}_m$ and $d$ is the
-        dimensionality. Evaluating the gradient and Laplacian of the single
+        dimensionality. Evaluating the gradient and the Laplacian of the single
         particle functions yields
@@
         \end{align}
-        where the Laplcian is given by \autoref{eq:laplacian_full_system}.  We
+        where the Laplacian is given by \autoref{eq:laplacian_full_system}.  We
         will avoid writing out the full expression for the local energy for this

### 2018-04-05 project_1/article/article.tex
Spell check section on scaling the system
Spell check MH algorithm section

@@
     \end{align}
-    If we have found the ``true'' trial wavefunction $\ket{\Psi_T}$ with the
+    If we have found the "true" trial wavefunction $\ket{\Psi_T}$ with the
     correct variational parameters, i.e., the eigenfunction to the Hamiltonian
@@
             \end{align}
-
-            Looking at the one-body part of the Hamiltonian we get
+            The one-body part of the Hamiltonian then yields
             \begin{align}
@@
                     ({x_k'}^2 + {y_k'}^2 + \beta^2{z_k'}^2)
-                }
-                \equiv
-                \phi(\vf{r}_k').
+                }.
                 \label{eq:scaled_spf}
@@
             \item The system is initialised by a certain number $N$ of randomly
-                generated positions, or particles. This allows us to evaluate
-                the wavefunction at these points and compute the local energy
-                $E_L$.
+                generated positions, $\vf{r} = (\vf{r}_1, \dots, \vf{r}_N)$,
+                which represent the particles. This allows us to evaluate the
+                wavefunction at these points and compute the local energy
+                $E_L(\vf{r})$.
 
@@
 
-            \item A ratio between new wavefunction density and the previous
-                (initial) density is computed and compared to a random number.
-                This acceptance probability decides if the particle move is
-                rejected or accepted.  The particle is only allowed to move a
+            \item A ratio, $q$, between the probability density of the new
+                wavefunction and the previous probability density is computed.
+                We accept the step if $q \geq 1$, else we accept the step with a
+                probability $p$. The latter is done by drawing a random uniform
+                number, $x \in [0, 1]$, and checking if $q > x$.  This
+                acceptance probability decides if the particle move is rejected
+                or accepted.  The particle is only allowed to move a
                 predetermined step length.
 
-            \item If the particle movement is accepted and the local energy
-                $E_L$ is computed for the new system.
+            \item If the particle movement is accepted then the local energy is
+                computed for the new system.
 
@@
         probability and the acceptance probability.  The strength of the
-        algorithm is that the transition algorithm need not be known. For
+        algorithm lies in that the transition algorithm need not be known. For
         example, the simplest case is to accept the new state, i.e., the new
@@
         where $\vf{r}_{i + 1}$ are all the positions at step $i + 1$, is greater
-        than a uniform probability $p \in [0, 1)$.
+        than a uniform probability $p \in [0, 1]$.
 
@@
                 \autoref{eq:parameter_difference} which will be further away and
-                might result in a local minimum ``trapping'' the method. To
+                might result in a local minimum "trapping" the method. To
                 prevent this from happening we have added a limit to the allowed
@@
         We will start without importance sampling, that is by employing the
-        simpler and more naïve ``brute force'' Monte Carlo sampling. Moreover, we
+        simpler and more naïve "brute force" Monte Carlo sampling. Moreover, we
         will compute the value for the local energy analytically as well as with
@@
             In \autoref{fig:gradient_descent_interacting} we can see how the
-            conjugate gradient method ``moves'' to find the optimal value of
+            conjugate gradient method "moves" to find the optimal value of
             $\alpha$.
@@
             interacting case, the one-body density is skewed towards a larger
-            $\abs{\vf{r}}$, as is expected. The ``ideal'' density, a system with spherical
+            $\abs{\vf{r}}$, as is expected. The "ideal" density, a system with spherical
             trap, is also included in the figure.
@@
                 densities correspond to the elliptical potential with and
-                without the Jastrow factor respectively. The ``ideal'' system is
+                without the Jastrow factor respectively. The "ideal" system is
                 the spherical potential with the non-interacting Gaussian trial
@@
         For the non-interacting harmonic oscillator system there is no
-        difference between a new ``particle'' and new dimensions
+        difference between a new particle and new dimensions
         \footnote{This is not entirely true in the case of our implementation of

### 2018-04-06 project_1/article/article.tex
Added Haakon as reference.
Add section on initial spread

@@
                 generated positions, $\vf{r} = (\vf{r}_1, \dots, \vf{r}_N)$,
-                which represent the particles. This allows us to evaluate the
+                which represent the particles\footnote{Also known as
+                \emph{random walkers}.}. This allows us to evaluate the
                 wavefunction at these points and compute the local energy
@@
             \item The initial configuration is changed by setting a new position
-                for one of these particles. The particle is picked at random.
+                for one of these particles. The particle is picked at random and
+                is only allowed to move a predetermined step length.
 
@@
                 acceptance probability decides if the particle move is rejected
-                or accepted.  The particle is only allowed to move a
-                predetermined step length.
-
-            \item If the particle movement is accepted then the local energy is
-                computed for the new system.
-
-            \item Repeat steps until convergence and an optimum is reached.
+                or accepted.
+
+            \item If the particle movement is accepted we compute the local
+                energy for the new system. If not, we reset the system to the
+                previous state and reuse the last local energy.
+
+            \item Accumulate the local energy. Repeat the steps until
+                convergence is reached or a set number of iterations has been
+                done. When finished divide the accumulated energy by the number
+                of cycles to compute the expectation value of the energy.
 
@@
         of the parameter space in order to find the optimal parameters.  Whether
-        a proposed move is accepted or not is determined by a transition
-        probability and the acceptance probability.  The strength of the
-        algorithm lies in that the transition algorithm need not be known. For
-        example, the simplest case is to accept the new state, i.e., the new
-        position for the random walker, if the ratio
+        a proposed move is accepted or not is determined by a \emph{transition
+        probability}. The strength of the algorithm lies in that the transition
+        algorithm need not be known. For example, the simplest case is to accept
+        the new state, i.e., the new position for the random walker, if the
+        ratio
         \begin{align}
@@
         where $\vf{r}_{i + 1}$ are all the positions at step $i + 1$, is greater
-        than a uniform probability $p \in [0, 1]$.
+        than a uniform probability $p \in [0, 1]$. This transition criteria
+        results in what we call the brute force Metropolis-Hastings algorithm.
+        By counting the number of accepted steps and dividing by the total
+        number of Monte Carlo cycles we get the \emph{acceptance probability}.
+        The acceptance probability helps in choosing a decent length for the
+        proposed new step. If the acceptance probability is very low we can
+        lower the step length to get more accepted steps and vice versa.
+
+        \subsubsection{Initial distribution of particles}
+            Prior to performing the Metropolis-Hastings algorithm we have to an
+            initial distribution of the particles. As the algorithm is supposed
+            to move particles in the right direction this distribution can be
+            \emph{somewhat} arbitrary. There are however a few caveats to this
+            approach.
+
+            \begin{enumerate}
+                \item The distribution is unphysical. This results in the
+                    particles taking up positions which can yield bogus values
+                    for the local energy. A remedy is to allow for some
+                    \emph{thermalization steps}, i.e., scatter the particles and
+                    run the algorithm for a set number of steps prior to
+                    starting the actual sampling.
+                \item The initial \emph{spread} of the particles is too small or
+                    too large. The former can lead to an acceptance ratio of
+                    zero if no steps are accepted due to overlap. This is
+                    something that can happen to the interacting system as an
+                    overlap between the particles makes the wavefunction
+                    evaluate to zero thus rejecting all steps in the ratio test.
+                    The latter can lead to the system never reaching a true
+                    state as they are scattered too far away from the
+                    interesting regions in the configuration space.
+                \item The system imposes some restrictions on the distribution.
+                    This can mean that we have to distribute the particles in a
+                    certain way to make sure that we don't break the
+                    restrictions. This resembles the former case when the spread
+                    is too small for our systems.
+            \end{enumerate}
+
+            In the case of the interacting system we are looking at we have to
+            take some measures to make sure that the initial spread does not
+            break the Monte Carlo cycles. We therefore distribute all the
+            particles on a grid with a distance larger than the diameter of the
+            hard sphere before letting the system thermalize. After the
+            thermaliztion we start the sampling.
 
         \subsubsection{Importance Sampling}
-            A problem with the naïve Metropolis-Hastings sampling approach is
-            that the sampling of position space is done with no regard for where
-            we are likely to find a particle. This problem can be remedied
-            through \emph{importance sampling}.  It is reasonable to assume that
-            the particles we erratically scatter in space are prone to movement
-            towards the peaks of the probability density as dictated by the wave
-            function. Consider therefore the Fokker-Planck equation,
+            A problem with the naïve Metropolis-Hastings sampling approach
+            described above is that the sampling of position space is done with
+            no regard for where we are likely to find a particle. This problem
+            can be remedied through \emph{importance sampling}.  It is
+            reasonable to assume that the particles we erratically scatter in
+            space are prone to movement towards the peaks of the probability
+            density as dictated by the wave function. Consider therefore the
+            Fokker-Planck equation,
             \begin{align}

### 2018-04-06 project_1/article/article.tex
Spell check section on initial distribution

@@
         \emph{one-body density} $\rho(\vf{r})$. The one-body density is defined
-        by the integral over all coordinates except one,
+        by the integral over all coordinates except
+        one\cite{kristiansen2017time},
         \begin{align}
@@
         \subsubsection{Initial distribution of particles}
-            Prior to performing the Metropolis-Hastings algorithm we have to an
-            initial distribution of the particles. As the algorithm is supposed
-            to move particles in the right direction this distribution can be
-            \emph{somewhat} arbitrary. There are however a few caveats to this
-            approach.
+            Prior to performing the Metropolis-Hastings algorithm we have to
+            have an initial distribution of the particles. As the algorithm is
+            supposed to move particles in the right direction this distribution
+            can be \emph{somewhat} arbitrary. There are however a few caveats to
+            this approach.
 
@@
                     starting the actual sampling.
+
                 \item The initial \emph{spread} of the particles is too small or
@@
                     zero if no steps are accepted due to overlap. This is
-                    something that can happen to the interacting system as an
-                    overlap between the particles makes the wavefunction
-                    evaluate to zero thus rejecting all steps in the ratio test.
-                    The latter can lead to the system never reaching a true
-                    state as they are scattered too far away from the
-                    interesting regions in the configuration space.
+                    something that can happen to the interacting system we are
+                    studying as an overlap between the particles makes the
+                    wavefunction evaluate to zero thus rejecting all steps in
+                    the ratio test.  The latter can lead to the system never
+                    reaching a true state as they are scattered too far away
+                    from the interesting regions in the configuration space.
+
                 \item The system imposes some restrictions on the distribution.

### 2018-04-06 project_1/article/article.tex
Spell check section on importance sampling

@@
             which describes the evolution in time of a probability density
-            function. In our case this is the trial wavefunction $\Psi_T$.
-            Originaly an equation that models diffusion, we have a diffusion
-            term $D$ and the drift force \autoref{eq:drift_force}. In our case
-            the diffusion term $D$ is simply $1/2$ from the kinetic energy (in
-            natural units).
+            function. In our case this is the trial wavefunction
+            $\Psi_T$\footnote{The trial wavefunction is not a probability
+            density, but the Fokker-Planck equation is robust enough to handle
+            the wavefunction as is.}.  Originaly an equation that models
+            diffusion, we have a diffusion term $D$ and the drift force
+            \autoref{eq:drift_force}. In our case the diffusion term $D$ is
+            simply $1/2$ from the kinetic energy (in natural units).
 
@@
 
-            Now we need to change the acceptance probability of the metropolis
-            algorithm to something that takes the new sampling method into
-            account.
+            Now we need to change the ratio computation in the
+            Metropolis-Hastings algorithm to something that takes the new
+            sampling method into account.
             \begin{align}
@@
                 \frac{
-                    G(\vf{r}_{i + 1}, \vf{r}_{i}, \delta t)
+                    G(\vf{r}_{i}, \vf{r}_{i + 1}, \delta t)
                     \left|\Psi_T(\vf{r}_{i + 1})\right|^2
@@
                 {
-                    G(\vf{r}_{i}, \vf{r}_{i + 1}, \delta t)
+                    G(\vf{r}_{i + 1}, \vf{r}_{i}, \delta t)
                     \left|\Psi_T(\vf{r}_{i})\right|^2
@@
             \begin{align}
-                G(\vf{r}_{i + 1}, \vf{r}_i, \delta t)
+                G(\vf{r}_{i}, \vf{r}_{i + 1}, \delta t)
                 &=
@@
                     -\frac{\bigl[
-                        \vf{r}_{i + 1}
-                        - \vf{r}_{i}
-                        - D\vf{F}(\vf{r}_i)\delta t
+                        \vf{r}_{i}
+                        - \vf{r}_{i + 1}
+                        - D\vf{F}(\vf{r}_{i + 1})\delta t
                     \bigr]^2}{4D\delta t}
@@
             \end{align}
-            where $d$ is the dimensionality.
+            where $d$ is the dimensionality. The new ratio, $q$, is still
+            subjected to the same ratio test as in the brute force
+            Metropolis-Hastings algortihm.
 

### 2018-04-06 project_1/article/article.tex
Spell check the block method

@@
         \end{equation}
-        where $N$ is the number of samples, or Monte-Carlo cycles, in the experiment.
-        However, it is reasonable to assume that the data we are dealing with in this
-        study is liable to suffer from \emph{autocorrelation} and \autoref{eq:standard_deviation}
-        does not hold. The prevailing
-        definition of autocorrelation in a data stream or signal is correlation
-        between a delay of the signal and the original signal. One would be interested to find the
-        delay, or lag, in the signal at which the "self-correlation" is highest. We refer to this
-        spacing as $d$, and define the following correlation function,
+        where $N$ is the number of samples, or Monte-Carlo cycles, in the
+        experiment.  However, it is reasonable to assume that the data we are
+        dealing with in this study is liable to suffer from
+        \emph{autocorrelation} and \autoref{eq:standard_deviation} does not
+        hold. The prevailing definition of autocorrelation in a data stream or
+        signal is correlation between a delay of the signal and the original
+        signal. One would be interested to find the delay, or lag, in the signal
+        at which the "self-correlation" is highest. We refer to this spacing as
+        $d$, and define the following correlation function,
         \begin{equation}
-            f_d = \frac{1}{n -d} \sum_{k=1}^{n-d} (x_k - \bar{x}_n)(x_{k+d} - \bar{x}_n).
+            f_d
+            = \frac{1}{n -d} \sum_{k=1}^{n-d}
+            (x_k - \bar{x}_n)(x_{k+d} - \bar{x}_n).
             \label{eq:correlation_f_d}
         \end{equation}
-        The keen reader would have noticed that the function $f_d$ in \autoref{eq:correlation_f_d}
-        would be equal to the sample variance for $d=0$. We can now define the \emph{autocorrelation function}
+        The keen reader would have noticed that the function $f_d$ in
+        \autoref{eq:correlation_f_d} would be equal to the sample variance for
+        $d=0$. We can now define the \emph{autocorrelation function}
         \begin{equation}
@@
         \end{equation}
-        which is equal to $1$ if the data exhibits no autocorrealtions, equating to $d=0$. From the autocorrelation
-        function (\ref{eq:autocorrelation_function}) we in turn define the \emph{autocorrelation time},
+        which is equal to $1$ if the data exhibits no autocorrealtions, equating
+        to $d=0$. From the autocorrelation function in
+        \autoref{eq:autocorrelation_function} we in turn define the
+        \emph{autocorrelation time},
         \begin{equation}
-            \tau = 1 + 2 \sum_{d=1}^{n-1}\kappa_d,
+            \tau = 1 + 2 \sum_{d=1}^{n-1}\kappa_d.
         \end{equation}
-        notice that the autocorrelation time is $1$ for a correlation free experiment.
-
-        We are now able to make a correction to the expression for the standard deviation
-        improving on \autoref{eq:standard_deviation} by taking correlation into account,
+        Notice that the autocorrelation time is $1$ for a correlation free
+        experiment.
+
+        We are now able to make a correction to the expression for the standard
+        deviation improving on \autoref{eq:standard_deviation} by taking
+        correlation into account,
         \begin{equation}
-            \sigma =\sqrt{\frac{1 + 2\tau / \Delta t}{N}(\ev{E^2_L} - \ev{E_L}^2)},
+            \sigma =\sqrt{\frac{1 + 2\tau / \Delta t}{N}
+            (\ev{E^2_L} - \ev{E_L}^2)},
         \end{equation}
-        where $\Delta t$ is the time between each sample. The main problem at this point is to
-        find $\tau$, which we do not know for any given system and it is generally very expensive to
-        compute. In order to find a good estimate of $\tau$ we use a procedure called \emph{blocking}.
+        where $\Delta t$ is the time between each sample. The main problem at
+        this point is to find $\tau$, which we do not know for any given system
+        and it is generally very expensive to compute. In order to find a good
+        estimate of $\tau$ we use a procedure called \emph{blocking}.
 
         \subsubsection{Blocking}
-        In the method of blocking we group the samples into blocks of increasing size. If one where to compute
-        the standard deviation for each block, one should see the variance increase with the block size.
-        The standard deviation would only increase up to a certain point, from whence it would stay
-        almost constant. What is happening is that we have reached a point where a particular
-        sample from one block is no longer correlated with a corresponding sample from an adjacent block.
-        The block size for this point of convergence now functions as an estimate for the autocorrelation
-        time $\tau$.
-
-         Instead of going by this "manual" method of looking at charts, we will construct a test statistic and
-         let the computer do the necessary considerations in a more "automatic way.
-
-        The easiest way to do this is by use of the \emph{automatic blocking scheme}, nicely illustrated
-        with a flow chart in \autoref{fig:blocking}. The input parameter for the method is an array of the
-        local energies, but it can be any array of time series data, which is why it is herein referred to as
-        $\mathbf{X}$. Any array of sequential data, where any particular data point depends on the previous
-        data point inhibits all attributes of a time series. Our array of local energies is therefor most definitely
-        a time series.
-
-        From $\mathbf{X}$ we compute the variance and sample covariance between the series itself and a
-        one-point lag of the series. This covariance is usually referred to as the first order autocovariance.
+            In the method of blocking we group the samples into blocks of
+            increasing size. If one where to compute the standard deviation for
+            each block, one should see the variance increase with the block
+            size.  The standard deviation would only increase up to a certain
+            point, from whence it would stay almost constant. What is happening
+            is that we have reached a point where a particular sample from one
+            block is no longer correlated with a corresponding sample from an
+            adjacent block.  The block size for this point of convergence now
+            functions as an estimate for the autocorrelation time $\tau$.
+
+            Instead of going by this "manual" method of looking at charts, we
+            will construct a test statistic and let the computer do the
+            necessary considerations in a more "automatic way.
+
+            The easiest way to do this is by use of the \emph{automatic blocking
+            scheme}, nicely illustrated with a flow chart in
+            \autoref{fig:blocking}. The input parameter for the method is an
+            array of the local energies, but it can be any array of time series
+            data, which is why it is herein referred to as $\mathbf{X}$. Any
+            array of sequential data, where any particular data point depends on
+            the previous data point inhibits all attributes of a time series.
+            Our array of local energies is therefore most definitely a time
+            series.
+
+            From $\mathbf{X}$ we compute the variance and sample covariance
+            between the series itself and a one-point lag of the series. This
+            covariance is usually referred to as the first order autocovariance.
             \begin{align}
@@
                     \sum_{k=1}^n
-                \left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)
+                    \left((\mathbf{X}_i)_k - \bar{\mathbf{X}}\right)
                 \right)
@@
                     \sum_{k=0}^{n-1}
-                \left((\mathbf{X}_i)_k - (\bar{\mathbf{X}}_i)_k\right)
+                    \left((\mathbf{X}_i)_k - \bar{\mathbf{X}}\right)
                 \right).
             \end{align}
-
-        Now we "block"! The next step is to perform a transformation of the our time series $\mathbf{x}_i$, where
-        we end up with an array that is half the length of the original time series,
+            Note that $i$ represents the iteration in the blocking process
+            whereas $k$ is the index into the array $\mathbf{X}_i$.
+
+            Now we "block"! The next step is to perform a transformation of the
+            our time series $\mathbf{X}_i$, where we end up with an array that
+            is half the length of the original time series,
             \begin{equation}
-            (\mathbf{X}_i)_k = \frac{1}{2}\left((\mathbf{X}_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right).
+                (\mathbf{X}_i)_k
+                = \frac{1}{2}\left((\mathbf{X}_i)_{2k-1}
+                + (\mathbf{X}_i)_{2k} \right).
             \end{equation}
-        This process is were the term "blocking" stems from. There are a number of ways to perform the blocking -
-        we could have picked the data points used in the new ($i+1$) array randomly, or more orderly. Here we we have
-        picked them sequentially. We have obtained $\frac{n}{2}$ new "primed" stochastic variables. Assuming that the
-        reader is playing close attention, it is now needless to say that it is much easier to perform these sequential
-        blocking steps if we have an array of length $n= 2^d$, where $d$ is integer. Otherwise, problems would arise
-        because one would have arrays of different lengths when the original array is "split in two". The ailment is
-        easy to remedy by excluding an observation from the longest resulting array after splitting.
-
-        For each new time series $\mathbf{X}_{i+1}$ a new sample variance and sample covariance is computed.
-        The blocking transformation is allowed to continue until we are we run out of data, i.e. the length
-        of $\mathbf{X}_i < 2$. In other words, there is only a single number left. We now compute a test statistic,
+            This process is were the term "blocking" stems from. There are a
+            number of ways to perform the blocking - we could have picked the
+            data points used in the new ($i+1$) array randomly, or more orderly.
+            Here we have picked them sequentially. We have obtained $n/2$ new
+            "primed" stochastic variables. Assuming that the reader is playing
+            close attention, it is now needless to say that it is much easier to
+            perform these sequential blocking steps if we have an array of
+            length $n= 2^d$, where $d$ is integer. Otherwise, problems would
+            arise because one would have arrays of different lengths when the
+            original array is split in two. The ailment is easy to remedy by
+            excluding an observation from the longest resulting array after
+            splitting.
+
+            For each new time series $\mathbf{X}_{i+1}$ a new sample variance
+            and sample covariance is computed.  The blocking transformation is
+            allowed to continue until we run out of data, i.e. the length of
+            $\mathbf{X}_i < 2$. In other words, there is only a single number
+            left.  We now compute a test statistic,
             \begin{equation}
-            M_j = \sum_{k=j}^{d-1} n_k \frac{\left[(n_k - 1)\frac{s^2_k}{n_k^2} y \gamma_k(1)\right] ^2}{s_k^4}.
+                M_j
+                = \sum_{k=j}^{d-1}
+                n_k \frac{\left[(n_k - 1)
+                \sigma^2_k/n_k^2 + \gamma_k\right] ^2}{\sigma_k^4},
             \end{equation}
-        As luck would have it, this number $M$ is   $\chi^2$-distributed. The following step would be to look up a
-        statistical table and find the first $k$, such that $M_k \leq q_{d-k}(1-\alpha)$, where $q$ is the statistical
-        limit and $\alpha$ is  the significance level that your hear desires. Finally, the correct sample standard deviation
-        is $\frac{s_k^2}{n_k}$. Again, see \autoref{fig:blocking} for a summary of the algorithm.
+            As luck would have it, this number $M$ is   $\chi^2$-distributed.
+            The following step would be to look up a statistical table and find
+            the first $k$, such that $M_k \leq q_{d-k}(1-\alpha)$, where $q$ is
+            the statistical limit and $\alpha$ is your favorite  significance
+            level. Finally, the correct sample variance is $\sigma_k^2/n_k$. In
+            the rest of the project we use the labeling
+            \begin{align}
+                \sigma_b^2 \equiv \frac{\sigma_k^2}{n_k},
+            \end{align}
+            for the correct sample variance found from the blocking method.
+            Again, see \autoref{fig:blocking} for a summary of the algorithm.
 

### 2018-04-06 project_1/article/article.tex
Spell check figure

@@
                 % Defining blocks of flowchart
-            \tikzstyle{decision} = [diamond, draw, fill=blue!20,
-                text width=5em, text badly centered, node distance=3cm, inner sep=0pt]
-            \tikzstyle{block} = [rectangle, draw, fill=blue!20,
-                    text width=11em, text centered, rounded corners, minimum height=4em]
+                \tikzstyle{decision} = [diamond, draw, fill=blue!20, text
+                width=5em, text badly centered, node distance=3cm, inner
+                sep=0pt]
+
+                \tikzstyle{block} = [rectangle, draw, fill=blue!20, text
+                width=11em, text centered, rounded corners, minimum height=4em]
+
                 \tikzstyle{line} = [draw, -latex']
@@
                     \node [block] (init) {Start with $\mathbf{X}_i$, $i=0$};
-                \node [block, below of=init] (variance) {Compute variance and autocovariance of $\mathbf{X}_i$, $s_i$ and $\gamma_i$};
-                \node [block, below of=variance] (transform)
-                    {Transform data $(\mathbf{X}_{i+1})_k = \frac{1}{2}\left((\mathbf{X})_i)_{2k-1} + (\mathbf{X}_i)_{2k} \right)$};
-                \node [block, right=1.0cm of transform] (i+1) {Next iteration: $i = i + 1$.} ;
-                \node [decision, below of = transform] (x_size) {Length of $\mathbf{X}_{i + 1} \geq 2$ ? };
-                \node [block, below of=x_size] (m_test) {Compute test statistic $M_j$ from $s_j$ and $\gamma_j$ for all $j$.};
-                \node [block, below of=m_test] (magic_numbers) {Find first $k$ so that $M_k \leq q_{d-k}(1-\alpha)$};
-                \node [block, below of=magic_numbers] (return) {Return $s_k^2/n_k$};
+
+                    \node [block, below of=init] (variance) {Compute variance
+                    and autocovariance of $\mathbf{X}_i$, $\sigma_i$ and
+                    $\gamma_i$};
+
+                    \node [block, below of=variance] (transform) {Transform data
+                    $(\mathbf{X}_{i+1})_k =
+                    \frac{1}{2}\left((\mathbf{X})_i)_{2k-1} +
+                    (\mathbf{X}_i)_{2k} \right)$};
+
+                    \node [block, right=1.0cm of transform] (i+1) {Next
+                    iteration: $i = i + 1$.} ;
+
+                    \node [decision, below of = transform] (x_size) {Length of
+                    $\mathbf{X}_{i + 1} \geq 2$ ? };
+
+                    \node [block, below of=x_size] (m_test) {Compute test
+                    statistic $M_j$ from $\sigma_j$ and $\gamma_j$ for all
+                    $j$.};
+
+                    \node [block, below of=m_test] (magic_numbers) {Find first
+                    $k$ such that $M_k \leq q_{d-k}(1-\alpha)$};
+
+                    \node [block, below of=magic_numbers] (return) {Return
+                    $\sigma_k^2/n_k$};
 
@@
                 \end{tikzpicture}
-            \caption{Flow chart depicting the automatic blocking method for finding a better estimate for sample variance.}
+                \caption{Flow chart depicting the automatic blocking method for
+                finding a better estimate for the sample variance.}
                 \label{fig:blocking}

### 2018-04-06 project_1/article/article.tex
Spell check gradient descent section

@@
         \subsection{Gradient Descent}
-
             Scanning the entire space of variational parameters can be a
@@
             for integation.}. In any optimum we generally have that the
-            derivative should equal zero. In our case, we want to find the
-            minimum of the expected local energy with respect to some
-            variational parameter $\alpha$, so that
+            derivative should equal zero. Our goal is to find the minimum of the
+            expected local energy with respect to the variational parameters
+            $\vfg{\alpha}$.
             \begin{equation}
-                \nabla_\alpha \expval{E_L(\mathbf{R}; \alpha)} = 0.
+                \nabla_{\vfg{\alpha}} \expval{E_L(\mathbf{R}; \vfg{\alpha})}
+                = 0.
             \end{equation}
@@
             that some analytic \footnote{We use the term analytic quite
-            recklessly.} function $\vf{F}(\alpha)$ decreases fastest in the direction
-            of $-\nabla_\alpha \cdot \vf{F}(\alpha)$. This means that we could eventually
-            get to an ensuing minimum by trailing along the pathway of the
-            following recursive relation,
+            recklessly.} function $\vf{F}(\vfg{\alpha})$ decreases fastest in
+            the direction of $-\nabla_{\vfg{\alpha}} \cdot
+            \vf{F}(\vfg{\alpha})$. This means that we could eventually get to an
+            ensuing minimum by trailing along the pathway of the following
+            recursive relation,
             \begin{equation}
                 \vfg{\alpha}_{k+1}
-                = \vfg{\alpha}_k - \gamma \nabla_{\alpha}\cdot \vf{F}(\alpha_k).
+                = \vfg{\alpha}_k
+                - \gamma \nabla_{\vfg{\alpha}}\cdot \vf{F}(\vfg{\alpha}_k).
                 \label{eq:parameter_difference}
             \end{equation}
-            If $\gamma$ is small enough we should have $\vf{F}(\alpha_k) \geq
-            \vf{F}(\alpha_{k+1})$ for all $k$ and hopefully we will experience
-            convergence to the desired minimum.
-
-            For our special case we only have one variational parameter,
-            $\alpha$.  The expression for the derivative is,
-            \begin{align}
-                \nabla_{\alpha}\expval{E_L}
+            If $\gamma$ is small enough we should have $\vf{F}(\vfg{\alpha}_k)
+            \geq \vf{F}(\vfg{\alpha}_{k+1})$ for all $k$ and hopefully we will
+            experience convergence to the desired minimum.
+
+            The variational gradient of the expectation value of the energy is
+            thus
+            \begin{align}
+                \nabla_{\vfg{\alpha}}\expval{E_L}
                 =
                 2\left(
-                    \expval{E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
-                    - \expval{\frac{1}{\Psi_T} \nabla_{\alpha}\Psi_T}
+                    \expval{E_L\frac{1}{\Psi_T}\nabla_{\vfg{\alpha}}\Psi_T}
+                    - \expval{\frac{1}{\Psi_T} \nabla_{\vfg{\alpha}}\Psi_T}
                     \expval{E_L}
@@
             \end{align}
-
             To evaluate this expression we sample the local energy, the
@@
     \begin{align}
-        \nabla_{\alpha}\expval{E_L}
-        &=
-        \nabla_{\alpha}
+        \nabla_{\vfg{\alpha}}\expval{E_L}
+        &=
+        \nabla_{\vfg{\alpha}}
         \para{
@@
         }{\para{\int\dd\vf{r}\Psi_T}^2}
-        \int\dd\vf{r}\Psi_T\nabla_{\alpha}\Psi_T
+        \int\dd\vf{r}\Psi_T\nabla_{\vfg{\alpha}}\Psi_T
         \\
@@
         \int\dd\vf{r}\Psi_TH\brak{
-            \nabla_{\alpha}\Psi_T
+            \nabla_{\vfg{\alpha}}\Psi_T
         }
@@
         \int\dd\vf{r}\Psi_T^2\para{
-            \frac{\nabla_{\alpha}\Psi_T}{\Psi_T}
+            \frac{\nabla_{\vfg{\alpha}}\Psi_T}{\Psi_T}
         }
@@
         \int\dd\vf{r}\Psi_T^2 E_L\para{
-            \frac{\nabla_{\alpha}\Psi_T}{\Psi_T}
+            \frac{\nabla_{\vfg{\alpha}}\Psi_T}{\Psi_T}
         }.
@@
     \begin{align}
-        \nabla_{\alpha}\expval{E_L}
+        \nabla_{\vfg{\alpha}}\expval{E_L}
         =
@@
             \expval{
-                E_L\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T
+                E_L\frac{1}{\Psi_T}\nabla_{\vfg{\alpha}}\Psi_T
             }
             - \expval{E_L}
-            \expval{\frac{1}{\Psi_T}\nabla_{\alpha}\Psi_T}
+            \expval{\frac{1}{\Psi_T}\nabla_{\vfg{\alpha}}\Psi_T}
         },

### 2018-04-06 project_1/vmc/vmc/src/wavefunctions/interacting_elliptical_gaussian.cpp
Use dimensionless SPF

@@ -71,7 +71,7 @@ double inline InteractingEllipticalGaussian::evaluate_single_particle_function(
             (i != 2) ? SQUARE(position[i]) : (m_beta*SQUARE(position[i]));
     }
 
-    return exp(-(alpha*HBAR/(m_mass*m_omega))*position_sum);
+    return exp(-alpha*position_sum);
 }
 
 // Private method. Interaction (correlation) contributions.

### 2018-04-06 project_1/article/article.tex
Remove scaling of SPF

@@
     to evaluate the ground state energy of a trapped, hard sphere Bose gas.
-    \cite{dubois2001bose} \cite{nilsen2005vortices}
 
@@
             where we have introduced the dimensionless frequency $\lambda =
-            \omega_z/\omega$. The single particle functions also get a scaling
-            factor
-            \begin{align}
-                \phi_k
-                &=
-                \exp\brak{
-                    -\frac{\alpha\hbar}{m\omega}
-                    ({x_k'}^2 + {y_k'}^2 + \beta^2{z_k'}^2)
-                }.
-                \label{eq:scaled_spf}
-            \end{align}
-            The correlation wavefunction and the interaction potential remain the
-            same, but using the scaled hard core radius $a/a_{\text{ho}} =
-            0.0043$.
+            \omega_z/\omega$. The wavefunction does not need to be scaled as the
+            exponentials are already dimensionless. We only need to use the
+            scaled hard core radius $a' = a/a_{\text{ho}}$. We are interested in
+            using the radius $a' = 0.0043$ as in the references
+            \cite{dubois2001bose} and \cite{nilsen2005vortices}.
 
@@
 
+                This is a so-called "hacky" solution, but it does the trick. It
+                is also the reason why the result figure in
+                \autoref{fig:gradient_descent_interacting}.
+
 \section{Results}

### 2018-04-06 project_1/article/article.tex
Explain look of steepest descent figure
Spell check results on non-interaction

@@
                 number, $x \in [0, 1]$, and checking if $q > x$.  This
-                acceptance probability decides if the particle move is rejected
+                transition probability decides if the particle move is rejected
                 or accepted.
@@
         By counting the number of accepted steps and dividing by the total
-        number of Monte Carlo cycles we get the \emph{acceptance probability}.
-        The acceptance probability helps in choosing a decent length for the
-        proposed new step. If the acceptance probability is very low we can
-        lower the step length to get more accepted steps and vice versa.
+        number of Monte Carlo cycles we get the \emph{acceptance probability},
+        denoted $A$.  The acceptance probability helps in choosing a decent
+        length for the proposed new step. If the acceptance probability is very
+        low we can lower the step length to get more accepted steps and vice
+        versa.
 
@@
                 This is a so-called "hacky" solution, but it does the trick. It
-                is also the reason why the result figure in
-                \autoref{fig:gradient_descent_interacting}.
+                is also the reason why the resultant figure in
+                \autoref{fig:gradient_descent_interacting} looks the way it
+                does for $\alpha_0 = 0.2$ and $\alpha_0 = 0.3$.
 
@@
         Initially we are interested in seeing how our implementation performs
-        and to what extent it functions properly.  the logical thing to do would
+        and to what extent it functions properly. The logical thing to do would
         be to look at a non-interacting system with a simple Gaussian
@@
         (\autoref{eq:closed_form_natural_units_local_energy}). This expression
-        is minimised for $\alpha = \frac{1}{2}\omega$. We also set $\omega = 1$
+        is minimised for $\alpha = \omega/2$. We also set $\omega = 1$
         and get an expression for the expected minimum local energy,
@@
 
-        We will start without importance sampling, that is by employing the
-        simpler and more naïve "brute force" Monte Carlo sampling. Moreover, we
-        will compute the value for the local energy analytically as well as with
-        a numerical approach.  In the numerical approach we employ the well-know
-        approximation for the second derivative,
+        We will start without importance sampling, that is by employing brute
+        force Monte Carlo sampling. Moreover, we will compute the value for the
+        local energy analytically as well as with a numerical approach.  In the
+        numerical approach we employ the second order central difference
+        approximation
         \begin{equation}
-                f''(x) \approx \frac{f(x-h) - 2f(x) + f(x+h)}{h^2}.
+            \dod[]{f(x)}{x} \approx \frac{f(x-h) - 2f(x) + f(x+h)}{h^2}.
         \end{equation}
@@
             \caption{One particle in one dimension for the analytic expression
-            with $2^{21}$ Monte Carlo cycles and step length of $0.5$.}
+            with $2^{21}$ Monte Carlo cycles and a step length of $0.5$.}
             \begin{ruledtabular}
@@
         \autoref{tab:1D1N_analytic} shows the expected local energy
-        $\expval{E_L}$, standard deviations $\sigma$, $\sigma_b$, the share of
-        accepted Monte Carlo reorientations $A$ and cpu time for the analytic
-        computation at different values for the variational parameter $\alpha$.
+        $\expval{E_L}$, the uncorrelated standard deviation, $\sigma$, the
+        standard deviation found from blocking, $\sigma_b$, the accpetance
+        probability, $A$, and the CPU time for the analytic computation at
+        different values for the variational parameter $\alpha$.
         \autoref{tab:1D1N_numeric} shows the same numbers but for the numeric
-        scheme. The first column of standard deviations $\sigma$ are the "naïve"
-        standard deviations, while $\sigma_b$ marks the standard deviations
-        computed with the blocking method. One can clearly see, in both the
-        analytic and numeric case, that the former standard deviations are lower
-        than the latter, meaning that autocorrelated data can lead to a lower
-        perceived uncertainty than what is reality. The blocking method takes
-        this fact into account and for this reason, all other standard
-        deviations will be computed with the blocking method from here on.
+        scheme.  One can clearly see, in both the analytic and numeric case,
+        that the uncorrelated standard deviations are lower than the standard
+        deviations found from the blocking method. This means that
+        autocorrelated data can lead to a lower perceived uncertainty than what
+        is the reality. The blocking method takes this fact into account and for
+        this reason, all other standard deviations will be computed with the
+        blocking method from here on.
 
         The first initial observation of note is that we have found an optimum
-        for the system in terms of a minimum energy at $\alpha=\frac{1}{2}$.
+        for the system in terms of a minimum energy at $\alpha = 1/2$.
         Moreover, notice how the standard deviation disappears completely.  This
@@
         \autoref{eq:closed_form_natural_units_local_energy}, as the term
-        involving particle positions disappears for the optimal $\alpha$.
+        involving particle positions disappears for the optimal $\alpha$ thus
+        rendering the entire Monte Carlo process of moving particles void.
 
@@
             \caption{Comparison of energy, standard deviation and the CPU time
-            for $\alpha = 0.5$ (the correct minimum) for the analytic and
-            numeric schemes in three dimensions and $N$ particles. The step
-            length is $0.5$.}
+            for $\alpha = 0.5$ for the analytic and numeric schemes in three
+            dimensions and $N$ particles. The step length is $0.5$.}
             \centering
@@
         results are quite predictable and monotonous; the expected local energy
-        $\expval{E_L}$ is perfectly proportional by a factor $\frac{3}{2}$ to
+        $\expval{E_L}$ is perfectly proportional by a factor $3/2$ to
         the number of particles and the standard deviation $\sigma_b$ is zero
@@
         particles, the energy misses its target critically - the true energy is
-        more than one standard deviation away. Lastly, the cpu time is notably
+        more than one standard deviation away. Lastly, the CPU time is notably
         higher for the numeric scheme.
@@
 \section{Brute Force Metropolis-Hastings}
+    In \autoref{fig:initial_problem_b} we see how the expectation value of the
+    energy for the analytic and numeric expression of the local energy using
+    Monte Carlo sampling matches the exact energy from \autoref{eq:exact_energy}
+    for $\alpha \in [0.3, 0.7]$.
     \begin{figure}[b]

### 2018-04-06 project_1/article/article.tex
Spell check results from importance sampling

@@
         \subsubsection{Introducing Importance Sampling}
-
-            Next we will introduce importance sampling, the first prediction for
-            importance sampling compared to brute, naïve method is that we
-            should approach an equilibrium much faster because we are now making
-            moves in a smarter way. This prediction is confirmed in
+            Next we will introduce importance sampling. The first prediction for
+            importance sampling compared to the brute force, naïve method is
+            that we should approach an equilibrium much faster because we are
+            now making moves in a smarter way. This prediction is confirmed in
             \autoref{fig:problem_c1} and \autoref{fig:problem_c2}.
@@
                     {../data/figures/problem_c1.pdf}
-                    \caption{As the number of Monte Carlo cycles increases the
-                    standard deviation of the local energies decrease. This
-                    effect is stronger when the importance sampling scheme is
-                    employed.}
+                    \caption{As the number of Monte Carlo cycles increases, the
+                    standard deviation of the expectation value of the energy
+                    for the brute force and importance sampling schemes
+                    decreases.}
                     \label{fig:problem_c1}
@@
                     \caption{The absolute difference between the expected energy
-                    and the exact energy for both brute force and importance
-                    sampling. }
+                    and the exact energy for both the brute force method and
+                    importance sampling.}
                     \label{fig:problem_c2}
@@
 
-            \autoref{fig:problem_c1} shows the standard deviation of the local
-            energies for increasing number of Monte Carlo cycles for both the
-            regular sampling and the importance sampling methods. We see that
-            the standard deviation decreases for both methods for a larger
-            number of Monte Carlo cycles, but the standard deviation is always
-            lower for the importance sampling method. This is in accordance with
-            our expectations.
+            In \autoref{fig:problem_c1} we can see the standard deviation of the
+            local energies for an increasing number of Monte Carlo cycles for
+            both the brute force sampling and the importance sampling methods.
+            We see that the standard deviation decreases for both methods for a
+            larger number of Monte Carlo cycles, but the standard deviation is
+            always lower for the importance sampling method. This is in
+            accordance with our expectations.
 
             We have also compared the expected energy with the exact energy for
-            different number of Monte Carlo cycles. The result of this analysis
-            is shown in \autoref{fig:problem_c2}. We see that both methods will
-            evnetually result in expected energies that are equal to the exact
-            energies as $\abs{E - \expval{E}} \to 0$. However, for the method of
-            importance sampling the error is consistently lower.
+            a different number of Monte Carlo cycles. The result of this
+            analysis is shown in \autoref{fig:problem_c2}. We see that both
+            methods will eventually result in expected energies that are equal
+            to the exact energies as $\abs{E - \expval{E}} \to 0$. However, for
+            the method of importance sampling the error is consistently lower.
 
             \begin{table}
-                \caption{Importance sampled simulations of $n=10$ particles in
+                \caption{Importance sampled simulations of $N=10$ particles in
                 $d=3$ dimensions. Average acceptance ratios and standard
-                deviations for eleven alpha values between $0.3$ and $0.7$.}
+                deviations for eleven values of $\alpha \in [0.3, 0.7]$.}
                 \centering
@@
             \autoref{tab:imp_time_step} shows the standard deviations and
-            acceptance ratios for different time steps. The simulations in this
-            tabel are from the same span $\alpha \in [0.3, 07]$ as in
-            \autoref{tab:1D1N_analytic} and \autoref{tab:1D1N_numeric}. The
-            optimum energies for $\alpha = \frac{1}{2}$ gives the same result
-            for importance sampling as for the naïve way of sampling. Notice
-            that by lowering the time step, the uncertainty starts to decrease
-            and then starts to increase again. Moreover, the acceptance ratios
-            increases as the time step decreases.
+            acceptance ratios for different time steps, $\delta t$ using the
+            importance sampling method. The simulations in this table are from
+            the same values of $\alpha$ as in \autoref{tab:1D1N_analytic} and
+            \autoref{tab:1D1N_numeric}. The optimum energies for $\alpha = 1/2$
+            gives the same result for importance sampling as for the naïve way
+            of sampling. Notice that by lowering $\delta t$, the uncertainty
+            starts to decrease and then starts to increase again. Moreover, the
+            acceptance ratios increases as $\delta t \to 0$. This happens as the
+            change in the wavefunction grows minuscule and $q(\vf{r}_{i + 1},
+            \vf{r}_i) \approx 1$.
 

### 2018-04-06 project_1/article/article.tex
Spell check interacting results

@@
     \subsection{Interacting elliptical harmonic oscillator}
-        We now make a change to the simple systems in the previous section in
+        We now make a change to the simple system in the previous section in
         order to make everything a bit more interesting. The potential in the
@@
         We also choose a new trial wavefunction with the elliptical gaussian
-        single particle functions in \autoref{eq:scaled_spf} and the Jastrow
-        factor. The hard sphere radius is set to $a/a_{\text{ho}} = 0.0043$.
+        single particle functions and the Jastrow factor. The hard sphere radius
+        is set to $a = 0.0043$.
 
@@
         additional $10\%$ of the Monte Carlo cycles for thermalization of the
-        system prior to doing any sampling. We have used $\delta t = 0.1$ yields
-        a high acceptance ratio, but slower convergence.
+        system prior to doing any sampling. We have used $\delta t = 0.1$ which
+        yields a high acceptance ratio, but slower convergence.
 
@@
         \autoref{tab:100_interacting} respectively. We observe that the energy
-        for systems of interacting particles increases rapidly as the size of
-        the system increases.
+        for systems of interacting particles increases rapidly with the size of
+        the system.
 

### 2018-04-06 project_1/article/article.tex
Spell check section on finding the minimum

@@
 
-            For this optimal, interacting $\alpha$ we have computed the expected
-            energy and standard deviation for systems of $n = \{10, 50, 100\}$
-            particles. These results are shown in \autoref{tab:10_optimal}, the
-            non-interacting case, excluding Jastrow factor is included as a
-            comparison.
+            For this value of $\alpha$ we have computed the expected energy and
+            standard deviation for systems of $N = \{10, 50, 100\}$ particles.
+            These results are shown in \autoref{tab:interacting_optimal}. We have
+            compared with the non-interacting case, i.e., where the Jastrow
+            factor is zero ($a = 0$).
 
@@
 
-        \subsubsection{One-Body Densities}
-            We now look at one-body densities for the value of $\alpha$ found in
-            \autoref{eq:optimal_interacting_alpha}. The results for a system consisting
-            of $n=100$ particles are displayed in \autoref{fig:one_body_density_100}.
-            In this figure we see the densities for for a system confined to a
-            perturbed elliptic trap with and without interaction. We see that for the
-            interacting case, the one-body density is skewed towards a larger
-            $\abs{\vf{r}}$, as is expected. The "ideal" density, a system with spherical
-            trap, is also included in the figure.
-
             \begin{table}
-                \caption{Comparison of the elliptical system with an elliptical
-                gaussian trial wavefunction with and without the Jastrow
-                factor.}
+                \caption{Comparison of an elliptical gaussian trial wavefunction
+                with and without the Jastrow factor using the elliptical trap.}
                 \centering
@@
                 \end{ruledtabular}
-                \label{tab:10_optimal}
+                \label{tab:interacting_optimal}
             \end{table}
+
+        \subsubsection{One-Body Densities}
+            We now look at one-body densities for the value of $\alpha$ found in
+            \autoref{eq:optimal_interacting_alpha}. The results for a system consisting
+            of $n=100$ particles are displayed in \autoref{fig:one_body_density_100}.
+            In this figure we see the densities for for a system confined to a
+            perturbed elliptic trap with and without interaction. We see that for the
+            interacting case, the one-body density is skewed towards a larger
+            $\abs{\vf{r}}$, as is expected. The "ideal" density, a system with spherical
+            trap, is also included in the figure.
 
@@
         larger of the interacting systems we have looked at, with $n=100$ particles,
-        The interaction add only a mere $9\%$ extra energy (see \autoref{tab:10_optimal}).
+        The interaction add only a mere $9\%$ extra energy (see \autoref{tab:interacting_optimal}).
         Studying \autoref{fig:one_body_density_100} will reveal the same.

### 2018-04-06 project_1/article/article.tex
Spell check section on one body densities

@@
             We now look at one-body densities for the value of $\alpha$ found in
-            \autoref{eq:optimal_interacting_alpha}. The results for a system consisting
-            of $n=100$ particles are displayed in \autoref{fig:one_body_density_100}.
-            In this figure we see the densities for for a system confined to a
-            perturbed elliptic trap with and without interaction. We see that for the
-            interacting case, the one-body density is skewed towards a larger
-            $\abs{\vf{r}}$, as is expected. The "ideal" density, a system with spherical
-            trap, is also included in the figure.
+            \autoref{eq:optimal_interacting_alpha}. The results for a system
+            consisting of $N=100$ particles are displayed in
+            \autoref{fig:one_body_density_100}.  In this figure we see the
+            densities for a system confined to a perturbed elliptic trap with
+            and without interaction. We see that for the interacting case, the
+            one-body density is skewed towards a larger $\abs{\vf{r}}$, as is
+            expected. The hard core prevents the bosons from stacking on top of
+            each other spatially. By increasing the hard core radius this trend
+            will be more noticeable. The one-body density of the "ideal" system
+            is also included in the figure.
 

### 2018-04-06 project_1/article/article.tex
Spell check energy per particle section

@@
         be fairly certain that the method can be expanded upon to include
-        attributes outside the realm problems that are analytically solvable.
-        The systems with elliptic traps and interacting particles are such
-        systems.
+        attributes outside the realm of problems that are analytically solvable.
+        The systems with the elliptic traps and the interacting particles have
+        just such attributes.
 
@@
         purpose of this study is first and foremost to introduce variational
-        Monte Carlo as a tool in computational physics. We assume, however,
-        that this study is easy enough to expand into a study of alkali atoms like
-        $^{87}Rb$, $^{23}Na$ or $^7Li$.
+        Monte Carlo as a tool in computational physics. We assume, however, that
+        this study is a good starting point for studies of alkali atoms like
+        $^{87}$Rb, $^{23}$Na or $^7Li$.
 
@@
         For the non-interacting harmonic oscillator system there is no
-        difference between a new particle and new dimensions
-        \footnote{This is not entirely true in the case of our implementation of
-        the Metropolis algorithms as we draw a single random particle and move
-        all dimensions instead of iterating over all particles and all
+        difference between a new particle and a new dimension\footnote{This is
+        not entirely true in the case of our implementation of the
+        Metropolis-Hastings algorithm as we draw a single random particle and
+        move all dimensions instead of iterating over all particles and all
         dimensions.}. This means that the inclusion of a new particle or
@@
         and in the expression for the exact energy \autoref{eq:exact_energy}.
+
         This, however, is \emph{not} the case in the interacting system due to
@@
             \includegraphics[width=244px]{../data/figures/problem_e.pdf}
-            \caption{Local energies as a function of variational parameter
-            $\alpha$ for different number of particles.}
+            \caption{Expected energy per particle as a function of variational
+            parameter $\alpha$ for a varying number of particles.}
             \label{fig:interacting}
@@
 
-        The reason for the increase in energy per particle can be ascribed to the
-        repulsive interaction added by the Jastrow factor. The repulsive force can
-        be explain thusly; The Jastrow factor  increases the value of the
-        wavefunctions for any given particle if that particle is some distance away
-         from any other particle. An increased value in the wavefucntion is equivalent
-         to an increase in probability amplitude, again correpsonding to an increase in
-         probability density, giving rise the the fact that particles are more likely to
-         appear further from one another.
+        The reason for the increase in energy per particle can be ascribed to
+        the repulsive interaction added by the Jastrow factor. The repulsive
+        force can be explained thusly: the Jastrow factor  increases the value
+        of the local energy for any given particle if that particle is some
+        distance away from any other particle. The closer the particles are, the
+        more the local energy increases. The Jastrow factor will lower the
+        probability for the particles to get too close to one another by
+        reducing the probability density in the vicinity of other particles.
+        However, if there are many particles in the system they will be closer
+        to another particle and therefore increase the energy of the system.
 

### 2018-04-06 project_1/article/article.tex
Review section on low density system

@@
     \subsection{Low density system}
-        It is safe to say that the effect of adding a Jastrow factor is significant,
-        i.e. the increase in
-            energy for the system as the interaction is switched on is much higher
-            than the uncertainty of the energy. A simple look at \autoref{fig:interacting}
-            will be proof enough as the error bars of the figure are not discernible.
-        On the other hand, the energy added to the system by including interaction
-        is not very high compared to the energy of the system as a whole. For the
-        larger of the interacting systems we have looked at, with $n=100$ particles,
-        The interaction add only a mere $9\%$ extra energy (see \autoref{tab:interacting_optimal}).
-        Studying \autoref{fig:one_body_density_100} will reveal the same.
-
-        The reason for this stems from the assumption which justifies that
-        adding only two-body interaction is sufficient to model all interaction to a
-        reasonable extent. This assumption is justified by the fact that alkali and
-        atomic hydrogen systems\footnote{Systems similar to our toy model}
-        are dilute. For gases of alkali atoms confined in magnetic or optical traps
-        the average distance between atoms is much larger than the range of
-        inter-atomic interaction. The probability is high that atoms very seldom
-        interact and if so, just with one other atom.
+        It is safe to say that the effect of adding a Jastrow factor is
+        significant, i.e. the increase in energy for the system as the
+        interaction is switched on is much higher than the uncertainty of the
+        energy. A simple look at \autoref{fig:interacting} will be proof enough
+        as the error bars of the figure are not discernible.  On the other hand,
+        the energy added to the system by including interaction is not very high
+        compared to the energy of the system as a whole. For the larger of the
+        interacting systems we have looked at, with $N=100$ particles, the
+        interaction add only a mere $9\%$ extra energy compared to the
+        non-interacting system (see \autoref{tab:interacting_optimal}).
+        Studying \autoref{fig:one_body_density_100} will reveal the same trend
+        in the distribution of the particles.
+
+        The reason for this is two-fold.
+        \begin{enumerate}
+            \item The hard sphere radius is vanishingly small. By increasing the
+                hard sphere radius the particles will contribute more
+                significantly to the local energy. The particles will also be
+                more spread out thus occupying a larger configuration space.
+
+            \item The assumption which justifies that adding only two-body
+                interaction is sufficient to model all interactions to a
+                reasonable extent. This assumption is justified by the fact that
+                alkali and atomic hydrogen systems\footnote{Systems similar to
+                our toy model.} are dilute. For gases of alkali atoms confined
+                in magnetic or optical traps the average distance between atoms
+                is much larger than the range of inter-atomic interaction. The
+                probability is high that atoms very seldom interact and if so,
+                just with one other atom.
+        \end{enumerate}
 

### 2018-04-06 project_1/article/article.tex
Add section on stability of the program

@@
 
+    \subsection{Stability of the program}
+        An interesting thing would be to model large systems of many particles
+        and with large hard sphere radii. Unfortunately, due to the shape of the
+        single particle functions this turns out to be quite tricky. As the
+        single particle function consists of a sum squared positions from a
+        given center, many particles and/or spread out systems will yield a
+        large sum. This means that the Slater permanent grows smaller and
+        smaller making it much harder to keep the system confined. By lowering
+        the frequency and thus lowering the minimum $\alpha$ we can remedy some
+        of this behaviour.
+
 \section{Summary Remarks}

### 2018-04-06 project_1/article/article.tex
Spell check summary remarks

@@
 
-    In this study we have built a functional variational Monte Carlo solver framework for
-    a gas following Bose-Einstein statistics confined in an elliptic trap. Monte Carlo
-    methods is a broad class of computational algorithms, whereas the variational
-    Monte Carlo is a subset. Other applications of such methods within the physical
-    sciences includes, but is not limited to, quantum chromodynamics, the modeling of
-    aearodynamic forms, particle physics detector design and galaxy evolution.
-    Needless to say, Monte Carlo method is a valuable tool for a computational physicist.
-
-    To prove the
-    validity of the framework we tested it against an analytically solved case, an ideal
-    non-interacting system. Thereafter we have shown how our implementation can
-    handle large systems of interacting particles and perturbed traps. Even though
-    this study is somewhat theoretic in nature, we have discussed the possibility of
-    expanding our machinery to a case bearing a larger resemblance to a system
-    found in nature. To this end we have computed one-body densities for the
-    interacting system, which can be compared to real physical properties as
-    measured in an experiment.
+    In this study we have built a functional variational Monte Carlo solver
+    framework for a gas following Bose-Einstein statistics confined in an
+    elliptic trap. Monte Carlo methods is a broad class of computational
+    algorithms, whereas the variational Monte Carlo is a subset. Other
+    applications of such methods within the physical sciences includes, but is
+    not limited to, quantum chromodynamics, the modeling of aerodynamic forms,
+    particle physics detector design and galaxy evolution.  Needless to say,
+    Monte Carlo methods is a valuable tool for a computational physicist.
+
+    To prove the validity of the framework we tested it against an analytically
+    solvable case, an ideal non-interacting system. Thereafter we have shown how
+    our implementation can handle large systems of interacting particles and
+    perturbed traps. Even though this study is somewhat theoretic in nature, we
+    have discussed the possibility of expanding our machinery to a case bearing
+    a larger resemblance to a system found in nature. To this end we have
+    computed one-body densities for the interacting system, which can be
+    compared to real physical properties as measured in an experiment.
 

### 2018-04-06 project_1/article/article.tex
Simplify appendix on exact energy
Review appendix B

@@
         \prod_{i = 1}^N
-        \int\dd\vf{r}_i\exp(-2\alpha\vf{r}_i^2)
-        \\
-        &=
-        \prod_{i = 1}^N
         \prod_{j = 1}^d
@@
         \int
-        \para{
-            \prod_{k = 1}^d
+        \dd x_{il}\exp(-2\alpha x_{il}^2)x_{il}^2
+        \nonumber \\
+        &\qquad
+        \times
+            \int
+            \prod_{k \neq l}^d
             \dd x_{ik}\exp(-2\alpha x_{ik}^2)
-        } x_{il}^2
-        \nonumber\\
+        \nonumber \\
         &\qquad
@@
 
-\section{Computing the drift force and local energy of the full system}
-    Computing the gradient of the wavefunction we get
+\section{Computing the drift force and local energy of the interacting system}
+    Using the trial wavefunction from \autoref{eq:initial_trial_wavefunction} we
+    start by computing the gradient.
     \begin{align}
@@
     \end{align}
-    The gradient of the Slater permament for particle $k$ is given by
+    The gradient of the Slater permanent for particle $k$ is given by
     \begin{align}
@@
 
-    The Laplacian of the trial wavefunction is found by finding the
-    divergence of \autoref{eq:gradient_full_wavefunction}.
+    The Laplacian of the trial wavefunction is found by finding the divergence
+    of \autoref{eq:gradient_full_wavefunction}.
     \begin{align}
@@
     where the squared term came from taking the gradient of the trial
-    wavefunction.  To further simplify we divide by the trial
-    wavefunction. This yields
+    wavefunction.  To further simplify we divide by the trial wavefunction. This
+    yields
     \begin{align}
@@
     \end{align}
-    From here, the next step is to find the gradient and the Laplacian of
-    the single particle functions, $\phi_k$, and the interaction
-    functions $u_{km}$. For the single particle functions we use
-    Cartesian coordinates when finding the derivatives while for
-    the interaction functions we use spherical coordinates and do a
-    variable substitution. Beginning with the gradient of the single
-    particle functions we get
+    From here, the next step is to find the gradient and the Laplacian of the
+    single particle functions, $\phi_k$, and the interaction functions $u_{km}$.
+    For the single particle functions we use Cartesian coordinates when finding
+    the derivatives while for the interaction functions we use spherical
+    coordinates and do a variable substitution. Beginning with the gradient of
+    the single particle functions we get
     \begin{align}
@@
     \end{align}
-    Note that the subscripts on the unit vectors $\vf{e}_i$ are
-    \emph{not} the same as the subscripts used for its components. The
-    Laplacian yields
+    Note that the subscripts on the unit vectors $\vf{e}_i$ are \emph{not} the
+    same as the subscripts used for its components. The Laplacian yields
     \begin{align}
@@
     \end{align}
-    with $d$ as the dimensionality of the problem.
-    In order to derive the interaction functions we have to do a
-    variable substitution using $r_{km} = |\vf{r}_k - \vf{r}_m|$. We can
-    then rewrite the $\nabla_k$-operator as
+    with $d$ as the dimensionality of the problem.  In order to derive the
+    interaction functions we have to do a variable substitution using $r_{km} =
+    |\vf{r}_k - \vf{r}_m|$. We can then rewrite the $\nabla_k$-operator as
     \begin{align}
@@
         \end{align}
-        where $d$ is again the dimensionality of the problem. In total
-        we can state an intermediate version of the Laplacian occuring
+        In total we can state an intermediate version of the Laplacian occuring
         in the local energy as
@@
     \end{widetext}
-    Moving on to the derivatives of the interaction terms, $u_{km}$, to
+    Finally we compute the derivatives of the interaction terms, $u_{km}$, to
     get an explicit expression for the Laplacian.

### 2018-04-06 project_1/vmc/test/test_interface.py
Remove test directory

@@ -1,128 +1,0 @@
-from vmc.interface import PySimpleGaussian, PyHarmonicOscillator
-import numpy as np
-import pytest
-
-@pytest.fixture
-def wavefunction_params():
-    spread = 2.0
-    num_particles = 100
-    num_parameters = 1
-    num_dimensions = 3
-    mass = 1
-    omega = 1
-
-    return spread, num_particles, num_parameters, num_dimensions, mass, omega
-
-@pytest.fixture()
-def one_simple_gauss():
-    spread = 1.0
-    num_particles = 1
-    num_parameters = 1
-    num_dimensions = 1
-    mass = 1
-    omega = 1
-
-    return spread, num_particles, num_parameters, num_dimensions, mass, omega
-
-@pytest.fixture()
-def two_simple_gauss():
-    spread = 1.0
-    num_particles = 1
-    num_parameters = 1
-    num_dimensions = 2
-    mass = 1
-    omega = 1
-
-    return spread, num_particles, num_parameters, num_dimensions, mass, omega
-
-@pytest.fixture()
-def three_simple_gauss():
-    spread = 1.0
-    num_particles = 1
-    num_parameters = 1
-    num_dimensions = 3
-    mass = 1
-    omega = 1
-
-    return spread, num_particles, num_parameters, num_dimensions, mass, omega
-
-@pytest.fixture
-def hamiltonian_params():
-    mass = 1
-    omega = 1
-
-    return mass, omega
-
-def test_wavefunction(wavefunction_params):
-    spread, num_particles, num_parameters, num_dimensions, mass, omega = \
-            wavefunction_params
-
-    parameters = np.array([0.5])
-
-    wavefunction = PySimpleGaussian(
-            num_particles, num_dimensions, num_parameters, mass, omega,
-            spread=spread)
-
-    particles = wavefunction.get_particles()
-    assert ((particles >= -spread) & (particles <= spread)).all()
-
-    wavefunction.set_parameters(parameters)
-    assert np.allclose(parameters, wavefunction.get_parameters())
-
-
-@pytest.mark.run(after="test_wavefunction")
-def test_one_dimensional_ho(hamiltonian_params, one_simple_gauss):
-    mass, omega = hamiltonian_params
-    spread, num_particles, num_parameters, num_dimensions, mass, omega = \
-            one_simple_gauss
-
-    ho = PyHarmonicOscillator(mass, omega)
-    wavefunction = PySimpleGaussian(
[…]

### 2018-04-08 project_2/article/article.tex
Initial commit project 2

--- 2018-04-06 project_1/article/article.tex
@@
 
-\title{Variational Monte Carlo on Bosonic Systems}
+\title{Ground state energy of quantum dots using the coupled cluster method}
 \author{Winther-Larsen, Sebastian Gregorius}
@@
 \begin{abstract}
-    In this project we use Quantum Variational Monte Carlo on a toy model of a
-    bosonic gas  in an elliptic trap. The framework we have built is first
-    tested against analytic solutions of simple non-interacting systems. From
-    there the system is extended to include perturbed harmonic oscillator traps
-    and particle interaction. We compute one-body densities for the interacting
-    case and discuss why this tool is important and how our toy model has
-    similar attributes to real-world systems.
+    Something about coupled-cluster... Preferably doubles.
 \end{abstract}
@@
 
-\section{Introduction}
-    We will in this project study the Variational Monte Carlo method, and use it
-    to evaluate the ground state energy of a trapped, hard sphere Bose gas.
-
-    By building up an increasingly intricate system we eloquently illustrate the
-    strength of the Variational Monte Carlo methods. Any Monte Carlo method
-    draws on the fact that the computer will not complain when being tasked to
-    perform a very similar task up to several million times in rapid succession.
-    Monte Carlo methods is in essence a series of guesses where the current
-    guess is only a minor alteration of the one prior to it. The trick lies in
-    only accepting the subsequent guesses that improves upon our configuration
-    up to a certain likelihood.
-
-    The method we have employed is called a \emph{Variational} Monte Carlo
-    method, because we make an educated guess at a trial wavefunction which we
-    allow certain degrees of freedom called \emph{variational parameters}. We
-    shall see that the system converges to a proper \footnote{Global, as far as
-    we can tell.} minimum for only a certain set of parameters.
-
-    First, we provide the theoretical foundation for the Variational Monte Carlo
-    method. This includes explaining how Monte Carlo integration works, and
-    introducing the important quantities \emph{local energy}, \emph{drift force}
-    and \emph{one-body densities}.
-
-    Second, we outline in more detailed algebra the nature of the systems we
-    will be studying. For the non-interacting system we are able to derive an
-    exact variational energy to which we can compare the results from our
-    simulations. We furthermore give a description of an interacting system
-    implemented as if each particle is a hard sphere.
-
-    Third, we provide a thorough description of algorithms and statistical
-    methods employed in this study. Here we introduce the Metropolis-Hastings
-    algorithm. This algorithm can by a deft trick be expanded with the method of
-    \emph{importance sampling} yielding a more intelligent choice for the next
-    step in the Monte Carlo cycling.  The main point of note with regards to the
-    statistical analysis is that the data generated by the Monte Carlo cycling
-    is autocorrelated and computation of variance must be done with the
-    \emph{blocking method} for time series data.  As an automated way of finding
-    the optimal variational parameter, we sketch out the method of
-    \emph{gradient descent}.
-
-    Fourth, we put the machinery described in the previous sections to work. We
-    test against the actual values in a non-interacting system before analyzing
-    the more interesting perturbed-trap system with interacting particles.
-
-    Lastly, we highlight interesting aspects of this study in a discussion of
-    the results before ending on some summary remarks.
-
-\section{Variational Monte Carlo}
-    The variational principle states that for a given \emph{trial wavefunction},
-    $\ket{\Psi_T}$, the expectation value of the Hamiltonian, $H$, will be an
-    upper bound to the ground state energy, $E_0$ of the system.
-    \begin{align}
-        E_0 \leq E
-        = \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}},
-    \end{align}
-    The complicated part of the variational principle is finding the trial
-    wavefunction that minimizes the energy. The strength in this method lies in
-    introducing variational parameters in the trial wavefunction. Using these as
-    degrees of freedom we can minimize the energy with respect to these in order
-    to find an infimum to the ground state energy for the given trial
-    wavefunction and Hamiltonian. For a select few systems we can calculate this
-    analytically, but for the interesting systems we quickly hit an intractable
-    wall as the systems become analytically unsolvable. To get around this we
-    try to calculate the energy for a given Hamiltonian and a trial wavefunction
-    for different variational parameters numerically.
-
-    To determine if we have located a minimum we can use the variance of the
-    variational energy.
-    \begin{align}
-        \sigma^2
-        &=
-        \frac{
-            \bra{\Psi_T}H^2\ket{\Psi_T}
-        }{
-            \bra{\Psi_T}\ket{\Psi_T}
-        }
-        - \para{
-            \frac{
-                \bra{\Psi_T}H\ket{\Psi_T}
-            }{
-                \bra{\Psi_T}\ket{\Psi_T}
-            }
-        }^2.
-    \end{align}
-    If we have found the "true" trial wavefunction $\ket{\Psi_T}$ with the
-    correct variational parameters, i.e., the eigenfunction to the Hamiltonian
-    we get $\sigma^2 = 0$. We are thus interested in trying to find the
-    variational parameters in the trial wavefunction that minimizes the
-    variance.
-
-    \subsection{Monte Carlo integration}
-        Unfortunately the evaluation of the expectation value of the energy for
-        a given trial wavefunction involves a large integral.
-        \begin{align}
-            E
-            &= \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
-            =
-            \frac{\int\dd\vf{r}\Psi_T^{*}(\vf{r})H\Psi_T(\vf{r})}
-            {\int\dd\vf{r}|\Psi_T|^2},
-        \end{align}
-        where $\vf{r} = (\vf{r}_1, \vf{r}_2, \dots \vf{r}_N)$ are the
-        positions\footnote{This can include both spatial and spin parts, but we
-        will in the remainder of this project ignore spin.} of the particles
-        contained in the system.
-
-        To get around\footnote{Not exactly \emph{get around} per se. We still
-        have to evaluate the integral, but we make the problem solvable at all.}
-        the trouble of evaluating the multidimensional-integral we employ the
-        method of Monte Carlo integration. This consists of approximating an
-        integral of the expectation value of a certain quantitity by a sum.
-        \begin{align}
-            \expval{x}
-            &=
-            \int_{\mathbb{R}}xp(x)\dd x
-            \approx
-            \frac{1}{N}\sum_{i = 1}^N x_ip(x_i),
-        \end{align}
-        where $x$ is a random variable with a probability density function
-        $p(x)$ and $\{x_i\}$ for $i \in \{1, \dots, N\}$ is the set of drawn
-        random variables. Here $N$ is the number of Monte Carlo samples chosen
-        to approximate the integral. By the law of large numbers we expect that
-        \begin{align}
-            \lim_{N \to \infty}\frac{1}{N}\sum_{i = 1}^N x_ip(x_i)
-            = \expval{x}.
-        \end{align}
-
-        For us to be able to take Monte Carlo integration to the quantum realm
-        we have to construct a probability density function from the
-        wavefunction. Luckily the wavefunction squared is just such a quantity.
-        Given a normalized wavefunction $\ket{\psi}$ we can compute the
-        expectation value of a quantum operator $\mathcal{O}$ using Monte
-        Carlo integration.
-        \begin{align}
-            \expval{\mathcal{O}}
-            &= \bra{\psi}\mathcal{O}\ket{\psi}
-            =
-            \int\dd\vf{r}
-            \psi^{*}(\vf{r})\mathcal{O}\psi(\vf{r})
-            \\
-            &=
-            \int\dd\vf{r}
-            |\psi(\vf{r})|^2
-            \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})},
-        \end{align}
-        where we have multiplied and divided by the wavefunction. We now define
-        the probability density function, $p(\vf{r})$, as
-        \begin{align}
-            p(\vf{r}) \equiv |\psi(\vf{r})|^2,
-        \end{align}
-        and the local quantum operator
-        \begin{align}
-            \mathcal{O}_L(\vf{r})
-            &\equiv \frac{\mathcal{O}\psi(\vf{r})}{\psi(\vf{r})}.
-        \end{align}
-        We see that if the chosen wavefunction is an eigenstate of the quantum
-        operator, the local quantum operator becomes a constant.  We can thus
-        approximate the expecatation value of the quantum operator using Monte
-        Carlo integration.
-        \begin{align}
-            \expval{O}
-            &=
-            \int\dd\vf{r}
-            p(\vf{r})
-            \mathcal{O}_L(\vf{r})
-            \approx
-            \frac{1}{N}\sum_{i = 1}^N
-            p(\vf{r}_i)\mathcal{O}_L(\vf{r}_i).
-        \end{align}
-        In the case of a non-normalized wavefunction we need to divide by a
-        normalization factor in the probability density function.
-
-    \subsection{Local energy}
-        As is mostly the case in physics, the quantity we are looking for is the
-        energy. We therefore define the \emph{local energy} by
-        \begin{align}
-            E_L(\vf{r})
-            &= \frac{H\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
-        \end{align}
-        The energy of the system is thus found by computing the expectation
-        value of the local energy.
-        \begin{align}
-            \expval{E_L}
-            =
-            \frac{
-                \int\dd \vf{r}|\Psi_T(\vf{r})|^2 E_L(\vf{r})
-            }{
-                \int\dd \vf{r}|\Psi_T(\vf{r})|^2
-            }
-            \approx
-            \frac{1}{N}\sum_{i = 1}^N
-            p(\vf{r}_i)E_L(\vf{r}_i).
-        \end{align}
-        To determine if we have found a minimum we compute the variance of the
-        expectation value of the local energy.
-        \begin{align}
-            \sigma^2 = \langle E_L^2\rangle - \expval{E_L}^2.
-        \end{align}
-        One of the most computationally intensive parts of the Variational Monte
-        Carlo algorithm will be to compute $E_L$. We will therefore find an
-        analytical expression for $E_L$ in terms of the trial wavefunctions and
-        the system we are exploring.
-
-    \subsection{The drift force}
-        A disadvantage in the use of the brute force Metropolis-Hastings
-        algorithm is that we might be spending much computational resources in
-        an uninteresting part of configuration space. To make smarter moves we
-        will add importance sampling (which will be discussed in due time) to
-        the algorithm. This alteration of the algorithm needs a new quantity
-        called the \emph{drift force} of the system\footnote{Also known as the
-        \emph{quantum force}.}.
-        \begin{align}
-            \vf{F}(\vf{r})
-            &=
-            \sum_{k = 1}^N
-            \vf{F}_k(\vf{r})
-            =
-            \sum_{k = 1}^N
-            \frac{2\nabla_k\Psi_T(\vf{r})}{\Psi_T(\vf{r})}.
-            \label{eq:drift_force}
-        \end{align}
-        Using the drift force we are able to move towards more interesting parts
-        of configuration space as we are following the gradient of the
-        wavefunction. We will mainly be interested in the drift force of a
-        single particle $k$ as the Monte Carlo sampling involves choosing a
-        single particle at a time.
-
-    \subsection{One-Body Densities}
-        In order to make some conclusion towards where the particles in our
-        systems are most prone to exist, we need a good way to visualise our
-        wavefunctions. If we are to visualize the combined wavefunction of two
-        three-dimensional particles we need four dimensions per particle and can
-        subtract one dimension as we look at a single probability density
-        yielding a grand total of seven dimensions. The fact that we will be
-        studying systems of up to $N = 500$ particles only adds to the headache.
-        We therefore introduce the single-particle density function, or
-        \emph{one-body density} $\rho(\vf{r})$. The one-body density is defined
-        by the integral over all coordinates except
-        one\cite{kristiansen2017time},
-        \begin{align}
-            \rho(\vf{r}_1)
-            \equiv
-            \int \dd\vf{r}_2 \dots \dd\vf{r}_N
-            \abs{\Psi(\vf{r}_1, \dots, \vf{r}_N)}^2.
-        \end{align}
-        The one-body density represents the distribution of all the particles in
-        the system. Furthermore, $\rho(\vf{r}_1)\dd\vf{r}_1$ will yield the
-        probability of finding any particle within the volume element
-        $\dd\vf{r}_1$. By convention the one-body density should be normalized
-        such that the following relation holds,
-        \begin{align}
-            N = \int \rho(\vf{r}_1)\dd\vf{r}_1.
-        \end{align}
-        In other words, the integral over the entire volume in
-        $\vf{r}_1$-direction gives us the total number of particles in our
-        system. This can be a bit odd at first sight as we are used to getting
-        unity when we integrate over the entire probability
-        density.\cite{hogberget2013quantum}
-
-        It is best to think of the one-body density as the multivariate
-        probability density for the entire system collapsed down to one
-        coordinate in a representation than looks very familiar to every
-        physicist with a basic knowledge of quantum mechanics.
-
-\section{Systems}
-    To model the trapped bosonic gas particles we use the potential
-    \begin{align}
-        v(\vf{r})
-        &=
-        \begin{cases}
-            \half m\omega^2r^2 & (\text{S}), \\
-            \half m \bigl[
-                \omega^2(x^2 + y^2) + \omega_z^2z^2
-            \bigr] & (\text{E}),
-        \end{cases}
-    \end{align}
-    where we can choose between a spherical (S) or an elliptical (E) harmonic
-    oscillator trap. The full Hamiltonian of the system is given by
-    \begin{align}
-        H = \sum_{i = 1}^{N}h(\vf{r}_i) + \sum_{i < j}^{N}w(\vf{r}_i, \vf{r}_j),
-    \end{align}
-    where the single particle one-body operator, $h(\vf{r}_i)$, is given by
-    \begin{align}
-        h(\vf{r}_i) = -\frac{\hbar^2}{2m}\nabla_i^2
-        + v(\vf{r}_i),
-    \end{align}
-    for equal masses. The two-body interaction operator, $w(\vf{r}_i,
-    \vf{r}_j)$, is
-    \begin{align}
-        w(\vf{r}_i, \vf{r}_j)
-        = \begin{cases}
-            \infty & |\vf{r}_i - \vf{r}_j| \leq a, \\
-            0 & |\vf{r}_i - \vf{r}_j| > a,
-        \end{cases}
-        \label{eq:two-body_interaction}
-    \end{align}
-    where $a$ is the hard sphere of the particle.  The trial wavefunction we
-    will be looking at is given by
-    \begin{align}
-        \Psi_T(\vf{r})
-        &= \Phi_T(\vf{r})
-        J(\vf{r}),
-        %\\
-        %&= \Biggl(
-        %    \prod_{i = 1}^N g(\alpha, \beta, \vf{r}_i)
-        %\Biggr)
-        %\prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k),
-        \label{eq:initial_trial_wavefunction}
-    \end{align}
-    where $\vf{r} = (\vf{r}_1, \dots, \vf{r}_N)$. The Slater permanent,
-    $\Phi_T(\vf{r})$, is given by the $N$ first single-particle functions
-    \begin{align}
-        \Phi_T(\vf{r})
-        &=
-        \prod_{i = 1}^N \phi(\vf{r}_i)
-        = \prod_{i = 1}^N\exp\bigl[
-            -\alpha(x_i^2 + y_i^2 + \beta z_i^2)
-        \bigr].
-    \end{align}
-    Here $\alpha$ is the variational parameter of the trial
-    wavefunction.\footnote{We can also treat $\beta$ as a variational parameter,
-    but we will in this project limit ourselves to a single variational
-    parameter, i.e., $\alpha$.} The Jastrow factor, $J(\vf{r})$ contributes with
-    correlations between the particles and is represented as
-    \begin{align}
-        J(\vf{r})
-        &=
-        \prod_{j < k}^N f(a, \vf{r}_j, \vf{r}_k),
-    \end{align}
-    where the correlation wavefunctions are given by
-    \begin{align}
-        f(a, \vf{r}_j, \vf{r}_k)
-        &=
-        \begin{cases}
-            0 & |\vf{r}_j - \vf{r}_k| \leq a, \\
-            \Bigl(
-                1 - \frac{a}{|\vf{r}_j - \vf{r}_k|}
-            \Bigr) & |\vf{r}_j - \vf{r}_k| > a.
-        \end{cases}
-        \label{eq:correlation_wavefunction}
-    \end{align}
-    We will for brevity define the notation
-    \begin{gather}
-        \phi_i \equiv \phi(\vf{r}_i),
-        \\
-        r_{jk} \equiv \abs{\vf{r}_j - \vf{r}_k}.
-    \end{gather}
-    In practice \autoref{eq:two-body_interaction} is
-    unnecessary as the situtation where $r_{jk} \leq a$ automatically yields a
-    probability density of zero from \autoref{eq:correlation_wavefunction} thus
-    rejecting the state completely.
-
-    \subsection{Non-interacting harmonic oscillators}
-        We start by looking at a simple system of non-interacting harmonic
-        oscillators, where $a = 0$ and $\beta = 1$. This means that there is no
-        hard-shell interaction between the particles and that we use a spherical
-        potential trap.  We thus get the trial  wavefunction
-        \begin{align}
-            \Psi_T(\vf{r})
-            = \Phi_T(\vf{r})
-            = \prod_{i = 1}^N \exp\para{
-                -\alpha |\vf{r}_i|^2
-            }.
-            \label{eq:trial_simple_gaussian}
-        \end{align}
-        As $a = 0$ the interaction term, $w(\vf{r}_i, \vf{r}_j)$, vanishes and
-        the Hamiltonian is given by (in the spherical case)
-        \begin{align}
-            H &= \sum_{i = 1}^N h(\vf{r}_i)
-            = \sum_{i = 1}^N \Biggl(
-                -\frac{\hbar^2}{2m}\nabla_i^2
-                + \half m \omega^2 r_i^2
-            \Biggr).
-        \end{align}
-        To find the drift force and the local energy we have to compute the
-        gradient and the Laplacian of the trial wavefunction. The gradient is
-        given by
-        \begin{align}
-            \nabla_k\Psi_T(\vf{r})
-            &= -2\alpha \vf{r}_k\Psi_T(\vf{r}),
-        \end{align}
-        whereas the Laplacian yields
-        \begin{align}
-            \nabla^2_k\Psi_T(\vf{r})
-            &= \big(-2d\alpha + 4\alpha^2 r_k^2\bigr)\Psi_T(\vf{r}).
-            \label{eq:laplacian_simple_gaussian}
-        \end{align}
-        Here $d$ is the dimensionality of the problem determined by $\vf{r}_k
-        \in \mathbb{R}^d$. We can thus use the gradient to find an expression
-        for the drift force for particle $k$.
-        \begin{align}
-            \vf{F}_k(\vf{r})
-            &= -4\alpha\vf{r}_k.
-        \end{align}
-        Using the Laplacian we can compute the kinetic term in the expression
-        for the local energy. We get
-        \begin{align}
-            E_L(\vf{r})
-            &=
-            \sum_{i = 1}^N
-            \Biggl(
-                -\frac{\hbar^2}{2m}
-                \bigl[
-                    -2d\alpha + 4\alpha^2 r_i^2
-                \bigr]
-                + \half m\omega^2 r_i^2
-            \Biggr).
-        \end{align}
-        In natural units, $\hbar = c = 1$, and with the mass set to unity this
-        reduces to
-        \begin{align}
-            E_L(\vf{r})
-            &=
-            \alpha dN
-            + \biggl(
-                \half\omega^2
-                - 2\alpha^2
-            \biggr)
-            \sum_{i = 1}^N r_i^2.
-            \label{eq:closed_form_natural_units_local_energy}
-        \end{align}
-        It is worth noting that for $\alpha = \omega/2$\footnote{Note that
-        $\alpha$ is required to be positive.} we get a stable value for the
-        local energy.
-        \begin{align}
-            E_L(\vf{r}; \alpha = \omega/2)
-            &\equiv
-            E_L
-            =
-            \frac{\omega d N}{2},
-            \label{eq:minimum_energy_simple}
-        \end{align}
-        which turns out to be the exact minimum for the ground state energy of
-        this system as we shall see.
-
-        \subsubsection{Exact variational energy}
-            As the system is non-interacting and consisting of Gaussians we can
-            find an expression for the exact energy as a function of the
-            variational parameter $\alpha$, i.e.,
-            \begin{align}
-                E(\alpha)
-                &=
-                \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
-                =
-                \left(
-                    \frac{\hbar^2 \alpha}{2m}
-                    + \frac{m\omega^2}{8\alpha}
-                \right)dN.
-                \label{eq:exact_energy}
-            \end{align}
-            By minimizing this expression with respect to $\alpha$ we get
-            \begin{align}
-                \dod[]{E(\alpha)}{\alpha} = 0
-                \implies
-                \alpha_0 = \frac{m\omega}{2\hbar},
-            \end{align}
-            which in natural units reduces to $\alpha_0 = \omega/2$. The energy
-            at this value of $\alpha$ (in natural units) is then
-            \begin{align}
-                E(\alpha_0)
-                &=
-                \frac{\omega dN}{2},
-            \end{align}
-            in perfect accordance with \autoref{eq:minimum_energy_simple} which
-            yields high hopes for the latter equation and the expected minimum.
-
-    \subsection{Interacting hard sphere bosons}
-        Moving to the full elliptical system for $\beta \neq 1$ and setting $a
-        \neq 0$ we use the trial wavefunction from
-        \autoref{eq:initial_trial_wavefunction}. We rewrite the Jastrow factor
-        to
-        \begin{align}
-            J(\vf{r})
-            &=
-            \exp\para{
-                \sum_{j < l}^N u(r_{jl})
-            },
-        \end{align}
-        where
-        \begin{align}
-            u(r_{jk})
-            = \ln\bigl[f(a, \vf{r}_j, \vf{r}_k)\bigr]
-            \equiv u_{jk}
-        \end{align}
-        The drift force for particle $k$ in this system becomes
-        \begin{align}
-            \vf{F}_k(\vf{r})
-            &=
-            2\Biggl(
-                \frac{\nabla_k\phi_k}{\phi_k}
-                + \sum_{m \neq k}^N
-                \nabla_k u_{km}
-            \Biggr).
-        \end{align}
-        The Laplacian of particle $k$ divided by the trial wavefunction is then
-        \begin{widetext}
-            \begin{align}
-                \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
-                &=
-                \frac{\nabla_k^2\phi_k}{\phi_k}
-                + 2\frac{\nabla_k\phi_k}{\phi_k}
-                \sum_{m \neq k}^N
-                \frac{\Delta\vf{r}_{km}}{r_{km}}
-                \dpd[]{u_{km}}{r_{km}}
-                + \sum_{m\neq k}^N
-                \para{
-                    \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
-                    + \dpd[2]{u_{km}}{r_{km}}
-                }
-                +
-                \para{
-                    \sum_{m \neq k}^N
-                    \frac{\Delta\vf{r}_{km}}{r_{km}}
-                    \dpd[]{u_{km}}{r_{km}}
-                }^2,
-                \label{eq:laplacian_full_system}
-            \end{align}
-        \end{widetext}
-        where $\Delta \vf{r}_{km} = \vf{r}_k - \vf{r}_m$ and $d$ is the
-        dimensionality. Evaluating the gradient and the Laplacian of the single
-        particle functions yields
-        \begin{gather}
-            \frac{\nabla_k\phi_k}{\phi_k}
-            =
-            -2\alpha
-            (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k),
-            \\
-            \frac{\nabla_k^2\phi_k}{\phi_k}
-            =
-            -2\alpha
-            \big(
-                d - 1 + \beta
-            \bigr)
-            + 4\alpha^2
-            \bigl(
-                x_k^2 + y_k^2 + \beta^2z_k^2
-            \bigr).
-        \end{gather}
-        Repeating this for the correlation functions yields
-        \begin{align}
-            \dpd[]{u_{km}}{r_{km}}
-            &=
-            \frac{a}{r_{km}(r_{km} - a)},
-            \\
-            \dpd[2]{u_{km}}{r_{km}}
-            &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.
-        \end{align}
-        The local energy is thus
-        \begin{align}
-            E_L(\vf{r})
-            &=
-            \sum_{k = 1}^N
-            \brac{
-                -\frac{\hbar^2}{2m}
-                \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
-                + v(\vf{r}_k)
-            },
-        \end{align}
-        where the Laplacian is given by \autoref{eq:laplacian_full_system}.  We
-        will avoid writing out the full expression for the local energy for this
-        system due to the size of the equation.
-
-        \subsubsection{Scaling the system}
-            We now introduce a scaled distance $\vf{r}' = \vf{r}/a_{\text{ho}}$
-            \cite{dubois2001bose}, where
-            \begin{align}
-                a_{\text{ho}}
-                &=
-                \sqrt{\frac{\hbar}{m\omega}}.
-            \end{align}
-            We can then rewrite the Hamiltonian for the elliptic potential in terms
-            of this new distance. By doing a variable substitution for each
-            direction in the Laplace operator we get
-            \begin{align}
-                \nabla_k^2
-                &=
-                \frac{1}{a_{\text{ho}}^2}{\nabla_k'}^2.
-            \end{align}
-            The one-body part of the Hamiltonian then yields
-            \begin{align}
-                H
-                &=
-                \sum_{k = 1}^N
-                \brac{
-                    -\frac{\hbar^2}{2m}\nabla_k^2
-                    + \half m\brak{
-                        \omega^2(x_k^2 + y_k^2)
-                        + \omega_z^2 z_k^2
-                    }
-                }
-                \\
-                &=
-                \frac{\hbar\omega}{2}
-                \sum_{k = 1}^N
-                \brac{
-                    -{\nabla_k'}^2
-                    +
-                    \brak{
-                        {x_k'}^2 + {y_k'}^2 + \lambda^2{z_k'}^2
-                    }
-                },
-                \label{eq:scaled_hamiltonian}
-            \end{align}
-            where we have introduced the dimensionless frequency $\lambda =
-            \omega_z/\omega$. The wavefunction does not need to be scaled as the
-            exponentials are already dimensionless. We only need to use the
-            scaled hard core radius $a' = a/a_{\text{ho}}$. We are interested in
-            using the radius $a' = 0.0043$ as in the references
-            \cite{dubois2001bose} and \cite{nilsen2005vortices}.
-
-\section{Algorithms}
-    In this project we rely on a Monte Carlo approach of random sampling to
-    obtain numerical results.  We simulate random walks over a volume in order
-    to find optimal parameters in our trial wavefunctions.  The most common of
-    such methods, which we make use of herein, is the Metropolis-Hastings
-    algorithm.
-
-    \subsection{The Metropolis-Hastings Algorithm}
-        The Metropolis-Hastings algorithm can in our particular situation be
-        condensed down to the following steps:
-
-        \begin{enumerate}
-            \item The system is initialised by a certain number $N$ of randomly
-                generated positions, $\vf{r} = (\vf{r}_1, \dots, \vf{r}_N)$,
-                which represent the particles\footnote{Also known as
-                \emph{random walkers}.}. This allows us to evaluate the
-                wavefunction at these points and compute the local energy
-                $E_L(\vf{r})$.
-
-            \item The initial configuration is changed by setting a new position
-                for one of these particles. The particle is picked at random and
-                is only allowed to move a predetermined step length.
-
-            \item A ratio, $q$, between the probability density of the new
-                wavefunction and the previous probability density is computed.
-                We accept the step if $q \geq 1$, else we accept the step with a
-                probability $p$. The latter is done by drawing a random uniform
-                number, $x \in [0, 1]$, and checking if $q > x$.  This
-                transition probability decides if the particle move is rejected
-                or accepted.
-
-            \item If the particle movement is accepted we compute the local
-                energy for the new system. If not, we reset the system to the
-                previous state and reuse the last local energy.
-
-            \item Accumulate the local energy. Repeat the steps until
-                convergence is reached or a set number of iterations has been
-                done. When finished divide the accumulated energy by the number
-                of cycles to compute the expectation value of the energy.
-
-        \end{enumerate}
-
-        The algorithm described above can be applied in an "exhaustive" search
-        of the parameter space in order to find the optimal parameters.  Whether
-        a proposed move is accepted or not is determined by a \emph{transition
-        probability}. The strength of the algorithm lies in that the transition
-        algorithm need not be known. For example, the simplest case is to accept
-        the new state, i.e., the new position for the random walker, if the
-        ratio
-        \begin{align}
-            q(\vf{r}_{i + 1}, \vf{r}_i)
-            &=
-            \frac{\left|\Psi_T(\vf{r}_{i + 1})\right|^2}
-            {\left|\Psi_T(\vf{r}_{i})\right|^2},
-        \end{align}
-        where $\vf{r}_{i + 1}$ are all the positions at step $i + 1$, is greater
-        than a uniform probability $p \in [0, 1]$. This transition criteria
-        results in what we call the brute force Metropolis-Hastings algorithm.
-        By counting the number of accepted steps and dividing by the total
-        number of Monte Carlo cycles we get the \emph{acceptance probability},
-        denoted $A$.  The acceptance probability helps in choosing a decent
-        length for the proposed new step. If the acceptance probability is very
-        low we can lower the step length to get more accepted steps and vice
-        versa.
-
-        \subsubsection{Initial distribution of particles}
-            Prior to performing the Metropolis-Hastings algorithm we have to
-            have an initial distribution of the particles. As the algorithm is
-            supposed to move particles in the right direction this distribution
-            can be \emph{somewhat} arbitrary. There are however a few caveats to
-            this approach.
-
-            \begin{enumerate}
-                \item The distribution is unphysical. This results in the
-                    particles taking up positions which can yield bogus values
-                    for the local energy. A remedy is to allow for some
-                    \emph{thermalization steps}, i.e., scatter the particles and
-                    run the algorithm for a set number of steps prior to
-                    starting the actual sampling.
-
-                \item The initial \emph{spread} of the particles is too small or
-                    too large. The former can lead to an acceptance ratio of
-                    zero if no steps are accepted due to overlap. This is
-                    something that can happen to the interacting system we are
-                    studying as an overlap between the particles makes the
-                    wavefunction evaluate to zero thus rejecting all steps in
-                    the ratio test.  The latter can lead to the system never
-                    reaching a true state as they are scattered too far away
-                    from the interesting regions in the configuration space.
-
-                \item The system imposes some restrictions on the distribution.
-                    This can mean that we have to distribute the particles in a
-                    certain way to make sure that we don't break the
-                    restrictions. This resembles the former case when the spread
-                    is too small for our systems.
-            \end{enumerate}
-
-            In the case of the interacting system we are looking at we have to
-            take some measures to make sure that the initial spread does not
-            break the Monte Carlo cycles. We therefore distribute all the
-            particles on a grid with a distance larger than the diameter of the
-            hard sphere before letting the system thermalize. After the
-            thermaliztion we start the sampling.
-
-        \subsubsection{Importance Sampling}
-            A problem with the naïve Metropolis-Hastings sampling approach
-            described above is that the sampling of position space is done with
-            no regard for where we are likely to find a particle. This problem
-            can be remedied through \emph{importance sampling}.  It is
-            reasonable to assume that the particles we erratically scatter in
-            space are prone to movement towards the peaks of the probability
-            density as dictated by the wave function. Consider therefore the
-            Fokker-Planck equation,
-            \begin{align}
-                \dpd[]{\Psi_T}{t}
-                &=
-                D \nabla\cdot
-                \left(
-                    \nabla
-                    - \vf{F}
-                \right) \Psi_T,
-                \label{eq:fokker_planck}
-            \end{align}
-            which describes the evolution in time of a probability density
-            function. In our case this is the trial wavefunction
-            $\Psi_T$\footnote{The trial wavefunction is not a probability
-            density, but the Fokker-Planck equation is robust enough to handle
-            the wavefunction as is.}.  Originaly an equation that models
-            diffusion, we have a diffusion term $D$ and the drift force
-            \autoref{eq:drift_force}. In our case the diffusion term $D$ is
-            simply $1/2$ from the kinetic energy (in natural units).
-
-            We use the Langevin equation to find the new position of the
-            particle.
-            \begin{align}
-                \dpd[]{\vf{r}}{t}
-                &=
-                D\vf{F}(\vf{r}) + \vfg{\eta},
-                \label{eq:langevin}
-            \end{align}
-            where $\vfg{\eta}$ is a uniformly distributed stochastic variable
-            for each dimenion.  Solving Langevin's equation by Euler's method
-            gives a recursive relation for the subsequent new positions of a
-            particle.
-            \begin{align}
-                \vf{r}_{i + 1}
-                &=
-                \vf{r}_i + D\vf{F}(\vf{r}_i)\delta t
-                + \vfg{\xi}\sqrt{\delta t},
-            \end{align}
-            for a given time step $\delta t$
-            \footnote{
-                Bear in mind that \autoref{eq:langevin} is only valid as $\Delta
-                t \to 0$, a property stemming from the use of Euler's method.
-            }
-            and a normally distributed stochastic variable $\vfg{\xi}$.
-
-            Now we need to change the ratio computation in the
-            Metropolis-Hastings algorithm to something that takes the new
-            sampling method into account.
-            \begin{align}
-                q(\vf{r}_{i + 1}, \vf{r}_i)
-                &=
-                \frac{
-                    G(\vf{r}_{i}, \vf{r}_{i + 1}, \delta t)
-                    \left|\Psi_T(\vf{r}_{i + 1})\right|^2
-                }
-                {
-                    G(\vf{r}_{i + 1}, \vf{r}_{i}, \delta t)
-                    \left|\Psi_T(\vf{r}_{i})\right|^2
-                },
-            \end{align}
-            where $G$ is the Green's function of the Fokker-Planck equation
-            given by
-            \begin{align}
-                G(\vf{r}_{i}, \vf{r}_{i + 1}, \delta t)
-                &=
-                \exp\Biggl(
-                    -\frac{\bigl[
-                        \vf{r}_{i}
-                        - \vf{r}_{i + 1}
-                        - D\vf{F}(\vf{r}_{i + 1})\delta t
-                    \bigr]^2}{4D\delta t}
-                \Biggr)
-                \nonumber \\
-                &\qquad
-                \times
-                \frac{1}{(4\pi D\delta t)^{dN/2}},
-            \end{align}
-            where $d$ is the dimensionality. The new ratio, $q$, is still
-            subjected to the same ratio test as in the brute force
-            Metropolis-Hastings algortihm.
-
-    \subsection{Statistical Analysis}
-        If the results of the metropolis sampling were completely uncorrelated,
-        it would be enough to compute the standard deviation in a familiar way,
-        \begin{equation}
-            \sigma = \sqrt{\frac{1}{N}(\ev{E^2_L} - \ev{E_L}^2},
-            \label{eq:standard_deviation}
-        \end{equation}
-        where $N$ is the number of samples, or Monte-Carlo cycles, in the
-        experiment.  However, it is reasonable to assume that the data we are
-        dealing with in this study is liable to suffer from
-        \emph{autocorrelation} and \autoref{eq:standard_deviation} does not
-        hold. The prevailing definition of autocorrelation in a data stream or
-        signal is correlation between a delay of the signal and the original
-        signal. One would be interested to find the delay, or lag, in the signal
-        at which the "self-correlation" is highest. We refer to this spacing as
-        $d$, and define the following correlation function,
-        \begin{equation}
-            f_d
-            = \frac{1}{n -d} \sum_{k=1}^{n-d}
-            (x_k - \bar{x}_n)(x_{k+d} - \bar{x}_n).
-            \label{eq:correlation_f_d}
-        \end{equation}
-        The keen reader would have noticed that the function $f_d$ in
-        \autoref{eq:correlation_f_d} would be equal to the sample variance for
-        $d=0$. We can now define the \emph{autocorrelation function}
-        \begin{equation}
-            \kappa_d = \frac{f_d}{\text{Var}(x)},
-            \label{eq:autocorrelation_function}
-        \end{equation}
-        which is equal to $1$ if the data exhibits no autocorrealtions, equating
-        to $d=0$. From the autocorrelation function in
-        \autoref{eq:autocorrelation_function} we in turn define the
-        \emph{autocorrelation time},
-        \begin{equation}
-            \tau = 1 + 2 \sum_{d=1}^{n-1}\kappa_d.
-        \end{equation}
-        Notice that the autocorrelation time is $1$ for a correlation free
-        experiment.
-
-        We are now able to make a correction to the expression for the standard
-        deviation improving on \autoref{eq:standard_deviation} by taking
-        correlation into account,
-        \begin{equation}
-            \sigma =\sqrt{\frac{1 + 2\tau / \Delta t}{N}
-            (\ev{E^2_L} - \ev{E_L}^2)},
-        \end{equation}
-        where $\Delta t$ is the time between each sample. The main problem at
-        this point is to find $\tau$, which we do not know for any given system
-        and it is generally very expensive to compute. In order to find a good
-        estimate of $\tau$ we use a procedure called \emph{blocking}.
-
-        \subsubsection{Blocking}
-            In the method of blocking we group the samples into blocks of
-            increasing size. If one where to compute the standard deviation for
-            each block, one should see the variance increase with the block
-            size.  The standard deviation would only increase up to a certain
-            point, from whence it would stay almost constant. What is happening
-            is that we have reached a point where a particular sample from one
-            block is no longer correlated with a corresponding sample from an
-            adjacent block.  The block size for this point of convergence now
-            functions as an estimate for the autocorrelation time $\tau$.
-
-            Instead of going by this "manual" method of looking at charts, we
-            will construct a test statistic and let the computer do the
-            necessary considerations in a more "automatic way.
-
-            The easiest way to do this is by use of the \emph{automatic blocking
-            scheme}, nicely illustrated with a flow chart in
-            \autoref{fig:blocking}. The input parameter for the method is an
-            array of the local energies, but it can be any array of time series
-            data, which is why it is herein referred to as $\mathbf{X}$. Any
-            array of sequential data, where any particular data point depends on
-            the previous data point inhibits all attributes of a time series.
-            Our array of local energies is therefore most definitely a time
-            series.
-
-            From $\mathbf{X}$ we compute the variance and sample covariance
-            between the series itself and a one-point lag of the series. This
-            covariance is usually referred to as the first order autocovariance.
-            \begin{align}
-                \gamma_i
-                &= \frac{1}{n}
-                \left(
-                    \sum_{k=1}^n
-                    \left((\mathbf{X}_i)_k - \bar{\mathbf{X}}\right)
-                \right)
-                \nonumber \\
-                &\qquad
-                \times
-                \left(
-                    \sum_{k=0}^{n-1}
-                    \left((\mathbf{X}_i)_k - \bar{\mathbf{X}}\right)
-                \right).
-            \end{align}
-            Note that $i$ represents the iteration in the blocking process
-            whereas $k$ is the index into the array $\mathbf{X}_i$.
-
-            Now we "block"! The next step is to perform a transformation of the
-            our time series $\mathbf{X}_i$, where we end up with an array that
-            is half the length of the original time series,
-            \begin{equation}
-                (\mathbf{X}_i)_k
-                = \frac{1}{2}\left((\mathbf{X}_i)_{2k-1}
-                + (\mathbf{X}_i)_{2k} \right).
-            \end{equation}
-            This process is were the term "blocking" stems from. There are a
-            number of ways to perform the blocking - we could have picked the
-            data points used in the new ($i+1$) array randomly, or more orderly.
-            Here we have picked them sequentially. We have obtained $n/2$ new
-            "primed" stochastic variables. Assuming that the reader is playing
-            close attention, it is now needless to say that it is much easier to
-            perform these sequential blocking steps if we have an array of
-            length $n= 2^d$, where $d$ is integer. Otherwise, problems would
-            arise because one would have arrays of different lengths when the
-            original array is split in two. The ailment is easy to remedy by
-            excluding an observation from the longest resulting array after
-            splitting.
-
-            For each new time series $\mathbf{X}_{i+1}$ a new sample variance
-            and sample covariance is computed.  The blocking transformation is
-            allowed to continue until we run out of data, i.e. the length of
-            $\mathbf{X}_i < 2$. In other words, there is only a single number
-            left.  We now compute a test statistic,
-            \begin{equation}
-                M_j
-                = \sum_{k=j}^{d-1}
-                n_k \frac{\left[(n_k - 1)
-                \sigma^2_k/n_k^2 + \gamma_k\right] ^2}{\sigma_k^4},
-            \end{equation}
-            As luck would have it, this number $M$ is   $\chi^2$-distributed.
-            The following step would be to look up a statistical table and find
-            the first $k$, such that $M_k \leq q_{d-k}(1-\alpha)$, where $q$ is
-            the statistical limit and $\alpha$ is your favorite  significance
-            level. Finally, the correct sample variance is $\sigma_k^2/n_k$. In
-            the rest of the project we use the labeling
-            \begin{align}
-                \sigma_b^2 \equiv \frac{\sigma_k^2}{n_k},
-            \end{align}
-            for the correct sample variance found from the blocking method.
-            Again, see \autoref{fig:blocking} for a summary of the algorithm.
-
-            % Flow chart for description of automatic blocking algorithm
-            \begin{figure}
-                \centering
-                % Defining blocks of flowchart
-                \tikzstyle{decision} = [diamond, draw, fill=blue!20, text
-                width=5em, text badly centered, node distance=3cm, inner
-                sep=0pt]
-
-                \tikzstyle{block} = [rectangle, draw, fill=blue!20, text
-                width=11em, text centered, rounded corners, minimum height=4em]
-
-                \tikzstyle{line} = [draw, -latex']
-
-                % Actual figure
-                \begin{tikzpicture}[node distance = 2.2cm, auto]
-
-                    % Nodes
-                    \node [block] (init) {Start with $\mathbf{X}_i$, $i=0$};
-
-                    \node [block, below of=init] (variance) {Compute variance
-                    and autocovariance of $\mathbf{X}_i$, $\sigma_i$ and
-                    $\gamma_i$};
-
-                    \node [block, below of=variance] (transform) {Transform data
-                    $(\mathbf{X}_{i+1})_k =
-                    \frac{1}{2}\left((\mathbf{X})_i)_{2k-1} +
-                    (\mathbf{X}_i)_{2k} \right)$};
-
-                    \node [block, right=1.0cm of transform] (i+1) {Next
-                    iteration: $i = i + 1$.} ;
-
-                    \node [decision, below of = transform] (x_size) {Length of
-                    $\mathbf{X}_{i + 1} \geq 2$ ? };
-
-                    \node [block, below of=x_size] (m_test) {Compute test
-                    statistic $M_j$ from $\sigma_j$ and $\gamma_j$ for all
-                    $j$.};
-
-                    \node [block, below of=m_test] (magic_numbers) {Find first
-                    $k$ such that $M_k \leq q_{d-k}(1-\alpha)$};
-
-                    \node [block, below of=magic_numbers] (return) {Return
-                    $\sigma_k^2/n_k$};
-
-                    % Edges
-                    \path [line] (init) -- (variance);
-                    \path [line] (variance) -- (transform);
-                    \path [line] (transform) -- (x_size);
-                        \path [line] (x_size) -| node [near start] {yes} (i+1);
-                        \path [line] (i+1) |- (variance);
-                    \path [line] (x_size) -- node [near start] {no} (m_test);
-                    \path [line] (m_test) -- (magic_numbers);
-                    \path [line] (magic_numbers) -- (return);
-                \end{tikzpicture}
-                \caption{Flow chart depicting the automatic blocking method for
-                finding a better estimate for the sample variance.}
-                \label{fig:blocking}
-            \end{figure}
-
-        \subsection{Gradient Descent}
-            Scanning the entire space of variational parameters can be a
-            somewhat despairing task. To aid us in the search we therefore
-            implement the method of gradient descent\footnote{Also known as
-            steepest descent, easily confused with the steepest descent method
-            for integation.}. In any optimum we generally have that the
-            derivative should equal zero. Our goal is to find the minimum of the
-            expected local energy with respect to the variational parameters
-            $\vfg{\alpha}$.
-            \begin{equation}
-                \nabla_{\vfg{\alpha}} \expval{E_L(\mathbf{R}; \vfg{\alpha})}
-                = 0.
-            \end{equation}
-            The foundation for  the method of gradient descent lies in the fact
-            that some analytic \footnote{We use the term analytic quite
-            recklessly.} function $\vf{F}(\vfg{\alpha})$ decreases fastest in
-            the direction of $-\nabla_{\vfg{\alpha}} \cdot
-            \vf{F}(\vfg{\alpha})$. This means that we could eventually get to an
-            ensuing minimum by trailing along the pathway of the following
-            recursive relation,
-            \begin{equation}
-                \vfg{\alpha}_{k+1}
-                = \vfg{\alpha}_k
-                - \gamma \nabla_{\vfg{\alpha}}\cdot \vf{F}(\vfg{\alpha}_k).
-                \label{eq:parameter_difference}
-            \end{equation}
-            If $\gamma$ is small enough we should have $\vf{F}(\vfg{\alpha}_k)
-            \geq \vf{F}(\vfg{\alpha}_{k+1})$ for all $k$ and hopefully we will
-            experience convergence to the desired minimum.
-
-            The variational gradient of the expectation value of the energy is
-            thus
-            \begin{align}
-                \nabla_{\vfg{\alpha}}\expval{E_L}
-                =
-                2\left(
-                    \expval{E_L\frac{1}{\Psi_T}\nabla_{\vfg{\alpha}}\Psi_T}
-                    - \expval{\frac{1}{\Psi_T} \nabla_{\vfg{\alpha}}\Psi_T}
-                    \expval{E_L}
-                \right).
-                \label{eq:variational_parameter_gradient}
-            \end{align}
-            To evaluate this expression we sample the local energy, the
-            parameter gradient of the wavefunction, and the combined expression
-            of these at each Monte Carlo cycle.
-
-            \subsubsection{Stability of the algorithm}
-                When employing the method of gradient descent to locate a
-                minimum value for the variational parameters we expect some
-                issues due to stability of the Monte Carlo methods and the
-                steepness of the expected energy curves. If our guess of the
-                location of the minimum is far off we can expect the method to
-                diverge as the gradient will become huge. This will yield a
-                new value for $\vfg{\alpha}$ in
-                \autoref{eq:parameter_difference} which will be further away and
-                might result in a local minimum "trapping" the method. To
-                prevent this from happening we have added a limit to the allowed
-                change in $\vfg{\alpha}$. If the change in $\vfg{\alpha}$ is
-                larger than a set order of magnitude we keep dividing
-                $\vfg{\alpha}$ with this number until this criteria is
-                fulfilled.
-
-                This is a so-called "hacky" solution, but it does the trick. It
-                is also the reason why the resultant figure in
-                \autoref{fig:gradient_descent_interacting} looks the way it
-                does for $\alpha_0 = 0.2$ and $\alpha_0 = 0.3$.
-
-\section{Results}
-
-    \subsection{Non-Interacting Systems}
-        Initially we are interested in seeing how our implementation performs
-        and to what extent it functions properly. The logical thing to do would
-        be to look at a non-interacting system with a simple Gaussian
-        wavefunction and harmonic oscillator potential. Such a system has a
-        simple closed-form analytic solution\footnote{A fact that every
-        undergraduate physics student should be well aware of after finishing
-        their first course in quantum physics}, which is further simplified by
-        the "god-given" analytic units
-        (\autoref{eq:closed_form_natural_units_local_energy}). This expression
-        is minimised for $\alpha = \omega/2$. We also set $\omega = 1$
-        and get an expression for the expected minimum local energy,
-        \begin{equation}
-            E_L(\vb{r}) = \frac{dN}{2},
-        \end{equation}
-        which will be used as the benchmark we are ultimately aiming for.
-
-        We will start without importance sampling, that is by employing brute
-        force Monte Carlo sampling. Moreover, we will compute the value for the
-        local energy analytically as well as with a numerical approach.  In the
-        numerical approach we employ the second order central difference
-        approximation
-        \begin{equation}
-            \dod[]{f(x)}{x} \approx \frac{f(x-h) - 2f(x) + f(x+h)}{h^2}.
-        \end{equation}
-
-        \begin{table}
-            \centering
-            \caption{One particle in one dimension for the analytic expression
-            with $2^{21}$ Monte Carlo cycles and a step length of $0.5$.}
-            \begin{ruledtabular}
-                \begin{tabular}{cccccc}
-                    $\alpha$ & $\langle  E_L\rangle$ & $\sigma$ & $\sigma_b$
-                    & $A$ & $t_{\text{CPU}} [\si{\second}]$ \\
-                    \hline
-                    0.30&0.56244&0.00026&0.00129&0.892&0.43\\
-                    0.34&0.53652&0.00019&0.00094&0.885&0.38\\
-                    0.38&0.51863&0.00013&0.00063&0.878&0.37\\
-                    0.42&0.50805&0.00009&0.00038&0.872&0.38\\
-                    0.46&0.50201&0.00004&0.00017&0.866&0.37\\
-                    0.50&0.50000&0.00000&0.00000&0.860&0.37\\
-                    0.54&0.50183&0.00004&0.00015&0.855&0.38\\
-                    0.58&0.50566&0.00007&0.00029&0.850&0.38\\
-                    0.62&0.51175&0.00011&0.00041&0.845&0.37\\
-                    0.66&0.51906&0.00014&0.00052&0.840&0.37\\
-                    0.70&0.52883&0.00017&0.00062&0.836&0.37\\
-                \end{tabular}
-            \end{ruledtabular}
-            \label{tab:1D1N_analytic}
-        \end{table}
-
-        \begin{table}
-            \centering
-            \caption{One particle in one dimension for the numeric expression
-            with $2^{21}$ Monte Carlo cycles and a step length of $0.5$.}
-            \begin{ruledtabular}
-                \begin{tabular}{cccccc}
-                    $\alpha$ & $\langle  E_L\rangle$ & $\sigma$ & $\sigma_b$
-                    & $A$&$t_{\text{CPU}} [\si{\second}]$ \\
-                    \hline
-                    0.30&0.56675&0.00026&0.00135&0.891&0.53\\
-                    0.34&0.53714&0.00019&0.00096&0.885&0.52\\
-                    0.38&0.51836&0.00013&0.00062&0.878&0.52\\
-                    0.42&0.50717&0.00009&0.00038&0.872&0.53\\
-                    0.46&0.50184&0.00004&0.00017&0.866&0.52\\
-                    0.50&0.50000&0.00001&0.00001&0.860&0.52\\
-                    0.54&0.50133&0.00004&0.00015&0.855&0.52\\
-                    0.58&0.50520&0.00007&0.00028&0.850&0.52\\
-                    0.62&0.51120&0.00011&0.00039&0.845&0.52\\
-                    0.66&0.51896&0.00014&0.00052&0.840&0.52\\
-                    0.70&0.52794&0.00017&0.00061&0.835&0.51\\
-                \end{tabular}
-            \end{ruledtabular}
-            \label{tab:1D1N_numeric}
-        \end{table}
-
-        \autoref{tab:1D1N_analytic} shows the expected local energy
-        $\expval{E_L}$, the uncorrelated standard deviation, $\sigma$, the
-        standard deviation found from blocking, $\sigma_b$, the accpetance
-        probability, $A$, and the CPU time for the analytic computation at
-        different values for the variational parameter $\alpha$.
-        \autoref{tab:1D1N_numeric} shows the same numbers but for the numeric
-        scheme.  One can clearly see, in both the analytic and numeric case,
-        that the uncorrelated standard deviations are lower than the standard
-        deviations found from the blocking method. This means that
-        autocorrelated data can lead to a lower perceived uncertainty than what
-        is the reality. The blocking method takes this fact into account and for
-        this reason, all other standard deviations will be computed with the
-        blocking method from here on.
-
-        The first initial observation of note is that we have found an optimum
-        for the system in terms of a minimum energy at $\alpha = 1/2$.
-        Moreover, notice how the standard deviation disappears completely.  This
-        should already be apparent from
-        \autoref{eq:closed_form_natural_units_local_energy}, as the term
-        involving particle positions disappears for the optimal $\alpha$ thus
-        rendering the entire Monte Carlo process of moving particles void.
-
-        \begin{table}
-            \caption{Comparison of energy, standard deviation and the CPU time
-            for $\alpha = 0.5$ for the analytic and numeric schemes in three
-            dimensions and $N$ particles. The step length is $0.5$.}
-            \centering
-            \begin{ruledtabular}
-                \begin{tabular}{r|rrr|rrr}
-                    & \multicolumn{3}{c|}{Analytic}
-                    & \multicolumn{3}{c}{Numeric} \\
-                    \hline
-                    $N$
-                    & $\langle E_L\rangle$ & $\sigma_b$
-                    & $t_{\text{CPU}} [\si{\second}]$
-                    & $\langle E_L\rangle$ & $\sigma_b$
-                    & $t_{\text{CPU}} [\si{\second}]$ \\
-                    \hline
-                    1 & 1.5 & 0.0 & 0.50
-                    & 1.49997 & 0.00008 & 0.73 \\
-                    10 & 15.0 & 0.0 & 0.73
-                    & 14.98278 & 0.00867 & 6.40 \\
-                    100 & 150.0 & 0.0 & 4.13
-                    & 149.81453 & 0.81501 & 387.60 \\
-                    500 & 750.0 & 0.0 & 19.16
-                    & 773.85435 & 20.69789 & 9595.02
-                \end{tabular}
-            \end{ruledtabular}
-            \label{tab:initial_optimum_comparison}
-        \end{table}
-
-        Systems of optimum variational parameter warrants further investigation,
-        as summed up in \autoref{tab:initial_optimum_comparison}. The analytical
-        results are quite predictable and monotonous; the expected local energy
-        $\expval{E_L}$ is perfectly proportional by a factor $3/2$ to
-        the number of particles and the standard deviation $\sigma_b$ is zero
-        for systems of any number of particles. The numerical results tells a
-        more tortuous tale. As the number of particles increase, the uncertainty
-        also increases rapidly. For the largest system, with $N = 500$
-        particles, the energy misses its target critically - the true energy is
-        more than one standard deviation away. Lastly, the CPU time is notably
-        higher for the numeric scheme.
-
-        In the appendices,  \autoref{fig:initial_problem_b} shows how both the
-        analytic and numeric approach matches the exact expression for the
-        energy as a function of $\alpha$ for systems of increasing size.  We see
-        also here that the uncertainty increases in the numeric scheme as the
-        system gets larger.
-
-        \subsubsection{Introducing Importance Sampling}
-            Next we will introduce importance sampling. The first prediction for
-            importance sampling compared to the brute force, naïve method is
-            that we should approach an equilibrium much faster because we are
-            now making moves in a smarter way. This prediction is confirmed in
-            \autoref{fig:problem_c1} and \autoref{fig:problem_c2}.
-
-            \begin{figure}
-                \centering
-                    \includegraphics[width=244px]
-                    {../data/figures/problem_c1.pdf}
-                    \caption{As the number of Monte Carlo cycles increases, the
-                    standard deviation of the expectation value of the energy
-                    for the brute force and importance sampling schemes
-                    decreases.}
-                    \label{fig:problem_c1}
-            \end{figure}
-
-            \begin{figure}
-                \centering
-                    \includegraphics[width=244px]
-                    {../data/figures/problem_c2.pdf}
-                    \caption{The absolute difference between the expected energy
-                    and the exact energy for both the brute force method and
-                    importance sampling.}
-                    \label{fig:problem_c2}
-            \end{figure}
-
-            In \autoref{fig:problem_c1} we can see the standard deviation of the
-            local energies for an increasing number of Monte Carlo cycles for
-            both the brute force sampling and the importance sampling methods.
-            We see that the standard deviation decreases for both methods for a
-            larger number of Monte Carlo cycles, but the standard deviation is
-            always lower for the importance sampling method. This is in
-            accordance with our expectations.
-
-            We have also compared the expected energy with the exact energy for
-            a different number of Monte Carlo cycles. The result of this
-            analysis is shown in \autoref{fig:problem_c2}. We see that both
-            methods will eventually result in expected energies that are equal
-            to the exact energies as $\abs{E - \expval{E}} \to 0$. However, for
-            the method of importance sampling the error is consistently lower.
-
-            \begin{table}
-                \caption{Importance sampled simulations of $N=10$ particles in
-                $d=3$ dimensions. Average acceptance ratios and standard
-                deviations for eleven values of $\alpha \in [0.3, 0.7]$.}
-                \centering
-                    \begin{ruledtabular}
-                        \begin{tabular}{cccc}
-                            $\delta t$ & $\bar{\sigma}_b$ & $\bar{A}$
-                            & $\expval{E}\left(\alpha = 1/2\right)$ \\
-                            \hline
-                            $2^{+3}$ & 0.09363 & 0.005 & 15.000 \\
-                            $2^{+2}$ & 0.01549  & 0.041 & 15.000 \\
-                            $2^{+1}$ & 0.00706 & 0.216 & 15.000 \\
-                            $2^{\ 0}$ & 0.00675 & 0.510 & 15.000 \\
-                            $2^{-1}$ & 0.00932 & 0.738 & 15.000 \\
-                            $2^{-2}$ & 0.01578 & 0.866 & 15.000 \\
-                            $2^{-3}$ & 0.02723 & 0.933 & 15.000 \\
-                            $2^{-4}$ & 0.05016 & 0.966 & 15.000 \\
-                            $2^{-5}$ & 0.08474 & 0.983 & 15.000 \\
-                            $2^{-6}$ & 0.11555 & 0.992 & 15.000 \\
-                            $2^{-7}$ & 0.17070 & 0.996 & 15.000 \\
-                        \end{tabular}
-                    \end{ruledtabular}
-                    \label{tab:imp_time_step}
-            \end{table}
-
-            \autoref{tab:imp_time_step} shows the standard deviations and
-            acceptance ratios for different time steps, $\delta t$ using the
-            importance sampling method. The simulations in this table are from
-            the same values of $\alpha$ as in \autoref{tab:1D1N_analytic} and
-            \autoref{tab:1D1N_numeric}. The optimum energies for $\alpha = 1/2$
-            gives the same result for importance sampling as for the naïve way
-            of sampling. Notice that by lowering $\delta t$, the uncertainty
-            starts to decrease and then starts to increase again. Moreover, the
-            acceptance ratios increases as $\delta t \to 0$. This happens as the
-            change in the wavefunction grows minuscule and $q(\vf{r}_{i + 1},
-            \vf{r}_i) \approx 1$.
-
-    \subsection{Interacting elliptical harmonic oscillator}
-        We now make a change to the simple system in the previous section in
-        order to make everything a bit more interesting. The potential in the
-        one-body Hamiltonian is perturbed in the $z$-direction according to
-        \autoref{eq:scaled_hamiltonian} by setting $\lambda = \sqrt{8} \approx
-        2.82843$. This corresponds to enclosing particles of the simulation in
-        an elliptical trap instead of a spherical trap.
-
-        We also choose a new trial wavefunction with the elliptical gaussian
-        single particle functions and the Jastrow factor. The hard sphere radius
-        is set to $a = 0.0043$.
-
-        Due to similarities with the spherical harmonic oscillator
-        system we make a guess that the true minimum of this new system should
-        be situated close to the minimum of the previous system. Choosing seven
-        values for $\alpha \in [0.2, 0.7]$ we run importance sampling on this
-        system for $N = \{10, 50, 100\}$ particles in $d = 3$ dimensions using
-        $2^{21}$ Monte Carlo cycles\footnote{This turned out to be a little bit
-        overkill due to the fast convergence of importance sampling.} and an
-        additional $10\%$ of the Monte Carlo cycles for thermalization of the
-        system prior to doing any sampling. We have used $\delta t = 0.1$ which
-        yields a high acceptance ratio, but slower convergence.
-
-        In \autoref{tab:10_interacting} the system is set to have $N = 10$
-        interacting particles. Notice first that the energy is higher in
-        the interacting system than for a comparable non-interacting system
-        (\autoref{tab:initial_optimum_comparison}).  Moreover, the standard
-        deviation gets lower as the system is close to an optimal variational
-        parameter.
-
-        \begin{table}
-            \caption{Simulation results for $N = 10$ bosonic interacting,
-            elliptical, harmonic oscillators.}
-            \centering
-            \begin{ruledtabular}
-                \begin{tabular}{ccccc}
-                    $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
-                    &$A$ & $t_{\text{CPU}} [\si{\second}]$ \\
-                    \hline
-                    0.2&35.17548&0.05399&0.990%98995
-                    &24.5\\%48868\\
-                    0.3&27.62004&0.02311&0.981%98187
-                    &13.6\\%65132\\
-                    0.4&24.97850&0.00827&0.972%97248
-                    &13.5\\%54666\\
-                    0.5&24.39877&0.00030&0.961%96140
-                    &13.4\\%38388\\
-                    0.6&24.83863&0.00604&0.950%94998
-                    &13.3\\%31570\\
-                    0.7&25.82855&0.01059&0.938%93766
-                    &13.2\\%20579\\
-                    0.8&27.22895&0.01436&0.924%92375
-                    &13.1\\%11933\\
-                \end{tabular}
-            \end{ruledtabular}
-            \label{tab:10_interacting}
-        \end{table}
-
-        The result of simulations with a higher number of particles, $N=50$ and
-        $N=100$, are  shown in \autoref{tab:50_interacting} and
-        \autoref{tab:100_interacting} respectively. We observe that the energy
-        for systems of interacting particles increases rapidly with the size of
-        the system.
-
-        \begin{table}
-            \caption{Simulation results for $N = 50$ bosonic interacting,
-            elliptical, harmonic oscillators.}
-            \centering
-            \begin{ruledtabular}
-                \begin{tabular}{ccccc}
-                    $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
-                    & $A$ & $t_{\text{CPU}} [\si{\second}]$ \\
-                    \hline
-                    0.2&181.69371&0.25533&0.987%98756
-                    &199.9\\%95366\\
-                    0.3&142.61591&0.10782&0.978%97849
-                    &195.8\\%80928\\
-                    0.4&129.88615&0.04051&0.969%96862
-                    &190.4\\%44553\\
-                    0.5&127.29926&0.00595&0.957%95749
-                    &195.1\\%10368\\
-                    0.6&129.97630&0.03300&0.946%94595
-                    &193.2\\%22755\\
-                    0.7&135.67382&0.05785&0.933%93330
-                    &192.0\\%191.96391\\
-                    0.8&143.23238&0.07286&0.921%92061
-                    &192.7\\%72399\\
-                \end{tabular}
-            \end{ruledtabular}
-            \label{tab:50_interacting}
-        \end{table}
-
-        \begin{table}
-            \caption{Simulation results for $N = 100$ bosonic interacting,
-            elliptical, harmonic oscillators.}
-            \centering
-            \begin{ruledtabular}
-                \begin{tabular}{ccccc}
-                    $\alpha$ & $\langle  E_L\rangle$ & $\sigma_b$
-                    &$A$ & $t_{\text{CPU}} [\si{\second}]$ \\
-                    \hline
-                    0.2&375.04401&0.51475&0.985%98527
-                    &745.2\\%21438\\
-                    0.3&296.76879&0.20214&0.975%97534
-                    &741.0\\%04425\\
-                    0.4&270.60830&0.08113&0.945%94522
-                    &723.1\\%06823\\
-                    0.5&266.37263&0.02020&0.953%95314
-                    &702.8\\%71000\\
-                    0.6&272.51171&0.07366&0.941%94112
-                    &689.1\\%12862\\
-                    0.7&285.10285&0.11381&0.929%92895
-                    &707.5\\%48248\\
-                    0.8&301.40609&0.15458&0.916%91648
-                    &707.9\\%86118\\
-                \end{tabular}
-            \end{ruledtabular}
-            \label{tab:100_interacting}
-        \end{table}
-
-        \subsubsection{Finding the minimum}
-            Using the method of gradient descent we choose a set of six starting
-            $\alpha_0 \in [0.2, 0.8]$ and try to locate a value for $\alpha$
-            where the gradient goes to zero. We looked at $N = 10$ with $d = 3$,
-            $\omega = 1$ and $\beta = \lambda = \sqrt{8}$. This yields the value
-            for the minimum $\alpha$ to be
-            \begin{align}
-                \alpha = 0.49744 \pm 0.00002.
-                \label{eq:optimal_interacting_alpha}
-            \end{align}
-            In \autoref{fig:gradient_descent_interacting} we can see how the
-            conjugate gradient method "moves" to find the optimal value of
-            $\alpha$.
-
-            For this value of $\alpha$ we have computed the expected energy and
-            standard deviation for systems of $N = \{10, 50, 100\}$ particles.
-            These results are shown in \autoref{tab:interacting_optimal}. We have
-            compared with the non-interacting case, i.e., where the Jastrow
-            factor is zero ($a = 0$).
-
-            \begin{figure}
-                \includegraphics[width=244px]{../data/figures/problem_f.pdf}
-                \caption{In this figure we can see the convergence of different
-                starting values for $\alpha$ towards the true value of $\alpha$
-                that minimizes the expected energy in the interacting case.}
-                \label{fig:gradient_descent_interacting}
-            \end{figure}
-
-            \begin{table}
-                \caption{Comparison of an elliptical gaussian trial wavefunction
-                with and without the Jastrow factor using the elliptical trap.}
-                \centering
-                \begin{ruledtabular}
-                    \begin{tabular}{r|rr|rr}
-                        & \multicolumn{2}{c|}{Interacting}
-                        & \multicolumn{2}{c}{Non-interacting} \\
-                        \hline
-                        $N$ & $\expval{E}$ & $\sigma_b$ & $\expval{E}$
-                        & $\sigma_b$ \\
-                        \hline
-                        10 & 24.3986 & 0.0003 & 24.1418 & 0.0002 \\
-                        50 & 127.2852 & 0.0077 & 120.7129 & 0.0012 \\
-                        100 & 266.2904 & 0.0297 & 241.4248 & 0.0025
-                    \end{tabular}
-                \end{ruledtabular}
-                \label{tab:interacting_optimal}
-            \end{table}
-
-        \subsubsection{One-Body Densities}
-            We now look at one-body densities for the value of $\alpha$ found in
-            \autoref{eq:optimal_interacting_alpha}. The results for a system
-            consisting of $N=100$ particles are displayed in
-            \autoref{fig:one_body_density_100}.  In this figure we see the
-            densities for a system confined to a perturbed elliptic trap with
-            and without interaction. We see that for the interacting case, the
-            one-body density is skewed towards a larger $\abs{\vf{r}}$, as is
-            expected. The hard core prevents the bosons from stacking on top of
-            each other spatially. By increasing the hard core radius this trend
-            will be more noticeable. The one-body density of the "ideal" system
-            is also included in the figure.
-
-            \begin{figure}
-                \includegraphics[width=244px]
-                {../data/figures/problem_g_100N.pdf}
-                \caption{In this figure the interacting and non-interacting
-                densities correspond to the elliptical potential with and
-                without the Jastrow factor respectively. The "ideal" system is
-                the spherical potential with the non-interacting Gaussian trial
-                wavefunction.}
-                \label{fig:one_body_density_100}
-            \end{figure}
-
-\section{Discussion}
-
-    \subsection{Validity of Results}
-        The non-interacting tests compared against the analytic "solution"
-        showed to a very high degree the validity of the method. One can see
-        this quite clearly from \autoref{fig:initial_problem_b}. Because we get
-        results from simulations that correspond to the analytic results we can
-        be fairly certain that the method can be expanded upon to include
-        attributes outside the realm of problems that are analytically solvable.
-        The systems with the elliptic traps and the interacting particles have
-        just such attributes.
-
-        That said, the way to get the best confirmation of the results presented
-        above is in comparison to experiment. This would be quite difficult
-        as the systems presented are very theoretical in nature and most likely
-        far from anything we would see in nature. Without being able to compare
-        our result with anything happening in the real world this entire study would
-        have no value except for the instrumental value, being no more than
-        a theoretical exercise.  However, the one-body
-        density is a measure of the role of many-body correlations which in
-        some cases can be related to experiment. For a system of charged
-        particles it can be used to determine a charge distribution, which in
-        turn can be linked to experiment. It would be very fruitful to simulate a
-        system of such particles that is more natural, i.e. real, but we realise the
-        purpose of this study is first and foremost to introduce variational
-        Monte Carlo as a tool in computational physics. We assume, however, that
-        this study is a good starting point for studies of alkali atoms like
-        $^{87}$Rb, $^{23}$Na or $^7Li$.
-
-    \subsection{Energy per particle}
-        For the non-interacting harmonic oscillator system there is no
-        difference between a new particle and a new dimension\footnote{This is
-        not entirely true in the case of our implementation of the
-        Metropolis-Hastings algorithm as we draw a single random particle and
-        move all dimensions instead of iterating over all particles and all
-        dimensions.}. This means that the inclusion of a new particle or
-        dimension yields a linear scaling in the energy. Thus if we compute the
-        energy per particle for systems of varying size we expect to get the
-        same energy. This is most easily seen in \autoref{fig:initial_problem_b}
-        and in the expression for the exact energy \autoref{eq:exact_energy}.
-
-        This, however, is \emph{not} the case in the interacting system due to
-        the Jastrow factor. Looking at \autoref{tab:10_interacting} and
-        \autoref{tab:100_interacting} we see that multiplying the expected
-        energy of the former with $10$ yields a lower number than the latter. We
-        can also see this behavior in \autoref{fig:interacting}, where the
-        energy per particle for $N = \{10, 50, 100\}$ have been plotted in the
-        same figure.
-
-        \begin{figure}
-            \includegraphics[width=244px]{../data/figures/problem_e.pdf}
-            \caption{Expected energy per particle as a function of variational
-            parameter $\alpha$ for a varying number of particles.}
-            \label{fig:interacting}
-        \end{figure}
-
-        The reason for the increase in energy per particle can be ascribed to
-        the repulsive interaction added by the Jastrow factor. The repulsive
-        force can be explained thusly: the Jastrow factor  increases the value
-        of the local energy for any given particle if that particle is some
-        distance away from any other particle. The closer the particles are, the
-        more the local energy increases. The Jastrow factor will lower the
-        probability for the particles to get too close to one another by
-        reducing the probability density in the vicinity of other particles.
-        However, if there are many particles in the system they will be closer
-        to another particle and therefore increase the energy of the system.
-
-    \subsection{Low density system}
-        It is safe to say that the effect of adding a Jastrow factor is
-        significant, i.e. the increase in energy for the system as the
-        interaction is switched on is much higher than the uncertainty of the
-        energy. A simple look at \autoref{fig:interacting} will be proof enough
-        as the error bars of the figure are not discernible.  On the other hand,
-        the energy added to the system by including interaction is not very high
-        compared to the energy of the system as a whole. For the larger of the
-        interacting systems we have looked at, with $N=100$ particles, the
-        interaction add only a mere $9\%$ extra energy compared to the
-        non-interacting system (see \autoref{tab:interacting_optimal}).
-        Studying \autoref{fig:one_body_density_100} will reveal the same trend
-        in the distribution of the particles.
-
-        The reason for this is two-fold.
-        \begin{enumerate}
-            \item The hard sphere radius is vanishingly small. By increasing the
-                hard sphere radius the particles will contribute more
-                significantly to the local energy. The particles will also be
-                more spread out thus occupying a larger configuration space.
-
-            \item The assumption which justifies that adding only two-body
-                interaction is sufficient to model all interactions to a
-                reasonable extent. This assumption is justified by the fact that
-                alkali and atomic hydrogen systems\footnote{Systems similar to
-                our toy model.} are dilute. For gases of alkali atoms confined
-                in magnetic or optical traps the average distance between atoms
-                is much larger than the range of inter-atomic interaction. The
-                probability is high that atoms very seldom interact and if so,
-                just with one other atom.
-        \end{enumerate}
-
-    \subsection{Stability of the program}
-        An interesting thing would be to model large systems of many particles
-        and with large hard sphere radii. Unfortunately, due to the shape of the
-        single particle functions this turns out to be quite tricky. As the
-        single particle function consists of a sum squared positions from a
-        given center, many particles and/or spread out systems will yield a
-        large sum. This means that the Slater permanent grows smaller and
-        smaller making it much harder to keep the system confined. By lowering
-        the frequency and thus lowering the minimum $\alpha$ we can remedy some
-        of this behaviour.
-
-\section{Summary Remarks}
-
-    In this study we have built a functional variational Monte Carlo solver framework for
-    a gas following Bose-Einstein statistics confined in an elliptic trap. Monte Carlo
-    methods is a broad class of computational algorithms, whereas the variational
-    Monte Carlo is a subset. Other applications of such methods within the physical
-    sciences includes, but is not limited to, quantum chromodynamics, the modeling of
-    aearodynamic forms, particle physics detector design and galaxy evolution.
-    Needless to say, Monte Carlo method is a valuable tool for a computational physicist.
-
-    To prove the
-    validity of the framework we tested it against an analytically solved case, an ideal
-    non-interacting system. Thereafter we have shown how our implementation can
-    handle large systems of interacting particles and perturbed traps. Even though
-    this study is somewhat theoretic in nature, we have discussed the possibility of
-    expanding our machinery to a case bearing a larger resemblance to a system
-    found in nature. To this end we have computed one-body densities for the
-    interacting system, which can be compared to real physical properties as
-    measured in an experiment.
-
 \bibliography{references}
-\clearpage
-
-\appendix
-\section{Computing the exact energy of the non-interacting system}
-    Here we show how to arrive at \autoref{eq:exact_energy}. We need to solve
-    two integrals consisting of $N$, $d$-dimensional, Gaussians from
-    \autoref{eq:trial_simple_gaussian}. We will label the position in the $j$-th
-    coordinate for the $i$-th particle by $x_{ij}$. Starting with the
-    normalization factor we get
-    \begin{align}
-        \bra{\Psi_T}\ket{\Psi_T}
-        &=
-        \int\dd\vf{r}\abs{\Psi_T(\vf{r})}^2
-        \\
-        &=
-        \prod_{i = 1}^N
-        \int\dd\vf{r}_i\exp(-2\alpha\vf{r}_i^2)
-        \\
-        &=
-        \prod_{i = 1}^N
-        \prod_{j = 1}^d
-        \int\dd x_{ij}\exp(-2\alpha x_{ij}^2)
-        \\
-        &=
-        \prod_{i = 1}^N
-        \prod_{j = 1}^d
-        \sqrt{\frac{\pi}{2\alpha}}
-        = \para{
-            \sqrt{\frac{\pi}{2\alpha}}
-        }^{dN}.
-    \end{align}
-    From the Hamiltonian we get two integrals containing $\vf{r}_i^2$ for each
-    particle $i$.
-    \begin{align}
-        \bra{\Psi_T}\vf{r}_i^2\ket{\Psi_T}
-        &=
-        \int\dd\vf{r}\abs{\Psi_T(\vf{r})}^2\vf{r}_i^2
-        \\
-        &=
-        \int\dd\vf{r}_i\exp(-2\alpha\vf{r}_i^2)\vf{r}_i^2
-        \nonumber \\
-        &\qquad
-        \times
-        \prod_{j \neq i}^N
-        \int\dd\vf{r}_j\exp(-2\alpha\vf{r}_j^2)
-        \\
-        &=
-        \sum_{l = 1}^d
-        \int
-        \para{
-            \prod_{k = 1}^d
-            \dd x_{ik}\exp(-2\alpha x_{ik}^2)
-        } x_{il}^2
-        \nonumber\\
-        &\qquad
-        \times
-        \para{
-            \sqrt{\frac{\pi}{2\alpha}}
-        }^{d(N - 1)}
-        \\
-        &=
-        \sum_{l = 1}^d
-        \frac{1}{4\alpha}\sqrt{\frac{\pi}{2\alpha}}
-        \para{
-            \sqrt{\frac{\pi}{2\alpha}}
-        }^{d - 1}
-        \nonumber\\
-        &\qquad
-        \times
-        \para{
-            \sqrt{\frac{\pi}{2\alpha}}
-        }^{d(N - 1)}
-        \\
-        &= \frac{d}{4\alpha}\para{
-            \sqrt{\frac{\pi}{2\alpha}}
-        }^{dN}.
-    \end{align}
-    We thus have that
-    \begin{align}
-        \frac{\bra{\Psi_T}\vf{r}_i^2\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
-        &=
-        \frac{d}{4\alpha}.
-    \end{align}
-    Using \autoref{eq:laplacian_simple_gaussian} we can find the exact energy
-    expression we are looking for.
-    \begin{align}
-        E(\alpha)
-        &=
-        \frac{\bra{\Psi_T}H\ket{\Psi_T}}{\bra{\Psi_T}\ket{\Psi_T}}
-        \\
-        &=
-        \sum_{i = 1}^N
-        \brac{
-            -\frac{\hbar^2}{2m}
-            \brak{
-                -2d\alpha
-                + 4\alpha^2\frac{d}{4\alpha}
-            }
-            + \frac{1}{2}m\omega^2\frac{d}{4\alpha}
-        }
-        \\
-        &=
-        \para{
-            \frac{\alpha\hbar^2}{2m}
-            + \frac{m\omega^2}{8\alpha}
-        }dN,
-    \end{align}
-    which is what we wanted to show.
-
-\section{Computing the drift force and local energy of the full system}
-    Computing the gradient of the wavefunction we get
-    \begin{align}
-        \nabla_k\Psi_T(\vf{r})
-        &=
-        \Bigl[
-            \nabla_k
-            \Phi_T(\vf{r})
-        \Bigr]
-        J(\vf{r})
-        + \Phi_T(\vf{r})
-        \nabla_k J(\vf{r}).
-    \end{align}
-    The gradient of the Slater permament for particle $k$ is given by
-    \begin{align}
-        \nabla_k
-        \Phi_T(\vf{r})
-        &=
-        \nabla_k\phi_k
-        \prod_{i \neq k}^N\phi_i
-        = \frac{\nabla_k\phi_k}{\phi_k}
-        \Phi_T(\vf{r}).
-    \end{align}
-    The gradient of the Jastrow factor is given by
-    \begin{align}
-        \nabla_k J(\vf{r})
-        &=
-        J(\vf{r})
-        \nabla_k\sum_{m < n}^N u_{mn} \\
-        &= J(\vf{r})
-        \Biggl(
-            \sum_{m = 1}^{k - 1}\nabla_k u_{mk}
-            \sum_{n = k + 1}^N\nabla_k u_{kn}
-        \Biggr)
-        \\
-        &=
-        J(\vf{r})
-        \sum_{m \neq k}^N\nabla_k u_{km},
-    \end{align}
-    where the gradient of the interaction term splits the antisymmetric
-    sum into two parts. As $r_{ij} = r_{ji}$ we can combine these sums
-    into a single sum. This in total yields the gradient
-    \begin{align}
-        \nabla_k\Psi_T(\vf{r})
-        &=
-        \Biggl(
-            \frac{\nabla_k\phi_k}{\phi_k}
-            + \sum_{m \neq k}^N
-            \nabla_k u_{km}
-        \Biggr)
-        \Psi_T(\vf{r}).
-        \label{eq:gradient_full_wavefunction}
-    \end{align}
-
-    The Laplacian of the trial wavefunction is found by finding the
-    divergence of \autoref{eq:gradient_full_wavefunction}.
-    \begin{align}
-        \nabla_k^2\Psi_T(\vf{r})
-        &=
-        \Biggl(
-            \nabla_k\Biggl[
-                \frac{\nabla_k\phi_k}{\phi_k}
-            \Biggr]
-            +
-            \sum_{m \neq k}^N \nabla_k^2 u_{km}
-        \Biggr)\Psi_T(\vf{r})
-        \\
-        &\qquad
-        +
-        \Biggl(
-            \frac{\nabla_k\phi_k}{\phi_k}
-            + \sum_{m \neq k}^N
-            \nabla_k u_{km}
-        \Biggr)^2
-        \Psi_T(\vf{r}),
-    \end{align}
-    where the squared term came from taking the gradient of the trial
-    wavefunction.  To further simplify we divide by the trial
-    wavefunction. This yields
-    \begin{align}
-        \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
-        &=
-        \frac{\nabla_k^2\phi_k}{\phi_k}
-        + 2\frac{\nabla_k\phi_k}{\phi_k}
-        \sum_{m \neq k}\nabla_k u_{km}
-        \nonumber \\
-        &\qquad
-        + \sum_{m\neq k}^N\nabla_k^2 u_{km}
-        + \Biggl(
-            \sum_{m \neq k}^N\nabla_k u_{km}
-        \Biggr)^2.
-    \end{align}
-    From here, the next step is to find the gradient and the Laplacian of
-    the single particle functions, $\phi_k$, and the interaction
-    functions $u_{km}$. For the single particle functions we use
-    Cartesian coordinates when finding the derivatives while for
-    the interaction functions we use spherical coordinates and do a
-    variable substitution. Beginning with the gradient of the single
-    particle functions we get
-    \begin{align}
-        \nabla_k\phi_k
-        &=
-        \nabla_k\exp\bigl[
-            -\alpha(x_k^2 + y_k^2 + \beta z_k^2)
-        \bigr] \\
-        &=
-        -2\alpha
-        (x_k\vf{e}_i + y_k\vf{e}_j + \beta z_k\vf{e}_k)
-        \phi_k.
-    \end{align}
-    Note that the subscripts on the unit vectors $\vf{e}_i$ are
-    \emph{not} the same as the subscripts used for its components. The
-    Laplacian yields
-    \begin{align}
-        \nabla_k^2\phi_k
-        &=
-        \Bigl[
-            -2\alpha
-            \big(
-                d - 1 + \beta
-            \bigr)
-            \nonumber \\
-            &\qquad
-            + 4\alpha^2
-            \bigl(
-                x_k^2 + y_k^2 + \beta^2z_k^2
-            \bigr)
-        \Bigr]
-        \phi_k,
-    \end{align}
-    with $d$ as the dimensionality of the problem.
-    In order to derive the interaction functions we have to do a
-    variable substitution using $r_{km} = |\vf{r}_k - \vf{r}_m|$. We can
-    then rewrite the $\nabla_k$-operator as
-    \begin{align}
-        \nabla_k
-        &=
-        \nabla_k
-        \dpd[]{r_{km}}{r_{km}}
-        =
-        \nabla_k r_{km} \dpd[]{}{r_{km}}
-        \\
-        &=
-        \frac{\vf{r}_k - \vf{r}_m}{r_{km}}\dpd[]{}{r_{km}}.
-    \end{align}
-    Applying this version of the $\nabla_k$-operator to $u_{km}$ yields
-    \begin{align}
-        \nabla_k u_{km}
-        &=
-        \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
-        \dpd[]{u_{km}}{r_{km}}.
-    \end{align}
-
-    \begin{widetext}
-        For the Laplacian we switch a little back and forth between the
-        two ways of representing the $\nabla_k$-operator. We thus get
-        \begin{align}
-            \nabla_k^2 u_{km}
-            &=
-            \frac{\nabla_k \vf{r}_k}{r_{km}}\dpd[]{u_{km}}{r_{km}}
-            + \Biggl[
-                \nabla_k \frac{1}{r_{km}}
-            \Biggr]
-            (\vf{r}_k - \vf{r}_m)\dpd[]{u_{km}}{r_{km}}
-            + \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
-            \nabla_k \dpd[]{u_{km}}{r_{km}} \\
-            &= \frac{d}{r_{km}}\dpd[]{u_{km}}{r_{km}}
-            - \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^3}
-            \dpd[]{u_{km}}{r_{km}}
-            + \frac{(\vf{r}_k - \vf{r}_m)^2}{r_{km}^2}
-            \dpd[2]{u_{km}}{r_{km}} \\
-            &=
-            \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
-            + \dpd[2]{u_{km}}{r_{km}},
-        \end{align}
-        where $d$ is again the dimensionality of the problem. In total
-        we can state an intermediate version of the Laplacian occuring
-        in the local energy as
-        \begin{align}
-            \frac{\nabla_k^2\Psi_T(\vf{r})}{\Psi_T(\vf{r})}
-            &=
-            \frac{\nabla_k^2\phi_k}{\phi_k}
-            + 2\frac{\nabla_k\phi_k}{\phi_k}
-            \sum_{m \neq k}^N
-            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
-            \dpd[]{u_{km}}{r_{km}}
-            + \sum_{m\neq k}^N
-            \Biggl(
-                \frac{d - 1}{r_{km}}\dpd[]{u_{km}}{r_{km}}
-                + \dpd[2]{u_{km}}{r_{km}}
-            \Biggr)
-            \nonumber \\
-            &\qquad
-            +
-            \sum_{m, n \neq k}^N
-            \frac{\vf{r}_k - \vf{r}_m}{r_{km}}
-            \frac{\vf{r}_k - \vf{r}_n}{r_{kn}}
-            \dpd[]{u_{km}}{r_{km}}
-            \dpd[]{u_{kn}}{r_{kn}}.
-        \end{align}
-    \end{widetext}
-    Moving on to the derivatives of the interaction terms, $u_{km}$, to
-    get an explicit expression for the Laplacian.
-    \begin{align}
-        \dpd[]{u_{km}}{r_{km}}
-        &=
-        \frac{a}{r_{km}(r_{km} - a)},
-        \\
-        \dpd[2]{u_{km}}{r_{km}}
-        &= \frac{a^2 - 2ar_{km}}{r_{km}^2(r_{km} - a)^2}.
-    \end{align}
-    The local energy and the drift force can now be found by combining these
-    expressions. We will not write out the explicit expressions as
-    these will be called by separated functions in our programs.
-
-\section{Variational parameter gradient of the expectation energy}
-    Here we show how to arrive at the expression shown in
-    \autoref{eq:variational_parameter_gradient}. This requires us to restrict
-    our view to real trial wavefunctions, i.e., $\Psi_T^{*} = \Psi_T$, which is
-    the case for the wavefunctions we are exploring.
-    \begin{align}
-        \nabla_{\vfg{\alpha}}\expval{E_L}
-        &=
-        \nabla_{\vfg{\alpha}}
-        \para{
-            \frac{
-                \int\dd\vf{r}\Psi_T^{*}H\Psi_T
-            }{
-                \int\dd\vf{r}|\Psi_T|^2
-            }
-        }
-        \\
-        &=
-        -
-        \frac{
-            2
-            \para{
-                \int\dd\vf{r}\Psi_T H\Psi_T
-            }
-        }{\para{\int\dd\vf{r}\Psi_T}^2}
-        \int\dd\vf{r}\Psi_T\nabla_{\vfg{\alpha}}\Psi_T
-        \\
-        &\qquad
-        +
-        \frac{2}{\int\dd\vf{r}\Psi_T^2}
-        \int\dd\vf{r}\Psi_TH\brak{
-            \nabla_{\vfg{\alpha}}\Psi_T
-        }
-        \\
-        &=
-        - 2\expval{E_L}
-        \frac{1}{
-            \int\dd\vf{r}\Psi_T^2
-        }
-        \int\dd\vf{r}\Psi_T^2\para{
-            \frac{\nabla_{\vfg{\alpha}}\Psi_T}{\Psi_T}
-        }
-        \\
-        &\qquad
-        +
-        \frac{2}{\int\dd\vf{r}\Psi_T^2}
-        \int\dd\vf{r}\Psi_T^2 E_L\para{
-            \frac{\nabla_{\vfg{\alpha}}\Psi_T}{\Psi_T}
-        }.
-    \end{align}
-    We now use the definition of the expectation value to group terms together.
-    We are thus left with
-    \begin{align}
-        \nabla_{\vfg{\alpha}}\expval{E_L}
-        =
-        2
-        \para{
-            \expval{
-                E_L\frac{1}{\Psi_T}\nabla_{\vfg{\alpha}}\Psi_T
-            }
-            - \expval{E_L}
-            \expval{\frac{1}{\Psi_T}\nabla_{\vfg{\alpha}}\Psi_T}
-        },
-    \end{align}
-    which is what we wanted to show.
-
-\clearpage
-\onecolumngrid
-\section{Brute Force Metropolis-Hastings}
-    In \autoref{fig:initial_problem_b} we see how the expectation value of the
-    energy for the analytic and numeric expression of the local energy using
-    Monte Carlo sampling matches the exact energy from \autoref{eq:exact_energy}
-    for $\alpha \in [0.3, 0.7]$.
-    \begin{figure}[b]
-        \makebox[\textwidth][c]{
-            \includegraphics[width=1.2\textwidth]{../data/figures/problem_b.pdf}
-        }
-        \caption{Comparison of analytic and numeric results with the exact
-        expression for the energy as a function of $\alpha$. The plots show the
-        standard deviation from the blocking method as error ticks. In the
-        figure $N$ is the number of particles and $d$ the number of dimensions.}
-        \label{fig:initial_problem_b}
-        \vspace{2cm}
-    \end{figure}
 

### 2018-04-09 project_2/article/article.tex
Add introduction section
Add system to the theory section

@@
 
+\section{Introduction}
+    In this project we will study the ground state energy of quantum dots.
+
+\section{Theory}
+    In this project we will study a system of $N$ interacting electrons. We will
+    be looking at a Hamiltonian consisting of a one-body and a two-body part.
+    The one-body part is given by
+    \begin{align}
+        h(\vf{r}_i)
+        &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
+    \end{align}
+    where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
+    The two-body part is the Coulomb interaction potential.
+    \begin{align}
+        v(\vf{r}_i, \vf{r}_j)
+        &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
+    \end{align}
+    We thus get the total Hamiltonian
+    \begin{align}
+        H &= \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N v(\vf{r}_i, \vf{r}_j).
+    \end{align}
+
 \bibliography{references}

### 2018-04-09 project_2/article/article.tex
Add start of second quantization

@@
 \usepackage{amsthm}
-% Import physics package to include bra-ket
-\usepackage{physics}
 \usepackage{enumerate}
@@
 
+\newcommand{\bra}[1]{\langle #1\lvert}
+\newcommand{\ket}[1]{\rvert #1\rangle}
+\newcommand{\braket}[2]{\langle #1 \vert #2 \rangle}
+
 \begin{document}
@@
     \end{align}
+    Next we define the reference Slater determinant as
+    \begin{align}
+        \ket{\Phi_0} &\equiv \ket{1, 2, \dots, N},
+    \end{align}
+    where $\ket{p}$ are the single particle functions that satisfies the
+    equation
+    \begin{align}
+        h\ket{p} = \varepsilon_p\ket{p}.
+    \end{align}
+    They are the eigenfunctions to the one-body part of the Hamiltonian and are
+    thus harmonic oscillator functions. Note that $\ket{\Phi_0}$ consists of the
+    $N$ first occupied single particle functions. We will denote the occupied
+    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the virtual states
+    with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices with
+    $p, q, r, s, \dots \in \{1, \dots, L\}$.
 

### 2018-04-10 project_2/article/article.tex
Add section on second quantization

@@
 
+\newcommand{\acr}[1]{a_{#1}^{\dagger}}
+\newcommand{\ade}[1]{a_{#1}}
+
 \begin{document}
@@
 
+    \subsection{Second quantization}
+        We now move over to a use second quantization. Employing the creation
+        operators, $\acr{p}$, and the destruction operators, $\ade{p}$, we can
+        write the Hamiltonian as
+        \begin{align}
+            H
+            &=
+            \sum_{pq}h_{q}^{p}\acr{p}\ade{q}
+            + \sum_{pqrs}w_{rs}^{pq}\acr{p}\acr{q}\ade{s}\ade{r},
+        \end{align}
+        where the sums are general indices over all $L$ basis states and the
+        matrix elements are defined as
+        \begin{gather}
+            h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
+            w_{rs}^{pq} \equiv \bra{pq}v\ket{rs}.
+        \end{gather}
+        In the case of the eigenstates we have that the one-body operator
+        reduces to a diagonal matrix.
+        \begin{align}
+            h^{p}_{q}
+            &= \bra{p}h\ket{q}
+            = \varepsilon_q\braket{p}{q}
+            = \varepsilon_q\delta_{pq},
+        \end{align}
+        where $\delta_{pq}$ is the Kronecker-Delta.
+
 \bibliography{references}

### 2018-04-10 project_2/article/article.tex
Add TODO
Add start of the coupled cluster section

@@
 \section{Introduction}
+    %TODO: Add explanation on why the many-body problem becomes intractable.
+    % This can be shown by demonstrating that the true wavefunction of the system
+    % must be a linear combination of all possible Slater detemerminants built up
+    % from permutations of the single particle functions in the basis. Thus the
+    % energy equation becomes too large.
     In this project we will study the ground state energy of quantum dots.
@@
 
+    \subsection{The coupled cluster approximation}
+        We approximate the true wavefunction, $\ket{\Psi}$, of the system by the
+        coupled cluster wavefunction, $\ket{\Psi_{\text{CC}}}$, defined as
+        \begin{align}
+            \ket{\Psi_{\text{CC}}}
+            &\equiv e^{T}\ket{\Phi_0}.
+        \end{align}
+
 \bibliography{references}

### 2018-04-12 project_2/article/article.tex
Add start of section on CC energy

@@
 \newcommand{\ade}[1]{a_{#1}}
+
+\newcommand{\kslat}{\ket{\Phi_0}}
+\newcommand{\bslat}{\bra{\Phi_0}}
+
+\newcommand{\kcc}{\ket{\Psi_{\text{CC}}}}
+\newcommand{\bcc}{\bra{\Psi_{\text{CC}}}}
 
@@
             \ket{\Psi_{\text{CC}}}
-            &\equiv e^{T}\ket{\Phi_0}.
+            &\equiv e^{T}\ket{\Phi_0}
+            = \para{
+                \sum_{i = 0}^n
+                \frac{1}{n!}T^n
+            }\ket{\Phi_0},
+        \end{align}
+        where the \emph{cluster operator}, $T$, is given by a sum of
+        $p$-excitation operators
+        \begin{align}
+            T &= T_1 + T_2 + \dots \\
+            &=
+            \sum_{ia}t_i^a\acr{a}\ade{i}
+            + \para{\frac{1}{2!}}^2\sum_{ijab}
+            t_{ij}^{ab}\acr{a}\acr{b}\ade{i}\ade{j}
+            + \dots.
+        \end{align}
+        We will be looking at the \emph{doubles} approximation, that is
+        \begin{align}
+            T = T_{\text{CCD}} \equiv T_2.
+        \end{align}
+
+        \subsubsection{Energy of the coupled cluster approximation}
+            When we're going to compute the energy of a system using the coupled
+            cluster approximation we would ideally want to find the expectation
+            value of the energy using the coupled cluster wavefunction.
+            \begin{align}
+                E = \bcc H\kcc.
+            \end{align}
+            As it turns out, this is an uncomfortable way of finding the energy
+            as $T \neq T^{\dagger}$. Instead we will define what we call the
+            \emph{similarity transformed Hamiltonian}. We plug the coupled
+            cluster wavefunction into the Schrödinger equation.
+            \begin{align}
+                H\kcc = E\kcc.
+            \end{align}
+            Next, we left multiply with the inverse of the cluster expansion,
+            i.e.,
+            \begin{align}
+                e^{-T}H\kcc = e^{-T}E\kcc
+                = E \kslat.
+            \end{align}
+            Projecting this equation on the reference state we get
+            \begin{align}
+                E = \bslat e^{-T}H\kcc
+                = \bslat e^{-T}He^{T}\kslat,
+            \end{align}
+            where in the latter inner-product we have located the similarity
+            transformed Hamiltonian defined by
+            \begin{align}
+                \bar{H} \equiv e^{-T}He^{T}.
+                \label{eq:similarity_transformed_hamiltonian}
             \end{align}

### 2018-04-12 project_2/article/article.tex
Add more theory on CC

@@
 \usepackage[toc, page]{appendix}
+\usepackage{simpler-wick}
 
@@
 
-        \subsubsection{Energy of the coupled cluster approximation}
+    \subsection{Energy of the coupled cluster approximation}
         When we're going to compute the energy of a system using the coupled
@@
 
+        \subsubsection{Normal ordered Hamiltonian}
+            To get to the coupled cluster equations we need to write the
+            Hamiltonian on a normal ordered form. We use Wick's theorem on the
+            Fermi vacuum\footnote{Fermi vacuum corresponds to the reference
+            Slater.}. The one-body part of the Hamiltonian thus becomes
+            \begin{align}
+                h &= \sum_{pq}h^p_q\acr{p}\ade{q}
+                = \sum_{pq}h^p_q\para{
+                    \{\acr{p}\ade{q}\}
+                    + \{
+                        \wick{\c a_{p}^{\dagger} \c a_{q}}
+                    \}
+                }
+                \\
+                &= h_N + \sum_{pq}\varepsilon_i.
+            \end{align}
+            The two-body part yields
+            \begin{align}
+                v &= \sum_{pqrs}w^{pq}_{rs}\acr{p}\acr{q}\ade{s}\ade{r}
+            \end{align}
+            % TODO: Fill in the remainder. Preferably in the appendix.
+            We can then write the Hamiltonian in terms of the reference energy
+            and the normal ordered Hamiltonian.
+            \begin{align}
+                H = H_N + \bslat H\kslat.
+            \end{align}
+            The energy equation thus becomes
+            \begin{align}
+                E &= \bslat\bar{H}\kslat
+                = E_{\text{ref}} + \bslat e^{-T}H_N e^T\kslat.
+            \end{align}
+
+        By expanding the exponentials the similarity transformed Hamiltonian and
+        recognizing the commutators we get the Baker-Campbell-Hausdorff
+        expansion.
+        \begin{align}
+            \bar{H}
+            &=
+            H + \com{H}{T} + \half\com{\com{H}{T}}{T} + \dots.
+        \end{align}
+
 \bibliography{references}

### 2018-04-16 project_2/article/article.tex
Remove assumption on the SPF

@@
     \end{align}
-    Next we define the reference Slater determinant as
+    Working in a basis of $L$ single particle functions, $\{\ket{p}\}_{p =
+    1}^L$. We define the reference Slater determinant as
     \begin{align}
@@
     \end{align}
-    where $\ket{p}$ are the single particle functions that satisfies the
-    equation
+    i.e., a tensorproduct of the $N$ first single particle functions, $\ket{i}$,
+    of the system. We call these single particle functions \emph{occupied} as
+    they are contained in the Slater determinant.  We will denote the occupied
+    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the \emph{virtual}
+    states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
+    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
+    functions we can write this as
     \begin{align}
-        h\ket{p} = \varepsilon_p\ket{p}.
+        \{\ket{p}\}_{p = 1}^L
+        = \{\ket{i}\}_{i = 1}^N
+        \cup
+        \{\ket{a}\}_{a = N + 1}^L,
     \end{align}
-    They are the eigenfunctions to the one-body part of the Hamiltonian and are
-    thus harmonic oscillator functions. Note that $\ket{\Phi_0}$ consists of the
-    $N$ first occupied single particle functions. We will denote the occupied
-    indices with $i, j, k, l, \dots \in \{1, \dots, N\}$, the virtual states
-    with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices with
-    $p, q, r, s, \dots \in \{1, \dots, L\}$.
+    i.e., the general indexed states consists of both occupied and virtual
+    states.
 
     \subsection{Second quantization}
-        We now move over to a use second quantization. Employing the creation
-        operators, $\acr{p}$, and the destruction operators, $\ade{p}$, we can
-        write the Hamiltonian as
+        Employing the creation operators, $\acr{p}$, and the destruction
+        operators, $\ade{p}$, we can write the Hamiltonian as
         \begin{align}
@@
         \end{gather}
-        In the case of the eigenstates we have that the one-body operator
-        reduces to a diagonal matrix.
-        \begin{align}
-            h^{p}_{q}
-            &= \bra{p}h\ket{q}
-            = \varepsilon_q\braket{p}{q}
-            = \varepsilon_q\delta_{pq},
-        \end{align}
-        where $\delta_{pq}$ is the Kronecker-Delta.
+        Note that the two-body matrix elements are not antisymmetric yet.
 
@@
         \begin{align}
-            T &= T_1 + T_2 + \dots \\
+            T &= T_1 + T_2 + \dots + T_p \\
             &=
@@
             t_{ij}^{ab}\acr{a}\acr{b}\ade{i}\ade{j}
-            + \dots.
+            + \dots,
         \end{align}
+        where $t_{i\dots}^{a\dots}$
         We will be looking at the \emph{doubles} approximation, that is
@@
             \end{align}
+            In this equation the unknowns are the cluster amplitudes
 

### 2018-04-16 project_2/article/article.tex
Add section on CCD energy

@@
 
+\newcommand{\ecc}{E_{\text{CC}}}
+\newcommand{\eccd}{E_{\text{CCD}}}
+
 \begin{document}
@@
     \begin{align}
-        H &= \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N v(\vf{r}_i, \vf{r}_j).
+        H &= h + W
+        =
+        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N v(\vf{r}_i, \vf{r}_j),
     \end{align}
-    Working in a basis of $L$ single particle functions, $\{\ket{p}\}_{p =
-    1}^L$. We define the reference Slater determinant as
+    where $h$ is the full one-body operator and $W$ the full two-body
+    operator, i.e., over the entire system.  Working in a basis of $L$ single
+    particle functions, $\{\ket{p}\}_{p = 1}^L$. We define the reference Slater
+    determinant as
     \begin{align}
@@
     i.e., the general indexed states consists of both occupied and virtual
-    states.
+    states. Note that the single particle functions are orthonormal, i.e.,
+    \begin{align}
+        \braket{p}{q} = \delta_{pq}.
+    \end{align}
+    We can construct other Slater determinants in this basis by exciting
+    or de-exciting the reference determinant. A general excitation is labeled
+    $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
+    particle functions with indices $i, j, \dots$ from the reference and added
+    $a, b, \dots$. Note that
+    \begin{align}
+        \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
+    \end{align}
+    for any excitation.
 
@@
         We approximate the true wavefunction, $\ket{\Psi}$, of the system by the
-        coupled cluster wavefunction, $\ket{\Psi_{\text{CC}}}$, defined as
+        coupled cluster wavefunction, $\ket{\Psi_{\text{CC}}}$, defined by
         \begin{align}
@@
         where the \emph{cluster operator}, $T$, is given by a sum of
-        $p$-excitation operators
+        $p$-excitation operators labeled $T_p$. They consist of \emph{cluster
+        amplitudes}, $t_{i\dots}^{a\dots}$, and creation and annihilation
+        operators.
         \begin{align}
@@
             t_{ij}^{ab}\acr{a}\acr{b}\ade{i}\ade{j}
-            + \dots,
+            + \dots.
         \end{align}
-        where $t_{i\dots}^{a\dots}$
-        We will be looking at the \emph{doubles} approximation, that is
+        In the doubles approximation we limit the cluster operator to
         \begin{align}
-            T = T_{\text{CCD}} \equiv T_2.
+            T \equiv T_2
+            = \frac{1}{4}\sum_{ijab}t_{ij}^{ab}\acr{a}\acr{b}\ade{j}\ade{i}.
+            \label{eq:T_2}
         \end{align}
+        The first part of the coupled cluster method consists of constructing
+        the cluster amplitudes using the \emph{amplitude equations}. After we
+        have found the amplitudes we can compute the energy.
 
@@
         \begin{align}
-            E = \bcc H\kcc.
+            \ecc = \bcc H\kcc.
         \end{align}
@@
         \begin{align}
-            H\kcc = E\kcc.
+            H\kcc = \ecc\kcc.
         \end{align}
@@
         \begin{align}
-            e^{-T}H\kcc = e^{-T}E\kcc
-            = E \kslat.
+            e^{-T}H\kcc = e^{-T}\ecc\kcc
+            = \ecc \kslat.
         \end{align}
@@
         \begin{align}
-            E = \bslat e^{-T}H\kcc
+            \ecc = \bslat e^{-T}H\kcc
             = \bslat e^{-T}He^{T}\kslat,
@@
 
-        \subsubsection{Normal ordered Hamiltonian}
-            To get to the coupled cluster equations we need to write the
-            Hamiltonian on a normal ordered form. We use Wick's theorem on the
-            Fermi vacuum\footnote{Fermi vacuum corresponds to the reference
-            Slater.}. The one-body part of the Hamiltonian thus becomes
+        To simplify the energy equation and the amplitude equations we use the
+        normal ordered Hamiltonian.
+        \begin{align}
+            H = H_N + \bslat H\kslat.
+        \end{align}
+        The energy equation thus becomes
+        \begin{align}
+            \ecc &= \bslat\bar{H}\kslat
+            = E_{\text{ref}} + \bslat e^{-T}H_N e^T\kslat,
+        \end{align}
+        where the reference energy is given by
+        \begin{align}
+            E_{\text{ref}} = \bslat H\kslat.
+        \end{align}
+        We now define the normal ordered similiarity transformed Hamiltonian as
+        \begin{align}
+            \bar{H}_N \equiv e^{-T}H_N e^T.
+        \end{align}
+        By expanding the exponentials of this Hamiltonian and recognizing the
+        commutators we get the Baker-Campbell-Hausdorff expansion.
+        \begin{align}
+            \bar{H}_N
+            &=
+            H_N + \com{H_N}{T} + \frac{1}{2!}\com{\com{H_N}{T}}{T} + \dots.
+        \end{align}
+        From the connected cluster theorem we know that the only nonzero terms
+        in the Baker-Campbell-Hausdorff expansion will be the terms where the
+        normal ordered Hamiltonian has at least one contraction\footnote{In the
+        Wick's theorem sense.} with every cluster operator on its right. This
+        lets us write the expansion as
+        \begin{align}
+            \bar{H}_N
+            &=
+            H_N + \para{H_N T}_c + \frac{1}{2!}\para{H_N T^2}_c + \dots,
+            \label{eq:sim_norm_hamiltionian_expansion}
+        \end{align}
+        where the subscript $c$ signifies that only contributions where at least
+        one contraction between $H_N$ and $T$ has been performed will be
+        included.
+
+        \subsubsection{Coupled cluster doubles energy equation}
+            Using the doubles approximation with the cluster operator $T_2$
+            defined in \autoref{eq:T_2} the energy equation becomes
+            \begin{align}
+                \eccd
+                &=
+                E_{\text{ref}}
+                + \bslat e^{-T_2}H_N e^{T_2}\kslat.
+            \end{align}
+            As the doubles cluster operator doubly excites the reference and
+            using the expansion in \autoref{eq:sim_norm_hamiltionian_expansion}
+            we see that we can write the energy equation as
+            \begin{align}
+                \eccd
+                &=
+                E_{\text{ref}}
+                + \bslat H_N(1 + T_2)\kslat,
+            \end{align}
+            as the Hamiltonian is only able to de-excite one pair of single
+            particle functions.
+
+\bibliography{references}
+
+\appendix
+\section{The normal ordered Hamiltonian}
+    When constructing the normal ordered Hamiltonian we use Wick's theorem to
+    write the one-body, $h$, and the two-body, $W$, operators onto a normal
+    ordered form. Specifically we define the normal ordered form in terms of
+    Fermi vacuum. That is, an operator on normal ordered form destroys the
+    reference Slater determinant.
     \begin{align}
@@
         \\
-                &= h_N + \sum_{pq}\varepsilon_i.
+        &= \sum_{pq}h^p_q\{
+            \acr{p}\ade{q}
+        \}
+        + \sum_{pq}h^p_q\delta_{p \in i}\delta_{pq}
+        \\
+        &= h_N + \sum_{i}h^i_i,
     \end{align}
-            The two-body part yields
-            \begin{align}
-                v &= \sum_{pqrs}w^{pq}_{rs}\acr{p}\acr{q}\ade{s}\ade{r}
-            \end{align}
-            % TODO: Fill in the remainder. Preferably in the appendix.
-            We can then write the Hamiltonian in terms of the reference energy
-            and the normal ordered Hamiltonian.
-            \begin{align}
-                H = H_N + \bslat H\kslat.
-            \end{align}
-            The energy equation thus becomes
-            \begin{align}
-                E &= \bslat\bar{H}\kslat
-                = E_{\text{ref}} + \bslat e^{-T}H_N e^T\kslat.
-            \end{align}
-            In this equation the unknowns are the cluster amplitudes
-
-        By expanding the exponentials the similarity transformed Hamiltonian and
-        recognizing the commutators we get the Baker-Campbell-Hausdorff
-        expansion.
-        \begin{align}
-            \bar{H}
-            &=
-            H + \com{H}{T} + \half\com{\com{H}{T}}{T} + \dots.
-        \end{align}
-
-\bibliography{references}
+    where we have used $\delta_{p \in i}$ to mean that $p$ must be an occupied
+    index.
+    \cite{crawford2007introduction}
+    % TODO: Fill in the rest of this expansion
 

### 2018-04-16 project_2/article/article.tex
Add start of CC amplitude equations

@@
         \braket{\Phi^{ab\dots}_{ij\dots}}{\Phi_0} = 0,
+        \label{eq:excited_overlap}
     \end{align}
@@
             = \ecc \kslat.
+            \label{eq:clean_schrodinger}
         \end{align}
@@
             \ecc &= \bslat\bar{H}\kslat
-            = E_{\text{ref}} + \bslat e^{-T}H_N e^T\kslat,
+            = E_0 + \bslat e^{-T}H_N e^T\kslat,
         \end{align}
@@
         \begin{align}
-            E_{\text{ref}} = \bslat H\kslat.
+            E_0 = \bslat H\kslat.
         \end{align}
@@
                 &=
-                E_{\text{ref}}
+                E_0
                 + \bslat e^{-T_2}H_N e^{T_2}\kslat.
@@
                 &=
-                E_{\text{ref}}
+                E_0
                 + \bslat H_N(1 + T_2)\kslat,
@@
             particle functions.
+
+    \subsection{Coupled cluster amplitude equations}
+        In order for us to solve the energy equation using the coupled cluster
+        approximation we need to figure out what the cluster amplitudes,
+        $t_{ij\dots}^{ab\dots}$, are. This is done by projecting
+        \autoref{eq:clean_schrodinger} onto an excited Slater determinant, i.e.,
+        \begin{align}
+            \bra{\Phi_{ij\dots}^{ab\dots}}e^{-T}He^{T}\kslat
+            = 0.
+        \end{align}
+        Note that in the amplitude equations we can use both the regular and the
+        normal ordered Hamiltonian. They are equal as the reference energy term
+        disappears due to \autoref{eq:excited_overlap}. The order of the
+        excitation in the projection determines the order of the amplitudes you
+        will find. In our case we are only interested in the second order
+        ampltiudes found in the doubles approximation, hence we will solve the
+        equation
+        \begin{align}
+            \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
+        \end{align}
+        to find $t_{ij}^{ab}$.
 

### 2018-04-17 project_2/article/article.tex
Start on two-body normal ordering

@@
     \begin{align}
-        v(\vf{r}_i, \vf{r}_j)
+        w(\vf{r}_i, \vf{r}_j)
         &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
@@
         =
-        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N v(\vf{r}_i, \vf{r}_j),
+        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j),
     \end{align}
@@
         \end{align}
-        to find $t_{ij}^{ab}$.
+        to find an expression that can be used to solve for $t_{ij}^{ab}$.
 
@@
     write the one-body, $h$, and the two-body, $W$, operators onto a normal
-    ordered form. Specifically we define the normal ordered form in terms of
-    Fermi vacuum. That is, an operator on normal ordered form destroys the
-    reference Slater determinant.
+    ordered form. Specifically we define the normal ordered form in terms of the
+    \emph{Fermi vacuum}\footnote{Fermi vacuum defines the reference state, i.e.,
+    $\kslat$, as the vacuum.}. That is, an operator on normal ordered form
+    destroys the reference Slater determinant.
+
+    We start by writing the one-body operator, $h$, to its normal-ordered form.
     \begin{align}
@@
     where we have used $\delta_{p \in i}$ to mean that $p$ must be an occupied
-    index.
+    index. Doing the same for the two-body operator is a slightly more tedious
+    endeavor. For brevity we will only write out the operator strings and only
+    keep the non-zero contributions.
+    \begin{align}
+        \acr{p}\acr{q}\ade{s}\ade{r}
+    \end{align}
     \cite{crawford2007introduction}
-    % TODO: Fill in the rest of this expansion
 

### 2018-04-17 project_2/article/article.tex
Include appendix on normal ordered Hamiltonian

@@
             \sum_{pq}h_{q}^{p}\acr{p}\ade{q}
-            + \sum_{pqrs}w_{rs}^{pq}\acr{p}\acr{q}\ade{s}\ade{r},
+            + \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}\acr{p}\acr{q}\ade{s}\ade{r},
         \end{align}
@@
             h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
-            w_{rs}^{pq} \equiv \bra{pq}v\ket{rs}.
+            \bra{pq}\ket{rs} \equiv \bra{pq}w\ket{rs} - \bra{pq}w\ket{sr}.
         \end{gather}
-        Note that the two-body matrix elements are not antisymmetric yet.
+        Note that we use the chemists notation to label the antisymmetric matrix
+        elements.
 
@@
         \acr{p}\acr{q}\ade{s}\ade{r}
+        &=
+        \{\acr{p}\acr{q}\ade{s}\ade{r}\}
+        + \{
+            \wick{\c a_p^{\dagger} a_{q}^{\dagger} \c a_{s} a_{r}}
+        \}
+        + \{
+            \wick{\c a_p^{\dagger} a_{q}^{\dagger} a_{s} \c a_{r}}
+        \}
+        \nonumber \\
+        &\qquad
+        + \{
+            \wick{a_p^{\dagger} \c a_{q}^{\dagger} \c a_{s} a_{r}}
+        \}
+        + \{
+            \wick{a_p^{\dagger} \c a_{q}^{\dagger} a_{s} \c a_{r}}
+        \}
+        \nonumber \\
+        &\qquad
+        + \{
+            \wick{\c1 a_p^{\dagger} \c2 a_{q}^{\dagger} \c1 a_{s} \c2 a_{r}}
+        \}
+        + \{
+            \wick{\c2 a_p^{\dagger} \c1 a_{q}^{\dagger} \c1 a_{s} \c2 a_{r}}
+        \}
+        \\
+        &=
+        \{\acr{p}\acr{q}\ade{s}\ade{r}\}
+        - \delta_{p \in i}\delta_{ps}\{\acr{q}\ade{r}\}
+        + \delta_{p \in i}\delta_{pr}\{\acr{q}\ade{s}\}
+        \nonumber \\
+        &\qquad
+        + \delta_{q \in i}\delta_{qs}\{\acr{p}\ade{r}\}
+        - \delta_{q \in i}\delta_{qr}\{\acr{p}\ade{s}\}
+        \nonumber \\
+        &\qquad
+        - \delta_{p \in i}\delta_{ps}\delta_{q \in j}\delta_{qr}
+        + \delta_{p \in i}\delta_{pr}\delta_{q \in j}\delta_{qs}.
     \end{align}
-    \cite{crawford2007introduction}
+    Inserted into the full two-body operator and sorting out the sums we get
+    \begin{align}
+        W
+        &=
+        \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
+        \{\acr{p}\acr{q}\ade{s}\ade{r}\}
+        - \frac{1}{4}\sum_{iqr}\bra{iq}\ket{ri}\{\acr{q}\ade{r}\}
+        \nonumber \\
+        &\qquad
+        + \frac{1}{4}\sum_{iqs}\bra{iq}\ket{is}\{\acr{q}\ade{s}\}
+        + \frac{1}{4}\sum_{pir}\bra{pi}\ket{ri}\{\acr{p}\ade{r}\}
+        \nonumber \\
+        &\qquad
+        - \frac{1}{4}\sum_{pis}\bra{pi}\ket{is}\{\acr{p}\ade{s}\}
+        - \frac{1}{4}\sum_{ij}\bra{ij}\ket{ji}
+        \nonumber \\
+        &\qquad
+        + \frac{1}{4}\sum_{ij}\bra{ij}\ket{ij}.
+    \end{align}
+    Using the antisymmetric properties of the two-body matrix elements,
+    \begin{align}
+        \bra{pq}\ket{rs}
+        = - \bra{pq}\ket{sr}
+        = - \bra{qp}\ket{rs}
+        = \bra{qp}\ket{sr},
+    \end{align}
+    and relabeling of the indices we can rearrange and collect some terms.
+    \begin{align}
+        W
+        &=
+        W_N + \sum_{pir}\bra{pi}\ket{ri}\{\acr{p}\ade{r}\}
+        + \half\sum_{ij}\bra{ij}\ket{ij},
+    \end{align}
+    where the normal ordered two-body operator is
+    \begin{align}
+        W_N = \frac{1}{4}\sum_{pqrs}
+        \bra{pq}\ket{rs}\{\acr{p}\acr{q}\ade{s}\ade{r}\}.
+    \end{align}
+    When we now construct the full Hamiltonian we can collect some terms. The
+    constants in both the one-body and the two-body operator in total constitues
+    the reference energy.
+    \begin{align}
+        \bslat H\kslat
+        = \sum_{i}h_i^i + \frac{1}{2}\sum_{ij}\bra{ij}\ket{ij}.
+    \end{align}
+    Combining the normal ordered one-body operator and the second term in the
+    two-body operator, i.e., the term with a single creation and annihilation
+    operator pair, we get the normal ordered Fock-operator.
+    \begin{align}
+        F_N
+        &=
+        \sum_{pq}h_{q}^{p}\{\acr{p}\ade{q}\}
+        + \sum_{pqi}\bra{pi}\ket{qi}\{\acr{p}\ade{q}\}
+        \\
+        &= \sum_{pq}f_{q}^{p}\{\acr{p}\ade{q}\},
+    \end{align}
+    where we have defined the Fock matrix elements as
+    \begin{align}
+        f_{q}^{p}
+        &=
+        h_q^p + \sum_{i}\bra{pi}\ket{qi}.
+    \end{align}
+    In total we get the full Hamiltonian
+    \begin{align}
+        H
+        &=
+        F_N + W_N + \bslat H\kslat
+        \\
+        &= H_N + \bslat H\kslat,
+    \end{align}
+    which is what we wanted to show.\cite{crawford2007introduction}
 

### 2018-04-17 project_2/article/article.tex
Add CCD energy equation

@@
                 E_0
-                + \bslat H_N(1 + T_2)\kslat,
+                + \bslat H_N \kslat + \bslat (H_N T_2)_c\kslat,
             \end{align}
             as the Hamiltonian is only able to de-excite one pair of single
-            particle functions.
+            particle functions. By construction we have that
+            \begin{align}
+                \bslat H_N \kslat = 0.
+            \end{align}
+            In the second term only the normal ordered two-body operator can
+            contribute as the cluster operator gives a total excitation of $+2$.
+            As we are projecting onto the reference we have to de-excite to zero
+            again. The normal ordered Fock operator is at most able to excite
+            and de-excite by $1$ and does therefore not contribute to the
+            overall expression.
+            \begin{align}
+                \bslat (W_N T_2)_c \kslat
+                &= \frac{1}{4}\sum_{ijab}\bra{ij}\ket{ab}t_{ij}^{ab}.
+            \end{align}
+            In total the energy equation reduces to
+            \begin{align}
+                \eccd
+                &= \sum_{i} h_i^i + \half \sum_{ij}\bra{ij}\ket{ij}
+                + \frac{1}{4}\sum_{ijab}\bra{ij}\ket{ab}t_{ij}^{ab},
+            \end{align}
+            where the first two terms come from the reference energy as shown in
+            \autoref{eq:reference_energy}.
 
@@
     \begin{align}
-        \bslat H\kslat
+        E_0 \equiv \bslat H\kslat
         = \sum_{i}h_i^i + \frac{1}{2}\sum_{ij}\bra{ij}\ket{ij}.
+        \label{eq:reference_energy}
     \end{align}

### 2018-04-26 project_1/vmc/vmc/runner.py
Numba test branch

@@ -3,6 +3,7 @@ import time
 import pandas as pd
 import numpy as np
 from .resampling_methods import bootstrap, blocking
+import numba
 
 def sample_local_energies(sampler, wavefunction, parameters, time=True,
         **sampler_kwargs):
@@ -25,6 +26,7 @@ def sample_local_energies(sampler, wavefunction, parameters, time=True,
 
     return df
 
+@numba.jit
 def run_all(sampler, parameters, parameter_names, bootstrap_samples,
         timeit=True, **sampler_kwargs):
 

### 2018-04-26 project_2/article/article.tex
Start on appendix for diagrams
Set mass to unity?
de-excitation?

@@
     where we use natural units $\hbar = c = e = 1$ and set the mass to unity.
+    % Why not write $\hbar = c = e = m = 1$?
     The two-body part is the Coulomb interaction potential.
@@
     We can construct other Slater determinants in this basis by exciting
-    or de-exciting the reference determinant. A general excitation is labeled
+    or relaxing the reference determinant. A general excitation is labeled
     $\ket{\Phi_{ij\dots}^{ab\dots}}$ which means that we have removed the single
@@
             \end{align}
-            as the Hamiltonian is only able to de-excite one pair of single
+            as the Hamiltonian is only able to relax one pair of single
             particle functions. By construction we have that
@@
             contribute as the cluster operator gives a total excitation of $+2$.
-            As we are projecting onto the reference we have to de-excite to zero
+            As we are projecting onto the reference we have to relax to zero
             again. The normal ordered Fock operator is at most able to excite
-            and de-excite by $1$ and does therefore not contribute to the
+            and relax by $1$ and does therefore not contribute to the
             overall expression.
@@
 
+\section{Coupled cluster doubles diagrams}
+    In order to get an expression for the energy equation and the amplitude
+    equations we use a diagrammatic approach.
+
 \end{document}

### 2018-04-27 project_2/article/article.tex

@@
 
+    \subsection{Constructing the matrix elements}
+        Having found the equations needed in order to find an estimate to the
+        ground state energy using the coupled cluster doubles approximation is a
+        well and dandy. But, we need basis functions to create the matrix
+        elements needed to feed into the coupled cluster code.  Often these
+        basis functions are not known and we have to use an approximation or
+        utilize Hartree-Fock to create more optimized basis functions.
+
+        \subsubsection{Harmonic oscillator basis}
+            We will be looking at a system of two-dimensional quantum dots with
+            a Coulomb repulsion.  If we assume, or make it so, that the
+            repulsive two-body part is small we can use the harmonic oscillator
+            functions as our basis functions and compute the matrix elements
+            before feeding it into the coupled cluster code. We will label these
+            basis functions using greek letters, viz.
+            \begin{align}
+                h\ket{\alpha} = \epsilon_{\alpha}\ket{\alpha}.
+            \end{align}
+            In polar coordinates we can write the harmonic oscillator
+            wavefunction for a single particle in two dimensions\footnote{Note
+            that this is without spin. As we are looking at fermions this means
+            that each mode of the harmonic oscillator functions will be repeated
+            twice.}.
+            \begin{align}
+                \phi_{nm}(r, \theta)
+                &=
+                N_{nm}
+                %a\sqrt{\frac{n!}{\pi(n + \abs{m})!}}
+                (ar)^{\abs{m}}L_n^{\abs{m}}(a^2 r^2)e^{-a^2 r^2/2}
+                e^{im\theta},
+            \end{align}
+            where $a = \sqrt{m\omega/\hbar}$ is the Bohr radius, $L_n^{\abs{m}}$
+            is the associated Laguerre polynomials, $n$ and $m$ are the
+            principal and azimuthal quantum numbers respectively and $N_{nm}$ is
+            a normalization constant given by
+            \begin{align}
+                N_{nm}
+                &= a\sqrt{\frac{n!}{\pi(n + \abs{m})!}}.
+            \end{align}
+
+            From E. Anisimovas and A. Matulis\cite{anisimovas1998energy} we can
+            get an analytical expression for the two-body matrix elements.
+
 \bibliography{references}

### 2018-04-30 project_2/article/article.tex
Add basis figure

@@
             a Coulomb repulsion.  If we assume, or make it so, that the
-            repulsive two-body part is small we can use the harmonic oscillator
-            functions as our basis functions and compute the matrix elements
-            before feeding it into the coupled cluster code. We will label these
-            basis functions using greek letters, viz.
-            \begin{align}
-                h\ket{\alpha} = \epsilon_{\alpha}\ket{\alpha}.
-            \end{align}
+            repulsive two-body part is small we can use eigenfunctions of the
+            one-body part our basis. In this case we have two-dimensional
+            harmonic oscillator functions as eigenfunctions. We can then compute
+            the matrix elements, $h_q^p$ and $u_{rs}^{pq}$, before feeding these
+            into the coupled cluster code.
+
             In polar coordinates we can write the harmonic oscillator
-            wavefunction for a single particle in two dimensions\footnote{Note
-            that this is without spin. As we are looking at fermions this means
-            that each mode of the harmonic oscillator functions will be repeated
-            twice.}.
+            wavefunction for a single particle in two dimensions
+            as\footnote{Note that this is without spin. As we are looking at
+            fermions this means that each mode of the harmonic oscillator
+            functions will be repeated twice.}.
             \begin{align}
@@
                 N_{nm}
-                %a\sqrt{\frac{n!}{\pi(n + \abs{m})!}}
                 (ar)^{\abs{m}}L_n^{\abs{m}}(a^2 r^2)e^{-a^2 r^2/2}
@@
             \end{align}
+            Included is also the spin, $\sigma$, of the wavefunction, which can
+            be either up or down. This means each level, $(n, m)$, is doubly
+            occupied. We also have that the wavefunctions are orthonormal
+            \begin{align}
+                \braket{n_1m_1\sigma_1}{n_2m_2\sigma_2}
+                = \delta_{n_1n_2}\delta_{m_1m_2}\delta_{\sigma_1\sigma_2}.
+            \end{align}
+            The eigenenergy of a single harmonic oscillator is given by
+            \begin{align}
+                \epsilon_{nm} &= \hbar\omega(2n + \abs{m} + 1).
+            \end{align}
+            As the wavefunctions are eigenfunctions to the one-body part we will
+            get a diagonal one-body matrix $h$,
+            \begin{align}
+                h\ket{\alpha} = \epsilon_{\alpha}\ket{\alpha},
+            \end{align}
+            where we have labeled the harmonic oscillator basis functions with
+            greek letters, i.e., $\alpha = (n_{\alpha}, m_{\alpha},
+            \sigma_{\alpha})$. Our next job is now to create a mapping between
+            the quantum number $\alpha$ and the three quantum numbers $n$, $m$
+            and $\sigma$.
+
+            \begin{figure}
+                \begin{center}
+                    \begin{tikzpicture}[scale=0.8]
+                        \begin{scope}
+                            \foreach \i in {1, 2, 3} {
+                                \draw(-1, \i - 1) node[anchor=east]
+                                {$\epsilon = \i$};
+                            }
+
+                            % Highest energy level
+                            \foreach \i in {0, 3, 6} {
+                                \draw (\i, 2) -- (\i + 2, 2);
+                                \node at (\i + 0.75, 2) {$\uparrow$};
+                                \node at (\i + 1.25, 2) {$\downarrow$};
+                            }
+                            \node[below, inner sep=.2cm] at (1, 2)
+                            {$n = 0, m = -2$};
+                            \node[below, inner sep=.2cm] at (4, 2)
+                            {$n = 1, m = 0$};
+                            \node[below, inner sep=.2cm] at (7, 2)
+                            {$n = 0, m = 2$};
+
+                            % Middle energy level
+                            \foreach \i in {1.5, 4.5} {
+                                \draw (\i, 1) -- (\i + 2, 1);
+                                \node at (\i + 0.75, 1) {$\uparrow$};
+                                \node at (\i + 1.25, 1) {$\downarrow$};
+                            }
+                            \node[below, inner sep=.2cm] at (2.5, 1)
+                            {$n = 0, m = -1$};
+                            \node[below, inner sep=.2cm] at (5.5, 1)
+                            {$n = 0, m = 1$};
+
+                            % Lowest energy level
+                            \draw (3, 0) -- (5, 0);
+                            \node at (3 + 0.75, 0) {$\uparrow$};
+                            \node at (3 + 1.25, 0) {$\downarrow$};
+                            \node[below, inner sep=.2cm] at (4, 0)
+                            {$n = 0, m = 0$};
+                        \end{scope}
+                    \end{tikzpicture}
+                \end{center}
+                \caption{In this plot we can see the energy degeneracy of the
+                lowest three energy levels in the two-dimensional quantum dot.
+                Each arrow representes a spin up or a spin down state with the
+                quantum numbers $n$ and $m$ as listed below.}
+                \label{fig:basis_states}
+            \end{figure}
 

### 2018-04-30 project_2/article/article.tex
Elaborate on the one-body matrices

@@
             \end{align}
-            As the wavefunctions are eigenfunctions to the one-body part we will
-            get a diagonal one-body matrix $h$,
-            \begin{align}
-                h\ket{\alpha} = \epsilon_{\alpha}\ket{\alpha},
-            \end{align}
-            where we have labeled the harmonic oscillator basis functions with
-            greek letters, i.e., $\alpha = (n_{\alpha}, m_{\alpha},
-            \sigma_{\alpha})$. Our next job is now to create a mapping between
-            the quantum number $\alpha$ and the three quantum numbers $n$, $m$
-            and $\sigma$.
-
+            Our next job is now to create a mapping from the three quantum
+            numbers $n$, $m$ and $\sigma$ to a single quantum number $\alpha$ as
+            the matrices $h$ and $u$ use single indices for each wavefunction.
+            In \autoref{fig:basis_states} we can see the energy levels that
+            needs to be mapped.
             \begin{figure}
@@
 
-            From E. Anisimovas and A. Matulis\cite{anisimovas1998energy} we can
-            get an analytical expression for the two-body matrix elements.
+            Starting from the bottom and working our way upwards from left to
+            right we can label each line from $0$ to $n$ in increasing
+            order. This enumeration will serve as our common quantum number
+            $\alpha$\footnote{Note that we use greek letters $\alpha, \beta,
+            \dots$ for the harmonic oscillator wavefunctions as opposed to latin
+            letters for the general indices.}.
+
+            The normalization condition now reads
+            \begin{align}
+                \braket{\alpha}{\beta} = \delta_{\alpha\beta},
+            \end{align}
+            when $\ket{\alpha}$ and $\ket{\beta}$ share the same spin direction.
+            The one-body matrix will now be a diagonal matrix with the
+            eigenenergies of single particle harmonic oscillator functions as
+            elements.
+            \begin{align}
+                \bra{\alpha}h\ket{\beta} = \epsilon_{\beta}\delta_{\alpha\beta}.
+            \end{align}
+
+            The two-body matrix elements are a little harder to work out as the
+            harmonic oscillator wavefunctions are not eigenfunctions to the
+            correlation operator. Luckily, from E. Anisimovas and A.
+            Matulis\cite{anisimovas1998energy} we can get an analytical
+            expression for the two-body matrix elements.
 

### 2018-04-30 project_2/article/article.tex
Move references after appendix
Add spin-deltas

@@
             \begin{align}
-                \braket{\alpha}{\beta} = \delta_{\alpha\beta},
+                \braket{\alpha}{\beta}
+                = \delta_{\alpha\beta}\delta_{\sigma_\alpha\sigma_\beta},
             \end{align}
-            when $\ket{\alpha}$ and $\ket{\beta}$ share the same spin direction.
             The one-body matrix will now be a diagonal matrix with the
@@
             \begin{align}
-                \bra{\alpha}h\ket{\beta} = \epsilon_{\beta}\delta_{\alpha\beta}.
+                \bra{\alpha}h\ket{\beta}
+                = \epsilon_{\beta}
+                \delta_{\alpha\beta}\delta_{\sigma_\alpha\sigma_\beta}.
             \end{align}
@@
             expression for the two-body matrix elements.
-
-\bibliography{references}
 
@@
 
+\bibliography{references}
+
 \end{document}

### 2018-05-09 project_2/article/article.tex
Add FIXME
Use u instead of W and w

@@
 
+% FIXME: Replace the two-body matrix elements in bra-ket notation with u-tensor.
+
 \section{Introduction}
@@
     \begin{align}
-        w(\vf{r}_i, \vf{r}_j)
+        u(\vf{r}_i, \vf{r}_j)
         &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
@@
     \begin{align}
-        H &= h + W
+        H &= h + u
         =
-        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N w(\vf{r}_i, \vf{r}_j),
+        \sum_{i = 1}^N h(\vf{r}_i) + \sum_{i < j}^N u(\vf{r}_i, \vf{r}_j),
     \end{align}
-    where $h$ is the full one-body operator and $W$ the full two-body
+    where $h$ is the full one-body operator and $u$ the full two-body
     operator, i.e., over the entire system.  Working in a basis of $L$ single
@@
             h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
-            \bra{pq}\ket{rs} \equiv \bra{pq}w\ket{rs} - \bra{pq}w\ket{sr}.
+            \bra{pq}\ket{rs} \equiv \bra{pq}u\ket{rs} - \bra{pq}u\ket{sr}.
         \end{gather}
@@
     When constructing the normal ordered Hamiltonian we use Wick's theorem to
-    write the one-body, $h$, and the two-body, $W$, operators onto a normal
+    write the one-body, $h$, and the two-body, $u$, operators onto a normal
     ordered form. Specifically we define the normal ordered form in terms of the
@@
     \begin{align}
-        W
+        u
         &=

### 2018-05-09 project_2/article/article.tex
Start on section on RHF

@@
             \dots$ for the harmonic oscillator wavefunctions as opposed to latin
-            letters for the general indices.}.
+            letters for the general indices.}. We will only work with full
+            \emph{shells}, i.e., we restrict our views to systems of $N$
+            particles where $N$ will be a magic number which we get by counting
+            all spin states for each energy level and the energy levels below.
+            In \autoref{fig:basis_states} we can see the magic number $N \in [2,
+            6, 12]$.
 
@@
             \end{align}
-            The one-body matrix will now be a diagonal matrix with the
-            eigenenergies of single particle harmonic oscillator functions as
-            elements.
+            The one-body matrix will be a diagonal matrix with the eigenenergies
+            of the single particle harmonic oscillator functions as elements.
             \begin{align}
@@
             harmonic oscillator wavefunctions are not eigenfunctions to the
-            correlation operator. Luckily, from E. Anisimovas and A.
-            Matulis\cite{anisimovas1998energy} we can get an analytical
-            expression for the two-body matrix elements.
+            correlation operator. Luckily, from E. Anisimovas and A.  Matulis
+            (Equation A2)\cite{anisimovas1998energy} we can get an analytical
+            expression for the two-body matrix elements. We can then construct
+            the antisymmetric two-body matrix elements in the harmonic
+            oscillator basis.
+            \begin{align}
+                \bra{\alpha\beta}\ket{\gamma\delta}
+                &= \braket{\alpha\beta}{\gamma\delta}
+                - \braket{\alpha\beta}{\delta\gamma}.
+            \end{align}
+            As we only need to compute these once it is a good idea to save all
+            non-zero values of $\bra{\alpha\beta}\ket{\gamma\delta}$ to a file.
+
+        \subsubsection{Constructing the Hartree-Fock basis}
+            Having found the matrix elements of $h$ and $u$ we can now use the
+            \emph{self-consistent field iteration} method to construct $h$ and
+            $u$ in a \emph{restricted Hartree-Fock} basis. This will yield a
+            better estimate to the actual, unknown basis functions of the
+            system.
 

### 2018-05-11 project_2/article/article.tex
Add link to crawdad
Begin doubles amplitudes

@@
             \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
+            \label{eq:doubles_amplitude_1}
         \end{align}
         to find an expression that can be used to solve for $t_{ij}^{ab}$.
+        By employing the Baker-Campbell-Haussdorf (BCH) expansion we find
+        \begin{equation}
+        \begin{aligned}
+        		\bar{H} 
+        			=& \Big(
+        					H_N + H_NT_1 + H_NT_2 + \frac{1}{2}H_NT_1^2 
+        					+ \frac{1}{2}H_NT_2^2 \\
+        				  &+ H_NT_1T_2 + \frac{1}{6}H_NT_1^3 
+        				    + \frac{1}{2}H_NT_1^2T_2 + \frac{1}{2}H_NT_1T_2^2 \\
+        				  &+ \frac{1}{6}H_NT_2^3 + \frac{1}{24}H_NT_1^4 
+        				    + \frac{1}{6}H_NT_1^3T_2 \\
+        				 &+ \frac{1}{6}H_NT_1^2T_2^2 + \frac{1}{6}H_NT_1T_2^3
+        				   + \frac{1}{24} H_NT_2^4
+        				 \Big)_c.
+        \end{aligned}
+        \label{eq:normal_order_expansion}
+        \end{equation}
+		The subscript $c$ indicates that only those terms in which the 
+		Hamiltonian is connected\footnote{In a Wick's theorem sense}
+		to every cluster operator on its right should be included. Since
+		the Hamiltonian contains at most four annihilation and creation 
+		operators, $H_N$ can connect to at most four cluster operators
+		at once. Therefor, the BCH expansion must truncate at the
+		fourth-order terms.
+
+		Now comes the rather tedious task of evaluating all the terms 
+		that arises from inserting \autoref{eq:normal_order_expansion}
+		into \autoref{eq:doubles_amplitude_1}. The first of such terms
+		is
+		\begin{equation}
+		\begin{aligned}
+			\bra{\Phi_{ij}^{ab}} &(F_N + V_N)\ket{\Phi_0} \\
+			 =& \sum_{pq}f_{pq}\bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa_q\}\ket{\Phi_0} \\
+			 &+\frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
+			 \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0}
+		\end{aligned}
+		\end{equation}
 
@@
             system.
+
+            % See Step 3 in:
+            % http://sirius.chem.vt.edu/wiki/doku.php?id=crawdad:programming:project4
 

### 2018-05-11 project_2/article/article.tex
First term of amplitude equations

@@
 			 &+\frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
-			 \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0}
+			 \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0}.
+		\end{aligned}
+		\end{equation}
+		The one-electron component does not have any full contractions, while the
+		two-electron component produces one contributing integral,
+		\begin{equation}
+		\begin{aligned}
+			\bra{\Phi_{ij}^{ab}} &(V_N)\ket{\Phi_0} \\
+				=& \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
+			 \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0} \\
+			 	=& \frac{1}{4} \sum_{pqrs} \bra{pq}\ket{rs} \\ &\times\Big(
+			 		\{ \wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a
+			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
+			 		+
+			 		\{ \wick{\c1 a^\dagger_i \c2 a^\dagger_j \c4 a_b \c3 a_a
+			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\} \\
+			 	  &\ +
+			 	 	\{ \wick{\c2 a^\dagger_i \c1 a^\dagger_j \c3 a_b \c4 a_a
+			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
+			 		+
+			 		\{ \wick{\c2 a^\dagger_i \c1 a^\dagger_j \c4 a_b \c3 a_a
+			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
+			 	\Big) \\
+			 	=&  \frac{1}{4} \sum_{pqrs} \bra{pq}\ket{rs} 
+			 	( \delta_{ap}\delta_{bq}\delta_{js}\delta_{ir}
+			 	 -\delta_{aq}\delta_{bp}\delta_{js}\delta_{ir} \\
+			     &\ -\delta_{ap}\delta_{bq}\delta_{jr}\delta_{is}
+			 	 +\delta_{aq}\delta_{bp}\delta_{jr}\delta_{is}) = \bra{ab}\ket{ij}.
 		\end{aligned}

### 2018-05-11 project_2/article/article.tex
This was CCD, not CCSD

@@
         to find an expression that can be used to solve for $t_{ij}^{ab}$.
-        By employing the Baker-Campbell-Haussdorf (BCH) expansion we find
+        By employing the Baker-Campbell-Haussdorf (BCH) expansion,
+        while setting $T= T_2$, we find
         \begin{equation}
@@
         			=& \Big(
-        					H_N + H_NT_1 + H_NT_2 + \frac{1}{2}H_NT_1^2 
+        					H_N + H_NT_2 
         					+ \frac{1}{2}H_NT_2^2 \\
-        				  &+ H_NT_1T_2 + \frac{1}{6}H_NT_1^3 
-        				    + \frac{1}{2}H_NT_1^2T_2 + \frac{1}{2}H_NT_1T_2^2 \\
-        				  &+ \frac{1}{6}H_NT_2^3 + \frac{1}{24}H_NT_1^4 
-        				    + \frac{1}{6}H_NT_1^3T_2 \\
-        				 &+ \frac{1}{6}H_NT_1^2T_2^2 + \frac{1}{6}H_NT_1T_2^3
+        				  	&\ + \frac{1}{6}H_NT_2^3 
         				   + \frac{1}{24} H_NT_2^4
-        				 \Big)_c.
+        				 \Big)_C.
         \end{aligned}

### 2018-05-14 project_2/coupled-cluster/coupled_cluster/matrix_elements/generate_matrices.py
Add function generating one-body matrix

@@ -1,7 +1,7 @@
 import sparse
 import pickle
 import numba
-from .coulomb_interface import get_coulomb_element
+from .coulomb_interface import get_coulomb_element, get_energy
 from .index_map import get_indices_nm
 
 ORBITAL_INTEGRALS = None
@@ -71,3 +71,11 @@ def get_antisymmetrized_elements(l: int, filename="") -> sparse.COO:
             pickle.dump(u, f)
 
     return u
+
+def get_one_body_elements(l: int) -> sparse.COO:
+    h = sparse.DOK((l, l))
+
+    for p in range(l):
+        h[p, p] = get_energy(*get_indices_nm(p//2))
+
+    return h.to_coo()

### 2018-05-14 project_2/coupled-cluster/tests/conftest.py
Add test of the one-body matrix

@@ -10,7 +10,9 @@ def spin_delta(p, q):
     return not ((p & 0x1) ^ (q & 0x1))
 
 def pytest_namespace():
-    return {"n": n, "l": l, "orbital_integrals": orbital_integrals, "u": u}
+    return {
+            "n": n, "l": l, "orbital_integrals": orbital_integrals, "u": u,
+            "h": h}
 
 def get_file_orbital_integrals():
     orbital_integrals = sparse.DOK((l//2, l//2, l//2, l//2))
@@ -55,6 +57,23 @@ def get_file_antisymmetrized_integrals():
     return u
 
 u = get_file_antisymmetrized_integrals()
+
+def one_body_integrals():
+    h = sparse.DOK((l, l))
+
+    for p in range(l):
+        if p < 2:
+            h[p, p] = 1.0
+        elif 2 <= p < 6:
+            h[p, p] = 2.0
+        elif 6 <= p < 12:
+            h[p, p] = 3.0
+        else:
+            raise NotImplementedError("Testing only supports l = 12")
+
+    return h.to_coo()
+
+h = one_body_integrals()
 
 @pytest.fixture
 def index_map():

### 2018-05-14 project_2/coupled-cluster/tests/test_matrix_elements.py

@@ -9,7 +9,8 @@ from coupled_cluster.matrix_elements.coulomb_interface import (
 )
 
 from coupled_cluster.matrix_elements.generate_matrices import (
-        get_coulomb_elements, get_antisymmetrized_elements
+        get_coulomb_elements, get_antisymmetrized_elements,
+        get_one_body_elements
 )
 
 def test_two_body_generation_one():
@@ -44,3 +45,10 @@ def test_two_body_antisymmetric_generation():
 
     sparse.utils.assert_eq(
             u, get_antisymmetrized_elements(l), atol=1e-5, rtol=1e-5)
+
+def test_one_body_generation():
+    h = pytest.h
+    l = pytest.l
+
+    sparse.utils.assert_eq(
+            h, get_one_body_elements(l), atol=1e-5, rtol=1e-5)

### 2018-05-15 project_2/coupled-cluster/coupled_cluster/matrix_elements/basis_transformation.py
Add two-body transformation

@@ -7,3 +7,10 @@ def transform_one_body_elements(h, c):
     _h[np.abs(_h) < 1e-8] = 0
 
     return sparse.COO.from_numpy(_h)
+
+def transform_two_body_elements(u, c):
+    _u = np.einsum("pi, qj, kr, ls, ijkl -> pqrs", c.conj(), c.conj(), c, c, u)
+
+    _u[np.abs(_u) < 1e-8] = 0
+
+    return sparse.COO.from_numpy(_u)

### 2018-05-15 project_2/coupled-cluster/coupled_cluster/schemes/cc.py
Add Coupled Cluster superclass

@@ -1,0 +1,50 @@
+import numpy as np
+
+class CoupledCluster:
+
+    def __init__(self, h, u, n, initial_guess=None):
+        self.n = n
+        self.l = len(h)
+        self.m = self.l - self.n
+
+        self.h = h
+        self.u = u
+
+        # Occupied (o) and virtual (v) array slices
+        self.o = slice(0, self.n)
+        self.v = slice(self.n, self.l)
+
+        self._initialize(initial_guess=initial_guess)
+
+    def __err(self):
+        raise NotImplementedError("Use approximation subclass")
+
+    def _initialize(self, initial_guess):
+        self.__err()
+
+    def _compute_energy(self):
+        self.__err()
+
+    def _compute_amplitudes(self):
+        self.__err()
+
+    def compute_reference_energy(self):
+        h, u, o, v = self.h, self.u, self.o, self.v
+        e_ref = np.einsum("ii ->", h[o, o]) \
+                + 0.5*np.einsum("ijij ->", u[o, o, o, o])
+
+        return e_ref
+
+    def compute_energy(self, max_iterations=100, tol=1e-4):
+        iterations = 0
+
+        energy_prev = 100
+        energy = self._compute_energy()
+
+        while abs(energy - energy_prev) > tol and iterations < max_iterations:
+            self._compute_amplitudes()
+            energy_prev = energy
+            energy = self._compute_energy()
+            iterations += 1
+
+        return energy

### 2018-05-15 project_2/coupled-cluster/coupled_cluster/schemes/ccd_sparse.py
Add sparse CCD code

@@ -1,0 +1,147 @@
+import numpy as np
+import sparse
+
+from .cc import CoupledCluster
+
+class CoupledClusterDoublesSparse(CoupledCluster):
+
+    def _initialize(self, initial_guess):
+        self.d = sparse.DOK((self.m, self.m, self.n, self.n))
+        self._compute_d_matrix()
+
+        self.h_sans_diag = sparse.DOK(self.h.shape)
+        for p in range(len(self.h)):
+            for q in range(len(self.h[p])):
+                if p == q:
+                    continue
+
+                self.h_sans_diag[p, q] = self.h[p, q]
+
+        self.h_sans_diag = self.h_sans_diag.to_coo()
+
+        if initial_guess is not None:
+            self.t = initial_guess
+        else:
+            self._compute_initial_guess()
+
+        self.u_bc = sparse.DOK((self.m, self.m))
+        self.u_kj = sparse.DOK((self.n, self.n))
+
+        for b in range(self.m):
+            for c in range(self.m):
+                val = 0
+                for k in range(self.n):
+                    val += self.u[b + self.n, k, c + self.n, k]
+                self.u_bc[b, c] = val
+
+        self.u_bc = self.u_bc.to_coo()
+
+        for k in range(self.n):
+            for j in range(self.n):
+                val = 0
+                for l in range(self.n):
+                    val += self.u[k, l, j, l]
+                self.u_kj[k, j] = val
+
+        self.u_kj = self.u_kj.to_coo()
+
+    def _compute_d_matrix(self):
+        h, n, m = self.h, self.n, self.m
+
+        for a in range(m):
+            for b in range(m):
+                for i in range(n):
+                    for j in range(n):
+                        res = h[i, i] + h[j, j] \
+                                - (h[n + a, n + a] + h[n + b, n + b])
+
+                        if abs(res) < 1e-8:
+                            continue
+
+                        self.d[a, b, i, j] = 1.0/res
+
+        self.d = self.d.to_coo()
+
+    def _compute_initial_guess(self):
+        u, d, o, v = self.u, self.d, self.o, self.v
+        n = self.n
+
[…]

### 2018-05-15 project_2/coupled-cluster/tests/test_hartree_fock.py
Add start of test of HF

@@ -1,0 +1,32 @@
+import numpy as np
+import pytest
+
+from coupled_cluster.schemes.ccd_sparse import CoupledClusterDoublesSparse
+from coupled_cluster.hartree_fock.scf_rhf import scf_rhf
+from coupled_cluster.matrix_elements.basis_transformation import (
+        transform_one_body_elements, transform_two_body_elements
+)
+
+def test_hartree_fock_energy():
+    h = pytest.h
+    u = pytest.u
+    l = pytest.l
+    n = pytest.n
+
+    c, energy = scf_rhf(h.todense(), u, np.eye(l), l, n)
+    _h = transform_one_body_elements(h, c)
+    print (_h.todense())
+
+    _u = transform_two_body_elements(u, c)
+
+    print (energy)
+
+    ccd = CoupledClusterDoublesSparse(h, u, n)
+
+    print (ccd.compute_reference_energy())
+
+    ccd = CoupledClusterDoublesSparse(_h, _u, n)
+
+    print (ccd.compute_reference_energy())
+
+    wat

### 2018-05-15 project_2/article/article.tex
Deltas on second amplitude equation term. Missing four or four to spare?

@@
 			     &\ -\delta_{ap}\delta_{bq}\delta_{jr}\delta_{is}
-			 	 +\delta_{aq}\delta_{bp}\delta_{jr}\delta_{is}) = \bra{ab}\ket{ij}.
+			 	 +\delta_{aq}\delta_{bp}\delta_{jr}\delta_{is}) \\
+			 	 =& \bra{ab}\ket{ij} = u^{ab}_{ij}.
 		\end{aligned}
 		\end{equation}
+
+		Next we would like to evaluate $\bra{\Phi_{ij}^{ab}}(F_N + V_N)T_2\ket{\Phi_0}$. Starting with the 
+		term involving the Fock operator $F_N$,
+		\begin{equation}
+		\begin{aligned}
+			\bra{\Phi_{ij}^{ab}}&F_NT_2\ket{\Phi_0} \\
+				=& \frac{1}{4}\sum_{pq}\sum_{klcd}f_{pq}t_{kl}^{cd} \bra{\Phi_0}
+					\{a_i^\dagger a_j^\dagger a_b a_a \}(\{a^\dagger_pa_q \}\{a^\dagger_c a^\dagger_d a_l a_k \})_c \ket{\Phi_0} \\
+				=& \frac{1}{4}\sum_{pq}\sum_{klcd}f_{pq} t_{kl}^{cd} \\
+				\times \Big(
+					&\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c1 a_l \c5 a_k } \}
+				 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c2 a_l \c5 a_k } \}	\\
+			    + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c1 a_l \c5 a_k } \}
+				 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c2 a_l \c5 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c5 a_l \c2 a_k } \}
+		          + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c5 a_l \c2 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c5 a_l \c1 a_k } \}
+		          + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c5 a_l \c1 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c4 a^\dagger_d \c2 a_l \c1 a_k } \}
+		     	 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c4 a^\dagger_d \c1 a_l \c2 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c3 a^\dagger_d \c2 a_l \c1 a_k } \}
+		          +\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c3 a^\dagger_d \c1 a_l \c2 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c1 a_l \c2 a_k } \}
+		          +\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c2 a_l \c1 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c1 a_l \c2 a_k } \}
+		          + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c3 a^\dagger_c \c5 a^\dagger_d \c2 a_l \c1 a_k } \}
+				\Big) \\
+				=& \sum_{pq}\sum_{klcd}f_{pq} t_{kl}^{cd} \\
+				\times&( - \delta_{a c} \delta_{b d} \delta_{i k} \delta_{j q} \delta_{l p} + \delta_{a c} \delta_{b d} \delta_{i l} \delta_{j q} \delta_{k p}  \\
+				 \ &+ \delta_{a c} \delta_{b d} \delta_{i q} \delta_{j k} \delta_{l p} - \delta_{a c} \delta_{b d} \delta_{i q} \delta_{j l} \delta_{k p} \\
+				 \ &+ \delta_{a c} \delta_{b p} \delta_{d q} \delta_{i k} \delta_{j l} - \delta_{a c} \delta_{b p} \delta_{d q} \delta_{i l} \delta_{j k} \\
+				 \ & + \delta_{a d} \delta_{b c} \delta_{i k} \delta_{j q} \delta_{l p} - \delta_{a d} \delta_{b c} \delta_{i l} \delta_{j q} \delta_{k p} \\
+				 \ &- \delta_{a d} \delta_{b c} \delta_{i q} \delta_{j k} \delta_{l p} + \delta_{a d} \delta_{b c} \delta_{i q} \delta_{j l} \delta_{k p} \\
+				 \ &- \delta_{a d} \delta_{b p} \delta_{c q} \delta_{i k} \delta_{j l} + \delta_{a d} \delta_{b p} \delta_{c q} \delta_{i l} \delta_{j k} \\
+				 \ &- \delta_{a p} \delta_{b c} \delta_{d q} \delta_{i k} \delta_{j l} + \delta_{a p} \delta_{b c} \delta_{d q} \delta_{i l} \delta_{j k} \\
+				 \ &+ \delta_{a p} \delta_{b d} \delta_{c q} \delta_{i k} \delta_{j l} - \delta_{a p} \delta_{b d} \delta_{c q} \delta_{i l} \delta_{j k}) \\
+				 =& (f_{bc}t^{ac}_{ij} - f_{ac}t^{bc}_{ij}) - (f_{jk}t^{ab}_{ik} - f_{ik}t^{ab}_{jk}) \\
+				 =& f_{bc}t^{ac}_{ij}P(ab) - f_{kj}t^{ab}_{ik} \\
+				 =& u_{cm}^{bm}t_{ij}^{ac}P(ab) + u_{jm}^{mk}t_{ik}^{ab}P(ij)
+		\end{aligned}
+		\end{equation}		
+		Where in the last steps we have consigned the sums by Einstein summation notation.
+		We have also only included the two-body part of the fock operator, because we are
+		dealing with double excitations (relaxations) only.
+
+		One could of course continue in this manner, employing Wick's general theorem to all
+		terms that arises from combining \autoref{eq:normal_order_expansion}
+		\autoref{eq:doubles_amplitude_1}, but because this task is quite strenuous, we employ
+		the second quantisation library from SymPy instead\footnote{This is also more in the
+		spirit of this project, as it is in the realm of \emph{Computational} Physics.}.
+		The CCD amplitude equation, from SymPy computation, is 
+		\begin{equation}
+		\begin{aligned}
+			0 =& u^{ab}_{ij} + t^{ac}_{ij} u^{bm}_{cm} P(ab) + t^{ab}_{ik} u^{mk}_{jsm} P(ij) \\
+				 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{nm} u^{nm}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
+				 -& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{nm}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
+				 -& t^{ac}_{im} t^{bd}_{jn} u^{nm}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
+				 +& \frac{1}{2}t^{ab}_{nm} u^{nm}_{ij}, 
+		\end{aligned}
+		\end{equation}
+
+		where we recognize the first three terms.
 

### 2018-05-24 project_2/article/article.tex
Move Wick's to appendix

@@
 		that arises from inserting \autoref{eq:normal_order_expansion}
-		into \autoref{eq:doubles_amplitude_1}. The first of such terms
-		is
-		\begin{equation}
-		\begin{aligned}
-			\bra{\Phi_{ij}^{ab}} &(F_N + V_N)\ket{\Phi_0} \\
-			 =& \sum_{pq}f_{pq}\bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa_q\}\ket{\Phi_0} \\
-			 &+\frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
-			 \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0}.
-		\end{aligned}
-		\end{equation}
-		The one-electron component does not have any full contractions, while the
-		two-electron component produces one contributing integral,
-		\begin{equation}
-		\begin{aligned}
-			\bra{\Phi_{ij}^{ab}} &(V_N)\ket{\Phi_0} \\
-				=& \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
-			 \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0} \\
-			 	=& \frac{1}{4} \sum_{pqrs} \bra{pq}\ket{rs} \\ &\times\Big(
-			 		\{ \wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a
-			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
-			 		+
-			 		\{ \wick{\c1 a^\dagger_i \c2 a^\dagger_j \c4 a_b \c3 a_a
-			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\} \\
-			 	  &\ +
-			 	 	\{ \wick{\c2 a^\dagger_i \c1 a^\dagger_j \c3 a_b \c4 a_a
-			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
-			 		+
-			 		\{ \wick{\c2 a^\dagger_i \c1 a^\dagger_j \c4 a_b \c3 a_a
-			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
-			 	\Big) \\
-			 	=&  \frac{1}{4} \sum_{pqrs} \bra{pq}\ket{rs} 
-			 	( \delta_{ap}\delta_{bq}\delta_{js}\delta_{ir}
-			 	 -\delta_{aq}\delta_{bp}\delta_{js}\delta_{ir} \\
-			     &\ -\delta_{ap}\delta_{bq}\delta_{jr}\delta_{is}
-			 	 +\delta_{aq}\delta_{bp}\delta_{jr}\delta_{is}) \\
-			 	 =& \bra{ab}\ket{ij} = u^{ab}_{ij}.
-		\end{aligned}
-		\end{equation}
-
-		Next we would like to evaluate $\bra{\Phi_{ij}^{ab}}(F_N + V_N)T_2\ket{\Phi_0}$. Starting with the 
-		term involving the Fock operator $F_N$,
-		\begin{equation}
-		\begin{aligned}
-			\bra{\Phi_{ij}^{ab}}&F_NT_2\ket{\Phi_0} \\
-				=& \frac{1}{4}\sum_{pq}\sum_{klcd}f_{pq}t_{kl}^{cd} \bra{\Phi_0}
-					\{a_i^\dagger a_j^\dagger a_b a_a \}(\{a^\dagger_pa_q \}\{a^\dagger_c a^\dagger_d a_l a_k \})_c \ket{\Phi_0} \\
-				=& \frac{1}{4}\sum_{pq}\sum_{klcd}f_{pq} t_{kl}^{cd} \\
-				\times \Big(
-					&\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c1 a_l \c5 a_k } \}
-				 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c2 a_l \c5 a_k } \}	\\
-			    + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c1 a_l \c5 a_k } \}
-				 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c2 a_l \c5 a_k } \} \\
-		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c5 a_l \c2 a_k } \}
-		          + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c5 a_l \c2 a_k } \} \\
-		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c5 a_l \c1 a_k } \}
-		          + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c5 a_l \c1 a_k } \} \\
-		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c4 a^\dagger_d \c2 a_l \c1 a_k } \}
-		     	 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c4 a^\dagger_d \c1 a_l \c2 a_k } \} \\
-		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c3 a^\dagger_d \c2 a_l \c1 a_k } \}
-		          +\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c3 a^\dagger_d \c1 a_l \c2 a_k } \} \\
-		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c1 a_l \c2 a_k } \}
-		          +\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c2 a_l \c1 a_k } \} \\
-		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c1 a_l \c2 a_k } \}
-		          + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c3 a^\dagger_c \c5 a^\dagger_d \c2 a_l \c1 a_k } \}
-				\Big) \\
-				=& \sum_{pq}\sum_{klcd}f_{pq} t_{kl}^{cd} \\
-				\times&( - \delta_{a c} \delta_{b d} \delta_{i k} \delta_{j q} \delta_{l p} + \delta_{a c} \delta_{b d} \delta_{i l} \delta_{j q} \delta_{k p}  \\
-				 \ &+ \delta_{a c} \delta_{b d} \delta_{i q} \delta_{j k} \delta_{l p} - \delta_{a c} \delta_{b d} \delta_{i q} \delta_{j l} \delta_{k p} \\
-				 \ &+ \delta_{a c} \delta_{b p} \delta_{d q} \delta_{i k} \delta_{j l} - \delta_{a c} \delta_{b p} \delta_{d q} \delta_{i l} \delta_{j k} \\
-				 \ & + \delta_{a d} \delta_{b c} \delta_{i k} \delta_{j q} \delta_{l p} - \delta_{a d} \delta_{b c} \delta_{i l} \delta_{j q} \delta_{k p} \\
-				 \ &- \delta_{a d} \delta_{b c} \delta_{i q} \delta_{j k} \delta_{l p} + \delta_{a d} \delta_{b c} \delta_{i q} \delta_{j l} \delta_{k p} \\
-				 \ &- \delta_{a d} \delta_{b p} \delta_{c q} \delta_{i k} \delta_{j l} + \delta_{a d} \delta_{b p} \delta_{c q} \delta_{i l} \delta_{j k} \\
-				 \ &- \delta_{a p} \delta_{b c} \delta_{d q} \delta_{i k} \delta_{j l} + \delta_{a p} \delta_{b c} \delta_{d q} \delta_{i l} \delta_{j k} \\
-				 \ &+ \delta_{a p} \delta_{b d} \delta_{c q} \delta_{i k} \delta_{j l} - \delta_{a p} \delta_{b d} \delta_{c q} \delta_{i l} \delta_{j k}) \\
-				 =& (f_{bc}t^{ac}_{ij} - f_{ac}t^{bc}_{ij}) - (f_{jk}t^{ab}_{ik} - f_{ik}t^{ab}_{jk}) \\
-				 =& f_{bc}t^{ac}_{ij}P(ab) - f_{kj}t^{ab}_{ik} \\
-				 =& u_{cm}^{bm}t_{ij}^{ac}P(ab) + u_{jm}^{mk}t_{ik}^{ab}P(ij)
-		\end{aligned}
-		\end{equation}		
-		Where in the last steps we have consigned the sums by Einstein summation notation.
-		We have also only included the two-body part of the fock operator, because we are
-		dealing with double excitations (relaxations) only.
-
-		One could of course continue in this manner, employing Wick's general theorem to all
-		terms that arises from combining \autoref{eq:normal_order_expansion}
-		\autoref{eq:doubles_amplitude_1}, but because this task is quite strenuous, we employ
-		the second quantisation library from SymPy instead\footnote{This is also more in the
-		spirit of this project, as it is in the realm of \emph{Computational} Physics.}.
+		into \autoref{eq:doubles_amplitude_1}. This can be done by applying
+		Wick's generalised theorem, but the task is a daunting and streneous one.
+		A few example computations of how this can be done is included in
+		\autoref{app:wickOnAmplitude}.
+	 	Instead of doing it in this manner, we employ the second quantisation library
+	 	from SymPy instead\footnote{This is also more in the spirit of this project,
+	 	as it is within the realm of \emph{Computational} Physics.}.
 		The CCD amplitude equation, from SymPy computation, is 
@@
 				 -& t^{ac}_{im} t^{bd}_{jn} u^{nm}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
-				 +& \frac{1}{2}t^{ab}_{nm} u^{nm}_{ij}, 
+				 +& \frac{1}{2}t^{ab}_{nm} u^{nm}_{ij}. 
 		\end{aligned}
 		\end{equation}
-
-		where we recognize the first three terms.
 
@@
 
+\section{Amplitude Equations}
+
+	\label{app:wickOnAmplitude}
+
+	In this section we have provided a few sample computations of how one would
+	evaluate the amplitude equations using wicks theorem. Starting with the simplest
+	term including only the normal-ordered Hamiltonian,
+
+	\begin{equation}
+		\begin{aligned}
+			\bra{\Phi_{ij}^{ab}} &(F_N + V_N)\ket{\Phi_0} \\
+			 =& \sum_{pq}f_{pq}\bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa_q\}\ket{\Phi_0} \\
+			 &+\frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
+			 \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0}.
+		\end{aligned}
+		\end{equation}
+		The one-electron component does not have any full contractions, while the
+		two-electron component produces one contributing integral,
+		\begin{equation}
+		\begin{aligned}
+			\bra{\Phi_{ij}^{ab}} &(V_N)\ket{\Phi_0} \\
+				=& \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
+			 \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0} \\
+			 	=& \frac{1}{4} \sum_{pqrs} \bra{pq}\ket{rs} \\ &\times\Big(
+			 		\{ \wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a
+			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
+			 		+
+			 		\{ \wick{\c1 a^\dagger_i \c2 a^\dagger_j \c4 a_b \c3 a_a
+			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\} \\
+			 	  &\ +
+			 	 	\{ \wick{\c2 a^\dagger_i \c1 a^\dagger_j \c3 a_b \c4 a_a
+			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
+			 		+
+			 		\{ \wick{\c2 a^\dagger_i \c1 a^\dagger_j \c4 a_b \c3 a_a
+			 		\c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
+			 	\Big) \\
+			 	=&  \frac{1}{4} \sum_{pqrs} \bra{pq}\ket{rs} 
+			 	( \delta_{ap}\delta_{bq}\delta_{js}\delta_{ir}
+			 	 -\delta_{aq}\delta_{bp}\delta_{js}\delta_{ir} \\
+			     &\ -\delta_{ap}\delta_{bq}\delta_{jr}\delta_{is}
+			 	 +\delta_{aq}\delta_{bp}\delta_{jr}\delta_{is}) \\
+			 	 =& \bra{ab}\ket{ij} = u^{ab}_{ij}.
+		\end{aligned}
+		\end{equation}
+
+		Next we would like to evaluate $\bra{\Phi_{ij}^{ab}}(F_N + V_N)T_2\ket{\Phi_0}$. Starting with the 
+		term involving the Fock operator $F_N$,
+		\begin{equation}
+		\begin{aligned}
+			\bra{\Phi_{ij}^{ab}}&F_NT_2\ket{\Phi_0} \\
+				=& \frac{1}{4}\sum_{pq}\sum_{klcd}f_{pq}t_{kl}^{cd} \bra{\Phi_0}
+					\{a_i^\dagger a_j^\dagger a_b a_a \}(\{a^\dagger_pa_q \}\{a^\dagger_c a^\dagger_d a_l a_k \})_c \ket{\Phi_0} \\
+				=& \frac{1}{4}\sum_{pq}\sum_{klcd}f_{pq} t_{kl}^{cd} \\
+				\times \Big(
+					&\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c1 a_l \c5 a_k } \}
+				 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c2 a_l \c5 a_k } \}	\\
+			    + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c1 a_l \c5 a_k } \}
+				 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c2 a_l \c5 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c5 a_l \c2 a_k } \}
+		          + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c5 a_l \c2 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c5 a_l \c1 a_k } \}
+		          + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c5 a_l \c1 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c4 a^\dagger_d \c2 a_l \c1 a_k } \}
+		     	 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c4 a^\dagger_d \c1 a_l \c2 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c3 a^\dagger_d \c2 a_l \c1 a_k } \}
+		          +\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c3 a^\dagger_d \c1 a_l \c2 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c1 a_l \c2 a_k } \}
+		          +\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c2 a_l \c1 a_k } \} \\
+		        + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c1 a_l \c2 a_k } \}
+		          + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c3 a^\dagger_c \c5 a^\dagger_d \c2 a_l \c1 a_k } \}
+				\Big) \\
+				=& \sum_{pq}\sum_{klcd}f_{pq} t_{kl}^{cd} \\
+				\times&( - \delta_{a c} \delta_{b d} \delta_{i k} \delta_{j q} \delta_{l p} + \delta_{a c} \delta_{b d} \delta_{i l} \delta_{j q} \delta_{k p}  \\
+				 \ &+ \delta_{a c} \delta_{b d} \delta_{i q} \delta_{j k} \delta_{l p} - \delta_{a c} \delta_{b d} \delta_{i q} \delta_{j l} \delta_{k p} \\
+				 \ &+ \delta_{a c} \delta_{b p} \delta_{d q} \delta_{i k} \delta_{j l} - \delta_{a c} \delta_{b p} \delta_{d q} \delta_{i l} \delta_{j k} \\
+				 \ & + \delta_{a d} \delta_{b c} \delta_{i k} \delta_{j q} \delta_{l p} - \delta_{a d} \delta_{b c} \delta_{i l} \delta_{j q} \delta_{k p} \\
+				 \ &- \delta_{a d} \delta_{b c} \delta_{i q} \delta_{j k} \delta_{l p} + \delta_{a d} \delta_{b c} \delta_{i q} \delta_{j l} \delta_{k p} \\
+				 \ &- \delta_{a d} \delta_{b p} \delta_{c q} \delta_{i k} \delta_{j l} + \delta_{a d} \delta_{b p} \delta_{c q} \delta_{i l} \delta_{j k} \\
+				 \ &- \delta_{a p} \delta_{b c} \delta_{d q} \delta_{i k} \delta_{j l} + \delta_{a p} \delta_{b c} \delta_{d q} \delta_{i l} \delta_{j k} \\
+				 \ &+ \delta_{a p} \delta_{b d} \delta_{c q} \delta_{i k} \delta_{j l} - \delta_{a p} \delta_{b d} \delta_{c q} \delta_{i l} \delta_{j k}) \\
+				 =& (f_{bc}t^{ac}_{ij} - f_{ac}t^{bc}_{ij}) - (f_{jk}t^{ab}_{ik} - f_{ik}t^{ab}_{jk}) \\
+				 =& f_{bc}t^{ac}_{ij}P(ab) - f_{kj}t^{ab}_{ik} \\
+				 =& (h^c_b + u_{cm}^{bm})t_{ij}^{ac}P(ab) + (h^j_m + u_{jm}^{mk})t_{ik}^{ab}P(ij)
+		\end{aligned}
+		\end{equation}		
+		Where in the last steps we have consigned the sums by Einstein summation notation,
+		and split the Fock operator into one- and two-body parts.
+
 \section{Coupled cluster doubles diagrams}

### 2018-05-24 project_2/coupled-cluster/coupled_cluster/matrix_elements/index_map.py
Use new index map generation

@@ -1,44 +1,52 @@
-import pandas as pd
-from .coulomb_interface import get_energy
+import numpy as np
 
-INDEX_MAP = None
+P_LIST = []
+NM_DICT = {}
+SHELL_ARR = None
 
-def _generate_index_map(n_range: int, m_range: int) -> None:
-    global INDEX_MAP
+def generate_index_map(num_shells: int) -> None:
+    global P_LIST, NM_DICT, SHELL_ARR
 
-    INDEX_MAP = pd.DataFrame(
-            index=[(n, m) for n in range(n_range)
-                for m in range(-m_range, m_range + 1)],
-            columns=["energy", "n", "m", "p"])
-    INDEX_MAP.n = list(map(lambda x: x[0], INDEX_MAP.index))
-    INDEX_MAP.m = list(map(lambda x: x[1], INDEX_MAP.index))
-    INDEX_MAP.energy = list(map(lambda x: get_energy(*x), INDEX_MAP.index))
-    INDEX_MAP = INDEX_MAP.sort_values(by=["energy", "m"])
-    INDEX_MAP.p = list(range(len(INDEX_MAP)))
+    P_LIST = []
+    NM_DICT = {}
+    SHELL_ARR = np.cumsum(range(1, num_shells + 1)) * 2
+
+    counter = 0
+    decrease = False
+
+    for shell in range(num_shells + 1):
+        n_max = shell // 2
+        n = 0
+
+        for m in range(-shell, shell + 1, 2):
+            P_LIST.append((n, m))
+            NM_DICT[(n, m)] = counter
+            counter += 1
+
+            if not decrease and n < n_max:
+                n += 1
+            elif not decrease and n == n_max:
+                decrease = True
+                if shell % 2 == 0:
+                    n -= 1
+            elif decrease:
+                n -= 1
+            else:
+                raise Exception((
+                    "Invalid value: decrease = {0}, n = {1}, " \
+                            + "n_max = {2}").format(decrease, n, n_max))
 
 def get_index_p(n: int, m: int) -> int:
-    if INDEX_MAP is None:
-        _generate_index_map(n + 1, m + 1)
-
-    n_mask = INDEX_MAP.n == n
-    if not any(n_mask):
-        _generate_index_map(n + 1, m + 1)
-        n_mask = INDEX_MAP.n == n
-
-    m_mask = INDEX_MAP.m == m
-    if not any(m_mask):
[…]

### 2018-05-24 project_2/coupled-cluster/tests/conftest.py
Add shells to testing

@@ -4,7 +4,9 @@ import sparse
 
 n = 2
 l = 12
+num_shells = 3
 large_l = 72
+large_num_shells = 8
 filename = "tests/dat/coulomb.dat"
 large_filename = "tests/dat/coulomb_L_72.dat"
 
@@ -13,8 +15,10 @@ def spin_delta(p, q):
 
 def pytest_namespace():
     return {
-            "n": n, "l": l, "orbital_integrals": orbital_integrals, "u": u,
-            "h": h, "large_oi": large_oi, "large_l": large_l}
+            "n": n, "l": l, "num_shells": num_shells,
+            "orbital_integrals": orbital_integrals, "u": u,
+            "h": h, "large_oi": large_oi, "large_l": large_l,
+            "large_num_shells": large_num_shells}
 
 def get_file_orbital_integrals(l, filename, tol=1e-8):
     orbital_integrals = sparse.DOK((l//2, l//2, l//2, l//2))

### 2018-05-24 project_2/coupled-cluster/tests/test_hartree_fock.py

@@ -10,11 +10,16 @@ from coupled_cluster.matrix_elements.generate_matrices import (
 from coupled_cluster.hartree_fock.basis_transformation import (
         transform_one_body_elements, transform_two_body_elements,
 )
+from coupled_cluster.matrix_elements.index_map import (
+        generate_index_map
+)
 
 def test_hartree_fock_energy():
     l = pytest.l
     n = pytest.n
+    num_shells = pytest.num_shells
 
+    generate_index_map(num_shells)
     h = get_one_body_elements(l)
     u = get_coulomb_elements(l)
 

### 2018-05-24 project_2/coupled-cluster/tests/test_index_map.py

@@ -1,8 +1,13 @@
+import pytest
+
 from coupled_cluster.matrix_elements.index_map import (
-        get_index_p, get_indices_nm
+        get_index_p, get_indices_nm, generate_index_map
 )
 
 def test_index_map(index_map):
+    num_shells = pytest.num_shells
+    generate_index_map(num_shells)
+
     for p in index_map:
         assert index_map[p] == get_indices_nm(p)
 

### 2018-05-24 project_2/coupled-cluster/tests/test_matrix_elements.py

@@ -1,7 +1,7 @@
 import pytest
 import sparse
 from coupled_cluster.matrix_elements.index_map import (
-        get_indices_nm
+        get_indices_nm, generate_index_map
 )
 
 from coupled_cluster.matrix_elements.coulomb_interface import (
@@ -16,6 +16,9 @@ from coupled_cluster.matrix_elements.generate_matrices import (
 def test_two_body_generation_one():
     orbital_integrals = pytest.orbital_integrals
     l = pytest.l
+    num_shells = pytest.num_shells
+
+    generate_index_map(num_shells)
 
     _p, _q, _r, _s = orbital_integrals.coords
 
@@ -35,6 +38,9 @@ def test_two_body_generation_one():
 def test_two_body_generation():
     orbital_integrals = pytest.orbital_integrals
     l = pytest.l
+    num_shells = pytest.num_shells
+
+    generate_index_map(num_shells)
 
     sparse.utils.assert_eq(
             orbital_integrals, get_coulomb_elements(l), atol=1e-5, rtol=1e-5)
@@ -42,6 +48,9 @@ def test_two_body_generation():
 def test_two_body_antisymmetric_generation():
     u = pytest.u
     l = pytest.l
+    num_shells = pytest.num_shells
+
+    generate_index_map(num_shells)
 
     sparse.utils.assert_eq(
             u, get_antisymmetrized_elements(l), atol=1e-5, rtol=1e-5)
@@ -49,6 +58,9 @@ def test_two_body_antisymmetric_generation():
 def test_one_body_generation():
     h = pytest.h
     l = pytest.l
+    num_shells = pytest.num_shells
+
+    generate_index_map(num_shells)
 
     sparse.utils.assert_eq(
             h, get_one_body_elements_spin(l), atol=1e-5, rtol=1e-5)
@@ -56,6 +68,27 @@ def test_one_body_generation():
 def test_large_file():
     l = pytest.large_l
     orbital_integrals = pytest.large_oi
+    num_shells = pytest.large_num_shells
 
-    sparse.utils.assert_eq(
-            orbital_integrals, get_coulomb_elements(l), atol=1e-5, rtol=1e-5)
+    generate_index_map(num_shells)
+
+    #sparse.utils.assert_eq(
+    #        orbital_integrals, get_coulomb_elements(l), atol=1e-5, rtol=1e-5)
+    u = get_coulomb_elements(l)
+
[…]

### 2018-05-24 project_2/article/article.tex

@@
 		\begin{aligned}
-			0 =& u^{ab}_{ij} + t^{ac}_{ij} u^{bm}_{cm} P(ab) + t^{ab}_{ik} u^{mk}_{jsm} P(ij) \\
-				 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{nm} u^{nm}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
-				 -& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{nm}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
-				 -& t^{ac}_{im} t^{bd}_{jn} u^{nm}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
-				 +& \frac{1}{2}t^{ab}_{nm} u^{nm}_{ij}. 
+			0 =& u^{ab}_{ij} + f_{bc}t^{ac}_{ij}P(ab) - f_{kj}t^{ab}_{ik}P(ij) \\
+				 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
+				 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
+				 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
+				 +& \frac{1}{2}t^{ab}_{mn} u^{mn}_{ij}. 
+		\label{eq:CCD_amp_1}
 		\end{aligned}
+		\end{equation}
+
+		\subsubsection{The Iterative Scheme}
+		Now we pick only diagonal elements of $f$ to be part of the unperturbed 
+		Hamiltonian and let the other parts be for a while as part of a perturbation. 
+		The second and third terms in \autoref{eq:CCD_amp_1} can now be 
+		rewritten,
+		\begin{equation}
+		\begin{aligned}
+			f_{bc}&t_{ij}^{ab}P(ab) - f_{kj}t_{ik}^{ab}P(ab) \\
+				&= f_{bb}t_{ij}^{ab} - f_{aa}t_{ij}^{ba} - f_{jj}t_{ij}^{ab} + f_{ii}t_{ji}^{ab} \\
+				&= (f_{aa} + f_{bb} - f_{ii} - f_{jj})t_{ij}^{ab} \\
+				&= (\epsilon_a + \epsilon_b - \epsilon_i - \epsilon_j)t_{ij}^{ab} \\
+				&= -(\epsilon_i + \epsilon_j - \epsilon_a - \epsilon_b)t_{ij}^{ab} \\				
+				&= -D_{ij}^{ab}t_{ij}^{ab}.
+		\end{aligned}
+		\end{equation}
+
+		Now, we define a new function consisting of all of \autoref{eq:CCD_amp_1} 
+		except term number two and three,
+		\begin{equation}
+		\begin{aligned}		
+			g(u, \tau) =&
+				    \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
+				 & \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
+				 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
+				 +& \frac{1}{2}t^{ab}_{mn} u^{mn}_{ij}, 
+		\end{aligned}
+		\end{equation}
+
+		so that $D_{ij}^{ab}t_{ij}^{ab} = g(u, \tau)$. This allows us to define an iterative scheme,
+		\begin{equation}
+			\tau^{(k + 1)} = \frac{g(u, \tau^{(k)})}{D_{ij}^{ab}},
+		\end{equation}
+		with the initial guess
+		\begin{equation}
+			\tau^{(0)} = \frac{u_{ij}^{ab}}{D_{ij}^{ab}}.
 		\end{equation}
@@
 				 =& (f_{bc}t^{ac}_{ij} - f_{ac}t^{bc}_{ij}) - (f_{jk}t^{ab}_{ik} - f_{ik}t^{ab}_{jk}) \\
-				 =& f_{bc}t^{ac}_{ij}P(ab) - f_{kj}t^{ab}_{ik} \\
-				 =& (h^c_b + u_{cm}^{bm})t_{ij}^{ac}P(ab) + (h^j_m + u_{jm}^{mk})t_{ik}^{ab}P(ij)
+				 =& f_{bc}t^{ac}_{ij}P(ab) - f_{kj}t^{ab}_{ik}P(ij)
 		\end{aligned}
 		\end{equation}		
-		Where in the last steps we have consigned the sums by Einstein summation notation,
-		and split the Fock operator into one- and two-body parts.
+		Where in the last steps we have consigned the sums by Einstein summation notation.
 

### 2018-05-25 project_2/article/article.tex
Intermediates

@@
 		Now we pick only diagonal elements of $f$ to be part of the unperturbed 
-		Hamiltonian and let the other parts be for a while as part of a perturbation. 
+		Hamiltonian and consider the rest of the terms a perturbation. 
 		The second and third terms in \autoref{eq:CCD_amp_1} can now be 
@@
 		\end{equation}
+
+		\subsubsection{Intermediate Computations}
+		Looking closely at the amplitude equation in (\ref{eq:CCD_amp_1}) one might come
+		to realize that it contains that this equation contains many of the same structures in
+		several of the terms. This warrants the search for an algebraic transformation of the
+		CCD amplitude equation that has the potential to reduce the amount of floating point
+		operations needed to compute it. As it turns out, such terms exist and they will decrease
+		the computing time necessary by an order of magnitude. We will define the following
+		"intermediates",
+		\begin{align}
+			 \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
+    			 \chi^b_c &= u^{bm}_{cm} - \frac{1}{2} t^{bd}_{mn} u^{mn}_{cd} \\
+   			 \chi^{bm}_{jc} &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd} \\
+    			 \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd}.
+		\end{align}
+		These intermediate structures will allow us to rewrite \autoref{eq:CCD_amp_1} to,
+		\begin{equation}
+		\begin{aligned}
+			0 =& \frac{1}{2}t^{ab}_{mn}u^{mn}_{ij} - t^{ab}_{im}u^{mn}_{jn}P(ij)
+      		+ t^{cd}_{ij}\chi^{ab}_{cd} + t^{ac}_{ij}\chi^b_c P(ab) \\
+      	  &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
+      		+ t^{ab}_{in}\chi^n_jP(ij) + u^{ab}_{ij}.
+		\end{aligned}
+		\end{equation}
+		The importance of this "trick" will become apparent in due time.
 

### 2018-06-03 project_2/article/article.tex
Begun intermediates control. Fock operator correction.

@@
         \begin{equation}
-        \begin{aligned}
         		\bar{H} 
-        			=& \Big(
+        			= \Big(
         					H_N + H_NT_2 
-        					+ \frac{1}{2}H_NT_2^2 \\
-        				  	&\ + \frac{1}{6}H_NT_2^3 
-        				   + \frac{1}{24} H_NT_2^4
-        				 \Big)_C.
-        \end{aligned}
+        					+ \frac{1}{2}H_NT_2^2 \Big)_C.
         \label{eq:normal_order_expansion}
@@
 		\begin{aligned}
-			0 =& u^{ab}_{ij} + f_{bc}t^{ac}_{ij}P(ab) - f_{kj}t^{ab}_{ik}P(ij) \\
+			0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
 				 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
@@
 		\subsubsection{The Iterative Scheme}
-		Now we pick only diagonal elements of $f$ to be part of the unperturbed 
+		At first, we pick only diagonal elements of $f$ to be part of the unperturbed 
 		Hamiltonian and consider the rest of the terms a perturbation. 
@@
 		\begin{aligned}
-			f_{bc}&t_{ij}^{ab}P(ab) - f_{kj}t_{ik}^{ab}P(ab) \\
-				&= f_{bb}t_{ij}^{ab} - f_{aa}t_{ij}^{ba} - f_{jj}t_{ij}^{ab} + f_{ii}t_{ji}^{ab} \\
-				&= (f_{aa} + f_{bb} - f_{ii} - f_{jj})t_{ij}^{ab} \\
+			f^b_c&t_{ij}^{ab}P(ab) - f^k_j t_{ik}^{ab}P(ij) \\
+				&\to f^b_bt_{ij}^{ab} - f^a_at_{ij}^{ba} - f^j_jt_{ij}^{ab} + f^i_it_{ji}^{ab} \\
+				&= (f^a_a + f^b_b - f^i_i - f^j_j)t_{ij}^{ab} \\
 				&= (\epsilon_a + \epsilon_b - \epsilon_i - \epsilon_j)t_{ij}^{ab} \\
@@
 
-		Now, we define a new function consisting of all of \autoref{eq:CCD_amp_1} 
+		The we can define a new function consisting of all of \autoref{eq:CCD_amp_1} 
 		except term number two and three,
@@
 			g(u, \tau) =&
-				    \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
-				 & \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
+					u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab) - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
+				 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
+				 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
 				 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
@@
 
-		so that $D_{ij}^{ab}t_{ij}^{ab} = g(u, \tau)$. This allows us to define an iterative scheme,
+		where $\tilde{t}$ are the non-diagonal parts of the Fock matrix.
+
+		Now we have $D_{ij}^{ab}t_{ij}^{ab} = g(u, \tau)$. 
+		This allows us to define an iterative scheme,
 		\begin{equation}
@@
 			 \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
-    			 \chi^b_c &= u^{bm}_{cm} - \frac{1}{2} t^{bd}_{mn} u^{mn}_{cd} \\
+    			 \chi^a_d &= \frac{1}{2} t^{ac}_{nm} u^{nm}_{cd} \\
    			 \chi^{bm}_{jc} &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd} \\
@@
 		These intermediate structures will allow us to rewrite \autoref{eq:CCD_amp_1} to,
+		THIS REMAINS TO BE CHECKED!! 
 		\begin{equation}

### 2018-06-03 project_2/article/article.tex
Intermediates: think I have it now.

@@
 			0 =& \frac{1}{2}t^{ab}_{mn}u^{mn}_{ij} - t^{ab}_{im}u^{mn}_{jn}P(ij)
-      		+ t^{cd}_{ij}\chi^{ab}_{cd} + t^{ac}_{ij}\chi^b_c P(ab) \\
+      		+ t^{cd}_{ij}\chi^{ab}_{cd} - t^{bd}_{ij}\chi^a_d P(ab) \\
       	  &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)

### 2018-06-03 project_2/coupled-cluster/tests/test_coupled_cluster.py
Add test of the coupled cluster energies

@@ -1,0 +1,41 @@
+import numpy as np
+import pytest
+
+from coupled_cluster.schemes.ccd_sparse import CoupledClusterDoublesSparse
+from coupled_cluster.schemes.ccd import CoupledClusterDoubles
+from coupled_cluster.hartree_fock.scf_rhf import scf_rhf
+from coupled_cluster.matrix_elements.generate_matrices import (
+        get_one_body_elements, get_coulomb_elements,
+        get_antisymmetrized_elements, add_spin_to_one_body_elements
+)
+from coupled_cluster.hartree_fock.basis_transformation import (
+        transform_one_body_elements, transform_two_body_elements
+)
+from coupled_cluster.matrix_elements.index_map import (
+        generate_index_map
+)
+
+def test_energy():
+    l = pytest.l
+    n = pytest.n
+    num_shells = pytest.num_shells
+
+    generate_index_map(num_shells)
+    h = get_one_body_elements(l)
+    u = get_coulomb_elements(l)
+
+    c, energy = scf_rhf(h.todense(), u, np.eye(l//2), n//2)
+
+    hi = transform_one_body_elements(h, c)
+    oi = transform_two_body_elements(u, c)
+
+    _h = add_spin_to_one_body_elements(hi, l)
+    _u = get_antisymmetrized_elements(l, oi=oi)
+
+    ccd = CoupledClusterDoubles(_h.todense(), _u.todense(), n)
+    ccd_sparse = CoupledClusterDoublesSparse(_h, _u, n)
+
+    print (ccd.compute_energy())
+    print (ccd_sparse.compute_energy())
+
+    wat

### 2018-06-03 project_2/coupled-cluster/coupled_cluster/schemes/ccd_sparse.py
Use the sparse version of the ampltiude scaling

@@ -59,7 +59,8 @@ class CoupledClusterDoublesSparse(CoupledCluster):
         t_two_body = self._compute_two_body_amplitude()
 
         _t = t_one_body + t_two_body
-        _amplitude_scaling_two_body(_t.coords, _t.data, self.h_dense, self.n)
+        _amplitude_scaling_two_body_sparse(
+                _t.coords, _t.data, self.h_dense, self.n)
 
         self.t = (1 - theta) * _t + theta * self.t
 

### 2018-06-03 project_2/coupled-cluster/coupled_cluster/schemes/ccd.py
Add first draft of CCD equations

@@ -6,10 +6,118 @@ from .cc_interface import _amplitude_scaling_two_body
 class CoupledClusterDoubles(CoupledCluster):
 
     def _initialize(self):
+        o, v = self.o, self.v
+        n, m, l = self.n, self.m, self.l
+
+        self.f = self.h + np.einsum("piqi -> pq", self.u[:, o, :, o])
+
         self._compute_initial_guess()
+
+        # Used for storing the previous value of self.t
+        self._t = np.zeros(self.t.shape)
+
+        # Used for intermediate terms
+        self.term = np.zeros(self.t.shape)
+
+
+        self.off_diag_f_bc = np.zeros((m, m))
+        self.off_diag_f_kj = np.zeros((n, n))
+
+        for b in range(m):
+            for c in range(m):
+                if b == c:
+                    continue
+
+                self.off_diag_f_bc[b, c] = self.f[b + n, c + n]
+
+        for k in range(n):
+            for j in range(n):
+                if k == j:
+                    continue
+
+                self.off_diag_f_kj[k, j] = self.f[k, j]
 
     def _compute_initial_guess(self):
         o, v = self.o, self.v
 
-        self.t = self.u[v, v, o, o]
-        _amplitude_scaling_two_body(self.t, self.h, self.m, self.n)
+        self.t = self.u[v, v, o, o].copy()
+        _amplitude_scaling_two_body(self.t, self.f, self.m, self.n)
+
+
+
+    def _compute_ccd_energy(self):
+        f, u, t, o, v = self.f, self.u, self.t, self.o, self.v
+
+        e_ref = np.einsum("ii ->", f[o, o])
+        e_ref -= 0.5 * np.einsum("ijij ->", u[o, o, o, o])
+        e_ref += 0.25 * np.einsum("abij, abij ->", u[v, v, o, o], t)
+
+        return e_ref
+
+    def _compute_energy(self):
+        return self._compute_ccd_energy()
+
+    def _compute_amplitudes(self, theta):
+        # TODO: Use DIIS
+
+        self._t.fill(0)
+
+        self._compute_one_body_amplitude()
+        self._compute_two_body_amplitude()
+
+        _amplitude_scaling_two_body(self._t, self.f, self.m, self.n)
+
+        self.t = (1 - theta) * self._t + theta * self.t
+
+    def _compute_one_body_amplitude(self):
+        self.term = np.einsum(
+                "bc, acij -> abij", self.off_diag_f_bc, self.t, out=self.term)
[…]

### 2018-06-04 project_2/coupled-cluster/tests/test_coupled_cluster.py
Make stronger test case

@@ -20,7 +20,7 @@ def test_compare_energy():
     n = pytest.n
     num_shells = pytest.num_shells
 
-    convergence_criteria = 1e-5
+    convergence_criteria = 1e-8
 
     generate_index_map(num_shells)
     h = get_one_body_elements(l)
@@ -38,6 +38,6 @@ def test_compare_energy():
     ccd_sparse = CoupledClusterDoublesSparse(_h, _u, n)
 
     energy, _ = ccd.compute_energy(tol=convergence_criteria)
-    energy_sparse, _ = ccd_sparse.compute_energy(tol=convergence_criteria)
+    energy_sparse, _sparse = ccd_sparse.compute_energy(tol=convergence_criteria)
 
     assert abs(energy_sparse - energy) < convergence_criteria

### 2018-06-04 project_2/article/article.tex
Update rewritten amplitude equation

@@
             \begin{aligned}
-			g(u, \tau) =&
+                g(u, t) =&
                         u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab) - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
@@
 
-		Now we have $D_{ij}^{ab}t_{ij}^{ab} = g(u, \tau)$. 
+            Now we have $D_{ij}^{ab}t_{ij}^{ab} = g(u, t)$.
             This allows us to define an iterative scheme,
             \begin{equation}
-			\tau^{(k + 1)} = \frac{g(u, \tau^{(k)})}{D_{ij}^{ab}},
+                t^{(k + 1)} = \frac{g(u, t^{(k)})}{D_{ij}^{ab}},
             \end{equation}
@@
             \begin{equation}
-			\tau^{(0)} = \frac{u_{ij}^{ab}}{D_{ij}^{ab}}.
+                t^{(0)} = \frac{u_{ij}^{ab}}{D_{ij}^{ab}}.
             \end{equation}
@@
             \end{align}
-		These intermediate structures will allow us to rewrite \autoref{eq:CCD_amp_1} to,
-		THIS REMAINS TO BE CHECKED!! 
-		\begin{equation}
-		\begin{aligned}
-			0 =& \frac{1}{2}t^{ab}_{mn}u^{mn}_{ij} - t^{ab}_{im}u^{mn}_{jn}P(ij)
-      		+ t^{cd}_{ij}\chi^{ab}_{cd} - t^{bd}_{ij}\chi^a_d P(ab) \\
-      	  &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
-      		+ t^{ab}_{in}\chi^n_jP(ij) + u^{ab}_{ij}.
-		\end{aligned}
-		\end{equation}
+            These intermediate structures will allow us to rewrite
+            \autoref{eq:CCD_amp_1} to
+            \begin{align}
+                0 &=
+                u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab)
+                - \tilde{f}^k_jt^{ab}_{ik}P(ij)
+                \nonumber\\
+                &\qquad
+                + t^{cd}_{ij}\chi^{ab}_{cd}
+                + t^{ab}_{in}\chi^n_jP(ij)
+                - t^{bd}_{ij}\chi^a_d P(ab)
+                \nonumber\\
+                &\qquad
+                + t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
+                + \frac{1}{2}t^{ab}_{mn}u^{mn}_{ij}.
+            \end{align}
             The importance of this "trick" will become apparent in due time.

### 2018-06-04 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Call super amplitude iteration

@@ -1,6 +1,7 @@
 import numpy as np
 
 from .ccd import CoupledClusterDoubles
+from .cc_interface import amplitude_scaling_two_body
 
 class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
 
@@ -17,8 +18,8 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
         # TODO: Use DIIS
 
         self._compute_intermediates()
+        super(CoupledClusterDoublesOptimized, self)._compute_amplitudes(theta)
 
-        self._t.fill(0)
 
         self._compute_one_body_amplitude()
         self._compute_two_body_amplitude()

### 2018-06-04 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Fix typo in shape

@@ -11,7 +11,7 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
         # Allocate memory for the intermediates
         self.chi_abcd = np.zeros((self.m, self.m, self.m, self.m))
         self.chi_ad = np.zeros((self.m, self.m))
-        self.chi_bmjc = np.zeros((self.m, self.n, self.n, self.n))
+        self.chi_bmjc = np.zeros((self.m, self.n, self.n, self.m))
         self.chi_nj = np.zeros((self.n, self.n))
 
     def _compute_amplitudes(self, theta):

### 2018-06-04 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Add two-body amplitude computation

@@ -20,15 +20,39 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
         self._compute_intermediates()
         super(CoupledClusterDoublesOptimized, self)._compute_amplitudes(theta)
 
-
-        self._compute_one_body_amplitude()
-        self._compute_two_body_amplitude()
-
-    def _compute_one_body_amplitude(self):
-        pass
-
     def _compute_two_body_amplitude(self):
-        pass
+        o, v = self.o, self.v
+
+        self._t += self.u[v, v, o, o]
+
+        self.term = np.einsum(
+                "cdij, abcd -> abij", self.t, self.chi_abcd,
+                out=self.term, optimize="optimal")
+        self._t += self.term
+
+        self.term = np.einsum(
+                "abin, nj -> abij", self.t, self.chi_nj,
+                out=self.term, optimize="optimal")
+        self.term -= self.term.swapaxes(2, 3)
+        self._t += self.term
+
+        self.term = - np.einsum(
+                "bdij, ad -> abij", self.t, self.chi_ad,
+                out=self.term, optimize="optimal")
+        self.term -= self.term.swapaxes(0, 1)
+        self._t += self.term
+
+        self.term = np.einsum(
+                "acim, bmjc -> abij", self.t, self.chi_bmjc,
+                out=self.term, optimize="optimal")
+        self.term -= self.term.swapaxes(0, 1)
+        self.term -= self.term.swapaxes(2, 3)
+        self._t += self.term
+
+        self.term = 0.5 * np.einsum(
+                "abmn, mnij -> abij", self.t, self.u[o, o, o, o],
+                out=self.term, optimize="optimal")
+        self._t += self.term
 
     def _compute_intermediates(self):
         o, v = self.o, self.v

### 2018-06-04 project_2/coupled-cluster/tests/test_coupled_cluster.py
Add test of optimized version

@@ -3,6 +3,7 @@ import pytest
 
 from coupled_cluster.schemes.ccd_sparse import CoupledClusterDoublesSparse
 from coupled_cluster.schemes.ccd import CoupledClusterDoubles
+from coupled_cluster.schemes.ccd_optimized import CoupledClusterDoublesOptimized
 from coupled_cluster.hartree_fock.scf_rhf import scf_rhf
 from coupled_cluster.matrix_elements.generate_matrices import (
         get_one_body_elements, get_coulomb_elements,
@@ -37,7 +38,36 @@ def test_compare_energy():
     ccd = CoupledClusterDoubles(_h.todense(), _u.todense(), n)
     ccd_sparse = CoupledClusterDoublesSparse(_h, _u, n)
 
-    energy, _ = ccd.compute_energy(tol=convergence_criteria)
-    energy_sparse, _sparse = ccd_sparse.compute_energy(tol=convergence_criteria)
+    energy, iterations = ccd.compute_energy(tol=convergence_criteria)
+    energy_sparse, iterations_sparse = ccd_sparse.compute_energy(
+            tol=convergence_criteria)
 
     assert abs(energy_sparse - energy) < convergence_criteria
+
+def test_optimized_energy():
+    l = pytest.l
+    n = pytest.n
+    num_shells = pytest.num_shells
+
+    convergence_criteria = 1e-8
+
+    generate_index_map(num_shells)
+    h = get_one_body_elements(l)
+    u = get_coulomb_elements(l)
+
+    c, energy = scf_rhf(h.todense(), u, np.eye(l//2), n//2)
+
+    hi = transform_one_body_elements(h, c)
+    oi = transform_two_body_elements(u, c)
+
+    _h = add_spin_to_one_body_elements(hi, l).todense()
+    _u = get_antisymmetrized_elements(l, oi=oi).todense()
+
+    ccd = CoupledClusterDoubles(_h, _u, n)
+    ccd_opt = CoupledClusterDoublesOptimized(_h, _u, n)
+
+    energy, iterations = ccd.compute_energy(tol=convergence_criteria)
+    energy_opt, iterations_opt = ccd_opt.compute_energy(
+            tol=convergence_criteria)
+
+    assert abs(energy - energy_opt) < convergence_criteria

### 2018-06-04 project_2/coupled-cluster/coupled_cluster/schemes/ccd.py
Remove whitespace

@@ -43,8 +43,6 @@ class CoupledClusterDoubles(CoupledCluster):
         self.t = self.u[v, v, o, o].copy()
         amplitude_scaling_two_body(self.t, self.f, self.m, self.n)
 
-
-
     def _compute_ccd_energy(self):
         f, u, t, o, v = self.f, self.u, self.t, self.o, self.v
 

### 2018-06-04 project_2/article/article.tex
Thorough intermediate computation in appendix.

--- 2018-06-03 project_2/article/article.tex
@@
 				 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
-				 +& \frac{1}{2}t^{ab}_{mn} u^{mn}_{ij}. 
+				 -& \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}. 
 		\label{eq:CCD_amp_1}
@@
 			g(u, \tau) =&
-					u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab) - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
+					u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
 				 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
@@
 				 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
-				 +& \frac{1}{2}t^{ab}_{mn} u^{mn}_{ij}, 
+				 -& \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
 		\end{aligned}
@@
 
-		where $\tilde{t}$ are the non-diagonal parts of the Fock matrix.
+		where $\tilde{f}$ are the non-diagonal parts of the Fock matrix.
 
@@
             \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
+            \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd} \\
             \chi^a_d &= \frac{1}{2} t^{ac}_{nm} u^{nm}_{cd} \\
-   			 \chi^{bm}_{jc} &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd} \\
-    			 \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd}.
+            \chi^{bm}_{jc} &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd} 
 		\end{align}
 		These intermediate structures will allow us to rewrite \autoref{eq:CCD_amp_1} to,
-		THIS REMAINS TO BE CHECKED!! 
 		\begin{equation}
 		\begin{aligned}
-			0 =& \frac{1}{2}t^{ab}_{mn}u^{mn}_{ij} - t^{ab}_{im}u^{mn}_{jn}P(ij)
-      		+ t^{cd}_{ij}\chi^{ab}_{cd} - t^{bd}_{ij}\chi^a_d P(ab) \\
-      	  &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
-      		+ t^{ab}_{in}\chi^n_jP(ij) + u^{ab}_{ij}.
+            0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
+              &+ t^{cd}_{ij}\chi^{ab}_{cd} + t^{ab}_{in}\chi^n_jP(ij) 
+               - t^{bd}_{ij}\chi^a_dP(ab) \\ 
+              &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij) + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
         \end{aligned}
@@
 		Where in the last steps we have consigned the sums by Einstein summation notation. 
+\section{Finding Amplitude Equation Intermediates}     
+    Starting from the CCD amplitude equation we factor out terms that will have the same
+    indices if we contract them,
+    \begin{align}
+        &\begin{aligned}
+        0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
+		  &+ \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
+		  &+ \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
+		  &+ t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
+		  &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn},            
+        \end{aligned}\\
+        &\begin{aligned}
+        0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
+		  &+ t^{cd}_{ij} \left(\frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2} u^{ab}_{cd} \right) \\
+          &+  t^{ab}_{in} \left(\frac{1}{2} t^{cd}_{jm} u^{mn}_{cd}\right) P(ij) \\
+          &- t^{bd}_{ij} \left(\frac{1}{2}t^{ac}_{nm}  u^{nm}_{cd}\right) P(ab) \\
+          &+ \frac{1}{2}t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{bc}_{im} t^{ad}_{jn} u^{mn}_{cd} P(ij)\\
+          &+ t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
+		  &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}
+        \end{aligned} \\
+        &\begin{aligned}
+        0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
+		  &+ t^{cd}_{ij} \left(\frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2} u^{ab}_{cd} \right) \\
+          &+  t^{ab}_{in} \left(\frac{1}{2} t^{cd}_{jm} u^{mn}_{cd}\right) P(ij) \\
+          &- t^{bd}_{ij} \left(\frac{1}{2}t^{ac}_{nm}  u^{nm}_{cd}\right) P(ab) \\
+          &+ \frac{1}{2}t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ab) P(ij) \\ 
+          &+ t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
+		  &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}
+        \end{aligned} \\
+        &\begin{aligned}
+        0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
+		  &+ t^{cd}_{ij} \left(\frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2} u^{ab}_{cd} \right) \\
+          &+  t^{ab}_{in} \left(\frac{1}{2} t^{cd}_{jm} u^{mn}_{cd}\right) P(ij) \\
+          &- t^{bd}_{ij} \left(\frac{1}{2}t^{ac}_{nm}  u^{nm}_{cd}\right) P(ab) \\
+          &+ t^{ac}_{im}\left(\frac{1}{2} t^{bd}_{jn} u^{mn}_{cd} + u^{bm}_{jc} \right)P(ab) P(ij) \\
+		  &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
+        \end{aligned}
+    \end{align}
+    Now we can introduce the intermediate $\chi$-terms,
+    \begin{align}
+        \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
+        \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd} \\
+        \chi^a_d &= \frac{1}{2} t^{ac}_{nm} u^{nm}_{cd} \\
+        \chi^{bm}_{jc} &= \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd} + u^{bm}_{jc},  
+    \end{align}
+    this gives us,
+    \begin{equation}
+		\begin{aligned}
+            0 =& u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab) - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
+              &+ t^{cd}_{ij}\chi^{ab}_{cd} + t^{ab}_{in}\chi^n_jP(ij) 
+               - t^{bd}_{ij}\chi^a_dP(ab) \\ 
+              &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij) + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
+        \end{aligned}
+	\end{equation}
 

### 2018-06-04 project_2/coupled-cluster/coupled_cluster/schemes/ccd.py
Reuse memory in addition

@@ -62,7 +62,7 @@ class CoupledClusterDoubles(CoupledCluster):
 
         amplitude_scaling_two_body(self._t, self.f, self.m, self.n)
 
-        self.t = (1 - theta) * self._t + theta * self.t
+        self.t = np.add((1 - theta) * self._t, theta * self.t, out=self.t)
 
     def _compute_one_body_amplitude(self):
         self.term = np.einsum(

### 2018-06-04 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Remove DIIS TODO

@@ -5,7 +5,7 @@ from .cc_interface import amplitude_scaling_two_body
 
 class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
 
-    def _initialize(self, max_diis_size=100, **kwargs):
+    def _initialize(self, **kwargs):
         super(CoupledClusterDoublesOptimized, self)._initialize(**kwargs)
 
         # Allocate memory for the intermediates
@@ -15,8 +15,6 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
         self.chi_nj = np.zeros((self.n, self.n))
 
     def _compute_amplitudes(self, theta):
-        # TODO: Use DIIS
-
         self._compute_intermediates()
         super(CoupledClusterDoublesOptimized, self)._compute_amplitudes(theta)
 

### 2018-06-04 project_2/scripts/comparison.py
Add new run

@@ -19,14 +19,13 @@ import os
 file_path = os.path.join("..", "dat")
 filename = os.path.join(file_path, "coulomb_{0}.pkl")
 
-num_shells = 12
+num_shells = 10
 generate_index_map(num_shells)
 
-omega = 1.0
+omega = 0.28
 l = IndexMap.shell_arr[-1]
-n = 2
-theta = 0.1
-#theta = 100
+n = 6
+theta = 0.9
 
 filename = filename.format(l)
 

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Add parallel computation of chi_nj

@@ -5,7 +5,7 @@ from .cc_interface import amplitude_scaling_two_body
 from .helper_ccd import (
         compute_chi_abcd, compute_chi_bmjc, compute_chi_abcd_contraction,
         compute_chi_bmjc_contraction, compute_t_u_contraction,
-        compute_chi_ad
+        compute_chi_ad, compute_chi_nj
 )
 
 class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
@@ -101,14 +101,9 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
         o, v = self.o, self.v
 
         compute_chi_abcd(self.chi_abcd, self.t, self.u, self.n, self.m)
-
         compute_chi_ad(self.chi_ad, self.t, self.u, self.n, self.m)
-
         compute_chi_bmjc(self.chi_bmjc, self.t, self.u, self.n, self.m)
-
-        self.chi_nj = 0.5 * np.einsum(
-                "cdjm, mncd -> nj", self.t, self.u[o, o, v, v],
-                out=self.chi_nj, optimize="optimal")
+        compute_chi_nj(self.chi_nj, self.t, self.u, self.n, self.m)
 
     def _compute_intermediates(self):
         o, v = self.o, self.v

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/helper_ccd.py

@@ -41,7 +41,6 @@ def compute_chi_ad(chi_ad, t, u, n_size, m_size):
 
             chi_ad[a, d] = 0.5 * val
 
-
 @numba.njit(nogil=True, parallel=True)
 def compute_chi_bmjc(chi_bmjc, t, u, n_size, m_size):
     for b in numba.prange(m_size):
@@ -58,6 +57,21 @@ def compute_chi_bmjc(chi_bmjc, t, u, n_size, m_size):
                             val += t[b, d, j, n] * u[m, n, c_virt, d_virt]
 
                     chi_bmjc[b, m, j, c] = 0.5 * val + u[b_virt, m, j, c_virt]
+
+@numba.njit(nogil=True, parallel=True)
+def compute_chi_nj(chi_nj, t, u, n_size, m_size):
+    for n in numba.prange(n_size):
+        for j in range(n_size):
+
+            val = 0
+            for c in range(m_size):
+                c_virt = c + n_size
+                for d in range(m_size):
+                    d_virt = d + n_size
+                    for m in range(n_size):
+                        val += t[c, d, j, m] * u[m, n, c_virt, d_virt]
+
+            chi_nj[n, j] = 0.5 * val
 
 @numba.njit(nogil=True, parallel=True)
 def compute_chi_abcd_contraction(term, t, chi_abcd, n_size, m_size):

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Move parallel functions together

@@ -63,6 +63,14 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
 
         self._t += self.term
 
+    def _compute_intermediates_parallel(self):
+        o, v = self.o, self.v
+
+        compute_chi_abcd(self.chi_abcd, self.t, self.u, self.n, self.m)
+        compute_chi_ad(self.chi_ad, self.t, self.u, self.n, self.m)
+        compute_chi_bmjc(self.chi_bmjc, self.t, self.u, self.n, self.m)
+        compute_chi_nj(self.chi_nj, self.t, self.u, self.n, self.m)
+
     def _compute_two_body_amplitude(self):
         o, v = self.o, self.v
 
@@ -96,14 +104,6 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
                 "abmn, mnij -> abij", self.t, self.u[o, o, o, o],
                 out=self.term, optimize="optimal")
         self._t += self.term
-
-    def _compute_intermediates_parallel(self):
-        o, v = self.o, self.v
-
-        compute_chi_abcd(self.chi_abcd, self.t, self.u, self.n, self.m)
-        compute_chi_ad(self.chi_ad, self.t, self.u, self.n, self.m)
-        compute_chi_bmjc(self.chi_bmjc, self.t, self.u, self.n, self.m)
-        compute_chi_nj(self.chi_nj, self.t, self.u, self.n, self.m)
 
     def _compute_intermediates(self):
         o, v = self.o, self.v

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Add parallel chi_nj contraction

@@ -3,9 +3,10 @@ import numpy as np
 from .ccd import CoupledClusterDoubles
 from .cc_interface import amplitude_scaling_two_body
 from .helper_ccd import (
-        compute_chi_abcd, compute_chi_bmjc, compute_chi_abcd_contraction,
+        compute_chi_abcd, compute_chi_bmjc, compute_chi_ad, compute_chi_nj,
+        compute_chi_abcd_contraction,
         compute_chi_bmjc_contraction, compute_t_u_contraction,
-        compute_chi_ad, compute_chi_nj
+        compute_chi_nj_contraction
 )
 
 class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
@@ -43,10 +44,8 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
                 self.term, self.t, self.chi_abcd, self.n, self.m)
         self._t += self.term
 
-        self.term = np.einsum(
-                "abin, nj -> abij", self.t, self.chi_nj,
-                out=self.term, optimize="optimal")
-        self.term -= self.term.swapaxes(2, 3)
+        compute_chi_nj_contraction(
+                self.term, self.t, self.chi_nj, self.n, self.m)
         self._t += self.term
 
         self.term = - np.einsum(

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/helper_ccd.py

@@ -91,6 +91,23 @@ def compute_chi_abcd_contraction(term, t, chi_abcd, n_size, m_size):
                     term[b, a, j, i] = val
 
 @numba.njit(nogil=True, parallel=True)
+def compute_chi_nj_contraction(term, t, chi_nj, n_size, m_size):
+    for a in numba.prange(m_size):
+        for b in range(a, m_size):
+            for i in range(n_size):
+                for j in range(i, n_size):
+
+                    val = 0
+                    for n in range(n_size):
+                        val += t[a, b, i, n] * chi_nj[n, j]
+                        val -= t[a, b, j, n] * chi_nj[n, i]
+
+                    term[a, b, i, j] = val
+                    term[b, a, i, j] = -val
+                    term[a, b, j, i] = -val
+                    term[b, a, j, i] = val
+
+@numba.njit(nogil=True, parallel=True)
 def compute_chi_bmjc_contraction(term, t, chi_bmjc, n_size, m_size):
     for a in numba.prange(m_size):
         for b in range(a, m_size):

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Add parallel chi_ad contraction

@@ -6,7 +6,7 @@ from .helper_ccd import (
         compute_chi_abcd, compute_chi_bmjc, compute_chi_ad, compute_chi_nj,
         compute_chi_abcd_contraction,
         compute_chi_bmjc_contraction, compute_t_u_contraction,
-        compute_chi_nj_contraction
+        compute_chi_nj_contraction, compute_chi_ad_contraction
 )
 
 class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
@@ -48,10 +48,8 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
                 self.term, self.t, self.chi_nj, self.n, self.m)
         self._t += self.term
 
-        self.term = - np.einsum(
-                "bdij, ad -> abij", self.t, self.chi_ad,
-                out=self.term, optimize="optimal")
-        self.term -= self.term.swapaxes(0, 1)
+        compute_chi_ad_contraction(
+                self.term, self.t, self.chi_ad, self.n, self.m)
         self._t += self.term
 
         compute_chi_bmjc_contraction(

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/helper_ccd.py

@@ -108,6 +108,23 @@ def compute_chi_nj_contraction(term, t, chi_nj, n_size, m_size):
                     term[b, a, j, i] = val
 
 @numba.njit(nogil=True, parallel=True)
+def compute_chi_ad_contraction(term, t, chi_ad, n_size, m_size):
+    for a in numba.prange(m_size):
+        for b in range(a, m_size):
+            for i in range(n_size):
+                for j in range(i, n_size):
+
+                    val = 0
+                    for d in range(m_size):
+                        val -= t[b, d, i, j] * chi_ad[a, d]
+                        val += t[a, d, i, j] * chi_ad[b, d]
+
+                    term[a, b, i, j] = val
+                    term[b, a, i, j] = -val
+                    term[a, b, j, i] = -val
+                    term[b, a, j, i] = val
+
+@numba.njit(nogil=True, parallel=True)
 def compute_chi_bmjc_contraction(term, t, chi_bmjc, n_size, m_size):
     for a in numba.prange(m_size):
         for b in range(a, m_size):

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Add parallel f_bc contraction

@@ -4,7 +4,7 @@ from .ccd import CoupledClusterDoubles
 from .cc_interface import amplitude_scaling_two_body
 from .helper_ccd import (
         compute_chi_abcd, compute_chi_bmjc, compute_chi_ad, compute_chi_nj,
-        compute_chi_abcd_contraction,
+        compute_chi_abcd_contraction, compute_f_bc_t_contraction,
         compute_chi_bmjc_contraction, compute_t_u_contraction,
         compute_chi_nj_contraction, compute_chi_ad_contraction
 )
@@ -25,7 +25,7 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
 
         if self.parallel:
             self._compute_intermediates_parallel()
-            self._compute_one_body_amplitude()
+            self._compute_one_body_amplitude_parallel()
             self._compute_two_body_amplitude_parallel()
         else:
             self._compute_intermediates()
@@ -34,6 +34,16 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
 
         amplitude_scaling_two_body(self._t, self.f, self.m, self.n)
         self.t = np.add((1 - theta) * self._t, theta * self.t, out=self.t)
+
+    def _compute_one_body_amplitude_parallel(self):
+        compute_f_bc_t_contraction(
+                self.term, self.off_diag_f_bc, self.t, self.n, self.m)
+        self._t += self.term
+
+        self.term = np.einsum(
+                "kj, abik -> abij", self.off_diag_f_kj, self.t, out=self.term)
+        self.term -= self.term.swapaxes(2, 3)
+        self._t += self.term
 
     def _compute_two_body_amplitude_parallel(self):
         o, v = self.o, self.v

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/helper_ccd.py

@@ -74,6 +74,23 @@ def compute_chi_nj(chi_nj, t, u, n_size, m_size):
             chi_nj[n, j] = 0.5 * val
 
 @numba.njit(nogil=True, parallel=True)
+def compute_f_bc_t_contraction(term, f, t, n_size, m_size):
+    for a in numba.prange(m_size):
+        for b in range(a, m_size):
+            for i in range(n_size):
+                for j in range(i, n_size):
+
+                    val = 0
+                    for c in range(m_size):
+                        val += f[b, c] * t[a, c, i, j]
+                        val -= f[a, c] * t[b, c, i, j]
+
+                    term[a, b, i, j] = val
+                    term[b, a, i, j] = -val
+                    term[a, b, j, i] = -val
+                    term[b, a, j, i] = val
+
+@numba.njit(nogil=True, parallel=True)
 def compute_chi_abcd_contraction(term, t, chi_abcd, n_size, m_size):
     for a in numba.prange(m_size):
         for b in range(a, m_size):

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Add parallel f_kj contraction

@@ -5,8 +5,9 @@ from .cc_interface import amplitude_scaling_two_body
 from .helper_ccd import (
         compute_chi_abcd, compute_chi_bmjc, compute_chi_ad, compute_chi_nj,
         compute_chi_abcd_contraction, compute_f_bc_t_contraction,
-        compute_chi_bmjc_contraction, compute_t_u_contraction,
-        compute_chi_nj_contraction, compute_chi_ad_contraction
+        compute_f_kj_t_contraction, compute_chi_bmjc_contraction,
+        compute_t_u_contraction, compute_chi_nj_contraction,
+        compute_chi_ad_contraction
 )
 
 class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
@@ -40,9 +41,8 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
                 self.term, self.off_diag_f_bc, self.t, self.n, self.m)
         self._t += self.term
 
-        self.term = np.einsum(
-                "kj, abik -> abij", self.off_diag_f_kj, self.t, out=self.term)
-        self.term -= self.term.swapaxes(2, 3)
+        compute_f_kj_t_contraction(
+                self.term, self.off_diag_f_kj, self.t, self.n, self.m)
         self._t += self.term
 
     def _compute_two_body_amplitude_parallel(self):

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/helper_ccd.py

@@ -91,6 +91,23 @@ def compute_f_bc_t_contraction(term, f, t, n_size, m_size):
                     term[b, a, j, i] = val
 
 @numba.njit(nogil=True, parallel=True)
+def compute_f_kj_t_contraction(term, f, t, n_size, m_size):
+    for a in numba.prange(m_size):
+        for b in range(a, m_size):
+            for i in range(n_size):
+                for j in range(i, n_size):
+
+                    val = 0
+                    for k in range(n_size):
+                        val += f[k, j] * t[a, b, i, k]
+                        val -= f[k, i] * t[a, b, j, k]
+
+                    term[a, b, i, j] = val
+                    term[b, a, i, j] = -val
+                    term[a, b, j, i] = -val
+                    term[b, a, j, i] = val
+
+@numba.njit(nogil=True, parallel=True)
 def compute_chi_abcd_contraction(term, t, chi_abcd, n_size, m_size):
     for a in numba.prange(m_size):
         for b in range(a, m_size):

### 2018-06-05 project_2/article/article.tex
SCF

@@
 
+            The neatest, yet arguably the most abstract, way to write the Hartree-Fock
+            equations for electron $i$\footnote{This is weird, as electron
+            should be indistinguishable and therefor impossible to label.} is
+            \begin{equation}
+                f_i \varphi_i = \varepsilon_i \varphi_i,
+            \end{equation}               
+            where $f_i$ is the Fock operator, $\varphi_i$ are eigenstates of the Fock 
+            operator consiting of a set of one-electorn wave funcitons, called the
+            Hartree-Fock molecular orbitals, and $\varepsilon_i$ are the eigenenergies
+            of the Fock operator. The fock operator, in matrix notation, is given by
+            \begin{equation}
+                F_{pa} = H_{pq} +  J(D)_{pq} - \frac{1}{2}K(D)_{pq},
+            \end{equation}
+            where $h_i$ is the one-body operatore, while the two-body operator is 
+            divided into what we call a direct part,
+            \begin{equation}
+                J(D)_{pq} = \braket{pq}{rs}D_{sr},
+            \end{equation}
+            and an exchange part,
+            \begin{equation}
+                K(D)_{pq} = \braket{ps}{rq}D_{sr}.
+            \end{equation}
+            The direct part of the two-body operator is comparable to classical Coloumb
+            repulsion, while the exchange does not have a classical analog as it arises 
+            from the antisymmetry requirement of the wavefunction. $D_{sr}$ is the 
+            density matrix of the system.
+
+            Introducing a basis set transforms the Hartree-Fock equations into the Rothaan
+            equations
+            \begin{equation}
+                FC = SC\varepsilon.
+            \end{equation}
+            This is a generalised eigenvalue problem where $S$ serves as an overlap matrix
+            that must be there in case of non-orthogonal basis. Since the Fock matrix $F$
+            depends on it's own solution through the orbitals, the eigenvalue problem must
+            be solved iteratively\footnote{This is also the reason why the
+            Hartree-Fock-Roothaan equations are often called the self-consistent-field 
+            procedure.}.
+
 \appendix

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Add cache-aligned parallel runs
        We are not able to beat einsum for the smaller contractions nor
        the intermediate calculations in parallel.

@@ -5,6 +5,7 @@ from .ccd import CoupledClusterDoubles
 from .cc_interface import amplitude_scaling_two_body
 from .helper_ccd import (
         compute_chi_abcd, compute_chi_bmjc, compute_chi_ad, compute_chi_nj,
+        compute_chi_bjcm,
         compute_chi_abcd_contraction, compute_f_bc_t_contraction,
         compute_f_kj_t_contraction, compute_chi_bmjc_contraction,
         compute_t_u_contraction, compute_chi_nj_contraction,
@@ -16,14 +17,28 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
     def _initialize(self, **kwargs):
         super(CoupledClusterDoublesOptimized, self)._initialize(**kwargs)
 
+        o, v = self.o, self.v
+
         # Allocate memory for the intermediates
         self.chi_abcd = np.zeros((self.m, self.m, self.m, self.m))
         self.chi_ad = np.zeros((self.m, self.m))
         self.chi_bmjc = np.zeros((self.m, self.n, self.n, self.m))
+        self.chi_bjcm = np.zeros((self.m, self.n, self.m, self.n))
         self.chi_nj = np.zeros((self.n, self.n))
 
         self._t_ijab = np.zeros((self.n, self.n, self.m, self.m))
+        self._t_iabj = np.zeros((self.n, self.m, self.m, self.n))
+        self._t_aibj = np.zeros((self.m, self.n, self.m, self.n))
+        self._u_hpph = np.zeros((self.n, self.m, self.m, self.n))
+        self._u_phph = np.zeros((self.m, self.n, self.m, self.n))
+        self._u_phph_2 = np.zeros((self.m, self.n, self.m, self.n))
+
         self._t_ijab[:] = self.t.transpose(2, 3, 0, 1)
+        self._t_iabj[:] = self.t.transpose(2, 0, 1, 3)
+        self._t_aibj[:] = self.t.transpose(0, 2, 1, 3)
+        self._u_hpph[:] = self.u[o, o, v, v].transpose(0, 2, 3, 1)
+        self._u_phph[:] = self.u[o, o, v, v].transpose(2, 0, 3, 1)
+        self._u_phph_2[:] = self.u[v, o, o, v].transpose(0, 2, 3, 1)
 
     def _compute_amplitudes(self, theta):
         self._t.fill(0)
[…]

### 2018-06-05 project_2/coupled-cluster/coupled_cluster/schemes/helper_ccd.py

@@ -42,21 +42,33 @@ def compute_chi_ad(chi_ad, t, u, n_size, m_size):
             chi_ad[a, d] = 0.5 * val
 
 @numba.njit(nogil=True, parallel=True)
-def compute_chi_bmjc(chi_bmjc, t, u, n_size, m_size):
+def compute_chi_bmjc(chi_bmjc, t_iabj, u_hpph, u_phhp, n_size, m_size):
     for b in numba.prange(m_size):
-        b_virt = b + n_size
         for m in range(n_size):
             for j in range(n_size):
                 for c in range(m_size):
-                    c_virt = c + n_size
 
                     val = 0
                     for d in range(m_size):
-                        d_virt = d + n_size
                         for n in range(n_size):
-                            val += t[b, d, j, n] * u[m, n, c_virt, d_virt]
+                            val += t_iabj[j, b, d, n] * u_hpph[m, c, d, n]
 
-                    chi_bmjc[b, m, j, c] = 0.5 * val + u[b_virt, m, j, c_virt]
+                    chi_bmjc[b, m, j, c] = 0.5 * val + u_phhp[b, m, j, c]
+
+
+@numba.njit(nogil=True, parallel=True)
+def compute_chi_bjcm(chi_bjcm, t_aibj, u_phph, u_phph_2, n_size, m_size):
+    for b in numba.prange(m_size):
+        for j in range(n_size):
+            for c in range(m_size):
+                for m in range(n_size):
+
+                    val = 0
+                    for d in range(m_size):
+                        for n in range(n_size):
+                            val += t_aibj[b, j, d, n] * u_phph[c, m, d, n]
+
+                    chi_bjcm[b, j, c, m] = 0.5 * val + u_phph_2[b, j, c, m]
 
 @numba.njit(nogil=True, parallel=True)
 def compute_chi_nj(chi_nj, t, u, n_size, m_size):

### 2018-06-05 project_2/article/article.tex
Begun implementation section.

@@
 \usepackage{simpler-wick}
+% Code listings
+\usepackage{listings, lstautogobble}
+% Colors
+\usepackage{xcolor}
 
@@
 \newcommand{\eccd}{E_{\text{CCD}}}
+
+% Defining color
+\definecolor{light-gray}{gray}{0.9}
+
+% Bash language specifier
+\lstnewenvironment{bash}
+{\lstset{
+    basicstyle=\small\ttfamily, language=bash, keywordstyle=\color{blue},
+    backgroundcolor=\color{light-gray}, showstringspaces=false,
+    frame=single, autogobble=true, morekeywords={make, git}}}
+{}
 
@@
 
+   \section{Implmentation}
+        We have costructed a flexible framework for making performing CCD computations,
+        consisting of package for python that is easy to install globally on any 
+        computer.
+
+        \subsection{Installation}
+        The software is easy to install by first cloning the GitHub reopsitory,
+        \begin{bash}
+            git clone git@github.com:Schoyen/FYS4411.git
+        \end{bash}
+        Then all you need to do is change directory to the project code folder
+        where a \lstinline{MakeFile} is included so you only need to build and
+        install,
+        \begin{bash}
+            cd FYS4411/project_2/coupled-cluster
+            make build
+            make install
+        \end{bash}
+
 \appendix

### 2018-06-05 project_2/article/article.tex
Begin implementation section in report

@@
 \usepackage{listings, lstautogobble}
+\usepackage{minted}
 % Colors
 \usepackage{xcolor}
+% Directory tree package
+\usepackage{dirtree, varwidth}
 
@@
     backgroundcolor=\color{light-gray}, showstringspaces=false,
-    frame=single, autogobble=true, morekeywords={make, git}}}
+    frame=single, autogobble=true, morekeywords={make, git},
+    alsoletter={.}}}
 {}
@@
         We have costructed a flexible framework for making performing CCD computations,
-        consisting of package for python that is easy to install globally on any 
+        consisting of package for \lstinline{python} that is easy to install globally on any 
         computer.
@@
         \end{bash}
+        Now everything should be able to run from anywhere on your computer.
+        To ensure that all requirements of the program are satified you can
+        install with \lstinline{make installr} instead.
+
+        \subsection{Program Structure}
+            There are three main subsetions within our program structure,
+            given in the directory tree below.
+
+            %% NONE OF THIS WORKED AS INTENDED.
+            %\begin{figure}
+            %\centering
+            %\framebox[0.5\textwidth]{%
+            %\begin{minipage}{0.49\textwidth}
+            %\begin{center}
+            %\begin{varwidth}{0.49\textwidth}
+            %\dirtree{%
+            %.1 coupled\_cluster.
+            %    .2 matrix\_elements.
+            %        .3 generate\_matrices. 
+            %        .3 index\_map.
+            %    .2 hartree\_fock.
+            %        .3 scf\_rhf.
+            %        .3 basis\_transformation.
+            %    .2 schemes.
+            %        .3 ccd.
+            %        .3 ccd\_sparse.
+            %        .3 ccd\_optimized.
+            %}
+            %\end{varwidth}
+            %\end{center}
+            %\end{minipage}  
+            %}  
+            %\end{figure}
+
+            \vspace{10pt}
+            \dirtree{%
+            .1 coupled\_cluster.
+                .2 matrix\_elements.
+                    .3 generate\_matrices. 
+                    .3 index\_map.
+                .2 hartree\_fock.
+                    .3 scf\_rhf.
+                    .3 basis\_transformation.
+                .2 schemes.
+                    .3 ccd.
+                    .3 ccd\_sparse.
+                    .3 ccd\_optimized.
+            }
+            \vspace{10pt}
+
+            The first subsection, \mintinline{python}{matrix_elemtents},
+            contains methods for computing matrix elements in harmonic oscillator
+            basis from an analytical expression\cite{anisimovas1998energy}. 
+            Computing the matrix elements is a very intensive task, and the central
+            functions are therefore implemented in C++ with a Cython interface to
+            enable use in Python.
+
+            The \mintinline{python}{hartree_fock} subsection contains methods for changing
+            from harmonic oscillator basis to Hartree-Fock basis as well as an 
+            implementation of the self-consistent-field algorithm to make this
+            transaction possible.
+
+            The \mintinline{python}{schemes} subsection is arguably the most important part
+            of this project. The subsection has three different classes that perform
+            the exact same computations, but in different and increasingly intelligent
+            ways. First, \mintinline{python}{CoupledClusterDoubles} is the most 
+            straightforward and naïve way to solve the CCD amplitude equations.
+            Second, because an overbearing amount of the elements in the operator matrices
+            in this problem are zero, we have implemented a sparse matrix CDD solver in
+            \mintinline{python}{CoupledClusterDoublesSparse}. Third,
+            \mintinline{python}{CoupledClusterDoublesOptimized} takes advantage of 
+            sparse matrices as well, but is also parallellized and optimized with
+            memory use and number of floating point operations in mind.
 
@@
 
+\section{Example Script}
+
+    \vspace{10pt}
+    \begin{widetext}
+    \inputminted[]{python}{../scripts/comparison.py}
+    \end{widetext}
+
 \bibliography{references}

### 2018-06-05 project_2/scripts/comparison.py

@@ -73,19 +73,23 @@ t0 = time.time()
 energy, iterations = ccd_hf_sparse.compute_energy(tol=1e-4, theta=theta)
 t1 = time.time()
 print ("Time spent computing CCD energy with HF basis: {0} sec".format(t1 - t0))
-print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}".format(energy, iterations, (t1 - t0)/iterations))
+print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}"
+    .format(energy, iterations, (t1 - t0)/iterations))
 
 t0 = time.time()
 ccd_hf = CoupledClusterDoublesOptimized(
         _h.todense(), _u.todense(), n, parallel=True)
 t1 = time.time()
-print ("Time spent setting up CCD (opt, parallel) code with HF basis: {0} sec".format(t1 - t0))
+print ("Time spent setting up CCD (opt, parallel) code with HF basis: {0} sec"
+    .format(t1 - t0))
 
 t0 = time.time()
 energy, iterations = ccd_hf.compute_energy(tol=1e-4, theta=theta)
 t1 = time.time()
-print ("Time spent computing CCD (opt, parallel) energy with HF basis: {0} sec".format(t1 - t0))
-print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}".format(energy, iterations, (t1 - t0)/iterations))
+print ("Time spent computing CCD (opt, parallel) energy with HF basis: {0} sec"
+   .format(t1 - t0))
+print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}"
+    .format(energy, iterations, (t1 - t0)/iterations))
 
 __h = omega * get_one_body_elements_spin(l)
 t0 = time.time()
@@ -101,7 +105,8 @@ t0 = time.time()
 energy, iterations = ccd.compute_energy(theta=theta)
 t1 = time.time()
 print ("Time spent computing CCD energy with HO basis: {0} sec".format(t1 - t0))
-print ("\tCCD Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}".format(energy, iterations, (t1 - t0)/iterations))
+print ("\tCCD Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}"
[…]

### 2018-06-06 project_2/coupled-cluster/coupled_cluster/schemes/ccd_optimized.py
Add optimized version of CCD using matmul

@@ -1,5 +1,4 @@
 import numpy as np
-import time
 
 from .ccd import CoupledClusterDoubles
 from .cc_interface import amplitude_scaling_two_body
@@ -22,114 +21,85 @@ class CoupledClusterDoublesOptimized(CoupledClusterDoubles):
         # Allocate memory for the intermediates
         self.chi_abcd = np.zeros((self.m, self.m, self.m, self.m))
         self.chi_ad = np.zeros((self.m, self.m))
-        self.chi_bmjc = np.zeros((self.m, self.n, self.n, self.m))
+        # We use different intermediates for the parallel situation and the
+        # serial implementation as the latter requires the intermediate on the
+        # form of a matrix.
+        if self.parallel:
+            self.chi_bmjc = np.zeros((self.m, self.n, self.n, self.m))
+        else:
+            self.chi_cmbj = np.zeros((m * n, m * n))
         self.chi_nj = np.zeros((self.n, self.n))
 
-        self.chi_cmbj = np.zeros((m * n, m * n))
+        # Allocate memory for blocks
+        if self.parallel:
+            self._t_ijab = self.t.transpose(2, 3, 0, 1).copy()
+            self._t_iabj = self.t.transpose(2, 0, 1, 3).copy()
+            self._u_hpph = self.u[o, o, v, v].transpose(0, 2, 3, 1).copy()
+        else:
+            self.u_hhhh = self.u[o, o, o, o]\
+                    .reshape(n * n, n * n)
+            self.u_hhpp = self.u[o, o, v, v]\
+                    .reshape(n * n, m * m)
+            self.u_hhpp_phh_p = self.u[o, o, v, v]\
+                    .transpose(2, 0, 1, 3)\
+                    .reshape(m * n * n, m)
+            self.u_hhpp_phph = self.u[o, o, v, v]\
+                    .transpose(3, 1, 2, 0)\
+                    .reshape(m * n, m * n)
+            self.u_phhp_phph = self.u[v, o, o, v]\
+                    .transpose(0, 2, 3, 1)\
+                    .reshape(m * n, m * n)
+            self.u_hhpp_h_pph = self.u[o, o, v, v]\
+                    .transpose(1, 2, 3, 0)\
+                    .reshape(n, m * m * n)
 
-        self.u_hhhh = self.u[o, o, o, o]\
-                .reshape(n * n, n * n)
-        self.u_hhpp = self.u[o, o, v, v]\
-                .reshape(n * n, m * m)
[…]

### 2018-06-06 project_2/scripts/comparison.py
Add large run with optimal code

@@ -20,7 +20,7 @@ import os
 file_path = os.path.join("..", "dat")
 filename = os.path.join(file_path, "coulomb_{0}.pkl")
 
-num_shells = 8
+num_shells = 12
 generate_index_map(num_shells)
 
 omega = 1.0
@@ -77,7 +77,7 @@ print ("Time spent antisymmetrizing two body elements: {0} sec".format(t1 - t0))
 
 t0 = time.time()
 ccd_hf = CoupledClusterDoublesOptimized(
-        _h.todense(), _u.todense(), n, parallel=True)
+        _h.todense(), _u.todense(), n, parallel=False)
 t1 = time.time()
 print ("Time spent setting up CCD (opt, parallel) code with HF basis: {0} sec"
     .format(t1 - t0))

### 2018-06-06 project_2/scripts/dat_to_pkl.py
Script to convert from .dat to .pkl (anything to anything)

@@ -1,0 +1,21 @@
+import numpy as np
+import sparse
+import os
+import pickle
+
+file_path = os.path.join("..", "dat")
+read_file = os.path.join(file_path, "coulomb_{0}.dat")
+write_file = os.path.join(file_path, "coulomb_{0}.pkl")
+
+l = 90
+read_file = read_file.format(l)
+write_file = write_file.format(l)
+
+numpy_from_file = np.loadtxt(read_file)
+
+data = numpy_from_file[:, 4]
+coords = np.transpose(numpy_from_file[:, 0:4])
+
+sparse_to_file = sparse.COO(coords, data, shape=(l//2, l//2, l//2, l//2))
+
+pickle.dump(sparse_to_file, open(write_file, "wb"))

### 2018-06-06 project_2/coupled-cluster/coupled_cluster/schemes/cc_interface.pyx
Insignificant changes

@@ -1,5 +1,5 @@
 import cython
-from cython.parallel import prange
+#from cython.parallel import prange
 
 import numpy as np
 cimport numpy as np
@@ -50,7 +50,7 @@ def amplitude_scaling_two_body_sparse(
 
     length = len(data)
 
-    for index in prange(length, nogil=True):
+    for index in range(length):
         a = a_arr[index]
         b = b_arr[index]
         i = i_arr[index]

### 2018-06-06 project_2/coupled-cluster/setup.py

@@ -60,7 +60,7 @@ extensions = [
         define_macros=define_macros,
         undef_macros=undef_macros,
         extra_compile_args=["-fopenmp"],
-        extra_link_args=["-fopenmp"]
+	    extra_link_args=["-fopenmp"]
     ),
     Extension(
         name="coupled_cluster.schemes.cc_interface",
@@ -71,7 +71,7 @@ extensions = [
         define_macros=define_macros,
         undef_macros=undef_macros,
         extra_compile_args=["-fopenmp"],
-        extra_link_args=["-fopenmp"]
+	    extra_link_args=["-fopenmp"]
     )
 ]
 

### 2018-06-06 project_2/scripts/comparison.py

@@ -20,7 +20,7 @@ import os
 file_path = os.path.join("..", "dat")
 filename = os.path.join(file_path, "coulomb_{0}.pkl")
 
-num_shells = 12
+num_shells = 10
 generate_index_map(num_shells)
 
 omega = 1.0
@@ -75,6 +75,7 @@ print ("Time spent antisymmetrizing two body elements: {0} sec".format(t1 - t0))
 #print ("Time spent computing CCD energy with HF basis: {0} sec".format(t1 - t0))
 #print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}".format(energy, iterations, (t1 - t0)/iterations))
 
+"""
 t0 = time.time()
 ccd_hf = CoupledClusterDoublesOptimized(
         _h.todense(), _u.todense(), n, parallel=False)
@@ -89,6 +90,7 @@ print ("Time spent computing CCD (opt, parallel) energy with HF basis: {0} sec"
    .format(t1 - t0))
 print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}"
     .format(energy, iterations, (t1 - t0)/iterations))
+"""
 
 __import__("sys").exit()
 

### 2018-06-06 project_2/coupled-cluster/coupled_cluster/schemes/cc_interface.pyx
Revert setup and cc_interface

@@ -1,5 +1,5 @@
 import cython
-#from cython.parallel import prange
+from cython.parallel import prange
 
 import numpy as np
 cimport numpy as np
@@ -50,7 +50,7 @@ def amplitude_scaling_two_body_sparse(
 
     length = len(data)
 
-    for index in range(length):
+    for index in prange(length, nogil=True):
         a = a_arr[index]
         b = b_arr[index]
         i = i_arr[index]

### 2018-06-06 project_2/coupled-cluster/setup.py

@@ -60,7 +60,7 @@ extensions = [
         define_macros=define_macros,
         undef_macros=undef_macros,
         extra_compile_args=["-fopenmp"],
-	    extra_link_args=["-fopenmp"]
+        extra_link_args=["-fopenmp"]
     ),
     Extension(
         name="coupled_cluster.schemes.cc_interface",
@@ -71,7 +71,7 @@ extensions = [
         define_macros=define_macros,
         undef_macros=undef_macros,
         extra_compile_args=["-fopenmp"],
-	    extra_link_args=["-fopenmp"]
+        extra_link_args=["-fopenmp"]
     )
 ]
 

### 2018-06-06 project_2/article/article.tex
Table templates
Table column headers
Add results for N = 12 particles
Table for N=2, missing 11 and 12 shells

@@
 
+\section{Results}
+
+    \begin{table}
+        \centering
+        \caption{N = 2}
+        \begin{ruledtabular}
+            \begin{tabular}{c|c|ccc}
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\ \hline
+                         &  1  & 0.596333 & 0.596333 & 0.596333 \\
+                         &  2  & 0.596333 & 0.512520 & 0.512520 \\
+                         &  3  & 0.526903 & 0.505972 & 0.442235 \\
+                         &  4  & 0.526903 & 0.499216 & 0.442011 \\
+                     0.1 &  5  & 0.525666 & 0.497172 & 0.443293 \\
+                         &  6  & 0.525666 & 0.494232 & 0.443145 \\
+                         &  7  & 0.525635 & 0.493142 & 0.443056 \\
+                         &  8  & 0.525635 & 0.491895 & 0.442981 \\
+                         &  9  & 0.525635 & 0.491262 & 0.491262 \\
+                         &  10 & 0.525635 & 0.490649 & 0.442886 \\
+                         &  11 & & & \\
+                         &  12 &  & & \\ \hline
+
+                         &  1  & 1.886227 & 1.886227 & 1.886227 \\
+                         &  2  & 1.886227 & 1.786914 & 1.786914 \\
+                         &  3  & 1.799856 & 1.778903 & 1.681979 \\
+                         &  4  & 1.799856 & 1.760117 & 1.673881 \\
+                     0.5 &  5  & 1.799748 & 1.754385 & 1.670053 \\
+                         &  6  & 1.799748 & 1.748232 & 1.667804 \\
+                         &  7  & 1.799745 & 1.745231 & 1.666474 \\
+                         &  8  & 1.799745 & 1.742548 & 1.665494 \\
+                         &  9  & 1.799743 & 1.740860 & 1.6648053 \\
+                         &  10 & 1.799743 & 1.739444 & 1.664270 \\
+                         &  11 & & & \\
+                         &  12 &  & & \\ \hline
+
+                         &  1  & 3.253314 & 3.253314 & 3.253314 \\
+                         &  2  & 3.253314 & 3.152328 & 3.152328 \\
+                         &  3  & 3.162691 & 3.141827 & 3.039048 \\
+                         &  4  & 3.162691 & 3.118679 & 3.025273 \\
+                     1.0 &  5  & 3.161921 & 3.110967 & 3.017944 \\
+                         &  6  & 3.161921 & 3.103338 & 3.013923 \\
+                         &  7  & 3.161909 & 3.099324 & 3.011405 \\
+                         &  8  & 3.161909 & 3.095916 & 3.009621 \\
+                         &  9  & 3.161909 & 3.093662 & 3.008343 \\
+                         &  10 & 3.161909 & 3.091818 & 3.007357 \\
+                         &  11 & & & \\
+                         &  12 &  & & \\ \hline
+
+                         &  1  & 5.772454 & 5.772454 & 5.772454 \\
+                         &  2  & 5.772454 & 5.671234 & 5.671234 \\
+                         &  3  & 5.679048 & 5.658272 & 5.553528 \\
+                         &  4  & 5.679048 & 5.631669 & 5.534333 \\
+                     2.0 &  5  & 5.677282 & 5.622092 & 5.523490 \\
+                         &  6  & 5.677282 & 5.613118 & 5.517552 \\
+                         &  7  & 5.677206 & 5.608130 & 5.513709 \\
+                         &  8  & 5.677206 & 5.604026 & 5.511012 \\
+                         &  9  & 5.677204 & 5.601216 & 5.509050 \\
+                         &  10 & 5.677204 & 5.598946 & 5.507540 \\
+                         &  11 &  & & \\
+                         &  12 &  & & \\
+            \end{tabular}
+        \end{ruledtabular}
+        \label{tab:N2}        
+    \end{table}
+
+    \begin{table}
+        \centering
+        \caption{N = 6}
+        \begin{ruledtabular}
+            \begin{tabular}{c|c|ccc}
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
+
+            \end{tabular}
+        \end{ruledtabular}
+        \label{tab:N6}        
+    \end{table}
+
+    \begin{table}
+        \centering
+        \caption{N = 12}
+        \begin{ruledtabular}
+            \begin{tabular}{c|c|ccc}
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
+
+            \end{tabular}
+        \end{ruledtabular}
+        \label{tab:N12}        
+    \end{table}
+
+    \begin{table}
+        \centering
+        \caption{N = 20}
+        \begin{ruledtabular}
+            \begin{tabular}{c|c|ccc}
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
+
+            \end{tabular}
+        \end{ruledtabular}
+        \label{tab:N20}        
+    \end{table}
+
 \appendix

### 2018-06-06 project_2/article/article.tex
Add mixing

--- 2018-06-05 project_2/article/article.tex
@@
 
+\usepackage{fontspec}
+\newfontfamily\DejaSans{DejaVu Sans}
+
+\newcommand{\nan}{\DejaSans 😱}
+
 % Macro for latin-letter vectors
@@
 			\tau^{(k + 1)} = \frac{g(u, \tau^{(k)})}{D_{ij}^{ab}},
+            \label{eq:iterative_amplitude}
 		\end{equation}
@@
 
-   \section{Implmentation}
-        We have costructed a flexible framework for making performing CCD computations,
-        consisting of package for \lstinline{python} that is easy to install globally on any 
-        computer.
+\section{Implementation}
+    We have costructed a flexible framework for making performing CCD
+    computations, consisting of package for \lstinline{python} that is easy
+    to install globally on any computer.
 
@@
         The first subsection, \mintinline{python}{matrix_elemtents},
-            contains methods for computing matrix elements in harmonic oscillator
-            basis from an analytical expression\cite{anisimovas1998energy}. 
-            Computing the matrix elements is a very intensive task, and the central
-            functions are therefore implemented in C++ with a Cython interface to
-            enable use in Python.
-
-            The \mintinline{python}{hartree_fock} subsection contains methods for changing
-            from harmonic oscillator basis to Hartree-Fock basis as well as an 
-            implementation of the self-consistent-field algorithm to make this
-            transaction possible.
-
-            The \mintinline{python}{schemes} subsection is arguably the most important part
-            of this project. The subsection has three different classes that perform
-            the exact same computations, but in different and increasingly intelligent
-            ways. First, \mintinline{python}{CoupledClusterDoubles} is the most 
+        contains methods for computing matrix elements in harmonic
+        oscillator basis from an analytical
+        expression\cite{anisimovas1998energy}.  Computing the matrix
+        elements is a very intensive task, and the central functions are
+        therefore implemented in C++ with a Cython interface to enable use
+        in Python.
+
+        The \mintinline{python}{hartree_fock} subsection contains methods
+        for changing from harmonic oscillator basis to Hartree-Fock basis as
+        well as an implementation of the self-consistent-field algorithm to
+        make this transaction possible.
+
+        The \mintinline{python}{schemes} subsection is arguably the most
+        important part of this project. The subsection has three different
+        classes that perform the exact same computations, but in different
+        and increasingly intelligent ways. First,
+        \mintinline{python}{CoupledClusterDoubles} is the most
         straightforward and naïve way to solve the CCD amplitude equations.
-            Second, because an overbearing amount of the elements in the operator matrices
-            in this problem are zero, we have implemented a sparse matrix CDD solver in
+        Second, because an overbearing amount of the elements in the
+        operator matrices in this problem are zero, we have implemented a
+        sparse matrix CDD solver in
         \mintinline{python}{CoupledClusterDoublesSparse}. Third,
-            \mintinline{python}{CoupledClusterDoublesOptimized} takes advantage of 
-            sparse matrices as well, but is also parallellized and optimized with
-            memory use and number of floating point operations in mind.
+        \mintinline{python}{CoupledClusterDoublesOptimized} takes advantage
+        of sparse matrices as well, but is also parallellized and optimized
+        with memory use and number of floating point operations in mind.
+
+    \subsection{Mixing}
+        A way to try ``massage'' convergence out of the CCD-method is to use
+        \emph{mixing}. A popular solution is Pulay mixing\footnote{Commonly
+        known as DIIS (direct inversion in the iterative subspace).}. We did not
+        implement Pulay mixing but used a more naïve and simpler solution.
+        \begin{align}
+            \tilde{t}^{k + 1} = \theta t^{k + 1} + (1 - \theta)t^k,
+        \end{align}
+        where $t^{k + 1}$ is the current value computed using
+        \autoref{eq:iterative_amplitude} and $t^k$ is the previous value for the
+        amplitude. Choosing $\theta \in [0, 1]$ we can tune how much of the
+        previous amplitude we wish to include in the new state. This allows for
+        a more gradual transition between the iterations. We now use
+        $\tilde{t}^{k + 1}$ as our estimate of the new amplitude.
+
+\section{Results}
+
+    \begin{table}
+        \centering
+        \caption{N = 2}
+        \begin{ruledtabular}
+            \begin{tabular}{c|c|ccc}
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
+
+            \end{tabular}
+        \end{ruledtabular}
+        \label{tab:N2}        
+    \end{table}
+
+    \begin{table}
+        \centering
+        \caption{N = 6}
+        \begin{ruledtabular}
+            \begin{tabular}{c|c|ccc}
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
+
+            \end{tabular}
+        \end{ruledtabular}
+        \label{tab:N6}        
+    \end{table}
+
+    \begin{table}
+        \centering
+        \caption{Here we look at $N = 12$ particles ($R = 3$ shells).}
+        \begin{ruledtabular}
+            \begin{tabular}{c|c|ccc}
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
+                \hline
+                      & $3$ & $46.361130$ & $46.361130$ & $46.361130$ \\
+                      & $4$ & $43.663266$ & $45.837079$ & $43.309845$ \\
+                      & $5$ & $41.108850$ & $45.456882$ & $40.654709$ \\
+                      & $6$ & $40.750512$ & \nan & $40.068340$ \\
+                $0.5$ & $7$ & $40.302718$ & \nan & $39.508500$ \\
+                      & $8$ & $40.263751$ & \nan & $39.399128$ \\
+                      & $9$ & $40.216688$ & \nan & $39.329310$ \\
+                      & $10$ & $40.216251$ & \nan & $39.309409$ \\
+                      & $11$ & $40.216194$ & \nan & $39.296007$ \\
+                      & $12$ & $40.216165$ & \nan & $39.285968$ \\
+                \hline
+                      & $3$ & $73.765549$ & $73.765549$ & $73.765549$ \\
+                      & $4$ & $70.673849$ & $73.314476$ & $70.324250$ \\
+                      & $5$ & $67.569930$ & $72.990679$ & $67.031096$ \\
+                      & $6$ & $67.296869$ & \nan & $66.526677$ \\
+                $1.0$ & $7$ & $66.934744$ & \nan & $66.049564$ \\
+                      & $8$ & $66.923094$ & \nan & $65.972157$ \\
+                      & $9$ & $66.912244$ & \nan & $65.921204$ \\
+                      & $10$ & $66.912035$ & \nan & $65.889281$ \\
+                      & $11$ & $66.911364$ & \nan & $65.866715$ \\
+                      & $12$ & $66.911364$ & \nan & $65.849775$ \\
+                \hline
+                      & $3$ & $120.722260$ & $120.722260$ & $120.722260$ \\
+                      & $4$ & $117.339641$ & $120.299999$ & $116.995036$ \\
+                      & $5$ & $113.660396$ & $120.012284$ & $113.048934$ \\
+                      & $6$ & $113.484866$ & $119.872007$ & $112.658821$ \\
+                $2.0$ & $7$ & $113.247601$ & $119.662201$ & $112.309482$ \\
+                      & $8$ & $113.246578$ & $119.584733$ & $112.235521$ \\
+                      & $9$ & $113.246302$ & $119.524417$ & $112.181828$ \\
+                      & $10$ & $113.245854$ & $119.472283$ & $112.140661$ \\
+                      & $11$ & $113.245255$ & $119.430435$ & $112.109972$ \\
+                      & $12$ & $113.245183$ & $119.394712$ & $112.085683$
+            \end{tabular}
+        \end{ruledtabular}
+        \label{tab:N12}        
+    \end{table}
+
+    \begin{table}
+        \centering
+        \caption{N = 20}
+        \begin{ruledtabular}
+            \begin{tabular}{c|c|ccc}
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
+
+            \end{tabular}
+        \end{ruledtabular}
+        \label{tab:N20}        
+    \end{table}
 

### 2018-06-06 project_2/article/article.tex
Add results for N = 12 and N = 20

@@
         \centering
-        \caption{Here we look at $N = 12$ particles ($R = 3$ shells).}
+        \caption{Here we look at $N = 12$ particles ($R = 3$ shells).  We did
+        not achieve convergence using the harmonic oscillator basis for the
+        lower frequency values and large number of shells.}
         \begin{ruledtabular}
@@
                       & $11$ & $113.245255$ & $119.430435$ & $112.109972$ \\
-                      & $12$ & $113.245183$ & $119.394712$ & $112.085683$
+                      & $12$ & $113.245183$ & $119.394712$ & $112.085683$ \\
+                \hline
+                $5.0$ & $3$ & $242.334878$ & $242.334878$ & $242.334878$ \\
+                      & $4$ & $238.739590$ & $241.927593$ & $238.394598$ \\
+                      & $5$ & $234.352741$ & $241.663950$ & $233.680649$ \\
+                      & $6$ & $234.282331$ & $241.507595$ & $233.425243$ \\
+                      & $7$ & $234.194820$ & $241.390524$ & $233.226529$ \\
+                      & $8$ & $234.194058$ & $241.293416$ & $233.137932$ \\
+                      & $9$ & $234.190796$ & $241.221056$ & $233.070206$ \\
+                      & $10$ & $234.190714$ & $241.158896$ & $233.020009$ \\
+                      & $11$ & $234.190665$ & $241.109926$ & $232.980634$ \\
+                      & $12$ & $234.190552$ & $241.068091$ & $232.948528$
             \end{tabular}
@@
         \centering
-        \caption{N = 20}
+        \caption{In this table we look at $N = 20$ particles ($R = 4$ shells).
+        We did not achieve convergence using the harmonic oscillator basis for
+        the lower frequency values and large number of shells.}
         \begin{ruledtabular}
             \begin{tabular}{c|c|ccc}
-                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
-
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
+                \hline
+                $1.0$ & $4$ & $177.963297$ & $177.963297$ & $177.963297$ \\
+                      & $5$ & $168.792442$ & $177.206535$ & $168.459124$ \\
+                      & $6$ & $161.339720$ & \nan & $160.594507$ \\
+                      & $7$ & $159.958722$ & \nan & $158.841119$ \\
+                      & $8$ & $158.400172$ & \nan & $157.038330$ \\
+                      & $9$ & $158.226030$ & \nan & $156.676039$ \\
+                      & $10$ & $158.017666$ & \nan & $156.367930$ \\
+                      & $11$ & $158.010276$ & \nan & $156.292422$ \\
+                      & $12$ & $158.004951$ & \nan & $156.238258$ \\
+                \hline
+                $2.0$ & $4$ & $286.825295$ & $286.825295$ & $286.825295$ \\
+                      & $5$ & $276.898195$ & $286.159165$ & $276.381708$ \\
+                      & $6$ & $267.269712$ & $266.413122$ & $285.614965$ \\
+                      & $7$ & $266.213200$ & \nan & $264.969415$ \\
+                      & $8$ & $264.933621$ & \nan & $263.434546$ \\
+                      & $9$ & $264.874009$ & \nan & $263.215451$ \\
+                      & $10$ & $264.809953$ & \nan & $263.046184$ \\
+                      & $11$ & $264.809900$ & \nan & $262.963702$ \\
+                      & $12$ & $264.809306$ & \nan & $262.899697$ \\
+                \hline
+                $5.0$ & $4$ & $563.773951$ & $563.773951$ & $563.773951$ \\
+                      & $5$ & $552.630093$ & $563.160136$ & $552.118708$ \\
+                      & $6$ & $540.804719$ & $562.692231$ & $539.824400$ \\
+                      & $7$ & $540.227792$ & $562.306124$ & $538.886073$ \\
+                      & $8$ & $539.499326$ & $562.114280$ & $537.925127$ \\
+                      & $9$ & $539.495940$ & \nan & $537.769045$ \\
+                      & $10$ & $539.494611$ & \nan & $537.646667$ \\
+                      & $11$ & $539.493512$ & \nan & $537.548827$ \\
+                      & $12$ & $539.491764$ & \nan & $537.470616$ \\
+                \hline
+                $10.0$ & $4$ & $973.032700$ & $973.032700$ & $973.032700$ \\
+                       & $5$ & $961.371081$ & $972.439477$ & $960.862053$ \\
+                       & $6$ & $948.057077$ & $972.002302$ & $947.019789$ \\
+                       & $7$ & $947.765473$ & $971.716015$ & $946.399546$ \\
+                       & $8$ & $947.410304$ & $971.508583$ & $945.827806$ \\
+                       & $9$ & $947.409439$ & $971.332132$ & $945.663819$ \\
+                       & $10$ & $947.404929$ & $971.193592$ & $945.528489$ \\
+                       & $11$ & $947.404360$ & $971.076617$ & $945.424174$ \\
+                       & $12$ & $947.403875$ & $970.978571$ & $945.339079$ \\
             \end{tabular}

### 2018-06-06 project_2/article/article.tex
Start on discussion

@@
 
+\section{Discussion}
+
+	\subsection{Validity of results}
+        All of our results have been compared with the master thesis of M. P.
+        Lohne\cite{lohne2010coupled} for up to 10 shells. Lohne gets two sets of
+        energies, one set from RHF and one set from CCSD code with harmonic
+        oscillator basis functions. Our CCD code with Hartree-Fock basis will
+        for some configurations\footnote{By configurations we mean frequency
+        $\omega$ and number of particles $N$.} beat CCSD with plain harmonic
+        oscillator basis. For 12 shells we can compare with the results from M.
+        P. Lohne et al.\cite{lohne2011ab}, but in this article an
+        \emph{effective interaction} for the Hamiltonian with a CCSD code has
+        been used. These results are therefore significanlty better than ours,
+        but we can get a ``ball-park'' idea to benchmark against.
+
+    \subsection{Convergence trouble}
+        For a large number of particles and a low frequency the CCD-method get
+        trouble with convergence. This happens as the confining potential $v
+        \propto \omega^2$ will not be able to confine the particles when the
+        interaction gets too strong. The CCD-method will in particular get a
+        hard time keep the particles together as the excitation operator $t$
+        only excites pairs of particles leading too strong increase in the
+        interaction when the particles increase their energy level. This effect
+        can potentially be somewhat alleviated by including the singles
+        excitation, e.g., using the CCSD-method.
+
+	\subsection{Sparse implementation}
+
 \appendix

### 2018-06-06 project_2/article/article.tex

@@
             \begin{tabular}{c|c|ccc}
-                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
-
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\ \hline
+                         &  1  & 0.596333 & 0.596333 & 0.596333 \\
+                         &  2  & 0.596333 & 0.512520 & 0.512520 \\
+                         &  3  & 0.526903 & 0.505972 & 0.442235 \\
+                         &  4  & 0.526903 & 0.499216 & 0.442011 \\
+                     0.1 &  5  & 0.525666 & 0.497172 & 0.443293 \\
+                         &  6  & 0.525666 & 0.494232 & 0.443145 \\
+                         &  7  & 0.525635 & 0.493142 & 0.443056 \\
+                         &  8  & 0.525635 & 0.491895 & 0.442981 \\
+                         &  9  & 0.525635 & 0.491262 & 0.491262 \\
+                         &  10 & 0.525635 & 0.490649 & 0.442886 \\
+                         &  11 & & & \\
+                         &  12 &  & & \\ \hline
+
+                         &  1  & 1.886227 & 1.886227 & 1.886227 \\
+                         &  2  & 1.886227 & 1.786914 & 1.786914 \\
+                         &  3  & 1.799856 & 1.778903 & 1.681979 \\
+                         &  4  & 1.799856 & 1.760117 & 1.673881 \\
+                     0.5 &  5  & 1.799748 & 1.754385 & 1.670053 \\
+                         &  6  & 1.799748 & 1.748232 & 1.667804 \\
+                         &  7  & 1.799745 & 1.745231 & 1.666474 \\
+                         &  8  & 1.799745 & 1.742548 & 1.665494 \\
+                         &  9  & 1.799743 & 1.740860 & 1.6648053 \\
+                         &  10 & 1.799743 & 1.739444 & 1.664270 \\
+                         &  11 & & & \\
+                         &  12 &  & & \\ \hline
+
+                         &  1  & 3.253314 & 3.253314 & 3.253314 \\
+                         &  2  & 3.253314 & 3.152328 & 3.152328 \\
+                         &  3  & 3.162691 & 3.141827 & 3.039048 \\
+                         &  4  & 3.162691 & 3.118679 & 3.025273 \\
+                     1.0 &  5  & 3.161921 & 3.110967 & 3.017944 \\
+                         &  6  & 3.161921 & 3.103338 & 3.013923 \\
+                         &  7  & 3.161909 & 3.099324 & 3.011405 \\
+                         &  8  & 3.161909 & 3.095916 & 3.009621 \\
+                         &  9  & 3.161909 & 3.093662 & 3.008343 \\
+                         &  10 & 3.161909 & 3.091818 & 3.007357 \\
+                         &  11 & & & \\
+                         &  12 &  & & \\ \hline
+
+                         &  1  & 5.772454 & 5.772454 & 5.772454 \\
+                         &  2  & 5.772454 & 5.671234 & 5.671234 \\
+                         &  3  & 5.679048 & 5.658272 & 5.553528 \\
+                         &  4  & 5.679048 & 5.631669 & 5.534333 \\
+                     2.0 &  5  & 5.677282 & 5.622092 & 5.523490 \\
+                         &  6  & 5.677282 & 5.613118 & 5.517552 \\
+                         &  7  & 5.677206 & 5.608130 & 5.513709 \\
+                         &  8  & 5.677206 & 5.604026 & 5.511012 \\
+                         &  9  & 5.677204 & 5.601216 & 5.509050 \\
+                         &  10 & 5.677204 & 5.598946 & 5.507540 \\
+                         &  11 &  & & \\
+                         &  12 &  & & \\
             \end{tabular}

### 2018-06-06 project_2/article/article.tex
Start on explanation of results

@@
 \section{Results}
+
+    \subsection{Ground state energies for two-dimensional quantum dots}
+        Here we show the ground state energies for the two-dimensional quantum
+        dots using restriced Hartree-Fock (RHF), coupled cluster doubles using
+        both the harmonic oscillator basis and the Hartree-Fock basis which we
+        construct after running the RHF-method. We have chosen the convergence
+        criteria to be $\num{1e-6}$ in our results for the sake of comparison
+        with M. P. Lohne\cite{lohne2011ab}.
 

### 2018-06-06 project_2/article/article.tex
Add results for N = 2

--- 2018-06-06 project_2/article/article.tex
@@
 
+    \subsection{Ground state energies for two-dimensional quantum dots}
+        Here we show the ground state energies for the two-dimensional quantum
+        dots using restriced Hartree-Fock (RHF), coupled cluster doubles using
+        both the harmonic oscillator basis and the Hartree-Fock basis which we
+        construct after running the RHF-method. We have chosen the convergence
+        criteria to be $\num{1e-6}$ in our results for the sake of comparison
+        with M. P. Lohne\cite{lohne2011ab}.
+
     \begin{table}
@@
             \begin{tabular}{c|c|ccc}
-                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
-
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
+                \hline
+                           &  $1$  & $0.596333$ & $0.596333$ & $0.596333$ \\
+                           &  $2$  & $0.596333$ & $0.512520$ & $0.512520$ \\
+                           &  $3$  & $0.526903$ & $0.505972$ & $0.442235$ \\
+                           &  $4$  & $0.526903$ & $0.499216$ & $0.442011$ \\
+                           &  $5$  & $0.525666$ & $0.497172$ & $0.443293$ \\
+                           &  $6$  & $0.525666$ & $0.494232$ & $0.443145$ \\
+                     $0.1$ &  $7$  & $0.525635$ & $0.493142$ & $0.443056$ \\
+                           &  $8$  & $0.525635$ & $0.491895$ & $0.442981$ \\
+                           &  $9$  & $0.525635$ & $0.491262$ & $0.442927$ \\
+                           &  $10$ & $0.525635$ & $0.490649$ & $0.442886$ \\
+                           &  $11$ & $0.525635$ & $0.490262$ & $0.442853$ \\
+                           &  $12$ & $0.525635$ & $0.489918$ & $0.442827$ \\
+                \hline
+
+                           &  $1$  & $1.886227$ & $1.886227$ & $1.886227$ \\
+                           &  $2$  & $1.886227$ & $1.786914$ & $1.786914$ \\
+                           &  $3$  & $1.799856$ & $1.778903$ & $1.681979$ \\
+                           &  $4$  & $1.799856$ & $1.760117$ & $1.673881$ \\
+                           &  $5$  & $1.799748$ & $1.754385$ & $1.670053$ \\
+                           &  $6$  & $1.799748$ & $1.748232$ & $1.667804$ \\
+                     $0.5$ &  $7$  & $1.799745$ & $1.745231$ & $1.666474$ \\
+                           &  $8$  & $1.799745$ & $1.742548$ & $1.665494$ \\
+                           &  $9$  & $1.799743$ & $1.740860$ & $1.664805$ \\
+                           &  $10$ & $1.799743$ & $1.739444$ & $1.664270$ \\
+                           &  $11$ & $1.799742$ & $1.738416$ & $1.663856$ \\
+                           &  $12$ & $1.799742$ & $1.737562$ & $1.663522$ \\
+                \hline
+
+                           &  $1$  & $3.253314$ & $3.253314$ & $3.253314$ \\
+                           &  $2$  & $3.253314$ & $3.152328$ & $3.152328$ \\
+                           &  $3$  & $3.162691$ & $3.141827$ & $3.039048$ \\
+                           &  $4$  & $3.162691$ & $3.118679$ & $3.025273$ \\
+                           &  $5$  & $3.161921$ & $3.110967$ & $3.017944$ \\
+                           &  $6$  & $3.161921$ & $3.103338$ & $3.013923$ \\
+                     $1.0$ &  $7$  & $3.161909$ & $3.099324$ & $3.011405$ \\
+                           &  $8$  & $3.161909$ & $3.095916$ & $3.009621$ \\
+                           &  $9$  & $3.161909$ & $3.093662$ & $3.008343$ \\
+                           &  $10$ & $3.161909$ & $3.091818$ & $3.007357$ \\
+                           &  $11$ & $3.161909$ & $3.090436$ & $3.006590$ \\
+                           &  $12$ & $3.161909$ & $3.089299$ & $3.005970$ \\
+                \hline
+
+                           &  $1$  & $5.772454$ & $5.772454$ & $5.772454$ \\
+                           &  $2$  & $5.772454$ & $5.671234$ & $5.671234$ \\
+                           &  $3$  & $5.679048$ & $5.658272$ & $5.553528$ \\
+                           &  $4$  & $5.679048$ & $5.631669$ & $5.534333$ \\
+                           &  $5$  & $5.677282$ & $5.622092$ & $5.523490$ \\
+                           &  $6$  & $5.677282$ & $5.613118$ & $5.517552$ \\
+                     $2.0$ &  $7$  & $5.677206$ & $5.608130$ & $5.513709$ \\
+                           &  $8$  & $5.677206$ & $5.604026$ & $5.511012$ \\
+                           &  $9$  & $5.677204$ & $5.601216$ & $5.509050$ \\
+                           &  $10$ & $5.677204$ & $5.598946$ & $5.507540$ \\
+                           &  $11$ & $5.677204$ & $5.597213$ & $5.506356$ \\
+                           &  $12$ & $5.677204$ & $5.595791$ & $5.505399$ \\
             \end{tabular}

### 2018-06-06 project_2/article/article.tex
Add more in discussion on convergence

@@
         hard time keep the particles together as the excitation operator $t$
-        only excites pairs of particles leading too strong increase in the
+        only excites pairs of particles leading to a too strong increase in the
         interaction when the particles increase their energy level. This effect
@@
         excitation, e.g., using the CCSD-method.
+
+        We can see this behaviour in our results. For $N = 2$ we get convergence
+        for all methods for $\omega = 0.1$ shown in \autoref{tab:N2}. By
+        increasing to $N = 6$ particles we immediately see how CCD with the HO
+        basis will get in trouble for small values of $\omega$. By increasing
+        the strength of the confinment even this method does converge.
+
+        For $N = 6$ none of the methods converged when $\omega =
+        0.1$\footnote{The RHF-method would probably have converged if we had
+        started tuning its mixing parameter, but as we have focused on CCD we did
+        not think it relevant to include these results.}. By increasing the
+        frequency we restored the convergence. For larger number of shells we
+        had to start increasing the value of $\theta \to 0.9$ in order to get
+        convergence within the threshold.
 

### 2018-06-06 project_2/article/article.tex
N=6 table, 11 and 12 shells missing.

--- 2018-06-06 project_2/article/article.tex
@@
 \usepackage{dirtree, varwidth}
+% Multirow (for table)
+\usepackage{multirow}
 
@@
 			\tau^{(k + 1)} = \frac{g(u, \tau^{(k)})}{D_{ij}^{ab}},
-            \label{eq:iterative_amplitude}
 		\end{equation}
@@
 
-\section{Implementation}
-    We have costructed a flexible framework for making performing CCD
-    computations, consisting of package for \lstinline{python} that is easy
-    to install globally on any computer.
+   \section{Implmentation}
+        We have costructed a flexible framework for making performing CCD computations,
+        consisting of package for \lstinline{python} that is easy to install globally on any 
+        computer.
 
@@
             The first subsection, \mintinline{python}{matrix_elemtents},
-        contains methods for computing matrix elements in harmonic
-        oscillator basis from an analytical
-        expression\cite{anisimovas1998energy}.  Computing the matrix
-        elements is a very intensive task, and the central functions are
-        therefore implemented in C++ with a Cython interface to enable use
-        in Python.
-
-        The \mintinline{python}{hartree_fock} subsection contains methods
-        for changing from harmonic oscillator basis to Hartree-Fock basis as
-        well as an implementation of the self-consistent-field algorithm to
-        make this transaction possible.
-
-        The \mintinline{python}{schemes} subsection is arguably the most
-        important part of this project. The subsection has three different
-        classes that perform the exact same computations, but in different
-        and increasingly intelligent ways. First,
-        \mintinline{python}{CoupledClusterDoubles} is the most
+            contains methods for computing matrix elements in harmonic oscillator
+            basis from an analytical expression\cite{anisimovas1998energy}. 
+            Computing the matrix elements is a very intensive task, and the central
+            functions are therefore implemented in C++ with a Cython interface to
+            enable use in Python.
+
+            The \mintinline{python}{hartree_fock} subsection contains methods for changing
+            from harmonic oscillator basis to Hartree-Fock basis as well as an 
+            implementation of the self-consistent-field algorithm to make this
+            transaction possible.
+
+            The \mintinline{python}{schemes} subsection is arguably the most important part
+            of this project. The subsection has three different classes that perform
+            the exact same computations, but in different and increasingly intelligent
+            ways. First, \mintinline{python}{CoupledClusterDoubles} is the most 
             straightforward and naïve way to solve the CCD amplitude equations.
-        Second, because an overbearing amount of the elements in the
-        operator matrices in this problem are zero, we have implemented a
-        sparse matrix CDD solver in
+            Second, because an overbearing amount of the elements in the operator matrices
+            in this problem are zero, we have implemented a sparse matrix CDD solver in
             \mintinline{python}{CoupledClusterDoublesSparse}. Third,
-        \mintinline{python}{CoupledClusterDoublesOptimized} takes advantage
-        of sparse matrices as well, but is also parallellized and optimized
-        with memory use and number of floating point operations in mind.
-
-    \subsection{Mixing}
-        A way to try ``massage'' convergence out of the CCD-method is to use
-        \emph{mixing}. A popular solution is Pulay mixing\footnote{Commonly
-        known as DIIS (direct inversion in the iterative subspace).}. We did not
-        implement Pulay mixing but used a more naïve and simpler solution.
-        \begin{align}
-            \tilde{t}^{k + 1} = \theta t^{k + 1} + (1 - \theta)t^k,
-        \end{align}
-        where $t^{k + 1}$ is the current value computed using
-        \autoref{eq:iterative_amplitude} and $t^k$ is the previous value for the
-        amplitude. Choosing $\theta \in [0, 1]$ we can tune how much of the
-        previous amplitude we wish to include in the new state. This allows for
-        a more gradual transition between the iterations. We now use
-        $\tilde{t}^{k + 1}$ as our estimate of the new amplitude.
+            \mintinline{python}{CoupledClusterDoublesOptimized} takes advantage of 
+            sparse matrices as well, but is also parallellized and optimized with
+            memory use and number of floating point operations in mind.
 
 \section{Results}
-
-    \subsection{Ground state energies for two-dimensional quantum dots}
-        Here we show the ground state energies for the two-dimensional quantum
-        dots using restriced Hartree-Fock (RHF), coupled cluster doubles using
-        both the harmonic oscillator basis and the Hartree-Fock basis which we
-        construct after running the RHF-method. We have chosen the convergence
-        criteria to be $\num{1e-6}$ in our results for the sake of comparison
-        with M. P. Lohne\cite{lohne2011ab}.
 
@@
                 $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\ \hline
-                         &  1  & 0.596333 & 0.596333 & 0.596333 \\
-                         &  2  & 0.596333 & 0.512520 & 0.512520 \\
-                         &  3  & 0.526903 & 0.505972 & 0.442235 \\
-                         &  4  & 0.526903 & 0.499216 & 0.442011 \\
-                     0.1 &  5  & 0.525666 & 0.497172 & 0.443293 \\
-                         &  6  & 0.525666 & 0.494232 & 0.443145 \\
-                         &  7  & 0.525635 & 0.493142 & 0.443056 \\
-                         &  8  & 0.525635 & 0.491895 & 0.442981 \\
-                         &  9  & 0.525635 & 0.491262 & 0.491262 \\
-                         &  10 & 0.525635 & 0.490649 & 0.442886 \\
-                         &  11 & & & \\
-                         &  12 &  & & \\ \hline
-
-                         &  1  & 1.886227 & 1.886227 & 1.886227 \\
-                         &  2  & 1.886227 & 1.786914 & 1.786914 \\
-                         &  3  & 1.799856 & 1.778903 & 1.681979 \\
-                         &  4  & 1.799856 & 1.760117 & 1.673881 \\
-                     0.5 &  5  & 1.799748 & 1.754385 & 1.670053 \\
-                         &  6  & 1.799748 & 1.748232 & 1.667804 \\
-                         &  7  & 1.799745 & 1.745231 & 1.666474 \\
-                         &  8  & 1.799745 & 1.742548 & 1.665494 \\
-                         &  9  & 1.799743 & 1.740860 & 1.6648053 \\
-                         &  10 & 1.799743 & 1.739444 & 1.664270 \\
-                         &  11 & & & \\
-                         &  12 &  & & \\ \hline
-
-                         &  1  & 3.253314 & 3.253314 & 3.253314 \\
-                         &  2  & 3.253314 & 3.152328 & 3.152328 \\
-                         &  3  & 3.162691 & 3.141827 & 3.039048 \\
-                         &  4  & 3.162691 & 3.118679 & 3.025273 \\
-                     1.0 &  5  & 3.161921 & 3.110967 & 3.017944 \\
-                         &  6  & 3.161921 & 3.103338 & 3.013923 \\
-                         &  7  & 3.161909 & 3.099324 & 3.011405 \\
-                         &  8  & 3.161909 & 3.095916 & 3.009621 \\
-                         &  9  & 3.161909 & 3.093662 & 3.008343 \\
-                         &  10 & 3.161909 & 3.091818 & 3.007357 \\
-                         &  11 & & & \\
-                         &  12 &  & & \\ \hline
-
-                         &  1  & 5.772454 & 5.772454 & 5.772454 \\
-                         &  2  & 5.772454 & 5.671234 & 5.671234 \\
-                         &  3  & 5.679048 & 5.658272 & 5.553528 \\
-                         &  4  & 5.679048 & 5.631669 & 5.534333 \\
-                     2.0 &  5  & 5.677282 & 5.622092 & 5.523490 \\
-                         &  6  & 5.677282 & 5.613118 & 5.517552 \\
-                         &  7  & 5.677206 & 5.608130 & 5.513709 \\
-                         &  8  & 5.677206 & 5.604026 & 5.511012 \\
-                         &  9  & 5.677204 & 5.601216 & 5.509050 \\
-                         &  10 & 5.677204 & 5.598946 & 5.507540 \\
-                         &  11 &  & & \\
-                         &  12 &  & & \\
+                         &  $1$  & $0.596333$ & $0.596333$ & $0.596333$ \\
+                         &  $2$  & $0.596333$ & $0.512520$ & $0.512520$ \\
+                         &  $3$  & $0.526903$ & $0.505972$ & $0.442235$ \\
+                         &  $4$  & $0.526903$ & $0.499216$ & $0.442011$ \\
+                         &  $5$  & $0.525666$ & $0.497172$ & $0.443293$ \\
+  \multirow{2}{*}{$0.1$} &  $6$  & $0.525666$ & $0.494232$ & $0.443145$ \\
+                         &  $7$  & $0.525635$ & $0.493142$ & $0.443056$ \\
+                         &  $8$  & $0.525635$ & $0.491895$ & $0.442981$ \\
+                         &  $9$  & $0.525635$ & $0.491262$ & $0.491262$ \\
+                         &  $10$ & $0.525635$ & $0.490649$ & $0.442886$ \\
+                         &  $11$ & & & \\
+                         &  $12$ & & & \\ \hline
+
+                         &  $1$  & $1.886227$ & $1.886227$ & $1.886227$ \\
+                         &  $2$  & $1.886227$ & $1.786914$ & $1.786914$ \\
+                         &  $3$  & $1.799856$ & $1.778903$ & $1.681979$ \\
+                         &  $4$  & $1.799856$ & $1.760117$ & $1.673881$ \\
+                         &  $5$  & $1.799748$ & $1.754385$ & $1.670053$ \\
+  \multirow{2}{*}{$0.5$} &  $6$  & $1.799748$ & $1.748232$ & $1.667804$ \\
+                         &  $7$  & $1.799745$ & $1.745231$ & $1.666474$ \\
+                         &  $8$  & $1.799745$ & $1.742548$ & $1.665494$ \\
+                         &  $9$  & $1.799743$ & $1.740860$ & $1.664805$ \\
+                         &  $10$ & $1.799743$ & $1.739444$ & $1.664270$ \\
+                         &  $11$ & & & \\
+                         &  $12$ & & & \\ \hline
+
+                         &  $1$  & $3.253314$ & $3.253314$ & $3.253314$ \\
+                         &  $2$  & $3.253314$ & $3.152328$ & $3.152328$ \\
+                         &  $3$  & $3.162691$ & $3.141827$ & $3.039048$ \\
+                         &  $4$  & $3.162691$ & $3.118679$ & $3.025273$ \\
+                         &  $5$  & $3.161921$ & $3.110967$ & $3.017944$ \\
+  \multirow{2}{*}{$1.0$} &  $6$  & $3.161921$ & $3.103338$ & $3.013923$ \\
+                         &  $7$  & $3.161909$ & $3.099324$ & $3.011405$ \\
+                         &  $8$  & $3.161909$ & $3.095916$ & $3.009621$ \\
+                         &  $9$  & $3.161909$ & $3.093662$ & $3.008343$ \\
+                         &  $10$ & $3.161909$ & $3.091818$ & $3.007357$ \\
+                         &  $11$ & & & \\
+                         &  $12$ & & & \\ \hline
+
+                         &  $1$  & $5.772454$ & $5.772454$ & $5.772454$ \\
+                         &  $2$  & $5.772454$ & $5.671234$ & $5.671234$ \\
+                         &  $3$  & $5.679048$ & $5.658272$ & $5.553528$ \\
+                         &  $4$  & $5.679048$ & $5.631669$ & $5.534333$ \\
+                         &  $5$  & $5.677282$ & $5.622092$ & $5.523490$ \\
+  \multirow{2}{*}{$2.0$} &  $6$  & $5.677282$ & $5.613118$ & $5.517552$ \\
+                         &  $7$  & $5.677206$ & $5.608130$ & $5.513709$ \\
+                         &  $8$  & $5.677206$ & $5.604026$ & $5.511012$ \\
+                         &  $9$  & $5.677204$ & $5.601216$ & $5.509050$ \\
+                         &  $10$ & $5.677204$ & $5.598946$ & $5.507540$ \\
+                         &  $11$ & & & \\
+                         &  $12$ & & & \\
             \end{tabular}
@@
             \begin{tabular}{c|c|ccc}
-                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\ \hline
+                         &  $1$  & \nan & \nan & \nan \\
+                         &  $2$  & $4.864244$ & $4.864244$ & $4.864244$ \\
+                         &  $3$  & $4.435740$ & \nan & $4.319901$ \\
+                         &  $4$  & $4.019787$ & \nan & $3.829962$ \\
+                         &  $5$  & $3.963149$ & \nan & $3.666722$ \\
+  \multirow{2}{*}{$0.1$} &  $6$  & $3.870617$ & \nan & $3.597876$ \\
+                         &  $7$  & $3.863135$ & \nan & $3.590388$ \\
+                         &  $8$  & $3.852880$ & \nan & $3.587711$ \\
+                         &  $9$  & $3.852591$ & \nan & $3.587291$ \\
+                         &  $10$ & $3.852393$ & \nan & $3.587136$ \\
+                         &  $11$ & & & \\
+                         &  $12$ & & & \\ \hline
+
+                         &  $1$  & \nan & \nan & \nan \\
+                         &  $2$  & $13.640713$ & $13.640713$ & $13.640713$ \\
+                         &  $3$  & $13.051620$ & $13.385987$ & $12.901520$ \\
+                         &  $4$  & $12.357471$ & $13.261097$ & $12.057345$ \\
+                         &  $5$  & $12.325128$ & $13.138572$ & $11.934988$ \\
+  \multirow{2}{*}{$0.5$} &  $6$  & $12.271499$ & $13.084158$ & $11.864098$ \\
+                         &  $7$  & $12.271375$ & $13.068399$ & $11.849762$ \\
+                         &  $8$  & $12.271361$ & $13.055561$ & $11.841330$ \\
+                         &  $9$  & $12.271337$ & $13.045386$ & $11.835470$ \\
+                         &  $10$ & $12.271326$ & $13.037878$ & $11.831351$ \\
+                         &  $11$ & & & \\
+                         &  $12$ & & & \\ \hline
+
+                         &  $1$  & \nan & \nan & \nan \\
+                         &  $2$  & $22.219813$ & $22.219813$ & $22.219813$ \\
+                         &  $3$  & $21.593198$ & $21.974675$ & $21.423811$ \\
+                         &  $4$  & $20.766919$ & $21.854191$ & $20.429265$ \\
+                         &  $5$  & $20.748402$ & $21.793624$ & $20.332454$ \\
+  \multirow{2}{*}{$1.0$} &  $6$  & $20.720257$ & $21.750091$ & $20.274013$ \\
+                         &  $7$  & $20.720132$ & $21.718843$ & $20.249849$ \\
+                         &  $8$  & $20.719248$ & $21.695224$ & $20.234705$ \\
+                         &  $9$  & $20.719248$ & $21.675931$ & $20.224389$ \\
+                         &  $10$ & $20.719217$ & $21.661830$ & $20.21707530$ \\
+                         &  $11$ & & & \\
+                         &  $12$ & & & \\ \hline
+
+                         &  $1$  & \nan & \nan & \nan \\
+                         &  $2$  & $37.281425$ & $37.281425$ & $37.281425$ \\
+                         &  $3$  & $36.637217$ & $37.042127$ & $36.450634$ \\
+                         &  $4$  & $35.689555$ & $36.925664$ & $35.328432$ \\
+                         &  $5$  & $35.681729$ & $36.864367$ & $35.250185$ \\
+  \multirow{2}{*}{$2.0$} &  $6$  & $35.672333$ & $36.812895$ & $35.200308$ \\
+                         &  $7$  & $35.671851$ & $36.775986$ & $35.168245$ \\
+                         &  $8$  & $35.670358$ & $36.747864$ & $35.147097$ \\
+                         &  $9$  & $35.670333$ & $36.725261$ & $35.131953$ \\
+                         &  $10$ & $35.670144$ & $36.708362$ & $35.121033$ \\
+                         &  $11$ & & & \\
+                         &  $12$ & & & \\
 
@@
         \centering
-        \caption{Here we look at $N = 12$ particles ($R = 3$ shells).  We did
-        not achieve convergence using the harmonic oscillator basis for the
-        lower frequency values and large number of shells.}
+        \caption{Here we look at $N = 12$ particles ($R = 3$ shells).}
         \begin{ruledtabular}
@@
                       & $11$ & $113.245255$ & $119.430435$ & $112.109972$ \\
-                      & $12$ & $113.245183$ & $119.394712$ & $112.085683$ \\
-                \hline
-                $5.0$ & $3$ & $242.334878$ & $242.334878$ & $242.334878$ \\
-                      & $4$ & $238.739590$ & $241.927593$ & $238.394598$ \\
-                      & $5$ & $234.352741$ & $241.663950$ & $233.680649$ \\
-                      & $6$ & $234.282331$ & $241.507595$ & $233.425243$ \\
-                      & $7$ & $234.194820$ & $241.390524$ & $233.226529$ \\
-                      & $8$ & $234.194058$ & $241.293416$ & $233.137932$ \\
-                      & $9$ & $234.190796$ & $241.221056$ & $233.070206$ \\
-                      & $10$ & $234.190714$ & $241.158896$ & $233.020009$ \\
-                      & $11$ & $234.190665$ & $241.109926$ & $232.980634$ \\
-                      & $12$ & $234.190552$ & $241.068091$ & $232.948528$
+                      & $12$ & $113.245183$ & $119.394712$ & $112.085683$
             \end{tabular}
@@
         \centering
-        \caption{In this table we look at $N = 20$ particles ($R = 4$ shells).
-        We did not achieve convergence using the harmonic oscillator basis for
-        the lower frequency values and large number of shells.}
+        \caption{N = 20}
         \begin{ruledtabular}
             \begin{tabular}{c|c|ccc}
-                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
-                \hline
-                $1.0$ & $4$ & $177.963297$ & $177.963297$ & $177.963297$ \\
-                      & $5$ & $168.792442$ & $177.206535$ & $168.459124$ \\
-                      & $6$ & $161.339720$ & \nan & $160.594507$ \\
-                      & $7$ & $159.958722$ & \nan & $158.841119$ \\
-                      & $8$ & $158.400172$ & \nan & $157.038330$ \\
-                      & $9$ & $158.226030$ & \nan & $156.676039$ \\
-                      & $10$ & $158.017666$ & \nan & $156.367930$ \\
-                      & $11$ & $158.010276$ & \nan & $156.292422$ \\
-                      & $12$ & $158.004951$ & \nan & $156.238258$ \\
-                \hline
-                $2.0$ & $4$ & $286.825295$ & $286.825295$ & $286.825295$ \\
-                      & $5$ & $276.898195$ & $286.159165$ & $276.381708$ \\
-                      & $6$ & $267.269712$ & $266.413122$ & $285.614965$ \\
-                      & $7$ & $266.213200$ & \nan & $264.969415$ \\
-                      & $8$ & $264.933621$ & \nan & $263.434546$ \\
-                      & $9$ & $264.874009$ & \nan & $263.215451$ \\
-                      & $10$ & $264.809953$ & \nan & $263.046184$ \\
-                      & $11$ & $264.809900$ & \nan & $262.963702$ \\
-                      & $12$ & $264.809306$ & \nan & $262.899697$ \\
-                \hline
-                $5.0$ & $4$ & $563.773951$ & $563.773951$ & $563.773951$ \\
-                      & $5$ & $552.630093$ & $563.160136$ & $552.118708$ \\
-                      & $6$ & $540.804719$ & $562.692231$ & $539.824400$ \\
-                      & $7$ & $540.227792$ & $562.306124$ & $538.886073$ \\
-                      & $8$ & $539.499326$ & $562.114280$ & $537.925127$ \\
-                      & $9$ & $539.495940$ & \nan & $537.769045$ \\
-                      & $10$ & $539.494611$ & \nan & $537.646667$ \\
-                      & $11$ & $539.493512$ & \nan & $537.548827$ \\
-                      & $12$ & $539.491764$ & \nan & $537.470616$ \\
-                \hline
-                $10.0$ & $4$ & $973.032700$ & $973.032700$ & $973.032700$ \\
-                       & $5$ & $961.371081$ & $972.439477$ & $960.862053$ \\
-                       & $6$ & $948.057077$ & $972.002302$ & $947.019789$ \\
-                       & $7$ & $947.765473$ & $971.716015$ & $946.399546$ \\
-                       & $8$ & $947.410304$ & $971.508583$ & $945.827806$ \\
-                       & $9$ & $947.409439$ & $971.332132$ & $945.663819$ \\
-                       & $10$ & $947.404929$ & $971.193592$ & $945.528489$ \\
-                       & $11$ & $947.404360$ & $971.076617$ & $945.424174$ \\
-                       & $12$ & $947.403875$ & $970.978571$ & $945.339079$ \\
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF)
+
             \end{tabular}
@@
     \end{table}
-
-\section{Discussion}
-
-    \subsection{Validity of results}
-        All of our results have been compared with the master thesis of M. P.
-        Lohne\cite{lohne2010coupled} for up to 10 shells. Lohne gets two sets of
-        energies, one set from RHF and one set from CCSD code with harmonic
-        oscillator basis functions. Our CCD code with Hartree-Fock basis will
-        for some configurations\footnote{By configurations we mean frequency
-        $\omega$ and number of particles $N$.} beat CCSD with plain harmonic
-        oscillator basis. For 12 shells we can compare with the results from M.
-        P. Lohne et al.\cite{lohne2011ab}, but in this article an
-        \emph{effective interaction} for the Hamiltonian with a CCSD code has
-        been used. These results are therefore significanlty better than ours,
-        but we can get a ``ball-park'' idea to benchmark against.
-
-    \subsection{Convergence trouble}
-        For a large number of particles and a low frequency the CCD-method get
-        trouble with convergence. This happens as the confining potential $v
-        \propto \omega^2$ will not be able to confine the particles when the
-        interaction gets too strong. The CCD-method will in particular get a
-        hard time keep the particles together as the excitation operator $t$
-        only excites pairs of particles leading too strong increase in the
-        interaction when the particles increase their energy level. This effect
-        can potentially be somewhat alleviated by including the singles
-        excitation, e.g., using the CCSD-method.
-
-    \subsection{Sparse implementation}
 

### 2018-06-06 project_2/coupled-cluster/coupled_cluster/schemes/cc_interface.pyx

@@ -1,5 +1,5 @@
 import cython
-from cython.parallel import prange
+#from cython.parallel import prange
 
 import numpy as np
 cimport numpy as np
@@ -15,7 +15,7 @@ def amplitude_scaling_two_body(
     cdef int a, b, i, j
     cdef double divisor, val
 
-    for a in prange(m, nogil=True):
+    for a in range(m):
         for b in range(a, m):
             for i in range(n):
                 for j in range(i, n):
@@ -50,7 +50,7 @@ def amplitude_scaling_two_body_sparse(
 
     length = len(data)
 
-    for index in prange(length, nogil=True):
+    for index in range(length):
         a = a_arr[index]
         b = b_arr[index]
         i = i_arr[index]

### 2018-06-06 project_2/scripts/comparison.py

@@ -20,12 +20,12 @@ import os
 file_path = os.path.join("..", "dat")
 filename = os.path.join(file_path, "coulomb_{0}.pkl")
 
-num_shells = 10
+num_shells = 1
 generate_index_map(num_shells)
 
-omega = 1.0
+omega = 2.0
 l = IndexMap.shell_arr[-1]
-n = 12
+n = 6
 theta = 0.3
 
 filename = filename.format(l)
@@ -46,7 +46,7 @@ t1 = time.time()
 print ("Time spent creating Coulomb elements: {0} sec".format(t1 - t0))
 
 t0 = time.time()
-c, energy = scf_rhf(h.todense(), u, np.eye(l//2), n//2)
+c, energy = scf_rhf(h.todense(), u, np.eye(l//2), n//2, tol=1e-6)
 t1 = time.time()
 print ("Time spent in SCF RHF: {0} sec".format(t1 - t0))
 
@@ -75,7 +75,6 @@ print ("Time spent antisymmetrizing two body elements: {0} sec".format(t1 - t0))
 #print ("Time spent computing CCD energy with HF basis: {0} sec".format(t1 - t0))
 #print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}".format(energy, iterations, (t1 - t0)/iterations))
 
-"""
 t0 = time.time()
 ccd_hf = CoupledClusterDoublesOptimized(
         _h.todense(), _u.todense(), n, parallel=False)
@@ -84,15 +83,14 @@ print ("Time spent setting up CCD (opt, parallel) code with HF basis: {0} sec"
     .format(t1 - t0))
 
 t0 = time.time()
-energy, iterations = ccd_hf.compute_energy(tol=1e-4, theta=theta)
+energy, iterations = ccd_hf.compute_energy(tol=1e-6, theta=theta)
 t1 = time.time()
 print ("Time spent computing CCD (opt, parallel) energy with HF basis: {0} sec"
    .format(t1 - t0))
 print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}"
     .format(energy, iterations, (t1 - t0)/iterations))
-"""
 
-__import__("sys").exit()
+#__import__("sys").exit()
 
 __h = omega * get_one_body_elements_spin(l)
 t0 = time.time()
@@ -101,11 +99,11 @@ t1 = time.time()
[…]

### 2018-06-06 project_2/scripts/comparison.py
Add new run and format output

@@ -20,13 +20,15 @@ import os
 file_path = os.path.join("..", "dat")
 filename = os.path.join(file_path, "coulomb_{0}.pkl")
 
-num_shells = 10
+num_shells = 12
 generate_index_map(num_shells)
 
-omega = 1.0
+omega = 2.0
 l = IndexMap.shell_arr[-1]
-n = 12
-theta = 0.3
+n = 2
+theta_hf = 0.3
+theta_ho = 0.3
+max_iterations = 1000
 
 filename = filename.format(l)
 
@@ -35,9 +37,10 @@ w = {0},
 num_shells = {1},
 l = {2},
 n = {3},
-theta = {4},
-filename = {5}
-""".format(omega, num_shells, l, n, theta, filename))
+theta_hf = {4},
+theta_ho = {5},
+filename = {6}
+""".format(omega, num_shells, l, n, theta_hf, theta_ho, filename))
 
 h = omega * get_one_body_elements(l)
 t0 = time.time()
@@ -46,11 +49,11 @@ t1 = time.time()
 print ("Time spent creating Coulomb elements: {0} sec".format(t1 - t0))
 
 t0 = time.time()
-c, energy = scf_rhf(h.todense(), u, np.eye(l//2), n//2)
+c, energy = scf_rhf(h.todense(), u, np.eye(l//2), n//2, tol=1e-6)
 t1 = time.time()
 print ("Time spent in SCF RHF: {0} sec".format(t1 - t0))
 
-print ("\tRHF Energy: {0}".format(energy))
+print ("\tRHF Energy: {0:.6f}".format(energy))
 
 hi = transform_one_body_elements(h, c)
 t0 = time.time()
@@ -75,7 +78,6 @@ print ("Time spent antisymmetrizing two body elements: {0} sec".format(t1 - t0))
 #print ("Time spent computing CCD energy with HF basis: {0} sec".format(t1 - t0))
 #print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}".format(energy, iterations, (t1 - t0)/iterations))
 
-"""
 t0 = time.time()
 ccd_hf = CoupledClusterDoublesOptimized(
         _h.todense(), _u.todense(), n, parallel=False)
@@ -84,15 +86,13 @@ print ("Time spent setting up CCD (opt, parallel) code with HF basis: {0} sec"
     .format(t1 - t0))
 
 t0 = time.time()
[…]

### 2018-06-06 project_2/article/article.tex
Recompilation

@@
             \tau^{(k + 1)} = \frac{g(u, \tau^{(k)})}{D_{ij}^{ab}},
+            \label{eq:iterative_amplitude}
         \end{equation}
@@
         The first subsection, \mintinline{python}{matrix_elemtents},
-            contains methods for computing matrix elements in harmonic oscillator
-            basis from an analytical expression\cite{anisimovas1998energy}. 
-            Computing the matrix elements is a very intensive task, and the central
-            functions are therefore implemented in C++ with a Cython interface to
-            enable use in Python.
-
-            The \mintinline{python}{hartree_fock} subsection contains methods for changing
-            from harmonic oscillator basis to Hartree-Fock basis as well as an 
-            implementation of the self-consistent-field algorithm to make this
-            transaction possible.
-
-            The \mintinline{python}{schemes} subsection is arguably the most important part
-            of this project. The subsection has three different classes that perform
-            the exact same computations, but in different and increasingly intelligent
-            ways. First, \mintinline{python}{CoupledClusterDoubles} is the most 
+        contains methods for computing matrix elements in harmonic
+        oscillator basis from an analytical
+        expression\cite{anisimovas1998energy}.  Computing the matrix
+        elements is a very intensive task, and the central functions are
+        therefore implemented in C++ with a Cython interface to enable use
+        in Python.
+
+        The \mintinline{python}{hartree_fock} subsection contains methods
+        for changing from harmonic oscillator basis to Hartree-Fock basis as
+        well as an implementation of the self-consistent-field algorithm to
+        make this transaction possible.
+
+        The \mintinline{python}{schemes} subsection is arguably the most
+        important part of this project. The subsection has three different
+        classes that perform the exact same computations, but in different
+        and increasingly intelligent ways. First,
+        \mintinline{python}{CoupledClusterDoubles} is the most
         straightforward and naïve way to solve the CCD amplitude equations.
-            Second, because an overbearing amount of the elements in the operator matrices
-            in this problem are zero, we have implemented a sparse matrix CDD solver in
+        Second, because an overbearing amount of the elements in the
+        operator matrices in this problem are zero, we have implemented a
+        sparse matrix CDD solver in
         \mintinline{python}{CoupledClusterDoublesSparse}. Third,
-            \mintinline{python}{CoupledClusterDoublesOptimized} takes advantage of 
-            sparse matrices as well, but is also parallellized and optimized with
-            memory use and number of floating point operations in mind.
+        \mintinline{python}{CoupledClusterDoublesOptimized} takes advantage
+        of sparse matrices as well, but is also parallellized and optimized
+        with memory use and number of floating point operations in mind.
+
+    \subsection{Mixing}
+        A way to try ``massage'' convergence out of the CCD-method is to use
+        \emph{mixing}. A popular solution is Pulay mixing\footnote{Commonly
+        known as DIIS (direct inversion in the iterative subspace).}. We did not
+        implement Pulay mixing but used a more naïve and simpler solution.
+        \begin{align}
+            \tilde{t}^{k + 1} = \theta t^{k + 1} + (1 - \theta)t^k,
+        \end{align}
+        where $t^{k + 1}$ is the current value computed using
+        \autoref{eq:iterative_amplitude} and $t^k$ is the previous value for the
+        amplitude. Choosing $\theta \in [0, 1]$ we can tune how much of the
+        previous amplitude we wish to include in the new state. This allows for
+        a more gradual transition between the iterations. We now use
+        $\tilde{t}^{k + 1}$ as our estimate of the new amplitude.
 
@@
     \end{table}
+
+\section{Discussion}
+
+    \subsection{Validity of results}
+        All of our results have been compared with the master thesis of M. P.
+        Lohne\cite{lohne2010coupled} for up to 10 shells. Lohne gets two sets of
+        energies, one set from RHF and one set from CCSD code with harmonic
+        oscillator basis functions. Our CCD code with Hartree-Fock basis will
+        for some configurations\footnote{By configurations we mean frequency
+        $\omega$ and number of particles $N$.} beat CCSD with plain harmonic
+        oscillator basis. For 12 shells we can compare with the results from M.
+        P. Lohne et al.\cite{lohne2011ab}, but in this article an
+        \emph{effective interaction} for the Hamiltonian with a CCSD code has
+        been used. These results are therefore significanlty better than ours,
+        but we can get a ``ball-park'' idea to benchmark against.
+
+    \subsection{Convergence trouble}
+        For a large number of particles and a low frequency the CCD-method get
+        trouble with convergence. This happens as the confining potential $v
+        \propto \omega^2$ will not be able to confine the particles when the
+        interaction gets too strong. The CCD-method will in particular get a
+        hard time keep the particles together as the excitation operator $t$
+        only excites pairs of particles leading too strong increase in the
+        interaction when the particles increase their energy level. This effect
+        can potentially be somewhat alleviated by including the singles
+        excitation, e.g., using the CCSD-method.
+
+    \subsection{Sparse implementation}
 

### 2018-06-06 project_2/article/article.tex
Fix merge conflict

@@
             \begin{tabular}{c|c|ccc}
-                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\ \hline
+                $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
+                \hline
                            &  $1$  & $0.596333$ & $0.596333$ & $0.596333$ \\
@@
                            &  $8$  & $0.525635$ & $0.491895$ & $0.442981$ \\
-                         &  $9$  & $0.525635$ & $0.491262$ & $0.491262$ \\
+                           &  $9$  & $0.525635$ & $0.491262$ & $0.442927$ \\
                            &  $10$ & $0.525635$ & $0.490649$ & $0.442886$ \\
-                         &  $11$ & & & \\
-                         &  $12$ & & & \\ \hline
+                           &  $11$ & $0.525635$ & $0.490262$ & $0.442853$ \\
+                           &  $12$ & $0.525635$ & $0.489918$ & $0.442827$ \\
+                \hline
 
@@
                            &  $10$ & $1.799743$ & $1.739444$ & $1.664270$ \\
-                         &  $11$ & & & \\
-                         &  $12$ & & & \\ \hline
+                           &  $11$ & $1.799742$ & $1.738416$ & $1.663856$ \\
+                           &  $12$ & $1.799742$ & $1.737562$ & $1.663522$ \\
+                \hline
 
@@
                            &  $10$ & $3.161909$ & $3.091818$ & $3.007357$ \\
-                         &  $11$ & & & \\
-                         &  $12$ & & & \\ \hline
+                           &  $11$ & $3.161909$ & $3.090436$ & $3.006590$ \\
+                           &  $12$ & $3.161909$ & $3.089299$ & $3.005970$ \\
+                \hline
 
@@
                            &  $10$ & $5.677204$ & $5.598946$ & $5.507540$ \\
-                         &  $11$ & & & \\
-                         &  $12$ & & & \\
+                           &  $11$ & $5.677204$ & $5.597213$ & $5.506356$ \\
+                           &  $12$ & $5.677204$ & $5.595791$ & $5.505399$ \\
             \end{tabular}
@@
                 $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\ \hline
-                         &  $1$  & \nan & \nan & \nan \\
                          &  $2$  & $4.864244$ & $4.864244$ & $4.864244$ \\
@@
 
-                         &  $1$  & \nan & \nan & \nan \\
                          &  $2$  & $13.640713$ & $13.640713$ & $13.640713$ \\
@@
 
-                         &  $1$  & \nan & \nan & \nan \\
                          &  $2$  & $22.219813$ & $22.219813$ & $22.219813$ \\
@@
 
-                         &  $1$  & \nan & \nan & \nan \\
                          &  $2$  & $37.281425$ & $37.281425$ & $37.281425$ \\

### 2018-06-06 project_2/article/article.tex
Start fixing truncated decimals

@@
                            &  $11$ & $5.677204$ & $5.597213$ & $5.506356$ \\
-                           &  $12$ & $5.677204$ & $5.595791$ & $5.505399$ \\
+                           &  $12$ & $5.677204$ & $5.595791$ & $5.505399$
             \end{tabular}
@@
                          &  $2$  & $4.864244$ & $4.864244$ & $4.864244$ \\
-                         &  $3$  & $4.435740$ & \nan & $4.319901$ \\
-                         &  $4$  & $4.019787$ & \nan & $3.829962$ \\
+                         &  $3$  & $4.435740$ & $4.446235$ & $4.319901$ \\
+                         &  $4$  & $4.019787$ & $4.383692$ & $3.829962$ \\
                          &  $5$  & $3.963149$ & \nan & $3.666722$ \\
-  \multirow{2}{*}{$0.1$} &  $6$  & $3.870617$ & \nan & $3.597876$ \\
-                         &  $7$  & $3.863135$ & \nan & $3.590388$ \\
+                         &  $6$  & $3.870617$ & \nan & $3.597876$ \\
+                  $0.1$  &  $7$  & $3.863135$ & \nan & $3.590388$ \\
                          &  $8$  & $3.852880$ & \nan & $3.587711$ \\
@@
                          &  $10$ & $3.852393$ & \nan & $3.587136$ \\
-                         &  $11$ & & & \\
-                         &  $12$ & & & \\ \hline
+                         &  $11$ & $3.852391$ & \nan & $3.586821$ \\
+                         &  $12$ & $3.852382$ & \nan & $3.586582$ \\ \hline
 
@@
                          &  $5$  & $12.325128$ & $13.138572$ & $11.934988$ \\
-  \multirow{2}{*}{$0.5$} &  $6$  & $12.271499$ & $13.084158$ & $11.864098$ \\
-                         &  $7$  & $12.271375$ & $13.068399$ & $11.849762$ \\
+                         &  $6$  & $12.271499$ & $13.084158$ & $11.864098$ \\
+                  $0.5$  &  $7$  & $12.271375$ & $13.068399$ & $11.849762$ \\
                          &  $8$  & $12.271361$ & $13.055561$ & $11.841330$ \\
@@
                          &  $10$ & $12.271326$ & $13.037878$ & $11.831351$ \\
-                         &  $11$ & & & \\
-                         &  $12$ & & & \\ \hline
+                         &  $11$ & $12.271324$ & \nan & $11.828242$ \\
+                         &  $12$ & $12.271320$ & \nan & $11.825835$ \\ \hline
 
@@
                          &  $5$  & $20.748402$ & $21.793624$ & $20.332454$ \\
-  \multirow{2}{*}{$1.0$} &  $6$  & $20.720257$ & $21.750091$ & $20.274013$ \\
-                         &  $7$  & $20.720132$ & $21.718843$ & $20.249849$ \\
+                         &  $6$  & $20.720257$ & $21.750091$ & $20.274013$ \\
+                  $1.0$  &  $7$  & $20.720132$ & $21.718843$ & $20.249849$ \\
                          &  $8$  & $20.719248$ & $21.695224$ & $20.234705$ \\
                          &  $9$  & $20.719248$ & $21.675931$ & $20.224389$ \\
-                         &  $10$ & $20.719217$ & $21.661830$ & $20.21707530$ \\
-                         &  $11$ & & & \\
-                         &  $12$ & & & \\ \hline
+                         &  $10$ & $20.719217$ & $21.661830$ & $20.217075$ \\
+                         &  $11$ & $20.719215$ & $21.649812$ & $20.211541$ \\
+                         &  $12$ & $20.719215$ & $21.640765$ & $20.207258$ \\
+                         \hline
 
@@
                          &  $5$  & $35.681729$ & $36.864367$ & $35.250185$ \\
-  \multirow{2}{*}{$2.0$} &  $6$  & $35.672333$ & $36.812895$ & $35.200308$ \\
-                         &  $7$  & $35.671851$ & $36.775986$ & $35.168245$ \\
+                         &  $6$  & $35.672333$ & $36.812895$ & $35.200308$ \\
+                  $2.0$  &  $7$  & $35.671851$ & $36.775986$ & $35.168245$ \\
                          &  $8$  & $35.670358$ & $36.747864$ & $35.147097$ \\
@@
                          &  $10$ & $35.670144$ & $36.708362$ & $35.121033$ \\
-                         &  $11$ & & & \\
-                         &  $12$ & & & \\
-
+                         &  $11$ & $35.670143$ & $36.694188$ & $35.112680$ \\
+                         &  $12$ & $35.670127$ & $36.683281$ & $35.106188$
             \end{tabular}
@@
                       & $3$ & $46.361130$ & $46.361130$ & $46.361130$ \\
-                      & $4$ & $43.663266$ & $45.837079$ & $43.309845$ \\
-                      & $5$ & $41.108850$ & $45.456882$ & $40.654709$ \\
+                      & $4$ & $43.663267$ & $45.837079$ & $43.309845$ \\
+                      & $5$ & $41.108851$ & $45.456883$ & $40.654710$ \\
                       & $6$ & $40.750512$ & \nan & $40.068340$ \\
-                $0.5$ & $7$ & $40.302718$ & \nan & $39.508500$ \\
-                      & $8$ & $40.263751$ & \nan & $39.399128$ \\
-                      & $9$ & $40.216688$ & \nan & $39.329310$ \\
-                      & $10$ & $40.216251$ & \nan & $39.309409$ \\
-                      & $11$ & $40.216194$ & \nan & $39.296007$ \\
+                $0.5$ & $7$ & $40.302719$ & \nan & $39.508500$ \\
+                      & $8$ & $40.263752$ & \nan & $39.399128$ \\
+                      & $9$ & $40.216688$ & \nan & $39.329311$ \\
+                      & $10$ & $40.216252$ & \nan & $39.309409$ \\
+                      & $11$ & $40.216195$ & \nan & $39.296007$ \\
                       & $12$ & $40.216165$ & \nan & $39.285968$ \\
@@
                       & $6$ & $67.296869$ & \nan & $66.526677$ \\
-                $1.0$ & $7$ & $66.934744$ & \nan & $66.049564$ \\
+                $1.0$ & $7$ & $66.934745$ & \nan & $66.049564$ \\
                       & $8$ & $66.923094$ & \nan & $65.972157$ \\
-                      & $9$ & $66.912244$ & \nan & $65.921204$ \\
+                      & $9$ & $66.912244$ & \nan & $65.921205$ \\
                       & $10$ & $66.912035$ & \nan & $65.889281$ \\
-                      & $11$ & $66.911364$ & \nan & $65.866715$ \\
-                      & $12$ & $66.911364$ & \nan & $65.849775$ \\
+                      & $11$ & $66.911365$ & \nan & $65.866715$ \\
+                      & $12$ & $66.911364$ & \nan & $65.849776$ \\
                 \hline
+                      % NO FURTHER!
                       & $3$ & $120.722260$ & $120.722260$ & $120.722260$ \\
@@
                       & $11$ & $113.245255$ & $119.430435$ & $112.109972$ \\
+                      % TODO: You were here going upwards ^
                       & $12$ & $113.245183$ & $119.394712$ & $112.085683$ \\

### 2018-06-06 project_2/scripts/comparison.py
Fix typo

@@ -1,4 +1,4 @@
-from coupled_cluster.schemes.ccd_sparse import CoupledClusterDoublesSparseD
+from coupled_cluster.schemes.ccd_sparse import CoupledClusterDoublesSparse
 from coupled_cluster.schemes.ccd_optimized import CoupledClusterDoublesOptimized
 from coupled_cluster.hartree_fock.scf_rhf import scf_rhf
 from coupled_cluster.matrix_elements.generate_matrices import (

### 2018-06-06 project_2/scripts/comparison.py
Add new run

@@ -25,9 +25,9 @@ generate_index_map(num_shells)
 
 omega = 2.0
 l = IndexMap.shell_arr[-1]
-n = 2
-theta_hf = 0.3
-theta_ho = 0.3
+n = 12
+theta_hf = 0.1
+theta_ho = 0.1
 max_iterations = 1000
 
 filename = filename.format(l)

### 2018-06-06 project_2/coupled-cluster/coupled_cluster/schemes/cc.py
Add progress output

@@ -43,6 +43,7 @@ class CoupledCluster:
         energy = self._compute_energy()
 
         while diff > tol and iterations < max_iterations:
+            print ("Iteration: {0}\tEnergy: {1}".format(iterations, energy))
             self._compute_amplitudes(theta)
             energy_prev = energy
             energy = self._compute_energy()

### 2018-06-07 project_2/scripts/generate_run_times.py
Add script for generating run times

@@ -1,0 +1,87 @@
+from coupled_cluster.schemes.ccd_sparse import CoupledClusterDoublesSparse
+from coupled_cluster.schemes.ccd_optimized import CoupledClusterDoublesOptimized
+from coupled_cluster.schemes.ccd import CoupledClusterDoubles
+from coupled_cluster.hartree_fock.scf_rhf import scf_rhf
+from coupled_cluster.matrix_elements.generate_matrices import (
+    get_one_body_elements, get_coulomb_elements,
+    get_antisymmetrized_elements, add_spin_to_one_body_elements,
+    get_one_body_elements_spin
+)
+from coupled_cluster.matrix_elements.index_map import (
+    generate_index_map, IndexMap
+)
+from coupled_cluster.hartree_fock.basis_transformation import (
+    transform_one_body_elements, transform_two_body_elements
+)
+
+import numpy as np
+import time
+import os
+
+ccd_method = CoupledClusterDoublesOptimized
+
+file_path = os.path.join("..", "dat")
+filename = os.path.join(file_path, "coulomb_{0}.pkl")
+
+num_shells = 5
+generate_index_map(num_shells)
+
+omega = 1.0
+l = IndexMap.shell_arr[-1]
+n = 2
+theta_hf = 0.1
+theta_ho = 0.1
+max_iterations = 1000
+convergence = 1e-2
+
+filename = filename.format(l)
+
+print ("""
+w = {0},
+num_shells = {1},
+l = {2},
+n = {3},
+theta_hf = {4},
+theta_ho = {5},
+filename = {6}
+""".format(omega, num_shells, l, n, theta_hf, theta_ho, filename))
+
+h = omega * get_one_body_elements(l)
+t0 = time.time()
+u = np.sqrt(omega) * get_coulomb_elements(l, filename=filename, tol=1e-12)
+t1 = time.time()
+print ("Time spent creating Coulomb elements: {0} sec".format(t1 - t0))
+
+t0 = time.time()
+c, energy = scf_rhf(h.todense(), u, np.eye(l//2), n//2, tol=convergence)
+t1 = time.time()
+print ("Time spent in SCF RHF: {0} sec".format(t1 - t0))
+
+print ("\tRHF Energy: {0:.6f}".format(energy))
+
+hi = transform_one_body_elements(h, c)
+t0 = time.time()
+oi = transform_two_body_elements(u, c)
+t1 = time.time()
[…]

### 2018-06-07 project_2/scripts/plot_running_time.py
Add savefig and labels

@@ -1,0 +1,28 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+
+import os
+
+sns.set(color_codes=True)
+
+file_path = os.path.join("..", "dat")
+filename = "iter_run_times_{0}.dat"
+
+fig_path = os.path.join(file_path, "figures")
+fig_name = "iter_run_times_{0}.pdf"
+
+n = 2
+filename = os.path.join(file_path, filename.format(n))
+fig_name = os.path.join(fig_path, fig_name.format(n))
+
+df = pd.read_csv(filename, sep=" *, *", engine="python")
+df.index = df["R"]
+del df["R"]
+
+df.plot()
+plt.xlabel(r"$R$")
+plt.ylabel(r"$t$ [s]")
+plt.title(r"Time spent per CCD iteration")
+plt.savefig(fig_name)
+plt.show()

### 2018-06-07 project_2/article/article.tex
Fix results for N = 12
Fix the last table
Add TODO's
Wrote more.

@@
         \begin{aligned}
-            g(u, \tau) =&
+            g(u, t) =&
                     u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
@@
 
-        Now we have $D_{ij}^{ab}t_{ij}^{ab} = g(u, \tau)$.
+        Now we have $D_{ij}^{ab}t_{ij}^{ab} = g(u, t)$.
         This allows us to define an iterative scheme,
         \begin{equation}
-            \tau^{(k + 1)} = \frac{g(u, \tau^{(k)})}{D_{ij}^{ab}},
+            t^{(k + 1)} = \frac{g(u, t^{(k)})}{D_{ij}^{ab}},
             \label{eq:iterative_amplitude}
@@
         \begin{equation}
-            \tau^{(0)} = \frac{u_{ij}^{ab}}{D_{ij}^{ab}}.
+            t^{(0)} = \frac{u_{ij}^{ab}}{D_{ij}^{ab}}.
         \end{equation}
@@
         \begin{align}
+            \label{eq:intermediate1}
             \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
+            \label{eq:intermediate2}
             \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd} \\
+            \label{eq:intermediate3}
             \chi^a_d &= \frac{1}{2} t^{ac}_{nm} u^{nm}_{cd} \\
+            \label{eq:intermediate4}
             \chi^{bm}_{jc} &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd}
@@
             Hartree-Fock-Roothaan equations are often called the self-consistent-field
-            procedure.}.
+            procedure.}
 
 \section{Implementation}
-    We have costructed a flexible framework for making performing CCD
-    computations, consisting of package for \lstinline{python} that is easy
-    to install globally on any computer.
-
-    \subsection{Installation}
+    We have costructed a flexible framework for making performing Hartree-Fock 
+    self-consistednt field (SCF) and coupled cluster with double excitations (CCD)
+    computations. The entire code base consists of package for \lstinline{python}
+    that is easy to install globally on any computer.
+
+    Because we have employed a mixture of Python and C++ with Cython interfaces,
+    the code sturcture may be difficult to undertand for someone without Cython
+    experience, but we have neverteless tried to make the usage of the software as
+    painless as possible.
+
+    \subsection{Installation and Usage}
+
         The software is easy to install by first cloning the GitHub reopsitory,
@@
 
+        We have included a sample python script in \autoref{app:sample_script}
+        that uses almost all methods and classes in the package.
+
     \subsection{Program Structure}
@@
         given in the directory tree below.
-
-        %% NONE OF THIS WORKED AS INTENDED.
-        %\begin{figure}
-        %\centering
-        %\framebox[0.5\textwidth]{%
-        %\begin{minipage}{0.49\textwidth}
-        %\begin{center}
-        %\begin{varwidth}{0.49\textwidth}
-        %\dirtree{%
-        %.1 coupled\_cluster.
-        %    .2 matrix\_elements.
-        %        .3 generate\_matrices.
-        %        .3 index\_map.
-        %    .2 hartree\_fock.
-        %        .3 scf\_rhf.
-        %        .3 basis\_transformation.
-        %    .2 schemes.
-        %        .3 ccd.
-        %        .3 ccd\_sparse.
-        %        .3 ccd\_optimized.
-        %}
-        %\end{varwidth}
-        %\end{center}
-        %\end{minipage}
-        %}
-        %\end{figure}
 
@@
 
-    \subsection{Mixing}
-        A way to try ``massage'' convergence out of the CCD-method is to use
-        \emph{mixing}. A popular solution is Pulay mixing\footnote{Commonly
-        known as DIIS (direct inversion in the iterative subspace).}. We did not
-        implement Pulay mixing but used a more naïve and simpler solution.
+        Both in \mintinline{python}{CoupledClusterDoublesSparse} and
+        \mintinline{python}{CoupledClusterDoublesOptimized} we make use
+        of the intermediates in equations \ref{eq:intermediate1} to
+        \ref{eq:intermediate4}. In the optimized class, special care was
+        taken to compute
+
+    \subsection{Convergence Problems and Mixing}
+
+        Iterative many-body methods are prone to convergence problems for some
+        configurations. Since Hartree-Fock is a variational method, SCF convergence
+        is found when the energy is stationary with respect to inifitesimal 
+        variations in the orbitals. Unfortunately, the SCF iteration scheme does 
+        not always converge. Luckily, numerous techniques exist for controlling
+        and accelerating convergence\cite{schlegel1991you}. The same kind of 
+        methods have proven useful to ensure convergence of coupled cluster methods
+        \cite{scuseria1986accelerating}.
+
+        The simplest way to try to ``massage'' convergence out of the CCD-method 
+        is to use \emph{damping} where you include a part of the result from the
+        previous iteration i.e.,
         \begin{align}
             \tilde{t}^{k + 1} = \theta t^{k + 1} + (1 - \theta)t^k,
+            \label{eq:mixing}
         \end{align}
@@
         $\tilde{t}^{k + 1}$ as our estimate of the new amplitude.
+
+        A more sophisticated mixing method is the direct inversion 
+        in the iterative subspace (DIIS), also known as Pulay mixing. While the
+        common mixing method is used for many other applications
+        \footnote{It is, for instance, called an alpha filter in data acquisition.},
+        the DIIS method is developed with the sole intent of accelerating
+        convergense in Hartree-Fock methods. In DIIS one would construct a
+        linear combinations of approximate errors from previous iterations,
+        analogous to a very clever weighted moving average. We have not implemented
+        this scheme, but it nevertheless warrants mention.
 
@@
         \centering
-        \caption{N = 2}
+        \caption{Results for N = 2 particles, where convergence was 
+            acchieved for all parameters. Taut's\cite{taut1994two}
+            analytic result for $\omega=1.0a.u.$ is $3$.}
         \begin{ruledtabular}
@@
         \centering
-        \caption{N = 6}
+        \caption{Results for $N = 6$ particles. No convergence for HO
+            basis for low frequency and large number of shells.}
         \begin{ruledtabular}
@@
     \subsection{Validity of results}
-        All of our results have been compared with the master thesis of M. P.
+        The analytical computed energy for $\omega=1.0a.u.$
+        and $N=2$ is $3$\cite{taut1994two}. From our tables we see 
+        that RHF is least successful in reaching this value, CCD in
+        HF basis is best, while CCD with HO basis is generally somewhere 
+        between the two. We see that we get even closer to the analytical
+        ``truth'' by increasing the number of shells.
+
+        The rest of results have been compared with the master thesis of M. P.
         Lohne\cite{lohne2010coupled} for up to 10 shells. Lohne gets two sets of
@@
         increasing to $N = 6$ particles we immediately see how CCD with the HO
-        basis will get in trouble for small values of $\omega$. By increasing
-        the strength of the confinment even this method does converge.
+        basis will get in trouble for small values of $\omega$. Even by increasing
+        the strength of the confinment this method does converge.
 
@@
         convergence within the threshold.
+
+        CCD with HO basis experiences the same convergence problems also for,
+        $N=12$ particles and for $N=20$ particles.
 
@@
 
+    \label{app:sample_script}
+
     \vspace{10pt}

### 2018-06-07 project_2/scripts/plot_running_time.py
Add number of particles to title

@@ -23,6 +23,6 @@ del df["R"]
 df.plot()
 plt.xlabel(r"$R$")
 plt.ylabel(r"$t$ [s]")
-plt.title(r"Time spent per CCD iteration")
+plt.title(r"Time spent per CCD iteration for $N = {0}$ particles".format(n))
 plt.savefig(fig_name)
 plt.show()

### 2018-06-07 project_2/scripts/generate_run_times.py
Add run

@@ -23,15 +23,15 @@ ccd_method = CoupledClusterDoublesOptimized
 file_path = os.path.join("..", "dat")
 filename = os.path.join(file_path, "coulomb_{0}.pkl")
 
-num_shells = 5
+num_shells = 12
 generate_index_map(num_shells)
 
 omega = 1.0
 l = IndexMap.shell_arr[-1]
-n = 2
+n = 6
 theta_hf = 0.1
 theta_ho = 0.1
-max_iterations = 1000
+max_iterations = 1
 convergence = 1e-2
 
 filename = filename.format(l)
@@ -72,6 +72,7 @@ t1 = time.time()
 print ("Time spent antisymmetrizing two body elements: {0} sec".format(t1 - t0))
 
 t0 = time.time()
+#ccd_hf = ccd_method(_h, _u, n)
 ccd_hf = ccd_method(_h.todense(), _u.todense(), n)
 t1 = time.time()
 print ("Time spent setting up {0} code with HF basis: {1} sec"

### 2018-06-07 project_2/scripts/plot_running_time.py
Change fontsize

@@ -12,7 +12,7 @@ filename = "iter_run_times_{0}.dat"
 fig_path = os.path.join(file_path, "figures")
 fig_name = "iter_run_times_{0}.pdf"
 
-n = 2
+n = 6
 filename = os.path.join(file_path, filename.format(n))
 fig_name = os.path.join(fig_path, fig_name.format(n))
 
@@ -21,8 +21,12 @@ df.index = df["R"]
 del df["R"]
 
 df.plot()
-plt.xlabel(r"$R$")
-plt.ylabel(r"$t$ [s]")
-plt.title(r"Time spent per CCD iteration for $N = {0}$ particles".format(n))
+plt.xlabel(r"$R$", fontsize=12)
+plt.ylabel(r"$t$ [s]", fontsize=12)
+plt.title(
+        r"Time spent per CCD iteration for $N = {0}$ particles".format(n),
+        fontsize=14)
+plt.xticks(fontsize=14)
+plt.yticks(fontsize=14)
 plt.savefig(fig_name)
 plt.show()

### 2018-06-07 project_2/article/article.tex
Indent table in Tex-document
Add running time and discussion on running time

@@
 
+        % TODO: Add short introduction to each table and link to them. Leave the
+        % discussion of the results in the discussion section.
+
         \begin{table}
@@
                     \hline
-                      % NO FURTHER!
                           & $3$ & $120.722260$ & $120.722260$ & $120.722260$ \\
-                      & $4$ & $117.339641$ & $120.299999$ & $116.995036$ \\
-                      & $5$ & $113.660396$ & $120.012284$ & $113.048934$ \\
-                      & $6$ & $113.484866$ & $119.872007$ & $112.658821$ \\
-                $2.0$ & $7$ & $113.247601$ & $119.662201$ & $112.309482$ \\
-                      & $8$ & $113.246578$ & $119.584733$ & $112.235521$ \\
-                      & $9$ & $113.246302$ & $119.524417$ & $112.181828$ \\
+                          & $4$ & $117.339642$ & $120.296556$ & $116.995036$ \\
+                          & $5$ & $113.660396$ & $120.007146$ & $113.048934$ \\
+                          & $6$ & $113.484866$ & $119.759037$ & $112.658821$ \\
+                    $2.0$ & $7$ & $113.247601$ & $119.662199$ & $112.309482$ \\
+                          & $8$ & $113.246579$ & $119.584733$ & $112.235521$ \\
+                          & $9$ & $113.246303$ & $119.524394$ & $112.181828$ \\
                           & $10$ & $113.245854$ & $119.472283$ & $112.140661$ \\
-                      & $11$ & $113.245255$ & $119.430435$ & $112.109972$ \\
-                      % TODO: You were here going upwards ^
+                          & $11$ & $113.245256$ & $119.430353$ & $112.109973$ \\
                           & $12$ & $113.245183$ & $119.394712$ & $112.085683$ \\
                     \hline
-                $5.0$ & $3$ & $242.334878$ & $242.334878$ & $242.334878$ \\
-                      & $4$ & $238.739590$ & $241.927593$ & $238.394598$ \\
+                    $5.0$ & $3$ & $242.334879$ & $242.334879$ & $242.334879$ \\
+                          & $4$ & $238.739591$ & $241.927593$ & $238.394598$ \\
                           & $5$ & $234.352741$ & $241.663950$ & $233.680649$ \\
                           & $6$ & $234.282331$ & $241.507595$ & $233.425243$ \\
-                      & $7$ & $234.194820$ & $241.390524$ & $233.226529$ \\
-                      & $8$ & $234.194058$ & $241.293416$ & $233.137932$ \\
-                      & $9$ & $234.190796$ & $241.221056$ & $233.070206$ \\
+                          & $7$ & $234.194820$ & $241.390525$ & $233.226529$ \\
+                          & $8$ & $234.194059$ & $241.293417$ & $233.137933$ \\
+                          & $9$ & $234.190797$ & $241.221056$ & $233.070205$ \\
                           & $10$ & $234.190714$ & $241.158896$ & $233.020009$ \\
                           & $11$ & $234.190665$ & $241.109926$ & $232.980634$ \\
-                      & $12$ & $234.190552$ & $241.068091$ & $232.948528$
+                          & $12$ & $234.190553$ & $241.068091$ & $232.948528$
                 \end{tabular}
@@
                     $1.0$ & $4$ & $177.963297$ & $177.963297$ & $177.963297$ \\
-                      & $5$ & $168.792442$ & $177.206535$ & $168.459124$ \\
-                      & $6$ & $161.339720$ & \nan & $160.594507$ \\
-                      & $7$ & $159.958722$ & \nan & $158.841119$ \\
+                          & $5$ & $168.792442$ & $177.206536$ & $168.459124$ \\
+                          & $6$ & $161.339721$ & \nan & $160.594507$ \\
+                          & $7$ & $159.958722$ & \nan & $158.841120$ \\
                           & $8$ & $158.400172$ & \nan & $157.038330$ \\
                           & $9$ & $158.226030$ & \nan & $156.676039$ \\
-                      & $10$ & $158.017666$ & \nan & $156.367930$ \\
+                          & $10$ & $158.017667$ & \nan & $156.367930$ \\
                           & $11$ & $158.010276$ & \nan & $156.292422$ \\
@@
                     $2.0$ & $4$ & $286.825295$ & $286.825295$ & $286.825295$ \\
-                      & $5$ & $276.898195$ & $286.159165$ & $276.381708$ \\
-                      & $6$ & $267.269712$ & $266.413122$ & $285.614965$ \\
+                          & $5$ & $276.898196$ & $286.159148$ & $276.381708$ \\
+                          & $6$ & $267.269712$ & $285.614958$ & $266.413122$ \\
                           & $7$ & $266.213200$ & \nan & $264.969415$ \\
-                      & $8$ & $264.933621$ & \nan & $263.434546$ \\
+                          & $8$ & $264.933622$ & \nan & $263.434546$ \\
                           & $9$ & $264.874009$ & \nan & $263.215451$ \\
-                      & $10$ & $264.809953$ & \nan & $263.046184$ \\
-                      & $11$ & $264.809900$ & \nan & $262.963702$ \\
-                      & $12$ & $264.809306$ & \nan & $262.899697$ \\
+                          & $10$ & $264.809954$ & \nan & $263.046195$ \\
+                          & $11$ & $264.809901$ & \nan & $262.963703$ \\
+                          & $12$ & $264.809306$ & \nan & $262.899698$ \\
                     \hline
-                $5.0$ & $4$ & $563.773951$ & $563.773951$ & $563.773951$ \\
+                    $5.0$ & $4$ & $563.773952$ & $563.773952$ & $563.773952$ \\
                           & $5$ & $552.630093$ & $563.160136$ & $552.118708$ \\
-                      & $6$ & $540.804719$ & $562.692231$ & $539.824400$ \\
-                      & $7$ & $540.227792$ & $562.306124$ & $538.886073$ \\
-                      & $8$ & $539.499326$ & $562.114280$ & $537.925127$ \\
-                      & $9$ & $539.495940$ & \nan & $537.769045$ \\
-                      & $10$ & $539.494611$ & \nan & $537.646667$ \\
-                      & $11$ & $539.493512$ & \nan & $537.548827$ \\
+                          & $6$ & $540.804720$ & $562.692231$ & $539.824400$ \\
+                          & $7$ & $540.227793$ & $562.306123$ & $538.886074$ \\
+                          & $8$ & $539.499326$ & $562.114279$ & $537.925127$ \\
+                          & $9$ & $539.495941$ & \nan & $537.769045$ \\
+                          & $10$ & $539.494611$ & \nan & $537.646668$ \\
+                          & $11$ & $539.493513$ & \nan & $537.548828$ \\
                           & $12$ & $539.491764$ & \nan & $537.470616$ \\
@@
                     $10.0$ & $4$ & $973.032700$ & $973.032700$ & $973.032700$ \\
-                       & $5$ & $961.371081$ & $972.439477$ & $960.862053$ \\
+                           & $5$ & $961.371081$ & $972.439478$ & $960.862053$ \\
                            & $6$ & $948.057077$ & $972.002302$ & $947.019789$ \\
-                       & $7$ & $947.765473$ & $971.716015$ & $946.399546$ \\
-                       & $8$ & $947.410304$ & $971.508583$ & $945.827806$ \\
-                       & $9$ & $947.409439$ & $971.332132$ & $945.663819$ \\
-                       & $10$ & $947.404929$ & $971.193592$ & $945.528489$ \\
-                       & $11$ & $947.404360$ & $971.076617$ & $945.424174$ \\
-                       & $12$ & $947.403875$ & $970.978571$ & $945.339079$ \\
+                           & $7$ & $947.765474$ & $971.716015$ & $946.399546$ \\
+                           & $8$ & $947.410305$ & $971.508584$ & $945.827806$ \\
+                           & $9$ & $947.409440$ & $971.332133$ & $945.663820$ \\
+                           & $10$ & $947.404930$ & $971.193592$ & $945.528489$ \\
+                           & $11$ & $947.404361$ & $971.076617$ & $945.424175$ \\
+                           & $12$ & $947.403875$ & $970.978571$ & $945.339080$ \\
                 \end{tabular}
@@
         \end{table}
+
+    \subsection{Running time}
+        Here we show how the three implementations of the CCD-method scales as a
+        function of the shell number $R$ for $N = 2$ and $N = 6$. The same
+        behaviour is shown for higher number of particles.
+
+        \begin{figure}
+            \centering
+            \includegraphics[width=244px]
+            {../dat/figures/iter_run_times_2.pdf}
+            \caption{In this figure we can see the running time per iteration
+            for $N = 2$ particles as a function of the shell number $R$. We did
+            not run the naïve implementation for $R = 12$.}
+            \label{fig:running_time_2}
+        \end{figure}
+
+        \begin{figure}
+            \centering
+            \includegraphics[width=244px]
+            {../dat/figures/iter_run_times_6.pdf}
+            \caption{Here we see the running time per iteration for $N = 6$
+            particles as a function of the shell number $R$. We did not run the
+            naïve implementation for $R > 9$ in this figure.}
+            \label{fig:running_time_6}
+        \end{figure}
 
@@
         but we can get a ``ball-park'' idea to benchmark against.
+
+    \subsection{Running time}
+        Looking at the running times shown in \autoref{fig:running_time_2} and
+        \autoref{fig:running_time_6} we can see how the naïve implementation
+        blows up very fast. As both the sparse and the optimized version uses
+        intermediate calculations these running times scales on the order of
+        $\mathcal{O}(l^6)$ whereas the naïve implementation goes as
+        $\mathcal{O}(l^8)$. Another important factor is that the naïve
+        implementation uses \texttt{np.einsum} to calculate the tensor
+        contractions whereas the sparse and optimized version uses
+        \texttt{sparse.tensordot} and \texttt{np.matmul} respectively. The two
+        latter function calls uses BLAS's implementation of the matrix dot which
+        is way faster than Numpys \texttt{np.einsum}.
 

### 2018-06-07 project_2/article/article.tex
Add table captions

@@
 
-\newcommand{\nan}{\DejaSans 😱}
+\newcommand{\nan}{{\DejaSans 😱}}
 
@@
             \centering
-            \caption{Results for N = 2 particles, where convergence was 
-                acchieved for all parameters. Taut's\cite{taut1994two}
-                analytic result for $\omega=1.0a.u.$ is $3$.}
+            \caption{Results for $N = 2$ particles, where convergence was
+            achieved for all parameters. Taut's\cite{taut1994two} analytic
+            result for $\omega = 1.0$ a.u. is $3$.}
             \begin{ruledtabular}
@@
             \centering
-            \caption{Results for $N = 6$ particles. No convergence for HO
-                basis for low frequency and large number of shells.}
+            \caption{Results for $N = 6$ particles. No convergence is marked as
+            {\nan} in the table. For low frequencies the CCD-method using the HO
+            basis has a hard time achieving convergence.}
             \begin{ruledtabular}
@@
             not achieve convergence using the harmonic oscillator basis for the
-            lower frequency values and large number of shells.}
+            lower frequency values and large number of shells. No convergence is
+            marked as {\nan} in the table.}
             \begin{ruledtabular}
@@
             \centering
-            \caption{In this table we look at $N = 20$ particles ($R = 4$ shells).
-            We did not achieve convergence using the harmonic oscillator basis for
-            the lower frequency values and large number of shells.}
+            \caption{In this table we look at $N = 20$ particles ($R = 4$
+            shells).  We did not achieve convergence using the harmonic
+            oscillator basis for the lower frequency values and large number of
+            shells. No convergence is marked as {\nan} in the table.}
             \begin{ruledtabular}

### 2018-06-07 project_2/article/article.tex

@@
         &= -\half\nabla_i^2 + \half\omega^2 \vf{r}_i^2,
+        \label{eq:one_body_hamiltonian}
     \end{align}
@@
     % Why not write $\hbar = c = e = m = 1$?
-    The two-body part is the Coulomb interaction potential.
+    The two-body part is the Coulomb interaction potential,
     \begin{align}
@@
         &= \frac{1}{\abs{\vf{r}_i - \vf{r}_j}}.
+        \label{eq:two_body_operator}
     \end{align}
@@
     for any excitation.
+
+    \subsection{Quantum Dots}
+
+        Some consideration should be given towards why the system described above
+        is a quantum dot, and why we should care about quantum dots.
+        There are at least two things we refer to as quantum dots. The
+        first one is a man-made device, also known as artificial atoms,
+        and the other is the simplest theoretical construct of such a device.
+
+        Aritifical atom quantum dots are semiconducting crystals, typically GaAs
+        relatively easy to build in a laboratory. Here, some free electrons are 
+        confined to a volume, either by a physical barrier like an insulator
+        or an electromagnetic field\cite{reimann2002electronic}. Due to the
+        confinement of the electrons, the energy levels of a quantum dot becomes
+        quantized, and they therefore hold many properties similar to naturally
+        occuring quantized systems, which is the reason why quantum dots have
+        attracted much interest.
+
+        The theoretical approximation of a quantum dot is impossible to make perfectly
+        accurate, yet simple. There are some important effects that should be
+        included such the potential arising as a result of the confinement, and
+        the interaction between particles. We therefore model our system with an
+        electron-electron interaction, given by the two-body operator in
+        \autoref{eq:eq:two_body_operator}, and with we approximate the confining
+        potential by the harmonic oscillator potential, given by the second term
+        in the one-body Hamiltonian in \autoref{eq:eq:one_body_hamiltonian}. The
+        last term in the one-body Hamiltonian is the kinetic energy of particle
+        in question, essential in any system.
 

### 2018-06-07 project_2/article/article.tex
Theory on quantum dots.

@@
 
-\newcommand{\nan}{{\DejaSans 😱}}
+\newcommand{\nan}{\DejaSans 😱}
 
@@
         \centering
-            \caption{Results for $N = 2$ particles, where convergence was
-            achieved for all parameters. Taut's\cite{taut1994two} analytic
-            result for $\omega = 1.0$ a.u. is $3$.}
+        \caption{Results for N = 2 particles, where convergence was 
+            acchieved for all parameters. Taut's\cite{taut1994two}
+            analytic result for $\omega=1.0a.u.$ is $3$.}
         \begin{ruledtabular}
@@
         \centering
-            \caption{Results for $N = 6$ particles. No convergence is marked as
-            {\nan} in the table. For low frequencies the CCD-method using the HO
-            basis has a hard time achieving convergence.}
+        \caption{Results for $N = 6$ particles. No convergence for HO
+            basis for low frequency and large number of shells.}
         \begin{ruledtabular}
@@
         not achieve convergence using the harmonic oscillator basis for the
-            lower frequency values and large number of shells. No convergence is
-            marked as {\nan} in the table.}
+        lower frequency values and large number of shells.}
         \begin{ruledtabular}
@@
         \centering
-            \caption{In this table we look at $N = 20$ particles ($R = 4$
-            shells).  We did not achieve convergence using the harmonic
-            oscillator basis for the lower frequency values and large number of
-            shells. No convergence is marked as {\nan} in the table.}
+        \caption{In this table we look at $N = 20$ particles ($R = 4$ shells).
+        We did not achieve convergence using the harmonic oscillator basis for
+        the lower frequency values and large number of shells.}
         \begin{ruledtabular}
@@
 
-    \subsection{Running time}
-        Here we show how the three implementations of the CCD-method scales as a
-        function of the shell number $R$ for $N = 2$ and $N = 6$. The same
-        behaviour is shown for higher number of particles.
-
-        \begin{figure}
-            \centering
-            \includegraphics[width=244px]
-            {../dat/figures/iter_run_times_2.pdf}
-            \caption{In this figure we can see the running time per iteration
-            for $N = 2$ particles as a function of the shell number $R$. We did
-            not run the naïve implementation for $R = 12$.}
-            \label{fig:running_time_2}
-        \end{figure}
-
-        \begin{figure}
-            \centering
-            \includegraphics[width=244px]
-            {../dat/figures/iter_run_times_6.pdf}
-            \caption{Here we see the running time per iteration for $N = 6$
-            particles as a function of the shell number $R$. We did not run the
-            naïve implementation for $R > 9$ in this figure.}
-            \label{fig:running_time_6}
-        \end{figure}
-
 \section{Discussion}
@@
         but we can get a ``ball-park'' idea to benchmark against.
-
-    \subsection{Running time}
-        Looking at the running times shown in \autoref{fig:running_time_2} and
-        \autoref{fig:running_time_6} we can see how the naïve implementation
-        blows up very fast. As both the sparse and the optimized version uses
-        intermediate calculations these running times scales on the order of
-        $\mathcal{O}(l^6)$ whereas the naïve implementation goes as
-        $\mathcal{O}(l^8)$. Another important factor is that the naïve
-        implementation uses \texttt{np.einsum} to calculate the tensor
-        contractions whereas the sparse and optimized version uses
-        \texttt{sparse.tensordot} and \texttt{np.matmul} respectively. The two
-        latter function calls uses BLAS's implementation of the matrix dot which
-        is way faster than Numpys \texttt{np.einsum}.
 

### 2018-06-07 project_2/article/article.tex

--- 2018-06-07 project_2/article/article.tex
@@
             \centering
-            \caption{Results for $N = 2$ particles, where convergence was
-            achieved for all parameters. Taut's\cite{taut1994two} analytic
-            result for $\omega = 1.0$ a.u. is $3$.}
+            \caption{Results for N = 2 particles, where convergence was 
+                acchieved for all parameters. Taut's\cite{taut1994two}
+                analytic result for $\omega=1.0a.u.$ is $3$.}
             \begin{ruledtabular}
@@
             \centering
-            \caption{Results for $N = 6$ particles. No convergence is marked as
-            {\nan} in the table. For low frequencies the CCD-method using the HO
-            basis has a hard time achieving convergence.}
+            \caption{Results for $N = 6$ particles. No convergence for HO
+                basis for low frequency and large number of shells.}
             \begin{ruledtabular}
@@
             not achieve convergence using the harmonic oscillator basis for the
-            lower frequency values and large number of shells. No convergence is
-            marked as {\nan} in the table.}
+            lower frequency values and large number of shells.}
             \begin{ruledtabular}
@@
             \centering
-            \caption{In this table we look at $N = 20$ particles ($R = 4$
-            shells).  We did not achieve convergence using the harmonic
-            oscillator basis for the lower frequency values and large number of
-            shells. No convergence is marked as {\nan} in the table.}
+            \caption{In this table we look at $N = 20$ particles ($R = 4$ shells).
+            We did not achieve convergence using the harmonic oscillator basis for
+            the lower frequency values and large number of shells.}
             \begin{ruledtabular}

### 2018-06-07 project_2/article/article.tex
More on HO -> HF conversion

--- 2018-06-07 project_2/article/article.tex
@@
             equations for electron $i$\footnote{This is weird, as electron
-            should be indistinguishable and therefor impossible to label.} is
+            should be indistinguishable and therefore impossible to label.} is
             \begin{equation}
@@
             procedure.}
+
+            Because the Hartree-Fock method is a variational method, the initial
+            wavefunction made up of the single-partrice wavefunctions $\varphi_i$
+            is varied until the minimal energy is found. The transformations necessary
+            to make all these variations will now be contained in $C$. So if we start from
+            a Harmonic Oscillator (HO) basis we can transform into a better basis, called the
+            Hartree-Fock (HF) basis,
+            \begin{equation}
+                \ket{\Psi}^{\text{HF}}_p = \sum_{k=1}^L C_{kp} \ket{\Psi}_k^{\text{HO}}.
+            \end{equation}
+            We will see how the coupled cluster method with double excitations works with
+            both the HO basis and the HF basis.
 

### 2018-06-07 project_2/article/article.tex
Performance comments in discussion

@@
 
+    \subsection{Performance comparison}
+        Somewhat surprisingly, the Hartree-Fock method outperforms the coupled
+        cluster doubles when the harmonic oscillator basis is used. This
+        happens for almost any configuration except the smallest system
+        with $N=2$ particles. This comes from the fact that the HO basis is 
+        not very good at representing the states in the system. As we have
+        already stated, because the Hartree-Fock SCF iterations is a variational
+        methdod, the basis is varied in order to find the minimal energy. The new
+        basis states will be linear combinations of the old basis states. When
+        the iterations are finished, the new transformed HF basis will then be 
+        much better at representing the possible quantum configurations compared 
+        with the HO basis. This effect is only apparent in system with a larger
+        number of particles and/or low frequency. This is because in a system with 
+        a large number of particles relative to the frequency particle-particle
+        interaction will be more prominent than the harmonic oscillator potential
+        the particles are subject to. I other words, the HO basis fits rather
+        well.
+
     \subsection{Validity of results}

### 2018-06-08 project_2/article/article.tex
Thorough work-through.

@@
 \begin{abstract}
-    Something about coupled-cluster... Preferably doubles.
+    The coupled cluster method for estimating the ground state energy of a many-body
+    system is arguably the most important post-Hartree-Fock method. We have implemented
+    the coupled cluster method with double excitations (CCD). We find that it does not
+    perform well with a simple harmonic oscillator basis, but the results vastly improve 
+    when it is provided a basis from a restricted Hatree-Fock self-consistent field method
+    run as a preliminary computation. For a system consisting of $N=2$ interacting
+    particles, using 12 orbital shells we get a ground stated energy of $3.005970$, very
+    close to the analytical ``true answer'' of $3$\cite{taut1994two}. 
 \end{abstract}
@@
 \section{Introduction}
-    %TODO: Add explanation on why the many-body problem becomes intractable.
-    % This can be shown by demonstrating that the true wavefunction of the system
-    % must be a linear combination of all possible Slater detemerminants built up
-    % from permutations of the single particle functions in the basis. Thus the
-    % energy equation becomes too large.
-    In this project we will study the ground state energy of quantum dots.
+    In this project we will compute the ground state energy of quantum dots, a
+    construct that has become a popular area of study in recent decades, both from
+    a theoretical and an experimental point of view. We achieve the ground state
+    computation by implementing a coupled cluster amplitude equation solver
+    which is used in succession with a restricted Hartree-Fock self-consistent
+    field method.
+
+    Regardless of how sophisticated the method used, the many-body problem remains
+    quite intractable. The true wavefunction of any system will be a linear 
+    combination of all possible slater-type orbitals one can model. This means that
+    in order to represent a system correctly, one would need a basis set of infinite
+    size. This is an obvious problem when we solve problems numerically, because we
+    obviously need to limit the size of the basis we use because we have limitations
+    in memeory and computating power. Luckily, great minds have bestowed a wide 
+    selection of very god numerical methods upon the world, which the coupled cluster
+    is a part of.
+
+    First, we provide a thorough theoretical outline of how we model a quantum dot, 
+    the mathematical formalism behind many-body problems and outline how the 
+    computational methods we employ both turn into iterative schemes that are 
+    easy to represent numerically. Second, we provide a brief overview of the
+    program we have constructed which is in the form an installable python package.
+    Third, we have employed the methods on a large parameter space presenting 
+    results for systems consisting of $N= \{2, 6, 12, 20\}$ particles. Fourth,
+    we discuss the most interesting aspects of the results and try to shed some
+    light over the shortcomings as well as the beneficial features of the different
+    metods. Lastly, we end on some summary remarks.
 
@@
 
-    \subsection{Quantum Dots}
-
-        Some consideration should be given towards why the system described above
-        is a quantum dot, and why we should care about quantum dots.
+    \subsection{\emph{Motivational Interlude}: Quantum Dots}
+
+        The system described above is a theoretical approximation of a
+        quantum dot. Some consideration should be given to an explanation of 
+        why the system described above is a quantum dot, and why we should
+        care about quantum dots.
         There are at least two things we refer to as quantum dots. The
@@
 
-        Aritifical atom quantum dots are semiconducting crystals, typically GaAs
-        relatively easy to build in a laboratory. Here, some free electrons are 
-        confined to a volume, either by a physical barrier like an insulator
+        Aritifical atom quantum dots are semiconducting crystals, typically GaAs,
+        and they are
+        relatively easy to build in a laboratory. Some free electrons are 
+        confined to a volume, either by a physical barrier, like an insulator,
         or an electromagnetic field\cite{reimann2002electronic}. Due to the
@@
         electron-electron interaction, given by the two-body operator in
-        \autoref{eq:eq:two_body_operator}, and with we approximate the confining
+        \autoref{eq:two_body_operator}, and with we approximate the confining
         potential by the harmonic oscillator potential, given by the second term
-        in the one-body Hamiltonian in \autoref{eq:eq:one_body_hamiltonian}. The
+        in the one-body Hamiltonian in \autoref{eq:one_body_hamiltonian}. The
         last term in the one-body Hamiltonian is the kinetic energy of particle
@@
         \end{gather}
-        Note that we use the chemists notation to label the antisymmetric matrix
+        Note that we use the ``chemists notation'' to label the antisymmetric matrix
         elements.
@@
         \end{align}
-        In the doubles approximation we limit the cluster operator to
+        In our approximation we limit the cluster operator to only include
+        double excitations,
         \begin{align}
@@
         \end{align}
-        to find an expression that can be used to solve for $t_{ij}^{ab}$.
+        to find an expression that can be used to solve for the amplitudes 
+        $t_{ij}^{ab}$.
         By employing the Baker-Campbell-Haussdorf (BCH) expansion,
@@
         The subscript $c$ indicates that only those terms in which the
-        Hamiltonian is connected\footnote{In a Wick's theorem sense}
+        Hamiltonian is connected\footnote{Again, in aWick's theorem sense}
         to every cluster operator on its right should be included. Since
@@
         operators, $H_N$ can connect to at most four cluster operators
-        at once. Therefor, the BCH expansion must truncate at the
+        at once. Therefore, the BCH expansion must truncate at the
         fourth-order terms.
@@
 
-        The we can define a new function consisting of all of \autoref{eq:CCD_amp_1}
-        except term number two and three,
+        Then we can define a new version of \autoref{eq:CCD_amp_1},
         \begin{equation}
@@
             g(u, t) =&
-                    u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
+                    u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab) - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
                  +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
@@
         Looking closely at the amplitude equation in (\ref{eq:CCD_amp_1}) one might come
-        to realize that it contains that this equation contains many of the same structures in
+        to realize that this equation contains many of the same structures in
         several of the terms. This warrants the search for an algebraic transformation of the
@@
             Having found the matrix elements of $h$ and $u$ we can now use the
-            \emph{self-consistent field iteration} method to construct $h$ and
-            $u$ in a \emph{restricted Hartree-Fock} basis. This will yield a
+            \emph{self-consistent field} (SCF) iteration method to construct $h$ and
+            $u$ in a \emph{restricted Hartree-Fock} (RHF) basis. This will yield a
             better estimate to the actual, unknown basis functions of the
             system.
-
-            % See Step 3 in:
-            % http://sirius.chem.vt.edu/wiki/doku.php?id=crawdad:programming:project4
 
@@
             where $f_i$ is the Fock operator, $\varphi_i$ are eigenstates of the Fock
-            operator consiting of a set of one-electorn wave funcitons, called the
+            operator consiting of a set of one-electorn wave funcitons called the
             Hartree-Fock molecular orbitals, and $\varepsilon_i$ are the eigenenergies
@@
 
-            Because the Hartree-Fock method is a variational method, the initial
-            wavefunction made up of the single-partrice wavefunctions $\varphi_i$
+            Because the RHF SCF method is a variational method, the initial
+            wavefunction made up of the single-particle wavefunctions $\varphi_i$
             is varied until the minimal energy is found. The transformations necessary
             to make all these variations will now be contained in $C$. So if we start from
-            a Harmonic Oscillator (HO) basis we can transform into a better basis, called the
+            a harmonic oscillator (HO) basis we can transform into a better basis, called a
             Hartree-Fock (HF) basis,
@@
             We will see how the coupled cluster method with double excitations works with
-            both the HO basis and the HF basis.
+            both the HO basis and the HF basis, but not equally well.
 
@@
     the code sturcture may be difficult to undertand for someone without Cython
-    experience, but we have neverteless tried to make the usage of the software as
+    experience, but we have tried to make the actual usage of the software as
     painless as possible.
+
+    Most of the computations that are done in a coupled cluster algorithm
+    is linear algebra and we found that NumPy, relying on
+    BLAS and LAPACK, is most likely as fast as it gets and, at the very least,
+    efficient enough for our purposes. The quicker and easier scripting capabilities
+    Python provides is worth what little extra speed-up writing the entire
+    code base in C++ could have provided.
 
@@
         To ensure that all requirements of the program are satified you can
-        install with \lstinline{make installr} instead.
+        install with \mintinline{bash}{make installr} instead.
 
         We have included a sample python script in \autoref{app:sample_script}
-        that uses almost all methods and classes in the package.
+        that uses almost all methods and classes in the package, and also compares
+        the time each methods takes to fininsh the computations.
 
@@
         of the intermediates in equations \ref{eq:intermediate1} to
-        \ref{eq:intermediate4}. In the optimized class, special care was
-        taken to compute
+        \ref{eq:intermediate4}. We have also relied upon implementations
+        in C++ for some of the heaviest computations. In the optimized
+        class special care was taken to ensure that every contraction
+        was computed in an optimal way, with the minimization of memory
+        use and CPU time as the overall goal. Moreover, several methods
+        in NumPy for doing linear algebra was tested, including but not limited
+        to \mintinline{python}{numpy.einsum()}, \mintinline{python}{numpy.tensordot},
+        and reshaping datastructures to two dimensions and using 
+        \mintinline{python}{numpy.dot}. Additionally, we tried to make use of 
+        the python package \mintinline{python}{dask} which allows for arrays that
+        are larger than the computer's memory buffer as well as automated
+        concurrency. We found that the dask arrays was a bad fit for our code
+        base, but we ended up using \mintinline{python}{numba} for automated
+        paralellization of some iterative loops as well as just-in-time
+        compialtion of some functions.
 
@@
 
-        % TODO: Add short introduction to each table and link to them. Leave the
-        % discussion of the results in the discussion section.
+        All the results from our simulations are displayed in tables \ref{tab:N2},
+        \ref{tab:N6}, \ref{tab:N12} and \ref{tab:N20}. They show the ground state
+        energy as a result of restricted Hartree-Fock SCF iterations, and coupled
+        cluster doubles with harmonic oscillator basis and Hatree-Fock basis for 
+        systems of increasing size, $N=2$, $N=6$, $N=12$ and $N=20$ particles
+        respectively. The results are thoroughtly discussed in the next section.
 
@@
                     \hline
-                    $5.0$ & $3$ & $242.334879$ & $242.334879$ & $242.334879$ \\
+                          & $3$ & $242.334879$ & $242.334879$ & $242.334879$ \\
                           & $4$ & $238.739591$ & $241.927593$ & $238.394598$ \\
@@
                           & $6$ & $234.282331$ & $241.507595$ & $233.425243$ \\
-                          & $7$ & $234.194820$ & $241.390525$ & $233.226529$ \\
+                    $5.0$ & $7$ & $234.194820$ & $241.390525$ & $233.226529$ \\
                           & $8$ & $234.194059$ & $241.293417$ & $233.137933$ \\
@@
                     \hline
-                    $1.0$ & $4$ & $177.963297$ & $177.963297$ & $177.963297$ \\
+                          & $4$ & $177.963297$ & $177.963297$ & $177.963297$ \\
                           & $5$ & $168.792442$ & $177.206536$ & $168.459124$ \\
@@
                           & $7$ & $159.958722$ & \nan & $158.841120$ \\
-                          & $8$ & $158.400172$ & \nan & $157.038330$ \\
+                    $1.0$ & $8$ & $158.400172$ & \nan & $157.038330$ \\
                           & $9$ & $158.226030$ & \nan & $156.676039$ \\
@@
                     \hline
-                    $2.0$ & $4$ & $286.825295$ & $286.825295$ & $286.825295$ \\
+                          & $4$ & $286.825295$ & $286.825295$ & $286.825295$ \\
                           & $5$ & $276.898196$ & $286.159148$ & $276.381708$ \\
@@
                           & $7$ & $266.213200$ & \nan & $264.969415$ \\
-                          & $8$ & $264.933622$ & \nan & $263.434546$ \\
+                    $2.0$ & $8$ & $264.933622$ & \nan & $263.434546$ \\
                           & $9$ & $264.874009$ & \nan & $263.215451$ \\
@@
                     \hline
-                    $5.0$ & $4$ & $563.773952$ & $563.773952$ & $563.773952$ \\
+                          & $4$ & $563.773952$ & $563.773952$ & $563.773952$ \\
                           & $5$ & $552.630093$ & $563.160136$ & $552.118708$ \\
@@
                           & $7$ & $540.227793$ & $562.306123$ & $538.886074$ \\
-                          & $8$ & $539.499326$ & $562.114279$ & $537.925127$ \\
+                    $5.0$ & $8$ & $539.499326$ & $562.114279$ & $537.925127$ \\
                           & $9$ & $539.495941$ & \nan & $537.769045$ \\
@@
                     \hline
-                    $10.0$ & $4$ & $973.032700$ & $973.032700$ & $973.032700$ \\
+                           & $4$ & $973.032700$ & $973.032700$ & $973.032700$ \\
                            & $5$ & $961.371081$ & $972.439478$ & $960.862053$ \\
@@
                            & $7$ & $947.765474$ & $971.716015$ & $946.399546$ \\
-                           & $8$ & $947.410305$ & $971.508584$ & $945.827806$ \\
+                    $10.0$ & $8$ & $947.410305$ & $971.508584$ & $945.827806$ \\
                            & $9$ & $947.409440$ & $971.332133$ & $945.663820$ \\
@@
         Here we show how the three implementations of the CCD-method scales as a
-        function of the shell number $R$ for $N = 2$ and $N = 6$. The same
-        behaviour is shown for higher number of particles.
+        function of the shell number $R$ for $N = 2$ and $N = 6$, in table 
+        \ref{fig:running_time_2} and \ref{fig:running_time_6} respectively. The same
+        behaviour is shown for higher number of particles. In \autoref{fig:running_time_6_zoom}
+        we have plotted the computation time for $N=6$ particles again, but we
+        have excluded the naïve method for better comparison of the sparse and
+        optimized method.
 
@@
             \label{fig:running_time_6}
+        \end{figure}
+
+        \begin{figure}
+           \centering
+            \includegraphics[width=244px]
+            {../dat/figures/iter_run_times_zoom.pdf}
+            \caption{Run time for $N=6$ particles, only the faster methods are
+            included for better comparison.}
+            \label{fig:running_time_6_zoom}            
         \end{figure}
@@
         much better at representing the possible quantum configurations compared 
-        with the HO basis. This effect is only apparent in system with a larger
+        to the HO basis. This effect is only apparent in system with a larger
         number of particles and/or low frequency. This is because in a system with 
@@
         interaction will be more prominent than the harmonic oscillator potential
-        the particles are subject to. I other words, the HO basis fits rather
+        the particles are subject to. I other words, the HO basis would fit rather
         well.
@@
     \subsection{Validity of results}
-        The analytical computed energy for $\omega=1.0a.u.$
+        The analytical computed energy for $\omega=1.0$ a.u.
         and $N=2$ is $3$\cite{taut1994two}. From our tables we see 
@@
 
-    \subsection{Sparse implementation}
-
+    %\subsection{Sparse implementation}
+    % Did you want to write something about memory use, Oyvind? Memory problems?
+
+    \section{Summary Remarks}
+
+        Our main product from this is a product that can do both computations with
+        restricted Hartree-Fock (RHF) self-consistent field and coupled cluster 
+        doubles (CCD). We have tried
+        to make a package in python that should be easy enough for anyone to test. 
+        Moreover we hav demonstrated that one can speed up the computation time with
+        a relatively simple algebraic transformation and finding the ``intermediates''.
+        The computational speed was further reduced using sparse matrices. It should
+        be noted, as a warning, that there are relatively few systems adequately sparese
+        for this to work. We added upon the program to increase it further with 
+        paralellization and optimal use of numerical functions.
+
+        In order to perform the computations we have constructed a harmonic oscillator
+        basis. We saw how the CCD method was outperformed RHF in most cases, and often
+        exhibited convergence problems. The remedy for these shortcomings is to use
+        the Hartree-Fock basis which is provided after successful convergence of 
+        SCF iterations. The results are improved markedly and classification of coupled
+        cluster methods as a post-Harte-Fock method is justified.
+
+        The coupled cluster method provides surprisingly good results even
+        though one only includes double esctiations. The reason for this is that
+        particle-particle interactions is the amongt the most important thing that 
+        happens in a ferminon system, second only to no interaction at all, i.e 
+        what particles do on their own. The way to improve the computations is to
+        include other types of excitations like singles (CCSD) and triples (CCSDT),
+        add more basis functions, and/or supplement the coupled cluster method with 
+        another method like a perturbative one. All of this will, needless to say,
+        increase the compute time, which must be balanced against the accuracy needed
+        for the situation at hand.
+
+\vfill
+\newpage
 \appendix
@@
                   + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c3 a^\dagger_c \c5 a^\dagger_d \c2 a_l \c1 a_k } \}
-                \Big) \\
+                \Big)
+        \end{aligned}
+        \end{equation}
+        \begin{equation}
+        \begin{aligned}
                 =& \sum_{pq}\sum_{klcd}f_{pq} t_{kl}^{cd} \\
@@
     indices if we contract them,
-    \begin{align}
-        &\begin{aligned}
+    \begin{equation}
+        \begin{aligned}
         0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
@@
           &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn},
-        \end{aligned}\\
-        &\begin{aligned}
+        \end{aligned}        
+    \end{equation}
+    \begin{equation}
+        \begin{aligned}
         0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
@@
           &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}
-        \end{aligned} \\
-        &\begin{aligned}
+        \end{aligned}
+    \end{equation}        
+    \begin{equation}
+        \begin{aligned}
         0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
@@
           &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}
-        \end{aligned} \\
-        &\begin{aligned}
+        \end{aligned}
+    \end{equation}
+    \begin{equation}
+        \begin{aligned}
         0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
@@
         \end{aligned}
-    \end{align}
+    \end{equation}
     Now we can introduce the intermediate $\chi$-terms,
@@
 
-\section{Coupled cluster doubles diagrams}
-    In order to get an expression for the energy equation and the amplitude
-    equations we use a diagrammatic approach.
-
+%\section{Coupled cluster doubles diagrams}
+%    In order to get an expression for the energy equation and the amplitude
+%    equations we use a diagrammatic approach.
+
+\vfill
+\clearpage
+\newpage
+\begin{widetext}
 \section{Example Script}
@@
     \vspace{10pt}
-    \begin{widetext}
-    \inputminted[]{python}{../scripts/comparison.py}
+
+    \inputminted[firstline=0, lastline=113]{python}{../scripts/comparison.py}
+
 \end{widetext}

### 2018-06-08 project_2/scripts/comparison.py

@@ -67,16 +67,16 @@ _u = get_antisymmetrized_elements(l, oi=oi, tol=1e-12)
 t1 = time.time()
 print ("Time spent antisymmetrizing two body elements: {0} sec".format(t1 - t0))
 
-#t0 = time.time()
-#ccd_hf_sparse = CoupledClusterDoublesSparse(_h, _u, n)
-#t1 = time.time()
-#print ("Time spent setting up CCD code with HF basis: {0} sec".format(t1 - t0))
-#
-#t0 = time.time()
-#energy, iterations = ccd_hf_sparse.compute_energy(tol=1e-4, theta=theta)
-#t1 = time.time()
-#print ("Time spent computing CCD energy with HF basis: {0} sec".format(t1 - t0))
-#print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}".format(energy, iterations, (t1 - t0)/iterations))
+t0 = time.time()
+ccd_hf_sparse = CoupledClusterDoublesSparse(_h, _u, n)
+t1 = time.time()
+print ("Time spent setting up CCD code with HF basis: {0} sec".format(t1 - t0))
+
+t0 = time.time()
+energy, iterations = ccd_hf_sparse.compute_energy(tol=1e-4, theta=theta)
+t1 = time.time()
+print ("Time spent computing CCD energy with HF basis: {0} sec".format(t1 - t0))
+print ("\tCCD (HF) Energy: {0}\n\tIterations: {1}\n\tSecond/iteration: {2}".format(energy, iterations, (t1 - t0)/iterations))
 
 t0 = time.time()
 ccd_hf = CoupledClusterDoublesOptimized(

### 2018-06-11 project_2/article/article.tex
Start on review of the theory section

@@
 \begin{abstract}
-    The coupled cluster method for estimating the ground state energy of a many-body
-    system is arguably the most important post-Hartree-Fock method. We have implemented
-    the coupled cluster method with double excitations (CCD). We find that it does not
-    perform well with a simple harmonic oscillator basis, but the results vastly improve 
-    when it is provided a basis from a restricted Hatree-Fock self-consistent field method
-    run as a preliminary computation. For a system consisting of $N=2$ interacting
-    particles, using 12 orbital shells we get a ground stated energy of $3.005970$, very
-    close to the analytical ``true answer'' of $3$\cite{taut1994two}. 
+    The coupled cluster method for estimating the ground state energy of a
+    many-body system is arguably the most important post-Hartree-Fock method. We
+    have implemented the coupled cluster method with double excitations (CCD).
+    We find that it does not perform well with a simple harmonic oscillator
+    basis, but the results vastly improve when it is provided a basis from a
+    restricted Hatree-Fock self-consistent field method run as a preliminary
+    computation. For a system consisting of $N=2$ interacting particles, using
+    12 orbital shells we get a ground stated energy of $3.005970$, very close to
+    the analytical ``true answer'' of $3$\cite{taut1994two}.
 \end{abstract}
@@
     In this project we will compute the ground state energy of quantum dots, a
-    construct that has become a popular area of study in recent decades, both from
-    a theoretical and an experimental point of view. We achieve the ground state
-    computation by implementing a coupled cluster amplitude equation solver
-    which is used in succession with a restricted Hartree-Fock self-consistent
-    field method.
-
-    Regardless of how sophisticated the method used, the many-body problem remains
-    quite intractable. The true wavefunction of any system will be a linear 
-    combination of all possible slater-type orbitals one can model. This means that
-    in order to represent a system correctly, one would need a basis set of infinite
-    size. This is an obvious problem when we solve problems numerically, because we
-    obviously need to limit the size of the basis we use because we have limitations
-    in memeory and computating power. Luckily, great minds have bestowed a wide 
-    selection of very god numerical methods upon the world, which the coupled cluster
-    is a part of.
-
-    First, we provide a thorough theoretical outline of how we model a quantum dot, 
-    the mathematical formalism behind many-body problems and outline how the 
-    computational methods we employ both turn into iterative schemes that are 
-    easy to represent numerically. Second, we provide a brief overview of the
-    program we have constructed which is in the form an installable python package.
-    Third, we have employed the methods on a large parameter space presenting 
-    results for systems consisting of $N= \{2, 6, 12, 20\}$ particles. Fourth,
-    we discuss the most interesting aspects of the results and try to shed some
-    light over the shortcomings as well as the beneficial features of the different
-    metods. Lastly, we end on some summary remarks.
+    construct that has become a popular area of study in recent decades, both
+    from a theoretical and an experimental point of view. We achieve the ground
+    state computation by implementing a coupled cluster amplitude equation
+    solver which is used in succession with a restricted Hartree-Fock
+    self-consistent field method.
+
+    Regardless of how sophisticated the method used, the many-body problem
+    remains quite intractable. The true wavefunction of any system will be a
+    linear combination of all possible slater-type orbitals one can model. This
+    means that in order to represent a system correctly, one would need a basis
+    set of infinite size. This is an obvious problem when we solve problems
+    numerically, because we obviously need to limit the size of the basis we use
+    because we have limitations in memeory and computating power. Luckily, great
+    minds have bestowed a wide selection of very god numerical methods upon the
+    world, which the coupled cluster is a part of.
+
+    First, we provide a thorough theoretical outline of how we model a quantum
+    dot, the mathematical formalism behind many-body problems and outline how
+    the computational methods we employ both turn into iterative schemes that
+    are easy to represent numerically. Second, we provide a brief overview of
+    the program we have constructed which is in the form an installable python
+    package.  Third, we have employed the methods on a large parameter space
+    presenting results for systems consisting of $N= \{2, 6, 12, 20\}$
+    particles. Fourth, we discuss the most interesting aspects of the results
+    and try to shed some light over the shortcomings as well as the beneficial
+    features of the different metods. Lastly, we end on some summary remarks.
 
@@
 
-        The system described above is a theoretical approximation of a
-        quantum dot. Some consideration should be given to an explanation of 
-        why the system described above is a quantum dot, and why we should
-        care about quantum dots.
-        There are at least two things we refer to as quantum dots. The
-        first one is a man-made device, also known as artificial atoms,
-        and the other is the simplest theoretical construct of such a device.
-
-        Aritifical atom quantum dots are semiconducting crystals, typically GaAs,
-        and they are
-        relatively easy to build in a laboratory. Some free electrons are 
-        confined to a volume, either by a physical barrier, like an insulator,
-        or an electromagnetic field\cite{reimann2002electronic}. Due to the
-        confinement of the electrons, the energy levels of a quantum dot becomes
-        quantized, and they therefore hold many properties similar to naturally
-        occuring quantized systems, which is the reason why quantum dots have
-        attracted much interest.
-
-        The theoretical approximation of a quantum dot is impossible to make perfectly
-        accurate, yet simple. There are some important effects that should be
-        included such the potential arising as a result of the confinement, and
-        the interaction between particles. We therefore model our system with an
-        electron-electron interaction, given by the two-body operator in
-        \autoref{eq:two_body_operator}, and with we approximate the confining
-        potential by the harmonic oscillator potential, given by the second term
-        in the one-body Hamiltonian in \autoref{eq:one_body_hamiltonian}. The
-        last term in the one-body Hamiltonian is the kinetic energy of particle
-        in question, essential in any system.
+        The system described above is a theoretical approximation of a quantum
+        dot. Some consideration should be given to an explanation of why the
+        system described above is a quantum dot, and why we should care about
+        quantum dots.  There are at least two things we refer to as quantum
+        dots. The first one is a man-made device, also known as artificial
+        atoms, and the other is the simplest theoretical construct of such a
+        device.
+
+        Aritifical atom quantum dots are semiconducting crystals, typically
+        GaAs, and they are relatively easy to build in a laboratory. Some free
+        electrons are confined to a volume, either by a physical barrier, like
+        an insulator, or an electromagnetic field\cite{reimann2002electronic}.
+        Due to the confinement of the electrons, the energy levels of a quantum
+        dot becomes quantized, and they therefore hold many properties similar
+        to naturally occuring quantized systems, which is the reason why quantum
+        dots have attracted much interest.
+
+        The theoretical approximation of a quantum dot is impossible to make
+        perfectly accurate, yet simple. There are some important effects that
+        should be included such the potential arising as a result of the
+        confinement, and the interaction between particles. We therefore model
+        our system with an electron-electron interaction, given by the two-body
+        operator in \autoref{eq:two_body_operator}, and with we approximate the
+        confining potential by the harmonic oscillator potential, given by the
+        second term in the one-body Hamiltonian in
+        \autoref{eq:one_body_hamiltonian}. The last term in the one-body
+        Hamiltonian is the kinetic energy of particle in question, essential in
+        any system.
 
@@
             &=
-            \sum_{pq}h_{q}^{p}\acr{p}\ade{q}
-            + \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}\acr{p}\acr{q}\ade{s}\ade{r},
-        \end{align}
-        where the sums are general indices over all $L$ basis states and the
-        matrix elements are defined as
+            h_{q}^{p}\acr{p}\ade{q}
+            + \frac{1}{4}u^{pq}_{rs}
+            \acr{p}\acr{q}\ade{s}\ade{r},
+        \end{align}
+        where we have used the Einstein summation convention with repeated
+        indices having an implicit sum. The matrix elements are defined as
         \begin{gather}
             h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
-            \bra{pq}\ket{rs} \equiv \bra{pq}u\ket{rs} - \bra{pq}u\ket{sr}.
+            u^{pq}_{rs} \equiv \bra{pq}u\ket{rs} - \bra{pq}u\ket{sr},
         \end{gather}
-        Note that we use the ``chemists notation'' to label the antisymmetric matrix
+        where the $u$-matrix consists of the the antisymmetric two-body
         elements.
@@
             &=
-            \sum_{ia}t_i^a\acr{a}\ade{i}
-            + \para{\frac{1}{2!}}^2\sum_{ijab}
+            t_i^a\acr{a}\ade{i}
+            + \para{\frac{1}{2!}}^2
             t_{ij}^{ab}\acr{a}\acr{b}\ade{i}\ade{j}
@@
             T \equiv T_2
-            = \frac{1}{4}\sum_{ijab}t_{ij}^{ab}\acr{a}\acr{b}\ade{j}\ade{i}.
+            = \frac{1}{4}t_{ij}^{ab}\acr{a}\acr{b}\ade{j}\ade{i}.
             \label{eq:T_2}
@@
         \end{align}
-        Projecting this equation on the reference state we get
+        Projecting this equation onto the reference state we get
         \begin{align}
@@
             \end{align}
-            As the doubles cluster operator doubly excites the reference and
-            using the expansion in \autoref{eq:sim_norm_hamiltionian_expansion}
-            we see that we can write the energy equation as
+            The doubles cluster operator will excite a pair of single particle
+            functions in the reference state. As the single particle functions
+            are orthonormal we have that
+            \begin{align}
+                \braket{\Phi^{X}}{\Phi_0} = 0,
+            \end{align}
+            where $X$ is an excitation different from zero. This means that for
+            a term in the Baker-Campbell-Hausdorff expansion contribute with a
+            nonzero value to the energy equation it must leave the reference
+            state in its original state. The two-body part of the Hamiltonian is
+            only able to relax a pair of single particle functions in the
+            reference state whereas the doubles cluster operator will excite a
+            pair. This means that we are only left with the following
+            contributions to the energy equation.
             \begin{align}
@@
                 E_0
-                + \bslat H_N \kslat + \bslat (H_N T_2)_c\kslat,
-            \end{align}
-            as the Hamiltonian is only able to relax one pair of single
-            particle functions. By construction we have that
+                + \bslat H_N \kslat + \bslat (H_N T_2)_c\kslat.
+            \end{align}
+            Further, by construction we have that
             \begin{align}
@@
                 \bslat (W_N T_2)_c \kslat
-                &= \frac{1}{4}\sum_{ijab}\bra{ij}\ket{ab}t_{ij}^{ab}.
+                &= \frac{1}{4}u^{ij}_{ab}t^{ab}_{ij}.
             \end{align}
@@
                 \eccd
-                &= \sum_{i} h_i^i + \half \sum_{ij}\bra{ij}\ket{ij}
-                + \frac{1}{4}\sum_{ijab}\bra{ij}\ket{ab}t_{ij}^{ab},
-            \end{align}
-            where the first two terms come from the reference energy as shown in
-            \autoref{eq:reference_energy}.
+                &= h^i_i + \half u^{ij}_{ij}
+                + \frac{1}{4} u^{ij}_{ab}t^{ab}_{ij},
+                \label{eq:ccd_energy}
+            \end{align}
+            where the first two terms come from the reference energy. See
+            Appendix \ref{app:energy_equation} for the calculation of this
+            expression.
 
@@
         excitation in the projection determines the order of the amplitudes you
-        will find. In our case we are only interested in the second order
-        ampltiudes found in the doubles approximation, hence we will solve the
-        equation
-        \begin{align}
-            \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
+        will find. As in the case of the energy equation we use the
+        Baker-Campbell-Hausdorff expansion when expanding the exponentials in
+        the similarity transformed Hamiltonian. We only keep the terms which
+        excites the reference state to the same degree as the amplitudes we are
+        trying to find.
+
+        \subsubsection{Coupled cluster doubles amplitude equations}
+            In our case we are only interested in the second order ampltiudes
+            found in the doubles approximation, hence we will solve the equation
+            \begin{align}
+                \bra{\Phi_{ij}^{ab}}e^{-T_2}H_N e^{T_2}\kslat = 0,
                 \label{eq:doubles_amplitude_1}
@@
             to find an expression that can be used to solve for the amplitudes
-        $t_{ij}^{ab}$.
-        By employing the Baker-Campbell-Haussdorf (BCH) expansion,
-        while setting $T= T_2$, we find
-        \begin{equation}
-                \bar{H}
-                    = \Big(
-                            H_N + H_NT_2
-                            + \frac{1}{2}H_NT_2^2 \Big)_C.
+            $t_{ij}^{ab}$. As the state we are projecting onto is doubly excited
+            we will only keep terms from the Baker-Campbell-Hausdorff expansions
+            which leaves the reference doubly excited.
+            \begin{equation}
+                \bar{H} = \para{
+                    H_N + H_NT_2 + \frac{1}{2}H_NT_2^2
+                }_c.
                 \label{eq:normal_order_expansion}
             \end{equation}
-        The subscript $c$ indicates that only those terms in which the
-        Hamiltonian is connected\footnote{Again, in aWick's theorem sense}
-        to every cluster operator on its right should be included. Since
-        the Hamiltonian contains at most four annihilation and creation
-        operators, $H_N$ can connect to at most four cluster operators
-        at once. Therefore, the BCH expansion must truncate at the
-        fourth-order terms.
-
-        Now comes the rather tedious task of evaluating all the terms
-        that arises from inserting \autoref{eq:normal_order_expansion}
-        into \autoref{eq:doubles_amplitude_1}. This can be done by applying
-        Wick's generalised theorem, but the task is a daunting and streneous one.
-        A few example computations of how this can be done is included in
-        \autoref{app:wickOnAmplitude}.
-         Instead of doing it in this manner, we employ the second quantisation library
-         from SymPy instead\footnote{This is also more in the spirit of this project,
-         as it is within the realm of \emph{Computational} Physics.}.
-        The CCD amplitude equation, from SymPy computation, is
+            Higher order terms will leave the reference in a too high excitation
+            state.
+
+            Now comes the rather tedious task of evaluating all the terms that
+            arises from inserting \autoref{eq:normal_order_expansion} into
+            \autoref{eq:doubles_amplitude_1}. This can be done by applying
+            Wick's generalised theorem, but the task is a daunting and streneous
+            one. A few example computations of how this can be done is included
+            in Appendix \ref{app:wick_on_amplitude}. Instead of doing it in this
+            manner, we employ the second quantization library from SymPy
+            \footnote{This is also more in the spirit of this project, as
+            it is within the realm of \emph{Computational} Physics.}.  The CCD
+            amplitude equation, from SymPy\footnote{Note that this is one of
+            many ways to write the amplitude equation as the ordering of the
+            indices can change drastically for every run of the SymPy-script.},
+            is
             \begin{equation}
                 \begin{aligned}
-            0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
-                 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
-                 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
-                 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
-                 -& \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
+                    0 &= u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab)
+                    - f^k_jt^{ab}_{ik}P(ij)
+                    \\
+                    &\qquad
+                    + \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd}
+                    + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd}
+                    \\
+                    &\qquad
+                    + \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij)
+                    - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab)
+                    \\
+                    &\qquad
+                    + t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij)
+                    + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij)
+                    \\
+                    &\qquad
+                    - \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
                     \label{eq:CCD_amp_1}
@@
             \end{equation}
+            Here $P(ij)$ is an exchange operator which interchanges two
+            particles with indices $i$ and $j$.
 
         \subsubsection{The Iterative Scheme}
-        At first, we pick only diagonal elements of $f$ to be part of the unperturbed
-        Hamiltonian and consider the rest of the terms a perturbation.
-        The second and third terms in \autoref{eq:CCD_amp_1} can now be
-        rewritten,
+            In order to find the amplitude $t$ we have to start with an initial
+            guess and use an iterative procedure to improve on the initial
+            guess. We start by picking the diagonal elements of $f$ to be a part
+            of the unperturbed Hamiltonian and consider the rest of the terms a
+            perturbation. The second and third terms in \autoref{eq:CCD_amp_1}
+            can now be rewritten,
             \begin{equation}
@@
                 f^b_c&t_{ij}^{ab}P(ab) - f^k_j t_{ik}^{ab}P(ij) \\
-                &\to f^b_bt_{ij}^{ab} - f^a_at_{ij}^{ba} - f^j_jt_{ij}^{ab} + f^i_it_{ji}^{ab} \\
+                    &\to f^b_bt_{ij}^{ab} - f^a_at_{ij}^{ba}
+                    - f^j_jt_{ij}^{ab} + f^i_it_{ji}^{ab} \\
                     &= (f^a_a + f^b_b - f^i_i - f^j_j)t_{ij}^{ab} \\
-                &= (\epsilon_a + \epsilon_b - \epsilon_i - \epsilon_j)t_{ij}^{ab} \\
-                &= -(\epsilon_i + \epsilon_j - \epsilon_a - \epsilon_b)t_{ij}^{ab} \\
-                &= -D_{ij}^{ab}t_{ij}^{ab}.
+                    &= (\epsilon_a + \epsilon_b - \epsilon_i
+                    - \epsilon_j)t_{ij}^{ab} \\
+                    &= -(\epsilon_i + \epsilon_j - \epsilon_a
+                    - \epsilon_b)t_{ij}^{ab} \\
+                    &= -D_{ij}^{ab}t_{ij}^{ab},
             \end{aligned}
             \end{equation}
-
-        Then we can define a new version of \autoref{eq:CCD_amp_1},
-        \begin{equation}
-        \begin{aligned}
-            g(u, t) =&
-                    u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab) - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
-                 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
-                 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
-                 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
-                 -& \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
-        \end{aligned}
-        \end{equation}
-
-        where $\tilde{f}$ are the non-diagonal parts of the Fock matrix.
-
-        Now we have $D_{ij}^{ab}t_{ij}^{ab} = g(u, t)$.
+            where the arrow signifies that we only look at the diagonal elements
+            of $f$. We now define the right hand side of \autoref{eq:CCD_amp_1}
+            to $g(f, u, t)$ where we make sure that we only use the off-diagonal
+            elements of $f$ in the computations.
+
+            By moving $D^{ab}_{ij}t^{ab}_{ij}$ to the left hand side we we get
+            \begin{align}
+                D_{ij}^{ab}t_{ij}^{ab} = g(u, t).
+            \end{align}
             This allows us to define an iterative scheme,
@@
             \includegraphics[width=244px]
-            {../dat/figures/iter_run_times_zoom.pdf}
+            {../dat/figures/iter_run_times_zoom_6.pdf}
             \caption{Run time for $N=6$ particles, only the faster methods are
@@
 
-\section{Amplitude Equations}
-
-    \label{app:wickOnAmplitude}
+\section{Energy equation}
+    \label{app:energy_equation}
+    Here we show how to compute the coupled cluster doubles energy from
+    \autoref{eq:ccd_energy}. The first two terms come from the reference energy
+    shown in \autoref{eq:reference_energy}. The last term we get by using Wick's
+    theorem.
+    \begin{align}
+        \bslat (H_N T_2)_c\kslat
+        &= \bslat (W_N T_2)_c\kslat,
+    \end{align}
+    as the normal ordered Fock operator can at most relax a single state, i.e.,
+    \begin{align}
+        \bslat (F_N T_2)_c\kslat = 0.
+    \end{align}
+    We are thus left with computing the contraction of the doubles cluster
+    operator and the normal ordered two-body operator. Note that we must keep
+    only the fully contracted terms.
+    \begin{align}
+        \bslat (W_N T_2)_c\kslat
+        &=
+        \frac{1}{16} u^{pq}_{rs} t^{ab}_{ij}
+        \bslat
+        A
+        \kslat,
+        \label{eq:third_term_ccd_energy}
+    \end{align}
+    where $A$ is the create and destruction operator strings. By using
+    generalized Wick's theorem we get
+    \begin{align}
+        A &=
+        \{\acr{p}\acr{q}\ade{s}\ade{r}\}
+        \{\acr{a}\acr{b}\ade{j}\ade{i}\}
+        \nonumber \\
+        &=
+        \{
+            \wick{
+                    \c4 a_p^{\dagger} \c3 a_q^{\dagger}
+                    \c2 a_s \c1 a_r
+                    \c1 a_a^{\dagger} \c2 a_b^{\dagger}
+                    \c3 a_j \c4 a_i
+            }
+        \}
+        +
+        \{
+            \wick{
+                    \c4 a_p^{\dagger} \c3 a_q^{\dagger}
+                    \c1 a_s \c2 a_r
+                    \c1 a_a^{\dagger} \c2 a_b^{\dagger}
+                    \c3 a_j \c4 a_i
+            }
+        \}
+        \nonumber \\
+        &\qquad
+        +
+        \{
+            \wick{
+                    \c3 a_p^{\dagger} \c4 a_q^{\dagger}
+                    \c2 a_s \c1 a_r
+                    \c1 a_a^{\dagger} \c2 a_b^{\dagger}
+                    \c3 a_j \c4 a_i
+            }
+        \}
+        +
+        \{
+            \wick{
+                    \c3 a_p^{\dagger} \c4 a_q^{\dagger}
+                    \c1 a_s \c2 a_r
+                    \c1 a_a^{\dagger} \c2 a_b^{\dagger}
+                    \c3 a_j \c4 a_i
+            }
+        \}
+        \nonumber \\
+        &=
+        \delta_{pi}\delta_{qj}\delta_{sb}\delta_{ra}
+        - \delta_{pi}\delta_{qj}\delta_{sa}\delta_{rb}
+        \nonumber \\
+        &\qquad
+        - \delta_{pj}\delta_{qi}\delta_{sb}\delta_{ra}
+        + \delta_{pj}\delta_{qi}\delta_{sa}\delta_{rb}.
+    \end{align}
+    Inserted back into \autoref{eq:third_term_ccd_energy} and summing the
+    indices in the Kronecker-Delta's we get
+    \begin{align}
+        \bslat (W_N T_2)_c\kslat
+        &=
+        \frac{1}{16}\para{
+            u^{ij}_{ab} - u^{ij}_{ba} - u^{ji}_{ab} + u^{ji}_{ba}
+        }
+        t^{ab}_{ij}
+        \nonumber \\
+        &=
+        \frac{1}{4}u^{ij}_{ab}t^{ab}_{ij},
+    \end{align}
+    where we have used the antisymmetry of the two-body matrix elements, thus
+    completing what we wanted to show.
+
+\section{Amplitude equations}
+
+    \label{app:wick_on_amplitude}
 

### 2018-06-11 project_2/coupled-cluster/coupled_cluster/hartree_fock/scf_rhf.py
Use same convention for indices as in article

@@ -2,14 +2,10 @@ import numpy as np
 import scipy.linalg
 
 def calculate_energy(F, W, D):
-    J = np.einsum("sr, qrps -> qp", D, W)
-    K = np.einsum("sr, qrsp -> qp", D, W)
+    J = np.einsum("prqs, sr -> pq", W, D)
+    K = np.einsum("prsq, sr -> pq", W, D)
 
-    energy = np.einsum("pq, qp ->", F, D)
-    energy -= 0.5*np.einsum("pq, qp ->", J, D)
-    energy += 0.25*np.einsum("pq, qp ->", K, D)
-
-    return energy
+    return np.einsum("pq, qp -> ", F - 0.5 * J + 0.25 * K, D)
 
 def build_density_matrix(U, num_occupied):
     """Function building a density matrix.
@@ -41,11 +37,11 @@ def build_rhf_fock_matrix(H, W, D):
 
         F_{pq} = H_{pq} + J(D)_{pq} - 0.5*K(D)_{pq},
 
-    where we build J and K from the electron repulsion integrals (W) (using
-    Mulliken notation) and the density matrix D.
+    where we build J and K from the electron repulsion integrals (W) and the
+    density matrix D.
 
-        J(D)_{pq} = (pq|rs)*D_{sr} = W[p][q][r][s]*D[s][r],
-        K(D)_{pq} = (ps|rq)*D_{sr} = W[p][s][r][q]*D[s][r].
+        J(D)_{pq} = <pr|u|qs> * D_{sr} = W[p, q, r, s] * D[s, r],
+        K(D)_{pq} = <pr|u|sq> * D_{sr} = W[p, s, r, q] * D[s, r].
 
     The NumPy-function einsum does this in a form which resemble the
     mathematical-product.
@@ -60,8 +56,8 @@ def build_rhf_fock_matrix(H, W, D):
     Returns:
         np.ndarray: The RHF-Fock matrix.
     """
-    J = np.einsum("sr, qrps -> qp", D, W)
-    K = np.einsum("sr, qrsp -> qp", D, W)
+    J = np.einsum("prqs, sr -> pq", W, D)
+    K = np.einsum("prsq, sr -> pq", W, D)
 
     return H + J - 0.5*K
 

### 2018-06-11 project_2/article/article.tex
Review intermediates and part of RHF

@@
 
-        \subsubsection{Intermediate Computations}
-        Looking closely at the amplitude equation in (\ref{eq:CCD_amp_1}) one might come
-        to realize that this equation contains many of the same structures in
-        several of the terms. This warrants the search for an algebraic transformation of the
-        CCD amplitude equation that has the potential to reduce the amount of floating point
-        operations needed to compute it. As it turns out, such terms exist and they will decrease
-        the computing time necessary by an order of magnitude. We will define the following
-        "intermediates",
+        \subsubsection{Intermediate computations}
+            Looking closely at the tensors in \autoref{eq:CCD_amp_1} one might
+            come to realize that many of the terms share a common structure.
+            This warrants the search for an algebraic transformation of the
+            amplitude equation that has the potential to reduce the amount of
+            floating point operations needed to compute it. As it turns out,
+            such terms exist and they will decrease the computing time necessary
+            by two orders of magnitude for the price of higher storage costs.
+            We will define the following "intermediates"\footnote{Note that are
+            more than one way of defining the intermediates.},
             \begin{align}
                 \label{eq:intermediate1}
-            \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
+                \chi^{ab}_{cd}
+                &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd}
+                + \frac{1}{2}u^{ab}_{cd} \\
                 \label{eq:intermediate2}
@@
                 \label{eq:intermediate4}
-            \chi^{bm}_{jc} &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd}
-        \end{align}
-        These intermediate structures will allow us to rewrite \autoref{eq:CCD_amp_1} to,
+                \chi^{bm}_{jc}
+                &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd}
+            \end{align}
+            See Appendix \ref{app:intermediates} for the derivation.  These
+            intermediate structures will allow us to rewrite
+            \autoref{eq:CCD_amp_1} to,
             \begin{equation}
                 \begin{aligned}
-            0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
-              &+ t^{cd}_{ij}\chi^{ab}_{cd} + t^{ab}_{in}\chi^n_jP(ij)
+                    0 &= u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab)
+                    - f^k_jt^{ab}_{ik}P(ij) \\
+                    &\qquad
+                    + t^{cd}_{ij}\chi^{ab}_{cd} + t^{ab}_{in}\chi^n_jP(ij)
                     - t^{bd}_{ij}\chi^a_dP(ab) \\
-              &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij) + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
+                    &\qquad
+                    + t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
+                    + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
                 \end{aligned}
             \end{equation}
-        The importance of this "trick" will become apparent in due time.
+            We will discuss the importance of doing this in
+            \autoref{subsec:ccd_scaling}.
 
@@
         Having found the equations needed in order to find an estimate to the
-        ground state energy using the coupled cluster doubles approximation is a
-        well and dandy. But, we need basis functions to create the matrix
+        ground state energy using the coupled cluster doubles approximation is
+        all well and dandy. But, we need basis functions to create the matrix
         elements needed to feed into the coupled cluster code.  Often these
@@
             a Coulomb repulsion.  If we assume, or make it so, that the
-            repulsive two-body part is small we can use eigenfunctions of the
-            one-body part our basis. In this case we have two-dimensional
-            harmonic oscillator functions as eigenfunctions. We can then compute
-            the matrix elements, $h_q^p$ and $u_{rs}^{pq}$, before feeding these
-            into the coupled cluster code.
+            repulsive two-body part is small we can use the eigenfunctions of
+            the one-body part as our basis of single particle functions. In this
+            case we have two-dimensional harmonic oscillator functions as
+            eigenfunctions. We can then pre-compute the matrix elements, $h_q^p$
+            and $u_{rs}^{pq}$, before feeding these into the coupled cluster
+            code.
 
@@
             fermions this means that each mode of the harmonic oscillator
-            functions will be repeated twice.}.
+            function will be repeated twice.}.
             \begin{align}
@@
             the matrices $h$ and $u$ use single indices for each wavefunction.
-            In \autoref{fig:basis_states} we can see the energy levels that
-            needs to be mapped.
+            In \autoref{fig:basis_states} we can see the lowest three energy
+            levels that needs to be mapped.
             \begin{figure}
@@
                 Each arrow representes a spin up or a spin down state with the
-                quantum numbers $n$ and $m$ as listed below.}
+                quantum numbers $n$ and $m$ as listed below. This pattern goes
+                on indefinitly with the addition of one bar (two oscillators)
+                per level.}
                 \label{fig:basis_states}
@@
             all spin states for each energy level and the energy levels below.
-            In \autoref{fig:basis_states} we can see the magic number $N \in [2,
-            6, 12]$.
+            In \autoref{fig:basis_states} we can see the magic numbers $N \in [2,
+            6, 12]$ by counting the levels in increasing order.
 
@@
             \begin{align}
-                \bra{\alpha\beta}\ket{\gamma\delta}
-                &= \braket{\alpha\beta}{\gamma\delta}
-                - \braket{\alpha\beta}{\delta\gamma}.
-            \end{align}
-            As we only need to compute these once it is a good idea to save all
-            non-zero values of $\bra{\alpha\beta}\ket{\gamma\delta}$ to a file.
+                u^{\alpha\beta}_{\gamma\delta}
+                &= \bra{\alpha\beta}u\ket{\gamma\delta}
+                - \bra{\alpha\beta}u\ket{\delta\gamma}.
+            \end{align}
+            Computing the Coulomb elements is an extremely costly operation and
+            we will thus only compute the non-zero orbital integrals, i.e., neglecting
+            spin and antisymmetry, before saving these to a file which we can
+            read several times.
 
@@
             Having found the matrix elements of $h$ and $u$ we can now use the
-            \emph{self-consistent field} (SCF) iteration method to construct $h$ and
-            $u$ in a \emph{restricted Hartree-Fock} (RHF) basis. This will yield a
-            better estimate to the actual, unknown basis functions of the
-            system.
-
-            The neatest, yet arguably the most abstract, way to write the Hartree-Fock
-            equations for electron $i$\footnote{This is weird, as electron
-            should be indistinguishable and therefore impossible to label.} is
-            \begin{equation}
-                f_i \varphi_i = \varepsilon_i \varphi_i,
-            \end{equation}
-            where $f_i$ is the Fock operator, $\varphi_i$ are eigenstates of the Fock
-            operator consiting of a set of one-electorn wave funcitons called the
-            Hartree-Fock molecular orbitals, and $\varepsilon_i$ are the eigenenergies
-            of the Fock operator. The fock operator, in matrix notation, is given by
-            \begin{equation}
-                F_{pa} = H_{pq} +  J(D)_{pq} - \frac{1}{2}K(D)_{pq},
-            \end{equation}
-            where $h_i$ is the one-body operatore, while the two-body operator is
-            divided into what we call a direct part,
-            \begin{equation}
-                J(D)_{pq} = \braket{pq}{rs}D_{sr},
+            \emph{self-consistent field} (SCF) iteration method to construct $h$
+            and $u$ in a \emph{restricted Hartree-Fock} (RHF) basis. This will
+            yield a better estimate to the actual, unknown basis functions of
+            the system.
+
+            The neatest, yet arguably the most abstract, way to write the
+            Hartree-Fock equations for electron $i$ is
+            \begin{equation}
+                f \varphi_p = \varepsilon_p \varphi_p,
+            \end{equation}
+            where $f$ is the Fock operator, $\varphi_p$ are eigenstates of the
+            Fock operator consisting of a set of one-electron wave functions
+            called the Hartree-Fock molecular orbitals, and $\varepsilon_p$ are
+            the eigenenergies of the Fock operator. The Fock operator, in matrix
+            notation, is given by
+            \begin{equation}
+                f^p_q = h^p_q + J(D)^p_q - \half K(D)^p_q,
+                %F_{pq} = H_{pq} +  J(D)_{pq} - \frac{1}{2}K(D)_{pq},
+            \end{equation}
+            where $h^p_q$ is the one-body operator, while the two-body operator
+            is divided into what we call a direct part,
+            \begin{equation}
+                J(D)^p_q = \bra{pr}u\ket{qs}D_{sr},
             \end{equation}
@@
             \begin{equation}
-                K(D)_{pq} = \braket{ps}{rq}D_{sr}.
-            \end{equation}
-            The direct part of the two-body operator is comparable to classical Coloumb
-            repulsion, while the exchange does not have a classical analog as it arises
-            from the antisymmetry requirement of the wavefunction. $D_{sr}$ is the
-            density matrix of the system.
-
-            Introducing a basis set transforms the Hartree-Fock equations into the Rothaan
-            equations
+                K(D)^p_q = \bra{pr}u\ket{sq}D_{sr}.
+            \end{equation}
+            Pay special attention to the non-antisymmetric two-body elements in
+            the direct and exchange part.  The direct part of the two-body
+            operator is comparable to classical Coloumb repulsion, while the
+            exchange does not have a classical analog as it arises from the
+            antisymmetry requirement of the wavefunction. The matrix $D_{sr}$ is
+            the density matrix of the system and will be defined shortly.
+
+            Introducing a basis set transforms the Hartree-Fock equations into
+            the Rothaan equations
             \begin{equation}
@@
             \end{equation}
-            This is a generalised eigenvalue problem where $S$ serves as an overlap matrix
-            that must be there in case of non-orthogonal basis. Since the Fock matrix $F$
-            depends on it's own solution through the orbitals, the eigenvalue problem must
-            be solved iteratively\footnote{This is also the reason why the
-            Hartree-Fock-Roothaan equations are often called the self-consistent-field
-            procedure.}
+            This is a generalised eigenvalue problem where
+            \begin{align}
+                S^{p}_{q} = \braket{p}{q},
+            \end{align}
+            i.e., the \emph{overlap matrix}. We now define the density matrix
+            $D$ by
+            \begin{align}
+                D^{p}_{q} \equiv C^p_{\alpha}{C^{\alpha}_q}^{*},
+            \end{align}
+            Since the Fock matrix $F$ depends on it's own solution through the
+            orbitals, the eigenvalue problem must be solved
+            iteratively\footnote{This is also the reason why the Roothan-Hall
+            equations are often called the self-consistent-field procedure.}
 
             Because the RHF SCF method is a variational method, the initial
-            wavefunction made up of the single-particle wavefunctions $\varphi_i$
-            is varied until the minimal energy is found. The transformations necessary
-            to make all these variations will now be contained in $C$. So if we start from
-            a harmonic oscillator (HO) basis we can transform into a better basis, called a
-            Hartree-Fock (HF) basis,
-            \begin{equation}
-                \ket{\Psi}^{\text{HF}}_p = \sum_{k=1}^L C_{kp} \ket{\Psi}_k^{\text{HO}}.
-            \end{equation}
-            We will see how the coupled cluster method with double excitations works with
-            both the HO basis and the HF basis, but not equally well.
+            wavefunction made up of the single-particle wavefunctions
+            $\varphi_i$ is varied until the minimal energy is found.
+            \begin{align}
+                E_{\text{RHF}}
+                &= \para{
+                    f^p_q - \half J(D)^p_q + \frac{1}{4}K(D)^p_q
+                }D^q_p.
+            \end{align}
+
+            The transformations necessary to make all these variations will now
+            be contained in $C$. So if we start from a harmonic oscillator (HO)
+            basis, $\ket{\alpha}$ we can transform into a better basis, called a
+            Hartree-Fock (HF) basis, $\ket{p}$.
+            \begin{equation}
+                \ket{p} = C^p_\alpha\ket{\alpha}.
+            \end{equation}
+            We will see how the coupled cluster method with double excitations
+            works with both the HO basis and the HF basis, but not equally well.
 
@@
         compialtion of some functions.
+
+    \subsection{Scaling of the coupled cluster doubles method}
+        \label{subsec:ccd_scaling}
 
@@
 \section{Finding Amplitude Equation Intermediates}
+    \label{app:intermediates}
     Starting from the CCD amplitude equation we factor out terms that will have the same

### 2018-06-13 project_2/article/article.tex
Elaborate on the HF-method
Link to example script on Github

@@
                 f^p_q = h^p_q + J(D)^p_q - \half K(D)^p_q,
-                %F_{pq} = H_{pq} +  J(D)_{pq} - \frac{1}{2}K(D)_{pq},
             \end{equation}
@@
             \begin{equation}
-                FC = SC\varepsilon.
-            \end{equation}
-            This is a generalised eigenvalue problem where
+                FC = SC\varepsilon,
+            \end{equation}
+            where $F$ is the Fock matrix constructed from $f^p_q$.  This is a
+            generalised eigenvalue problem where
             \begin{align}
@@
             \begin{equation}
-                \ket{p} = C^p_\alpha\ket{\alpha}.
+                \ket{p}
+                = \para{\ket{\alpha}\bra{\alpha}}\ket{p}
+                = C^p_\alpha\ket{\alpha}.
             \end{equation}
@@
 \section{Implementation}
-    We have costructed a flexible framework for making performing Hartree-Fock
-    self-consistednt field (SCF) and coupled cluster with double excitations (CCD)
-    computations. The entire code base consists of package for \lstinline{python}
-    that is easy to install globally on any computer.
+    We have costructed a flexible framework for performing Hartree-Fock
+    self-consistent field iterations and coupled cluster with double excitations
+    computations. The entire code base consists of package for
+    \lstinline{Python} that is easy to install globally on any computer.
 
     Because we have employed a mixture of Python and C++ with Cython interfaces,
-    the code sturcture may be difficult to undertand for someone without Cython
+    the code sturcture may be difficult to understand for someone without Cython
     experience, but we have tried to make the actual usage of the software as
@@
 
-    Most of the computations that are done in a coupled cluster algorithm
-    is linear algebra and we found that NumPy, relying on
-    BLAS and LAPACK, is most likely as fast as it gets and, at the very least,
-    efficient enough for our purposes. The quicker and easier scripting capabilities
-    Python provides is worth what little extra speed-up writing the entire
-    code base in C++ could have provided.
+    Most of the computations that are done in a coupled cluster algorithm is
+    linear algebra and we found that NumPy, relying on BLAS and LAPACK, is quite
+    fast and at the very least, efficient enough for our purposes. The quicker
+    and easier scripting capabilities Python provides is worth what little extra
+    speed-up writing the entire code base in C++ could have provided.
 
@@
         Then all you need to do is change directory to the project code folder
-        where a \lstinline{MakeFile} is included so you only need to build and
+        where a \lstinline{Makefile} is included so you only need to build and
         install,
@@
         \end{bash}
-        Now everything should be able to run from anywhere on your computer.
-        To ensure that all requirements of the program are satified you can
-        install with \mintinline{bash}{make installr} instead.
-
-        We have included a sample python script in \autoref{app:sample_script}
-        that uses almost all methods and classes in the package, and also compares
-        the time each methods takes to fininsh the computations.
+        Now everything should be able to run from anywhere on your computer.  To
+        ensure that all requirements of the program are satisfied you can
+        install with \mintinline{bash}{make installr} instead. Running
+        \mintinline{bash}{py.test} on the command-line in the
+        top-directory\footnote{That is, in
+        \texttt{FYS4411/project\_2/coupled-cluster}.} will run the test scripts
+        located in the \texttt{tests}-directory to make sure that the
+        installation passed successfully.
+
+        The script that is used to generate most of the data is located in
+        \url{https://github.com/Schoyen/FYS4411/blob/master/project_2/scripts/comparison.py}
+        and provides a good example on how to use the code base.
 
@@
 
-\vfill
+%\vfill
 \clearpage
-\newpage
-\begin{widetext}
-\section{Example Script}
-
-    \label{app:sample_script}
-
-    \vspace{10pt}
-
-    \inputminted[firstline=0, lastline=113]{python}{../scripts/comparison.py}
-
-\end{widetext}
+%\newpage
+%\begin{widetext}
+%\section{Example Script}
+%
+%    \label{app:sample_script}
+%
+%    \vspace{10pt}
+%
+%    \inputminted[firstline=0, lastline=113]{python}{../scripts/comparison.py}
+%
+%\end{widetext}
 

### 2018-06-13 project_2/article/article.tex
Add more discussion in the implementation section

@@
 % SI-units
-\usepackage{siunitx}
+\usepackage[binary-units]{siunitx}
 % Enhanced math formatting
@@
     states with $a, b, c, d, \dots \in \{N + 1, \dots, L\}$ and general indices
-    with $p, q, r, s, \dots \in \{1, \dots, L\}$. In terms of sets of basis
-    functions we can write this as
+    with $p, q, r, s, \dots \in \{1, \dots, L\}$. We define $M = L - N$ as the
+    number of virtual states. In terms of sets of basis functions we can write
+    this as
     \begin{align}
@@
             \end{equation}
-            We will discuss the importance of doing this in
-            \autoref{subsec:ccd_scaling}.
+            Doing this reduces the number of FLOPS from $\mathcal{O}(M^4N^4)$
+            for the heaviest contraction, i.e., the contraction
+            \begin{align}
+                g(f, u, t) \gets
+                \frac{1}{4}t^{cd}_{ij}t^{ab}_{nm}u^{mn}_{cd},
+            \end{align}
+            to $\mathcal{O}(M^4N^2)$ when using the largest intermediate to do
+            the contraction
+            \begin{align}
+                g(f, u, t) \gets t^{cd}_{ij}\chi^{ab}_{cd},
+            \end{align}
+            for the price of storing $M^4$ more matrix elements and
+            pre-computing the intermediate.
 
@@
                 = \para{\ket{\alpha}\bra{\alpha}}\ket{p}
-                = C^p_\alpha\ket{\alpha}.
+                = C^{\alpha}_p\ket{\alpha}.
             \end{equation}
@@
     \subsection{Program Structure}
-        There are three main subsetions within our program structure,
-        given in the directory tree below.
+        There are three main subsections within our program structure, given in
+        the directory tree below.
 
@@
 
-        The first subsection, \mintinline{python}{matrix_elemtents},
+        The first subsection, \mintinline{bash}{matrix_elemtents},
         contains methods for computing matrix elements in harmonic
@@
         elements is a very intensive task, and the central functions are
-        therefore implemented in C++ with a Cython interface to enable use
-        in Python.
-
-        The \mintinline{python}{hartree_fock} subsection contains methods
+        therefore implemented in C++ with a Cython interface to enable use in
+        Python\footnote{The C++ code is inherited from Alocias Mariadason from
+        his project in FYS4411 last year with slight modifications. The original
+        source code can be found here
+        \url{https://github.com/Oo1Insane1oO/FYS4411/blob/master/project1/src/Coulomb_Functions.cpp}
+        and
+        \url{https://github.com/Oo1Insane1oO/FYS4411/blob/master/project1/src/Coulomb_Functions.hpp}.
+        }.
+
+        The \mintinline{bash}{hartree_fock} subsection contains methods
         for changing from harmonic oscillator basis to Hartree-Fock basis as
@@
 
-        The \mintinline{python}{schemes} subsection is arguably the most
-        important part of this project. The subsection has three different
-        classes that perform the exact same computations, but in different
-        and increasingly intelligent ways. First,
-        \mintinline{python}{CoupledClusterDoubles} is the most
-        straightforward and naïve way to solve the CCD amplitude equations.
-        Second, because an overbearing amount of the elements in the
-        operator matrices in this problem are zero, we have implemented a
-        sparse matrix CDD solver in
-        \mintinline{python}{CoupledClusterDoublesSparse}. Third,
-        \mintinline{python}{CoupledClusterDoublesOptimized} takes advantage
-        of sparse matrices as well, but is also parallellized and optimized
-        with memory use and number of floating point operations in mind.
-
+        The \mintinline{bash}{schemes} subsection is arguably the most important
+        part of this project. The subsection has three different classes that
+        perform the exact same computations, but in different and increasingly
+        intelligent ways. First, \mintinline{python}{CoupledClusterDoubles} is
+        the most straightforward and naïve way to solve the CCD amplitude
+        equations.  Second, because an overbearing amount of the elements in the
+        operator matrices in this problem are zero, we have implemented a sparse
+        matrix CCD solver in \mintinline{python}{CoupledClusterDoublesSparse}.
+        Third, \mintinline{python}{CoupledClusterDoublesOptimized} is
+        parallellized and optimized with memory use and number of floating point
+        operations in mind.
+
+    \subsection{Index mapping}
+        The index mapping generating a $(n, m) \mapsto p$ and $p \mapsto (n, m)$
+        is a Python adaption of the code found here:
+        \url{https://github.com/ManyBodyPhysics/LectureNotesPhysics/blob/master/Programs/Chapter8-programs/cpp/CCD/src/qDots2DSPBasis.cpp#L18-L68}.
+
+    \subsection{Generating the antisymmetric two-body elements}
+        Constructing the Coulomb elements is by far the most intensive task
+        that is done. Luckily, we only need to do this once for each basis size
+        as the elements do not change and can be reused. We therefore compute
+        the elements once and store all the non-zero elements as Python
+        pickle-objects. If we wish to do a new run we first look for the
+        relevant Coulomb elements in the \texttt{dat}-directory before
+        attempting to create them anew.
+
+        After the elements are created/read we have to scale them with the
+        current frequency, i.e.,
+        \begin{align}
+            \bra{\alpha\beta}u\ket{\gamma\delta}
+            = \sqrt{\omega}\bra{12}u\ket{34},
+        \end{align}
+        where $\bra{12}u\ket{34}$ is found from E. Anisimovas and A.  Matulis
+        (Equation A2)\cite{anisimovas1998energy}. Next we antisymmetrize the
+        two-body elements and include spin which doubles each dimension of the
+        orbital integrals.
+        \begin{align}
+            u^{\alpha\beta}_{\gamma\delta}
+            &=
+            \delta_{\sigma_{\alpha}\sigma_{\gamma}}
+            \delta_{\sigma_{\beta}\sigma_{\delta}}
+            \bra{\alpha\beta}u\ket{\gamma\delta}
+            \nonumber \\
+            &\qquad
+            -
+            \delta_{\sigma_{\alpha}\sigma_{\delta}}
+            \delta_{\sigma_{\beta}\sigma_{\gamma}}
+            \bra{\alpha\beta}u\ket{\delta\gamma}.
+        \end{align}
+        The antisymmetrization process uses four nested loops and is therefore
+        implemented in Cython to reduce the computation time.
+
+    \subsection{Reducing the complexity of the basis transformation}
+        Having computed the RHF energy and the basis transformation matrix $C$
+        we can transform from the HO basis to the HF basis. For the one-body
+        matrix elements this becomes
+        \begin{align}
+            \bra{p}h\ket{q}
+            &= {C^p_{\alpha}}^{\dagger}C^{\beta}_q\bra{\alpha}h\ket{\beta}.
+        \end{align}
+        Note the usage of the Hermitian conjugate and complex conjugate. We
+        define them in such a manner
+        \begin{align}
+            {C^p_{\alpha}}^{\dagger}
+            = {C^{\alpha}_p}^{*}.
+        \end{align}
+        For the two-body elements this is a more complicated
+        task\footnote{Imagine the poor fellows that needs to deal with
+        three-body elements...}.
+        \begin{align}
+            \bra{pq}u\ket{rs}
+            &=
+            {C^p_{\alpha}}^{\dagger}{C^q_{\beta}}^{\dagger}
+            C^{\gamma}_r C^{\delta}_s
+            \bra{\alpha\beta}u\ket{\gamma\delta},
+        \end{align}
+        which translates into eight nested loops when programming the
+        expression as it stands. Luckily, by noticing that no indices are
+        repeated more than twice we can do each loop separately viz.
+        \begin{align}
+            \bra{pq}u\ket{rs}
+            &=
+            {C^p_{\alpha}}^{\dagger}
+            \brak{
+                {C^q_{\beta}}^{\dagger}
+                \brak{
+                    C^{\gamma}_r
+                    \brak{
+                        C^{\delta}_s
+                        \bra{\alpha\beta}u\ket{\gamma\delta}
+                    }
+                }
+            }.
+        \end{align}
+        This reduces the computations to four quintuply nested
+        loops\footnote{This trick was made aware to us through
+        \url{http://sirius.chem.vt.edu/wiki/doku.php?id=crawdad:programming:project4}.
+        }.
+
+    \subsection{Implementation details}
         Both in \mintinline{python}{CoupledClusterDoublesSparse} and
-        \mintinline{python}{CoupledClusterDoublesOptimized} we make use
-        of the intermediates in equations \ref{eq:intermediate1} to
-        \ref{eq:intermediate4}. We have also relied upon implementations
-        in C++ for some of the heaviest computations. In the optimized
-        class special care was taken to ensure that every contraction
-        was computed in an optimal way, with the minimization of memory
-        use and CPU time as the overall goal. Moreover, several methods
-        in NumPy for doing linear algebra was tested, including but not limited
-        to \mintinline{python}{numpy.einsum()}, \mintinline{python}{numpy.tensordot},
-        and reshaping datastructures to two dimensions and using
-        \mintinline{python}{numpy.dot}. Additionally, we tried to make use of
-        the python package \mintinline{python}{dask} which allows for arrays that
-        are larger than the computer's memory buffer as well as automated
-        concurrency. We found that the dask arrays was a bad fit for our code
-        base, but we ended up using \mintinline{python}{numba} for automated
-        paralellization of some iterative loops as well as just-in-time
-        compialtion of some functions.
+        \mintinline{python}{CoupledClusterDoublesOptimized} we make use of the
+        intermediates in equations \ref{eq:intermediate1} to
+        \ref{eq:intermediate4}.  In the optimized class special care was taken
+        to ensure that every contraction was computed in an optimal way, with
+        the minimization of memory use and CPU time as the overall goal.
+        Moreover, several methods in NumPy for doing linear algebra was tested,
+        including but not limited to \mintinline{python}{numpy.einsum},
+        \mintinline{python}{numpy.tensordot}, and reshaping datastructures to
+        two dimensions and using \mintinline{python}{numpy.matmul}.
+        Additionally, we tried to make use of the python package
+        \mintinline{python}{dask} which allows for arrays that are larger than
+        the computer's memory buffer as well as automated concurrency. We found
+        that the dask arrays was a bad fit for our code base, but we ended up
+        using \mintinline{python}{numba} for paralellization of some
+        iterative loops as well as just-in-time compilation of some functions.
 
     \subsection{Scaling of the coupled cluster doubles method}
-        \label{subsec:ccd_scaling}
+        When talking about the scaling of the coupled cluster method there are
+        two aspects to consider.
+        \begin{enumerate}
+            \item The number of FLOPS to compute $g(f, u, t)$.
+            \item The number of bytes needed to store the arrays of matrix
+                elements.
+        \end{enumerate}
+        For the CCD method the naïve solution using the form of $g(f, u, t)$
+        without the intermediates results in $\mathcal{O}(M^4N^4)$ FLOPS each
+        iteration. Using the form of the intermediates we have set up this
+        reduces to $\mathcal{O}(M^4N^2)$, i.e., two order of magnitude less. The
+        effect of this is \emph{dramatic} and will be shown in the results
+        section in due time.
+
+        A downside of using the intermediates\footnote{Particularly our form for
+        the intermediates as we store $\chi^{ab}_{cd}$ instead of a rewriting
+        this intermediate as $\chi^{ij}_{kl}$ which is also possible.} is the
+        higher storage costs. Lucky enough for us, the largest storage cost
+        comes from storing $u^{pq}_{rs}$ and overshadows all other terms as it
+        scales as $L^4$. See \autoref{tab:memory_u} for some example storage
+        costs. For a normal desktop computer we quickly reach a threshold where
+        $u^{pq}_{rs}$ will eat up all memory and activate disk storage thus
+        triggering a massive slowdown in computational speed.
+
+        \begin{table}
+            \centering
+            \caption{Storage costs of $u^{pq}_{rs}$ in gigabytes for a given
+            shell $R$ and the corresponding number of basis functions $L$
+            assuming $\SI{8}{\byte}$ (\texttt{double}) precision.}
+            \begin{ruledtabular}
+                \begin{tabular}{ccc}
+                    $R$ & $L$ & $u^{pq}_{rs}$ $[\si{\giga\byte}]$ \\
+                    \hline
+                    $8$ & $72$ & $0.2$ \\
+                    $9$ & $90$ & $0.49$ \\
+                    $10$ & $110$ & $1.1$ \\
+                    $11$ & $132$ & $2.3$ \\
+                    $12$ & $156$ & $4.4$ \\
+                    $13$ & $182$ & $8.2$ \\
+                    $14$ & $210$ & $14.5$
+                \end{tabular}
+            \end{ruledtabular}
+            \label{tab:memory_u}
+        \end{table}
+
+        For larger shell values we have to be particularly careful in terms of
+        the memory usage. This can be semi-problematic when using a programming
+        language such as Python as memory usage is camouflaged. Using NumPy
+        allows us to reuse memory buffers quite efficiently, but we have to be
+        careful what type of operations we do and if they incur copying.
+
+        \subsubsection{Sparse implementation}
+            It turns out that most of the elements in both
+            $u^{\alpha\beta}_{\gamma\delta}$ (the HO basis) and $u^{pq}_{rs}$
+            (the HF basis) are zero. Roughly $0.014\%$ and $0.082\%$ of the
+            elements are non-zero respectively (for $R = 12$). This inspired us
+            to look for a way to do sparse CCD. Using the
+            \texttt{sparse}-library in Python\footnote{Found here:
+            \url{https://github.com/pydata/sparse}.} we were able to set up a
+            sparse implementation. Unfortunately being efficient in terms of
+            re-using memory is quite complicated and there are definitely
+            improvements to be made. Nonetheless the performance of the method
+            is comparable to the optimized code.
+
+        \subsubsection{Parallelizing the contractions}
+            The easiest way to parallelize the coupled cluster code is to use
+            OpenMP on the contractions. In particular the heavy contractions. We
+            solved this using Numba with the \texttt{prange} function which
+            essentially translates into an OpenMP directive. This turned out to
+            be slower than the built-in NumPy calls, especially since NumPy
+            comes with the option of running in parallel if the code is built
+            with BLAS by setting the environment variable
+            \texttt{OMP\_NUM\_THREADS}. For example:
+            \begin{bash}
+                export OMP_NUM_THREADS=4
+            \end{bash}
+
+            Therefore, our most optimized code uses the matrix multiplication
+            code (\texttt{np.matmul}) from NumPy which translates into the BLAS
+            call \texttt{DGEMM} for matrix-matrix multiplication. This requires
+            us to reshape the four-dimensional tensors as matrices and write
+            them in such a way that we can do the contraction using matrix
+            multiplication. Here NumPy comes to our rescue as a reshape is done
+            inplace and all memory is stored as a contiguous array.
 

### 2018-06-13 project_2/article/article.tex
Start on review of results

@@
 \usepackage{listings, lstautogobble}
-\usepackage{minted}
+%\usepackage{minted}
 % Colors
@@
         ensure that all requirements of the program are satisfied you can
-        install with \mintinline{bash}{make installr} instead. Running
-        \mintinline{bash}{py.test} on the command-line in the
+        install with \texttt{make installr} instead. Running
+        \texttt{py.test} on the command-line in the
         top-directory\footnote{That is, in
@@
 
-        The first subsection, \mintinline{bash}{matrix_elemtents},
+        The first subsection, \texttt{matrix\_elements},
         contains methods for computing matrix elements in harmonic
@@
 
-        The \mintinline{bash}{hartree_fock} subsection contains methods
+        The \texttt{hartree\_fock} subsection contains methods
         for changing from harmonic oscillator basis to Hartree-Fock basis as
@@
 
-        The \mintinline{bash}{schemes} subsection is arguably the most important
+        The \texttt{schemes} subsection is arguably the most important
         part of this project. The subsection has three different classes that
         perform the exact same computations, but in different and increasingly
-        intelligent ways. First, \mintinline{python}{CoupledClusterDoubles} is
+        intelligent ways. First, \texttt{CoupledClusterDoubles} is
         the most straightforward and naïve way to solve the CCD amplitude
@@
         operator matrices in this problem are zero, we have implemented a sparse
-        matrix CCD solver in \mintinline{python}{CoupledClusterDoublesSparse}.
-        Third, \mintinline{python}{CoupledClusterDoublesOptimized} is
+        matrix CCD solver in \texttt{CoupledClusterDoublesSparse}.
+        Third, \texttt{CoupledClusterDoublesOptimized} is
         parallellized and optimized with memory use and number of floating point
@@
     \subsection{Implementation details}
-        Both in \mintinline{python}{CoupledClusterDoublesSparse} and
-        \mintinline{python}{CoupledClusterDoublesOptimized} we make use of the
+        Both in \texttt{CoupledClusterDoublesSparse} and
+        \texttt{CoupledClusterDoublesOptimized} we make use of the
         intermediates in equations \ref{eq:intermediate1} to
@@
         Moreover, several methods in NumPy for doing linear algebra was tested,
-        including but not limited to \mintinline{python}{numpy.einsum},
-        \mintinline{python}{numpy.tensordot}, and reshaping datastructures to
-        two dimensions and using \mintinline{python}{numpy.matmul}.
+        including but not limited to \texttt{numpy.einsum},
+        \texttt{numpy.tensordot}, and reshaping datastructures to
+        two dimensions and using \texttt{numpy.matmul}.
         Additionally, we tried to make use of the python package
-        \mintinline{python}{dask} which allows for arrays that are larger than
+        \texttt{dask} which allows for arrays that are larger than
         the computer's memory buffer as well as automated concurrency. We found
         that the dask arrays was a bad fit for our code base, but we ended up
-        using \mintinline{python}{numba} for paralellization of some
+        using \texttt{numba} for paralellization of some
         iterative loops as well as just-in-time compilation of some functions.
@@
         \begin{align}
-            \tilde{t}^{k + 1} = \theta t^{k + 1} + (1 - \theta)t^k,
+            \tilde{t}^{k + 1} = (1 - \theta) t^{k + 1} + \theta t^k,
             \label{eq:mixing}
@@
 
-        All the results from our simulations are displayed in tables \ref{tab:N2},
-        \ref{tab:N6}, \ref{tab:N12} and \ref{tab:N20}. They show the ground state
-        energy as a result of restricted Hartree-Fock SCF iterations, and coupled
-        cluster doubles with harmonic oscillator basis and Hatree-Fock basis for
-        systems of increasing size, $N=2$, $N=6$, $N=12$ and $N=20$ particles
-        respectively. The results are thoroughtly discussed in the next section.
+        All the results from our simulations are displayed in tables
+        \ref{tab:N2}, \ref{tab:N6}, \ref{tab:N12} and \ref{tab:N20}. The tables
+        show the results for systems of increasing size, $N=2$, $N=6$, $N=12$
+        and $N=20$ particles respectively. The results are thoroughtly discussed
+        in the next section.
 
@@
             \centering
-            \caption{Here we look at $N = 12$ particles ($R = 3$ shells).  We did
-            not achieve convergence using the harmonic oscillator basis for the
-            lower frequency values and large number of shells. No convergence is
+            \caption{Here we look at $N = 12$ particles.  We did not achieve
+            convergence using the harmonic oscillator basis for the lower
+            frequency values and large number of shells. No convergence is
             marked as {\nan} in the table.}
@@
             \centering
-            \caption{In this table we look at $N = 20$ particles ($R = 4$
-            shells).  We did not achieve convergence using the harmonic
-            oscillator basis for the lower frequency values and large number of
-            shells. No convergence is marked as {\nan} in the table.}
+            \caption{In this table we look at $N = 20$ particles.  We did not
+            achieve convergence using the harmonic oscillator basis for the
+            lower frequency values and large number of shells. No convergence is
+            marked as {\nan} in the table.}
             \begin{ruledtabular}
@@
         Here we show how the three implementations of the CCD-method scales as a
-        function of the shell number $R$ for $N = 2$ and $N = 6$, in table
-        \ref{fig:running_time_2} and \ref{fig:running_time_6} respectively. The same
-        behaviour is shown for higher number of particles. In \autoref{fig:running_time_6_zoom}
-        we have plotted the computation time for $N=6$ particles again, but we
-        have excluded the naïve method for better comparison of the sparse and
-        optimized method.
+        function of the shell number $R$ for $N = 2$ and $N = 6$, in figures
+        \ref{fig:running_time_2} and \ref{fig:running_time_6} respectively. The
+        same behaviour is shown for higher number of particles. In
+        \autoref{fig:running_time_6_zoom} we have plotted the computation time
+        for $N=6$ particles again, but we have excluded the naïve method for
+        better comparison of the sparse and optimized method.
 
@@
         Somewhat surprisingly, the Hartree-Fock method outperforms the coupled
-        cluster doubles when the harmonic oscillator basis is used. This
-        happens for almost any configuration except the smallest system
-        with $N=2$ particles. This comes from the fact that the HO basis is
-        not very good at representing the states in the system. As we have
-        already stated, because the Hartree-Fock SCF iterations is a variational
-        methdod, the basis is varied in order to find the minimal energy. The new
-        basis states will be linear combinations of the old basis states. When
-        the iterations are finished, the new transformed HF basis will then be
-        much better at representing the possible quantum configurations compared
-        to the HO basis. This effect is only apparent in system with a larger
-        number of particles and/or low frequency. This is because in a system with
-        a large number of particles relative to the frequency particle-particle
-        interaction will be more prominent than the harmonic oscillator potential
-        the particles are subject to. I other words, the HO basis would fit rather
-        well.
+        cluster doubles when the harmonic oscillator basis is used. This happens
+        for almost any configuration except the smallest system with $N=2$
+        particles. This comes from the fact that the HO basis is not very good
+        at representing the states in the system. Because the Hartree-Fock SCF
+        iterations is a variational method, the basis is varied in order to find
+        the minimal energy. The new basis states will be linear combinations of
+        the old basis states. When the iterations are finished, the new
+        transformed HF basis will then be much better at representing the
+        possible quantum configurations compared to the HO basis. This effect is
+        only apparent in systems with a larger number of particles and/or low
+        frequency. This is because in a system with a large number of particles
+        relative to the frequency will experience that the particle-particle
+        interaction will be more prominent than the harmonic oscillator
+        potential the particles are subject to. The trapped
+        particles in the harmonic oscillator potential will be pushed out of the
+        well unless we increase the frequency thus increasing the strength of
+        the trap.
+
+        In other words, our first guess using the harmonic oscillator basis
+        functions as an approximation to the exact eigenfunction of the quantum
+        dot Hamiltonian is only a good first attempt when the number of
+        particles is low and/or the frequency is very high. In those situations
+        we can treat the particle-particle interaction as a small perturbation.
 
@@
 
-        The rest of results have been compared with the master thesis of M. P.
-        Lohne\cite{lohne2010coupled} for up to 10 shells. Lohne gets two sets of
-        energies, one set from RHF and one set from CCSD code with harmonic
-        oscillator basis functions. Our CCD code with Hartree-Fock basis will
-        for some configurations\footnote{By configurations we mean frequency
-        $\omega$ and number of particles $N$.} beat CCSD with plain harmonic
-        oscillator basis. For 12 shells we can compare with the results from M.
-        P. Lohne et al.\cite{lohne2011ab}, but in this article an
-        \emph{effective interaction} for the Hamiltonian with a CCSD code has
-        been used. These results are therefore significanlty better than ours,
-        but we can get a ``ball-park'' idea to benchmark against.
+        The rest of the results have been compared with the master thesis of M.
+        P.  Lohne\cite{lohne2010coupled} for up to $R = 10$ shells. Lohne gets
+        two sets of energies, one set from RHF and one set from CCSD code with
+        harmonic oscillator basis functions. Our CCD code with Hartree-Fock
+        basis will for some configurations\footnote{By configurations we mean
+        frequency $\omega$ and number of particles $N$.} beat CCSD with plain
+        harmonic oscillator basis. For $R = 12$ shells we can compare with the
+        results from M. P. Lohne et al.\cite{lohne2011ab}, but in this article
+        an \emph{effective interaction} for the Hamiltonian with a CCSD code
+        using the HF basis has been used. These results are therefore
+        significantly better than ours, but we can get a ``ball-park'' idea to
+        benchmark against.
 

### 2018-06-14 project_2/article/article.tex
Add reviewed project

@@
     because we have limitations in memeory and computating power. Luckily, great
-    minds have bestowed a wide selection of very god numerical methods upon the
+    minds have bestowed a wide selection of very good numerical methods upon the
     world, which the coupled cluster is a part of.
@@
     are easy to represent numerically. Second, we provide a brief overview of
-    the program we have constructed which is in the form an installable python
+    the program we have constructed which is in the form an installable Python
     package.  Third, we have employed the methods on a large parameter space
@@
         perfectly accurate, yet simple. There are some important effects that
-        should be included such the potential arising as a result of the
+        should be included such as the potential arising as a result of the
         confinement, and the interaction between particles. We therefore model
         our system with an electron-electron interaction, given by the two-body
-        operator in \autoref{eq:two_body_operator}, and with we approximate the
-        confining potential by the harmonic oscillator potential, given by the
-        second term in the one-body Hamiltonian in
-        \autoref{eq:one_body_hamiltonian}. The last term in the one-body
+        operator in \autoref{eq:two_body_operator}, and with it we approximate
+        the confining potential by the harmonic oscillator potential, given by
+        the second term in the one-body Hamiltonian in
+        \autoref{eq:one_body_hamiltonian}. The first term in the one-body
         Hamiltonian is the kinetic energy of particle in question, essential in
@@
         two dimensions and using \texttt{numpy.matmul}.
-        Additionally, we tried to make use of the python package
+        Additionally, we tried to make use of the Python package
         \texttt{dask} which allows for arrays that are larger than
@@
         \end{enumerate}
-        For the CCD method the naïve solution using the form of $g(f, u, t)$
+        For the CCD-method the naïve solution using the form of $g(f, u, t)$
         without the intermediates results in $\mathcal{O}(M^4N^4)$ FLOPS each
@@
     \subsection{Validity of results}
-        The analytical computed energy for $\omega=1.0$ a.u.
-        and $N=2$ is $3$\cite{taut1994two}. From our tables we see
-        that RHF is least successful in reaching this value, CCD in
-        HF basis is best, while CCD with HO basis is generally somewhere
-        between the two. We see that we get even closer to the analytical
-        ``truth'' by increasing the number of shells.
+        The analytical computed energy for $\omega=1.0$ a.u.  and $N=2$ is $3$,
+        from Taut\cite{taut1994two}. From our tables we see that RHF is the
+        least successful in reaching this value, CCD using the HF basis is best,
+        while CCD with the HO basis is generally somewhere between the two. We
+        see that we get even closer to the analytical ``truth'' by increasing
+        the number of shells.
+
+        As predicted in the previous subsection we see that the CCD-method using
+        the HO basis gradually deteroriates compared with RHF and CCD using the
+        HF basis as the number of particles increases. In fact, it quickly
+        reaches a point where we get trouble with convergence due to the
+        unstable states with a large number of particles packed together in a
+        too weak trap.
+
+        When the states become too unstable we have to increase the mixing
+        parameter $\theta \to 1$ in \autoref{eq:mixing}. This will sometimes
+        make the method converge, but it can be at the cost of a wrong value of
+        the energy as we might get trapped in unstable configurations which with
+        the help of mixing will converge for our criteria.
 
@@
         P.  Lohne\cite{lohne2010coupled} for up to $R = 10$ shells. Lohne gets
-        two sets of energies, one set from RHF and one set from CCSD code with
+        two sets of energies, one set from RHF and one set from CCSD code using
         harmonic oscillator basis functions. Our CCD code with Hartree-Fock
@@
         intermediate calculations these running times scales on the order of
-        $\mathcal{O}(l^6)$ whereas the naïve implementation goes as
-        $\mathcal{O}(l^8)$. Another important factor is that the naïve
+        $\mathcal{O}(M^4N^2)$ whereas the naïve implementation goes as
+        $\mathcal{O}(M^4N^4)$. The exponential difference in scaling is readily
+        apparent from these plots. Another important factor is that the naïve
         implementation uses \texttt{np.einsum} to calculate the tensor
@@
         interaction gets too strong. The CCD-method will in particular get a
-        hard time keep the particles together as the excitation operator $t$
+        hard time keeping the particles together as the excitation operator $t$
         only excites pairs of particles leading to a too strong increase in the
@@
         increasing to $N = 6$ particles we immediately see how CCD with the HO
-        basis will get in trouble for small values of $\omega$. Even by increasing
-        the strength of the confinment this method does converge.
+        basis will get in trouble for small values of $\omega$. Even by
+        increasing the strength of the confinment this method does not always
+        converge.
 
@@
         frequency we restored the convergence. For larger number of shells we
-        had to start increasing the value of $\theta \to 0.9$ in order to get
+        had to start increasing the value of $\theta \to 1$ in order to get
         convergence within the threshold.
 
-        CCD with HO basis experiences the same convergence problems also for,
-        $N=12$ particles and for $N=20$ particles.
-
-    %\subsection{Sparse implementation}
-    % Did you want to write something about memory use, Oyvind? Memory problems?
+        The CCD-method with the HO basis experiences the same convergence
+        problems also for, $N=12$ particles and for $N=20$ particles.
+
+    \subsection{Sparse implementation}
+        From \autoref{fig:running_time_6_zoom} we can see that the sparse
+        implementation is close to the optimized version of the CCD-method. This
+        is a sobering idea as the sparse implementation is by no means optimized
+        properly and more resembles the naïve implementation. This brings faith
+        in the fact that a more optimized version of the sparse implementation
+        using block matrices will be able to achieve much better performance.
+        This is however something that will be experimented with in the future.
 
 \section{Summary Remarks}
-
-        Our main product from this is a product that can do both computations with
-        restricted Hartree-Fock (RHF) self-consistent field and coupled cluster
-        doubles (CCD). We have tried
-        to make a package in python that should be easy enough for anyone to test.
-        Moreover we hav demonstrated that one can speed up the computation time with
-        a relatively simple algebraic transformation and finding the ``intermediates''.
-        The computational speed was further reduced using sparse matrices. It should
-        be noted, as a warning, that there are relatively few systems adequately sparese
-        for this to work. We added upon the program to increase it further with
+    Our main product from this project is a code that can do both computations
+    using the restricted Hartree-Fock (RHF) self-consistent field and coupled
+    cluster doubles (CCD). We have tried to make a package in Python that should
+    be easy enough for anyone to test.  Moreover we have demonstrated that one
+    can speed up the computation time with a relatively simple algebraic
+    transformation and finding the ``intermediates''. The computational speed
+    was further reduced using sparse matrices. It should be noted, as a warning,
+    that not all systems are sparse enough for a sparse implementation to make
+    sense.  We added upon the program to increase it further with
     paralellization and optimal use of numerical functions.
 
-        In order to perform the computations we have constructed a harmonic oscillator
-        basis. We saw how the CCD method was outperformed RHF in most cases, and often
-        exhibited convergence problems. The remedy for these shortcomings is to use
-        the Hartree-Fock basis which is provided after successful convergence of
-        SCF iterations. The results are improved markedly and classification of coupled
-        cluster methods as a post-Harte-Fock method is justified.
-
-        The coupled cluster method provides surprisingly good results even
-        though one only includes double esctiations. The reason for this is that
-        particle-particle interactions is the amongt the most important thing that
-        happens in a ferminon system, second only to no interaction at all, i.e
+    In order to perform the computations we have constructed a harmonic
+    oscillator basis. We saw how the CCD-method using this basis was
+    outperformed by RHF in most cases, and often exhibited convergence problems.
+    The remedy for these shortcomings is to use the Hartree-Fock basis which is
+    provided after successful convergence of SCF iterations. The results are
+    improved markedly and the classification of coupled cluster methods as a
+    post-Harte-Fock method is justified.
+
+    The coupled cluster method provides surprisingly good results even though
+    one only includes double exitations. The reason for this is that
+    particle-particle interactions is among the most important thing that
+    happens in a fermionic system, second only to no interaction at all, i.e
     what particles do on their own. The way to improve the computations is to
@@
     add more basis functions, and/or supplement the coupled cluster method with
-        another method like a perturbative one. All of this will, needless to say,
-        increase the compute time, which must be balanced against the accuracy needed
-        for the situation at hand.
+    another method like perturbation theory. All of this will, needless to say,
+    increase the compute time, which must be balanced against the accuracy
+    needed for the situation at hand.
 
@@
     When constructing the normal ordered Hamiltonian we use Wick's theorem to
-    write the one-body, $h$, and the two-body, $u$, operators onto a normal
-    ordered form. Specifically we define the normal ordered form in terms of the
-    \emph{Fermi vacuum}\footnote{Fermi vacuum defines the reference state, i.e.,
-    $\kslat$, as the vacuum.}. That is, an operator on normal ordered form
-    destroys the reference Slater determinant.
-
-    We start by writing the one-body operator, $h$, to its normal-ordered form.
-    \begin{align}
-        h &= \sum_{pq}h^p_q\acr{p}\ade{q}
-        = \sum_{pq}h^p_q\para{
+    write $h$ and $u$ onto a normal ordered form. Specifically we define the
+    normal ordered form in terms of the \emph{Fermi vacuum}\footnote{Fermi
+    vacuum defines the reference state, i.e., $\kslat$, as the vacuum.}. That
+    is, an operator on normal ordered form destroys the reference Slater
+    determinant.
+
+    We start by writing the one-body operator $h$, to its normal-ordered form.
+    \begin{align}
+        h &= h^p_q\acr{p}\ade{q}
+        = h^p_q\para{
             \{\acr{p}\ade{q}\}
@@
         \\
-        &= \sum_{pq}h^p_q\{
+        &= h^p_q\{
             \acr{p}\ade{q}
         \}
-        + \sum_{pq}h^p_q\delta_{p \in i}\delta_{pq}
+        + h^p_q\delta_{p \in i}\delta_{pq}
         \\
-        &= h_N + \sum_{i}h^i_i,
+        &= h_N + h^i_i,
     \end{align}
@@
     index. Doing the same for the two-body operator is a slightly more tedious
-    endeavor. For brevity we will only write out the operator strings and only
-    keep the non-zero contributions.
+    endeavor. For brevity we will only write out the operator strings.
     \begin{align}
@@
         &=
-        \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
+        \frac{1}{4}u^{pq}_{rs}
         \{\acr{p}\acr{q}\ade{s}\ade{r}\}
-        - \frac{1}{4}\sum_{iqr}\bra{iq}\ket{ri}\{\acr{q}\ade{r}\}
-        \nonumber \\
-        &\qquad
-        + \frac{1}{4}\sum_{iqs}\bra{iq}\ket{is}\{\acr{q}\ade{s}\}
-        + \frac{1}{4}\sum_{pir}\bra{pi}\ket{ri}\{\acr{p}\ade{r}\}
-        \nonumber \\
-        &\qquad
-        - \frac{1}{4}\sum_{pis}\bra{pi}\ket{is}\{\acr{p}\ade{s}\}
-        - \frac{1}{4}\sum_{ij}\bra{ij}\ket{ji}
-        \nonumber \\
-        &\qquad
-        + \frac{1}{4}\sum_{ij}\bra{ij}\ket{ij}.
+        - \frac{1}{4}u^{iq}_{ri}\{\acr{q}\ade{r}\}
+        \nonumber \\
+        &\qquad
+        + \frac{1}{4}u^{iq}_{is}\{\acr{q}\ade{s}\}
+        + \frac{1}{4}u^{pi}_{ri}\{\acr{p}\ade{r}\}
+        \nonumber \\
+        &\qquad
+        - \frac{1}{4}u^{pi}_{is}\{\acr{p}\ade{s}\}
+        - \frac{1}{4}u^{ij}_{ji}
+        \nonumber \\
+        &\qquad
+        + \frac{1}{4}u^{ij}_{ij}.
     \end{align}
@@
     \begin{align}
-        \bra{pq}\ket{rs}
-        = - \bra{pq}\ket{sr}
-        = - \bra{qp}\ket{rs}
-        = \bra{qp}\ket{sr},
+        u^{pq}_{rs}
+        = - u^{pq}_{sr}
+        = - u^{qp}_{rs}
+        = u^{qp}_{sr},
     \end{align}
@@
         &=
-        W_N + \sum_{pir}\bra{pi}\ket{ri}\{\acr{p}\ade{r}\}
-        + \half\sum_{ij}\bra{ij}\ket{ij},
+        W_N + u^{pi}_{ri}\{\acr{p}\ade{r}\}
+        + \half u^{ij}_{ij},
     \end{align}
@@
     \begin{align}
-        W_N = \frac{1}{4}\sum_{pqrs}
-        \bra{pq}\ket{rs}\{\acr{p}\acr{q}\ade{s}\ade{r}\}.
+        W_N = \frac{1}{4}
+        u^{pq}_{rs}\{\acr{p}\acr{q}\ade{s}\ade{r}\}.
     \end{align}
     When we now construct the full Hamiltonian we can collect some terms. The
-    constants in both the one-body and the two-body operator in total constitues
-    the reference energy.
+    constants in both the one-body and the two-body operator in total
+    constitutes the reference energy.
     \begin{align}
         E_0 \equiv \bslat H\kslat
-        = \sum_{i}h_i^i + \frac{1}{2}\sum_{ij}\bra{ij}\ket{ij}.
+        = h_i^i + \frac{1}{2}u^{ij}_{ij}.
         \label{eq:reference_energy}
@@
         &=
-        \sum_{pq}h_{q}^{p}\{\acr{p}\ade{q}\}
-        + \sum_{pqi}\bra{pi}\ket{qi}\{\acr{p}\ade{q}\}
+        h_{q}^{p}\{\acr{p}\ade{q}\}
+        + u^{pi}_{qi}\{\acr{p}\ade{q}\}
         \\
-        &= \sum_{pq}f_{q}^{p}\{\acr{p}\ade{q}\},
+        &= f_{q}^{p}\{\acr{p}\ade{q}\},
     \end{align}
@@
         &=
-        h_q^p + \sum_{i}\bra{pi}\ket{qi}.
+        h_q^p + u^{pi}_{qi}.
     \end{align}
@@
     \end{align}
-    where $A$ is the create and destruction operator strings. By using
+    where $A$ is the creation and destruction operator strings. By using the
     generalized Wick's theorem we get
@@
     In this section we have provided a few sample computations of how one would
-    evaluate the amplitude equations using wicks theorem. Starting with the simplest
-    term including only the normal-ordered Hamiltonian,
+    evaluate the amplitude equations using Wick's theorem. Starting with the
+    simplest term including only the normal-ordered Hamiltonian,
 
@@
         \begin{aligned}
-            \bra{\Phi_{ij}^{ab}} &(F_N + V_N)\ket{\Phi_0} \\
-             =& \sum_{pq}f_{pq}\bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa_q\}\ket{\Phi_0} \\
-             &+\frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
-             \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0}.
+            \bra{\Phi_{ij}^{ab}} &(F_N + V_N)\ket{\Phi_0}
+             = f^p_q\bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}
+             \{a^\dagger_pa_q\}\ket{\Phi_0} \\
+             &\qquad
+             + \frac{1}{4}u^{pq}_{rs}
+             \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}
+             \{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0}.
         \end{aligned}
@@
         \begin{aligned}
-            \bra{\Phi_{ij}^{ab}} &(V_N)\ket{\Phi_0} \\
-                =& \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}
-             \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}\{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0} \\
-                 =& \frac{1}{4} \sum_{pqrs} \bra{pq}\ket{rs} \\ &\times\Big(
+            \bra{\Phi_{ij}^{ab}} &(V_N)\ket{\Phi_0}
+            = \frac{1}{4}u^{pq}_{rs}
+            \bra{\Phi_0}\{a_i^\dagger a_j^\dagger a_ba_a\}
+            \{a^\dagger_pa^\dagger_qa_sa_r\}\ket{\Phi_0} \\
+                 &= \frac{1}{4} u^{pq}_{rs} \\
+                 &\times
+                 \Biggl(
                      \{ \wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a
@@
                      \{ \wick{\c1 a^\dagger_i \c2 a^\dagger_j \c4 a_b \c3 a_a
-                     \c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\} \\
-                   &\ +
+                     \c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
+                     \\
+                     &+
                       \{ \wick{\c2 a^\dagger_i \c1 a^\dagger_j \c3 a_b \c4 a_a
@@
                      \c4 a^\dagger_p \c3 a^\dagger_q \c2 a_s \c1 a_r}\}
-                 \Big) \\
-                 =&  \frac{1}{4} \sum_{pqrs} \bra{pq}\ket{rs}
-                 ( \delta_{ap}\delta_{bq}\delta_{js}\delta_{ir}
+                 \Biggr) \\
+                 &=  \frac{1}{4} u^{pq}_{rs}
+                 \biggl( \delta_{ap}\delta_{bq}\delta_{js}\delta_{ir}
                  -\delta_{aq}\delta_{bp}\delta_{js}\delta_{ir} \\
-                 &\ -\delta_{ap}\delta_{bq}\delta_{jr}\delta_{is}
-                  +\delta_{aq}\delta_{bp}\delta_{jr}\delta_{is}) \\
-                  =& \bra{ab}\ket{ij} = u^{ab}_{ij}.
+                 & -\delta_{ap}\delta_{bq}\delta_{jr}\delta_{is}
+                 + \delta_{aq}\delta_{bp}\delta_{jr}\delta_{is}\biggr) \\
+                &= u^{ab}_{ij}.
         \end{aligned}
@@
 
-        Next we would like to evaluate $\bra{\Phi_{ij}^{ab}}(F_N + V_N)T_2\ket{\Phi_0}$. Starting with the
-        term involving the Fock operator $F_N$,
+        \begin{widetext}
+            Next we would like to evaluate the normal ordered Hamiltonian connected
+            with a single doubles cluster operator. Starting with the term involving
+            the Fock operator $F_N$,
+            \begin{align}
+                \bra{\Phi_{ij}^{ab}}(F_NT_2)_c\ket{\Phi_0}
+                    &= \frac{1}{4}f^p_qt_{kl}^{cd} \bra{\Phi_0}
+                    \{a_i^\dagger a_j^\dagger a_b a_a \}
+                    \para{
+                        \{a^\dagger_pa_q \}\{a^\dagger_c a^\dagger_d a_l a_k \}
+                    }_c \ket{\Phi_0}
+                    \nonumber \\
+                    &= \frac{1}{4}f^p_q t_{kl}^{cd}
+                    \Biggl(
+                        \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c5 a^\dagger_p \c2 a_q \c3 a^\dagger_c
+                                \c4 a^\dagger_d \c1 a_l \c5 a_k
+                            }
+                        \}
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c5 a^\dagger_p \c1 a_q \c3 a^\dagger_c
+                                \c4 a^\dagger_d \c2 a_l \c5 a_k
+                            }
+                        \}
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a
+                                \c5 a^\dagger_p \c2 a_q \c4 a^\dagger_c
+                                \c3 a^\dagger_d \c1 a_l \c5 a_k
+                            }
+                        \}
+                        \nonumber \\
+                        &\qquad
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c5 a^\dagger_p \c1 a_q \c4 a^\dagger_c
+                                \c3 a^\dagger_d \c2 a_l \c5 a_k
+                            }
+                        \}
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c5 a^\dagger_p \c1 a_q \c3 a^\dagger_c
+                                \c4 a^\dagger_d \c5 a_l \c2 a_k
+                            }
+                        \}
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c5 a^\dagger_p \c1 a_q \c4 a^\dagger_c
+                                \c3 a^\dagger_d \c5 a_l \c2 a_k
+                            }
+                        \}
+                        \nonumber \\
+                        &\qquad
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c5 a^\dagger_p \c2 a_q \c3 a^\dagger_c
+                                \c4 a^\dagger_d \c5 a_l \c1 a_k
+                            }
+                        \}
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c5 a^\dagger_p \c2 a_q
+                                \c4 a^\dagger_c \c3 a^\dagger_d \c5 a_l \c1 a_k
+                            }
+                        \}
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c3 a^\dagger_p \c5 a_q
+                                \c5 a^\dagger_c \c4 a^\dagger_d \c2 a_l \c1 a_k
+                            }
+                        \}
+                        \nonumber \\
+                        &\qquad
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c3 a^\dagger_p \c5 a_q
+                                \c5 a^\dagger_c \c4 a^\dagger_d \c1 a_l \c2 a_k
+                            }
+                        \}
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c4 a^\dagger_p \c5 a_q \c5 a^\dagger_c
+                                \c3 a^\dagger_d \c2 a_l \c1 a_k
+                            }
+                        \}
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c4 a^\dagger_p \c5 a_q \c5 a^\dagger_c
+                                \c3 a^\dagger_d \c1 a_l \c2 a_k
+                            }
+                        \}
+                        \nonumber \\
+                        &\qquad
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a
+                                \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c
+                                \c5 a^\dagger_d \c1 a_l \c2 a_k
+                            }
+                        \}
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c
+                                \c5 a^\dagger_d \c2 a_l \c1 a_k
+                            }
+                        \}
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c
+                                \c5 a^\dagger_d \c1 a_l \c2 a_k
+                            }
+                        \}
+                        \nonumber \\
+                        &\qquad
+                        + \{
+                            \wick{
+                                \c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b
+                                \c4 a_a \c4 a^\dagger_p \c5 a_q \c3 a^\dagger_c
+                                \c5 a^\dagger_d \c2 a_l \c1 a_k
+                            }
+                        \}
+                    \Biggr)
+                    \nonumber \\
+                    &= f^p_q t_{kl}^{cd}
+                    \Bigl(
+                        - \delta_{a c} \delta_{b d} \delta_{i k}
+                        \delta_{j q} \delta_{l p}
+                        + \delta_{a c} \delta_{b d} \delta_{i l}
+                        \delta_{j q} \delta_{k p}
+                        + \delta_{a c} \delta_{b d} \delta_{i q}
+                        \delta_{j k} \delta_{l p}
+                        - \delta_{a c} \delta_{b d} \delta_{i q}
+                        \delta_{j l} \delta_{k p}
+                        \nonumber \\
+                        &\qquad
+                        + \delta_{a c} \delta_{b p} \delta_{d q}
+                        \delta_{i k} \delta_{j l}
+                        - \delta_{a c} \delta_{b p} \delta_{d q}
+                        \delta_{i l} \delta_{j k}
+                        + \delta_{a d} \delta_{b c} \delta_{i k}
+                        \delta_{j q} \delta_{l p}
+                        - \delta_{a d} \delta_{b c} \delta_{i l}
+                        \delta_{j q} \delta_{k p}
+                        \nonumber \\
+                        &\qquad
+                        - \delta_{a d} \delta_{b c} \delta_{i q}
+                        \delta_{j k} \delta_{l p}
+                        + \delta_{a d} \delta_{b c} \delta_{i q}
+                        \delta_{j l} \delta_{k p}
+                        - \delta_{a d} \delta_{b p} \delta_{c q}
+                        \delta_{i k} \delta_{j l}
+                        + \delta_{a d} \delta_{b p} \delta_{c q}
+                        \delta_{i l} \delta_{j k}
+                        \nonumber \\
+                        &\qquad
+                        - \delta_{a p} \delta_{b c} \delta_{d q}
+                        \delta_{i k} \delta_{j l}
+                        + \delta_{a p} \delta_{b c} \delta_{d q}
+                        \delta_{i l} \delta_{j k}
+                        + \delta_{a p} \delta_{b d} \delta_{c q}
+                        \delta_{i k} \delta_{j l}
+                        - \delta_{a p} \delta_{b d} \delta_{c q}
+                        \delta_{i l} \delta_{j k}
+                    \Big)
+                    \nonumber \\
+                    &= (f^b_c t^{ac}_{ij} - f^a_c t^{bc}_{ij})
+                    - (f^j_k t^{ab}_{ik} - f^i_k t^{ab}_{jk})
+                    \nonumber \\
+                    &= f^b_c t^{ac}_{ij}P(ab) - f^k_j t^{ab}_{ik}P(ij).
+            \end{align}
+            This procedure can be repeated for the remaining steps, or one can
+            resort to using SymPy.
+        \end{widetext}
+
+\section{Finding amplitude equation intermediates}
+    \label{app:intermediates}
+    Starting from the CCD amplitude equation we factor out terms to create
+    pre-computed intermediates. We are particularly interested in all terms
+    where there is a contraction between three tensors as they are the
+    computationaly most intensive.
+    \begin{align}
+        g(f, u, t) &\gets
+        \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd}
+        + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd}
+        \\
+        &= t^{cd}_{ij} \para{
+            \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2} u^{ab}_{cd}
+        } \\
+        &= t^{cd}_{ij}\chi^{ab}_{cd}.
+    \end{align}
+    The two next intermediates only consists of pre-computing contractions in a
+    single term instead of combining two terms.
+    \begin{align}
+        g(f, u, t) &\gets
+        \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij)
+        \\
+        &= t^{ab}_{in} \left(\frac{1}{2} t^{cd}_{jm} u^{mn}_{cd}\right) P(ij)
+        \\
+        &= t^{ab}_{in}\chi^n_j P(ij),
+        \\
+        g(f, u, t) &\gets
+        \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab)
+        \\
+        &=
+        t^{bd}_{ij} \left(\frac{1}{2}t^{ac}_{nm}  u^{nm}_{cd}\right) P(ab)
+        \\
+        &= t^{bd}_{ij}\chi^a_d P(ab).
+    \end{align}
+    The last intermediate is not so straight-forward. We have insert an extra
+    exchange operator in order to group two terms into an intermediate.
+    \begin{align}
+        g(f, u, t) &\gets
+        t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij)
+        \nonumber \\
+        &\qquad
+        + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
+        &=
+        t^{ac}_{im}\left(\frac{1}{2} t^{bd}_{jn} u^{mn}_{cd}
+        + u^{bm}_{jc} \right)P(ab) P(ij) \\
+        &=
+        t^{ac}_{im}\chi^{bm}_{jc} P(ab)P(ij).
+    \end{align}
+    The total expression for $g(f, u, t)$ is then
     \begin{equation}
         \begin{aligned}
-            \bra{\Phi_{ij}^{ab}}&F_NT_2\ket{\Phi_0} \\
-                =& \frac{1}{4}\sum_{pq}\sum_{klcd}f_{pq}t_{kl}^{cd} \bra{\Phi_0}
-                    \{a_i^\dagger a_j^\dagger a_b a_a \}(\{a^\dagger_pa_q \}\{a^\dagger_c a^\dagger_d a_l a_k \})_c \ket{\Phi_0} \\
-                =& \frac{1}{4}\sum_{pq}\sum_{klcd}f_{pq} t_{kl}^{cd} \\
-                \times \Big(
-                    &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c1 a_l \c5 a_k } \}
-                 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c2 a_l \c5 a_k } \}    \\
-                + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c1 a_l \c5 a_k } \}
-                 + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c2 a_l \c5 a_k } \} \\
-                + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c5 a_l \c2 a_k } \}
-                  + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c1 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c5 a_l \c2 a_k } \} \\
-                + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c3 a^\dagger_c \c4 a^\dagger_d \c5 a_l \c1 a_k } \}
-                  + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c5 a^\dagger_p \c2 a_q \c4 a^\dagger_c \c3 a^\dagger_d \c5 a_l \c1 a_k } \} \\
-                + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c4 a^\dagger_d \c2 a_l \c1 a_k } \}
-                  + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c4 a^\dagger_d \c1 a_l \c2 a_k } \} \\
-                + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c3 a^\dagger_d \c2 a_l \c1 a_k } \}
-                  +\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c5 a^\dagger_c \c3 a^\dagger_d \c1 a_l \c2 a_k } \} \\
-                + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c1 a_l \c2 a_k } \}
-                  +\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c2 a_l \c1 a_k } \} \\
-                + &\{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c3 a^\dagger_p \c5 a_q \c4 a^\dagger_c \c5 a^\dagger_d \c1 a_l \c2 a_k } \}
-                  + \{\wick{\c1 a^\dagger_i \c2 a^\dagger_j \c3 a_b \c4 a_a \c4 a^\dagger_p \c5 a_q \c3 a^\dagger_c \c5 a^\dagger_d \c2 a_l \c1 a_k } \}
-                \Big)
-        \end{aligned}
-        \end{equation}
-        \begin{equation}
-        \begin{aligned}
-                =& \sum_{pq}\sum_{klcd}f_{pq} t_{kl}^{cd} \\
-                \times&( - \delta_{a c} \delta_{b d} \delta_{i k} \delta_{j q} \delta_{l p} + \delta_{a c} \delta_{b d} \delta_{i l} \delta_{j q} \delta_{k p}  \\
-                 \ &+ \delta_{a c} \delta_{b d} \delta_{i q} \delta_{j k} \delta_{l p} - \delta_{a c} \delta_{b d} \delta_{i q} \delta_{j l} \delta_{k p} \\
-                 \ &+ \delta_{a c} \delta_{b p} \delta_{d q} \delta_{i k} \delta_{j l} - \delta_{a c} \delta_{b p} \delta_{d q} \delta_{i l} \delta_{j k} \\
-                 \ & + \delta_{a d} \delta_{b c} \delta_{i k} \delta_{j q} \delta_{l p} - \delta_{a d} \delta_{b c} \delta_{i l} \delta_{j q} \delta_{k p} \\
-                 \ &- \delta_{a d} \delta_{b c} \delta_{i q} \delta_{j k} \delta_{l p} + \delta_{a d} \delta_{b c} \delta_{i q} \delta_{j l} \delta_{k p} \\
-                 \ &- \delta_{a d} \delta_{b p} \delta_{c q} \delta_{i k} \delta_{j l} + \delta_{a d} \delta_{b p} \delta_{c q} \delta_{i l} \delta_{j k} \\
-                 \ &- \delta_{a p} \delta_{b c} \delta_{d q} \delta_{i k} \delta_{j l} + \delta_{a p} \delta_{b c} \delta_{d q} \delta_{i l} \delta_{j k} \\
-                 \ &+ \delta_{a p} \delta_{b d} \delta_{c q} \delta_{i k} \delta_{j l} - \delta_{a p} \delta_{b d} \delta_{c q} \delta_{i l} \delta_{j k}) \\
-                 =& (f_{bc}t^{ac}_{ij} - f_{ac}t^{bc}_{ij}) - (f_{jk}t^{ab}_{ik} - f_{ik}t^{ab}_{jk}) \\
-                 =& f_{bc}t^{ac}_{ij}P(ab) - f_{kj}t^{ab}_{ik}P(ij)
-        \end{aligned}
-        \end{equation}
-        Where in the last steps we have consigned the sums by Einstein summation notation.
-\section{Finding Amplitude Equation Intermediates}
-    \label{app:intermediates}
-    Starting from the CCD amplitude equation we factor out terms that will have the same
-    indices if we contract them,
-    \begin{equation}
-        \begin{aligned}
-        0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
-          &+ \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
-          &+ \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
-          &+ t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
-          &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn},
-        \end{aligned}
-    \end{equation}
-    \begin{equation}
-        \begin{aligned}
-        0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
-          &+ t^{cd}_{ij} \left(\frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2} u^{ab}_{cd} \right) \\
-          &+  t^{ab}_{in} \left(\frac{1}{2} t^{cd}_{jm} u^{mn}_{cd}\right) P(ij) \\
-          &- t^{bd}_{ij} \left(\frac{1}{2}t^{ac}_{nm}  u^{nm}_{cd}\right) P(ab) \\
-          &+ \frac{1}{2}t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{bc}_{im} t^{ad}_{jn} u^{mn}_{cd} P(ij)\\
-          &+ t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
-          &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}
-        \end{aligned}
-    \end{equation}
-    \begin{equation}
-        \begin{aligned}
-        0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
-          &+ t^{cd}_{ij} \left(\frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2} u^{ab}_{cd} \right) \\
-          &+  t^{ab}_{in} \left(\frac{1}{2} t^{cd}_{jm} u^{mn}_{cd}\right) P(ij) \\
-          &- t^{bd}_{ij} \left(\frac{1}{2}t^{ac}_{nm}  u^{nm}_{cd}\right) P(ab) \\
-          &+ \frac{1}{2}t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ab) P(ij) \\
-          &+ t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
-          &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}
-        \end{aligned}
-    \end{equation}
-    \begin{equation}
-        \begin{aligned}
-        0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
-          &+ t^{cd}_{ij} \left(\frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2} u^{ab}_{cd} \right) \\
-          &+  t^{ab}_{in} \left(\frac{1}{2} t^{cd}_{jm} u^{mn}_{cd}\right) P(ij) \\
-          &- t^{bd}_{ij} \left(\frac{1}{2}t^{ac}_{nm}  u^{nm}_{cd}\right) P(ab) \\
-          &+ t^{ac}_{im}\left(\frac{1}{2} t^{bd}_{jn} u^{mn}_{cd} + u^{bm}_{jc} \right)P(ab) P(ij) \\
-          &+ \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
-        \end{aligned}
-    \end{equation}
-    Now we can introduce the intermediate $\chi$-terms,
-    \begin{align}
-        \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
-        \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd} \\
-        \chi^a_d &= \frac{1}{2} t^{ac}_{nm} u^{nm}_{cd} \\
-        \chi^{bm}_{jc} &= \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd} + u^{bm}_{jc},
-    \end{align}
-    this gives us,
-    \begin{equation}
-        \begin{aligned}
-            0 =& u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab) - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
-              &+ t^{cd}_{ij}\chi^{ab}_{cd} + t^{ab}_{in}\chi^n_jP(ij)
-               - t^{bd}_{ij}\chi^a_dP(ab) \\
-              &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij) + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
+            g(f, u, t) &=
+            u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab)
+            - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
+            &\qquad
+            + t^{cd}_{ij}\chi^{ab}_{cd} + t^{ab}_{in}\chi^n_jP(ij)
+            - t^{bd}_{ij}\chi^a_dP(ab)
+            \\
+            &\qquad
+            + t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
+            + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
         \end{aligned}
@@
 %\vfill
-\clearpage
+%\clearpage
 %\newpage

### 2018-10-13 project_2/article/article.tex
Fix an index error in amplitude equations

@@
                     &\qquad
-                    - \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
+                    + \frac{1}{2}t^{ab}_{mn} u^{mn}_{ij}.
                     \label{eq:CCD_amp_1}
@@
                     + t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
-                    + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
+                    + \frac{1}{2}t^{ab}_{mn}u^{mn}_{ij}.
                 \end{aligned}
